```{r}
library(tidyverse)
library(here)
#BiocManager::install("topGO")
library(topGO)
library(cowplot)
library(gprofiler2)

source(here("analysis", "gea", "go.R"))
plotpath <- here("analysis", "gea", "plots")
outpath <- here("analysis", "gea", "outputs")
```

```{r}
# Get GEA genes
genes_org <- read_csv(here("analysis", "gea", "outputs", "bio1ndvi_genes_list.csv"))

# Get all annotated genes
all_genes <-read_csv(here("analysis", "gea", "outputs", "all_genes_list.csv"))

# Check the number of unique genes
print(paste("Number of genes:", nrow(genes_org)))
print(paste("Number of unique genes:", nrow(genes_org %>% distinct(uniprot_id))))
print(paste("Number of all genes:", nrow(all_genes)))
print(paste("Number of unique all genes:", nrow(all_genes %>% distinct(uniprot_id))))

#https://cran.r-project.org/web/packages/gprofiler2/vignettes/gprofiler2.html
#install.packages("gprofiler2")
library(gprofiler2)

genes_org %>% count(organism) %>% arrange(desc(n)) %>% head(20)
```

# Find gene orthologs
Note: warnings are normal, they are just saying that some genes don't have orthologs in the other species
```{r}
genes_hsapiens <- get_ortho(genes_org, target_organism = "hsapiens", file_prefix = "hsapiens")
genes_ggallus <- get_ortho(genes_org, "ggallus", file_prefix = "ggallus")
all_genes_hsapiens <- get_ortho(all_genes, "hsapiens", file_prefix = "hsapiens_all")
all_genes_ggallus <- get_ortho(all_genes, "ggallus", file_prefix = "ggallus_all")

genes_hsapiens <- genes_org %>% mutate(query = uniprot_id)
genes_ggallus <- genes_org%>% mutate(query = uniprot_id)
all_genes_hsapiens <- all_genes%>% mutate(query = uniprot_id)
all_genes_ggallus <- all_genes%>% mutate(query = uniprot_id)
```

# Run GO 

Column descriptions for gprofiler output:
 
query: The name or identifier of the query. This indicates which query the results are associated with.
significant: A logical value (TRUE or FALSE) indicating whether the term is significantly enriched.
p_value: The p-value for the enrichment of the term. It indicates the statistical significance of the enrichment.
term_size: The total number of genes associated with the GO term in the background set.
query_size: The total number of genes in the query set.
intersection_size: The number of genes in the query set that are also associated with the GO term.
*precision:* The precision of the enrichment, calculated as the ratio of intersection_size to query_size. Precision measures the proportion of relevant genes (i.e., genes associated with the GO term) in the query set.
*recall:* The recall of the enrichment, calculated as the ratio of intersection_size to term_size. Recall measures the proportion of genes associated with the GO term that are present in the reference set.
term_id: The identifier of the GO term or other ontology term.
source: The source of the term, such as "GO:BP" for Gene Ontology Biological Process, "GO:MF" for Gene Ontology Molecular Function, or "GO:CC" for Gene Ontology Cellular Component.
term_name: The name of the GO term or other ontology term.
effective_domain_size: The effective domain size, which represents the number of genes in the background set that are considered for the enrichment analysis.
source_order: The order of the source in the results, which can be used for sorting or prioritizing results.
parents: The parent terms of the GO term or other ontology term. This provides hierarchical context for the term.
highlighted: A logical value (TRUE or FALSE) indicating whether the term is highlighted in the results. This can be used for visualization or emphasis in the output.
Run GO analysis for hsapiens genes

```{r}
# Load correlations between SNPs and variables
cors <- 
  read_csv(here(outpath, "bio1ndvi_genes_cortest.csv")) %>%
  # Drop loci without significant correlations with NDVI or BIO1
  drop_na(group) %>%
  dplyr::select(locus, group) %>%
  # [TODO: FIX] Some loci are currently duplicated so right now i am just taking the first one [NEED TO FIX]
  group_by(locus) %>%
  dplyr::slice(1) %>%
  ungroup()
```

### hsapiens
```{r}
if (!file.exists(here("analysis", "gea", "outputs", "all_genes_hsapiens.csv"))) {
  go_result_all <-
    gost(
      query = all_genes_hsapiens$query, 
      organism = "hsapiens", ordered_query = FALSE, 
      multi_query = FALSE, significant = FALSE, exclude_iea = FALSE, 
      measure_underrepresentation = FALSE, evcodes = FALSE, 
      user_threshold = 0.05, correction_method = "g_SCS", 
      domain_scope = "annotated", custom_bg = NULL, 
      numeric_ns = "", sources = NULL, as_short_link = FALSE, highlight = TRUE
    )
  
  write_csv(go_result_all$result, here("analysis", "gea", "outputs", "all_genes_go_hsapiens.csv"))
}

all_genes_go_hsapiens <- 
  read_csv(here("analysis", "gea", "outputs", "all_genes_go_hsapiens.csv")) %>% 
  as_tibble() %>%
  # Just keep GO terms
  filter(grepl("GO", source)) %>%
  dplyr::select(term_id, intersection_size, query_size, term_size, p_value)

# Add _all to the names of the columns
names(all_genes_go_hsapiens)[-1] <- paste0(names(all_genes_go_hsapiens), "_all")[-1]
```

```{r}
# Link genes and snps
rda_hsapiens_genes <-
  # file from intersect_genes.R
  read_csv(here("analysis", "gea", "outputs", "bio1ndvi_gea_gene_snp.csv")) %>%
  dplyr::select(locus, original_entry = full_name) %>%
  # Join with genes
  left_join(dplyr::select(genes_org, original_entry, uniprot_id)) %>%
  drop_na(uniprot_id) %>%
  # Join with orthologs
  left_join(genes_hsapiens, by = "uniprot_id") %>%
  # Some genes don't have hsapiens orthologs/query names, so drop them
  drop_na(query) %>%
  # Join with the correlation results 
  # DISTINCT IS TEMPORARY TO DEAL WITH DUPLICATES (LUCKILY GROUPS ARE SAME FOR DUPLICATES)
  # Drop group NA to remove loci without significant correlations with NDVI or BIO1
  left_join(cors, by = "locus") 

bio1_query_hsapiens <- rda_hsapiens_genes %>% filter(grepl("bio1", group)) %>% pull(query)
length(bio1_query_hsapiens)
ndvi_query_hsapiens <- rda_hsapiens_genes %>% filter(grepl("ndvi", group)) %>% pull(query)
length(ndvi_query_hsapiens)
both_query_hsapiens <- rda_hsapiens_genes %>% pull(query)

go_bio1_hsapiens <- run_go(bio1_query_hsapiens, "hsapiens")
go_ndvi_hsapiens <- run_go(ndvi_query_hsapiens, "hsapiens")
go_both_hsapiens <- run_go(both_query_hsapiens, "hsapiens")

sig_bio1_hsapiens <- go_correction(go_bio1_hsapiens, all_genes_go_hsapiens)
sig_ndvi_hsapiens <- go_correction(go_ndvi_hsapiens, all_genes_go_hsapiens)
sig_both_hsapiens <- go_correction(go_both_hsapiens, all_genes_go_hsapiens)
# TODO: check out if  HSP genes are associated with BIO1/NDVI or both
```

```{r}
go_topbp(sig_bio1_hsapiens)
go_topbp(sig_ndvi_hsapiens)
go_topbp(sig_both_hsapiens)
```

## ggallus

```{r}
if (!file.exists(here("analysis", "gea", "outputs", "all_genes_go_ggallus.csv"))) {
  go_result_all <-
    gost(
      query = all_genes_ggallus$query, 
      organism = "ggallus", ordered_query = FALSE, 
      multi_query = FALSE, significant = FALSE, exclude_iea = FALSE, 
      measure_underrepresentation = FALSE, evcodes = FALSE, 
      user_threshold = 0.05, correction_method = "g_SCS", 
      domain_scope = "annotated", custom_bg = NULL, 
      numeric_ns = "", sources = NULL, as_short_link = FALSE, highlight = TRUE
    )
  
  write_csv(go_result_all$result, here("analysis", "gea", "outputs", "all_genes_go_ggallus.csv"))
}

all_genes_go_ggallus <- 
  read_csv(here("analysis", "gea", "outputs", "all_genes_go_ggallus.csv")) %>% 
  as_tibble() %>%
  # Just keep GO terms
  filter(grepl("GO", source)) %>%
  dplyr::select(term_id, intersection_size, query_size, term_size, p_value)

names(all_genes_go_ggallus)[-1] <- paste0(names(all_genes_go_ggallus), "_all")[-1]
```

```{r}
# Link genes and snps
rda_ggallus_genes <-
  # file from intersect_genes.R
  read_csv(here("analysis", "gea", "outputs", "bio1ndvi_gea_gene_snp.csv")) %>%
  dplyr::select(locus, original_entry = full_name) %>%
  # Join with genes
  left_join(dplyr::select(genes_org, original_entry, uniprot_id)) %>%
  drop_na(uniprot_id) %>%
  # Join with orthologs
  left_join(genes_ggallus, by = "uniprot_id") %>%
  # Some genes don't have human orthologs/query names, so drop them
  drop_na(query) %>%
  # Join with the correlation results 
  # DISTINCT IS TEMPORARY TO DEAL WITH DUPLICATES (LUCKILY GROUPS ARE SAME FOR DUPLICATES)
  # Drop group NA to remove loci without significant correlations with NDVI or BIO1
  left_join(cors, by = "locus") 


bio1_query_ggallus <- rda_ggallus_genes %>% filter(grepl("bio1", group)) %>% pull(query)
length(bio1_query_ggallus)
ndvi_query_ggallus <- rda_ggallus_genes %>% filter(grepl("ndvi", group)) %>% pull(query)
length(ndvi_query_ggallus)
both_query_ggallus <- rda_ggallus_genes %>% pull(query)

go_bio1_ggallus <- run_go(bio1_query_ggallus, "ggallus")
go_ndvi_ggallus <- run_go(ndvi_query_ggallus, "ggallus")
go_both_ggallus <- run_go(both_query_ggallus, "ggallus")

sig_bio1_ggallus <- go_correction(go_bio1_ggallus, all_genes_go_ggallus)
sig_ndvi_ggallus <- go_correction(go_ndvi_ggallus, all_genes_go_ggallus)
sig_both_ggallus <- go_correction(go_both_ggallus, all_genes_go_ggallus)
```

# Compare results
```{r}
bp_bio1_hsapiens <- sig_bio1_hsapiens %>% filter(grepl("GO:BP", source))
bp_ndvi_hsapiens <- sig_ndvi_hsapiens %>% filter(grepl("GO:BP", source))
bp_both_hsapiens <- sig_both_hsapiens %>% filter(grepl("GO:BP", source))
bp_bio1_ggallus <- sig_bio1_ggallus %>% filter(grepl("GO:BP", source))
bp_ndvi_ggallus <- sig_ndvi_ggallus %>% filter(grepl("GO:BP", source))
bp_both_ggallus <- sig_both_ggallus %>% filter(grepl("GO:BP", source))

go_topbp(sig_bio1_hsapiens)
go_topbp(sig_ndvi_hsapiens)
go_topbp(sig_both_hsapiens)
go_topbp(sig_bio1_ggallus)
go_topbp(sig_ndvi_ggallus)
go_topbp(sig_both_ggallus)

library(gt)
gtsave(go_table(bp_bio1_hsapiens), here(plotpath, "bp_bio1_hsapiens.html"))
gtsave(go_table(bp_ndvi_hsapiens), here(plotpath, "bp_ndvi_hsapiens.html"))
gtsave(go_table(bp_both_hsapiens), here(plotpath, "bp_both_hsapiens.html"))
gtsave(go_table(bp_ndvi_ggallus), here(plotpath, "bp_ndvi_ggallus.html"))
gtsave(go_table(bp_bio1_ggallus), here(plotpath, "bp_bio1_ggallus.html"))
gtsave(go_table(bp_bio1_ggallus), here(plotpath, "bp_bio1_ggallus.html"))


gtsave(go_table(arrange(bp_both_hsapiens, desc(precision))), here(plotpath, "bp_both_hsapiens.html"))
gtsave(go_table(arrange(bp_both_hsapiens, fish_p)), here(plotpath, "bp_both_hsapiens.html"))
gtsave(go_table(arrange(bp_both_hsapiens, desc(recall)), n = 10), here(plotpath, "bp_both_hsapiens.html"))


gtsave(go_table(arrange(bp_both_ggallus, desc(precision))), here(plotpath, "bp_both_ggallus.html"))
gtsave(go_table(arrange(bp_both_ggallus, fish_p)), here(plotpath, "bp_both_ggallus.html"))
gtsave(go_table(arrange(bp_both_ggallus, desc(recall)), n = 10), here(plotpath, "bp_both_ggallus.html"))

gtsave(go_table(arrange(bp_bio1_ggallus, desc(recall)), n = 10), here(plotpath, "bp_bio1_ggallus.html"))
gtsave(go_table(arrange(bp_ndvi_ggallus, desc(recall)), n = 10), here(plotpath, "bp_ndvi_ggallus.html"))

```
