```{r}
library(terra)
library(here)
library(sf)
library(tidyverse)
source(here("general_functions.R"))
source(here("analysis", "genetic_diversity", "functions_genetic_diversity.R"))
source(here("analysis", "gea", "functions_selection_stats.R"))

coords <- get_coords(sf = TRUE)
plotpath <- here("analysis", "gea", "plots")
adaptive_path <- here("analysis", "adaptive", "outputs")

adapind_load <- rast(here(adaptive_path, "58-Sceloporus_bio1ndvi_gea_adapind_load.tif"))
offset <- rast(here(adaptive_path, "58-Sceloporus_bio1ndvi_gea_offset_load_RCP85.tif"))
offset_total <- sum(offset)
adapind_rgb <- map(1:nlyr(adapind_load), function(i) {
  layer <- adapind_load[[i]]
  stats <- global(layer, fun = range, na.rm = TRUE)
  vmin <- stats[1, 1]
  vmax <- stats[1, 2]
  ((layer - vmin) / (vmax - vmin)) * 255
}) %>% rast()

adapind_rgb <- c(adapind_rgb, null_lyr = adapind_rgb[[1]] * 0)
names(adapind_rgb) <- c("RDA1", "RDA2", "NULL")

rgb_vals <- drop_na(terra::extract(adapind_rgb, coords, ID = FALSE))
adapind_vals <- terra::extract(adapind_load, coords, ID = FALSE)
names(adapind_vals) <- c("RDA1", "RDA2")
rgb_df <- bind_cols(coords, adapind_vals) %>% drop_na()
rgb_df$rgb <- rgb(red = rgb_vals[,1], green = rgb_vals[,3], blue = rgb_vals[,2], maxColorValue = 255)

rda_varloadings <- read_csv(here("analysis", "adaptive", "outputs", "RDA_biplot_58-Sceloporus_bio1ndvi_gea_mod.csv"))

plot_rgb <- function(df) {
  ggplot(df, aes(x = RDA1, y = RDA2)) +
    annotate("text", x = max(df$RDA1) + 0.5, y = 0, label = "RDA1", hjust = 1.2, vjust = -0.5, size = 3, col = "darkgray") +
    annotate("text", x = 0, y = max(df$RDA2) + 0.5, label = "RDA2", hjust = -0.2, vjust = 1.2, size = 3, col = "darkgray") +
    geom_hline(yintercept = 0, linewidth = 0.2, col = "gray") +
    geom_vline(xintercept = 0, linewidth = 0.2, col = "gray") +
    geom_point(aes(col = rgb)) +
    geom_text(data = rda_varloadings, aes(x = RDA1 * 2.5, y = RDA2 * 2.5, label = var), size = 4, vjust = 1) +
    geom_segment(data = rda_varloadings, aes(x = 0, y = 0, xend = RDA1 * 2.5, yend = RDA2 * 2.5),
                 arrow = arrow(length = unit(0.2, "cm"))) +
    scale_color_identity() +
    scale_x_continuous(expand = expansion(mult = 0.1)) +  # 5% margin
    scale_y_continuous(expand = expansion(mult = 0.1)) +
    coord_equal() +
    theme_void()


pdf(here(plotpath, "adaptive_index.pdf"), width = 5, height = 5)
plot_rgb(rgb_df) 
plotRGB(adapind_rgb, r = 1, g = 3, b = 2, axes = FALSE)
plotRGB(adapind_rgb, r = 1, g = 3, b = 2, axes = FALSE)
plotRGB(adapind_rgb, r = 1, g = 3, b = 2, axes = FALSE)
points(coords, pch = 1)
plotRGB(adapind_rgb, r = 1, g = 3, b = 3, axes = FALSE) # RDA1
plotRGB(adapind_rgb, r = 3, g = 3, b = 2, axes = FALSE) # RDA2
plot(offset, main = "Adaptive index offset", axes = FALSE)
plot(offset_total, main = "Adaptive index offset (total)", axes = FALSE, col = viridis::inferno(100))

plot(adapind_rgb[[1]], col = colorRampPalette(c("black", "red"))(4))
plot(adapind_rgb[[2]], col = colorRampPalette(c("black", "blue"))(4))
dev.off()

pdf(here(plotpath, "adaptive_index_test.pdf"), width = 5, height = 5)
plotRGB(adapind_rgb, r = 1, g = 2, b = 3, main = "Adaptive index loadings", axes = FALSE)
plotRGB(adapind_rgb, r = 1, g = 3, b = 2, main = "Adaptive index loadings", axes = FALSE)
plotRGB(adapind_rgb, r = 3, g = 1, b = 2, main = "Adaptive index loadings", axes = FALSE)
plotRGB(adapind_rgb, r = 2, g = 1, b = 3, main = "Adaptive index loadings", axes = FALSE)
plotRGB(adapind_rgb, r = 2, g = 3, b = 1, main = "Adaptive index loadings", axes = FALSE)
plotRGB(adapind_rgb, r = 2, g = 1, b = 1, main = "Adaptive index loadings", axes = FALSE)
plot(adapind_load, main = "Adaptive index loadings", axes = FALSE)
plot(offset, main = "Adaptive index offset", axes = FALSE)
plot(offset_total, main = "Adaptive index offset (total)", axes = FALSE, col = viridis::rocket(100))
dev.off()

```


```{}
```{r}
coords <- get_coords(sf = TRUE)

genes <- 
  format_het(here("analysis", "gea", "outputs", "genes.het"), callable_sites = 1) %>%
  mutate(Ho_genes = 1 - `O(HOM)`/`N(NM)`) %>%
  dplyr::rename(SampleID = IID) %>%
  dplyr::select(SampleID, Ho_genes) %>%
  left_join(get_het())

pop_df <- get_pops() %>% mutate(pop = gsub("pop", "", cluster))

pop_cols <- get_pop_cols()

offset_df <- 
  terra::extract(offset_total, coords, ID = FALSE) %>%
  mutate(SampleID = coords$SampleID) %>%
  as_tibble() %>%
  dplyr::rename(offset = sum) %>%
  left_join(genes) %>%
  drop_na(Ho, Ho_genes) %>%
  mutate(resids = residuals(lm(Ho_genes ~ Ho, data = .))) %>%
  left_join(pop_df) %>%
  drop_na(pop)


#devtools::source_url("https://raw.githubusercontent.com/austinwpearce/SoilTestCocaCola/main/quad_plateau.R")

offset_mean <- 
  offset_df %>% 
  group_by(pop) %>% 
  summarise(
    Ho_sd = sd(Ho, na.rm = TRUE),
    offset_sd = sd(offset, na.rm = TRUE),
    offset = mean(offset, na.rm = TRUE), 
    Ho = mean(Ho, na.rm = TRUE)
  ) %>%
  drop_na(pop) %>%
  mutate(
    Ho_min = Ho - Ho_sd,
    Ho_max = Ho + Ho_sd,
    offset_min = offset - offset_sd,
    offset_max = offset + offset_sd
  )

pdf(here(plotpath, "offset_vs_ho.pdf"), width = 4, height = 4)
ggplot(offset_df, aes(y = (offset), x = (Ho))) +
  geom_point(aes(col = pop), shape = 16, size = 1) +
  geom_point(data = offset_mean, aes(col = pop), size = 3, shape = 8) +
  geom_smooth(method = "lm", col = "black") +
  xlab("Genome-wide heterozygosity") +
  ylab("Future genomic offset") +
  scale_color_manual(values = pop_cols) +
  ggpubr::stat_cor() +
  theme_classic() +
  theme(legend.position = "none")
ggplot(offset_mean, aes(y = (offset), x = (Ho))) +
  #geom_errorbar(aes(ymin = offset_min, ymax = offset_max, xmin = Ho_min, xmax = Ho_max), width = 0.1) +
  geom_point(aes(col = pop)) +
  geom_smooth(method = "lm", col = "black") +
  xlab("Genome-wide heterozygosity") +
  ylab("Future genomic offset") +
  scale_color_manual(values = pop_cols) +
  ggpubr::stat_cor() +
  theme_classic() +
  theme(legend.position = "none")
dev.off()

ggplot(offset_df, aes(y = (offset), x = Ho_genes, col = offset)) +
  geom_point() +
  geom_smooth(method = "lm", col = "black") +
  xlab("Adaptive heterozygosity") +
  ylab("Future genomic offset") +
  ggpubr::stat_cor() +
  scale_color_viridis_c(option = "inferno") +
  theme_classic()+
  theme(legend.position = "none")

ggplot(offset_df, aes(y = (offset), x = resids, col = offset)) +
  geom_point() +
  geom_smooth(method = "lm", col = "black") +
  xlab("Residuals of adaptive heterozygosity") +
  ylab("Future genomic offset") +
  ggpubr::stat_cor() +
  scale_color_viridis_c(option = "inferno") +
  theme_classic()+
  theme(legend.position = "none")
dev.off()
```

```{r}
library(biscale)
library(cowplot)
# Step 1: Classify into bivariate groups
offset_df_bi <- 
  bi_class(offset_df, x = Ho, y = offset, style = "quantile", dim = 3) %>%
  left_join(get_coords(sf = TRUE)) %>%
  st_as_sf()

ca <- get_ca()
# Step 2: Create the scatterplot with bi_class fill
map <- 
  ggplot(offset_df_bi) +
  geom_sf(data = ca, fill = "lightgray") +
  geom_sf(aes(col = bi_class), size = 3, alpha = 1) +  # use color for points
  bi_scale_color(pal = "GrPink", dim = 3) +  # use color instead of fill for points
  theme_void()+ # clean theme from biscale +
  theme(legend.position = "none")

# Step 3: Create the legend
legend <- bi_legend(
  pal = "GrPink",
  dim = 3,
  xlab = "Higher Heterozygosity →",
  ylab = "↑ Higher Genomic Offset",
  size = 8
)

# Step 4: Combine plot and legend
final_plot <- 
  ggdraw() +
  draw_plot(map, 0, 0, 1, 1) +
  draw_plot(legend, x = 0.6, y = 0.6, width = 0.35, height = 0.34)

# Show the result
pdf(here(plotpath, "offset_bivariate.pdf"), width = 8, height = 6)
print(final_plot)
dev.off()
```
```{r}
source(here("analysis", "adaptive", "adaptive_index.R"))
bio1ndvi_envlayers <- import_env_files(type = "ind_layers", future = TRUE)
change = (bio1ndvi_envlayers$env_fut$CHELSA_bio1_2071.2100_gfdl.esm4_ssp585_V.2.1/10 - 273.15) -  bio1ndvi_envlayers$env_pres$BIO1 
pdf(here(plotpath, "env_future.pdf"), width = 10, height = 10)
plot(change, main = "Future environmental change", axes = FALSE, col = viridis::inferno(100))
dev.off()


change_df <- coords %>% mutate(change = terra::extract(change, coords, ID = FALSE)) %>% left_join(genes)

pdf(here(plotpath, "change_vs_env.pdf"), width = 4, height = 4)

plot(change, main = "Future environmental change", axes = FALSE, col = viridis::inferno(100))
ggplot(change_df, aes(y = (change), x = Ho_genes, col = change)) +
  geom_point() +
  geom_smooth(method = "lm", col = "black") +
  xlab("Adaptive heterozygosity") +
  ylab("Future temperature change") +
  ggpubr::stat_cor() +
  scale_color_viridis_c(option = "inferno") +
  theme_classic()+
  theme(legend.position = "none")
dev.off()
```
