```{r}
library(here)
library(tidyverse)
library(sf)
library(terra)
library(ggpubr)
library(algatr)
library(viridis)
library(tigris)
library(raster)
library(igraph)
library(sp)
library(gridExtra)
library(wingen)
library(furrr)
source(here("analysis", "chromenv", "functions_chromenv.R"))
source(here("analysis", "gea", "functions_selection_stats.R"))
source(here("general_functions.R"))
outpath <- here("analysis", "gea", "outputs")
plotpath <- here("analysis", "gea", "plots")

#  Number of RDA + Linked SNPs: 958850
#  Number of SNPs in genes: 396217
# Number of non-synonymous SNPs in genes: 20344

396217/958850
20344/958850 * 5.81
20344/396217 * 5.81

```

```{r}
# GEA
gea <- get_gea_genes(nonsyn = FALSE) %>% distinct()
gea_nonsyn <- get_gea_genes(nonsyn = TRUE) %>% distinct() # FIGURE OUT HOW TO DEAL WITH DUPLICATES: GNX-391/GNX-359 (overlapping genes, same SNP)

# Tajima's D
tjd <- get_tajimad(windowkb = 10) 

# Non-synonymous SNPs
nonsyn <- get_nonsyn() %>% rename(start = position) # NOTE: position = SNP position= start

# Chromenv
chromenv <- get_chromenv() %>% filter(p_adj < 0.05) %>% filter(var %in% c("csi_past", "tmean_dif"))
chromenv_neg <- chromenv %>% filter(estimate < 0) %>% filter(var %in% c("csi_past", "tmean_dif"))

# CHECKING BOTTOM 5% CALCULATION
tjd_neg <- tjd %>% group_by(pop) %>% mutate(bottom = TajimaD < quantile(TajimaD, 0.05, na.rm = TRUE))
tjd_neg %>% ungroup() %>% filter(bottom) %>% summarize(mean(outliers == "Bottom 5%")) == 1
tjd_neg %>% ungroup() %>% filter(!bottom) %>% summarize(mean(outliers == "Bottom 5%", na.rm = TRUE)) == 0
# Using bottom and top 10% 
tjd_neg <- tjd %>% group_by(pop) %>% mutate(bottom = TajimaD < quantile(TajimaD, 0.10, na.rm = TRUE))
tjd_neg <- tjd_neg %>% ungroup() %>% filter(bottom)
tjd_pos <- tjd %>% group_by(pop) %>% mutate(top = TajimaD > quantile(TajimaD, 0.10, na.rm = TRUE))
tjd_pos <- tjd_pos %>% ungroup() %>% filter(top)

tjd_quant <- tjd %>% group_by(pop) %>% summarize(q10 = quantile(TajimaD, 0.10, na.rm = TRUE))
```

```{r}
plan(multisession, workers = 10)
combo_df <- 
  gea_nonsyn %>%
  # # NOTE: some loci fall in more than one gene, so using distinct() to pull out unique set of loci
  distinct(locus, start, scaffold) %>%
  group_by(locus) %>%
  group_split() %>%
  future_map(~{
    # Get position and scaffold
    start <- .x$start
    scaffold <- .x$scaffold

    # ...is in Bottom 5% of Tajima's D
    tjd_lgl <- tjd_neg %>% filter(scaffold == !!scaffold, BIN_START <= start, BIN_END >= start) %>% nrow(.) > 0

    # ...is in a region with negative association between diversity and an env variable
    chromenv_neg_lgl <- chromenv_neg %>% filter(scaffold == !!scaffold, BIN_START <= start, BIN_END >= start) %>% nrow(.) > 0

    # ...is in a region with any association between diversity and an env variable
    chromenv_lgl <- chromenv %>% filter(scaffold == !!scaffold, BIN_START <= start, BIN_END >= start) %>% nrow(.) > 0

    # Combine into df
    data.frame(locus = .x$locus, tjd = tjd_lgl, chromenv = chromenv_lgl, chromenv_neg = chromenv_neg_lgl)
  }, .progress = TRUE) 
plan(sequential)

combo_df <- bind_rows(combo_df)

summary_df <-
  combo_df %>%
  bind_rows() %>%
  summarize(
    both_count = sum(tjd & chromenv_neg), 
    tjd_count = sum(tjd & !chromenv_neg), 
    chromenv_count = sum(chromenv_neg & !tjd), 
    neither_count = sum(!tjd & !chromenv_neg),
    both_mean = mean(tjd & chromenv_neg),
    tjd_mean = mean(tjd & !chromenv_neg),
    chromenv_mean = mean(chromenv_neg & !tjd),
    neither_mean = mean(!tjd & !chromenv_neg)
  ) %>%
  mutate(across(ends_with("_mean"), ~round(.x * 100, 0)))
summary_df

gea_combo_df <- left_join(gea_nonsyn, combo_df, by = "locus")

goi <- get_goi(type = "genes")
goi_combo <- gea_combo_df %>% filter(full_name %in% goi$full_name) 

# Look at all SNPs with very negative TJD & chromenv
gea_combo_df %>% 
  filter(tjd & chromenv_neg) %>% 
  distinct(name) %>%
  drop_na()

# Look at all SNPs with very negative TJD
gea_combo_df %>% 
  filter(tjd) %>% 
  distinct(name) %>%
  drop_na()

# Look at genes of interests: 
# No genes of interest have both negative TJD and chromenv
goi_combo %>% filter(tjd & chromenv_neg) %>% dplyr::select(scaffold, name, tjd, chromenv_neg)

# Many HSP have negative TJD
goi_combo %>% filter(tjd | chromenv_neg) %>% distinct(name, tjd, chromenv_neg) %>% arrange(chromenv_neg, tjd)

# TPRM8 is not in a negative tajima's D region
goi_combo %>% filter(grepl("TRPM8", full_name))
```
```

```{r}
# Get full information about associations with SNPs of interest
# For chromenv - get the variable with the negative assocation
# For TJD - get the population with the negative association
plan(multisession, workers = 10)
details <- 
  gea_combo_df %>%
  # # NOTE: some loci fall in more than one gene, so using distinct() to pull out unique set of loci
  distinct(locus, start, scaffold) %>%
  group_by(locus) %>%
  group_split() %>%
  future_map(~{
    # Get position and scaffold
    start <- .x$start
    scaffold <- .x$scaffold

    # Get region with association
    result1 <-
      chromenv %>% 
      filter(scaffold == !!scaffold, BIN_START <= start, BIN_END >= start) %>% 
      dplyr::select(var, estimate) %>%
      mutate(locus = .x$locus)
    
    # Get TJD
    result2 <-
      tjd %>% 
      filter(scaffold == !!scaffold, BIN_START <= start, BIN_END >= start) %>% 
      dplyr::select(pop, TajimaD) %>%
      mutate(locus = .x$locus)

    if (nrow(result1) == 0) result1 <- NULL
    if (nrow(result2) == 0) result2 <- NULL
    
    return(list(chromenv = result1, tjd = result2))
    # Combine
  }, .progress = TRUE) 
plan(sequential)

# Pull out details
chromenv_var <- map(details, "chromenv") %>% compact() %>% bind_rows()
tjd_pop <- map(details, "tjd") %>% compact() %>% bind_rows()

# Chromenv:
left_join(goi_combo, chromenv_var) %>% 
  drop_na(var) %>% 
  dplyr::select(var, name, estimate)
## Tmean_dif:
left_join(gea_combo_df, chromenv_var, relationship = "many-to-many") %>% 
  drop_na(name) %>% 
  filter(estimate < 0) %>%
  filter(var == "tmean_dif") %>% 
  distinct(var, name)

# TJD
## TJD for all SNPs
all_tjd <- left_join(gea_combo_df, tjd_pop, relationship = "many-to-many")
all_tjd %>% 
  drop_na(name) %>% 
  filter(TajimaD < 0) %>%
  distinct(TajimaD, name, pop)

## TJD for GOI
goi_tjd <- left_join(goi_combo, tjd_pop)
goi_tjd %>% 
  drop_na(TajimaD) %>% 
  filter(TajimaD < 0) %>%
  dplyr::select(pop, name, TajimaD) %>% 
  left_join(tjd_quant)

goi_tjd <-  left_join(goi_combo, tjd_pop)
hsp <- 
  goi_tjd %>% 
  drop_na(TajimaD) %>% 
  dplyr::select(pop, name, TajimaD) %>% 
  left_join(tjd_quant) %>%
  filter(grepl("Heat shock", name)) # Most of the negative Tajima's D are not in pop9
```

Run models of Tajima's D vs environmental variables:
```{r}
pop_tmean_dif <- 
  read_csv(here("analysis", "genetic_diversity", "outputs", "model_df.csv")) %>% 
  left_join(get_pops()) %>%
  group_by(cluster) %>%
  summarize(across(c("NDVI", "glacier", "tmean_dif", "gHM", "bio1", "csi_past"), \(x) mean(x, na.rm = TRUE))) %>%
  mutate(pop = paste0("pop", as.character(cluster))) %>%
  dplyr::select(-cluster) %>%
  ungroup() %>%
  mutate(across(c("NDVI", "glacier", "tmean_dif", "gHM", "bio1", "csi_past"), \(x) scale(x, center = TRUE, scale = TRUE)))

mod_df <- 
  left_join(all_tjd, pop_tmean_dif, by = "pop") %>% 
  distinct(name, pop, TajimaD, NDVI, glacier, tmean_dif, gHM, bio1, csi_past)

lm_result <- 
  mod_df %>%
  group_by(name) %>%
  group_split() %>%
  map(~{
    if (drop_na(.x, TajimaD, NDVI, glacier, tmean_dif, gHM, bio1, csi_past) %>% nrow() == 0) return(NULL)
    lm(TajimaD ~ NDVI + glacier + tmean_dif + gHM + bio1 + csi_past , data = .x) %>% 
      broom::tidy() %>% 
      filter(term != "(Intercept)") %>%
      mutate(name = unique(.x$name))
  }, .progress = TRUE) %>%
  bind_rows()

lm_result %>% filter(name %in% goi_combo$name) %>% filter(p.adjust(p.value, "fdr") < 0.05) #nada

# A lot of significants
lm_result %>% filter(p.adjust(p.value, "fdr") < 0.05) 

# Look for negative associations with tmean_dif
lm_result %>% 
  filter(p.adjust(p.value, "fdr") < 0.05)  %>% 
  filter(estimate < 0, term == "tmean_dif") %>% 
  drop_na(name) %>%
  arrange(estimate) %>%
  dplyr::select(name, term, estimate, p.value)

# Look for negative associations with csi_past
lm_result %>% 
  filter(p.adjust(p.value, "fdr") < 0.05)  %>% 
  filter(estimate < 0, term == "csi_past") %>% 
  drop_na(name) %>%
  arrange(estimate) %>%
  dplyr::select(name, term, estimate, p.value)

```

# Exploration of SNPs/genes of interest
```{r}
# Count different kinds of non-synonymous SNPs
nonsyn %>% group_by(csq) %>% count() %>% arrange(desc(n))

# Count different kinds of non-synonymous GEA SNPs
gea_nonsyn %>% dplyr::select(scaffold, start) %>% left_join(nonsyn) %>% count(csq) %>% arrange(desc(n))

# Each SNP has one annotation
nonsyn %>% filter(scaffold == "chr2") %>% group_by(scaffold, start) %>% count() %>% arrange(desc(n))

itpr1_snps <- 
  gea_combo_df %>% 
  filter(name == "Inositol 1,4,5-trisphosphate receptor type 1 ") %>% 
  distinct(scaffold, start) %>%
  left_join(nonsyn, by = c("start", "scaffold"))

itpr1_snps %>% count(csq) %>% arrange(desc(n))

tjd_pop9 <- 
  tjd %>% 
  filter(pop == "pop9", scaffold == "chr2") %>% 
  filter(BIN_START >= min(itpr1_snps$start) - 500000, BIN_END <= max(itpr1_snps$start) + 500000)

itpr1_gene <-
  all_genes %>% 
  filter(grepl("Inositol 1,4,5-trisphosphate receptor type 1 ", full_name)) %>% 
  distinct(scaffold, start, end) %>%
  filter(scaffold == "chr2")

tjd_quant
pdf(here(plotpath, "itpr1_pop9.pdf"), width = 10, height = 3)
ggplot(tjd_pop9) +
  geom_rect(data = itpr1_gene, aes(xmin = start, xmax = end, fill = "ITPR1"), ymin = -Inf, ymax = Inf) +
  geom_vline(data = itpr1_snps, aes(xintercept = start, col = csq)) +
  geom_line(aes(x = BIN_START, y = TajimaD), alpha = 0.6, lwd = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", lwd = 0.5, col = "black") +
  geom_hline(yintercept = -1.13, linetype = "dashed", lwd = 0.5, col = "red") +
  scaffold_theme() +
  scale_x_continuous(expand = c(0, 0), breaks = seq(min(tjd_pop9$BIN_START), max(tjd_pop9$BIN_START), by = 500000), labels = scales::comma) +
  ggtitle("Tajima's D (pop9)") +
  labs(fill = "Gene", color = "Non-synonymous GEA SNPs") +
  scale_color_manual(values = c("purple2", "gray", "green2", "cyan3")) +
  theme(axis.text.x = element_text(), axis.ticks.x = element_line(), axis.line = element_line())
dev.off()

```

```{r}
pop9_top_locus <-
 goi_tjd %>% 
 filter(pop == "pop9") %>% 
 left_join(gea_combo_df) %>% 
 drop_na(uniprotid) %>%
 arrange(TajimaD) %>%
 head(10)

pop9_top_locus %>% dplyr::select(name, scaffold)

top_locus = "Heat shock protein beta-1 " 
top_snps <- 
  gea_combo_df %>% 
  filter(name == top_locus) %>% 
  distinct(scaffold, start) %>%
  left_join(nonsyn, by = c("start", "scaffold"))

top_snps %>% count(csq) %>% arrange(desc(n))

tjd_pop9 <- 
  tjd %>% 
  filter(pop == "pop9", scaffold == "chr7") %>% 
  filter(BIN_START >= min(top_snps$start) - 500000, BIN_END <= max(top_snps$start) + 500000)

top_gene <-
  all_genes %>% 
  filter(grepl(top_locus, full_name)) %>% 
  distinct(scaffold, start, end) %>%
  filter(scaffold == "chr7")

pdf(here(plotpath, "top_gene_pop9.pdf"), width = 10, height = 3)
ggplot(tjd_pop9) +
  geom_rect(data = top_gene, aes(xmin = start, xmax = end, fill = !!top_locus), ymin = -Inf, ymax = Inf) +
  geom_vline(data = top_snps, aes(xintercept = start, col = csq)) +
  geom_line(aes(x = BIN_START, y = TajimaD), alpha = 0.6, lwd = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", lwd = 0.5, col = "black") +
  geom_hline(yintercept = -1.13, linetype = "dashed", lwd = 0.5, col = "red") +
  scaffold_theme() +
  scale_x_continuous(expand = c(0, 0), breaks = seq(min(tjd_pop9$BIN_START), max(tjd_pop9$BIN_START), by = 500000), labels = scales::comma) +
  ggtitle("Tajima's D (pop9)") +
  labs(fill = "Gene", color = "Non-synonymous GEA SNPs") +
  scale_color_manual(values = c("purple2", "gray", "green2", "cyan3")) +
  theme(axis.text.x = element_text(), axis.ticks.x = element_line(), axis.line = element_line())
dev.off()

```











```{r}
plotchrom_c <- function(data, fill){
  scaff_rect <- get_scaff_rect(data)
  ggplot(data) +
    geom_rect(aes(xmin = BIN_START, xmax = BIN_END, ymin = 0, ymax = 1, fill = .data[[fill]])) +
    geom_rect(data = scaff_rect, aes(xmin = BIN_START, xmax = BIN_END), ymin = 0, ymax = 1, col = "black", fill = NA) +
    scaffold_theme() +
    scale_fill_viridis_c(option = "magma")
}

tjd <- get_tajimad()

goi <- get_goi() %>% filter(scaffold %in% unique(tjd$scaffold)) %>% mutate(scaffold = factor(scaffold, levels = unique(tjd$scaffold)))
```
```{r}
# Get genes
genes <- 
  get_genes() %>%
  mutate(BIN_START = floor(start / 500000) * 500000) %>%
  mutate(BIN_END = BIN_START + 500000) %>%
  filter(scaffold %in% unique(sem_df$scaffold)) %>%
  mutate(scaffold = factor(scaffold, levels = unique(sem_df$scaffold)))

genes_bin <-
  genes %>%
  group_by(scaffold, BIN_START, BIN_END) %>%
  summarise(ngenes = n()) %>%
  mutate(scaffold = factor(scaffold, levels = unique(sem_df$scaffold)))

# Create bins for BIN_START
nonsyn <- 
  get_nonsyn() %>%
  mutate(BIN_START = floor(position / 500000) * 500000)  %>%
  mutate(BIN_END = BIN_START + 500000) %>%
  filter(scaffold %in% unique(sem_df$scaffold)) %>%
  mutate(scaffold = factor(scaffold, levels = unique(sem_df$scaffold)))

# Calculate number of nonsynonymous mutations per bin of 100k
nonsyn_bin <- 
  nonsyn %>% 
  group_by(scaffold, BIN_START, BIN_END) %>% 
  summarise(nonsyn = n()) %>%
  filter(scaffold %in% unique(sem_df$scaffold))

# Create bins for BIN_START
syn <- 
  get_syn() %>%
  mutate(BIN_START = floor(position / 500000) * 500000)  %>%
  mutate(BIN_END = BIN_START + 500000) %>%
  filter(scaffold %in% unique(sem_df$scaffold)) %>%
  mutate(scaffold = factor(scaffold, levels = unique(sem_df$scaffold)))

# Calculate number of nonsynonymous mutations per bin of 100k
syn_bin <- 
  syn %>% 
  group_by(scaffold, BIN_START, BIN_END) %>% 
  summarise(syn = n()) %>%
  filter(scaffold %in% unique(sem_df$scaffold))

syn_nonsyn <-
  full_join(syn_bin, nonsyn_bin, by = c("scaffold", "BIN_START", "BIN_END")) %>%
  # DON"T MAKE SYN 0 OR YOU WILL GET INFINITY
 # mutate(syn = ifelse(is.na(syn), 0, syn)) %>%
  mutate(nonsyn = ifelse(is.na(nonsyn), 0, nonsyn)) %>%
  mutate(ratio = nonsyn / syn)

scaff_rect <- get_scaff_rect(tjd)

#fst <- get_fst()

```

```{r}
pdf(here(plotpath, "stats.pdf"), width = 12, height = 10)
ggplot(syn_nonsyn) +
  geom_rect(data = goi, aes(xmin = start, xmax = end), ymin = min(syn_nonsyn$ratio, na.rm = TRUE), ymax = max(syn_nonsyn$ratio, na.rm = TRUE), col = "red", fill = NA) +
  geom_line(aes(x = BIN_START, y = ratio), alpha = 0.6, lwd = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", lwd = 0.5, col = "white") +
  geom_rect(data = scaff_rect, aes(xmin = BIN_START, xmax = BIN_END), ymin = min(syn_nonsyn$ratio, na.rm = TRUE), ymax = max(syn_nonsyn$ratio, na.rm = TRUE), col = "black", fill = NA) +
  scaffold_theme() +
  ggtitle("Non-synonymous : Synonymous") +
  theme(axis.text.y = element_text(), axis.ticks.y = element_line())

ggplot(genes_bin) +
  geom_line(aes(x = BIN_START, y = ngenes), alpha = 0.6, lwd = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", lwd = 0.5, col = "white") +
  geom_rect(data = scaff_rect, aes(xmin = BIN_START, xmax = BIN_END), ymin = min(genes_bin$ngenes, na.rm = TRUE), ymax = max(genes_bin$ngenes, na.rm = TRUE), col = "black", fill = NA) +
  scaffold_theme() +
  ggtitle("Gene count") +
  theme(axis.text.y = element_text(), axis.ticks.y = element_line())

ggplot(tjd) +
  geom_rect(data = goi, aes(xmin = start, xmax = end), ymin = min(tjd$TajimaD, na.rm = TRUE), ymax = max(tjd$TajimaD, na.rm = TRUE), col = "red", fill = NA) +
  geom_line(aes(x = BIN_START, y = TajimaD, col = pop), alpha = 0.6, lwd = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", lwd = 0.5, col = "white") +
  geom_rect(data = scaff_rect, aes(xmin = BIN_START, xmax = BIN_END), ymin = 0, ymax = 0, col = "black", fill = NA, lty = "dashed") +
  geom_rect(data = scaff_rect, aes(xmin = BIN_START, xmax = BIN_END), ymin = min(tjd$TajimaD, na.rm = TRUE), ymax = max(tjd$TajimaD, na.rm = TRUE), col = "black", fill = NA) +
  scaffold_theme() +
  ggtitle("Tajima's D") 

tjd2 <- filter(tjd, pop %in% c("pop1", "pop6"), scaffold %in% c("chr1", "chr2", "chr3", "chr4", "chr5"))
ggplot(tjd2) +
  geom_rect(data = goi, aes(xmin = start, xmax = end), ymin = min(tjd2$TajimaD, na.rm = TRUE), ymax = max(tjd2$TajimaD, na.rm = TRUE), col = "red", fill = NA) +
  geom_line(aes(x = BIN_START, y = TajimaD, col = pop), alpha = 0.7, lwd = 0.5) +
  geom_rect(data = scaff_rect, aes(xmin = BIN_START, xmax = BIN_END), ymin = 0, ymax = 0, col = "black", fill = NA, lty = "dashed") +
  geom_rect(data = scaff_rect, aes(xmin = BIN_START, xmax = BIN_END), ymin = min(tjd2$TajimaD, na.rm = TRUE), ymax = max(tjd2$TajimaD, na.rm = TRUE), col = "black", fill = NA) +
  scaffold_theme() +
  ggtitle("Tajima's D (pop1 vs pop6)") 
dev.off()
```

```{r}
test <- tjd %>% filter(scaffold == "chr1") %>% mutate(scaffold = factor(scaffold, levels = unique(scaffold)))
pdf(here(plotpath, "tjd_bin.pdf"), width = 10, height = 20)
ggplot(tjd) +
  geom_tile(aes(x = BIN_START, y = pop, fill = TajimaD)) +
  facet_wrap(~scaffold, ncol = 1) +
  scale_fill_gradient2(midpoint = 0, low = "cyan3", mid = "white", high = "orange2", limits = c(-0.55, 0.55)) +
 # geom_rect(data = goi, aes(xmin = start, xmax = end), ymin = -Inf, ymax = Inf, col = "red", fill = NA) +
 theme_void() +
 theme(axis.text.y = element_text()) +
 scale_x_continuous(expand = c(0, 0))
dev.off()

tjd %>% filter(scaffold == "chr1", pop == "pop9")
```