```{r}
library(tidyverse)
library(here)
#BiocManager::install("topGO")
library(topGO)
library(cowplot)
library(gprofiler2)

source(here("analysis", "gea", "functions_go.R"))
plotpath <- here("analysis", "gea", "plots")
outpath <- here("analysis", "gea", "outputs")
```

```{r}
# Get non-synonymous GEA SNPs and genes
genes_nonsyn <- read_csv(here(outpath, "bio1ndvi_gea_genes_nonsyn.csv"))

# Get all annotated genes
all_genes <-read_csv(here("analysis", "gea", "outputs", "all_genes_list.csv"))

# Filter to only include genes with non-synonymous GEA variants
genes_org <- 
  all_genes %>% 
  # TODO: this shouldn't reall need to be two steps
  filter(full_name %in% genes_nonsyn$full_name) %>%
  drop_na(gene_name)

# Check the number of unique genes
print(paste("Number of GEA genes with unique gene names:", nrow(genes_org %>% distinct(gene_name)))) # 3521
print(paste("Number of all genes with names or descriptions:", nrow(all_genes))) # 18670

#https://cran.r-project.org/web/packages/gprofiler2/vignettes/gprofiler2.html
#install.packages("gprofiler2")
library(gprofiler2)
```

# Run GO 

Column descriptions for gprofiler output:
 
query: The name or identifier of the query. This indicates which query the results are associated with.
significant: A logical value (TRUE or FALSE) indicating whether the term is significantly enriched.
p_value: The p-value for the enrichment of the term. It indicates the statistical significance of the enrichment.
term_size: The total number of genes associated with the GO term in the background set.
query_size: The total number of genes in the query set.
intersection_size: The number of genes in the query set that are also associated with the GO term.
*precision:* The precision of the enrichment, calculated as the ratio of intersection_size to query_size. Precision measures the proportion of relevant genes (i.e., genes associated with the GO term) in the query set.
*recall:* The recall of the enrichment, calculated as the ratio of intersection_size to term_size. Recall measures the proportion of genes associated with the GO term that are present in the reference set.
term_id: The identifier of the GO term or other ontology term.
source: The source of the term, such as "GO:BP" for Gene Ontology Biological Process, "GO:MF" for Gene Ontology Molecular Function, or "GO:CC" for Gene Ontology Cellular Component.
term_name: The name of the GO term or other ontology term.
effective_domain_size: The effective domain size, which represents the number of genes in the background set that are considered for the enrichment analysis.
source_order: The order of the source in the results, which can be used for sorting or prioritizing results.
parents: The parent terms of the GO term or other ontology term. This provides hierarchical context for the term.
highlighted: A logical value (TRUE or FALSE) indicating whether the term is highlighted in the results. This can be used for visualization or emphasis in the output.
Run GO analysis for hsapiens genes

```{r}
# Load correlations between SNPs and variables (from _seperate_vars.R)
cors <- 
  read_csv(here(outpath, "bio1ndvi_genes_cortest.csv")) %>%
  # Drop loci without significant correlations with NDVI or BIO1
  drop_na(group) %>%
  dplyr::select(locus, group) %>%
  # [TODO: FIX] Some loci are currently duplicated so right now i am just taking the first one [NEED TO FIX]
  group_by(locus) %>%
  dplyr::slice(1) %>%
  ungroup()
```

# CUSTOM BG

query: The name or identifier of the query. This indicates which query the results are associated with.
significant: A logical value (TRUE or FALSE) indicating whether the term is significantly enriched.
p_value: The p-value for the enrichment of the term. It indicates the statistical significance of the enrichment.
term_size: The total number of genes associated with the GO term in the background set.
query_size: The total number of genes in the query set.
intersection_size: The number of genes in the query set that are also associated with the GO term.
*precision:* The precision of the enrichment, calculated as the ratio of intersection_size to query_size. Precision measures the proportion of relevant genes (i.e., genes associated with the GO term) in the query set.
*recall:* The recall of the enrichment, calculated as the ratio of intersection_size to term_size. Recall measures the proportion of genes associated with the GO term that are present in the reference set.
term_id: The identifier of the GO term or other ontology term.
source: The source of the term, such as "GO:BP" for Gene Ontology Biological Process, "GO:MF" for Gene Ontology Molecular Function, or "GO:CC" for Gene Ontology Cellular Component.
term_name: The name of the GO term or other ontology term.
effective_domain_size: The effective domain size, which represents the number of genes in the background set that are considered for the enrichment analysis.
source_order: The order of the source in the results, which can be used for sorting or prioritizing results.
parents: The parent terms of the GO term or other ontology term. This provides hierarchical context for the term.
highlighted: A logical value (TRUE or FALSE) indicating whether the term is highlighted in the results. This can be used for visualization or emphasis in the output.
Run GO analysis for hsapiens genes

## Prepare queries and background
```{r}
# Link genes and snps
rda_genes <-
  # file from csq.R
  genes_nonsyn %>%
  dplyr::select(locus, full_name) %>%
  # Join with genes
  left_join(dplyr::select(genes_org, full_name, gene_name)) %>%
  # !Drop rows without a gene name - note that this drops genes without a description too
  drop_na(gene_name) %>%
  # Join with the correlation results 
  # DISTINCT IS TEMPORARY TO DEAL WITH DUPLICATES (LUCKILY GROUPS ARE SAME FOR DUPLICATES)
  # Drop group NA to remove loci without significant correlations with NDVI or BIO1
  left_join(cors, by = "locus") 

bio1_query <- rda_genes %>% filter(grepl("bio1", group)) %>% pull(gene_name)
length(bio1_query)
ndvi_query <- rda_genes %>% filter(grepl("ndvi", group)) %>% pull(gene_name)
length(ndvi_query)
both_query <- rda_genes %>% pull(gene_name)

background <- unique(all_genes$gene_name)
```

## Human reference
```{r}
go_hsapiens_defaultbg <- gost(
    query = both_query, 
    organism = "hsapiens", 
    ordered_query = FALSE, 
    multi_query = FALSE, 
    significant = FALSE, 
    exclude_iea = FALSE, 
    measure_underrepresentation = FALSE, 
    evcodes = TRUE, 
    user_threshold = 0.05, 
    correction_method = "fdr",
    numeric_ns = "", 
    sources = NULL, 
    as_short_link = FALSE
  )
go_both_hsapiens <- run_go(both_query, custom_bg = background, organism = "hsapiens")
go_bio1_hsapiens <- run_go(bio1_query, custom_bg = background, organism = "hsapiens")
go_ndvi_hsapiens <- run_go(ndvi_query, custom_bg = background, organism = "hsapiens")

write_csv(go_both_hsapiens$result, here(outpath, "hsapiens_custom_go_both.csv"))
write_csv(go_bio1_hsapiens$result, here(outpath, "hsapiens_custom_go_bio1.csv"))
write_csv(go_ndvi_hsapiens$result, here(outpath, "hsapiens_custom_go_ndvi.csv"))

go_both_hsapiens_result <- read_csv(here(outpath, "hsapiens_custom_go_both.csv"))
go_bio1_hsapiens_result <- read_csv(here(outpath, "hsapiens_custom_go_bio1.csv"))
go_ndvi_hsapiens_result <- read_csv(here(outpath, "hsapiens_custom_go_ndvi.csv"))
```

```{r}
# Compare variables
go_both_hsapiens_result %>% filter(p_value < 0.05) %>% dplyr::select(term_name, source, precision, intersection_size) %>% arrange(desc(precision))
go_bio1_hsapiens_result %>% filter(p_value < 0.05) %>% dplyr::select(term_name, source, precision, intersection_size) %>% arrange(desc(precision))
go_ndvi_hsapiens_result %>% filter(p_value < 0.05) %>% dplyr::select(term_name, source, precision, intersection_size) %>% arrange(desc(precision))

go_sigbp(go_both_hsapiens_result) 
go_sigmf(go_both_hsapiens_result) 

go_sigbp(go_bio1_hsapiens_result) 
go_sigmf(go_bio1_hsapiens_result) 

go_sigbp(go_ndvi_hsapiens_result)
go_sigmf(go_ndvi_hsapiens_result) 
```

```{r}
# Check for significant terms
go_sigbp(go_both_hsapiens_result) 
go_sigmf(go_both_hsapiens_result) 

# Check top 10 GO:BP terms based on precision (categories where at least 10% of the genes in the query are associated with the term)
go_topbp(go_both_hsapiens_result, precision_cutoff = 0.10) 
go_topmf(go_both_hsapiens_result, precision_cutoff = 0.10) 

# Check what genes are associated with the significant GO terms
sig_go <- go_both_hsapiens_result %>% filter(p_value < 0.05)
result_genes <- 
  go_both_hsapiens_result %>%
  filter(term_name %in% sig_go$term_name) %>%
  pull(intersection) %>% 
  strsplit(",") %>% 
  unlist()

# Get gene information
go_gea_genes <- genes_org %>% filter(gene_name %in% result_genes)

# Get genes of interest
source(here("analysis", "gea", "functions_selection_stats.R"))
goi <- get_goi_names()

# Filter GO genes for genes of interest
go_goi <- go_gea_genes %>% filter(grepl(paste(goi, collapse = "|"), full_name)) %>% distinct(gene_name)
go_goi 

heat_shock <- go_gea_genes %>% filter(grepl("heat shock", full_name)) %>% distinct(gene_name)

# Check what GO terms HSPA2 (hsp70) and heat shock proteins is associated with
go_both_hsapiens_result %>% 
  filter(p_value < 0.05) %>% 
  filter(grepl("HSPA2", intersection)) %>% 
  distinct(term_name, source)

go_both_hsapiens_result %>% 
  filter(p_value < 0.05) %>% 
  filter(grepl(paste0(heat_shock, "|"), intersection)) %>% 
  filter(source %in% c("GO:BP", "GO:MF")) %>%
  distinct(term_name, source)
```

## Chicken reference
```{r}
# CHECK IF THIS IS DIFFERENT 
go_both_ggallus <- run_go(both_query, custom_bg = background, organism = "ggallus")
go_bio1_ggallus <- run_go(bio1_query, custom_bg = background, organism = "ggallus")
go_ndvi_ggallus <- run_go(ndvi_query, custom_bg = background, organism = "ggallus")

write_csv(go_both_ggallus$result, here(outpath, "ggallus_custom_go_both.csv"))
write_csv(go_bio1_ggallus$result, here(outpath, "ggallus_custom_go_bio1.csv"))
write_csv(go_ndvi_ggallus$result, here(outpath, "ggallus_custom_go_ndvi.csv"))

go_both_ggallus_result <- read_csv(here(outpath, "ggallus_custom_go_both.csv"))
go_bio1_ggallus_result <- read_csv(here(outpath, "ggallus_custom_go_bio1.csv"))
go_ndvi_ggallus_result <- read_csv(here(outpath, "ggallus_custom_go_ndvi.csv"))
```

```{r}
# Compare variables
go_both_ggallus_result %>% filter(p_value < 0.05) %>% dplyr::select(term_name, source, precision) %>% arrange(desc(precision))
go_bio1_ggallus_result %>% filter(p_value < 0.05) %>% dplyr::select(term_name, source, precision) %>% arrange(desc(precision))
go_ndvi_ggallus_result %>% filter(p_value < 0.05) %>% dplyr::select(term_name, source, precision) %>% arrange(desc(precision))

go_sigbp(go_both_ggallus_result) 
go_sigmf(go_both_ggallus_result) 

go_sigbp(go_bio1_ggallus_result) 
go_sigmf(go_bio1_ggallus_result) 

go_sigbp(go_ndvi_ggallus_result) 
go_sigmf(go_ndvi_ggallus_result) 
```

```{r}
# Check for significant terms
ggallus_result <- go_both_ggallus_result
go_sigbp(ggallus_result) # 0 significant GO:BP terms
go_sigmf(ggallus_result) 

# Check top 10 GO:BP terms based on precision (categories where at least 10% of the genes in the query are associated with the term)
go_topbp(ggallus_result, precision_cutoff = 0.10) 
go_topmf(ggallus_result, precision_cutoff = 0.10) 

# Check what genes are associated with the significant GO terms
ggallus_result_genes <- 
  ggallus_result %>% 
  filter(p_value < 0.05) %>% 
  pull(intersection) %>% 
  strsplit(",") %>% 
  unlist()

# Get gene information
go_gea_genes <- genes_org %>% filter(gene_name %in% result_genes)
go_gea_genes %>% distinct(gene_name) %>% write_csv(here(outpath, "hsapiens_atp_dependent_activity_genes.csv"))

# Get genes of interest
source(here("analysis", "gea", "functions_selection_stats.R"))
goi <- get_goi_names()

# Filter GO genes for genes of interest
go_goi <- go_gea_genes %>% filter(grepl(paste(goi, collapse = "|"), full_name)) %>% distinct(full_name)
go_goi 

go_gea_genes %>% filter(grepl("heat", full_name))

# Check what GO terms HSPA2 (hsp70) and heat shock proteins is associated with
go_both_ggallus_result %>% 
  filter(p_value < 0.05) %>% 
  filter(grepl("HSPA2", intersection)) %>% 
  distinct(term_name, source)

go_both_ggallus_result %>% 
  filter(p_value < 0.05) %>% 
  filter(grepl(paste0(heat_shock, "|"), intersection)) %>% 
  filter(source %in% c("GO:BP", "GO:MF")) %>%
  distinct(term_name, source)
```

# Compare results
```{r}
sig_both_hsapiens <- go_both_hsapiens_result %>% filter(p_value < 0.05)
sig_both_ggallus <- go_both_ggallus_result %>% filter(p_value < 0.05)

bp_both_hsapiens <- sig_both_hsapiens %>% filter(grepl("GO:BP", source))
bp_both_ggallus <- sig_both_ggallus %>% filter(grepl("GO:BP", source))

go_topbp(sig_both_hsapiens)
go_topbp(sig_both_ggallus)

library(gt)
gtsave(go_table(bp_both_hsapiens), here(plotpath, "bp_both_hsapiens.html"))
gtsave(go_table(bp_both_ggallus), here(plotpath, "bp_both_ggallus.html"))
```