```{r}
library(tidyverse)
library(here)
#BiocManager::install("topGO")
library(topGO)
library(cowplot)
library(gprofiler2)

source(here("analysis", "gea", "functions_go.R"))
plotpath <- here("analysis", "gea", "plots")
outpath <- here("analysis", "gea", "outputs")
```

```{r}
# Get non-synonymous GEA SNPs and genes
genes_nonsyn <- read_csv(here(outpath, "bio1ndvi_gea_genes_nonsyn.csv"))

# Get all annotated genes
all_genes <-read_csv(here("analysis", "gea", "outputs", "all_genes_list.csv"))

# Filter to only include genes with non-synonymous GEA variants
genes_org <- 
  all_genes %>% 
  # TODO: this shouldn't reall need to be two steps
  filter(full_name %in% genes_nonsyn$full_name) %>%
  drop_na(uniprot_id)

# Check the number of unique genes
print(paste("Number of GEA genes with unique UniprotIDs:", nrow(genes_org %>% distinct(uniprot_id)))) # 4958
print(paste("Number of all genes with UniprotIDs:", nrow(all_genes))) # 20638
print(paste("Number of all genes with unique UniprotIDs:", nrow(all_genes %>% distinct(uniprot_id)))) # 20638

#https://cran.r-project.org/web/packages/gprofiler2/vignettes/gprofiler2.html
#install.packages("gprofiler2")
library(gprofiler2)
```

# Find gene orthologs
Note: warnings are normal, they are just saying that some genes don't have orthologs in the other species
```{r}
all_genes_hsapiens <- get_ortho(all_genes, "hsapiens", file_prefix = "hsapiens_all", cache = TRUE)
all_genes_ggallus <- get_ortho(all_genes, "ggallus", file_prefix = "ggallus_all", cache = TRUE)
genes_hsapiens <- all_genes_hsapiens %>% filter(uniprot_id %in% genes_org$uniprot_id)
genes_ggallus <- all_genes_ggallus %>% filter(uniprot_id %in% genes_org$uniprot_id)
# genes_hsapiens <- genes_org %>% mutate(query = uniprot_id)
# genes_ggallus <- genes_org %>% mutate(query = uniprot_id)
# all_genes_hsapiens <- all_genes %>% mutate(query = uniprot_id)
# all_genes_ggallus <- all_genes %>% mutate(query = uniprot_id)
```

# Run GO 

Column descriptions for gprofiler output:
 
query: The name or identifier of the query. This indicates which query the results are associated with.
significant: A logical value (TRUE or FALSE) indicating whether the term is significantly enriched.
p_value: The p-value for the enrichment of the term. It indicates the statistical significance of the enrichment.
term_size: The total number of genes associated with the GO term in the background set.
query_size: The total number of genes in the query set.
intersection_size: The number of genes in the query set that are also associated with the GO term.
*precision:* The precision of the enrichment, calculated as the ratio of intersection_size to query_size. Precision measures the proportion of relevant genes (i.e., genes associated with the GO term) in the query set.
*recall:* The recall of the enrichment, calculated as the ratio of intersection_size to term_size. Recall measures the proportion of genes associated with the GO term that are present in the reference set.
term_id: The identifier of the GO term or other ontology term.
source: The source of the term, such as "GO:BP" for Gene Ontology Biological Process, "GO:MF" for Gene Ontology Molecular Function, or "GO:CC" for Gene Ontology Cellular Component.
term_name: The name of the GO term or other ontology term.
effective_domain_size: The effective domain size, which represents the number of genes in the background set that are considered for the enrichment analysis.
source_order: The order of the source in the results, which can be used for sorting or prioritizing results.
parents: The parent terms of the GO term or other ontology term. This provides hierarchical context for the term.
highlighted: A logical value (TRUE or FALSE) indicating whether the term is highlighted in the results. This can be used for visualization or emphasis in the output.
Run GO analysis for hsapiens genes

```{r}
# Load correlations between SNPs and variables (from _seperate_vars.R)
cors <- 
  read_csv(here(outpath, "bio1ndvi_genes_cortest.csv")) %>%
  # Drop loci without significant correlations with NDVI or BIO1
  drop_na(group) %>%
  dplyr::select(locus, group) %>%
  # [TODO: FIX] Some loci are currently duplicated so right now i am just taking the first one [NEED TO FIX]
  group_by(locus) %>%
  dplyr::slice(1) %>%
  ungroup()
```

# CUSTOM BG

query: The name or identifier of the query. This indicates which query the results are associated with.
significant: A logical value (TRUE or FALSE) indicating whether the term is significantly enriched.
p_value: The p-value for the enrichment of the term. It indicates the statistical significance of the enrichment.
term_size: The total number of genes associated with the GO term in the background set.
query_size: The total number of genes in the query set.
intersection_size: The number of genes in the query set that are also associated with the GO term.
*precision:* The precision of the enrichment, calculated as the ratio of intersection_size to query_size. Precision measures the proportion of relevant genes (i.e., genes associated with the GO term) in the query set.
*recall:* The recall of the enrichment, calculated as the ratio of intersection_size to term_size. Recall measures the proportion of genes associated with the GO term that are present in the reference set.
term_id: The identifier of the GO term or other ontology term.
source: The source of the term, such as "GO:BP" for Gene Ontology Biological Process, "GO:MF" for Gene Ontology Molecular Function, or "GO:CC" for Gene Ontology Cellular Component.
term_name: The name of the GO term or other ontology term.
effective_domain_size: The effective domain size, which represents the number of genes in the background set that are considered for the enrichment analysis.
source_order: The order of the source in the results, which can be used for sorting or prioritizing results.
parents: The parent terms of the GO term or other ontology term. This provides hierarchical context for the term.
highlighted: A logical value (TRUE or FALSE) indicating whether the term is highlighted in the results. This can be used for visualization or emphasis in the output.
Run GO analysis for hsapiens genes

## Human reference

```{r}
# Link genes and snps
rda_hsapiens_genes <-
  # file from csq.R
  genes_nonsyn %>%
  dplyr::select(locus, full_name) %>%
  # Join with genes
  left_join(dplyr::select(genes_org, full_name, uniprot_id)) %>%
  drop_na(uniprot_id) %>%
  # Join with orthologs
  left_join(genes_hsapiens, by = "uniprot_id") %>%
  # Some genes don't have hsapiens orthologs/query names, so drop them
  drop_na(query) %>%
  # Join with the correlation results 
  # DISTINCT IS TEMPORARY TO DEAL WITH DUPLICATES (LUCKILY GROUPS ARE SAME FOR DUPLICATES)
  # Drop group NA to remove loci without significant correlations with NDVI or BIO1
  left_join(cors, by = "locus") 
  
bio1_query_hsapiens <- rda_hsapiens_genes %>% filter(grepl("bio1", group)) %>% pull(query)
length(bio1_query_hsapiens)
ndvi_query_hsapiens <- rda_hsapiens_genes %>% filter(grepl("ndvi", group)) %>% pull(query)
length(ndvi_query_hsapiens)
both_query_hsapiens <- rda_hsapiens_genes %>% pull(query)

hsapiens_bg <- unique(all_genes_hsapiens$query)
go_bio1_hsapiens <- run_go(bio1_query_hsapiens, custom_bg = hsapiens_bg, organism = "hsapiens")
go_ndvi_hsapiens <- run_go(ndvi_query_hsapiens, custom_bg = hsapiens_bg, organism = "hsapiens")
go_both_hsapiens <- run_go(both_query_hsapiens, custom_bg = hsapiens_bg, organism = "hsapiens")

write_csv(go_both_hsapiens$result, here(outpath, "hsapiens_custom_go_both.csv"))
write_csv(go_bio1_hsapiens$result, here(outpath, "hsapiens_custom_go_bio1.csv"))
write_csv(go_ndvi_hsapiens$result, here(outpath, "hsapiens_custom_go_ndvi.csv"))

go_both_hsapiens_result <- read_csv(here(outpath, "hsapiens_custom_go_both.csv"))
go_bio1_hsapiens_result <- read_csv(here(outpath, "hsapiens_custom_go_bio1.csv"))
go_ndvi_hsapiens_result <- read_csv(here(outpath, "hsapiens_custom_go_ndvi.csv"))
```

```{r}
# Compare variables
go_both_hsapiens_result %>% filter(p_value < 0.05) %>% dplyr::select(term_name, source, precision, intersection_size) %>% arrange(desc(precision))
go_bio1_hsapiens_result %>% filter(p_value < 0.05) %>% dplyr::select(term_name, source, precision, intersection_size) %>% arrange(desc(precision))
go_ndvi_hsapiens_result %>% filter(p_value < 0.05) %>% dplyr::select(term_name, source, precision, intersection_size) %>% arrange(desc(precision))

go_sigbp(go_both_hsapiens_result) # 0 significant GO:BP terms
go_sigmf(go_both_hsapiens_result) # 2 significant GO:MF terms (ATP-dependent activity and ATP hydrolysisq activity)

go_sigbp(go_bio1_hsapiens_result) # 0 significant GO:BP terms
go_sigmf(go_bio1_hsapiens_result) # 0 significant GO:MF terms

go_sigbp(go_ndvi_hsapiens_result) # 5 significant GO:BP terms 
#                              term_name term_size query_size intersection_size
# 1                    response to dsRNA        27       1037                12
# 2           cellular response to dsRNA        15       1037                 9
# 3          icosanoid metabolic process        61       1037                18
# 4          response to exogenous dsRNA        22       1037                10
# 5 cellular response to exogenous dsRNA        11       1037                 7
go_sigmf(go_ndvi_hsapiens_result) # 0 significant GO:MF terms
```

```{r}
# Check for significant terms
hsapiens_result <- go_both_hsapiens_result
go_sigbp(hsapiens_result) # 0 significant GO:BP terms
go_sigmf(hsapiens_result) # 2 significant GO:MF terms (ATP-dependent activity and ATP hydrolysisq activity)

go_sigbp(hsapiens_result, p = 0.10) # 0 significant GO:BP terms
go_sigmf(hsapiens_result, p = 0.10) # 4 significant GO:MF terms (ATP-dependent activity, ATP hydrolysisq activity, dipaptidyl-peptidase activity, carbohydrate binding)

# Check top 10 GO:BP terms based on precision (categories where at least 10% of the genes in the query are associated with the term)
go_topbp(hsapiens_result, precision_cutoff = 0.10) 
go_topmf(hsapiens_result, precision_cutoff = 0.10) 

# Check what genes are associated with the significant GO terms
hsapiens_result_genes <- 
  hsapiens_result %>% 
  filter(term_name == "ATP-dependent activity") %>%
  pull(intersection) %>% 
  strsplit(",") %>% 
  unlist()

# Get original uniprot IDs for the genes
# Necessary step because orthologs were used for running GO
hsapiens_result_uniprot_ids <- genes_hsapiens %>% filter(query %in% hsapiens_result_genes) %>% pull(uniprot_id)

# Get gene information
hsapiens_go_gea_genes <- genes_org %>% filter(uniprot_id %in% hsapiens_result_uniprot_ids)
hsapiens_go_gea_genes %>% distinct(gene_name) %>% write_csv(here(outpath, "hsapiens_atp_dependent_activity_genes.csv"))

# Get genes of interest
goi <- get_goi_names()

# Filter GO genes for genes of interest
hsapiens_go_goi <- hsapiens_go_gea_genes %>% filter(grepl(paste(goi, collapse = "|"), full_name)) %>% distinct(gene_name)
hsapiens_go_goi # Heat shock 70 kDa protein 13 and Heat shock 70 kDa protein 1A
```

## Chicken reference
```{r}
# Link genes and snps
rda_ggallus_genes <-
  # file from csq.R
  genes_nonsyn %>%
  dplyr::select(locus, full_name) %>%
  # Join with genes
  left_join(dplyr::select(genes_org, full_name, uniprot_id)) %>%
  drop_na(uniprot_id) %>%
  # Join with orthologs
  left_join(genes_ggallus, by = "uniprot_id") %>%
  # Some genes don't have ggallus orthologs/query names, so drop them
  drop_na(query) %>%
  # Join with the correlation results 
  # DISTINCT IS TEMPORARY TO DEAL WITH DUPLICATES (LUCKILY GROUPS ARE SAME FOR DUPLICATES)
  # Drop group NA to remove loci without significant correlations with NDVI or BIO1
  left_join(cors, by = "locus") 
  
bio1_query_ggallus <- rda_ggallus_genes %>% filter(grepl("bio1", group)) %>% pull(query)
length(bio1_query_ggallus)
ndvi_query_ggallus <- rda_ggallus_genes %>% filter(grepl("ndvi", group)) %>% pull(query)
length(ndvi_query_ggallus)
both_query_ggallus <- rda_ggallus_genes %>% pull(query)

ggallus_bg <- unique(all_genes_ggallus$query)
go_bio1_ggallus <- run_go(bio1_query_ggallus, custom_bg = ggallus_bg, organism = "ggallus")
go_ndvi_ggallus <- run_go(ndvi_query_ggallus, custom_bg = ggallus_bg, organism = "ggallus")
go_both_ggallus <- run_go(both_query_ggallus, custom_bg = ggallus_bg, organism = "ggallus")

write_csv(go_both_ggallus$result, here(outpath, "ggallus_custom_go_both.csv"))
write_csv(go_bio1_ggallus$result, here(outpath, "ggallus_custom_go_bio1.csv"))
write_csv(go_ndvi_ggallus$result, here(outpath, "ggallus_custom_go_ndvi.csv"))

go_both_ggallus_result <- read_csv(here(outpath, "ggallus_custom_go_both.csv"))
go_bio1_ggallus_result <- read_csv(here(outpath, "ggallus_custom_go_bio1.csv"))
go_ndvi_ggallus_result <- read_csv(here(outpath, "ggallus_custom_go_ndvi.csv"))
```


```{r}
# Compare variables

go_both_ggallus_result %>% filter(p_value < 0.05) %>% dplyr::select(term_name, source, precision) %>% arrange(desc(precision))
go_bio1_ggallus_result %>% filter(p_value < 0.05) %>% dplyr::select(term_name, source, precision) %>% arrange(desc(precision))
go_ndvi_ggallus_result %>% filter(p_value < 0.05) %>% dplyr::select(term_name, source, precision) %>% arrange(desc(precision))


go_sigbp(go_both_ggallus_result) # 0 significant GO:BP terms
go_sigmf(go_both_ggallus_result) # 1 significant GO:MF terms (ATP-dependent activity)

go_sigbp(go_bio1_ggallus_result) # 0 significant GO:BP terms
go_sigmf(go_bio1_ggallus_result) # 0 significant GO:MF terms

go_sigbp(go_ndvi_ggallus_result) # 0 significant GO:BP terms
go_sigmf(go_ndvi_ggallus_result) # 0 significant GO:MF terms
```

```{r}
# Check for significant terms
ggallus_result <- ggallus_custom_bg$result 
go_sigbp(ggallus_result) # 0 significant GO:BP terms
go_sigmf(ggallus_result) # 1 significant GO:MF terms (ATP-dependent activity)

# Check top 10 GO:BP terms based on precision (categories where at least 10% of the genes in the query are associated with the term)
go_topbp(ggallus_result, precision_cutoff = 0.10) 
go_topmf(ggallus_result, precision_cutoff = 0.10) 

# Check what genes are associated with the significant GO terms
ggallus_result_genes <- 
  ggallus_result %>% 
  filter(p_value < 0.05) %>% 
  pull(intersection) %>% 
  strsplit(",") %>% 
  unlist()

# Get original uniprot IDs for the genes
# Necessary step because orthologs were used for running GO
ggallus_result_uniprot_ids <- genes_ggallus %>% filter(query %in% ggallus_result_genes) %>% pull(uniprot_id)

# Get gene information
ggallus_go_gea_genes <- genes_org %>% filter(uniprot_id %in% ggallus_result_uniprot_ids)

# Get genes of interest
goi <- get_goi_names()

# Filter GO genes for genes of interest
ggallus_go_goi <- ggallus_go_gea_genes %>% filter(grepl(paste(goi, collapse = "|"), full_name)) %>% distinct(gene_name)
ggallus_go_goi # Heat shock 70 kDa protein 13 and Heat shock 70 kDa protein 1A
```

# Looking into NDVI for H. sapiends

```{r}
go_ndvi <- go_ndvi_hsapiens_result %>% filter(source == "GO:BP", p_value < 0.05)
go_ndvi_genes <- 
  go_ndvi %>% 
  pull(intersection) %>% 
  strsplit(",") %>% 
  unlist()


# Get original uniprot IDs for the genes
ndvi_uniprot_ids <- genes_hsapiens %>% filter(query %in% go_ndvi_genes) %>% pull(uniprot_id)

# Get gene information
ndvi_go_gea_genes <- genes_org %>% filter(uniprot_id %in% ndvi_uniprot_ids)

print(unique(ndvi_go_gea_genes$gene_name)) 

tick_goi <- c(
  "Interferon epsilon",
  "Interferon beta",
  "Antiviral innate immune response receptor RIG-I",
  "Interferon-induced protein with tetratricopeptide repeats 1",
  "Interferon regulatory factor 3",
  "MHC class II transactivator"
)

ndvi_go_gea_genes <- 
  ndvi_go_gea_genes %>% 
  filter(grepl(paste(tick_goi, collapse = "|"), full_name)) 

# Get SNPs associated with the genes
ndvi_snps <- get_gea_genes(nonsyn = TRUE) %>% filter(full_name %in% ndvi_go_gea_genes$full_name) 

# Load dosage data
dos <- read_table(here("analysis", "gea", "outputs", "genes.raw"))

# Load coords and model dataframe
coords <- 
  get_coords(sf = TRUE) %>% 
  filter(SampleID %in% dos$IID) %>% 
  st_transform(3310) %>%
  # REMOVE CHANNEL ISLAND
  filter(SampleID != "Scelocci_CCGPMC_MW01-3-14") 

# Subset to only samples in coords
dos <- dos[dos$IID %in% coords$SampleID, ]

# Confirm orders are the same
stopifnot(all(dos$IID == coords$SampleID))

# CORRECT DOS NAMES
colnames(dos) <- gsub("_[A-Z]$", "", colnames(dos))

# Get loci for the NDVI genes
gene_loci <- ndvi_snps %>% pull(locus)

# Get dosage data for the NDVI genes
gene_dos <- dos[, gene_loci]
gene_dos$SampleID <- dos$IID

# Impute missing values
mean(is.na(gene_dos)) * 100 # < 1% NA
nona_dos <- 
  gene_dos %>%
  mutate(across(everything(), ~ ifelse(is.na(.), median(., na.rm = TRUE), .)))

# Make into df for plotting
gene_df <- 
  coords %>% 
  left_join(gene_dos, by = "SampleID")  %>%
  pivot_longer(-c(SampleID, geometry), names_to = "locus", values_to = "dos")

# Run PCA
gene_pca <- prcomp(dplyr::select(nona_dos, -SampleID))

# Add PCA results to the gene_df
gene_df <- 
  coords %>% 
  left_join(gene_dos, by = "SampleID") %>%
  mutate(PC1 = gene_pca$x[, 1]) %>%
  pivot_longer(-c(SampleID, geometry, PC1), names_to = "locus", values_to = "dos")

ca <- get_ca()
nloci <- length(unique(gene_df$locus))
pca_plt <-
  ggplot(gene_df) +
  geom_sf(data = ca) +
  geom_sf(aes(fill = PC1), pch = 21, cex = 2, col = rgb(0,0,0,0)) +
  scale_fill_viridis_c(option = "turbo") +
  theme_void() +
  theme(legend.position = "bottom") +
  labs(title = paste0("NDVI GO genes", " (# SNPs = ", nloci, ", # Genes = ", length(unique(ndvi_go_gea_genes$full_name)), ")."))

pdf(here(plotpath, "ndvi_go_genes_pca.pdf"), width = 8, height = 6)
print(pca_plt)
dev.off()
```

# OLD GO
### hsapiens
```{r}
if (!file.exists(here("analysis", "gea", "outputs", "all_genes_go_hsapiens.csv"))) {
  go_result_all <-
    gost(
      query = all_genes_hsapiens$query, 
      organism = "hsapiens", ordered_query = FALSE, 
      multi_query = FALSE, significant = FALSE, exclude_iea = FALSE, 
      measure_underrepresentation = FALSE, evcodes = FALSE, 
      user_threshold = 0.05, correction_method = "g_SCS", 
      domain_scope = "annotated", custom_bg = NULL, 
      numeric_ns = "", sources = NULL, as_short_link = FALSE, highlight = TRUE
    )
  
  write_csv(go_result_all$result, here("analysis", "gea", "outputs", "all_genes_go_hsapiens.csv"))
}

all_genes_go_hsapiens <- 
  read_csv(here("analysis", "gea", "outputs", "all_genes_go_hsapiens.csv")) %>% 
  as_tibble() %>%
  # Just keep GO terms
  filter(grepl("GO", source)) %>%
  dplyr::select(term_id, intersection_size, query_size, term_size, p_value)

# Add _all to the names of the columns
names(all_genes_go_hsapiens)[-1] <- paste0(names(all_genes_go_hsapiens), "_all")[-1]
```

```{r}
# Link genes and snps
rda_hsapiens_genes <-
  # file from csq.R
  genes_nonsyn %>%
  dplyr::select(locus, full_name) %>%
  # Join with genes
  left_join(dplyr::select(genes_org, full_name, uniprot_id)) %>%
  drop_na(uniprot_id) %>%
  # Join with orthologs
  left_join(genes_hsapiens, by = "uniprot_id") %>%
  # Some genes don't have hsapiens orthologs/query names, so drop them
  drop_na(query) %>%
  # Join with the correlation results 
  # DISTINCT IS TEMPORARY TO DEAL WITH DUPLICATES (LUCKILY GROUPS ARE SAME FOR DUPLICATES)
  # Drop group NA to remove loci without significant correlations with NDVI or BIO1
  left_join(cors, by = "locus") 

bio1_query_hsapiens <- rda_hsapiens_genes %>% filter(grepl("bio1", group)) %>% pull(query)
length(bio1_query_hsapiens)
ndvi_query_hsapiens <- rda_hsapiens_genes %>% filter(grepl("ndvi", group)) %>% pull(query)
length(ndvi_query_hsapiens)
both_query_hsapiens <- rda_hsapiens_genes %>% pull(query)

go_bio1_hsapiens <- run_go(bio1_query_hsapiens, "hsapiens")
go_ndvi_hsapiens <- run_go(ndvi_query_hsapiens, "hsapiens")
go_both_hsapiens <- run_go(both_query_hsapiens, "hsapiens")

sig_bio1_hsapiens <- go_correction(go_bio1_hsapiens, all_genes_go_hsapiens)
sig_ndvi_hsapiens <- go_correction(go_ndvi_hsapiens, all_genes_go_hsapiens)
sig_both_hsapiens <- go_correction(go_both_hsapiens, all_genes_go_hsapiens)
# TODO: check out if  HSP genes are associated with BIO1/NDVI or both
```

```{r}
go_topbp(sig_bio1_hsapiens)
go_topbp(sig_ndvi_hsapiens)
go_topbp(sig_both_hsapiens)
```

```{r}
full_hsapiens <-
  bind_rows(
    data.frame(sig_bio1_hsapiens, run_group = "bio1"),
    data.frame(sig_ndvi_hsapiens, run_group = "ndvi"),
    data.frame(sig_both_hsapiens, run_group = "both")
  )
write_csv(here(outpath, "hsapiens_go.csv"))
```

## ggallus

```{r}
if (!file.exists(here("analysis", "gea", "outputs", "all_genes_go_ggallus.csv"))) {
  go_result_all <-
    gost(
      query = all_genes_ggallus$query, 
      organism = "ggallus", ordered_query = FALSE, 
      multi_query = FALSE, significant = FALSE, exclude_iea = FALSE, 
      measure_underrepresentation = FALSE, evcodes = FALSE, 
      user_threshold = 0.05, correction_method = "g_SCS", 
      domain_scope = "annotated", custom_bg = NULL, 
      numeric_ns = "", sources = NULL, as_short_link = FALSE, highlight = TRUE
    )
  
  write_csv(go_result_all$result, here("analysis", "gea", "outputs", "all_genes_go_ggallus.csv"))
}

all_genes_go_ggallus <- 
  read_csv(here("analysis", "gea", "outputs", "all_genes_go_ggallus.csv")) %>% 
  as_tibble() %>%
  # Just keep GO terms
  filter(grepl("GO", source)) %>%
  dplyr::select(term_id, intersection_size, query_size, term_size, p_value)

names(all_genes_go_ggallus)[-1] <- paste0(names(all_genes_go_ggallus), "_all")[-1]
```

```{r}
# Link genes and snps
rda_ggallus_genes <-
  # file from intersect_genes.R
  read_csv(here("analysis", "gea", "outputs", "bio1ndvi_gea_genes_snp.csv")) %>%
  dplyr::select(locus, full_name) %>%
  # Join with genes
  left_join(dplyr::select(genes_org, full_name, uniprot_id)) %>%
  drop_na(uniprot_id) %>%
  # Join with orthologs
  left_join(genes_ggallus, by = "uniprot_id") %>%
  # Some genes don't have human orthologs/query names, so drop them
  drop_na(query) %>%
  # Join with the correlation results 
  # DISTINCT IS TEMPORARY TO DEAL WITH DUPLICATES (LUCKILY GROUPS ARE SAME FOR DUPLICATES)
  # Drop group NA to remove loci without significant correlations with NDVI or BIO1
  left_join(cors, by = "locus") 


bio1_query_ggallus <- rda_ggallus_genes %>% filter(grepl("bio1", group)) %>% pull(query)
length(bio1_query_ggallus)
ndvi_query_ggallus <- rda_ggallus_genes %>% filter(grepl("ndvi", group)) %>% pull(query)
length(ndvi_query_ggallus)
both_query_ggallus <- rda_ggallus_genes %>% pull(query)

go_bio1_ggallus <- run_go(bio1_query_ggallus, "ggallus")
go_ndvi_ggallus <- run_go(ndvi_query_ggallus, "ggallus")
go_both_ggallus <- run_go(both_query_ggallus, "ggallus")

sig_bio1_ggallus <- go_correction(go_bio1_ggallus, all_genes_go_ggallus)
sig_ndvi_ggallus <- go_correction(go_ndvi_ggallus, all_genes_go_ggallus)
sig_both_ggallus <- go_correction(go_both_ggallus, all_genes_go_ggallus)
```

```{r}
full_ggallus <-
  bind_rows(
    data.frame(sig_bio1_ggallus, run_group = "bio1"),
    data.frame(sig_ndvi_ggallus, run_group = "ndvi"),
    data.frame(sig_both_ggallus, run_group = "both")
  )
write_csv(full_ggallus, here(outpath, "ggallus_go.csv"))
```

# Compare results
```{r}
bp_bio1_hsapiens <- sig_bio1_hsapiens %>% filter(grepl("GO:BP", source))
bp_ndvi_hsapiens <- sig_ndvi_hsapiens %>% filter(grepl("GO:BP", source))
bp_both_hsapiens <- sig_both_hsapiens %>% filter(grepl("GO:BP", source))
bp_bio1_ggallus <- sig_bio1_ggallus %>% filter(grepl("GO:BP", source))
bp_ndvi_ggallus <- sig_ndvi_ggallus %>% filter(grepl("GO:BP", source))
bp_both_ggallus <- sig_both_ggallus %>% filter(grepl("GO:BP", source))

go_topbp(sig_bio1_hsapiens)
go_topbp(sig_ndvi_hsapiens)
go_topbp(sig_both_hsapiens)
go_topbp(sig_bio1_ggallus)
go_topbp(sig_ndvi_ggallus)
go_topbp(sig_both_ggallus)

library(gt)
gtsave(go_table(bp_bio1_hsapiens), here(plotpath, "bp_bio1_hsapiens.html"))
gtsave(go_table(bp_ndvi_hsapiens), here(plotpath, "bp_ndvi_hsapiens.html"))
gtsave(go_table(bp_both_hsapiens), here(plotpath, "bp_both_hsapiens.html"))
gtsave(go_table(bp_ndvi_ggallus), here(plotpath, "bp_ndvi_ggallus.html"))
gtsave(go_table(bp_bio1_ggallus), here(plotpath, "bp_bio1_ggallus.html"))
gtsave(go_table(bp_bio1_ggallus), here(plotpath, "bp_bio1_ggallus.html"))


# gtsave(go_table(arrange(bp_both_hsapiens, desc(precision))), here(plotpath, "bp_both_hsapiens.html"))
# gtsave(go_table(arrange(bp_both_hsapiens, fish_p)), here(plotpath, "bp_both_hsapiens.html"))
# gtsave(go_table(arrange(bp_both_hsapiens, desc(recall)), n = 10), here(plotpath, "bp_both_hsapiens.html"))

# gtsave(go_table(arrange(bp_both_ggallus, desc(precision))), here(plotpath, "bp_both_ggallus.html"))
# gtsave(go_table(arrange(bp_both_ggallus, fish_p)), here(plotpath, "bp_both_ggallus.html"))
# gtsave(go_table(arrange(bp_both_ggallus, desc(recall)), n = 10), here(plotpath, "bp_both_ggallus.html"))

# gtsave(go_table(arrange(bp_bio1_ggallus, desc(recall)), n = 10), here(plotpath, "bp_bio1_ggallus.html"))
# gtsave(go_table(arrange(bp_ndvi_ggallus, desc(recall)), n = 10), here(plotpath, "bp_ndvi_ggallus.html"))

```