---
output: html_document
date: "2023-10-06"
---
```{r}
library(rvertnet)
library(readxl)
library(tidyverse)
library(sf)
library(tigris)
```
```{r}
# Download the California shapefile
california <- states(cb = TRUE, resolution = "20m", class = "sf") %>%
  filter(STUSPS == "CA")

# Plot the California map
ggplot() +
  geom_sf(data = california) +
  labs(title = "California Map")
```

# Scanning plan
P1 - MVZ CCGP specimen 
P2 - LACNHM/CalAcad CCGP specimen →  check age class/ sex and that there is a specimen, it is not just a tissue → loan request actual specimens
Everything remaining will not have CCGP genomic data
P3 - find MVZ specimens that are close to the CCGP specimens (5 - 10km) - that have tissues (that will give us specimens to quantify phenotypic diversity nearby and then also potentially get more genomic data) - INCLUDE ones with and without specimens so that we know what we have for sites, MAKE SURE THE SUGGESTED ONES ARE NOT OTHER CCGP SPECIMENS
P3A ADULT/MALE/5km
P3B NO AGE CLASS DATA OR ADULT/ANY SEX/10km 
P4 - find LACNHM/CalAcad specimens that are close to the CCGP specimens (5 - 10km)
P5 - look for any uncaptured environmental diversity across the state (unlikely) for which we would get specimens


# Read in CCGP sequencing data and CT specimens
```{r}
ccgp_seq <- read_table("data/58-Sceloporus.coords.txt", col_names = FALSE) 
names(ccgp_seq) <- c("SampleID", "x", "y")
tracker <- ccgp_seq

ct_scanned <- 
  read_csv("data/Copy of CT_scanning - Sheet1.csv") %>%
  mutate(SampleID = paste0(museum_ID, specimen_no))
```
# VertNet

Filter to SCOC, California, MVZ, has coordinates, age class = adult, sex = M (?), year > 1995, has tissues
```{r}
# read in scelop data
cas <-
  searchbyterm(
    scientificname = "Sceloporus occidentalis",
    stateprovince = "California",
    mappable = TRUE,
    vntype = "specimen",
    #sex = "male",
    #year = '>=1995',
    #lifestage = "adult",
    institutioncode = "CAS",
    limit = 10000,
    hastissue = TRUE
  )

cas$data <- 
  cas$data %>% 
  mutate(x = as.numeric(decimallongitude), 
         y = as.numeric(decimallatitude),
         hasSpecimen = TRUE, 
         SampleID = paste0(institutioncode, catalognumber))

lacm <-
  searchbyterm(
    scientificname = "Sceloporus occidentalis",
    stateprovince = "California",
    mappable = TRUE,
    vntype = "specimen",
    #sex = "male",
    #year = '>=1995',
    #lifestage = "adult",
    institutioncode = "LACM",
    limit = 10000,
    hastissue = TRUE
  )

lacm$data <- 
  lacm$data %>% 
  mutate(x = as.numeric(decimallongitude), 
         y = as.numeric(decimallatitude),
         hasSpecimen = TRUE, 
         SampleID = paste0(institutioncode, catalognumber)) %>%
  # remove any without catalogs (just tissue)
  filter(!is.na(catalognumber))

# LOOKING AT NOTES EVERYTHING IS TISSUE

# bigsearch(
#     scientificname = "Sceloporus occidentalis",
#     stateprovince = "California",
#     mappable = TRUE,
#     vntype = "specimen",
#     # couldn't filter by sex:
#     #sex = "male",
#     year = '>=1995',
#     institutioncode = "MVZ",
#     email = "anusha.bishop@berkeley.edu",
#     rfile = "MVZ_specimens",
#     hastissue = TRUE
# )

mvz <- read_csv("data/MVZ_rvertnet.csv")
mvz <- 
  mvz %>%
  mutate(
    x = as.numeric(decimallongitude),
    y = as.numeric(decimallatitude),
    hasSpecimen = TRUE,
    SampleID = gsub(":Herp:", "", catalognumber)
  )


```

# P1: CCGP scanned (All MVZ)
```{r}
p1 <- 
  left_join(ccgp_seq, ct_scanned, by = "SampleID") %>%
  drop_na(specimen_no)

tracker1 <- 
  tracker %>%
  mutate(instid = gsub("[0-9]", "", SampleID)) %>%
  mutate(p1 = case_when(SampleID %in% p1$SampleID ~ "scanned", TRUE ~ NA))
```

# P2: CCGP not scanned, from LACM/CAS
```{r}
ccgp_lacm_samples <- 
  ccgp_seq %>%
  filter(grepl("LACM", SampleID)) %>%
  left_join(lacm$data, by = "SampleID")

stopifnot(all(ccgp_lacm_samples$hasSpecimen))
unique(ccgp_lacm_samples$lifestage)

ccgp_cas_samples <- 
  ccgp_seq %>%
  filter(grepl("CAS", SampleID)) %>%
  left_join(cas$data, by = "SampleID")

stopifnot(all(ccgp_cas_samples$hasSpecimen))
unique(ccgp_cas_samples$lifestage)

tracker2 <-
  tracker1 %>%
  mutate(p2 = case_when(SampleID %in% ccgp_cas_samples$SampleID ~ "request from CAS", 
                        SampleID %in% ccgp_lacm_samples$SampleID ~ "request from LACM"))
```

```{r}

close_specimens <- function(i, samples_sf, specimens, buffer = 10000){
  sample <- samples_sf[i,]
  
  # get distances
  dists <- as.numeric(st_distance(sample, specimens))
  names(dists) <- specimens$SampleID
  dists <- dists[as.numeric(dists) < buffer]
  
  if (length(dists) ==0) return(data.frame(SampleID = sample$SampleID))
  
  # get ID
  dists <- dists[order(dists)]
  sid <- names(dists)
  sid <- sid[!(sid %in% samples_sf$SampleID)]
  if (length(sid) == 0) return(data.frame(SampleID = sample$SampleID))
  sid <- t(sid)
  colnames(sid) <- paste0("Specimen", 1:ncol(sid))
  
  return(data.frame(SampleID = sample$SampleID, sid))
}

close_museum <- function(samples, museums){
  samples_sf <- 
    st_as_sf(samples, coords = c("x", "y")) %>% 
    st_set_crs(st_crs(4326)) %>%
    st_transform(st_crs(3310))

  museums_sf <- 
    st_as_sf(museums, coords = c("x", "y")) %>% 
    st_set_crs(st_crs(4326)) %>%
    st_transform(st_crs(3310))

  get_samples <- 
    map(1:nrow(samples_sf), ~close_specimens(.x, samples_sf, museums_sf)) %>%
    bind_rows() 
  
  samples_df <- 
    samples %>%
    left_join(get_samples, by = "SampleID") %>%
    mutate(p3 = case_when(is.na(p1) & is.na(p2) & !is.na(Specimen1) ~ "nearby"),
           near_specimen = Specimen1)
  
  near_coords <-
    samples_df %>%
    left_join(museums, by = c("near_specimen" = "SampleID")) %>%
    select(x.y, y.y, near_specimen, lifestage, sex) %>%
    mutate(near_instid = gsub("[0-9]", "", near_specimen)) %>%
    rename(near_x = x.y, near_y = y.y, near_lifestage = lifestage, near_sex = sex) %>%
    distinct()
  
  samples_df <-
    # need to figure out how to deal with duplicates
    left_join(samples_df, near_coords, by = "near_specimen")
  
  return(samples_df)
}


museums <- 
  bind_rows(select(mvz, x, y, SampleID, lifestage, sex),
            select(lacm$data, x, y, SampleID, lifestage, sex),
            select(cas$data, x, y, SampleID, lifestage, sex)) %>%
  # remove duplicates
  # check for duplicates: 
  #museums %>% filter(duplicated(.))
  #filter(lifestage == "Adult", sex == "male") %>%
  filter(!(lifestage %in% c("juvenile", "hatchling", "subadult", "Juv"))) %>%
  distinct()

tracker3 <- 
  close_museum(tracker2, museums) %>%
  mutate(p3 = case_when(!is.na(p3) ~ paste(p3, near_instid))) %>%
  mutate(note = case_when(!is.na(p1) ~ p1,
                          !is.na(p2) ~ p2,
                          !is.na(p3) ~ p3)) %>%
  left_join(select(museums, SampleID, sex, lifestage), by = "SampleID") %>%
  mutate(genome_and_specimen = !is.na(p1) | !is.na(p2),
         genome_and_specimenORnearby = !is.na(note))

ggplot() + 
  geom_sf(data = california) +
  geom_point(data = tracker3, aes(x = x, y = y, col = note)) +
  theme_void()

ggplot() + 
  geom_sf(data = california) +
  geom_point(data = tracker3, aes(x = x, y = y, col = genome_and_specimen)) +
  theme_void()

ggplot() + 
  geom_sf(data = california) +
  geom_point(data = tracker3[!is.na(tracker3$p3),],
             aes(x = near_x, y = near_y, col = near_lifestage, pch = near_sex)) +
  theme_void()

ggplot() + 
  geom_sf(data = california) +
  geom_point(data = tracker3[is.na(tracker3$p3),],
             aes(x = x, y = y, col = lifestage, pch = sex)) +
  theme_void()
```


```{r}
ggplot() + 
  geom_sf(data = california) +
  geom_point(data = lacm$data[lacm$data$x < 110, ],
             aes(x = x, y = y, col = lifestage, pch = sex)) +
  theme_void()

ggplot() + 
  geom_sf(data = california) +
  geom_point(data = mvz,
             aes(x = x, y = y, col = lifestage, pch = sex)) +
  theme_void()

ggplot() + 
  geom_sf(data = california) +
  geom_point(data = cas$data,
             aes(x = x, y = y, col = lifestage, pch = sex)) +
  theme_void()
```

