---
output: html_document
date: "2023-10-06"
---
```{r}
library(rvertnet)
library(readxl)
library(tidyverse)
library(sf)
```
```{r}
# Load the tigris package
library(tigris)

# Download the California shapefile
california <- states(cb = TRUE, resolution = "20m", class = "sf") %>%
  filter(STUSPS == "CA")

# Plot the California map
ggplot() +
  geom_sf(data = california) +
  labs(title = "California Map")
```

# Scanning plan
P1 - MVZ CCGP specimen 
P2 - LACNHM/CalAcad CCGP specimen →  check age class/ sex and that there is a specimen, it is not just a tissue → loan request actual specimens
Everything remaining will not have CCGP genomic data
P3 - find MVZ specimens that are close to the CCGP specimens (5 - 10km) - that have tissues (that will give us specimens to quantify phenotypic diversity nearby and then also potentially get more genomic data) - INCLUDE ones with and without specimens so that we know what we have for sites, MAKE SURE THE SUGGESTED ONES ARE NOT OTHER CCGP SPECIMENS
P3A ADULT/MALE/5km
P3B NO AGE CLASS DATA OR ADULT/ANY SEX/10km 
P4 - find LACNHM/CalAcad specimens that are close to the CCGP specimens (5 - 10km)
P5 - look for any uncaptured environmental diversity across the state (unlikely) for which we would get specimens


# Read in CCGP sequencing data and CT specimens
```{r}
ccgp_seq <- read_table("data/58-Sceloporus.coords.txt", col_names = FALSE) 
names(ccgp_seq) <- c("SampleID", "x", "y")
tracker <- ccgp_seq

ct_scanned <- 
  read_csv("data/Copy of CT_scanning - Sheet1.csv") %>%
  mutate(SampleID = paste0(museum_ID, specimen_no))
```
# VertNet

Filter to SCOC, California, MVZ, has coordinates, age class = adult, sex = M (?), year > 1995, has tissues
```{r}
# read in scelop data
cas <-
  searchbyterm(
    scientificname = "Sceloporus occidentalis",
    stateprovince = "California",
    mappable = TRUE,
    vntype = "specimen",
    #sex = "male",
    #year = '>=1995',
    #lifestage = "adult",
    institutioncode = "CAS",
    limit = 10000,
    hastissue = TRUE
  )

lacm <-
  searchbyterm(
    scientificname = "Sceloporus occidentalis",
    stateprovince = "California",
    mappable = TRUE,
    vntype = "specimen",
    #sex = "male",
    #year = '>=1995',
    #lifestage = "adult",
    institutioncode = "LACM",
    limit = 10000,
    hastissue = TRUE
  )

mvz <- read_csv("data/MVZ_rvertnet.csv")

bigsearch(
    scientificname = "Sceloporus occidentalis",
    stateprovince = "California",
    mappable = TRUE,
    vntype = "specimen",
    # couldn't filter by sex:
    #sex = "male",
    year = '>=1995',
    institutioncode = "MVZ",
    email = "anusha.bishop@berkeley.edu",
    rfile = "MVZ_specimens",
    hastissue = TRUE
)

bigsearch(
  Genus = "Sceloporus",
  SpecificEpithet = "occidentalis",
  stateprovince = "California",
  mappable = TRUE,
  vntype = "specimen",
  # couldn't filter by sex:
  #sex = "male",
  year = '>=1995',
  institutioncode = "LACM",
  email = "anusha.bishop@berkeley.edu",
  rfile = "LACM_specimens",
  hastissue = TRUE
)
bigsearch(
  Genus = "Sceloporus",
  SpecificEpithet = "occidentalis",
  stateprovince = "California",
  mappable = TRUE,
  vntype = "specimen",
  # couldn't filter by sex:
  #sex = "male",
  year = '>=1995',
  institutioncode = "CAS",
  email = "anusha.bishop@berkeley.edu",
  rfile = "CAS_specimens",
  hastissue = TRUE
)
```

# P1: CCGP scanned (All MVZ)
```{r}
p1 <- 
  left_join(ccgp_seq, ct_scanned, by = "SampleID") %>%
  drop_na(specimen_no)

tracker1 <- 
  tracker %>%
  mutate(instid = gsub("[0-9]", "", SampleID)) %>%
  mutate(p1 = case_when(SampleID %in% p1$SampleID ~ "scanned", TRUE ~ NA))
```

# P2: CCGP not scanned, from LACM/CAS
```{r}
ccgp_lacm_samples <- 
  ccgp_seq %>%
  filter(grepl("LACM", SampleID))

lacm$data$hasSpecimen <- TRUE
lacm$data$SampleID <- paste0(lacm$data$institutioncode, lacm$data$catalognumber)

ccgp_lacm_samples <- 
  left_join(ccgp_lacm_samples, lacm$data)

stopifnot(all(ccgp_lacm_samples$hasSpecimen))

ccgp_cas_samples <- 
  ccgp_seq %>%
  filter(grepl("CAS", SampleID))

cas$data$hasSpecimen <- TRUE
cas$data$SampleID <- paste0(cas$data$institutioncode, cas$data$catalognumber)

ccgp_cas_samples <- 
  left_join(ccgp_cas_samples, cas$data)

stopifnot(all(ccgp_cas_samples$hasSpecimen))

tracker2 <-
  tracker1 %>%
  mutate(p2 = case_when(SampleID %in% ccgp_cas_samples$SampleID ~ "request from CAS", 
                        SampleID %in% ccgp_lacm_samples$SampleID ~ "request from LACM"))
```

```{r}

tracker_sf <- 
  st_as_sf(tracker2, coords = c("x", "y")) %>% 
  st_set_crs(st_crs(4326)) %>%
  st_transform(st_crs(3310))

mvz$SampleID <- gsub(":Herp:", "", mvz$catalognumber)
mvz_sf <- 
  st_as_sf(mvz, coords = c("decimallongitude", "decimallatitude")) %>% 
  st_set_crs(st_crs(4326)) %>%
  st_transform(st_crs(3310))

close_specimen <- function(i, tracker_sf, mvz_sf, buffer = 10000){
  ccgp_sample <- tracker_sf[i,]
  
  # get distances
  dists <- as.numeric(st_distance(ccgp_sample, mvz_sf))
  names(dists) <- mvz_sf$SampleID
  dists <- dists[as.numeric(dists) < buffer]
  
  if (length(dists) ==0) return(data.frame(SampleID = ccgp_sample$SampleID))
  
  # get ID
  dists <- dists[order(dists)]
  sid <- names(dists)
  sid <- sid[!(sid %in% tracker_sf$SampleID)]
  if (length(sid) == 0) return(data.frame(SampleID = ccgp_sample$SampleID))
  sid <- t(sid)
  colnames(sid) <- paste0("Specimen", 1:ncol(sid))
  
  return(data.frame(SampleID = ccgp_sample$SampleID, sid))
}

get_samples <- 
  map(1:nrow(tracker_sf), ~close_specimen(.x, tracker_sf, mvz_sf)) %>%
  bind_rows() 

tracker3 <- 
  tracker2 %>%
  left_join(get_samples) %>%
  mutate(p3 = case_when(is.na(p1) & is.na(p2) & !is.na(Specimen1) ~ "nearby MVZ specimen"),
         near_specimen = Specimen1)

near_coords <-
  tracker3 %>%
  mutate(SampleID = near_specimen) %>%
  left_join(mvz) %>%
  select(decimallatitude, decimallongitude, near_specimen) %>%
  rename(near_x = decimallongitude, near_y = decimallatitude)

tracker3 <-
  left_join(tracker3, near_coords)

tracker3 <- 
  tracker3 %>%
  mutate(note = case_when(!is.na(p1) ~ p1,
                           !is.na(p2) ~ p2,
                           !is.na(p3) ~ p3))
ggplot() + 
  geom_sf(data = california) +
  geom_point(data = tracker3, aes(x = x, y = y, col = note)) +
  theme_void()

```


# next steps : check near LACM
# check near: CAS

