---
output: 
  html_document
---

```{r setup, include=FALSE}
# Set default chunk options for all code chunks
knitr::opts_chunk$set(
  echo = FALSE,      # Do not display code by default
  message = FALSE,   # Do not display messages by default
  warning = FALSE, # Do not display warnings by default
  result = FALSE
  )
```

```{r}
library(rvertnet)
library(readxl)
library(tidyverse)
library(sf)
library(tigris)
library(ggpubr)
```

```{r, include = FALSE}
# Download California shapefile for plotting
california <-
  states(cb = TRUE, resolution = "20m", class = "sf") %>%
  filter(STUSPS == "CA")
```

# Scanning plan

P1 - MVZ CCGP specimen 

P2 - LACNHM/CalAcad CCGP specimen →  check age class/ sex and that there is a specimen, it is not just a tissue → loan request actual specimens

*Everything remaining will not have CCGP genomic data*

P3 - find museum specimens that are close to the CCGP specimens (5 - 10km) - that have tissues (that will give us specimens to quantify phenotypic diversity nearby and then also potentially get more genomic data) - INCLUDE ones with and without specimens so that we know what we have for sites, MAKE SURE THE SUGGESTED ONES ARE NOT OTHER CCGP SPECIMENS

P3A ADULT/MALE/5km

P3B NO AGE CLASS DATA OR ADULT/ANY SEX/10km 

P4 - look for any uncaptured environmental diversity across the state (unlikely) for which we would get specimens


# Read in CCGP sequencing data and CT specimens
```{r}
# CCGP samples 
ccgp_seq <- read_table("data/58-Sceloporus.coords.txt", col_names = FALSE) 
names(ccgp_seq) <- c("SampleID", "x", "y")

# Erin's CT specimens
ct_scanned <- 
  read_csv("data/Copy of CT_scanning - Sheet1.csv") %>%
  mutate(SampleID = paste0(museum_ID, specimen_no))

# I am going to use this df called "tracker" to keep track of the specimens and what type they are)
tracker <- ccgp_seq
```

# VertNet Data

Filters:

- Species

- California

- MVZ/CAS/LACM

- has coordinates

- year > 1995

- has tissues

- Late on I will filter for age class = adult, sex = M

```{r, cache = TRUE}
# CAS
cas <-
  searchbyterm(
    scientificname = "Sceloporus occidentalis",
    stateprovince = "California",
    mappable = TRUE,
    vntype = "specimen",
    year = '>=1995',
    institutioncode = "CAS",
    limit = 10000,
    hastissue = TRUE
  )

cas$data <- 
  cas$data %>% 
  mutate(x = as.numeric(decimallongitude), 
         y = as.numeric(decimallatitude),
         hasSpecimen = TRUE, 
         # format ID to match CCGP
         SampleID = paste0(institutioncode, catalognumber))

# LACM
lacm <-
  searchbyterm(
    scientificname = "Sceloporus occidentalis",
    stateprovince = "California",
    mappable = TRUE,
    vntype = "specimen",
    year = '>=1995',
    institutioncode = "LACM",
    limit = 10000,
    hastissue = TRUE
  )

# note: the LACM query doesn't seem to always properly filter full specimens only, 
# make sure to double check that there are actually specimens
lacm$data <- 
  lacm$data %>% 
  mutate(x = as.numeric(decimallongitude), 
         y = as.numeric(decimallatitude),
         hasSpecimen = TRUE, 
         # format ID to match CCGP
         SampleID = paste0(institutioncode, catalognumber)) %>%
  # remove any without catalogs (just tissue samples)
  filter(!is.na(catalognumber))


mvz <-
  searchbyterm(
    scientificname = "Sceloporus occidentalis",
    stateprovince = "California",
    mappable = TRUE,
    vntype = "specimen",
    year = '>=1995',
    institutioncode = "MVZ",
    limit = 10000,
    hastissue = TRUE
  )

mvz$data <- 
  mvz$data %>% 
  # format ID to match CCGP
  mutate(x = as.numeric(decimallongitude), 
         y = as.numeric(decimallatitude),
         hasSpecimen = TRUE, 
         SampleID = gsub(":Herp:", "", catalognumber)) 
```

# P1: CCGP scanned (All MVZ)
```{r}
# join CCGP and Erin CT data to determine scanned specimens
p1 <- 
  left_join(ccgp_seq, ct_scanned, by = "SampleID") %>%
  drop_na(specimen_no)

tracker1 <- 
  tracker %>%
  # update tracker with information about which samples have been scanned
  mutate(p1 = case_when(SampleID %in% p1$SampleID ~ "scanned", TRUE ~ NA))

# TODO: EW why does it seem like there are so few scanned specimens? Can you double check I have formatted the SampleIDs correctly?
tracker1 %>% filter(!is.na(p1))
```

# P2: CCGP not scanned, from LACM/CAS
```{r}
# join CCGP data with LACM records to get LACM specimens
ccgp_lacm <- 
  ccgp_seq %>%
  filter(grepl("LACM", SampleID)) %>%
  left_join(lacm$data, by = "SampleID")

# check if all of the samples have specimens
stopifnot(all(ccgp_lacm$basisofrecord == "PreservedSpecimen"))

# check lifestages
unique(ccgp_lacm$lifestage)

# join CCGP data with CAS records to get CAS specimens
ccgp_cas <- 
  ccgp_seq %>%
  filter(grepl("CAS", SampleID)) %>%
  left_join(cas$data, by = "SampleID")

# check if all of the samples have specimens
stopifnot(all(ccgp_cas$basisofrecord == "PreservedSpecimen"))

# check lifestages
# TODO: EW is Juvenile/Adult (J/A) okay or should we find a replacement?
unique(ccgp_cas$lifestage)

# update tracker to include specimens that need to be requested
tracker2 <-
  tracker1 %>%
  mutate(p2 = case_when(SampleID %in% ccgp_cas$SampleID ~ "request CAS", 
                        SampleID %in% ccgp_lacm$SampleID ~ "request LACM"))
```

# P3: Finding close museum specimens

```{r}
# Function to find close museum specimens
#
# samples - samples data frame
# museums - museum specimens data frame
close_museum <- function(samples, museums){
  
  # Convert samples and museums to spatial objects with a common projected CRS
  # (this is necessary for calculating distances)
  samples_sf <- 
    st_as_sf(samples, coords = c("x", "y")) %>% 
    # WGS 84 - lat/long
    st_set_crs(st_crs(4326)) %>%
    # CA Albers
    st_transform(st_crs(3310))

  museums_sf <- 
    st_as_sf(museums, coords = c("x", "y")) %>% 
    # WGS 84 - lat/long
    st_set_crs(st_crs(4326)) %>%
    # CA Albers
    st_transform(st_crs(3310))

  # Get close specimens and make into df
  get_samples <- 
    map(1:nrow(samples_sf), ~close_specimens(.x, samples_sf, museums_sf)) %>%
    bind_rows() 
  
  # Merge close samples with original samples data frame
  samples_df <- 
    samples %>%
    left_join(get_samples, by = "SampleID") %>%
    # add tracker information that these are nearby specimens
    # only adds this if p1/p2 are NA (i.e., the original specimen is not available)
    mutate(p3 = case_when(is.na(p1) & is.na(p2) & !is.na(Specimen1) ~ "request nearby"),
           # make a column for the nearest specimen (Specimen1)
           near_specimen = Specimen1)
  
  # Extract coordinates and attributes of nearest specimen
  near_coords <-
    samples_df %>%
    # get data on near specimens from the museum df
    left_join(museums, by = c("near_specimen" = "SampleID")) %>%
    # take only the cols you need
    select(x.y, y.y, near_specimen, lifestage, sex) %>%
    # get institution ID from SampleID
    mutate(near_instid = gsub("[0-9]", "", near_specimen)) %>%
    # rename to specify this is data for the near specimen 
    rename(near_x = x.y, near_y = y.y, near_lifestage = lifestage, near_sex = sex) %>%
    # sometimes the nearest sample will be the same for multiple samples
    # only keep distinct rows so that the left_join doesn't give you the multiple x/y warning
    distinct()
  
  # Merge back into the samples data frame
  samples_df <- left_join(samples_df, near_coords, by = "near_specimen")
  
  return(samples_df)
}


# Function to find nearest specimens to a sample
# i - sample index (used in a loop)
# samples_sf - sf object of CCGP samples (tracker df)
# museums_sf - sf onject of museum specimens (musem df)
# buffer - furthest distance from which to take a sample (defaults to 10000m = 10km)
close_specimens <- function(i, samples_sf, museums_sf, buffer = 10000){
  # get single sample
  sample <- samples_sf[i,]
  
  # get distances from sample to museum specimens
  dists <- as.numeric(st_distance(sample, museums_sf))
  names(dists) <- museums_sf$SampleID
  
  # get samples that fall within buffer distance
  dists <- dists[as.numeric(dists) < buffer]
  
  # check if there are nearby ccgp samples with specimens
  ccgp_dist <- as.numeric(st_distance(sample,
                                      samples_sf[!is.na(samples_sf$p1) | !is.na(samples_sf$p2),]))
  
  # get second closest dist (first is the same sample)
  nearest_dist <- ccgp_dist[order(ccgp_dist)][2]
  
  # if there are no samples that fall within the buffer, return a df with just the ID/near dist
  # (to keep track of what was checkd)
  if (length(dists) == 0) return(data.frame(SampleID = sample$SampleID, nearest_dist = nearest_dist))
  
  # order distances from shortest to longest
  dists <- dists[order(dists)]
  
  # get SampleIDs from distances
  sid <- names(dists)
  
  # remove any museum specimens that are already CCGP samples
  sid <- sid[!(sid %in% samples_sf$SampleID)]
  
  # if there are no samples that are not CCGP samples, return a df with just the ID/near dist
  if (length(sid) == 0) return(data.frame(SampleID = sample$SampleID, nearest_dist = nearest_dist))
  
  # transpose so that specimen options are columns
  sid <- t(sid)
  colnames(sid) <- paste0("Specimen", 1:ncol(sid))
  
  return(data.frame(SampleID = sample$SampleID, sid, nearest_dist = nearest_dist))
}
```

```{r}

museums <- 
  # combine museum data
  bind_rows(select(mvz$data, x, y, SampleID, lifestage, sex),
            select(lacm$data, x, y, SampleID, lifestage, sex),
            select(cas$data, x, y, SampleID, lifestage, sex)) %>%
  # remove any juvis
  filter(!(lifestage %in% c("juvenile", "hatchling", "subadult", "Juv"))) %>%
  # remove any duplicates
  # if you want to check for duplicates: museums %>% filter(duplicated(.))
  distinct()

# get museum specimens that are close to ccgp samples
close_museum_specimens <- close_museum(tracker2, museums)

# update tracker to with close specimens data
tracker3 <- 
  close_museum_specimens %>%
  # this formats p3 like "nearby [MUSEUM CODE]"
  mutate(p3 = case_when(!is.na(p3) ~ paste(p3, near_instid)))
```

# Visualize

```{r}
# clean up tracker for plotting
tracker_final <- 
  tracker3 %>%
  # get sex and lifestage from museums df
  left_join(select(museums, SampleID, sex, lifestage), by = "SampleID") %>%
  mutate(
    # create museum code column
    instid = gsub("[0-9]", "", SampleID),
    # create note that combines P1/P2/P3 into one column
    note = case_when(!is.na(p1) ~ p1, !is.na(p2) ~ p2, !is.na(p3) ~ p3),
    # create priority column
    priority = case_when(!is.na(p1) ~ "P1",!is.na(p2) ~ "P2",!is.na(p3) ~ "P3"),
    # keep track of museums for requests
    near_museum = case_when(is.na(p1) ~ gsub(" ", "", gsub("nearby", "", gsub("request", "", note) ))),
    # keep track of which CCGP samples have specimens that we can access
    genome_and_specimen = !is.na(p1) | !is.na(p2),
    # keep track of which CCGP samples have specimens or nearby options
    genome_and_specimenORnearby = !is.na(note),
    # determine samples that could potentially be dropped from scanning because 
    # there is another CCGP sample with a specimen nearby
    potential_drop = case_when(nearest_dist < 50000 & (!is.na(p3) | is.na(note)) ~ TRUE, TRUE ~ FALSE)
  ) 
```

```{r}
# Count CCGP samples with specimens
table(tracker_final$genome_and_specimen)

# Count CCGP samples with specimens OR nearby options
table(tracker_final$genome_and_specimenORnearby)
```

Main notes:
- we can get a good chunk of coverage from specimens requested from CAS
- LACM doesn't have a ton of good specimens, but they fill in important gaps
- We need samples from S. CA and along the coast
```{r}
# plot of all samples and color coded by what we need to do for them
ggplot() + 
  geom_sf(data = california) +
  geom_point(data = tracker_final, aes(x = x, y = y, col = note)) +
  theme_void()

# color coded by priority
# P1 = scanned
# P2 = need to request CCGP specimen
# P3 = need to request nearby specimen
ggplot() + 
  geom_sf(data = california) +
  geom_point(data = tracker_final, aes(x = x, y = y, col = priority)) +
  theme_void()

# color coded by museum for request
ggplot() + 
  geom_sf(data = california) +
  geom_point(data = tracker_final, aes(x = x, y = y, col = near_museum)) +
  theme_void()

# color coded by what we have specimens and CCGP data for
ggplot() + 
  geom_sf(data = california) +
  geom_point(data = tracker_final, aes(x = x, y = y, col = genome_and_specimen)) +
  theme_void()
```

# Digging more into specimen requests


```{r}

requests <-
  tracker_final %>%
  # keep only P3
  filter(!is.na(p3)) %>% 
  # reformatting so that there is a column with all near specimen IDs (not just the nearest)
  pivot_longer(cols = starts_with("Specimen"), names_to = "Specimen", values_to = "near_SampleID") %>%
  # getting coords for near specimens
  left_join(select(museums, SampleID, x, y), by = c("near_SampleID" = "SampleID")) %>%
  mutate(near_x = x.y, near_y = y.y, x = x.x, y = y.x) %>%
  # removing NA rows
  drop_na(near_SampleID) %>%
  # clean up lifestage/sex data
  mutate(
    lifestage = case_when(
      lifestage == "Adult" ~ "adult",
      lifestage == "J/A" ~ "adult/juvenile",
      is.na(lifestage) ~ "undetermined" ,
      TRUE ~ lifestage
    ),
    near_lifestage = case_when(
      near_lifestage == "Adult" ~ "adult",
      near_lifestage == "J/A" ~ "adult/juvenile",
      is.na(near_lifestage) ~ "undetermined",
      TRUE ~ near_lifestage
    ),
    sex = case_when(is.na(sex) ~ "undetermined", TRUE ~ sex),
    near_sex = case_when(is.na(near_sex) ~ "undetermined", TRUE ~ near_sex),
    sex = factor(sex),
    near_sex = factor(near_sex),
    lifestage = factor(lifestage),
    near_lifestage = factor(near_lifestage))

```

```{r}
# checking that none of the near specimens have already been scanned
ct_requests <- 
  requests %>% 
  left_join(mutate(ct_scanned, scanned = TRUE), 
            by = c("near_SampleID" = "SampleID"))

stopifnot(!any(ct_requests$scanned, na.rm = TRUE))
```


It seems like including some females and unknown lifestages is potentially unavoidable because they fill important gaps, however we could probably exclude the adult/juveniles.  
```{r}
# lifestage and sex of nearby specimens to request
ggplot() + 
  geom_sf(data = california) +
  geom_point(data = requests,
             aes(x = near_x, y = near_y, pch = near_lifestage, col = near_sex)) +
  theme_void()

```

There are also some specimens we might not need to replace because there are others close by (depends on how densely we want to sample)
```{r}
ggplot() + 
  geom_sf(data = california) +
  geom_point(data = tracker_final,
             aes(x = x, y = y, col = "CCGP samples w/o any options"), cex = 1) +
  geom_point(data = tracker_final[!is.na(tracker_final$p1) | !is.na(tracker_final$p2),],
               aes(x = x, y = y, col = "CCGP samples w/ specimens"), cex = 1) +
  geom_point(data = requests,
             aes(x = x, y = y, col = "CCGP samples that need replacement")) +
   geom_point(data = requests[requests$potential_drop,],
             aes(x = x, y = y, col = "CCGP samples that need replacement,\nbut have other CCGP specimens within 50km")) +
  theme_void() +
  scale_color_manual(values = c("orange2", "limegreen", "black", "gray"))
```


Looking at each point individually, it looks like Specimen 1 (the closest specimen; outlined in black in the plot below) is a good pick in all cases (i.e., there aren't cases where another close specimen is better because it is a male or has a classified life stage)
```{r, fig.width = 10, fig.height = 10}
# yes I know this is disgusting figure code...but it works
# Create a list of ggplot objects
plts <- map(1:length(unique(requests$SampleID)), ~ {
  
  # Subset for a specific 'SampleID'
  tmp <- requests %>% filter(SampleID == unique(requests$SampleID)[.x])
  
  # Create a plot  object for the current 'SampleID'
  plt <- ggplot() +
  
    # Add CA background
    geom_sf(data = california) +
    
    # Add points for females (hidden, but helps w/ legend formatting)
    geom_point(
      data = requests[requests$near_sex == "female", ],
      aes(x = near_x, y = near_y, col = "Female replacement"),
      cex = 2
    ) +
    
    # Add points for undetermined lifestage females (hidden, but helps w/ legend formatting)
    geom_point(
      data = requests[requests$near_lifestage == "undetermined", ],
      aes(x = near_x, y = near_y, col = "Replacement w/ undetermined lifestage"),
      cex = 2
    ) +
    
    # Add points:
    geom_point(data = tracker_final[is.na(tracker_final$note),],
               aes(x = x, y = y, col = "CCGP samples w/o any options"),
               cex = 1) +
    geom_point(data = tracker_final[!is.na(tracker_final$p1) |
                                    !is.na(tracker_final$p2), ],
               aes(x = x, y = y, col = "CCGP samples w/ specimens"),
               cex = 1) +
    geom_point(data = tmp,
               aes(x = x, y = y, col = "Focal CCGP sample"),
               cex = 4) +
    geom_point(
      data = tmp,
      aes(x = near_x, y = near_y, col = "Perfect replacements"),
      cex = 3
    ) +
    geom_point(
      data = tmp[tmp$near_sex == "female", ],
      aes(x = near_x, y = near_y, col = "Female replacement"),
      cex = 3
    ) +
    geom_point(
      data = tmp[tmp$near_lifestage == "undetermined", ],
      aes(x = near_x, y = near_y, col = "Replacement w/ undetermined lifestage"),
      cex = 3
    ) +
    
    # Add black outline for the currently selected top specimen 
    geom_point(
      data = tmp[tmp$Specimen == "Specimen1", ],
      aes(x = near_x, y = near_y, pch = "Current replacement selection"),
      col = "black",
      cex = 3
    ) +
    
    # Set the plot theme to void (no background)
    theme_void() +
    
    # Set the x and y axis limits based on the current 'SampleID' data to zoom in on the map
    xlim(tmp$x[1] - 0.2,  tmp$x[1] + 0.2) +
    ylim(tmp$y[1] - 0.2,  tmp$y[1] + 0.2) +
    
    # Set the plot title to the current 'SampleID'
    ggtitle(tmp$SampleID) +
    
    # Customize color scales
    scale_color_manual(
      values = c(
        "CCGP samples w/o any options" = "gray",
        "CCGP samples w/ specimens" = "black",
        "Focal CCGP sample" = "orange",
        "Perfect replacements" = "cyan3",
        "Female replacement" = "red",
        "Replacement w/ undetermined lifestage" = "magenta"
      ),
      breaks = c(
        "CCGP samples w/o any options",
        "CCGP samples w/ specimens",
        "Focal CCGP sample",
        "Perfect replacements",
        "Female replacement",
        "Replacement w/ undetermined lifestage"
      )
    ) +
    
    scale_shape_manual(values = c("Current replacement selection" = 1))
  
  # Return the current ggplot object
  return(plt)
})
suppressWarnings(do.call(ggarrange, c(plts, nrow = 4, ncol = 4, common.legend = TRUE, legend = "top")))
```

# P4: Identifying gaps

Big gaps:

- Southern CA

- Coastal region just below Bay Area

```{r}
# color coded by priority
# P1 = scanned
# P2 = need to request CCGP specimen
# P3 = need to request nearby specimen
ggplot() + 
  geom_sf(data = california) +
  geom_point(data = tracker_final, aes(x = x, y = y, col = priority)) +
  theme_void()

# color coded by museum for nearest specimen request
ggplot() + 
  geom_sf(data = california) +
  geom_point(data = tracker_final, aes(x = x, y = y, col = near_museum)) +
  theme_void()
```
```{r}
# Counts for each P
tracker_final %>% group_by(note) %>% summarize(n = n())

# TRUE = samples we should have if we can get all the Ps
tracker_final %>% group_by(!is.na(note)) %>% summarize(n = n())
```
```{r}

```

Next steps:
1. Request from CAS
2. Request from LACM
3. Think about field work in the Spring (check for funding)

# NEXT: CHECK to see if close specimens have been scanned by Erin
