
```{r}
library("here")
library("tidyverse")
library("ggpubr")
library("terra")
library("sf")
library("gridExtra")
library("rpaleoclim")
library("spdep")
library("spatialreg")
library("MuMIn")
library("cowplot")
library("performance")

source(here("general_functions.R"))
source(here("analysis", "genetic_diversity", "genetic_diversity.R"))
source(here("analysis", "genetic_diversity", "paleovars.R"))
ca <- get_ca()
outpath <- here("analysis", "genetic_diversity", "outputs")
plotpath <- here("analysis", "genetic_diversity", "plots")

coords <- get_coords(sf = TRUE)
het <- get_het()
dem <- get_dem(r= TRUE)
```

# Hypothesis 1: Genetic diversity explained by paleoclimate

```{r}
# Get current climate data (NOTE: not using paleoclim data because they use V1.2 instead of V2.1)
bioclim <- rast(here("data", "env", "california_chelsa_bioclim_1981-2010_V.2.1.tif"))
cur <- bioclim[["CHELSA_bio1_1981-2010_V.2.1"]]

# Get LGM climate data (NOTE: not using paleoclim data because they use V1.2 instead of V2.1)
lgm <- rast(here("data", "env", "california_chelsa_pmip_ccsm4.tif"))

# Get paleoclimate data for LIG
lig <- get_paleovars("lig")$lig$bio_1
# Convert from C/10 to C
lig <- lig/10

# Calculate climate change
cur_lgm <- cur - lgm
# Note: LIG is lower resolution than current and LGM data, so we need to resample it
lgm_lig <- lgm - resample(lig, lgm)

# Make a data frame with the climate change values
paleoclim_mod_df <- 
  data.frame(
    SampleID = coords$SampleID,
    cur_lgm = terra::extract(cur_lgm, coords, ID = FALSE)[,1],
    lgm_lig = terra::extract(lgm_lig, coords, ID = FALSE)[,1]
    )

# All paleovariables are highly correlated
cor(paleoclim_mod_df[, - 1], use = "pairwise.complete.obs")
```

# Hypothesis 2: Genetic diversity explained by past climate stability

### Climate stability (CSI)

Original paper: 
Herrando-Moraira, S., Nualart, N., Galbany-Casals, M. et al. Climate Stability Index maps, a global high resolution cartography of climate stability from Pliocene to 2100. Sci Data 9, 48 (2022). https://doi.org/10.1038/s41597-022-01144-5

Note: Bioclimatic variables do not contain Bio 2 or Bio 5, but they do contain other variables correlated with those variables

```{r, fig.width = 10, fig.height = 2.5}
csi_past <- rast(here("data", "env", "csi", "Layers", "past", "csi_past.tif"))
csi_past <- crop(csi_past, ca)
csi_past <- mask(csi_past, ca)
csi_past <- 1 - csi_past

files <- list.files(here("data", "env", "csi", "Layers", "future"), full.names = TRUE, pattern = "csi*")
csi_future <- rast(map(files, rast))
csi_future <- crop(csi_future, ca)
csi_future <- mask(csi_future, ca)
csi_future <- 1 - csi_future

csi_stack <- c(csi_past, csi_future)
names(csi_stack) <- c("Past",  "SSP 1-2.6", "SSP 2-4.5", "SSP 3-7.0", "SSP 5-8.5")
writeRaster(csi_stack, here("data", "env", "csi", "csi.tif"), overwrite = TRUE)
csi_df <- 
  terra::as.data.frame(csi_stack, xy = TRUE) %>% 
  pivot_longer(-c(x, y), 
  names_to = "model", 
  values_to = "CSI")
```

```{r, fig.height = 4, fig.width = 4.5}
csi_vals <- extract(csi_past, coords, ID = FALSE)
df_past <- data.frame(csi_past = csi_vals[, "csi_past"], coords)
```

```{r, fig.height = 4, fig.width = 16}
csi_vals <- extract(csi_future, coords, ID = FALSE)
df_future <- 
  data.frame(csi_vals, coords) %>% 
  pivot_longer(starts_with("csi"), names_to = "future_model", values_to = "csi") %>%
  mutate(model_name = case_when(
    future_model == "csi_future_ssp126" ~ "SSP 1-2.6",
    future_model == "csi_future_ssp245" ~ "SSP 2-4.5",
    future_model == "csi_future_ssp370" ~ "SSP 3-7.0",
    future_model == "csi_future_ssp585" ~ "SSP 5-8.5"
  ))
```

```{r}
coords <- get_coords(sf = TRUE)
# Do 1 - to make the values more intuitive
csi_mod_df <- data.frame(csi = extract(csi_past, coords, ID = FALSE), coords)
```

# Hypothesis 3: Genetic diversity explained by admixture
```{r}
q <- read_csv(here("analysis", "admixture", "outputs", "Q9.csv"))

q_mod_df <- 
  q %>%
  dplyr::select(SampleID, Q) %>%
  # Do 1- to make the values more intuitive
  mutate(Q = 1 - Q)
```

# Hypothesis 4: Genetic diversity explained by current environment 
```{r}
bioclim <- rast(here("data", "env", "california_chelsa_bioclim_1981-2010_V.2.1.tif"))
ndvi <- rast(here("data", "env", "california_ndvi_mean_2000_2020.tif"))
ghm <- rast(here("data", "env", "california_global_human_modification_2016.tif"))

env_mod_df <- data.frame(
  SampleID = coords$SampleID,
  NDVI = terra::extract(ndvi, coords, ID = FALSE)[, "NDVI"],
  gHM = terra::extract(ghm, coords, ID = FALSE)[,"gHM"],
  bio1 = terra::extract(bioclim, coords, ID = FALSE)[,"CHELSA_bio1_1981-2010_V.2.1"]
)
```

# Hypothesis 5: Genetic diversity explained by glaciation
```{r}
lgm_glaciers <- st_read(here("data", "env", "Glaciotect_NA_lgm", "shp", "LGM_maxextent.shp"))
# Note: have to first intersect and then transform otherwise you get a node error
lgm_glaciers <- st_intersection(lgm_glaciers, st_transform(ca, crs = st_crs(lgm_glaciers)))
lgm_glaciers <- st_transform(lgm_glaciers, crs = st_crs(ca))
# Add a new column to coords indicating whether each point is inside the lgm shapefile
coords <- get_coords(sf = TRUE)
glacier_mod_df <- 
  coords %>%
  mutate(glacier = as.numeric((lengths(st_within(coords, lgm_glaciers)) > 0))) 
```

# Hypothesis 6: Genetic diversity explained by recent climate change
PRISM mean temperature data from historical data: https://prism.oregonstate.edu/historical/

```{r}
coords <- get_coords(sf = TRUE)
# Get prism data
get_prism <- function(year, output = "df") {
  # Create PRISM path
  prism_path <- here("data", "env", "prism", "PRISM_tmean_stable_4kmM3", year)

  # List annual file
  file <- list.files(prism_path, pattern = paste0(year, "_*.bil.bil$"), full.names = TRUE)

  # Load data
  raster <- terra::rast(file)

  if (output == "raster"){
    return(raster)
  } else {
    vals <- terra::extract(raster, st_transform(coords, st_crs(raster)), ID = FALSE)
    names(vals) <- "tmean"
    df <- data.frame(year = year, tmean = vals, SampleID = coords$SampleID)
    return(df)
  }
}

# Loop over the years of interest
tmean_his <- map(1895:1905, get_prism, .progress =TRUE) %>% bind_rows()  
tmean_his <- tmean_his %>% group_by(SampleID) %>% summarise(tmean_his = mean(tmean, na.rm = TRUE))
tmean_cur <- map(2010:2020, ~get_prism(year = .x), .progress =TRUE) %>% bind_rows()  
tmean_cur <- tmean_cur %>% group_by(SampleID) %>% summarise(tmean_cur = mean(tmean, na.rm = TRUE))

tmean_his_r <- map(1895:1905, get_prism, .progress =TRUE, output= "raster") %>% rast() %>% mean()
tmean_cur_r <- map(2010:2020, get_prism, .progress =TRUE, output= "raster") %>% rast() %>% mean()
tmean_dif_r <- tmean_cur_r - tmean_his_r
tmean_dif_r <- crop(tmean_dif_r, ca)
tmean_dif_r <- mask(tmean_dif_r, st_transform(ca, st_crs(tmean_dif_r)))

writeRaster(tmean_dif_r, here("data", "env", "tmean_dif.tif"), overwrite = TRUE)

tmean_mod_df <- 
  left_join(tmean_his, tmean_cur) %>% 
  mutate(tmean_dif = tmean_cur - tmean_his) %>%
  right_join(coords) %>% 
  dplyr::select(SampleID, tmean_dif, tmean_cur)
```

# Models
```{r}
mod_vars <- c("NDVI", "gHM", "bio1", "Q", "csi_past", "glacier", "tmean_dif", "cur_lgm", "lgm_lig", "x", "y")

mod_df <- 
  get_coords() %>% 
  left_join(get_het()) %>%
  left_join(env_mod_df) %>%
  left_join(paleoclim_mod_df) %>%
  left_join(q_mod_df) %>%
  left_join(csi_mod_df) %>%
  left_join(glacier_mod_df) %>%
  left_join(tmean_mod_df) %>%
  dplyr::select(Ho, SampleID, geometry, all_of(mod_vars)) %>%
  drop_na(Ho) %>%
  # REMOVE CHANNEL ISLAND SAMPLE
  filter(SampleID != "Scelocci_CCGPMC_MW01-3-14") 

# Count NAs in each column
have_nas <- colSums(is.na(mod_df))[colSums(is.na(mod_df)) > 0]
have_nas

# 1 point along the coast has NA values, impute to the value of the closest sample
mod_sf <- mod_df %>% st_as_sf() %>% st_transform(3310)
close_impute <- function(var, mod_sf){
  map_dbl(1:nrow(mod_sf), \(i){ 
    if (!is.na(st_drop_geometry(mod_sf[i, var]))) return(mod_sf[i, var][[1]])
    # Get the closest sample that is not NA
    nearest <- st_nearest_feature(mod_sf[i,], drop_na(mod_sf[-i,]), 1)
    # Get the value of the closest sample
    return(drop_na(mod_sf)[nearest, var][[1]])
  }) 
}

vars <- names(have_nas)
names(vars) <- vars
mod_sf_imputed <- map(vars, ~close_impute(.x, mod_sf), .progress = TRUE) 
mod_df_imputed <- bind_cols(dplyr::select(mod_df, -all_of(vars)), bind_cols(mod_sf_imputed))

# Confirm no NAs:
colSums(is.na(mod_df_imputed))[colSums(is.na(mod_df_imputed)) > 0]

# Scale the variables (except glacier)
mod_df_imputed_scaled <- 
  mod_df_imputed %>% 
  mutate(across(c(all_of(mod_vars), -glacier), scale))

# Convert to sf
mod_sf <- st_as_sf(mod_df) %>% st_transform(3310)

# Check correlations
# Compute the correlation matrix
r <- cor(mod_df_imputed_scaled[,mod_vars], use = "pairwise.complete.obs")
# Set the diagonal and lower triangle to NA
r[lower.tri(r, diag = TRUE)] <- NA
# Find the pairs with correlations greater than 0.6
high_corr <- which(abs(r) > 0.6, arr.ind = TRUE)
# Print the pairs
high_corr_pairs <- data.frame(
  Var1 = rownames(r)[high_corr[, 1]],
  Var2 = rownames(r)[high_corr[, 2]],
  Correlation = r[high_corr]
)
high_corr_pairs
# cur_lgm is correlated with bio1, csi_past, lgm_lig; csi_past is correlated with lgm_lig
```

```{r}
export_df <- 
  st_drop_geometry(mod_df_imputed) %>% 
  select(SampleID, x, y, everything(), -geometry)
write_csv(export_df, here(outpath, "model_df.csv"))
```

```{r}
library(nlme)
library(ape)

coords <- get_coords(sf = TRUE)
dframe <- mod_df_imputed_scaled %>% st_as_sf()

# Construct IDW matrix
# Calculate pairwise geographic distances
geoMat <- st_distance(dframe)

# Convert to a numeric matrix while preserving the original matrix structure
geoMat <- matrix(as.numeric(units::drop_units(geoMat)), nrow = nrow(geoMat), ncol = ncol(geoMat))

# Apply inverse distance weighting
spdata.IDW <- 1 / geoMat
diag(spdata.IDW) <- 0  # Set diagonal to 0 to avoid division by zero in IDW

# Perform a global test of Moran's I on the point observations
# Ho is spatially autocorrelated
Moran.I(as.numeric(dframe$Ho), spdata.IDW)

# Add a dummy variable for the random effect
dummy <- rep(1, nrow(dframe))
dframe <- cbind(dframe, dummy) # Add it to the data frame

# Fit a null model
mod_vars <- c("NDVI", "gHM", "bio1", "cur_lgm", "lgm_lig", "csi_past", "glacier", "tmean_dif", "Q")
f <- as.formula(paste("Ho ~", paste(mod_vars, collapse = " + ")))
null.model <- lme(fixed = f, data = dframe, random = ~ 1 | dummy, method = "ML")
summary(null.model)

# Residuals are spatially autocorrelated
null_resids <- residuals(null.model)
null_M <- Moran.I(null_resids, spdata.IDW)
null_M

# Fit spatial models
exp.sp <- update(null.model, correlation = corExp(1, form = ~ x + y), method = "ML")
writeLines("Results for exponential spatial correlation:\n")
summary(exp.sp)

#gau.sp <- update(null.model, correlation = corGaus(1, form = ~ x + y), method = "ML")
#writeLines("\nResults for Gaussian spatial correlation:\n")
#summary(gau.sp)

sph.sp <- update(null.model, correlation = corSpher(1, form = ~ x + y), method = "ML")
writeLines("\nResults for spherical spatial correlation:\n")
summary(sph.sp)

# Get the AIC for all models
#aic <- AIC(null.model, exp.sp, gau.sp, sph.sp)
aic <- AIC(null.model, exp.sp, sph.sp)
best <- rownames(aic)[which.min(aic[, "AIC"])]
best

# the best model is exp.sp
best_model <- get(best)

summary(best_model)
car::vif(best_model) # cur_lgm VIF = 9

# Residuals are spatially autocorrelated
resids <- residuals(best_model)
M <- Moran.I(resids, spdata.IDW)

# Residual autocorrelation is stronger than the null model, so using OLS instead
null_M$observed
M$observed

dframe$resids <- resids
dframe$null_resids <- null_resids

gg_df <- dframe %>% pivot_longer(cols = contains("resids"), names_to = "resid_type", values_to = "resids")

pdf(here(plotpath, "full_model_residuals.pdf"), width = 12, height = 6)
ggplot(gg_df) + 
  geom_sf(data = ca) + 
  geom_sf(aes(col = resids)) +
  facet_wrap(~resid_type) +
  scale_color_viridis_c(option = "magma") +
  theme_void()
dev.off()
```

```{r}
# Reminder of the correlations
high_corr_pairs 

# csi_past, cur_lgm, and lgm_lig are all correlated with each other
# Choosing to keep cur_lgm 
mod_vars <- c("NDVI", "gHM", "bio1", "cur_lgm", "glacier", "tmean_dif")
```

```{r}
f <- as.formula(paste("Ho ~", paste(mod_vars, collapse = " + ")))

# This uses parallelization, but can also be run without it (remove plan() calls)
plan(multisession, workers = 10)
dredge_result <- spatial_dredge(full_formula = f, data = dframe)

mod_vars <- c("NDVI", "gHM", "bio1", "cur_lgm", "lgm_lig", "glacier", "tmean_dif")
f <- as.formula(paste("Ho ~", paste(mod_vars, collapse = " + ")))
mod <- lm(f, data = dframe, na.action = na.fail)
dredge_result <- dredge(mod)
best_model_f <- MuMIn::get.models(dredge_result, subset = 1)[[1]]$call
summary(lm(best_model_f, data = dframe))
plan(sequential)

# Look at top 3 models
dredge_result %>% head(3)

# Get best model
best_model_f <- dredge_result %>% filter(AIC == min(AIC))
```

```{r}
# TEST OF SP 1
library(spdep)
library(spatialreg)
mod_sf <- st_as_sf(mod_df_imputed_scaled) %>% st_transform(3310)


mod_vars <- c("NDVI", "gHM", "bio1", "cur_lgm", "lgm_lig", "glacier", "tmean_dif")
f <- as.formula(paste("Ho ~", paste(mod_vars, collapse = " + ")))
null <- lm(f, data = mod_sf)
nb <- dnearneigh(mod_sf, d1 = 0, d2 = 100000)
listw <- nb2listw(nb, style = "W", zero.policy = TRUE)
moran.test(residuals(null), listw)$estimate

testd <- map(c(5000, 10000, 50000, 100000, 500000), ~{
  nb <- dnearneigh(mod_sf, d1 = 0, d2 = .x)
  listw <- nb2listw(nb, style = "W", zero.policy = TRUE)
  mod <- errorsarlm(f, data = mod_sf, listw = listw, na.action = "na.omit", zero.policy = TRUE)
  c(moran.test(residuals(mod), listw)$estimate, AIC = AIC(mod), d = .x)
}) %>% bind_rows()
testd %>% arrange(AIC)

nb <- dnearneigh(mod_sf, d1 = 0, d2 = 100000)
listw <- nb2listw(nb, style = "W", zero.policy = TRUE)

pdf(here(plotpath, "nb.pdf"))
edges <- nb2lines(nb, coords = st_coordinates(mod_sf))
plot(edges)
dev.off()

mod_vars <- c("NDVI", "gHM", "bio1", "cur_lgm", "lgm_lig", "csi_past", "glacier", "tmean_dif", "Q")
f <- as.formula(paste("Ho ~", paste(mod_vars, collapse = " + ")))
plan(multisession, workers = 10)
dredge_result <- sem_dredge(full_formula = f, data = mod_sf, listw = listw)
plan(sequential)

dredge_result %>% arrange(AIC) %>% head(3)
best_model_f <- dredge_result %>% filter(AIC == min(AIC)) %>% pull(model_formula)
mod <- errorsarlm(best_model_f, data = mod_sf, listw = listw, na.action = "na.omit", zero.policy = TRUE)
summary(mod)   
AIC(mod) 

null <- lm(best_model_f, data = mod_sf)
lm_tests <- lm.RStests(null, listw, test = "all")

# calc Moran's I - residuals are equally spatially autocorrelated bfore and after
null <- lm(best_model_f, data = mod_sf)
Moran.I(residuals(null), spdata.IDW)$observed
Moran.I(residuals(mod), spdata.IDW)$observed
cor(residuals(null), residuals(mod))

summary(mod)
AIC(mod)
```

```{r}
# Fit the best model
f <- formula(best_model_f$model_formula)
null.model.best <- lme(fixed = f, data = dframe, random = ~ 1 | dummy, method = "ML")
exp.sp.best <- update(null.model.best, correlation = corExp(1, form = ~ x + y), method = "ML")
summary(exp.sp.best)
best_model <- exp.sp.best
sig_vars <- 
  summary(exp.sp.best)$tTable %>% 
  as.data.frame() %>%
  rownames_to_column("var") %>%
  filter(`p-value` < 0.05 & var != "(Intercept)") %>% 
  pull(var)

# Model diagnostics
pdf(here(plotpath, "model_check.pdf"), width = 8, height = 6)
plot(check_model(best_model, check = c("linearity", "vif", "normality")))
# Compare to normal linear model
plot(check_model(lm(f, data = dframe), check = c("linearity", "vif", "normality")))
dev.off()
```

# Variance partioning analysis
```{r}
lm_variance_partition <- function(predictor_sets, response, scale_vars = TRUE) {
  # Validate inputs
  if (length(predictor_sets) < 2) {
    stop("At least two predictor sets are required")
  }
  
  # Convert everything to data frames
  predictor_sets <- map(predictor_sets, as.data.frame)
  
  # Validate dimensions
  n_obs <- length(response)
  walk2(
    predictor_sets,
    names(predictor_sets),
    ~if (nrow(.x) != n_obs) {
      stop(sprintf("Predictor set %s has different number of observations than response", .y))
    }
  )
  
  # Scale variables if requested
  if (scale_vars) {
    response <- scale(response) %>% as.vector()
    predictor_sets <- map(predictor_sets, ~scale(.x) %>% as.data.frame())
  }
  
  # Function to calculate R2 and adjusted R2
  get_r2 <- function(model) {
    sum <- summary(model)
    c(
      r2 = sum$r.squared,
      adj_r2 = sum$adj.r.squared
    )
  }
  
  # Individual models for each predictor set
  individual_models <- predictor_sets %>%
    map(~lm(response ~ ., data = .x))
  
  # Full model with all predictors
  all_predictors <- as.data.frame(predictor_sets)
  names(all_predictors) <- names(predictor_sets)
  full_model <- lm(response ~ ., data = all_predictors)
  
  # Get R2 values for individual models
  individual_r2 <- map(individual_models, get_r2) %>%
    bind_rows(.id = "predictor_set")
  
  # Function to get all possible combinations
  get_combinations <- function(n) {
    combn(names(predictor_sets), n, simplify = FALSE)
  }
  
  # Get all combinations of predictor sets
  all_combinations <- map(
    2:length(predictor_sets),
    get_combinations
  ) %>%
    flatten()
  
  # Calculate R2 for all combinations
  combination_results <- all_combinations %>%
    set_names(map_chr(., ~paste(.x, collapse = "+"))) %>%
    map(~{
      combined_predictors <- predictor_sets[.x] %>% as.data.frame()
      names(combined_predictors) <- .x
      model <- lm(response ~ ., data = combined_predictors)
      get_r2(model)
    }) %>%
    bind_rows(.id = "combination")
  
  # Calculate unique contributions (Type III SS)
  unique_contributions <- map2_dfr(
    names(predictor_sets),
    predictor_sets,
    ~{
      # Model without this predictor set
      other_predictors <- predictor_sets[names(predictor_sets) != .x] %>% as.data.frame()
      names(other_predictors) <- names(predictor_sets)[names(predictor_sets) != .x]
      reduced_model <- lm(response ~ ., data = other_predictors)
      
      # Calculate unique contribution
      unique_r2 <- get_r2(full_model)[2] - get_r2(reduced_model)[2]
      
      tibble(
        predictor_set = .x,
        unique_contribution = unique_r2
      )
    }
  )
  
  # Calculate shared variance
  total_r2 <- get_r2(full_model)[2]
  shared_variance <- total_r2 - sum(unique_contributions$unique_contribution)
  
  # Prepare summary statistics
  summary_stats <- tibble(
    total_explained = total_r2,
    shared = shared_variance,
    unexplained = 1 - total_r2
  )
  
  
  # Print summary
  cat("\nVariance Partitioning Summary:\n")
  cat("----------------------------\n")
  ind <- 
    individual_r2 %>% 
    mutate(across(where(is.numeric), ~scales::percent(.x, accuracy = 1)))
  uniq <- 
    unique_contributions %>% 
    mutate(unique_contribution = scales::percent(unique_contribution, accuracy = 1))
  summary_table <- left_join(ind, uniq, by = "predictor_set")
  print(summary_table)

  cat("\nSummary statistics:\n")
  print(summary_stats %>% 
        mutate(across(everything(), ~scales::percent(.x, accuracy = 1))))
  
  
  # Create results list
  results <- list(
    summary_table = summary_table,
    individual_models = individual_models,
    full_model = full_model,
    individual_r2 = individual_r2,
    unique_contributions = unique_contributions,
    combination_results = combination_results,
    summary_stats = summary_stats
  )
  return(results)
}

```
```{r}
mod_vars <- c("NDVI", "gHM", "bio1", "cur_lgm", "lgm_lig", "csi_past", "glacier", "tmean_dif", "Q")
predictor_sets <- mod_df_imputed_scaled %>% select(all_of(mod_vars)) %>% as.list()
response <- mod_df_imputed_scaled$Ho
varpart_result <- lm_variance_partition(predictor_sets, response, scale_vars =TRUE)
```

# Future climate change
```{r}
future_ho <- 
  # NOTE: CSI was inverted (1 - CSI) before
  df_future %>%
  left_join(mod_df) %>%
  # REMOVE CHANNEL ISLAND
  filter(SampleID != "Scelocci_CCGPMC_MW01-3-14") 

plt <- 
  ggplot(future_ho, aes(x = csi, y = Ho, col = csi)) +
  geom_point() +
  facet_wrap(~model_name, scales = "free_x", nrow = 1) +
  theme_classic() +
  labs(x = "Climate stability", y = "Ho") +
  geom_smooth(method = "lm", se = TRUE) +
  stat_cor(label.y = min(future_ho$Ho, na.rm =TRUE)) +
  scale_color_viridis_c(option = "mako") +
  theme(
    strip.background = element_blank(),
    strip.text.x = element_text(hjust = 0, size = 14),
    axis.title = element_text(size = 14))

pdf(here(plotpath, "future_plot.pdf"), width = 12, height = 3.5)
plt
dev.off()
```

# Plots and tables

```{r}
# Turn into a nice gt table
library(gt)

best_model_table <- 
  summary(best_model)$tTable %>%
  data.frame() %>%
  rownames_to_column(var = "term") %>%
  rename(estimate = Value, std.error = `Std.Error`) %>%
  mutate(term = map_chr(term, ~make_pretty_names(.x))) %>%
  filter(term != "(Intercept)") %>%
  gt() %>%
  tab_header(title = "Full model") %>%
  gtExtras::gt_hulk_col_numeric(estimate, trim = TRUE) %>%
  fmt_scientific(
    columns = c("estimate", "std.error", "t.value", "p.value"),
    decimals = 2
  ) %>%
  # bold significant rows
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = c("p.value"),
      rows = p.value < 0.05
    )
  )

# export to html
gt::gtsave(best_model_table, filename = here(plotpath, "best_model_table.html"))
```

```{r}
# Make pretty plots 
mod <- best_model

# Plot predicted vs observed
plt1 <- 
  map(mod_vars, ~{
    ggplot(mod_df_imputed) +
      geom_point(aes(x = predict(mod), y = Ho, col = .data[[.x]])) +
      #geom_abline(intercept = 0, slope = 1) +
      geom_smooth(aes(x = predict(mod), y = Ho), method = "lm") +
      theme_classic() +
      labs(x = "Predicted", y = "Observed") +
      scale_color_viridis_c(option = "mako", end = 0.1, begin = 0.9)
  })

# Plot values across california
plt2 <-
  map(mod_vars, ~{
    ggplot(mod_df_imputed) +
      geom_sf(data = ca) +
      geom_point(aes(x = x, y = y, col = .data[[.x]])) +
      theme_void() +
      scale_color_viridis_c(option = "mako", end = 0.1, begin = 0.9)
  })

# Plot Ho vs variable
plt3 <- map(mod_vars, ~{
  ggplot(mod_df_imputed, aes(x = .data[[.x]], y = Ho)) +
    geom_point(aes(col = .data[[.x]])) +
    geom_smooth(method = "lm") +
    theme_classic() +
    labs(x = gsub("\n", " ", .x), y = "Ho") +
    scale_color_viridis_c(option = "mako", end = 0.1, begin = 0.9) +
    theme(legend.position = "none") +
    stat_cor(label.y = min(gg_df[["Ho"]], na.rm = TRUE))
})

# Plot Ho
plt4 <-
  map(1:length(mod_vars), ~{
    ggplot(mod_df_imputed) +
      geom_sf(data = ca) +
      geom_point(aes(x = x, y = y, col = Ho)) +
      theme_void() +
      scale_color_viridis_c(option ="mako", end = 0.1, begin = 0.9, direction = -1)
})

pdf(here(plotpath, "ho.pdf"), width = 4.5, height = 4.5)
ggplot(mod_df_imputed) +
  geom_sf(data = ca) +
  geom_point(aes(x = x, y = y, col = Ho)) +
  theme_void() +
  scale_color_viridis_c(option ="mako") +
  theme(plot.margin = margin(t = 1, r = 1, b = 1, l = 1, unit = "mm"))
dev.off()

# use cowplot to combine plots
pdf(here(plotpath, "linear_model_plot.pdf"), width = 10, height = 25)
ordered_plts <- pmap(list(plt3, plt1, plt2), ggpubr::ggarrange, common.legend = TRUE, legend = "right", nrow = 1)
do.call(cowplot::plot_grid, c(ordered_plts, list(nrow = length(plt1))))
ordered_plts <- pmap(list(plt3, plt1, plt4), ggpubr::ggarrange, nrow = 1)
do.call(cowplot::plot_grid, c(ordered_plts, list(nrow = length(plt1))))
dev.off()
```

# Figures
```{r}
# Variance partitioning plot
gg_df <- 
  varpart_result$summary_table %>%
  select(-r2) %>%
  pivot_longer(cols = c("adj_r2", "unique_contribution"), names_to = "metric", values_to = "value") %>%
  mutate(metric = recode(metric, adj_r2 = "Total", unique_contribution = "Unique")) %>%
  mutate(value = as.numeric(gsub("%", "", value))/100) %>%
  mutate(predictor_set = make_pretty_names(predictor_set))

ylgnbl <- RColorBrewer::brewer.pal(9, "YlGnBu")
ylgnbl_lite <- (ylgnbl[-c(1, 7:9)])

varpart_plt <- 
  ggplot(gg_df, aes(x = metric, y = predictor_set, fill = value)) +
  geom_tile(color = "white", lwd = 1) +
  geom_text(aes(label = scales::percent(value, accuracy = 1)), color = "black") +
  scale_fill_gradientn(colors = ylgnbl_lite) +
  theme_minimal() +
  labs(x = "", y = "", fill = "Variance\nExplained") +
  scale_x_discrete(position = "top") +
  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12))

# Correlation plot
mod_vars <- c("NDVI", "gHM", "bio1", "cur_lgm", "lgm_lig", "csi_past", "glacier", "tmean_dif", "Q")
cormat <- cor(mod_df_imputed[, mod_vars])

cor_df <- 
  cormat %>%
  as.data.frame() %>%
  rownames_to_column(var = "var1") %>%
  pivot_longer(cols = -var1, names_to = "var2", values_to = "cor") %>%
  filter(!is.na(cor)) %>%
  mutate(var1_pretty = make_pretty_names(var1), var2_pretty = make_pretty_names(var2)) %>%
  mutate(cor = case_when(var1 == var2 ~ NA, TRUE ~ cor))

cor_plt <- 
  ggplot(cor_df, aes(x = var1_pretty, y = var2_pretty, fill = cor)) +
  geom_tile(color = "white", lwd = 1) +
  geom_text(aes(label = round(cor, 2)), color = "black") +
  scale_fill_gradient2(
    low = "cyan3",
    mid = "white",
    high = "orange",
    midpoint = 0,
    limits = c(-1, 1)
  ) +
  theme_minimal() +
  labs(x = "", y = "", fill = "Correlation") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 12),
        axis.text.y = element_text(size = 12))

# Plot variables correlated with Ho that are not in the final model
names(mod_vars) <- mod_vars
confounded_vars <- map(mod_vars, ~cor.test(mod_df_imputed$Ho, mod_df_imputed[[.x]])$p.value) < 0.05
confounded_vars <- names(confounded_vars[confounded_vars])
confounded_vars <- confounded_vars[!(confounded_vars %in% names(best_model$coefficients$fixed))]

missig_plt <-
  mod_df_imputed %>%
  pivot_longer(all_of(confounded_vars)) %>%
  mutate(name = make_pretty_names(name)) %>%
  ggplot(aes(x = value, y = Ho)) +
  geom_point(aes(col = lgm_lig)) +
  geom_smooth(method = "lm") +
  theme_classic() +
  labs(x = "", y = "Ho", col = "Paleoclimate change\n(LGM - LIG)") +
  scale_color_viridis_c(option = "mako") +
  facet_wrap(~name, scales = "free", strip.position = "bottom", nrow = 1) +
  theme(strip.placement = "outside", strip.background = element_blank()) +
  stat_cor(label.y = min(mod_df_imputed$Ho, na.rm = TRUE)) 

combo <- plot_grid(cor_plt, varpart_plt, nrow = 1, rel_widths = c(5, 3), labels = c("A", "B"))
pdf(here(plotpath, "confounding.pdf"), width = 18, height = 9)
plot_grid(combo, missig_plt, nrow = 2, rel_heights = c(0.6, 0.35), labels = c("", "C"))
dev.off()
```

```{r}
sig_vars <- c("Q", "lgm_lig")
plot_map <- function(var){
  ggplot(st_as_sf(mod_df_imputed)) +
    geom_sf(data = ca) +
    geom_sf(aes(fill = .data[[var]]), pch = 21, col = rgb(0,0,0,0), cex = 2) +
    theme_void() +
    scale_fill_viridis_c(option = "mako") +
    labs(fill = gsub(" ", "\n", make_pretty_names(var)))
}

mod_df_imputed$predicted <- best_model$fitted[, "fixed"]


mod_plt <-
  ggplot(mod_df_imputed, aes(x = predicted, y = Ho)) +
  geom_point(aes(col = Ho), alpha = 0.8) +
  geom_smooth(method = "lm", color = "black") +
  theme_classic() +
  labs(x = "Predicted", y = "Observed") +
  scale_color_viridis_c(option = "mako") +
  annotate("text", x = min(mod_df_imputed$predicted) * 1.1, y = min(mod_df_imputed$Ho), label = paste("Adj. R² =", round(summary(lm(f, data = mod_df_imputed))$adj.r.squared, 2))) +
  theme(legend.position = "none")


sp_resids <- function(f, data){
  # Run spatial model
  null.model <- lme(fixed = formula(f), data = data, random = ~ 1 | dummy, method = "ML")
  exp.model <- update(null.model, correlation = corExp(1, form = ~ x + y), method = "ML")

  # Get residuals
  resids <- residuals(exp.model)
}

resids_df <- data.frame(
  q_x = sp_resids("Q ~ lgm_lig", dframe),
  q_y = sp_resids("Ho ~ lgm_lig", dframe),
  lgm_lig_x = sp_resids("lgm_lig ~ Q", dframe),
  lgm_lig_y = sp_resids("Ho ~ Q", dframe)
)


map_plts <- map(c("Ho", sig_vars), plot_map)
lm_plts <- map(sig_vars, ~gglm(.x, "Ho", mod_df_imputed))
partial_plts <- map(sig_vars, ~ggpartial(.x, "Ho", sig_vars, mod_df_imputed))
#partial_plts <- list(gglm("q_x", "q_y", resids_df),gglm("lgm_lig_x", "lgm_lig_y", resids_df))

plts <- c(map_plts, list(mod_plt), lm_plts, list(blank = ggplot() + theme_void()), partial_plts)
labels <- LETTERS[1:(length(plts) - 1)]
labels[which(names(plts) == "blank")] <- ""

pdf(here(plotpath, "main_figure.pdf"), width = 12.5, height = 11) 
do.call(plot_grid, c(plts, nrow = 3))
dev.off()

summary(lm(f, data = mod_df_imputed_scaled))
```

# Test of model without correlated variables

recent climate change (tmean_dif) seems uncorrelated with paleo_change so I wanted to build a model with all variables not correlated with paleo_change to see what pops out as significant. In the lm() tmean_dif is significant but in the spatial model it is not.
```{r}
uncor_vars <-
  cor_df %>% 
  filter(var2 == "paleo_change_lgm_lig" | var1 == "paleo_change_lgm_lig") %>% 
  filter(abs(cor) < 0.6) %>%
  select(var1, var2) %>%
  pivot_longer(cols = c(var1, var2)) %>% 
  distinct(value)

f <- as.formula(paste("Ho ~", paste(uncor_vars$value, collapse = " + ")))
lmod <- summary(lm(f, data = mod_df_imputed_scaled))

data <- mod_df_imputed_scaled
data$dummy <- 1
coords <- c("x", "y")
random <- "~ 1 | dummy"
method = "ML"
spmod <-
  summary(
  lme(fixed = f, 
  data = data, 
  random = as.formula(random), 
  correlation = corExp(1, form = as.formula(paste("~", paste(coords, collapse = " + ")))),
  method = method))

spmod2 <-
  summary(
  lme(fixed = formula("Ho ~ Q + paleo_change + tmean_dif"), 
  data = data, 
  random = as.formula(random), 
  correlation = corExp(1, form = as.formula(paste("~", paste(coords, collapse = " + ")))),
  method = method))

# paleo_change_lgm_lig and Q are significant
spmod
spmod2

# paleo_change_lgm_lig, Q, and tmean_dif are significant
lmod


mod <- lm("Ho ~ lgm_lig + tmean_dif + bio1", data = mod_df_imputed_scaled) 
```

So once you account for spatial correlation structure, the signal from recent climate change disapears.