
```{r}
library("rpaleoclim")
library("here")
library("tidyverse")
library("ggpubr")
library("terra")
library("sf")
library("gridExtra")
library("rpaleoclim")
source("general_functions.R")
source(here("analysis", "genetic_diversity", "genetic_diversity.R"))
source(here("analysis", "sdm", "paleovars.R"))
ca <- get_ca()
pdfpath <- here("analysis", "sdm", "pdfs")

coords <- get_coords(sf = TRUE)
het <- get_het()
dem <- get_dem(r= TRUE)
```

# Hypothesis 1: Genetic diversity explained by paleoclimate

```{r}
# Get paleoclimate data
# Created using paleoclimate.Rmd
paleovars <- rast(here("data", "env", "paleoclim", "paleoclim.tif"))

# Get important var for change in climate (based on paleoclimate.Rmd)
bio1_dif_lig <- paleovars$cur_bio_1 - paleovars$lig_bio_1
paleoclim_mod_df <- 
  data.frame(
    SampleID = coords$SampleID,
    paleo_change = terra::extract(bio1_dif_lig, coords, ID = FALSE)[,1]
    )

# Plots of climate change
gg_df <-
  as.data.frame(paleovars[[c("cur_bio_1", "lh_bio_1", "lgm_bio_1", "lig_bio_1")]]/10, xy = TRUE) %>%
  mutate(dif_lh = cur_bio_1 - lh_bio_1,
         dif_lgm = cur_bio_1 - lgm_bio_1,
         dif_lig = cur_bio_1 - lig_bio_1) 

gg_df1 <- 
  gg_df %>% 
  select(-contains("dif"), x, y) %>% 
  pivot_longer(c(-x, -y)) %>% 
  mutate(name = factor(name, levels = unique(name))) %>%
  mutate(name = case_when(
    name == "cur_bio_1" ~ "Current",
    name == "lh_bio_1" ~ "Late Holocene",
    name == "lgm_bio_1" ~ "LGM",
    name == "lig_bio_1" ~ "LIG"
  ))
  
gg_df2 <- 
  gg_df %>% 
  select(contains("dif"), x, y) %>% 
  pivot_longer(c(-x, -y)) %>% 
  mutate(name = factor(name, levels = unique(name)))%>%
  mutate(name = case_when(
    name == "dif_lh" ~ "Current - Late Holocene",
    name == "dif_lgm" ~ "Current - LGM",
    name == "dif_lig" ~ "Current - LIG"
  ))

plt1 <-
  ggplot(gg_df1) +
  geom_sf(data = ca) +
  geom_raster(aes(x = x, y = y, fill = value)) +
  scale_fill_viridis_c(option = "magma", na.value = "gray") +
  facet_wrap(~name, nrow = 1) +
  theme_void() +
  labs(fill = "Mean\nTemperature")

plt2 <-
  ggplot(gg_df2) +
  geom_sf(data = ca) +
  geom_raster(aes(x = x, y = y, fill = value)) +
  scale_fill_viridis_c(option = "magma", na.value = "gray") +
  facet_wrap(~name, nrow = 1) +
  theme_void() +
  labs(fill = "Mean\nTemperature\nChange")

gg_df3 <- gg_df2 %>% group_by(name) %>% mutate(value = scale(value))
plt3 <-
  ggplot(gg_df3) +
  geom_sf(data = ca) +
  geom_raster(aes(x = x, y = y, fill = value))+
  scale_fill_viridis_c(option = "magma", na.value = "gray") +
  facet_wrap(~name, nrow = 1) +
  theme_void() +
  labs(fill = "Mean\nTemperature\nChange\n(scaled)")

cowplot::plot_grid(plt1, plt2, plt3, nrow = 3)
```


# Hypothesis 2: Genetic diversity explained by past climate stability

### Climate stability (CSI)

Original paper: 
Herrando-Moraira, S., Nualart, N., Galbany-Casals, M. et al. Climate Stability Index maps, a global high resolution cartography of climate stability from Pliocene to 2100. Sci Data 9, 48 (2022). https://doi.org/10.1038/s41597-022-01144-5

Note: Bioclimatic variables do not contain Bio 2 or Bio 5, but they do contain other variables correlated with those variables

```{r, fig.width = 10, fig.height = 2.5}
csi_past <- rast(here("data", "env", "stability", "Layers", "past", "csi_past.tif"))
csi_past <- crop(csi_past, ca)
csi_past <- mask(csi_past, ca)

files <- list.files(here("data", "env", "stability", "Layers", "future"), full.names = TRUE, pattern = "csi*")
csi_future <- rast(map(files, rast))
csi_future <- crop(csi_future, ca)
csi_future <- mask(csi_future, ca)

csi_stack <- c(csi_past, csi_future)
names(csi_stack) <- c("Past",  "SSP 1-2.6", "SSP 2-4.5", "SSP 3-7.0", "SSP 5-8.5")
writeRaster(csi_stack, here("data", "env", "stability", "csi.tif"), overwrite = TRUE)
csi_df <- 
  terra::as.data.frame(csi_stack, xy = TRUE) %>% 
  pivot_longer(-c(x, y), 
  names_to = "model", 
  values_to = "CSI")

ggplot(csi_df) +
  geom_sf(data = ca) +
  geom_raster(aes(x = x, y = y, fill = CSI)) +
  scale_fill_viridis_c(option = "magma", na.value = NA) + 
  facet_wrap(~model, nrow = 1) +
  theme_void() 

ggplot(csi_df) +
  geom_sf(data = ca) +
  geom_raster(aes(x = x, y = y, fill = CSI)) +
  scale_fill_stepsn(colours = viridis::magma(100), n.breaks = 20) +
  facet_wrap(~model, nrow = 1) +
  theme_void() 

ggplot(filter(csi_df, model == "SSP 3-7.0")) +
  geom_sf(data = ca) +
  geom_raster(aes(x = x, y = y, fill = CSI)) +
  labs(fill = "Future climate\nstability index\n(CSI)") +
  scale_fill_viridis_c(option = "magma", na.value = "gray", limits = c(NA, 0.2)) +
  facet_wrap(~model, nrow = 2) +
  theme_void() 
```

```{r, fig.height = 4, fig.width = 4.5}
csi_vals <- extract(csi_past, coords)
df_past <- data.frame(csi_past = csi_vals[,1], coords)
```

```{r, fig.height = 4, fig.width = 16}
csi_vals <- extract(csi_future, coords, ID = FALSE)
df_future <- 
  data.frame(1 - csi_vals, coords) %>% 
  pivot_longer(starts_with("csi"), names_to = "future_model", values_to = "csi") %>%
  mutate(model_name = case_when(
    future_model == "csi_future_ssp126" ~ "SSP 1-2.6",
    future_model == "csi_future_ssp245" ~ "SSP 2-4.5",
    future_model == "csi_future_ssp370" ~ "SSP 3-7.0",
    future_model == "csi_future_ssp585" ~ "SSP 5-8.5"
  ))
```

```{r}
coords <- get_coords(sf = TRUE)
# Do 1 - to make the values more intuitive
csi_mod_df <- data.frame(csi = 1 - extract(csi_past, coords, ID = FALSE), coords)
```

# Hypothesis 3: Genetic diversity explained by admixture
```{r}
q <- read_csv(here("analysis", "admixture", "outputs", "Q6.csv"))
q_mod_df <- 
  q %>%
  dplyr::select(SampleID, Q) %>%
  # Do 1- to make the values more intuitive
  mutate(Q = 1 - Q)
```

# Hypothesis 4: Genetic diversity explained by current environment 
```{r}
env_data <- rast(here("data", "env", "california_env_stack.tif"))
env_mod_df <- data.frame(terra::extract(env_data, coords, ID = FALSE))
env_mod_df$SampleID <- coords$SampleID
env_mod_df$BurnDate <- as.numeric(env_mod_df$BurnDate > 0)
```

```{r}
# Test buffering for gHM
# Get buffered human modification layer values
gHM_buffers <- map(c(0, 1000, 10000, 100000), .progress = TRUE, ~{
  coords_buffer <- st_buffer(coords, .x)
  gHM_buffer <- extract(env_data[["gHM"]], coords_buffer, fun = mean, na.rm = TRUE)
  return(data.frame(gHM = gHM_buffer$gHM, SampleID = coords$SampleID, buffer = .x))
})

buffer_df <- 
  bind_rows(gHM_buffers) %>%
  left_join(get_het()) %>%
  left_join(get_coords())

cor_df <-
  buffer_df %>%
  group_by(buffer) %>%
  summarise(cor = cor(gHM, Ho, use = "pairwise.complete.obs"))

ggplot(cor_df, aes(x = buffer, y = cor)) +
  geom_line() +
  theme_classic() +
  labs(x = "Buffer distance (m)", y = "Correlation between gHM and Ho") 

polymod <- 
  buffer_df %>%
  group_by(buffer) %>%
  group_split() %>%
  map(~{
    mod_df <- drop_na(.x)
    # Fit a quadratic model (parabola)
    model <- lm(Ho ~ poly(gHM, 2), data = mod_df)
    # Visualize the fitted model
    mod_df$fitted <- predict(model, mod_df)
    # Get model stats
    mod_df$AIC_poly <- AIC(model)
    mod_df$AIC <- AIC(lm(Ho ~ gHM, data = mod_df))
    return(mod_df)
  }) %>%
  bind_rows()

ggplot(polymod) +
  geom_line(aes(x = buffer, y = AIC, col = "Regular")) +
  geom_line(aes(x = buffer, y = AIC_poly, col = "Polynomial")) +
  theme_classic() +
  labs(x = "Buffer distance (m)", y = "AIC")

best_mod_gHM <- polymod %>% filter(AIC_poly == min(AIC_poly))
ggplot(best_mod_gHM) + 
  geom_point(aes(x = gHM, y = Ho)) +
  geom_line(aes(x = gHM, y = fitted), color = "blue") +
  geom_smooth(aes(x = gHM, y = Ho), method = "lm", se = FALSE) +
  ggtitle("Quadratic Fit to Sample Data") +
  facet_wrap(~buffer, scales = "free_x") +
  theme_classic() 
unique(best_mod_gHM$AIC)
unique(best_mod_gHM$AIC_poly)

gg_df <- left_join(best_mod_gHM, get_coords(sf = TRUE))
ggplot(gg_df) +
  geom_sf(data = ca) +
  geom_point(aes(x = x, y = y, col = gHM)) +
  theme_void() +
  scale_color_viridis_c(option = "plasma", na.value = "gray") +
  labs(col = "Human modification") +
  facet_wrap(~buffer)
```

```{r}
env_mod_df <- 
  env_mod_df %>%
  select(-gHM) %>% 
  left_join(select(best_mod_gHM, SampleID, buffer, gHM))
```

# Hypothesis 5: Genetic diversity explained by glaciation
```{r}
lgm_glaciers <- st_read(here("data", "env", "Glaciotect_NA_lgm", "shp", "LGM_maxextent.shp"))
# Note: have to first intersect and then transform otherwise you get a node error
lgm_glaciers <- st_intersection(lgm_glaciers, st_transform(ca, crs = st_crs(lgm_glaciers)))
lgm_glaciers <- st_transform(lgm_glaciers, crs = st_crs(ca))
# Add a new column to coords indicating whether each point is inside the lgm shapefile
coords <- get_coords(sf = TRUE)
glacier_mod_df <- 
  coords %>%
  mutate(glacier = as.numeric((lengths(st_within(coords, lgm_glaciers)) > 0)))
```

# Hypothesis 6: Genetic diversity explained by recent climate change
Download PRISM mean temperature data from historical data: https://prism.oregonstate.edu/historical/

Data can be downloaded for all years using:
wget -r ftp://prism.oregonstate.edu/monthly/tmean/

mv prism.oregonstate.edu/monthly/tmean/* .

for year in {1895..2024}
do
  unzip '/media/wanglab/798f0e01-89d1-4d0f-8ed8-ef323be70ab91/Anusha/GitHub/ccgpscelop/data/env/prism/TEMP/'$year'/*.zip' -d '/media/wanglab/798f0e01-89d1-4d0f-8ed8-ef323be70ab91/Anusha/GitHub/ccgpscelop/data/env/prism/TEMP/'$year
done

```{r}
coords <- get_coords(sf = TRUE)
year = 1895

# Get prism data
get_prism <- function(year  = 1895, output = "df") {
  # Create PRISM path
  prism_path <- here("data", "env", "prism", "PRISM_tmean_stable_4kmM3", year)

  # List annual file
  file <- list.files(prism_path, pattern = paste0(year, "_*.bil.bil$"), full.names = TRUE)

  # Load data
  raster <- terra::rast(file)

  if (output == "raster"){
    return(raster)
  } else {
    vals <- terra::extract(raster, st_transform(coords, st_crs(raster)), ID = FALSE)
    names(vals) <- "tmean"
    df <- data.frame(year = year, tmean = vals, SampleID = coords$SampleID)
    return(df)
  }
}

# Loop over the years of interest
tmean_his <- map(1895:1905, get_prism, .progress =TRUE) %>% bind_rows()  
tmean_his <- tmean_his %>% group_by(SampleID) %>% summarise(tmean_his = mean(tmean, na.rm = TRUE))
tmean_cur <- map(2010:2020, ~get_prism(year = .x), .progress =TRUE) %>% bind_rows()  
tmean_cur <- tmean_cur %>% group_by(SampleID) %>% summarise(tmean_cur = mean(tmean, na.rm = TRUE))

tmean_mod_df <- 
  left_join(tmean_his, tmean_cur) %>% 
  mutate(tmean_dif = tmean_cur - tmean_his) %>%
  right_join(coords) %>% 
  dplyr::select(SampleID, tmean_dif)

```

# Models
```{r}
mod_vars <- c("BurnDate", "NDVI", "gHM", "paleo_change", "CHELSA_bio1_1981.2010_V.2.1", "Q", "csi_past", "glacier", "tmean_dif")

mod_df <- 
  get_coords() %>% 
  left_join(get_het()) %>%
  left_join(env_mod_df) %>%
  left_join(paleoclim_mod_df) %>%
  left_join(q_mod_df) %>%
  left_join(csi_mod_df) %>%
  left_join(glacier_mod_df) %>%
  left_join(tmean_mod_df) %>%
  dplyr::select(Ho, SampleID, geometry, all_of(mod_vars)) %>%
  drop_na(Ho) %>%
  # REMOVE CHANNEL ISLAND
  filter(SampleID != "Scelocci_CCGPMC_MW01-3-14")
  
# Count NAs in each column
have_nas <- colSums(is.na(mod_df))[colSums(is.na(mod_df)) > 0]

# NDVI, gHM, resdist has some NA values (< 3), impute to the value of the closest sample
mod_sf <- mod_df %>% st_as_sf() %>% st_transform(3310)
close_impute <- function(var, mod_sf){
  map_dbl(1:nrow(mod_sf), \(i){ 
    if (!is.na(st_drop_geometry(mod_sf[i, var]))) return(mod_sf[i, var][[1]])
    # Get the closest sample that is not NA
    nearest <- st_nearest_feature(mod_sf[i,], drop_na(mod_sf[-i,]), 1)
    # Get the value of the closest sample
    return(drop_na(mod_sf)[nearest, var][[1]])
  }) 
}

vars <- names(have_nas)
names(vars) <- vars
mod_sf_imputed <- map(vars, ~close_impute(.x, mod_sf), .progress = TRUE) 
mod_df_imputed <- bind_cols(dplyr::select(mod_df, -all_of(vars)), bind_cols(mod_sf_imputed))

# Confirm no NAs:
colSums(is.na(mod_df_imputed))[colSums(is.na(mod_df_imputed)) > 0]

# Scale the variables (except glacier)
mod_df_imputed_scaled <- mod_df_imputed %>% mutate(across(c(all_of(mod_vars), -glacier), scale))

# Convert to sf
mod_sf <- st_as_sf(mod_df) %>% st_transform(3310)

# Check correlations
# Compute the correlation matrix
r <- cor(mod_df_imputed_scaled[,mod_vars], use = "pairwise.complete.obs")
# Set the diagonal and lower triangle to NA
r[lower.tri(r, diag = TRUE)] <- NA
# Find the pairs with correlations greater than 0.6
high_corr <- which(abs(r) > 0.6, arr.ind = TRUE)
# Print the pairs
high_corr_pairs <- data.frame(
  Var1 = rownames(r)[high_corr[, 1]],
  Var2 = rownames(r)[high_corr[, 2]],
  Correlation = r[high_corr]
)
high_corr_pairs
# paleo_change is correlated with CHELSA_bio1_1981.2010_V.2.1, gHM, and csi_past
```

```{r}
make_pretty_names <- function(x){
  if (x == "BurnDate") return("Burned area (yes/no)")
  if (x == "paleo_change") return("Paleoclimate temperature change")
  if (grepl("CHELSA", x)) return("Contemporary temperature")
  if (grepl("csi", x)) return("Past climate stability")
  if (grepl("gHM", x)) return("Human modification")
  if (grepl("glacier", x)) return("Glacier (inside/outside)")
  if (grepl("Q", x)) return("Admixture")
  if (grepl("tmean_dif", x)) return("Contemporary temperature change")
  return(x)
}

# Plot all variables
gg_df <- 
  mod_df %>%
  pivot_longer(all_of(mod_vars)) %>%
  mutate(name =  map_chr(name, ~make_pretty_names(.x))) 

ggplot(gg_df, aes(x = value, y = Ho)) +
  geom_point() +
  facet_wrap(~name, scales = "free_x") +
  theme_classic() +
  labs(x = "Value", y = "Ho") +
  geom_smooth(method = "lm", se = TRUE) +
  stat_regline_equation(
        aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
        formula = y ~ x
      ) +
  stat_cor(label.y = 0) +
  theme(strip.background = element_blank(), strip.text.x = element_text(hjust = 0, size = 12))

# Make mixed corplot
vars_df <- mod_df[,mod_vars]
names(vars_df) <- map_chr(names(vars_df), ~make_pretty_names(.x))
r <- cor(vars_df, use = "pairwise.complete.obs")
corrplot::corrplot.mixed(
  r,
  lower = "number",
  upper = "ellipse",
  #tl.pos = c("d", "lt", "n"),
  tl.pos = "lt",
  #diag = c("n", "l", "u"),
  diag = "n"
  )

# Get variables with a significant correlation with Ho
named_mod_vars <- mod_vars
names(named_mod_vars) <- named_mod_vars
Ho_r <- map_dbl(named_mod_vars, ~cor.test(mod_df$Ho, mod_df[[.x]])$p.value) < 0.05

```

```{r}
# Plot all variables
gg_df <- 
  get_coords(sf = TRUE) %>%
  left_join(mod_df)

plt <- 
  map(mod_vars, ~{
    ggplot(gg_df) +
      geom_sf(data = ca) +
      geom_sf(aes(col = .data[[.x]])) +
      labs(col = make_pretty_names(.x)) +
      scale_color_viridis_c(option = "magma", begin = 0.05, end = 0.95) +
      theme_void() 
  })


cowplot::plot_grid(plotlist = plt)
```
# Model selection
```{r}
library(MuMIn)

# Included both regular and poly gHM to see which was the better fit using dredge
mod <- lm(Ho ~ BurnDate + NDVI + paleo_change + poly(gHM, 2) + gHM + CHELSA_bio1_1981.2010_V.2.1 + csi_past + factor(glacier) + Q + tmean_dif, data = mod_df_imputed_scaled, na.action = na.fail)

summary(mod)

# Use dredge for model selection
dredge_result <- dredge(mod)

# Select the models with delta AIC < 2
best_models <- get.models(dredge_result, subset = delta < 2)

# Get a model average
model_avg <- model.avg(best_models)
model_avg_summary <- summary(model_avg)

# Get coefficient and p-values
coeffs <- 
  model_avg_summary[["coefmat.subset"]] %>% 
  data.frame(., predictor = rownames(.))

# Get confidence intervals
confints <-
  confint(model_avg) %>% 
  data.frame(., predictor = rownames(.))

# Get sum of weights (relative importance)
sws <- 
  sw(model_avg_summary) %>% 
  data.frame() %>%
  t() %>%
  data.frame() %>%
  pivot_longer(everything(), names_to = "predictor", values_to = "Sum of weights") %>%
  drop_na(`Sum of weights`)

# Combine everything
model_df <-
  left_join(coeffs, confints, by = "predictor")%>%
  left_join(sws, by = "predictor") %>%
  rename(p = `Pr...z..`) 

# For the model average, paleo_change, Q, and gHM are significant
sigvars <-
  model_df %>% 
  filter(predictor != "(Intercept)") %>%
  filter(p < 0.05)
print(sigvars)

# Look at the best models
# Get the best model
best_model <- get.models(dredge_result, subset = delta < 2)[[1]]
second_best_model <- get.models(dredge_result, subset = delta < 2)[[2]]
third_best_model <- get.models(dredge_result, subset = delta < 2)[[3]]
AIC(best_model)
AIC(second_best_model)
AIC(third_best_model)

# Get the p-values (* for significant)
summary(best_model) # ~ glacier + gHM* + paleo_change* + Q*
summary(second_best_model) # ~ BurnDate + glacier* + gHM* + paleo_change* + Q*
summary(third_best_model) # ~ glacier + gHM* + paleo_change* + Q* + tmean_dif*

# Adj RSQ
summary(best_model)$adj.r.squared

# ggplot predicted vs observed for the best_model
mod_df_imputed_scaled$predict <- predict(best_model)
ggplot(mod_df_imputed_scaled, aes(x = predict, y = Ho)) +
  geom_point() +
  #geom_abline(intercept = 0, slope = 1) +
  geom_smooth(method = "lm") +
  theme_classic() +
  labs(x = "Predicted Ho", y = "Observed Ho")

# Turn into a nice gt table
library(gt)

full_model_table <- 
  mod %>%
  broom::tidy() %>%
  mutate(term = map_chr(term, ~make_pretty_names(.x))) %>%
  filter(term != "(Intercept)") %>%
  gt() %>%
  tab_header(title = "Full model") %>%
  gtExtras::gt_hulk_col_numeric(estimate, trim = TRUE) %>%
  fmt_scientific(
    columns = c("estimate", "std.error", "statistic", "p.value"),
    decimals = 2
  ) %>%
  # bold significant rows
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = c("p.value"),
      rows = p.value < 0.05
    )
  )

best_model_table <- 
  best_model %>%
  broom::tidy() %>%
  mutate(term = map_chr(term, ~make_pretty_names(.x))) %>%
  filter(term != "(Intercept)") %>%
  gt() %>%
  tab_header(title = "Best model") %>%
  data_color(
    columns = estimate,
    method = "numeric",
    palette = "viridis",
    reverse = TRUE, domain = c(-0.00011,0)
  ) %>%
  fmt_scientific(
    columns = c("estimate", "std.error", "statistic", "p.value"),
    decimals = 2
  )
# export to html
gt::gtsave(best_model_table, filename = here("analysis", "genetic_diversity", "best_model_table.html"))

gt::gtsave(full_model_table, filename = here("analysis", "genetic_diversity", "full_model_table.html"))
```

```{r}
mod <- best_model
gg_df <- 
  left_join(mod_df, get_coords()) %>%
  rename(Admixture = Q, `Paleoclimate\ntemp change` = paleo_change, `Human\nmodification` = gHM)

# Plot predicted vs observed
qplt1 <- 
  ggplot(gg_df) +
  geom_point(aes(x = predict(mod), y = Ho, col = Admixture)) +
  geom_abline(intercept = 0, slope = 1) +
  geom_smooth(aes(x = predict(mod), y = Ho), method = "lm") +
  theme_classic() +
  labs(x = "Predicted", y = "Observed") +
  scale_color_viridis_c(option = "viridis", end = 0.1, begin = 0.9)

gplt1 <- 
  ggplot(gg_df) +
  geom_point(aes(x = predict(mod), y = Ho, col = `Human\nmodification`)) +
  geom_smooth(aes(x = predict(mod), y = Ho), method = "lm") +
  theme_classic() +
  labs(x = "Predicted", y = "Observed") +
  scale_color_viridis_c(option = "rocket", end = 0.1, begin = 0.9)

bio2plt1 <-
  ggplot(gg_df) +
  geom_point(aes(x = predict(mod), y = Ho, col = `Paleoclimate\ntemp change`)) +
  geom_smooth(aes(x = predict(mod), y = Ho), method = "lm") +
  geom_smooth(aes(x = predict(mod), y = Ho), method = "lm") +
  theme_classic() +
  labs(x = "Predicted", y = "Observed") +
  scale_color_viridis_c(option = "inferno", end = 0.1, begin = 0.9)

# Plot values across california
qplt2 <-
  ggplot(gg_df) +
  geom_sf(data = ca) +
  geom_point(aes(x = x, y = y, col = Admixture)) +
  theme_void() +
  scale_color_viridis_c(option = "viridis", end = 0.1, begin = 0.9)

gplt2 <-
  ggplot(gg_df) +
  geom_sf(data = ca) +
  geom_point(aes(x = x, y = y, col = `Human\nmodification`)) +
  theme_void() +
  scale_color_viridis_c(option = "rocket", end = 0.1, begin = 0.9)

bio2plt2 <-
  ggplot(gg_df) +
  geom_sf(data = ca) +
  geom_point(aes(x = x, y = y, col = `Paleoclimate\ntemp change`)) +
  theme_void() +
  scale_color_viridis_c(option = "inferno", end = 0.1, begin = 0.9)

# Plot Ho vs variable
qplt3 <-
  ggplot(gg_df, aes(x = Admixture, y = Ho, col = Admixture)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_classic() +
  scale_color_viridis_c(option = "viridis", end = 0.1, begin = 0.9)

gplt3 <- 
  ggplot(gg_df, aes(x = `Human\nmodification`, y = Ho, col = `Human\nmodification`)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_classic() +
  scale_color_viridis_c(option = "rocket", end = 0.1, begin = 0.9)

bio2plt3 <-
  ggplot(gg_df, aes(x =  `Paleoclimate\ntemp change`, y = Ho, col = `Paleoclimate\ntemp change`)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_classic() +
  scale_color_viridis_c(option = "inferno", end = 0.1, begin = 0.9)

# use cowplot to combine plots
pdf(here("analysis", "genetic_diversity", "linear_model_plot.pdf"), width = 13, height = 10)
cowplot::plot_grid(bio2plt2, bio2plt3, bio2plt1, qplt2, qplt3, qplt1, gplt2, gplt3, gplt1, ncol = 3)
dev.off() 
```

```{r}
# Plot all variables with significant associations with Ho & paleo_change, gHM, or Q
paleo_change_r <- map_dbl(named_mod_vars, ~cor.test(mod_df$paleo_change, mod_df[[.x]])$p.value) < 0.05
q_r <- map_dbl(named_mod_vars, ~cor.test(mod_df$Q, mod_df[[.x]])$p.value) < 0.05
gHM_r <- map_dbl(named_mod_vars, ~cor.test(mod_df$gHM, mod_df[[.x]])$p.value) < 0.05

cor_vars <- mod_vars[Ho_r & (paleo_change_r | q_r | gHM_r)]
gg_df2 <- 
  mod_df %>%
  pivot_longer(c(all_of(cor_vars), -paleo_change, -Q, -gHM)) %>%
  mutate(name =  map_chr(name, ~make_pretty_names(.x))) %>%
  filter(!(name %in% c("Glacier (inside/outside)")))

cor_df <- 
  data.frame(
    q_r = q_r,
    paleo_change_r = paleo_change_r,
    gHM_r = gHM_r,
    Ho_r = Ho_r
  ) %>%
  mutate(name = rownames(.)) %>%
  filter(Ho_r) %>%
  select(-Ho_r) %>%
  filter(!(name %in% c("paleo_change", "Q", "gHM"))) %>%
  rename("Admixture" = q_r, "Paleoclimate temperature change" = paleo_change_r, "Human modification" = gHM_r) %>%
  mutate(descrip = apply(., 1, function(x) paste0("Correlated with:\n", paste(names(x[x == TRUE]), collapse = "\n")))) %>%
  mutate(name = map_chr(name, ~make_pretty_names(.x))) %>%
  filter(!(name %in% c("Glacier (inside/outside)")))

ggplot(gg_df2, aes(x = value, y = Ho)) +
  geom_point() +
  facet_wrap(~name, scales = "free") +
  geom_text(data = cor_df, aes(y = Inf, x = Inf, label = descrip), hjust = 1, col = "red", vjust =1)+
  theme_classic() +
  labs(x = "Value", y = "Ho") +
  geom_smooth(method = "lm", se = TRUE) +
  stat_cor(label.y = 0) +
  theme(strip.background = element_blank(), strip.text.x = element_text(hjust = 0, size = 12))

```
# Evaluating Q
```{r}
# with Q
summary(lm(Ho ~ BurnDate + NDVI + bio2_dif_lig + paleo_change + elev + gHM + Q + glacier + resdist + tmean_dif, data = mod_df_imputed))
# paleo_change, Q, and glacier are all significant

# without Q
summary(lm(Ho ~ BurnDate + NDVI + bio2_dif_lig + paleo_change + elev + gHM + glacier + resdist + tmean_dif, data = mod_df_imputed))
# paleo_change, Q, and glacier + bio2_dif_lig and BurnDate are all significant

# Note how Q is predicted by BurnDateTRUE/bio2_dif_lig  which are the variables that drop out in the with Q model
summary(lm(Q ~ BurnDate + NDVI + gHM + CHELSA_bio1_1981.2010_V.2.1 + bio2_dif_lig + paleo_change + elev + csi_past + glacier + resdist + tmean_dif, data = mod_df))
# we may want to add Q models however to get models of predicted admixture

plot(mod_df$gHM, mod_df$Ho)
plot(mod_df$BurnDate, mod_df$Ho)
plot(mod_df$CHELSA_bio1_1981.2010_V.2.1, mod_df$Ho)

# Plot predicted vs observed values
mod <- lm(Ho ~ BurnDate + NDVI + gHM + CHELSA_bio1_1981.2010_V.2.1 + bio2_dif_lig + paleo_change + elev + csi_past + glacier + resdist + Q  + tmean_dif, data = mod_df_imputed)
mod_df$Ho_pred <- predict(mod)
ggplot(mod_df, aes(x = Ho, y = Ho_pred)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Observed Ho", y = "Predicted Ho")  

# Plot residuals on a map
mod_sf$residuals <- residuals(mod)
ggplot(mod_sf) +
  geom_sf(data = ca) +
  geom_sf(aes(col = residuals)) +
  scale_color_viridis_c(option = "magma") +
  theme_void()
```

# Spatial model
```{r}
mod_sr <- bind_cols(mod_df, st_coordinates(mod_sf))

library(spdep)
library(spatialreg)
# Build SER model
# Create neighborhood using coordinates
# 100 km distance
nbdist <- 100000
nb <- dnearneigh(mod_sf, d1 = 0, d2 = nbdist)
listw <- nb2listw(nb, style = "W")

# plot neighbors
plot(nb, mod_sr[,c("X", "Y")], main = "Neighborhoods")

# Guides: 
# https://www.emilyburchfield.org/courses/gsa/spatial_regression_lab#:~:text=To%20determine%20whether%20to%20estimate,are%20the%20same%2C%20go%20robust.
# https://pages.cms.hu-berlin.de/EOL/gcg_quantitative-methods/Lab15_SpatialRegression.html 

# Ordinary least squares model
ols_model <- lm(Ho ~ paleo_change + glacier + Q, data = mod_df)
# test for evidence of autocorrelation
lm_tests <- lm.RStests(ols_model, listw, test = "all")
lm_tests
# Significant: RSerr, RSlag, SARMA, adjRSlag
# Not significant: adjRSerr
# because all are significant I am testing both an error and a lag model

# lag model:
# CAN'T INCLUDE RESDIST IN THIS MODEL SINCE IT IS CORRELATED WITH GEODIST
slm_model <- lagsarlm(Ho ~ paleo_change + glacier + Q, data = mod_df, listw = listw, na.action = "na.omit")

# error model:
sem_model <- errorsarlm(Ho ~ paleo_change + glacier + Q, data = mod_df, listw = listw, na.action = "na.omit")

# Compare AIC
writeLines(paste("AIC OLS:", round(AIC(ols_model), 2)))
writeLines(paste("AIC SEM:", round(AIC(sem_model), 2)))
writeLines(paste("AIC SLM:", round(AIC(slm_model), 2)))
# SLM is the best model based on AIC
```

```{r, fig.width = 8, fig.height = 6}
# Plot residuals
resids <- 
  mod_sf %>% 
  mutate(OLS = residuals(ols_model), SEM = residuals(sem_model), SLM = residuals(slm_model)) %>% 
  pivot_longer(c(OLS, SEM, SLM), values_to = "spmod_residuals")

ggplot(resids) + 
  geom_sf(data = ca) +
  geom_sf(aes(col = spmod_residuals), pch = 16, cex = 2) + 
  facet_grid(~name, switch = "y") +
  scale_color_viridis_c(option = "plasma") +
  ggtitle("Comparison of model residuals") +
  theme_void() + 
  theme(plot.title = element_text(hjust = 0.5), strip.text.y = element_text(angle = 90))
# SEM model is the best model based on residuals
```

```{r}
# print model summary
writeLines("Final model summary:")
summary(slm_model)

coeff_df <- 
  summary(slm_model)$Coef %>% 
  data.frame() %>%
  filter(`Pr...z..` < 0.05) 

coeff_df
```

# HTML Maps
```{r}
coords <- 
  get_coords() %>% 
  left_join(get_het()) %>% 
  select(SampleID, x, y, Ho)

write_csv(coords, here("data", "coords_ho.csv"))

library(leaflet)
library(readr)
library(here)

# Read the data
coords <- read_csv(here("data", "coords_ho.csv"))

# Create a color palette for the Ho variable
pal <- colorNumeric(palette = "viridis", domain = coords$Ho)

# Create the map
leaflet(coords) %>%
  addTiles() %>%
  addCircleMarkers(
    ~x, ~y,
    color = ~pal(Ho),
    radius = 5,
    stroke = FALSE,
    fillOpacity = 1,
    popup = ~paste0("SampleID: ", SampleID, "<br>Ho: ", Ho)
  ) %>%
  addLegend(
    "bottomright",
    pal = pal,
    values = ~Ho,
    title = "Ho",
    opacity = 1
  )

#USGS.USImageryTopo
# Create the map with a topographic base map
map <- leaflet(coords) %>%
  addProviderTiles(leaflet::providers$USGS.USImageryTopo, options = providerTileOptions(noWrap = TRUE)) %>%
  #addProviderTiles(leaflet::providers$Esri.WorldTopoMap, options = providerTileOptions(noWrap = TRUE)) %>%
  addCircleMarkers(
    ~x, ~y,
    fillColor = ~pal(Ho),
    radius = 5,
    stroke = TRUE,
    color = "black",
    fillOpacity = 1,
    weight = 1,
    popup = ~paste0("SampleID: ", SampleID, "<br>Ho: ", Ho)
  ) %>%
  addLegend(
    "bottomright",
    pal = pal,
    values = ~Ho,
    title = "Ho",
    opacity = 1
  )

# Save the map as an HTML file
htmlwidgets::saveWidget(map, "map.html")

# Open the HTML file in your web browser
browseURL("map.html")
```

# Future climate change
```{r}
future_ho <- 
  df_future %>%
  left_join(get_het()) 

ggplot(future_ho, aes(x = csi, y = Ho)) +
  geom_point() +
  facet_wrap(~model_name, scales = "free_x") +
  theme_classic() +
  labs(x = "Value", y = "Ho") +
  geom_smooth(method = "lm", se = TRUE) +
  stat_regline_equation(
        aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
        formula = y ~ x
      ) +
  stat_cor(label.y = min(future_ho$Ho, na.rm =TRUE)) +
  theme(strip.background = element_blank(), strip.text.x = element_text(hjust = 0, size = 12))


```

# ROH models

## Models
```{r}
mod_df2 <- 
  mod_df_imputed_scaled %>% 
  left_join(get_roh()) %>%
  left_join(dplyr::select(q, SampleID, cluster), by = "SampleID") %>% 
  drop_na(cluster) %>%
  drop_na(froh0)

mod <- lm(froh0 ~ BurnDate + NDVI + paleo_change + gHM + CHELSA_bio1_1981.2010_V.2.1 + csi_past + factor(glacier) + Q + tmean_dif, data = mod_df2, na.action = na.fail)

summary(mod)

# Use dredge for model selection
dredge_result <- dredge(mod)

# Select the models with delta AIC < 2
best_models <- get.models(dredge_result, subset = delta < 2)

# Get a model average
model_avg <- model.avg(best_models)
model_avg_summary <- summary(model_avg)

# Get coefficient and p-values
coeffs <- 
  model_avg_summary[["coefmat.subset"]] %>% 
  data.frame(., predictor = rownames(.))

# Get confidence intervals
confints <-
  confint(model_avg) %>% 
  data.frame(., predictor = rownames(.))

# Get sum of weights (relative importance)
sws <- 
  sw(model_avg_summary) %>% 
  data.frame() %>%
  t() %>%
  data.frame() %>%
  pivot_longer(everything(), names_to = "predictor", values_to = "Sum of weights") %>%
  drop_na(`Sum of weights`)

# Combine everything
model_df <-
  left_join(coeffs, confints, by = "predictor")%>%
  left_join(sws, by = "predictor") %>%
  rename(p = `Pr...z..`) 

# For the model average, paleo_change, gHM, and glacier are significant
sigvars <-
  model_df %>% 
  filter(predictor != "(Intercept)") %>%
  filter(p < 0.05)
print(sigvars)

# Look at the best models
# Get the best model
best_model <- get.models(dredge_result, subset = delta < 2)[[1]]
second_best_model <- get.models(dredge_result, subset = delta < 2)[[2]]
third_best_model <- get.models(dredge_result, subset = delta < 2)[[3]]
AIC(best_model)
AIC(second_best_model)
AIC(third_best_model)

# Get the p-values (* for significant)
summary(best_model) # ~ glacier* + paleo_change* + gHM*
summary(second_best_model) # ~ glacier* + paleo_change* + gHM* + NDVI
summary(third_best_model) # ~ glacier* + paleo_change* + gHM* + BurnDate

# ggplot predicted vs observed for the best_model
mod_df_imputed_scaled$predict <- predict(best_model)
ggplot(mod_df_imputed_scaled, aes(x = predict, y = Ho)) +
  geom_point() +
  #geom_abline(intercept = 0, slope = 1) +
  geom_smooth(method = "lm") +
  theme_classic() +
  labs(x = "Predicted Ho", y = "Observed Ho")

# Turn into a nice gt table
library(gt)

best_model_df <- 
  best_model %>%
  broom::tidy() %>%
  mutate(name = map_chr(term, ~make_pretty_names(.x))) %>%
  filter(name != "(Intercept)")
  
best_model_table <- 
  best_model_df %>%
  dplyr::select(-term) %>%
  rename(term = name) %>%
  gt() %>%
  tab_header(title = "Best model") %>%
  data_color(
    columns = estimate,
    method = "numeric",
    palette = "viridis",
    reverse = TRUE, domain = c(-0.00011,0)
  ) %>%
  fmt_scientific(
    columns = c("estimate", "std.error", "statistic", "p.value"),
    decimals = 2
  )
# export to html
gt::gtsave(best_model_table, filename = here("analysis", "genetic_diversity", "roh_best_model_table.html"))
```

```{r}
gg_df <-
  mod_df %>%
  dplyr::select(any_of(best_model_df$term), glacier, froh0, cluster) 

ggplot(gg_df, aes(x = gHM, y = log(froh0 + 1))) +
  geom_point() +
  theme_classic() +
  labs(x = "Value", y = "Froh0") +
  geom_smooth(method = "lm", se = TRUE) +
  #facet_wrap(~cluster) +
  stat_regline_equation(
        aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
        formula = y ~ x
      ) +
  stat_cor(label.y = -0.05) +
  theme(strip.background = element_blank(), strip.text.x = element_text(hjust = 0, size = 12))


gg_df <-
  mod_df %>%
  dplyr::select(any_of(best_model_df$term), glacier, froh0) 

gg_df2 <- 
  gg_df %>% 
  pivot_longer(-froh0) %>%
  mutate(name =  map_chr(name, ~make_pretty_names(.x)))

ggplot(gg_df2, aes(x = value, y = log(froh0 + 1))) +
  geom_point() +
  theme_classic() +
  labs(x = "Value", y = "Froh0") +
  geom_smooth(method = "lm", se = TRUE) +
  facet_wrap(~name, scale = "free_x") +
  stat_regline_equation(
        aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
        formula = y ~ x
      ) +
  stat_cor(label.y = 0) +
  ylim(0, NA) +
  theme(strip.background = element_blank(), strip.text.x = element_text(hjust = 0, size = 12))
```
