
```{r}
library("rpaleoclim")
library("here")
library("tidyverse")
library("ggpubr")
library("terra")
library("sf")
library("gridExtra")
library("rpaleoclim")
source("general_functions.R")
source(here("analysis", "genetic_diversity", "genetic_diversity.R"))
source(here("analysis", "sdm", "paleovars.R"))
ca <- get_ca()
pdfpath <- here("analysis", "sdm", "pdfs")

coords <- get_coords(sf = TRUE)
het <- get_het()
dem <- get_dem(r= TRUE)
```

# Hypothesis 1: Genetic diversity explained by geography of range expansion
Get ecoregions:
```{r}
ecol3 <- st_read(here("data", "env", "ecoregions", "ca_eco_l3.shp")) %>% st_transform(crs = 3310)
ecol4 <- st_read(here("data", "env", "ecoregions", "ca_eco_l4.shp")) %>% st_transform(crs = 3310)
ecol30 <- st_read(here("data", "env", "ecoregions", "ACE_Ecoregions_BaileyDerived_2022.shp")) %>% st_transform(crs = 3310)

#ggplot() + geom_sf(data = ecol3, aes(fill = US_L3NAME)) 
#ggplot() + geom_sf(data = ecol4, aes(fill = US_L4CODE)) +theme(legend.position = "none")
#ggplot() + geom_sf(data = ecol30, aes(fill = ECOREGION_))
```
wget https://gis.data.cnra.ca.gov/api/download/v1/items/d5ab2f9c5f7c45e682fd7e9a6ba8ad59/shapefile?layers=0

wget https://gaftp.epa.gov/EPADataCommons/ORD/Ecoregions/ca/ca_eco_l3.zip

Ecoregions: https://www.epa.gov/eco-research/ecoregion-download-files-state-region-9#pane-04

Making resistance layer based on elevation + removing central valley
```{r}
# Higher elevation = higher resistance = lower connectivity
# rescale dem from 0 to 1
ca <- get_ca()

coords_sf <- get_coords(sf = TRUE)

coords_sf_buffer <- st_buffer(coords_sf, 50000)

dem <- elevatr::get_elev_raster(coords_sf_buffer, z = 5)

dem <- crop(dem, coords_sf_buffer)

dem[dem < 0] <- NA

# Plot the result
#plot(dem, zlim = c(0,4000))
#points(coords_sf)
dem <- rast(dem)
minr <- global(dem, "min", na.rm = TRUE)[,1]
maxr <- global(dem, "max", na.rm = TRUE)[,1]
dem01 <- (dem - minr) / (maxr - minr)

connectivity <- 1 - dem01

#plot(connectivity)
#points(coords_sf)
# INVERSE IS WEIRD FIGURE OUT LAYER

valley <- ecol3 %>% filter(US_L3NAME == "Central California Valley") %>% st_transform(st_crs(dem))
connectivity_mask <- mask(connectivity, valley, inverse = TRUE) 
connectivity_mask[is.na(connectivity_mask)] <- global(connectivity, "min", na.rm = TRUE)[,1]
connectivity_mask <- mask(connectivity_mask, dem)
connectivity_mask <- project(connectivity_mask, "EPSG:3310")
coords_sf <- st_transform(get_coords(sf =TRUE), "EPSG:3310")
#plot(connectivity_mask)
#points(coords_sf)

# Calculate resistance distance
# COME BACK AND COMPARE TO BIO2 AND EACH OTHER
lyr <- aggregate(connectivity_mask, 6, fun = mean, na.rm = TRUE)
#plot(lyr, col = viridis::inferno(100))
# we know the central valley was a barrier
#lyr <- aggregate(connectivity, 5, fun = mean, na.rm = TRUE)

# convert coords to matrix and rename
coords_mat <- 
  get_coords(sf = TRUE) %>% 
  st_transform(crs = st_crs(lyr)) %>% 
  st_coordinates() %>% 
  as.matrix()

# create transition surface
trSurface <- gdistance::transition(raster::raster(lyr), transitionFunction = mean, directions = 8)

# note: type = "c" is for least cost distances
trSurface <- gdistance::geoCorrection(trSurface, type = "c", scl = FALSE)

# plot transition surface
#plot(raster::raster(trSurface), col = viridis::inferno(100))

# NOTE: If you get an out of bounds error that means you have coordinates on the edge, increase the edge aroudn your raster (buffer it)
resdist <- as.matrix(gdistance::commuteDistance(trSurface, coords_mat))

# Get southern-most point
southern_point <- which(coords_mat[,2] == min(coords_mat[,2]))

# Get the resistance distance from the southern-most point to all other points
southern_resdist <- resdist[southern_point,]

# Get the geographic distance from the southern-most point to all other points
southern_dist <- as.vector(st_distance(coords_sf[southern_point,], coords_sf))


df <- 
  get_coords() %>%
  mutate(resdist = southern_resdist, geodist = southern_dist) %>% 
  # remove point on channel island
  filter(resdist != max(resdist, na.rm = TRUE)) %>%
  # remove central valley outliers
  filter(resdist != max(resdist, na.rm = TRUE)) %>%
  filter(resdist != max(resdist, na.rm = TRUE)) %>%
  left_join(get_het()) 

ggdf <- 
  df %>%
  pivot_longer(c(resdist, geodist), names_to = "distance_type", values_to = "distance") %>%
  mutate(distance_type = case_when(
    distance_type == "resdist" ~ "B. Resistance distance",
    distance_type == "geodist" ~ "A. Geographic distance"
  ))

ggplot(ggdf, aes(x = distance, y = Ho)) +
  geom_point() + 
  theme_classic() +
  labs(x = "Distance from southern-most point", y = "Ho") +
  facet_wrap(~distance_type, scales = "free_x") +
  # add correlation
  geom_smooth(method = "lm", se = TRUE) +
  stat_regline_equation(
        aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
        formula = y ~ x
      ) +
  stat_cor(label.y = 0) +
  theme(strip.background = element_blank(), strip.text.x = element_text(hjust = 0, size = 12))

# Resdist has a slighlty lower AIC
AIC(lm(Ho ~ resdist, data = df)) #-2252.156
AIC(lm(Ho ~ geodist, data = df)) #-2251.572
```

```{r}
resist_mod_df <- 
  get_coords(sf = TRUE) %>%
  left_join(df) %>%
  dplyr::select(SampleID, resdist, geodist)
```

# Hypothesis 2: Genetic diversity explained by paleoclimate

```{r}
# Get paleoclimate data
# Created using paleoclimate.Rmd
paleovars <- rast(here("data", "env", "paleoclim", "paleoclim.tif"))

# Get important vars for change in climate (based on paleoclimate.Rmd)
bio2_dif_lh <- paleovars$cur_bio_2 - paleovars$lh_bio_2
bio2_dif_lig <- paleovars$cur_bio_2 - paleovars$lig_bio_2
paleoclim_mod_df <- 
  data.frame(
    SampleID = coords$SampleID,
    bio2_dif_lh = terra::extract(bio2_dif_lh, coords, ID = FALSE)[,1],
    bio2_dif_lig = terra::extract(bio2_dif_lig, coords, ID = FALSE)[,1]
    )
```


# Hypothesis 3: Genetic diversity explained by past climate stability

### Climate stability (CSI)

Original paper: 
Herrando-Moraira, S., Nualart, N., Galbany-Casals, M. et al. Climate Stability Index maps, a global high resolution cartography of climate stability from Pliocene to 2100. Sci Data 9, 48 (2022). https://doi.org/10.1038/s41597-022-01144-5

Note: Bioclimatic variables do not contain Bio 2 or Bio 5, but they do contain other variables correlated with those variables

```{r, fig.width = 10, fig.height = 2.5}
csi_past <- rast(here("data", "env", "stability", "Layers", "past", "csi_past.tif"))
csi_past <- crop(csi_past, ca)
csi_past <- mask(csi_past, ca)

files <- list.files(here("data", "env", "stability", "Layers", "future"), full.names = TRUE, pattern = "csi*")
csi_future <- rast(map(files, rast))
csi_future <- crop(csi_future, ca)
csi_future <- mask(csi_future, ca)

csi_stack <- c(csi_past, csi_future)
names(csi_stack) <- c("Past",  "SSP 1-2.6", "SSP 2-4.5", "SSP 3-7.0", "SSP 5-8.5")
writeRaster(csi_stack, here("data", "env", "stability", "csi.tif"), overwrite = TRUE)
csi_df <- 
  terra::as.data.frame(csi_stack, xy = TRUE) %>% 
  pivot_longer(-c(x, y), 
  names_to = "model", 
  values_to = "CSI")

ggplot(csi_df) +
  geom_sf(data = ca) +
  geom_raster(aes(x = x, y = y, fill = CSI)) +
  scale_fill_viridis_c(option = "magma", na.value = NA) + 
  facet_wrap(~model, nrow = 1) +
  theme_void() 

ggplot(csi_df) +
  geom_sf(data = ca) +
  geom_raster(aes(x = x, y = y, fill = CSI)) +
  scale_fill_stepsn(colours = viridis::magma(100), n.breaks = 20) +
  facet_wrap(~model, nrow = 1) +
  theme_void() 

ggplot(filter(csi_df, model == "SSP 3-7.0")) +
  geom_sf(data = ca) +
  geom_raster(aes(x = x, y = y, fill = CSI)) +
  labs(fill = "Future climate\nstability index\n(CSI)") +
  scale_fill_viridis_c(option = "magma", na.value = "gray", limits = c(NA, 0.2)) +
  facet_wrap(~model, nrow = 2) +
  theme_void() 
```

```{r, fig.height = 4, fig.width = 4.5}
csi_vals <- extract(csi_past, coords)
df_past <- data.frame(csi_past = csi_vals[,1], coords)
```

```{r, fig.height = 4, fig.width = 16}
csi_vals <- extract(csi_future, coords, ID = FALSE)
df_future <- 
  data.frame(csi_vals, coords) %>% 
  pivot_longer(starts_with("csi"), names_to = "future_model", values_to = "csi") %>%
  mutate(model_name = case_when(
    future_model == "csi_future_ssp126" ~ "SSP 1-2.6",
    future_model == "csi_future_ssp245" ~ "SSP 2-4.5",
    future_model == "csi_future_ssp370" ~ "SSP 3-7.0",
    future_model == "csi_future_ssp585" ~ "SSP 5-8.5"
  ))
```

```{r}
coords <- get_coords(sf = TRUE)
csi_mod_df <- data.frame(csi = extract(csi_past, coords, ID = FALSE), coords)
```

# Hypothesis 4: Genetic diversity explained by admixture
```{r}
q <- read_csv(here("analysis", "admixture", "outputs", "Q6.csv"))
q_mod_df <- 
  q %>%
  dplyr::select(SampleID, Q)
```

# Hypothesis 5: Genetic diversity explained by current environment 
```{r}
env_data <- rast(here("data", "env", "california_env_stack.tif"))
env_mod_df <- data.frame(terra::extract(env_data, coords, ID = FALSE))
env_mod_df$SampleID <- coords$SampleID
env_mod_df$BurnDate <- as.numeric(env_mod_df$BurnDate > 0)
```

# Hypothesis 6: Genetic diversity explained by elevation
```{r}
dem_mod_df <- get_dem()
```


# Hypothesis 7: Genetic diversity explained by glaciation
```{r}
lgm_glaciers <- st_read(here("data", "env", "Glaciotect_NA_lgm", "shp", "LGM_maxextent.shp"))
# Note: have to first intersect and then transform otherwise you get a node error
lgm_glaciers <- st_intersection(lgm_glaciers, st_transform(ca, crs = st_crs(lgm_glaciers)))
lgm_glaciers <- st_transform(lgm_glaciers, crs = st_crs(ca))
# Add a new column to coords indicating whether each point is inside the lgm shapefile
coords <- get_coords(sf = TRUE)
glacier_mod_df <- 
  coords %>%
  mutate(glacier = as.numeric((lengths(st_within(coords, lgm_glaciers)) > 0)))
```

# Hypothesis 8: Genetic diversity explained by recent climate change
Download PRISM mean temperature data from historical data: https://prism.oregonstate.edu/historical/

Data can be downloaded for all years using:
wget -r ftp://prism.oregonstate.edu/monthly/tmean/

mv prism.oregonstate.edu/monthly/tmean/* .

for year in {1895..2024}
do
  unzip '/media/wanglab/798f0e01-89d1-4d0f-8ed8-ef323be70ab91/Anusha/GitHub/ccgpscelop/data/env/prism/TEMP/'$year'/*.zip' -d '/media/wanglab/798f0e01-89d1-4d0f-8ed8-ef323be70ab91/Anusha/GitHub/ccgpscelop/data/env/prism/TEMP/'$year
done

```{r}
coords <- get_coords(sf = TRUE)
year = 1895

# Get prism data
get_prism <- function(year  = 1895, output = "df") {
  # Create PRISM path
  prism_path <- here("data", "env", "prism", "PRISM_tmean_stable_4kmM3", year)

  # List annual file
  file <- list.files(prism_path, pattern = paste0(year, "_*.bil.bil$"), full.names = TRUE)

  # Load data
  raster <- terra::rast(file)

  if (output == "raster"){
    return(raster)
  } else {
    vals <- terra::extract(raster, st_transform(coords, st_crs(raster)), ID = FALSE)
    names(vals) <- "tmean"
    df <- data.frame(year = year, tmean = vals, SampleID = coords$SampleID)
    return(df)
  }
}

# Loop over the years of interest
tmean_his <- map(1895:1905, get_prism, .progress =TRUE) %>% bind_rows()  
tmean_his <- tmean_his %>% group_by(SampleID) %>% summarise(tmean_his = mean(tmean, na.rm = TRUE))
tmean_cur <- map(2010:2020, ~get_prism(year = .x), .progress =TRUE) %>% bind_rows()  
tmean_cur <- tmean_cur %>% group_by(SampleID) %>% summarise(tmean_cur = mean(tmean, na.rm = TRUE))

tmean_mod_df <- 
  left_join(tmean_his, tmean_cur) %>% 
  mutate(tmean_dif = tmean_cur - tmean_his) %>%
  right_join(coords) %>% 
  dplyr::select(SampleID, tmean_dif)

```

# Hypothesis 9: Transverse ranges [ NOT USED CURRENTLY ]
```{r}
# Get the transverse ranges
transverse <- ecol4 %>% filter(grepl("Transverse", US_L4NAME))
coords <- get_coords(sf = TRUE) %>% left_join(get_het())
transverse <- st_transform(ecol4, crs = st_crs(coords))
coords_transverse <- st_join(filter(coords, Ho > 0.0011), transverse)
transverse_coords <- transverse %>% filter(US_L4NAME %in% coords_transverse$US_L4NAME)
ggplot() + 
  geom_sf(data = ca) + 
  geom_sf(data = transverse_coords, aes(fill = US_L4NAME)) +
  geom_sf(data = coords_transverse, aes(col = Ho)) + 
  scale_color_viridis_c(option = "mako") +
  theme_void()

ggplot() + 
  geom_sf(data = ca) + 
  geom_sf(data = ecol30, aes(fill = ECOREGION_)) +
  geom_sf(data = coords, aes(col = Ho)) + 
  scale_color_viridis_c(option = "mako") +
  theme_void()


# Add a new column to coords indicating whether each point is inside the transverse ranges
coords_transverse <- st_join(coords, transverse)
transverse_mod_df <- 
  coords_transverse %>%
  st_drop_geometry() %>%
  select(SampleID, US_L4NAME) %>%
  mutate(transverse = as.numeric(!is.na(US_L4NAME)))
```

# Models
```{r}
mod_vars <- c("BurnDate", "NDVI", "gHM", "bio2_dif_lig", "bio2_dif_lh", "elev", "CHELSA_bio1_1981.2010_V.2.1", "Q", "csi_past", "glacier", "resdist", "tmean_dif")
mod_df <- 
  get_coords() %>% 
  left_join(get_het()) %>%
  left_join(get_roh()) %>%
  left_join(env_mod_df) %>%
  left_join(paleoclim_mod_df) %>%
  left_join(dem_mod_df) %>%
  left_join(q_mod_df) %>%
  left_join(csi_mod_df) %>%
  left_join(resist_mod_df) %>%
  left_join(glacier_mod_df) %>%
  left_join(tmean_mod_df) %>%
  dplyr::select(Ho, froh0, SampleID, geometry, all_of(mod_vars)) %>%
  drop_na(Ho)
  
# Count NAs in each column
colSums(is.na(mod_df))[colSums(is.na(mod_df)) > 0]

# Elevation, NDVI, gHM, resdist has some NA values, impute to the value of the closest sample
mod_sf <- mod_df %>% st_as_sf() %>% st_transform(3310)
close_impute <- function(var, mod_sf){
  map_dbl(1:nrow(mod_sf), \(i){ 
    if (!is.na(st_drop_geometry(mod_sf[i, var]))) return(mod_sf[i, var][[1]])
    # Get the closest sample that is not NA
    nearest <- st_nearest_feature(mod_sf[i,], drop_na(mod_sf[-i,]), 1)
    # Get the value of the closest sample
    return(drop_na(mod_sf)[nearest, var][[1]])
  }) 
}

vars <- c("elev", "NDVI", "gHM", "resdist", "tmean_dif")
names(vars) <- vars
mod_sf_imputed <- map(vars, ~close_impute(.x, mod_sf), .progress = TRUE) 
mod_df_imputed <- bind_cols(dplyr::select(mod_df, -all_of(vars)), bind_cols(mod_sf_imputed))

# Confirm no NAs:
colSums(is.na(mod_df_imputed))[colSums(is.na(mod_df_imputed)) > 0]

# Scale the variables (except glacier)
mod_df_imputed_scaled <- mod_df_imputed %>% mutate(across(c(all_of(mod_vars), -glacier), scale))

# Convert to sf
mod_sf <- st_as_sf(mod_df) %>% st_transform(3310)

# Check correlations
# Compute the correlation matrix
r <- cor(mod_df_imputed_scaled[,mod_vars], use = "pairwise.complete.obs")
# Set the diagonal and lower triangle to NA
r[lower.tri(r, diag = TRUE)] <- NA
# Find the pairs with correlations greater than 0.6
high_corr <- which(abs(r) > 0.6, arr.ind = TRUE)
# Print the pairs
high_corr_pairs <- data.frame(
  Var1 = rownames(r)[high_corr[, 1]],
  Var2 = rownames(r)[high_corr[, 2]],
  Correlation = r[high_corr]
)
high_corr_pairs
# Elevation is correlated with CHELSA_bio1_1981.2010_V.2.1 and csi_past, which makes sense
```

```{r}
make_pretty_names <- function(x){
  if (x == "BurnDate") return("Burned area (yes/no)")
  if (x == "bio2_dif_lh") return("LH change in diurnal range")
  if (x == "bio2_dif_lig") return("LIG change in diurnal range")
  if (grepl("CHELSA", x)) return("Contemporary temperature")
  if (grepl("csi", x)) return("Past climate stability")
  if (grepl("elev", x)) return("Elevation")
  if (grepl("gHM", x)) return("Human modification")
  if (grepl("glacier", x)) return("Glacier (inside/outside)")
  if (grepl("resdis", x)) return("Resistance dist from south") 
  if (grepl("Q", x)) return("Admixture")
  if (grepl("tmean_dif", x)) return("Contemporary temp change")
  return(x)
}

# Plot all variables
gg_df <- 
  mod_df %>%
  pivot_longer(all_of(mod_vars)) %>%
  mutate(name =  map_chr(name, ~make_pretty_names(.x))) 

ggplot(gg_df, aes(x = value, y = Ho)) +
  geom_point() +
  facet_wrap(~name, scales = "free_x") +
  theme_classic() +
  labs(x = "Value", y = "Ho") +
  geom_smooth(method = "lm", se = TRUE) +
  stat_regline_equation(
        aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
        formula = y ~ x
      ) +
  stat_cor(label.y = 0) +
  theme(strip.background = element_blank(), strip.text.x = element_text(hjust = 0, size = 12))

# Make mixed corplot
vars_df <- mod_df[,mod_vars]
names(vars_df) <- map_chr(names(vars_df), ~make_pretty_names(.x))
r <- cor(vars_df, use = "pairwise.complete.obs")
corrplot::corrplot.mixed(
  r,
  lower = "number",
  upper = "ellipse",
  #tl.pos = c("d", "lt", "n"),
  tl.pos = "lt",
  #diag = c("n", "l", "u"),
  diag = "n"
  )

# Get variables with a significant correlation with Ho
named_mod_vars <- mod_vars
names(named_mod_vars) <- named_mod_vars
Ho_r <- map_dbl(named_mod_vars, ~cor.test(mod_df$Ho, mod_df[[.x]])$p.value) < 0.05

# Plot all variables with significant associations with Ho & bio2_dif_lh
bio2_dif_lh_r <- map_dbl(named_mod_vars, ~cor.test(mod_df$bio2_dif_lh, mod_df[[.x]])$p.value) < 0.05
bio2_vars <- mod_vars[Ho_r & bio2_dif_lh_r]
gg_df2 <- 
  mod_df %>%
  pivot_longer(c(all_of(bio2_vars), -bio2_dif_lh)) %>%
  mutate(name =  map_chr(name, ~make_pretty_names(.x))) %>%
  filter(!(name %in% c("LIG change in diurnal range", "Glacier (inside/outside)")))

ggplot(gg_df2, aes(x = value, y = bio2_dif_lh)) +
  geom_point() +
  facet_wrap(~name, scales = "free_x") +
  theme_classic() +
  labs(x = "Value", y = "LH change in diurnal range") +
  geom_smooth(method = "lm", se = TRUE) +
  stat_regline_equation(
        aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
        formula = y ~ x
      ) +
  stat_cor(label.y = -8) +
  theme(strip.background = element_blank(), strip.text.x = element_text(hjust = 0, size = 12))

ggplot(gg_df2, aes(x = value, y = Ho)) +
  geom_point() +
  facet_wrap(~name, scales = "free_x") +
  theme_classic() +
  labs(x = "Value", y = "Ho") +
  geom_smooth(method = "lm", se = TRUE) +
  stat_regline_equation(
        aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
        formula = y ~ x
      ) +
  stat_cor(label.y = 0) +
  theme(strip.background = element_blank(), strip.text.x = element_text(hjust = 0, size = 12))

# Plot variables associated with Q and Ho
q_r <- map_dbl(named_mod_vars, ~cor.test(mod_df$Q, mod_df[[.x]])$p.value) < 0.05
q_vars <- mod_vars[Ho_r & q_r]
gg_df3 <- 
  mod_df %>%
  pivot_longer(c(all_of(q_vars), -Q)) %>%
  mutate(name =  map_chr(name, ~make_pretty_names(.x)))

ggplot(gg_df3, aes(x = value, y = Q)) +
  geom_point() +
  facet_wrap(~name, scales = "free_x") +
  theme_classic() +
  labs(x = "Value", y = "Ho") +
  geom_smooth(method = "lm", se = TRUE) +
  stat_regline_equation(
        aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
        formula = y ~ x
      ) +
  stat_cor(label.y = 0.4) +
  theme(strip.background = element_blank(), strip.text.x = element_text(hjust = 0, size = 12))

```

```{r}
# Plot all variables
gg_df <- 
  get_coords(sf = TRUE) %>%
  left_join(mod_df)

plt <- 
  map(mod_vars, ~{
    ggplot(gg_df) +
      geom_sf(data = ca) +
      geom_sf(aes(col = .data[[.x]])) +
      labs(col = make_pretty_names(.x)) +
      scale_color_viridis_c(option = "magma", begin = 0.05, end = 0.95) +
      theme_void() 
  })


plot(gg_df$tmean_dif, gg_df$Ho)

cowplot::plot_grid(plotlist = plt)
```
# Model selection
```{r}
library(MuMIn)

mod <- lm(Ho ~ BurnDate + NDVI + bio2_dif_lig + bio2_dif_lh + gHM + CHELSA_bio1_1981.2010_V.2.1 + csi_past + factor(glacier) + Q + tmean_dif, data = mod_df_imputed_scaled, na.action = na.fail)

summary(mod)

# Use dredge for model selection
dredge_result <- dredge(mod)

# Select the models with delta AIC < 2
best_models <- get.models(dredge_result, subset = delta < 2)

# Get a model average
model_avg <- model.avg(best_models)

# Get the best model
best_model <- get.models(dredge_result, subset = delta == 0)[[1]]

# Get the p-values
summary(best_model)

# Get the adj RSQ
summary(best_model)$adj.r.squared

# ggplot predicted vs observed for the best_model
mod_df_imputed_scaled$predict <- predict(best_model)
ggplot(mod_df_imputed_scaled, aes(x = predict, y = Ho)) +
  geom_point() +
  #geom_abline(intercept = 0, slope = 1) +
  geom_smooth(method = "lm") +
  theme_classic() +
  labs(x = "Predicted Ho", y = "Observed Ho")
# Turn into a nice gt table
library(gt)


full_model_table <- 
  mod %>%
  broom::tidy() %>%
  mutate(term = map_chr(term, ~make_pretty_names(.x))) %>%
  filter(term != "(Intercept)") %>%
  gt() %>%
  tab_header(title = "Full model") %>%
  gtExtras::gt_hulk_col_numeric(estimate, trim = TRUE) %>%
  fmt_scientific(
    columns = c("estimate", "std.error", "statistic", "p.value"),
    decimals = 2
  ) %>%
  # bold significant rows
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = c("p.value"),
      rows = p.value < 0.05
    )
  )

best_model_table <- 
  best_model %>%
  broom::tidy() %>%
  mutate(term = map_chr(term, ~make_pretty_names(.x))) %>%
  filter(term != "(Intercept)") %>%
  gt() %>%
  tab_header(title = "Best model") %>%
  data_color(
    columns = estimate,
    method = "numeric",
    palette = "viridis",
    reverse = TRUE, domain = c(-0.00011,0)
  ) %>%
  fmt_scientific(
    columns = c("estimate", "std.error", "statistic", "p.value"),
    decimals = 2
  )
# export to html
gt::gtsave(best_model_table, filename = here("analysis", "genetic_diversity", "best_model_table.html"))

gt::gtsave(full_model_table, filename = here("analysis", "genetic_diversity", "full_model_table.html"))
```


# Plot predicted vs observed values

```{r}
gg_df <- 
  left_join(mod_df, get_coords()) %>%
  mutate(glacier = case_when(glacier == 0 ~ "Outside", glacier == 1 ~ "Inside")) %>%
  rename(Admixture = Q, `LH change in\ndiurnal range` = bio2_dif_lh, Glacier = glacier)

# Plot predicted vs observed
qplt1 <- 
  ggplot(gg_df) +
  geom_point(aes(x = predict(mod), y = Ho, col = Admixture)) +
  geom_abline(intercept = 0, slope = 1) +
  geom_smooth(aes(x = predict(mod), y = Ho), method = "lm") +
  theme_classic() +
  labs(x = "Predicted", y = "Observed") +
  scale_color_viridis_c(option = "viridis", end = 0.1, begin = 0.9)

gplt1 <- 
  ggplot(gg_df) +
  geom_point(aes(x = predict(mod), y = Ho, col = Glacier)) +
  geom_smooth(aes(x = predict(mod), y = Ho), method = "lm") +
  theme_classic() +
  labs(x = "Predicted", y = "Observed") +
  scale_color_manual(values = c("skyblue2", "lightgray"))

bio2plt1 <-
  ggplot(gg_df) +
  geom_point(aes(x = predict(mod), y = Ho, col = `LH change in\ndiurnal range`)) +
  geom_smooth(aes(x = predict(mod), y = Ho), method = "lm") +
  geom_smooth(aes(x = predict(mod), y = Ho), method = "lm") +
  theme_classic() +
  labs(x = "Predicted", y = "Observed") +
  scale_color_viridis_c(option = "inferno", end = 0.1, begin = 0.9)

# Plot values across california
qplt2 <-
  ggplot(gg_df) +
  geom_sf(data = ca) +
  geom_point(aes(x = x, y = y, col = Admixture)) +
  theme_void() +
  scale_color_viridis_c(option = "viridis", end = 0.1, begin = 0.9)

gplt2 <-
  ggplot(gg_df) +
  geom_sf(data = ca) +
  geom_point(aes(x = x, y = y, col = Glacier)) +
  theme_void() +
  scale_color_manual(values = c("skyblue2", "gray"))

bio2plt2 <-
  ggplot(gg_df) +
  geom_sf(data = ca) +
  geom_point(aes(x = x, y = y, col = `LH change in\ndiurnal range`)) +
  theme_void() +
  scale_color_viridis_c(option = "inferno", end = 0.1, begin = 0.9)

# Plot Ho vs variable
qplt3 <-
  ggplot(gg_df, aes(x = Admixture, y = Ho, col = Admixture)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_classic() +
  scale_color_viridis_c(option = "viridis", end = 0.1, begin = 0.9)

gplt3 <- 
  ggplot(gg_df) +
  geom_boxplot(aes(x = Glacier, y = Ho, fill = Glacier)) +
  geom_jitter(aes(x = Glacier, y = Ho), alpha = 0.5) +
  theme_classic() +
  scale_fill_manual(values = c("skyblue2", "lightgray"))

bio2plt3 <-
  ggplot(gg_df, aes(x =  `LH change in\ndiurnal range`, y = Ho, col = `LH change in\ndiurnal range`)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_classic() +
  scale_color_viridis_c(option = "inferno", end = 0.1, begin = 0.9)

# use cowplot to combine plots
pdf(here("analysis", "genetic_diversity", "linear_model_plot.pdf"), width = 13, height = 10)
cowplot::plot_grid(bio2plt2, bio2plt3, bio2plt1, qplt2, qplt3, qplt1, gplt2, gplt3, gplt1, ncol = 3)
dev.off() 
```

# Evaluating Q
```{r}
# with Q
summary(lm(Ho ~ BurnDate + NDVI + bio2_dif_lig + bio2_dif_lh + elev + gHM + Q + glacier + resdist + tmean_dif, data = mod_df_imputed))
# bio2_dif_lh, Q, and glacier are all significant

# without Q
summary(lm(Ho ~ BurnDate + NDVI + bio2_dif_lig + bio2_dif_lh + elev + gHM + glacier + resdist + tmean_dif, data = mod_df_imputed))
# bio2_dif_lh, Q, and glacier + bio2_dif_lig and BurnDate are all significant

# Note how Q is predicted by BurnDateTRUE/bio2_dif_lig  which are the variables that drop out in the with Q model
summary(lm(Q ~ BurnDate + NDVI + gHM + CHELSA_bio1_1981.2010_V.2.1 + bio2_dif_lig + bio2_dif_lh + elev + csi_past + glacier + resdist + tmean_dif, data = mod_df))
# we may want to add Q models however to get models of predicted admixture

plot(mod_df$gHM, mod_df$Ho)
plot(mod_df$BurnDate, mod_df$Ho)
plot(mod_df$CHELSA_bio1_1981.2010_V.2.1, mod_df$Ho)

# Plot predicted vs observed values
mod <- lm(Ho ~ BurnDate + NDVI + gHM + CHELSA_bio1_1981.2010_V.2.1 + bio2_dif_lig + bio2_dif_lh + elev + csi_past + glacier + resdist + Q  + tmean_dif, data = mod_df_imputed)
mod_df$Ho_pred <- predict(mod)
ggplot(mod_df, aes(x = Ho, y = Ho_pred)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Observed Ho", y = "Predicted Ho")  

# Plot residuals on a map
mod_sf$residuals <- residuals(mod)
ggplot(mod_sf) +
  geom_sf(data = ca) +
  geom_sf(aes(col = residuals)) +
  scale_color_viridis_c(option = "magma") +
  theme_void()
```

# Spatial model
```{r}
mod_sr <- bind_cols(mod_df, st_coordinates(mod_sf))

library(spdep)
library(spatialreg)
# Build SER model
# Create neighborhood using coordinates
# 100 km distance
nbdist <- 100000
nb <- dnearneigh(mod_sf, d1 = 0, d2 = nbdist)
listw <- nb2listw(nb, style = "W")

# plot neighbors
plot(nb, mod_sr[,c("X", "Y")], main = "Neighborhoods")

# Guides: 
# https://www.emilyburchfield.org/courses/gsa/spatial_regression_lab#:~:text=To%20determine%20whether%20to%20estimate,are%20the%20same%2C%20go%20robust.
# https://pages.cms.hu-berlin.de/EOL/gcg_quantitative-methods/Lab15_SpatialRegression.html 

# Ordinary least squares model
ols_model <- lm(Ho ~ bio2_dif_lh + glacier + Q, data = mod_df)
vif(ols_model)
ols_model <- lm(Ho ~ bio2_dif_lh + glacier + Q, data = mod_df)

# test for evidence of autocorrelation
lm_tests <- lm.RStests(ols_model, listw, test = "all")
lm_tests
# Significant: RSerr, RSlag, SARMA, adjRSlag
# Not significant: adjRSerr
# because all are significant I am testing both an error and a lag model

# lag model:
# CAN'T INCLUDE RESDIST IN THIS MODEL SINCE IT IS CORRELATED WITH GEODIST
slm_model <- lagsarlm(Ho ~ bio2_dif_lh + glacier + Q, data = mod_df, listw = listw, na.action = "na.omit")

# error model:
sem_model <- errorsarlm(Ho ~ bio2_dif_lh + glacier + Q, data = mod_df, listw = listw, na.action = "na.omit")

# Compare AIC
writeLines(paste("AIC OLS:", round(AIC(ols_model), 2)))
writeLines(paste("AIC SEM:", round(AIC(sem_model), 2)))
writeLines(paste("AIC SLM:", round(AIC(slm_model), 2)))
# SLM is the best model based on AIC
```

```{r, fig.width = 8, fig.height = 6}
# Plot residuals
resids <- 
  mod_sf %>% 
  mutate(OLS = residuals(ols_model), SEM = residuals(sem_model), SLM = residuals(slm_model)) %>% 
  pivot_longer(c(OLS, SEM, SLM), values_to = "spmod_residuals")

ggplot(resids) + 
  geom_sf(data = ca) +
  geom_sf(aes(col = spmod_residuals), pch = 16, cex = 2) + 
  facet_grid(~name, switch = "y") +
  scale_color_viridis_c(option = "plasma") +
  ggtitle("Comparison of model residuals") +
  theme_void() + 
  theme(plot.title = element_text(hjust = 0.5), strip.text.y = element_text(angle = 90))
# SEM model is the best model based on residuals
```

```{r}
# print model summary
writeLines("Final model summary:")
summary(slm_model)

coeff_df <- 
  summary(slm_model)$Coef %>% 
  data.frame() %>%
  filter(`Pr...z..` < 0.05) 

coeff_df
```

# HTML Maps
```{r}
coords <- 
  get_coords() %>% 
  left_join(get_het()) %>% 
  select(SampleID, x, y, Ho)

write_csv(coords, here("data", "coords_ho.csv"))

library(leaflet)
library(readr)
library(here)

# Read the data
coords <- read_csv(here("data", "coords_ho.csv"))

# Create a color palette for the Ho variable
pal <- colorNumeric(palette = "viridis", domain = coords$Ho)

# Create the map
leaflet(coords) %>%
  addTiles() %>%
  addCircleMarkers(
    ~x, ~y,
    color = ~pal(Ho),
    radius = 5,
    stroke = FALSE,
    fillOpacity = 1,
    popup = ~paste0("SampleID: ", SampleID, "<br>Ho: ", Ho)
  ) %>%
  addLegend(
    "bottomright",
    pal = pal,
    values = ~Ho,
    title = "Ho",
    opacity = 1
  )

#USGS.USImageryTopo
# Create the map with a topographic base map
map <- leaflet(coords) %>%
  addProviderTiles(leaflet::providers$USGS.USImageryTopo, options = providerTileOptions(noWrap = TRUE)) %>%
  #addProviderTiles(leaflet::providers$Esri.WorldTopoMap, options = providerTileOptions(noWrap = TRUE)) %>%
  addCircleMarkers(
    ~x, ~y,
    fillColor = ~pal(Ho),
    radius = 5,
    stroke = TRUE,
    color = "black",
    fillOpacity = 1,
    weight = 1,
    popup = ~paste0("SampleID: ", SampleID, "<br>Ho: ", Ho)
  ) %>%
  addLegend(
    "bottomright",
    pal = pal,
    values = ~Ho,
    title = "Ho",
    opacity = 1
  )

# Save the map as an HTML file
htmlwidgets::saveWidget(map, "map.html")

# Open the HTML file in your web browser
browseURL("map.html")
```

# Future climate change
```{r}
future_ho <- 
  df_future %>%
  left_join(get_het()) 

ggplot(future_ho, aes(x = csi, y = Ho)) +
  geom_point() +
  facet_wrap(~model_name, scales = "free_x") +
  theme_classic() +
  labs(x = "Value", y = "Ho") +
  geom_smooth(method = "lm", se = TRUE) +
  stat_regline_equation(
        aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
        formula = y ~ x
      ) +
  stat_cor(label.y = min(future_ho$Ho, na.rm =TRUE)) +
  theme(strip.background = element_blank(), strip.text.x = element_text(hjust = 0, size = 12))


```

# BIO2 DIF LH vs BIO1 DIF LIG
```{r}
paleovars <- rast(here("data", "env", "paleoclim", "paleoclim.tif"))
bio2_dif_cur_lh <- paleovars$cur_bio_2 - paleovars$lh_bio_2
bio1_dif_cur_lig <- paleovars$cur_bio_1 - paleovars$lig_bio_1
bio1_dif_lgm_lig <- paleovars$lgm_bio_1 - paleovars$lig_bio_1
bio1_dif_lh_lgm <- paleovars$lh_bio_1 - paleovars$lgm_bio_1
bio1_dif_cur_lh <- paleovars$cur_bio_1 - paleovars$lh_bio_1
bio1_dif_cur_lgm <- paleovars$cur_bio_1 - paleovars$lgm_bio_1


make_pretty2 <- function(x){
  if (x == "bio2_dif_cur_lh") return("A. CUR - LH diurnal range")
  if (x == "bio1_dif_cur_lig") return("B. CUR - LIG temp")
  if (x == "bio1_dif_cur_lgm") return("C. CUR - LGM temp")
  if (x == "bio1_dif_cur_lh") return("D. CUR - LH temp")
  if (x == "bio1_dif_lgm_lig") return("E. LGM - LIG temp")
  if (x == "bio1_dif_lh_lgm") return("F. LH - LGM temp")
  return(x)
}

ggdf <-
  data.frame(
    bio2_dif_cur_lh = terra::extract(bio2_dif_cur_lh, coords, ID = FALSE)[,1],
    bio1_dif_cur_lig = terra::extract(bio1_dif_cur_lig, coords, ID = FALSE)[,1],
    bio1_dif_lgm_lig = terra::extract(bio1_dif_lgm_lig, coords, ID = FALSE)[,1],
    bio1_dif_lh_lgm = terra::extract(bio1_dif_lh_lgm, coords, ID = FALSE)[,1],
    bio1_dif_cur_lh = terra::extract(bio1_dif_cur_lh, coords, ID = FALSE)[,1],
    bio1_dif_cur_lgm = terra::extract(bio1_dif_cur_lgm, coords, ID = FALSE)[,1],
    coords
  ) %>%
  left_join(het) %>%
  pivot_longer(contains("bio")) %>%
  mutate(pretty_name = map_chr(name, ~make_pretty2(.x)))

# Figuring out arc pattern in LGM based calcs:
ggplot(filter(ggdf, name %in% c("bio1_dif_lh_lgm"))) +
  geom_sf(data = ca) +
  geom_sf(aes(geometry = geometry, col = value)) +
  facet_wrap(~pretty_name) +
  theme_void() +
  scale_color_viridis_c(option = "magma", na.value = NA)
ggplot(filter(ggdf, name %in% c("bio1_dif_cur_lh"))) +
  geom_sf(data = ca) +
  geom_sf(aes(geometry = geometry, col = value)) +
  facet_wrap(~pretty_name) +
  theme_void() +
  scale_color_viridis_c(option = "magma", na.value = NA)

ggplot(filter(ggdf, name %in% c("bio2_dif_cur_lh", "bio1_dif_cur_lig")), aes(x = value, y = Ho)) +
  geom_point() +
  facet_wrap(~pretty_name, scales = "free_x") +
  theme_classic() +
  labs(x = "Value", y = "Ho") +
  geom_smooth(method = "lm", se = TRUE) +
  stat_regline_equation(
        aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
        formula = y ~ x
      ) +
  stat_cor(label.y = 0) +
  theme(strip.background = element_blank(), strip.text.x = element_text(hjust = 0, size = 12))


ggplot(filter(ggdf, !grepl("bio2", name)), aes(x = value, y = Ho)) +
  geom_point() +
  facet_wrap(~pretty_name, scales = "free_x") +
  theme_classic() +
  labs(x = "Value", y = "Ho") +
  geom_smooth(method = "lm", se = TRUE) +
  stat_regline_equation(
        aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
        formula = y ~ x
      ) +
  stat_cor(label.y = 0) +
  theme(strip.background = element_blank(), strip.text.x = element_text(hjust = 0, size = 12))
```

# ROH models

## Models
```{r}
mod_df2 <- 
  mod_df_imputed_scaled %>% 
  left_join(dplyr::select(q, SampleID, cluster), by = "SampleID") %>% 
  drop_na(cluster)

mod <- lm(froh0 ~ BurnDate + NDVI + bio2_dif_lig + bio2_dif_lh + elev + gHM + CHELSA_bio1_1981.2010_V.2.1 + csi_past + factor(glacier) + resdist + Q + tmean_dif, data = mod_df2, na.action = na.fail)


summary(mod)

# Use dredge for model selection
dredge_result <- dredge(mod)

# Select the models with delta AIC < 2
best_models <- get.models(dredge_result, subset = delta < 2)

# Get a model average
model_avg <- model.avg(best_models)

# Get the best model
best_model <- get.models(dredge_result, subset = delta == 0)[[1]]

# Get the p-values
summary(best_model)

# Get the adj RSQ
summary(best_model)$adj.r.squared

# ggplot predicted vs observed for the best_model
mod_df_imputed_scaled$predict <- predict(best_model)
ggplot(mod_df_imputed_scaled, aes(x = predict, y = Ho)) +
  geom_point() +
  #geom_abline(intercept = 0, slope = 1) +
  geom_smooth(method = "lm") +
  theme_classic() +
  labs(x = "Predicted Ho", y = "Observed Ho")

# Turn into a nice gt table
library(gt)

best_model_df <- 
  best_model %>%
  broom::tidy() %>%
  mutate(name = map_chr(term, ~make_pretty_names(.x))) %>%
  filter(name != "(Intercept)")
  
best_model_table <- 
  best_model_df %>%
  dplyr::select(-term) %>%
  rename(term = name) %>%
  gt() %>%
  tab_header(title = "Best model") %>%
  data_color(
    columns = estimate,
    method = "numeric",
    palette = "viridis",
    reverse = TRUE, domain = c(-0.00011,0)
  ) %>%
  fmt_scientific(
    columns = c("estimate", "std.error", "statistic", "p.value"),
    decimals = 2
  )
# export to html
gt::gtsave(best_model_table, filename = here("analysis", "genetic_diversity", "roh_best_model_table.html"))
```

```{r}
gg_df <-
  mod_df %>%
  dplyr::select(any_of(best_model_df$term), glacier, froh0, cluster) 

ggplot(gg_df, aes(x = gHM, y = log(froh0 + 1))) +
  geom_point() +
  theme_classic() +
  labs(x = "Value", y = "Froh0") +
  geom_smooth(method = "lm", se = TRUE) +
  #facet_wrap(~cluster) +
  stat_regline_equation(
        aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
        formula = y ~ x
      ) +
  stat_cor(label.y = -0.05) +
  theme(strip.background = element_blank(), strip.text.x = element_text(hjust = 0, size = 12))


gg_df <-
  mod_df %>%
  dplyr::select(any_of(best_model_df$term), glacier, froh0) 

gg_df2 <- 
  gg_df %>% 
  pivot_longer(-froh0) %>%
  mutate(name =  map_chr(name, ~make_pretty_names(.x)))

ggplot(gg_df2, aes(x = value, y = log(froh0 + 1))) +
  geom_point() +
  theme_classic() +
  labs(x = "Value", y = "Froh0") +
  geom_smooth(method = "lm", se = TRUE) +
  facet_wrap(~name, scale = "free_x") +
  stat_regline_equation(
        aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
        formula = y ~ x
      ) +
  stat_cor(label.y = 0) +
  ylim(0, NA) +
  theme(strip.background = element_blank(), strip.text.x = element_text(hjust = 0, size = 12))
```
