
```{r}
library("here")
library("tidyverse")
library("ggpubr")
library("terra")
library("sf")
library("gridExtra")
library("rpaleoclim")
library("spdep")
library("spatialreg")
library("MuMIn")

source(here("general_functions.R"))
source(here("analysis", "genetic_diversity", "genetic_diversity.R"))
source(here("analysis", "sdm", "paleovars.R"))
ca <- get_ca()
pdfpath <- here("analysis", "sdm", "pdfs")

coords <- get_coords(sf = TRUE)
het <- get_het()
dem <- get_dem(r= TRUE)
```

# Hypothesis 1: Genetic diversity explained by paleoclimate

```{r}
# Get paleoclimate data
# Created using paleoclimate.Rmd
paleovars <- rast(here("data", "env", "paleoclim", "paleoclim.tif"))

# Get important var for change in climate (based on paleoclimate.Rmd)
bio1_dif_lig <- paleovars$cur_bio_1 - paleovars$lig_bio_1
bio1_dif_lgm_lig <- paleovars$lgm_bio_1 - paleovars$lig_bio_1
bio1_dif_cur_lgm <- paleovars$cur_bio_1 - paleovars$lgm_bio_1
paleoclim_mod_df <- 
  data.frame(
    SampleID = coords$SampleID,
    paleo_change = terra::extract(bio1_dif_lig, coords, ID = FALSE)[,1],
    paleo_change_lgm = terra::extract(bio1_dif_lgm_lig, coords, ID = FALSE)[,1],
    paleo_change_cur = terra::extract(bio1_dif_cur_lgm, coords, ID = FALSE)[,1]
    )

# Plots of climate change
gg_df <-
  as.data.frame(paleovars[[c("cur_bio_1", "lh_bio_1", "lgm_bio_1", "lig_bio_1")]]/10, xy = TRUE) %>%
  mutate(dif_lh = cur_bio_1 - lh_bio_1,
         dif_lgm = cur_bio_1 - lgm_bio_1,
         dif_lig = cur_bio_1 - lig_bio_1) 

gg_df1 <- 
  gg_df %>% 
  select(-contains("dif"), x, y) %>% 
  pivot_longer(c(-x, -y)) %>% 
  mutate(name = factor(name, levels = unique(name))) %>%
  mutate(name = case_when(
    name == "cur_bio_1" ~ "Current",
    name == "lh_bio_1" ~ "Late Holocene",
    name == "lgm_bio_1" ~ "LGM",
    name == "lig_bio_1" ~ "LIG"
  ))
  
gg_df2 <- 
  gg_df %>% 
  select(contains("dif"), x, y) %>% 
  pivot_longer(c(-x, -y)) %>% 
  mutate(name = factor(name, levels = unique(name)))%>%
  mutate(name = case_when(
    name == "dif_lh" ~ "Current - Late Holocene",
    name == "dif_lgm" ~ "Current - LGM",
    name == "dif_lig" ~ "Current - LIG"
  ))

plt1 <-
  ggplot(gg_df1) +
  geom_sf(data = ca) +
  geom_raster(aes(x = x, y = y, fill = value)) +
  scale_fill_viridis_c(option = "magma", na.value = "gray") +
  facet_wrap(~name, nrow = 1) +
  theme_void() +
  labs(fill = "Mean\nTemperature")

plt2 <-
  ggplot(gg_df2) +
  geom_sf(data = ca) +
  geom_raster(aes(x = x, y = y, fill = value)) +
  scale_fill_viridis_c(option = "magma", na.value = "gray") +
  facet_wrap(~name, nrow = 1) +
  theme_void() +
  labs(fill = "Mean\nTemperature\nChange")

gg_df3 <- gg_df2 %>% group_by(name) %>% mutate(value = scale(value))
plt3 <-
  ggplot(gg_df3) +
  geom_sf(data = ca) +
  geom_raster(aes(x = x, y = y, fill = value))+
  scale_fill_viridis_c(option = "magma", na.value = "gray") +
  facet_wrap(~name, nrow = 1) +
  theme_void() +
  labs(fill = "Mean\nTemperature\nChange\n(scaled)")

#cowplot::plot_grid(plt1, plt2, plt3, nrow = 3)
```

# Hypothesis 2: Genetic diversity explained by past climate stability

### Climate stability (CSI)

Original paper: 
Herrando-Moraira, S., Nualart, N., Galbany-Casals, M. et al. Climate Stability Index maps, a global high resolution cartography of climate stability from Pliocene to 2100. Sci Data 9, 48 (2022). https://doi.org/10.1038/s41597-022-01144-5

Note: Bioclimatic variables do not contain Bio 2 or Bio 5, but they do contain other variables correlated with those variables

```{r, fig.width = 10, fig.height = 2.5}
csi_past <- rast(here("data", "env", "stability", "Layers", "past", "csi_past.tif"))
csi_past <- crop(csi_past, ca)
csi_past <- mask(csi_past, ca)

files <- list.files(here("data", "env", "stability", "Layers", "future"), full.names = TRUE, pattern = "csi*")
csi_future <- rast(map(files, rast))
csi_future <- crop(csi_future, ca)
csi_future <- mask(csi_future, ca)

csi_stack <- c(csi_past, csi_future)
names(csi_stack) <- c("Past",  "SSP 1-2.6", "SSP 2-4.5", "SSP 3-7.0", "SSP 5-8.5")
writeRaster(csi_stack, here("data", "env", "stability", "csi.tif"), overwrite = TRUE)
csi_df <- 
  terra::as.data.frame(csi_stack, xy = TRUE) %>% 
  pivot_longer(-c(x, y), 
  names_to = "model", 
  values_to = "CSI")
```

```{r, fig.height = 4, fig.width = 4.5}
csi_vals <- extract(csi_past, coords)
df_past <- data.frame(csi_past = csi_vals[,1], coords)
```

```{r, fig.height = 4, fig.width = 16}
csi_vals <- extract(csi_future, coords, ID = FALSE)
df_future <- 
  data.frame(1 - csi_vals, coords) %>% 
  pivot_longer(starts_with("csi"), names_to = "future_model", values_to = "csi") %>%
  mutate(model_name = case_when(
    future_model == "csi_future_ssp126" ~ "SSP 1-2.6",
    future_model == "csi_future_ssp245" ~ "SSP 2-4.5",
    future_model == "csi_future_ssp370" ~ "SSP 3-7.0",
    future_model == "csi_future_ssp585" ~ "SSP 5-8.5"
  ))
```

```{r}
coords <- get_coords(sf = TRUE)
# Do 1 - to make the values more intuitive
csi_mod_df <- data.frame(csi = 1 - extract(csi_past, coords, ID = FALSE), coords)
```

# Hypothesis 3: Genetic diversity explained by admixture
```{r}
q <- read_csv(here("analysis", "admixture", "outputs_old_nov16", "Q9.csv"))

q_mod_df <- 
  q %>%
  dplyr::select(SampleID, Q) %>%
  # Do 1- to make the values more intuitive
  mutate(Q = 1 - Q)
```

# Hypothesis 4: Genetic diversity explained by current environment 
```{r}
env_data <- rast(here("data", "env", "california_env_stack.tif"))
env_data$BurnDate <- as.numeric(env_data$BurnDate > 0)
env_mod_df <- data.frame(terra::extract(env_data, coords, ID = FALSE))
env_mod_df$SampleID <- coords$SampleID
```

```{r}
coords_buffer <- st_buffer(coords, 10000)
gHM_buffer <- extract(env_data[["gHM"]], coords_buffer, fun = mean, na.rm = TRUE)

env_mod_df$gHM <- gHM_buffer$gHM
env_mod_df$buffer <- 10000
```

# Hypothesis 5: Genetic diversity explained by glaciation
```{r}
lgm_glaciers <- st_read(here("data", "env", "Glaciotect_NA_lgm", "shp", "LGM_maxextent.shp"))
# Note: have to first intersect and then transform otherwise you get a node error
lgm_glaciers <- st_intersection(lgm_glaciers, st_transform(ca, crs = st_crs(lgm_glaciers)))
lgm_glaciers <- st_transform(lgm_glaciers, crs = st_crs(ca))
# Add a new column to coords indicating whether each point is inside the lgm shapefile
coords <- get_coords(sf = TRUE)
glacier_mod_df <- 
  coords %>%
  mutate(glacier = as.numeric((lengths(st_within(coords, lgm_glaciers)) > 0))) 
```

# Hypothesis 6: Genetic diversity explained by recent climate change
Download PRISM mean temperature data from historical data: https://prism.oregonstate.edu/historical/

Data can be downloaded for all years using:
wget -r ftp://prism.oregonstate.edu/monthly/tmean/

mv prism.oregonstate.edu/monthly/tmean/* .

for year in {1895..2024}
do
  unzip '/media/wanglab/798f0e01-89d1-4d0f-8ed8-ef323be70ab91/Anusha/GitHub/ccgpscelop/data/env/prism/TEMP/'$year'/*.zip' -d '/media/wanglab/798f0e01-89d1-4d0f-8ed8-ef323be70ab91/Anusha/GitHub/ccgpscelop/data/env/prism/TEMP/'$year
done

```{r}
coords <- get_coords(sf = TRUE)
year = 1895

# Get prism data
get_prism <- function(year  = 1895, output = "df") {
  # Create PRISM path
  prism_path <- here("data", "env", "prism", "PRISM_tmean_stable_4kmM3", year)

  # List annual file
  file <- list.files(prism_path, pattern = paste0(year, "_*.bil.bil$"), full.names = TRUE)

  # Load data
  raster <- terra::rast(file)

  if (output == "raster"){
    return(raster)
  } else {
    vals <- terra::extract(raster, st_transform(coords, st_crs(raster)), ID = FALSE)
    names(vals) <- "tmean"
    df <- data.frame(year = year, tmean = vals, SampleID = coords$SampleID)
    return(df)
  }
}

# Loop over the years of interest
tmean_his <- map(1895:1905, get_prism, .progress =TRUE) %>% bind_rows()  
tmean_his <- tmean_his %>% group_by(SampleID) %>% summarise(tmean_his = mean(tmean, na.rm = TRUE))
tmean_cur <- map(2010:2020, ~get_prism(year = .x), .progress =TRUE) %>% bind_rows()  
tmean_cur <- tmean_cur %>% group_by(SampleID) %>% summarise(tmean_cur = mean(tmean, na.rm = TRUE))

tmean_his_r <- map(1895:1905, get_prism, .progress =TRUE, output= "raster") %>% rast() %>% mean()
tmean_cur_r <- map(2010:2020, get_prism, .progress =TRUE, output= "raster") %>% rast() %>% mean()
tmean_dif_r <- tmean_cur_r - tmean_his_r
tmean_dif_r <- crop(tmean_dif_r, ca)
tmean_dif_r <- mask(tmean_dif_r, st_transform(ca, st_crs(tmean_dif_r)))

tmean_mod_df <- 
  left_join(tmean_his, tmean_cur) %>% 
  mutate(tmean_dif = tmean_cur - tmean_his) %>%
  right_join(coords) %>% 
  dplyr::select(SampleID, tmean_dif, tmean_cur)
```

```{r}
env_list <- list(
  paleo_change = bio1_dif_lig, 
  csi_past = 1 - csi_past, 
  env_data = env_data, 
  glacier = lgm_glaciers, 
  tmean_dif = tmean_dif_r
  )
```

# Models
```{r}
mod_vars <- c("NDVI", "gHM", "paleo_change", "CHELSA_bio1_1981.2010_V.2.1", "CHELSA_bio12_1981.2010_V.2.1", "Q", "csi_past", "glacier", "tmean_dif", "paleo_change_lgm", "paleo_change_cur", "x", "y")

mod_df <- 
  get_coords() %>% 
  left_join(get_het()) %>%
  left_join(env_mod_df) %>%
  left_join(paleoclim_mod_df) %>%
  left_join(q_mod_df) %>%
  left_join(csi_mod_df) %>%
  left_join(glacier_mod_df) %>%
  left_join(tmean_mod_df) %>%
  dplyr::select(Ho, SampleID, geometry, all_of(mod_vars)) %>%
  drop_na(Ho) %>%
  # REMOVE CHANNEL ISLAND
  filter(SampleID != "Scelocci_CCGPMC_MW01-3-14")
  
# Count NAs in each column
have_nas <- colSums(is.na(mod_df))[colSums(is.na(mod_df)) > 0]
have_nas

# NDVI has some NA values (< 3), impute to the value of the closest sample
mod_sf <- mod_df %>% st_as_sf() %>% st_transform(3310)
close_impute <- function(var, mod_sf){
  map_dbl(1:nrow(mod_sf), \(i){ 
    if (!is.na(st_drop_geometry(mod_sf[i, var]))) return(mod_sf[i, var][[1]])
    # Get the closest sample that is not NA
    nearest <- st_nearest_feature(mod_sf[i,], drop_na(mod_sf[-i,]), 1)
    # Get the value of the closest sample
    return(drop_na(mod_sf)[nearest, var][[1]])
  }) 
}

vars <- names(have_nas)
names(vars) <- vars
mod_sf_imputed <- map(vars, ~close_impute(.x, mod_sf), .progress = TRUE) 
mod_df_imputed <- bind_cols(dplyr::select(mod_df, -all_of(vars)), bind_cols(mod_sf_imputed))

# Confirm no NAs:
colSums(is.na(mod_df_imputed))[colSums(is.na(mod_df_imputed)) > 0]

# Scale the variables (except glacier)
mod_df_imputed_scaled <- 
  mod_df_imputed %>% 
  mutate(across(c(all_of(mod_vars), -glacier), scale))

# Convert to sf
mod_sf <- st_as_sf(mod_df) %>% st_transform(3310)

# Check correlations
# Compute the correlation matrix
r <- cor(mod_df_imputed_scaled[,mod_vars], use = "pairwise.complete.obs")
# Set the diagonal and lower triangle to NA
r[lower.tri(r, diag = TRUE)] <- NA
# Find the pairs with correlations greater than 0.6
high_corr <- which(abs(r) > 0.6, arr.ind = TRUE)
# Print the pairs
high_corr_pairs <- data.frame(
  Var1 = rownames(r)[high_corr[, 1]],
  Var2 = rownames(r)[high_corr[, 2]],
  Correlation = r[high_corr]
)
high_corr_pairs
# paleo_change is correlated with CHELSA_bio1_1981.2010_V.2.1, gHM, and csi_past
```

# Model selection
```{r}
# Full model
# Note: included both regular and poly gHM to see which was the better fit using dredge
mod <- lm(Ho ~  NDVI + paleo_change + poly(gHM, 2) + gHM + CHELSA_bio1_1981.2010_V.2.1 + CHELSA_bio12_1981.2010_V.2.1 + csi_past + factor(glacier) + Q + tmean_dif + x + y, data = mod_df_imputed_scaled, na.action = na.fail)

summary(mod)

# Use dredge for model selection
dredge_result <- dredge(mod)

# Select the models with delta AIC < 2
best_models <- get.models(dredge_result, subset = delta < 2)

# Get a model average
model_avg <- model.avg(best_models)
model_avg_summary <- summary(model_avg)

# Get coefficient and p-values
coeffs <- 
  model_avg_summary[["coefmat.subset"]] %>% 
  data.frame(., predictor = rownames(.))

# Get confidence intervals
confints <-
  confint(model_avg) %>% 
  data.frame(., predictor = rownames(.))

# Get sum of weights (relative importance)
sws <- 
  sw(model_avg_summary) %>% 
  data.frame() %>%
  t() %>%
  data.frame() %>%
  pivot_longer(everything(), names_to = "predictor", values_to = "Sum of weights") %>%
  drop_na(`Sum of weights`)

# Combine everything
model_df <-
  left_join(coeffs, confints, by = "predictor")%>%
  left_join(sws, by = "predictor") %>%
  rename(p = `Pr...z..`) 

# For the model average, paleo_change, Q, and glacier are significant
sigvars <-
  model_df %>% 
  filter(predictor != "(Intercept)") %>%
  filter(p < 0.05)
print(sigvars)

# Look at the best model
best_model <- get.models(dredge_result, subset = delta < 2)[[1]]
AIC(best_model)

# Get the p-values 
summary(best_model) 

# Adj RSQ
summary(best_model)$adj.r.squared
```

```{r}
make_pretty_names <- function(x){
  if (x == "BurnDate") return("Burned area (yes/no)")
  if (x == "paleo_change") return("Paleoclimate temperature change")
  if (grepl("CHELSA_bio12", x)) return("Contemporary precipitation")
  if (grepl("CHELSA_bio1", x)) return("Contemporary temperature")
  if (grepl("csi", x)) return("Past climate stability")
  if (grepl("gHM", x)) return("Human modification")
  if (grepl("glacier", x)) return("Glacier (inside/outside)")
  if (grepl("Q", x)) return("Admixture")
  if (grepl("tmean_dif", x)) return("Contemporary temperature change")
  return(x)
}

# Plot all variables
gg_df <- 
  mod_df %>%
  pivot_longer(all_of(mod_vars)) %>%
  mutate(name =  map_chr(name, ~make_pretty_names(.x))) 

ggplot(gg_df, aes(x = value, y = Ho)) +
  geom_point() +
  facet_wrap(~name, scales = "free_x") +
  theme_classic() +
  labs(x = "Value", y = "Ho") +
  geom_smooth(method = "lm", se = TRUE) +
  stat_regline_equation(
        aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
        formula = y ~ x
      ) +
  stat_cor(label.y = 0) +
  theme(strip.background = element_blank(), strip.text.x = element_text(hjust = 0, size = 12))

# Plot all variables across CA
gg_df <- 
  get_coords(sf = TRUE) %>%
  left_join(mod_df)

plt <- 
  map(mod_vars, ~{
    ggplot(gg_df) +
      geom_sf(data = ca) +
      geom_sf(aes(col = .data[[.x]])) +
      labs(col = make_pretty_names(.x)) +
      scale_color_viridis_c(option = "magma", begin = 0.05, end = 0.95) +
      theme_void() 
  })

cowplot::plot_grid(plotlist = plt)
```

```{r}
# ggplot predicted vs observed for the best_model
mod_df_imputed_scaled$predict <- predict(best_model)
mod_df_imputed_scaled$residuals <- residuals(best_model)

ggplot(mod_df_imputed_scaled, aes(x = predict, y = Ho)) +
  geom_point() +
  #geom_abline(intercept = 0, slope = 1) +
  geom_smooth(method = "lm") +
  theme_classic() +
  labs(x = "Predicted Ho", y = "Observed Ho")

gg_df <-
  mod_df_imputed_scaled %>%
  pivot_longer(c(Ho, predict)) %>%
  mutate(name = case_when(
    name == "Ho" ~ "Observed",
    name == "predict" ~ "Predicted"
  ))

# Plot residuals
ggplot(gg_df) +
  geom_sf(data = ca, fill = "#5b5b5b") +
  geom_sf(aes(geometry = geometry, col = residuals), cex = 2) +
  theme_void() +
  scale_color_gradient2(low = "#ff5f33", mid = "#fffcec", high = "#00c6af", midpoint = 0)

# Turn into a nice gt table
library(gt)

full_model_table <- 
  mod %>%
  broom::tidy() %>%
  mutate(term = map_chr(term, ~make_pretty_names(.x))) %>%
  filter(term != "(Intercept)") %>%
  gt() %>%
  tab_header(title = "Full model") %>%
  gtExtras::gt_hulk_col_numeric(estimate, trim = TRUE) %>%
  fmt_scientific(
    columns = c("estimate", "std.error", "statistic", "p.value"),
    decimals = 2
  ) %>%
  # bold significant rows
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = c("p.value"),
      rows = p.value < 0.05
    )
  )

best_model_table <- 
  best_model %>%
  broom::tidy() %>%
  mutate(term = map_chr(term, ~make_pretty_names(.x))) %>%
  filter(term != "(Intercept)") %>%
  gt() %>%
  tab_header(title = "Best model") %>%
  gtExtras::gt_hulk_col_numeric(columns = estimate, trim = TRUE, domain = c(-0.00011, 0.00011)) %>%
  fmt_scientific(
    columns = c("estimate", "std.error", "statistic", "p.value"),
    decimals = 2
  )
# export to html
gt::gtsave(best_model_table, filename = here("analysis", "genetic_diversity", "best_model_table.html"))

gt::gtsave(full_model_table, filename = here("analysis", "genetic_diversity", "full_model_table.html"))
```


```{r}
# Plot all variables with significant associations with Ho & a intermediate+ correlation with the sig vars
# Get significant variables from the model
sigvars <- 
  data.frame(summary(best_model)$coefficients) %>%
  mutate(var = row.names(.)) %>%
  filter(!(var %in% c("(Intercept)"))) %>%
  filter(`Pr...t..` < 0.05) %>%
  filter(!grepl("glacier", var)) %>%
  pull(var)

named_mod_vars <- mod_vars
names(named_mod_vars) <- mod_vars
# drop glacier
named_mod_vars <- named_mod_vars[-which(named_mod_vars == "glacier")]
Ho_r <- map_dbl(named_mod_vars, ~cor.test(mod_df$Ho, mod_df[[.x]])$p.value) < 0.05
Ho_r <- Ho_r[Ho_r]

# Test for significant correlations with model variables
cor_vars <-
  expand.grid(var1 = mod_vars, var2 = sigvars) %>%
  mutate(p.value = pmap_dbl(list(var1, var2), ~ cor.test(mod_df_imputed_scaled[[..1]], mod_df_imputed_scaled[[..2]], use = "complete.obs")$p.value)) %>%
  filter(!(var1 %in% sigvars)) %>%
  filter(var1 %in% names(Ho_r)) %>%
  filter(p.value < 0.05) %>%
  pull(var1)

gg_df2 <- 
  mod_df %>%
  pivot_longer(all_of(cor_vars)) %>%
  mutate(name =  map_chr(name, ~make_pretty_names(.x))) %>%
  filter(!(name %in% c("Glacier (inside/outside)")))

ggplot(gg_df2, aes(x = value, y = Ho)) +
  geom_point() +
  facet_wrap(~name, scales = "free") +
  theme_classic() +
  labs(x = "Value", y = "Ho") +
  geom_smooth(method = "lm", se = TRUE) +
  stat_cor(label.y = 0) +
  theme(strip.background = element_blank(), strip.text.x = element_text(hjust = 0, size = 12))
```


```{r}
# Make pretty plots 
mod <- best_model
gg_df <- 
  left_join(mod_df, get_coords()) %>%
  # TEMP
  #filter(Ho > 0.0005) %>%
  mutate(glacier = factor(glacier, levels = c(0, 1), labels = c("Outside", "Inside"))) %>%
  rename(
      Admixture = Q, 
      `Paleoclimate\ntemperature\nchange` = paleo_change, 
      `Contemporary\ntemperature` = CHELSA_bio1_1981.2010_V.2.1,
      `Human\nmodification` = gHM,
      `Past climate\nstability` = csi_past,
      `Contemporary\ntemperature\nchange` = tmean_dif,
      `Glacier` = glacier) 

vars <- c("Paleoclimate\ntemperature\nchange", "Contemporary\ntemperature", "Admixture", "Human\nmodification", "Past climate\nstability", "Contemporary\ntemperature\nchange")
varcols <- c("inferno", "rocket", "viridis", "rocket", "mako", "inferno")
varcols <- rep("mako", length(vars))

# Plot predicted vs observed
plt1 <- 
  map2(vars, varcols, ~{
    ggplot(gg_df) +
      geom_point(aes(x = predict(mod), y = Ho, col = .data[[.x]])) +
      geom_abline(intercept = 0, slope = 1) +
      geom_smooth(aes(x = predict(mod), y = Ho), method = "lm") +
      theme_classic() +
      labs(x = "Predicted", y = "Observed") +
      scale_color_viridis_c(option = .y, end = 0.1, begin = 0.9)
  })

# Plot values across california
plt2 <-
  map2(vars, varcols, ~{
    ggplot(gg_df) +
      geom_sf(data = ca) +
      geom_point(aes(x = x, y = y, col = .data[[.x]])) +
      theme_void() +
      scale_color_viridis_c(option = .y, end = 0.1, begin = 0.9)
  })

# Plot Ho vs variable
plt3 <- map2(vars, varcols, ~{
  ggplot(gg_df, aes(x = .data[[.x]], y = Ho)) +
    geom_point(aes(col = .data[[.x]])) +
    geom_smooth(method = "lm") +
    theme_classic() +
    labs(x = gsub("\n", " ", .x), y = "Ho") +
    scale_color_viridis_c(option = .y, end = 0.1, begin = 0.9) +
    theme(legend.position = "none") +
    stat_cor(label.y = min(gg_df[["Ho"]], na.rm = TRUE))
})

# Plot Ho
plt4 <-
  map(1:(length(vars)+1), ~{
    ggplot(gg_df) +
      geom_sf(data = ca) +
      geom_point(aes(x = x, y = y, col = Ho)) +
      theme_void() +
      scale_color_viridis_c(option ="mako", end = 0.1, begin = 0.9, direction = -1)
})

pdf(here("analysis", "genetic_diversity", "ho.pdf"), width = 4.5, height = 4.5)
ggplot(gg_df) +
  geom_sf(data = ca) +
  geom_point(aes(x = x, y = y, col = Ho)) +
  theme_void() +
  scale_color_viridis_c(option ="mako") +
  theme(plot.margin = margin(t = 1, r = 1, b = 1, l = 1, unit = "mm"))
dev.off()

# Make glacier plots
plt2_glacier <-
  ggplot(gg_df) +
    geom_sf(data = ca) +
    geom_sf(data = lgm_glaciers, col = "cyan3", fill = "cyan3", alpha = 0.3, lty = "solid", lwd = 0.7) + 
    geom_point(aes(x = x, y = y, col =  Glacier)) +
    scale_color_manual(values = c("darkgray", "cyan4")) +
    theme_void() +
    theme(legend.position.inside = c(0, 0))

plt1_glacier <-
  ggplot(gg_df) +
    geom_point(aes(x = predict(mod), y = Ho, col = Glacier)) +
    geom_abline(intercept = 0, slope = 1) +
    geom_smooth(aes(x = predict(mod), y = Ho), method = "lm") +
    theme_classic() +
    labs(x = "Predicted", y = "Observed") +
    scale_color_manual(values = c("darkgray", "cyan4"))

plt3_glacier <- 
  ggplot(gg_df) +
    geom_jitter(aes(x = `Glacier`, y = Ho, col = `Glacier`)) +
    geom_boxplot(aes(x = `Glacier`, y = Ho, fill = `Glacier`), alpha = 0.3) +
    theme_classic() +
    labs(x = "Glacier", y = "Ho") +
    scale_color_manual(values = c("darkgray", "cyan4")) +
    scale_fill_manual(values = c("darkgray", "cyan3")) +
    theme(legend.position = "none")

plt1_ls <- c(plt1, list(plt1_glacier))
plt2_ls <- c(plt2, list(plt2_glacier))
plt3_ls <- c(plt3, list(plt3_glacier))

# use cowplot to combine plots
pdf(here("analysis", "genetic_diversity", "linear_model_plot1.pdf"), width = 11, height = 24)
ordered_plts <- pmap(list(plt3_ls, plt1_ls, plt2_ls), ggpubr::ggarrange, common.legend = TRUE, legend = "right", nrow = 1)
do.call(cowplot::plot_grid, c(ordered_plts, list(nrow = length(plt1_ls))))
ordered_plts <- pmap(list(plt3_ls, plt2_ls, plt4), ggpubr::ggarrange, nrow = 1)
do.call(cowplot::plot_grid, c(ordered_plts, list(nrow = length(plt1_ls))))
dev.off()

pdf(here("analysis", "genetic_diversity", "linear_model_plot.pdf"), width = 8.5, height = 24)
ordered_plts <- pmap(list(plt3_ls, plt2_ls), ggpubr::ggarrange, common.legend = TRUE, legend = "right", nrow = 1)
do.call(cowplot::plot_grid, c(ordered_plts, list(nrow = length(plt1_ls))))
dev.off()
```

```{r}
plot_resid <- function(x){
  vars <- c("glacier", "paleo_change", "Q")
  mod_vars <- vars[vars != x]
  mod <- lm(formula(paste0("Ho ~ ", paste(mod_vars, collapse = " + "))), data = mod_df_imputed_scaled)
  resids <- residuals(mod)

  gg_df <- 
    mod_df_imputed_scaled %>%
    mutate(residuals = resids) %>%
    pivot_longer(all_of(x))
  return(gg_df)
}

resid_plts <- 
  map(c("glacier", "paleo_change", "Q"), plot_resid) %>% 
  bind_rows() %>%
  mutate(name = map_chr(name, ~make_pretty_names(.x)))

ggplot(resid_plts, aes(x = value, y = residuals)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_classic() +
  facet_wrap(~name, scales = "free") +
  labs(x = "Variable", y = "Residual Ho") +
  theme(strip.background = element_blank(), strip.text.x = element_text(hjust = 0, size = 12))
```

# Evaluating Q

```{r}
# Full model
# Note: included both regular and poly gHM to see which was the better fit using dredge
mod <- lm(Ho ~  NDVI + paleo_change + poly(gHM, 2) + gHM + CHELSA_bio1_1981.2010_V.2.1 + csi_past + factor(glacier) + tmean_dif, data = mod_df_imputed_scaled, na.action = na.fail)

summary(mod)

# Use dredge for model selection
dredge_result <- dredge(mod)

# Select the models with delta AIC < 2
best_models <- get.models(dredge_result, subset = delta < 2)

# Get a model average
model_avg <- model.avg(best_models)
model_avg_summary <- summary(model_avg)

# Get coefficient and p-values
coeffs <- 
  model_avg_summary[["coefmat.subset"]] %>% 
  data.frame(., predictor = rownames(.))

# Get confidence intervals
confints <-
  confint(model_avg) %>% 
  data.frame(., predictor = rownames(.))

# Get sum of weights (relative importance)
sws <- 
  sw(model_avg_summary) %>% 
  data.frame() %>%
  t() %>%
  data.frame() %>%
  pivot_longer(everything(), names_to = "predictor", values_to = "Sum of weights") %>%
  drop_na(`Sum of weights`)

# Combine everything
model_df <-
  left_join(coeffs, confints, by = "predictor")%>%
  left_join(sws, by = "predictor") %>%
  rename(p = `Pr...z..`) 

# For the model average, paleo_change, Q, and glacier are significant
sigvars <-
  model_df %>% 
  filter(predictor != "(Intercept)") %>%
  filter(p < 0.05)
print(sigvars)

# Look at the best models
# Get the best model
best_model <- get.models(dredge_result, subset = delta < 2)[[1]]
second_best_model <- get.models(dredge_result, subset = delta < 2)[[2]]
third_best_model <- get.models(dredge_result, subset = delta < 2)[[3]]
AIC(best_model)
AIC(second_best_model)
AIC(third_best_model)

# Get the p-values 
summary(best_model) 
summary(second_best_model)
summary(third_best_model) 

# Adj RSQ
summary(best_model)$adj.r.squared
```

```{r}
# Plot variables with non-zero Q
gg_df <- 
  filter(mod_df_imputed_scaled, Q > 0) %>%
  pivot_longer(c(all_of(mod_vars), -Q)) 

ggplot(gg_df, aes(x = value, y = Q)) +
  geom_point() +
  facet_wrap(~name, scales = "free_x") +
  theme_classic() +
  labs(x = "Value", y = "Q") +
  geom_smooth(method = "lm", se = TRUE) +
  stat_cor(label.y = 0) +
  theme(strip.background = element_blank(), strip.text.x = element_text(hjust = 0, size = 12))

# Plot variables with zero Q
gg_df <- 
  mod_df %>%
  pivot_longer(c(all_of(mod_vars), -Q)) 

ggplot(gg_df, aes(x = value, y = Q)) +
  geom_point() +
  facet_wrap(~name, scales = "free_x") +
  theme_classic() +
  labs(x = "Value", y = "Q") +
  geom_smooth(method = "lm", se = TRUE) +
  stat_cor(label.y = 0) +
  theme(strip.background = element_blank(), strip.text.x = element_text(hjust = 0, size = 12))
```

# Spatial model

```{r}
mod_sr <- bind_cols(mod_df, st_coordinates(mod_sf))

nb_test <- map(c(10000, 50000, 100000, 200000), ~{
  nb <- dnearneigh(mod_sf, d1 = 0, d2 = .x)
  listw <- nb2listw(nb, style = "W", zero.policy = TRUE)
  slm_model <- lagsarlm(Ho ~ paleo_change + CHELSA_bio1_1981.2010_V.2.1 + Q, data = mod_df, listw = listw, na.action = "na.omit", zero.policy = TRUE)
  sem_model <- errorsarlm(Ho ~ paleo_change + CHELSA_bio1_1981.2010_V.2.1 + Q, data = mod_df, listw = listw, na.action = "na.omit", zero.policy = TRUE)
  return(data.frame(slm = AIC(slm_model), sem = AIC(sem_model), d2 = .x))
}, .progress = TRUE) %>% bind_rows()

ggplot(nb_test) +
  geom_line(aes(x = d2, y = slm, col = "SLM")) +
  geom_line(aes(x = d2, y = sem, col = "SEM")) +
  theme_classic() +
  labs(x = "Distance", y = "AIC") 

nb_test %>% filter(slm == min(slm) | sem == min(sem))
```

```{r}

build_spatial_models <- function(nbdist) {
  # Assuming mod_sf, mod_df, and best_model are available in the environment
  # Create neighborhood using coordinates
  nb <- dnearneigh(mod_sf, d1 = 0, d2 = nbdist)
  listw <- nb2listw(nb, style = "W", zero.policy = TRUE)

  # Calculate Moran's I on residuals of best mod
  mod_df_imputed_scaled$residuals <- residuals(best_model)
  morans_i <- moran.test(mod_df_imputed_scaled$residuals, listw)
  print(morans_i)

  # plot neighbors - This will execute but not return in the function's output
  plot(nb, mod_sr[,c("X", "Y")], main = "Neighborhoods")

  # Ordinary least squares model
  ols_model <- lm(Ho ~ paleo_change + factor(glacier) + Q, data = mod_df)
  # test for evidence of autocorrelation
  lm_tests <- lm.RStests(ols_model, listw, test = "all")
  print(lm_tests)

  # Spatial lag model
  slm_model <- lagsarlm(Ho ~ paleo_change + Q + factor(glacier), data = mod_df, listw = listw, na.action = "na.omit", zero.policy = TRUE)

  # Spatial error model
  sem_model <- errorsarlm(Ho ~ paleo_change + Q + factor(glacier), data = mod_df, listw = listw, na.action = "na.omit", zero.policy = TRUE)

  # Compare AIC
  writeLines(paste("AIC OLS:", round(AIC(ols_model), 2)))
  writeLines(paste("AIC SEM:", round(AIC(sem_model), 2)))
  writeLines(paste("AIC SLM:", round(AIC(slm_model), 2)))

  # Return a list of models
  return(list(ols_model = ols_model, sem_model = sem_model, slm_model = slm_model))
}


sp50k <- build_spatial_models(50000)
sp100k <- build_spatial_models(100000)

# Compare AIC
writeLines(paste("AIC OLS 50k:", round(AIC(sp50k$sem_model), 2)))
writeLines(paste("AIC OLS 100k:", round(AIC(sp100k$sem_model), 2)))

summary(sp50k$sem_model)
summary(sp100k$sem_model)
```
```{r}
nbdist <- 100000
nb <- dnearneigh(mod_sf, d1 = 0, d2 = nbdist)
listw <- nb2listw(nb, style = "B", zero.policy = TRUE)
# Assuming 'mod_df' is your data frame
# Calculate spatial lag
neighbor_ids <- nb2listw(nb, style = "B", zero.policy = TRUE)$neighbours

# Find the southernmost point
st_coords <- st_coordinates(mod_sf)
southernmost <- which.max(st_coords[,2])

# Create a graph
library(igraph)
g <- graph_from_adj_list(neighbor_ids)

# Calculate shortest paths from the south to each point
shortest_paths <- shortest_paths(g, from = southernmost)$vpath

# Calculate total distance along the shortest path for each point
distances_to_south <- map_dbl(shortest_paths, \(path) {
  if (length(path) == 0) return(NA)
  if (length(path) == 1) return(0)
  path_points <- mod_sf[as.vector(path), ]
  path_distances <- sapply(2:length(path), function(i) {
    st_distance(path_points[i - 1, ], path_points[i, ])
  })
  sum(path_distances)
}, .progress = TRUE)

df <- mod_sf
df$distances_to_south <- distances_to_south

ggplot(df) +
  geom_sf(data = ca) +
  geom_sf(aes(col = distances_to_south)) +
  scale_color_viridis_c(option = "magma") +
  theme_void()

ggplot(df) +
  geom_point(aes(x = distances_to_south, y = Ho)) +
  geom_smooth(aes(x = distances_to_south, y = Ho), method = "lm") +
  theme_classic() 

ggplot(df) +
  geom_point(aes(x = distances_to_south, y = paleo_change)) +
  geom_smooth(aes(x = distances_to_south, y = paleo_change), method = "lm") +
  theme_classic() 

ggplot(df) +
  geom_point(aes(x = distances_to_south, y = CHELSA_bio1_1981.2010_V.2.1)) +
  geom_smooth(aes(x = distances_to_south, y = CHELSA_bio1_1981.2010_V.2.1), method = "lm") +
  theme_classic() 

summary(lm(Ho ~ distances_to_south + paleo_change + CHELSA_bio1_1981.2010_V.2.1 + Q, data = df))
```
```{r, fig.width = 8, fig.height = 6}
# Plot residuals
resids <- 
  mod_sf %>% 
  mutate(OLS = residuals(ols_model), SEM = residuals(sem_model), SLM = residuals(slm_model)) %>% 
  pivot_longer(c(OLS, SEM, SLM), values_to = "spmod_residuals")

ggplot(resids) + 
  geom_sf(data = ca) +
  geom_sf(aes(col = spmod_residuals), pch = 16, cex = 2) + 
  facet_grid(~name, switch = "y") +
  scale_color_viridis_c(option = "plasma") +
  ggtitle("Comparison of model residuals") +
  theme_void() + 
  theme(plot.title = element_text(hjust = 0.5), strip.text.y = element_text(angle = 90))
# SEM model is the best model based on residuals
```

```{r}
# print model summary
writeLines("Final model summary:")
summary(slm_model)

coeff_df <- 
  summary(slm_model)$Coef %>% 
  data.frame() %>%
  filter(`Pr...z..` < 0.05) 

coeff_df
```

# Spatial Model 2

```{r}
library(nlme)
library(ape)
```
```{r}
coords <- get_coords(sf = TRUE)
dframe <- mod_df_imputed_scaled %>% st_as_sf()
```
```{r}
# Construct IDW matrix
# Calculate pairwise geographic distances
geoMat <- st_distance(dframe)

# Convert to a numeric matrix while preserving the original matrix structure
geoMat <- matrix(as.numeric(units::drop_units(geoMat)), nrow = nrow(geoMat), ncol = ncol(geoMat))

# Ensure the matrix is symmetrical
geoMat[upper.tri(geoMat)] <- geoMat[lower.tri(geoMat)]

# Apply inverse distance weighting
spdata.IDW <- 1 / geoMat
diag(spdata.IDW) <- 0  # Set diagonal to 0 to avoid division by zero in IDW

# Perform a global test of Moran's I on the point observations
# All are spatially autocorrelated
Moran.I(as.numeric(dframe$paleo_change), spdata.IDW)
Moran.I(as.numeric(dframe$Q), spdata.IDW)
Moran.I(as.numeric(dframe$glacier), spdata.IDW)
```

```{r}
dummy <- rep(1, nrow(dframe))
dframe <- cbind(dframe, dummy) # Add it to the data frame
```

```{r}
mod_vars <- c("NDVI", "gHM", "paleo_change", "CHELSA_bio1_1981.2010_V.2.1", "csi_past", "glacier", "tmean_dif", "Q")
f <- as.formula(paste("Ho ~", paste(mod_vars, collapse = " + ")))
null.model <- lme(fixed = f, data = dframe, random = ~ 1 | dummy, method = "ML")
summary(null.model)

# Residuals are spatially autocorrelated
null_resids <- residuals(null.model)
null_M <- Moran.I(null_resids, spdata.IDW)
```

Next, we can run the same model with spatial correlation structures.  We can specify such a structure with the correlation option of lme().  The structures are generated using corExp for exponential, corGaus for Gaussian, or corSpher for spherical.  
```{r}
exp.sp <- update(null.model, correlation = corExp(1, form = ~ x + y), method = "ML")
writeLines("Results for exponential spatial correlation:\n")
summary(exp.sp)

#gau.sp <- update(null.model, correlation = corGaus(1, form = ~ x + y), method = "ML")
#writeLines("\nResults for Gaussian spatial correlation:\n")
#summary(gau.sp)

sph.sp <- update(null.model, correlation = corSpher(1, form = ~ x + y), method = "ML")
writeLines("\nResults for spherical spatial correlation:\n")
summary(sph.sp)
```

```{r}
# Get the AIC for all models
#aic <- AIC(null.model, exp.sp, gau.sp, sph.sp)
aic <- AIC(null.model, exp.sp, sph.sp)
best <- rownames(aic)[which.min(aic[, "AIC"])]
best

# the best model is exp.sp
best_model <- get(best)

summary(best_model)

# Residuals are spatially autocorrelated
resids <- residuals(best_model)
M <- Moran.I(resids, spdata.IDW)

# Residual autocorrelation is somehow stronger than the null model?
null_M$observed
M$observed

dframe$resids <- resids
dframe$null_resids <- null_resids

gg_df <- dframe %>% pivot_longer(cols = contains("resids"), names_to = "resid_type", values_to = "resids")

ggplot(gg_df) + 
  geom_sf(data = ca) + 
  geom_sf(aes(col = resids)) +
  facet_wrap(~resid_type) +
  scale_color_viridis_c(option = "magma") +
  theme_void()
```

```{r}
# Reminder of the correlations
high_corr_pairs 
high_corr_pairs %>% pivot_longer(cols = c(Var1, Var2)) %>% distinct(value)
```

```{r}
spatial_dredge <- function(full_formula, data, coords = c("x", "y"), 
                          random = "~ 1 | dummy", method = "ML") {
  
  # Get all predictor variables from formula
  predictors <- all.vars(full_formula[[3]])
  
  # Create all possible combinations of predictors
  n_preds <- length(predictors)
  combinations <- lapply(1:n_preds, 
                        function(k) combn(predictors, k, simplify = FALSE))
  combinations <- unlist(combinations, recursive = FALSE)
  
  # Add null model (intercept only)
  combinations <- c(list(character(0)), combinations)
  # Initialize storage for models and their metrics
  model_results <- data.frame(
    model_formula = character(),
    AIC = numeric(),
    BIC = numeric(),
    logLik = numeric(),
    stringsAsFactors = FALSE
  )
  
  # Function to fit a model and return its metrics
  fit_model <- function(predictors) {
    tryCatch({
      # Construct formula
      if(length(predictors) == 0) {
        # Null model (intercept only)
        current_formula <- as.formula(paste("Ho ~ 1"))
      } else {
        # Model with predictors
        current_formula <- as.formula(
          paste("Ho ~", paste(predictors, collapse = " + "))
        )
      }
      
      # Fit spatial model
      current_model <- lme(fixed = current_formula,
                          data = data,
                          random = as.formula(random),
                          correlation = corExp(1, form = as.formula(
                            paste("~", paste(coords, collapse = " + "))
                          )),
                          method = method)
      
      # Return results
      return(data.frame(
        model_formula = paste(deparse(current_formula), collapse = ""),
        AIC = AIC(current_model),
        BIC = BIC(current_model),
        logLik = logLik(current_model),
        stringsAsFactors = FALSE
      ))
      
    }, error = function(e) {
      cat("Error fitting model:", deparse(current_formula), "\n")
      cat("Error message:", conditionMessage(e), "\n")
      return(NULL)
    })
  }
  
  # Run models for each combination using purrr::map
  model_results <- map_dfr(combinations, fit_model, .progress = TRUE)

  # Sort results by AIC
  model_results <- model_results %>%
    arrange(AIC)
  
  return(model_results)
}
mod_vars <- c("NDVI", "gHM", "paleo_change", "CHELSA_bio1_1981.2010_V.2.1", "csi_past", "glacier", "tmean_dif", "Q")
f <- as.formula(paste("Ho ~", paste(mod_vars, collapse = " + ")))
dredge_result <- spatial_dredge(full_formula = f, data = dframe, method = "ML")

# Look at top 3 models
dredge_result %>% head(3)

# Get best model
best_model <- dredge_result %>% filter(AIC == min(AIC))
```

```{r}
# Fit the best model
f <- as.formula(best_model$model_formula)
null.model.best <- lme(fixed = f, data = dframe, random = ~ 1 | dummy, method = "ML")
exp.sp.best <- update(null.model.best, correlation = corExp(1, form = ~ x + y), method = "ML")
summary(exp.sp.best)

```

# Variance partioning analysis
```{r}
lm_variance_partition <- function(predictor_sets, response, scale_vars = TRUE) {
  # Validate inputs
  if (length(predictor_sets) < 2) {
    stop("At least two predictor sets are required")
  }
  
  # Convert everything to data frames
  predictor_sets <- map(predictor_sets, as.data.frame)
  
  # Validate dimensions
  n_obs <- length(response)
  walk2(
    predictor_sets,
    names(predictor_sets),
    ~if (nrow(.x) != n_obs) {
      stop(sprintf("Predictor set %s has different number of observations than response", .y))
    }
  )
  
  # Scale variables if requested
  if (scale_vars) {
    response <- scale(response) %>% as.vector()
    predictor_sets <- map(predictor_sets, ~scale(.x) %>% as.data.frame())
  }
  
  # Function to calculate R2 and adjusted R2
  get_r2 <- function(model) {
    sum <- summary(model)
    c(
      r2 = sum$r.squared,
      adj_r2 = sum$adj.r.squared
    )
  }
  
  # Individual models for each predictor set
  individual_models <- predictor_sets %>%
    map(~lm(response ~ ., data = .x))
  
  # Full model with all predictors
  all_predictors <- as.data.frame(predictor_sets)
  names(all_predictors) <- names(predictor_sets)
  full_model <- lm(response ~ ., data = all_predictors)
  
  # Get R2 values for individual models
  individual_r2 <- map(individual_models, get_r2) %>%
    bind_rows(.id = "predictor_set")
  
  # Function to get all possible combinations
  get_combinations <- function(n) {
    combn(names(predictor_sets), n, simplify = FALSE)
  }
  
  # Get all combinations of predictor sets
  all_combinations <- map(
    2:length(predictor_sets),
    get_combinations
  ) %>%
    flatten()
  
  # Calculate R2 for all combinations
  combination_results <- all_combinations %>%
    set_names(map_chr(., ~paste(.x, collapse = "+"))) %>%
    map(~{
      combined_predictors <- predictor_sets[.x] %>% as.data.frame()
      names(combined_predictors) <- .x
      model <- lm(response ~ ., data = combined_predictors)
      get_r2(model)
    }) %>%
    bind_rows(.id = "combination")
  
  # Calculate unique contributions (Type III SS)
  unique_contributions <- map2_dfr(
    names(predictor_sets),
    predictor_sets,
    ~{
      # Model without this predictor set
      other_predictors <- predictor_sets[names(predictor_sets) != .x] %>% as.data.frame()
      names(other_predictors) <- names(predictor_sets)[names(predictor_sets) != .x]
      reduced_model <- lm(response ~ ., data = other_predictors)
      
      # Calculate unique contribution
      unique_r2 <- get_r2(full_model)[2] - get_r2(reduced_model)[2]
      
      tibble(
        predictor_set = .x,
        unique_contribution = unique_r2
      )
    }
  )
  
  # Calculate shared variance
  total_r2 <- get_r2(full_model)[2]
  shared_variance <- total_r2 - sum(unique_contributions$unique_contribution)
  
  # Prepare summary statistics
  summary_stats <- tibble(
    total_explained = total_r2,
    shared = shared_variance,
    unexplained = 1 - total_r2
  )
  
  
  # Print summary
  cat("\nVariance Partitioning Summary:\n")
  cat("----------------------------\n")
  ind <- 
    individual_r2 %>% 
    mutate(across(where(is.numeric), ~scales::percent(.x, accuracy = 1)))
  uniq <- 
    unique_contributions %>% 
    mutate(unique_contribution = scales::percent(unique_contribution, accuracy = 1))
  summary_table <- left_join(ind, uniq, by = "predictor_set")
  print(summary_table)

  cat("\nSummary statistics:\n")
  print(summary_stats %>% 
        mutate(across(everything(), ~scales::percent(.x, accuracy = 1))))
  
  
  # Create results list
  results <- list(
    summary_table = summary_table,
    individual_models = individual_models,
    full_model = full_model,
    individual_r2 = individual_r2,
    unique_contributions = unique_contributions,
    combination_results = combination_results,
    summary_stats = summary_stats
  )
  return(results)
}

```
```{r}
# results <- lm_variance_partition(
#   climate = climate_vars,
#   soil = soil_vars,
#   topography = topo_vars,
#   response = genetic_diversity
# )

mod_vars <- c("NDVI", "gHM", "paleo_change", "CHELSA_bio1_1981.2010_V.2.1", "csi_past", "glacier", "tmean_dif", "Q")
predictor_sets <- mod_df_imputed_scaled %>% select(all_of(mod_vars)) %>% as.list()
response <- mod_df_imputed_scaled$Ho
results <- lm_variance_partition(predictor_sets, response, scale_vars =TRUE)
```


# Comparison of x/y versus paleo change

```{r}
r <- env_data[[1]]
r[] <- 1:ncell(r)

mod_df2 <- 
  get_coords() %>% 
  left_join(get_het()) %>%
  left_join(paleoclim_mod_df)

mod_df2$r <- extract(r,  get_coords(sf = TRUE), ID = FALSE)[,1]

AIC(lm(Ho ~ x + y, data = mod_df2))
AIC(lm(Ho ~ paleo_change, data = mod_df2))
AIC(lm(Ho ~ r, data = mod_df2))
```

# Moving window
```{r heterozygosity window, fig.width = 7, fig.height = 7, eval = FALSE}
library(wingen)
# Project coordinates and boundary to equal area projection (California Albers)
coords_proj <- st_transform(mod_sf, 3310)
ca_proj <- st_transform(ca, 3310)

# Convert coordinates to a raster layer with 10km resolution
lyr <- coords_to_raster(coords_proj, buffer = 10, res = 10000)

# Calculate windowed mean of Ho
window_Ho <- window_general(coords_proj$Ho, coords = coords_proj, lyr = lyr, wdim = 9, stat = mean, na.rm = TRUE)


ggplot_gd(window_Ho, col = viridis::mako(100, direction = -1), bkg = st_transform(ca, 3310))

writeRaster(window_Ho, here("analysis", "genetic_diversity", "outputs", "Ho_window.tif"), overwrite = TRUE)
window_Ho <- rast(here("analysis", "genetic_diversity", "outputs", "Ho_window.tif"))

# Turn into a dataframe
window_df <- 
  window_Ho %>% 
  as.data.frame(xy = TRUE)

# Convert glacier to raster
glacier <- st_transform(env_list$glacier, st_crs(window_Ho))
glacier_r <- raster::rasterize(glacier, window_Ho)
glacier_r[is.na(glacier_r)] <- 0
env_list$glacier <- glacier_r

# Read in Q raster
admixture <- 1 - rast(here("data", "Scelop_bounds_K6_maxQ.tif"))
env_list$admixture <- admixture

# Extract env vals
window_env <- 
  imap(env_list, \(x, i){
    r <- terra::project(x, crs(window_Ho))
    if (nlyr(r) == 1) names(r) <- i
    vals <- terra::extract(r, window_df[,c("x", "y")], ID = FALSE) 
    data.frame(vals) 
  }, .progress = TRUE) %>% 
  bind_cols()

mod_vars <- c("BurnDate", "NDVI", "gHM", "paleo_change", "CHELSA_bio1_1981.2010_V.2.1", "csi_past", "glacier", "tmean_dif", "admixture")

window_mod_df <- 
  bind_cols(window_df, window_env) %>% 
  # drop moving window NAs
  drop_na(custom) %>%
  # drop NA for predictors
  drop_na(admixture, BurnDate, NDVI, gHM, paleo_change, CHELSA_bio1_1981.2010_V.2.1, csi_past, glacier, tmean_dif) %>%
  rename(Ho = custom)

# Count NAs in each column to confirm no NAs
colSums(is.na(window_mod_df))

# Run lm 
mod <- lm(Ho ~ BurnDate + NDVI + gHM + paleo_change + CHELSA_bio1_1981.2010_V.2.1 + csi_past + glacier + tmean_dif + admixture, data = window_mod_df, na.action = "na.fail")

# Use dredge for model selection
dredge_result <- dredge(mod)

# Select the models with delta AIC < 2
best_models <- get.models(dredge_result, subset = delta < 2)

# Only one best model
best_model <- get.models(dredge_result, subset = delta < 2)[[1]]

# Get the p-values (* for significant)
summary(best_model)

# Run spatial models
window_sf <- window_mod_df %>% st_as_sf(coords = c("x", "y"), crs = 3310)
nbdist <- 90000
nb <- dnearneigh(window_sf, d1 = 0, d2 = nbdist)
listw <- nb2listw(nb, style = "W", zero.policy = TRUE)

# OLS model:
ols_model <- lm(Ho ~ BurnDate + NDVI + gHM + paleo_change + CHELSA_bio1_1981.2010_V.2.1 + csi_past + glacier + tmean_dif + admixture, data = window_mod_df, na.action = "na.fail")

# lag model:
slm_model <- lagsarlm(Ho ~ BurnDate + NDVI + gHM + paleo_change + CHELSA_bio1_1981.2010_V.2.1 + csi_past + glacier + tmean_dif + admixture, data = window_mod_df, listw = listw, na.action = "na.omit", zero.policy = TRUE)

# error model:
sem_model <- errorsarlm(Ho ~ BurnDate + NDVI + gHM + paleo_change + CHELSA_bio1_1981.2010_V.2.1 + csi_past + glacier + tmean_dif + admixture, data = window_mod_df, listw = listw, na.action = "na.omit", zero.policy = TRUE)

# Compare AIC
writeLines(paste("AIC OLS:", round(AIC(ols_model), 2)))
writeLines(paste("AIC SEM:", round(AIC(sem_model), 2)))
writeLines(paste("AIC SLM:", round(AIC(slm_model), 2)))

# SLM is best model
model_df <- 
  slm_model %>%
  broom::tidy() %>%
  mutate(name = map_chr(term, ~make_pretty_names(.x))) %>%
  filter(name != "(Intercept)")

sigvars <- 
  model_df %>% 
  filter(p.value < 0.05) %>% 
  filter(term != "rho") %>% 
  pull(term)

summary(slm_model)

gg_df <-
 window_mod_df %>% 
 select(all_of(sigvars), Ho) %>%
 pivot_longer(-Ho) %>%
 mutate(name = map_chr(name, ~make_pretty_names(.x)))

ggplot(gg_df, aes(x = value, y = Ho)) +
  geom_point(alpha = 0.3, pch = 16) +
  facet_wrap(~name, scales = "free_x") +
  theme_classic() +
  labs(x = "Value", y = "Ho") +
  geom_smooth(method = "lm", se = TRUE) +
  stat_regline_equation(
        aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
        formula = y ~ x
      ) +
  stat_cor(label.y = 0) +
  theme(strip.background = element_blank(), strip.text.x = element_text(hjust = 0, size = 12))

```

# HTML Maps
```{r}
coords <- 
  get_coords() %>% 
  left_join(get_het()) %>% 
  select(SampleID, x, y, Ho)

write_csv(coords, here("data", "coords_ho.csv"))

library(leaflet)
library(readr)
library(here)

# Read the data
coords <- read_csv(here("data", "coords_ho.csv"))

# Create a color palette for the Ho variable
pal <- colorNumeric(palette = "viridis", domain = coords$Ho)

# Create the map
leaflet(coords) %>%
  addTiles() %>%
  addCircleMarkers(
    ~x, ~y,
    color = ~pal(Ho),
    radius = 5,
    stroke = FALSE,
    fillOpacity = 1,
    popup = ~paste0("SampleID: ", SampleID, "<br>Ho: ", Ho)
  ) %>%
  addLegend(
    "bottomright",
    pal = pal,
    values = ~Ho,
    title = "Ho",
    opacity = 1
  )

#USGS.USImageryTopo
# Create the map with a topographic base map
map <- leaflet(coords) %>%
  addProviderTiles(leaflet::providers$USGS.USImageryTopo, options = providerTileOptions(noWrap = TRUE)) %>%
  #addProviderTiles(leaflet::providers$Esri.WorldTopoMap, options = providerTileOptions(noWrap = TRUE)) %>%
  addCircleMarkers(
    ~x, ~y,
    fillColor = ~pal(Ho),
    radius = 5,
    stroke = TRUE,
    color = "black",
    fillOpacity = 1,
    weight = 1,
    popup = ~paste0("SampleID: ", SampleID, "<br>Ho: ", Ho)
  ) %>%
  addLegend(
    "bottomright",
    pal = pal,
    values = ~Ho,
    title = "Ho",
    opacity = 1
  )

# Save the map as an HTML file
htmlwidgets::saveWidget(map, "map.html")

# Open the HTML file in your web browser
browseURL("map.html")
```

# Future climate change
```{r}
future_ho <- 
  # NOTE: CSI was inverted (1 - CSI) before
  df_future %>%
  left_join(mod_df) %>%
  # REMOVE CHANNEL ISLAND
  filter(SampleID != "Scelocci_CCGPMC_MW01-3-14")

plt <- 
  ggplot(future_ho, aes(x = csi, y = Ho, col = csi)) +
  geom_point() +
  facet_wrap(~model_name, scales = "free_x", nrow = 1) +
  theme_classic() +
  labs(x = "Climate stability", y = "Ho") +
  geom_smooth(method = "lm", se = TRUE) +
  stat_cor(label.y = min(future_ho$Ho, na.rm =TRUE)) +
  scale_color_viridis_c(option = "mako") +
  theme(
    strip.background = element_blank(),
    strip.text.x = element_text(hjust = 0, size = 14),
    axis.title = element_text(size = 14))

pdf(here("analysis", "genetic_diversity", "future_plot.pdf"), width = 12, height = 3.5)
plt
dev.off()
```

# ROH models

```{r}
mod_df2 <- 
  mod_df_imputed_scaled %>% 
  left_join(get_roh()) %>%
  left_join(dplyr::select(q, SampleID, cluster), by = "SampleID") %>% 
  drop_na(cluster) %>%
  drop_na(froh0) %>%
  mutate(froh0_log = log(froh0 + 1e-6))

mod <- lm(froh0 ~ BurnDate + NDVI + paleo_change + gHM + poly(gHM, 2)  + CHELSA_bio1_1981.2010_V.2.1 + csi_past + factor(glacier) + Q + tmean_dif, data = mod_df2, na.action = na.fail)

summary(mod)

# Use dredge for model selection
dredge_result <- dredge(mod)

# Select the models with delta AIC < 2
best_models <- get.models(dredge_result, subset = delta < 2)

# Get a model average
model_avg <- model.avg(best_models)
model_avg_summary <- summary(model_avg)

# Get coefficient and p-values
coeffs <- 
  model_avg_summary[["coefmat.subset"]] %>% 
  data.frame(., predictor = rownames(.))

# Get confidence intervals
confints <-
  confint(model_avg) %>% 
  data.frame(., predictor = rownames(.))

# Get sum of weights (relative importance)
sws <- 
  sw(model_avg_summary) %>% 
  data.frame() %>%
  t() %>%
  data.frame() %>%
  pivot_longer(everything(), names_to = "predictor", values_to = "Sum of weights") %>%
  drop_na(`Sum of weights`)

# Combine everything
model_df <-
  left_join(coeffs, confints, by = "predictor")%>%
  left_join(sws, by = "predictor") %>%
  rename(p = `Pr...z..`) 

# For the model average, paleo_change, gHM, and glacier are significant
sigvars <-
  model_df %>% 
  filter(predictor != "(Intercept)") %>%
  filter(p < 0.05)
print(sigvars)

# Look at the best models
# Get the best model
best_model <- get.models(dredge_result, subset = delta < 2)[[1]]
second_best_model <- get.models(dredge_result, subset = delta < 2)[[2]]
third_best_model <- get.models(dredge_result, subset = delta < 2)[[3]]
AIC(best_model)
AIC(second_best_model)
AIC(third_best_model)

# Get the p-values (* for significant)
summary(best_model)
summary(second_best_model)
summary(third_best_model) 

# ggplot predicted vs observed for the best_model
mod_df_imputed_scaled$predict <- predict(best_model)
ggplot(mod_df_imputed_scaled, aes(x = predict, y = Ho)) +
  geom_point() +
  #geom_abline(intercept = 0, slope = 1) +d
  geom_smooth(method = "lm") +
  theme_classic() +
  labs(x = "Predicted Ho", y = "Observed Ho")

# Turn into a nice gt table
library(gt)

best_model_df <- 
  best_model %>%
  broom::tidy() %>%
  mutate(name = map_chr(term, ~make_pretty_names(.x))) %>%
  filter(name != "(Intercept)")
  
best_model_table <- 
  best_model_df %>%
  dplyr::select(-term) %>%
  rename(term = name) %>%
  gt() %>%
  tab_header(title = "Best model") %>%
  data_color(
    columns = estimate,
    method = "numeric",
    palette = "viridis",
    reverse = TRUE, domain = c(-0.00011,0)
  ) %>%
  fmt_scientific(
    columns = c("estimate", "std.error", "statistic", "p.value"),
    decimals = 2
  )
# export to html
gt::gtsave(best_model_table, filename = here("analysis", "genetic_diversity", "roh_best_model_table.html"))
```

```{r}
gg_df <-
  mod_df2 %>%
  dplyr::select(any_of(best_model_df$term), glacier, froh0, froh0_log, cluster, gHM) 
 
gHM_poly <- lm(froh0 ~ poly(gHM, 2), mod_df2)
summary(gHM_poly)
mod_df2$gHM_poly <- predict(gHM_poly)
ggplot(mod_df2) +
  geom_point(aes(x = gHM, y = froh0)) +
  geom_line(aes(x = gHM, y = gHM_poly), col = "blue") + 
  theme_classic() +
  labs(x = "gHM", y = "FROH") +
  theme(strip.background = element_blank(), strip.text.x = element_text(hjust = 0, size = 12))

gg_df2 <- 
  mod_df2 %>% 
  pivot_longer(all_of(mod_vars)) %>%
  mutate(name =  map_chr(name, ~make_pretty_names(.x)))

ggplot(gg_df2, aes(x = value, y = froh0)) +
  geom_point() +
  theme_classic() +
  labs(x = "Value", y = "FROH") +
  geom_smooth(method = "lm", se = TRUE) +
  facet_wrap(~name, scale = "free_x") +
  stat_cor(label.y = 0.14) +
  ylim(0, NA) +
  theme(strip.background = element_blank(), strip.text.x = element_text(hjust = 0, size = 12))

ggplot(gg_df2, aes(x = value, y = froh0_log)) +
  geom_point() +
  theme_classic() +
  labs(x = "Value", y = "log(FROH + 1e-6)") +
  geom_smooth(method = "lm", se = TRUE) +
  facet_wrap(~name, scale = "free_x") +
  stat_cor(label.y = 0.14) +
  #ylim(0, NA) +
  theme(strip.background = element_blank(), strip.text.x = element_text(hjust = 0, size = 12))
```

# Evaluating different change variables

```{r}
# Full model
# Note: included both regular and poly gHM to see which was the better fit using dredge
mod <- lm(Ho ~  NDVI + paleo_change + poly(paleo_change_cur, 2) + poly(paleo_change_lgm, 2) + poly(gHM, 2) + gHM + CHELSA_bio1_1981.2010_V.2.1 + csi_past + factor(glacier) + Q + tmean_dif, data = mod_df_imputed_scaled, na.action = na.fail)

summary(mod)

# Use dredge for model selection
dredge_result <- dredge(mod)

# Select the models with delta AIC < 2
best_models <- get.models(dredge_result, subset = delta < 2)

# Get a model average
model_avg <- model.avg(best_models)
model_avg_summary <- summary(model_avg)

# Get coefficient and p-values
coeffs <- 
  model_avg_summary[["coefmat.subset"]] %>% 
  data.frame(., predictor = rownames(.))

# Get confidence intervals
confints <-
  confint(model_avg) %>% 
  data.frame(., predictor = rownames(.))

# Get sum of weights (relative importance)
sws <- 
  sw(model_avg_summary) %>% 
  data.frame() %>%
  t() %>%
  data.frame() %>%
  pivot_longer(everything(), names_to = "predictor", values_to = "Sum of weights") %>%
  drop_na(`Sum of weights`)

# Combine everything
model_df <-
  left_join(coeffs, confints, by = "predictor")%>%
  left_join(sws, by = "predictor") %>%
  rename(p = `Pr...z..`) 

# For the model average, paleo_change, Q, and glacier are significant
sigvars <-
  model_df %>% 
  filter(predictor != "(Intercept)") %>%
  filter(p < 0.05)
print(sigvars)

# Look at the best models
# Get the best model
best_model <- get.models(dredge_result, subset = delta < 2)[[1]]
second_best_model <- get.models(dredge_result, subset = delta < 2)[[2]]
third_best_model <- get.models(dredge_result, subset = delta < 2)[[3]]
AIC(best_model)
AIC(second_best_model)
AIC(third_best_model)

# Get the p-values 
summary(best_model) 
summary(second_best_model)
summary(third_best_model) 

# Adj RSQ
summary(best_model)$adj.r.squared

plot(mod_df_imputed_scaled$paleo_change_lgm, mod_df_imputed_scaled$Ho)

gg_df <- 
  mod_df_imputed_scaled %>% 
  pivot_longer(c(paleo_change_lgm, paleo_change_cur, paleo_change)) %>%
  mutate(name = 
    case_when(name == "paleo_change_lgm" ~ "LGM - LIG", 
    name == "paleo_change_cur" ~ "CUR - LGM",
    name == "paleo_change" ~ "CUR - LIG")
  )

plt1 <-
  ggplot(gg_df, aes(x = value, y = Ho)) +
  geom_point() +
  facet_wrap(~name, scales = "free") +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2)) +
  geom_smooth(method = "lm", col = "red")
  
# plot paleo_change_cur across CA
gg_df <- 
  mod_sf %>% 
  mutate_at(c("paleo_change_cur", "paleo_change_lgm", "paleo_change"), ~scale(.)) %>%
  pivot_longer(c(paleo_change_cur, paleo_change_lgm, paleo_change)) %>%
  mutate(name = 
    case_when(name == "paleo_change_lgm" ~ "LGM - LIG", 
    name == "paleo_change_cur" ~ "CUR - LGM",
    name == "paleo_change" ~ "CUR - LIG")
  )
plt2 <-
  ggplot(gg_df) +
  geom_sf(data = ca) +
  geom_sf(aes(col = value)) +
  facet_wrap(~name) +
  scale_color_viridis_c(option = "magma", begin = 0.05, end = 0.95) +
  theme_void()

plot_grid(plt1, plt2, ncol = 1)
```
