

# Libraries

```{r}
library("here")
library("tidyverse")
library("ggpubr")
library("terra")
library("sf")
library("gridExtra")
library("rpaleoclim")
library("MuMIn")
library("cowplot")
library("performance")
library("gstat")
library("sp")
library("furrr")
library("MetBrewer")
library("nlme")
library("broom.mixed")
```

# Paths for outputs
```{r}
outpath <- here("analysis", "genetic_diversity", "outputs")
plotpath <- here("analysis", "genetic_diversity", "plots")
```

# Custom functions
```{r}
source(here("general_functions.R"))
source(here("analysis", "genetic_diversity", "functions_genetic_diversity.R"))
source(here("analysis", "genetic_diversity", "functions_paleovars.R"))
```

# California polygon for plotting
```{r}
ca <- get_ca()
ca_proj <- ca %>% st_transform(3310)
```

# Coordinates
```{r}
coords <- get_coords(sf = TRUE)
coords_proj <- coords %>% st_transform(crs = st_crs(ca_proj))
```

# Heterozygosity data
```{r}
# Get genome-wide heterozygosity
het <- get_het("58-Sceloporus_chr.het") 

# Make sure you have the expected number of samples
stopifnot(nrow(het) == 259)

# Count number of sites used to calculate heterozygosity
mean(het$`N(NM)`, na.rm = TRUE) # 59955192
```

# Hypothesis 1: Genetic diversity explained by paleoclimate stability

```{r}
# Get current climate data (NOTE: not using paleoclim data because they use CHELSA V1.2 instead of V2.1)
bioclim <- rast(here("data", "env", "california_chelsa_bioclim_1981-2010_V.2.1.tif"))
cur <- bioclim[["CHELSA_bio1_1981-2010_V.2.1"]]
names(cur) <- "cur"

# Get LGM climate data (NOTE: not using paleoclim data because they use CHELSA V1.2 instead of V2.1)
lgm <- rast(here("data", "env", "california_chelsa_pmip_ccsm4.tif"))
names(lgm) <- "lgm"

# Get paleoclimate data (BIO1) from rpaleoclim and convert from C/10 to C
# Using time periods back until about ~ 1 mya (when Sceloporus diverged in S. CA)
#
# code	period	bp	source
# cur	Current (1979 – 2013)		CHELSA
# lh	Late Holocene: Meghalayan	4.2-0.3 ka	Fordham  et al. 2017
# mh	Mid Holocene: Northgrippian	8.326-4.2 ka	Fordham  et al. 2017
# eh	Early Holocene: Greenlandian	11.7-8.326 ka	Fordham  et al. 2017
# yds	Pleistocene: Younger Dryas Stadial	12.9-11.7 ka	Fordham  et al. 2017
# ba	Pleistocene: Bølling-Allerød	14.7-12.9 ka	Fordham  et al. 2017
# hs1	Pleistocene: Heinrich Stadial 1	17.0-14.7 ka	Fordham  et al. 2017
# lgm	Pleistocene: Last Glacial Maximum	ca. 21 ka	CHELSA
# lig	Pleistocene: Last Interglacial	ca. 130 ka	Otto-Bliesner et al. 2006
# mis19	Pleistocene: MIS19	ca. 787 ka	Brown et al. 2018
paleovars <- get_paleovars(c("lh", "mh", "eh", "yds", "ba", "hs1", "lig", "mis19"))
paleobio1 <- map(paleovars, ~.x$bio_1) %>% rast()/10

# Stack CHELSA rasters
chelsabio1 <- list(cur, lgm)

# Resample the higher resolution data (CHELSA) to match the lower resolution (rpaleoclim)
# This is the most conservative approach to avoid over-interpreting fine-scale variation 
csi_stack <- 
  c(as.list(chelsabio1), as.list(paleobio1)) %>% 
  map(~resample(.x, paleobio1[[1]]), method = "bilinear") %>% 
  rast()

# Calculate climate stability based on the standard deviation over the stack of rasters
# Don't remove NAs, use imputation later (NA values are at some points along the coast)
# If you remove NAs you will get some outliers because some time periods have data and others do not
csi_past <- app(csi_stack, sd, na.rm = FALSE)

# Scale CSI raster from 0 to 1 and subtract from 1 to get higher values = higher stability
csi_past[] <- 1 - scales::rescale(csi_past[], to = c(0, 1), na.rm = TRUE)

# Make a data frame with the climate stability values and holocene values for each sample
paleoclim_mod_df <- 
  data.frame(
    SampleID = coords$SampleID,
    csi_past = terra::extract(csi_past, coords, ID = FALSE)[,1],
    lh = terra::extract(paleobio1[["lh"]], coords, ID = FALSE)[,1],
    mh = terra::extract(paleobio1[["mh"]], coords, ID = FALSE)[,1],
    eh = terra::extract(paleobio1[["eh"]], coords, ID = FALSE)[,1],
    lgm = terra::extract(lgm, coords, ID = FALSE)[,1],
    lig = terra::extract(paleobio1[["lig"]], coords, ID = FALSE)[,1]
  )

# Check NA values (should be <10)
print(colSums(is.na(paleoclim_mod_df)))

# Some coastal points have NA values, so use a bilinear approach to get take the average of the four nearest raster cells
paleoclim_mod_df$csi_past <- bilinear_impute(paleoclim_mod_df$csi_past, csi_past, coords)
paleoclim_mod_df$mh <- bilinear_impute(paleoclim_mod_df$mh, csi_past, coords)
paleoclim_mod_df$eh <- bilinear_impute(paleoclim_mod_df$eh, csi_past, coords)
paleoclim_mod_df$lgm <- bilinear_impute(paleoclim_mod_df$lgm, csi_past, coords)

# Re-check NA values (should be 0)
print(colSums(is.na(paleoclim_mod_df)))
```

# Hypothesis 2: Genetic diversity explained by admixture
```{r}
# Read in admixture results produced by analysis/admixture/notebook_admixture.Rmd
q <- read_csv(here("analysis", "admixture", "outputs", "Q8.csv"))
q_r <- rast(here("analysis", "admixture", "outputs", "max_admix.tif"))

q_mod_df <- 
  q %>%
  dplyr::select(SampleID, Q) %>%
  # Calculate 1 - Q to make an "admixture score" whereby higher values = more admixture
  mutate(Q = 1 - Q)
```

# Hypothesis 3: Genetic diversity explained by current environment 
```{r}
# Read in contemporary environmental layers downloaded using the scripts in data_processing/env
bio1 <- bioclim[["CHELSA_bio1_1981-2010_V.2.1"]] # mean annual temperature
ndvi <- rast(here("data", "env", "california_ndvi_mean_2000_2020.tif")) # mean NDVI
ghm <- rast(here("data", "env", "california_global_human_modification_2016.tif")) # global human modification index

# Combine into df
env_mod_df <- data.frame(
  SampleID = coords$SampleID,
  NDVI = terra::extract(ndvi, coords, ID = FALSE)[, "NDVI"],
  gHM = terra::extract(ghm, coords, ID = FALSE)[,"gHM"],
  bio1 = terra::extract(bio1, coords, ID = FALSE)[,"CHELSA_bio1_1981-2010_V.2.1"]
)

# Check NA values
print(colSums(is.na(env_mod_df)))

# Some coastal points have NA values for gHM and bio1, so use a bilinear approach to get take the average of the four nearest raster cells
# Note: This function only imputes missing values, not all values
env_mod_df$gHM <- bilinear_impute(env_mod_df$gHM, ghm, coords)
env_mod_df$bio1 <- bilinear_impute(env_mod_df$bio1, bio1, coords)

# Double check NA values (should be 0)
print(colSums(is.na(env_mod_df)))
```

# Hypothesis 4: Genetic diversity explained by glaciation
```{r}
# Read in LGM polygon provided by Michelle Koo (MVZ)
lgm_glaciers <- st_read(here("data", "env", "Glaciotect_NA_lgm", "shp", "LGM_maxextent.shp"))

# Transform to same CRS as coords
# Note: have to first intersect and then transform otherwise you get a node error
lgm_glaciers <- st_intersection(lgm_glaciers, st_transform(ca, crs = st_crs(lgm_glaciers)))
lgm_glaciers <- st_transform(lgm_glaciers, crs = st_crs(ca))

# Add a new column to coords indicating whether each point is inside the LGM glacial extent
glacier_coords <- coords %>% st_transform(st_crs(lgm_glaciers))
glacier_mod_df <- 
  coords %>%
  mutate(glacier = as.numeric(lengths(st_within(glacier_coords, lgm_glaciers)) > 0)) %>%
  st_drop_geometry()
```

# Hypothesis 5: Genetic diversity explained by recent climate change
PRISM mean temperature data from historical data: https://prism.oregonstate.edu/historical/
```{r}
# Get historical prism data (https://prism.oregonstate.edu/historical/)
# Processed by scripts in data_processing/env
get_prism <- function(year, output = "df") {
  # Create PRISM path
  prism_path <- here("data", "env", "prism", "PRISM_tmean_stable_4kmM3", year)

  # List annual file
  file <- list.files(prism_path, pattern = paste0(year, "_*.bil.bil$"), full.names = TRUE)

  # Load data
  raster <- terra::rast(file)

  if (output == "raster"){
    return(raster)
  } else {
    vals <- terra::extract(raster, st_transform(coords, st_crs(raster)), ID = FALSE)
    names(vals) <- "tmean"
    df <- data.frame(year = year, tmean = vals, SampleID = coords$SampleID)
    return(df)
  }
}

# Loop over the years of interest
historical_years <- 1895:1905 # Historical decade
current_years <- 2010:2020 # Contemporary decade

# Make dataframe of average temperature by year for each sample
tmean_his <- 
  map(
    historical_years, 
    ~get_prism(year = .x), 
    .progress =TRUE
  ) %>% 
  bind_rows()  

tmean_cur <- 
  map(
    current_years, 
    ~get_prism(year = .x), 
    .progress =TRUE
  ) %>% 
  bind_rows()  

# Calculate mean temperature for each sample over the decade
tmean_his <- 
  tmean_his %>% 
  group_by(SampleID) %>% 
  summarise(tmean_his = mean(tmean, na.rm = TRUE))

tmean_cur <- 
  tmean_cur %>% 
  group_by(SampleID) %>% 
  summarise(tmean_cur = mean(tmean, na.rm = TRUE))

# Get rasters per year and average over the decade
tmean_his_r <-
   map(historical_years, ~get_prism(year = .x, output = "raster"), .progress =TRUE) %>% 
   rast() %>% 
   mean()

tmean_cur_r <- 
  map(current_years, ~get_prism(year = .x, output = "raster"), .progress =TRUE) %>% 
  rast() %>% 
  mean()

# Calculate difference between current and historical
tmean_dif_r <- tmean_cur_r - tmean_his_r

# Crop to california and mask
tmean_dif <- crop(tmean_dif_r, ca)
tmean_dif <- mask(tmean_dif, st_transform(ca, st_crs(tmean_dif_r)))

# Calculate change
tmean_mod_df <- 
  left_join(tmean_his, tmean_cur) %>% 
  mutate(tmean_dif = tmean_cur - tmean_his) %>%
  right_join(coords) %>% 
  dplyr::select(SampleID, tmean_dif, tmean_cur)

# Check for NAs
tmean_mod_df %>% 
  summarise(across(everything(), ~sum(is.na(.))))
```

# Hypothesis 6: Genetic diversity explained by current vegetation type
```{r}
ca_proj_buffer <-  ca_proj %>% st_transform(5070) %>% st_buffer(dist = 10000) # buffer to ensure full coverage when cropping 

evt_og <- rast("analysis/wildfire/LF2016_EVT_200_CONUS/Tif/LC16_EVT_200.tif") 
evt <- crop(evt_og, ca_proj_buffer)
activeCat(evt) <- "EVT_PHYS"

# Extract EVT values in a 100 m buffer around each point
coords_buffer <- coords_proj %>% st_buffer(dist = 100)

# Get counts of each EVT class per ID
evt_vals <- 
  terra::extract(evt, coords_buffer, na.rm = FALSE, ID = TRUE) %>%
  rename(evt = EVT_PHYS)

evt_counts <-
  evt_vals  %>% 
  # Remove open water and roads
  filter(!evt %in% c("Open Water", "Developed-Roads")) %>%
  count(ID, evt, name = "n")     # counts per ID x class

# This keeps all top classes per ID (i.e., ties are retained)
evt_mod_df <-
  evt_counts %>%
  group_by(ID) %>%
  slice_max(order_by = n, n = 1, with_ties = TRUE) %>%
  ungroup()

# IDs that have ties for the top class
ties <-
  evt_mod_df %>%
  add_count(ID, name = "n_top") %>%  # how many "winners" for that ID
  filter(n_top > 1) %>%
  arrange(ID, desc(n), evt)

# For ties, take the class of the original coordinates
tie_ids <- unique(ties$ID)
tie_coords <- coords_proj[tie_ids, ]
tie_evt_vals <- 
  terra::extract(evt, tie_coords, na.rm = FALSE) %>%
  rename(evt = EVT_PHYS) %>%
  mutate(ID = tie_ids)

# Replace tied classes with the value from the original coordinates
evt_mod_df <-
  evt_mod_df %>%
  filter(!(ID %in% tie_ids)) %>%  # remove tied IDs
  bind_rows(tie_evt_vals %>% select(ID, evt)) %>%  # add back in tied IDs with original coord values
  arrange(ID) %>%
  mutate(SampleID = coords_proj$SampleID) %>%
  select(-ID) 

unique(evt_mod_df$evt)

# Replace any evt classes that are developed with just "Developed"
# Group similar vegetation types (to match groupveg categories)
evt_mod_df <-
  evt_mod_df %>%
  mutate(evt = as.character(evt)) %>%
  mutate(evt = ifelse(grepl("Developed", evt), "Developed", evt)) %>%
  # Replace exotic with corresponding native vegetation type
  mutate(evt = case_when(
    evt == "Exotic Herbaceous" ~ "Grassland",
    evt == "Exotic Tree-Shrub" ~ "Shrubland",
    TRUE ~ evt
  )) %>%
  select(-n)

# Print unique cats
unique(evt_mod_df$evt)

# Check NAs
evt_mod_df %>%
  summarise(across(everything(), ~sum(is.na(.))))
```

# Hypothesis 7: Genetic diversity explained by historical fire regime
```{r}
# Historical fire regime data
# Fire Return Interval (FRI)
ca_proj_buffer <-  ca_proj %>% st_transform(5070) %>% st_buffer(dist = 10000) # buffer to ensure full coverage when cropping 

fri_og <- rast("analysis/wildfire/LF2016_FRI_200_CONUS/Tif/LC16_FRI_200.tif") 
fri <- crop(fri_og, ca_proj_buffer)

# Percent Fire Severity (PFS) 
pfs_og <- rast("analysis/wildfire/LF2016_PFS_200_CONUS/Tif/LC16_PFS_200.tif")
pfs <- crop(pfs_og, ca_proj_buffer)

# Vegetation Departure (VDep)
vdep_og <- rast("analysis/wildfire/LF2016_VDep_200_CONUS/Tif/LC16_VDep_200.tif")
vdep <- crop(vdep_og, ca_proj_buffer)

# Fire regime group (FRG) + vegetation group (BPS)
bps_og <- rast("analysis/wildfire/LF2016_BPS_200_CONUS/Tif/LC16_BPS_200.tif")
bps <- crop(bps_og, ca_proj_buffer)

# Longest fire free interval from 1984-2022
lffi_og <- rast("analysis/wildfire/conus_fire_history_metrics_1984_2022/conus_1984_2022_LFFI.tif")
lffi <- crop(lffi_og, ca_proj_buffer)

# Recent fire frequency (FRQ) from 1984-2022
frq_og <- rast("analysis/wildfire/conus_fire_history_metrics_1984_2022/conus_1984_2022_FRQ.tif")
frq <- crop(frq_og, ca_proj_buffer)

# Activate relevant categories for each raster
activeCat(fri)<- "FRI_ALLFIR"
activeCat(vdep) <- "LABEL"
activeCat(bps) <- "GROUPVEG"

# Only using percent replacement for fire severity (high severity fire) - other PFS categories are low/moderate and together those percentages (low + moderate + high) sum to 100, so high severity captures the relevant variation (high = 100 - low + moderate)
pfs_replac <- pfs
activeCat(pfs_replac) <- "PRC_REPLAC"
```

```{r}
# Create list of historical fire variables
fire_list <- list(pfs_replac, fri)
names(fire_list) <- c("pfs_replac", "fri")

# Make buffered points
coords_buffer <- coords_proj %>% st_buffer(dist = 100)

# Get vegetation data
groupveg_vals <- terra::extract(bps, coords_buffer, na.rm = FALSE)

# Count each vegetation type within buffer
groupveg_count <- 
  groupveg_vals %>%
  filter(GROUPVEG != "Open Water") %>%  # remove open water points
  count(ID, GROUPVEG) %>%               # count points per vegetation type
  group_by(ID) 

# Get mode vegetation type per ID
groupveg_mode_with_ties <- 
  groupveg_count %>%
  group_by(ID) %>%
  slice_max(order_by = n, n = 1, with_ties = TRUE) %>%
  ungroup() %>%
  select(ID, GROUPVEG)

# Pull out tied values and replace with the coordinate value
tied_ids <- 
  groupveg_mode_with_ties %>%
  group_by(ID) %>%
  filter(n() > 1) %>%
  pull(ID) %>%
  unique()

if (length(tied_ids) > 0) {
  tied_coords <- coords_proj[tied_ids, ]
  tied_groupveg_vals <- terra::extract(bps, tied_coords, na.rm = FALSE)
  tied_groupveg_vals$ID <- tied_ids # ensure ID column is correct
  # Replace in groupveg_mode
  groupveg_mode <- 
    groupveg_mode_with_ties %>%
    filter(!(ID %in% tied_ids)) %>%
    bind_rows(tied_groupveg_vals) %>%
    arrange(ID)
}

# Confirm that number of rows matches number of coords
stopifnot(nrow(groupveg_mode) == nrow(coords_proj))

# Get unique vegetation values for each coordinate to get a vector of allowed GROUPVEG types
allowed_groupveg <- unique(terra::extract(bps, coords_proj, na.rm = FALSE)$GROUPVEG)
# Remove open water 
allowed_groupveg <- allowed_groupveg[allowed_groupveg != "Open Water"]
# Anything that is -9999 and NOT in allowed_groupveg will be coded as NA (since it is not sceloporus habitat), any other value of -9999 will get a 0 (no fire regime)
print(allowed_groupveg)

# Create fire regime groups based on FRI and PFS based on an adaptation of bins from LANDFIRE
# NEW FRG bins (Too fine scale): https://www.landfire.gov/sites/default/files/DataDictionary/LF2016/LF16_FRGADD.pdf
# Adapted based on OLD FRG bins: https://www.landfire.gov/sites/default/files/DataDictionary/LF2014/LF14_FRGADD.pdf
frg_mode_with_ties <-
  # Extract values
  extract(rast(fire_list), coords_buffer, ID = TRUE)  %>%
  # Get mode vegetation type in buffer
  mutate(groupveg = groupveg_vals$GROUPVEG) %>%
  # Convert to numeric
  mutate(pfs_replac = as.numeric(as.character(pfs_replac)), fri = as.numeric(as.character(fri))) %>%
  # Replace -9999 with NAs
  mutate(replac_clean = ifelse(pfs_replac == -9999, NA, pfs_replac), fri_clean = ifelse(fri == -9999, NA, fri)) %>%
  mutate(
    frg = 
      case_when(
        # If vegetation group is not in allowed list, code as NA
        !(groupveg %in% allowed_groupveg) ~ NA_character_,
        # If FRI is NA and vegetation type is in allowed list, code as No Fire Regime
        is.na(fri_clean) & (groupveg %in% allowed_groupveg) ~ "No Fire Regime",
        # New FRG bins
        fri_clean <= 35 & replac_clean >= 66.7 ~ "High Frequency + High Severity",
        fri_clean <= 35 & replac_clean < 66.7 ~ "High Frequency + Low/Moderate Severity",
        fri_clean > 35 & fri_clean < 200 & replac_clean >= 66.7 ~ "Intermediate Frequency + High Severity",
        fri_clean > 35 & fri_clean < 200 & replac_clean < 66.7 ~ "Intermediate Frequency + Low/Moderate Severity",
        fri_clean >= 200 ~ "Infrequent Fire Regime",
        TRUE ~ NA_character_
      )
  ) %>%
  # Remove NA values before counting
  drop_na(frg) %>%
  # Count each fire regime group per ID
  count(ID, frg, name = "n") %>%
  group_by(ID) %>%
  # Get mode fire regime group per ID (keeping ties)
  slice_max(order_by = n, n = 1, with_ties = TRUE) %>%
  ungroup()

# Identify IDs that have ties
frg_ties <-
  frg_mode_with_ties %>%
  add_count(ID, name = "n_top") %>%   # number of "winners" for that ID
  filter(n_top > 1) %>%
  arrange(ID, desc(n), frg)

tie_ids <- unique(frg_ties$ID)

# For tied IDs, use the original (unbuffered) coordinate value as the tiebreaker
if (length(tie_ids) > 0) {
  tie_coords <- coords_proj[tie_ids, ]

  tie_frg_vals <-
    terra::extract(rast(fire_list), tie_coords, na.rm = FALSE) %>%
    mutate(pfs_replac = as.numeric(as.character(pfs_replac)), fri = as.numeric(as.character(fri))) %>%
    mutate(replac_clean = ifelse(pfs_replac == -9999, NA, pfs_replac), fri_clean = ifelse(fri == -9999, NA, fri)) %>%
    # No vegetation check here because these are all in allowed vegetation types since they were selected from the main coords
    mutate(
      frg = 
        case_when(
          is.na(fri_clean) ~ "No Fire Regime",
          fri_clean <= 35 & replac_clean >= 66.7 ~ "High Frequency + High Severity",
          fri_clean <= 35 & replac_clean < 66.7 ~ "High Frequency + Low/Moderate Severity",
          fri_clean > 35 & fri_clean < 200 & replac_clean >= 66.7 ~ "Intermediate Frequency + High Severity",
          fri_clean > 35 & fri_clean < 200 & replac_clean < 66.7 ~ "Intermediate Frequency + Low/Moderate Severity",
          fri_clean >= 200 ~ "Infrequent Fire Regime",
          TRUE ~ NA_character_
        ),
      ID = tie_ids
    ) %>%
    select(ID, frg)

  # Replace tied IDs in the buffer-mode result with point-based category
  frg_mode_resolved <-
    frg_mode_with_ties %>%
    filter(!(ID %in% tie_ids)) %>%          # drop tied IDs
    bind_rows(tie_frg_vals) %>%             # add back resolved values
    arrange(ID)
} else {
  frg_mode_resolved <- frg_mode_with_ties %>% arrange(ID)
}

stopifnot(nrow(frg_mode_resolved) == nrow(coords_proj))  # confirm number of rows matches number of coords

# Create dataframe with SampleID and fire regime group
frg_df <- 
  frg_mode_resolved %>%
  mutate(SampleID = coords_proj$SampleID) %>%
  select(SampleID, frg)

# Get vegetation departure data
vdep_vals <-
  terra::extract(vdep, coords_buffer) %>%
  rename(vdep = LABEL) %>%
  # Convert vdep to fully numeric by coding categories (urban/agriculture/etc.) as 101 (highest departure)
  # A warning is normal here because of coercion to numeric
  mutate(
    vdep = as.numeric(as.character(vdep)),
    vdep = ifelse(is.na(vdep), 101, vdep)
  ) %>%
  group_by(ID) %>%
  summarise(vdep = median(vdep, na.rm = TRUE))

# Get recent fire data
# (Not same scale as evt or vegetation group, so just take median over buffer)
contemporary_fire_vals <-
  terra::extract(c(lffi, frq), coords_buffer, na.rm = FALSE, ID = TRUE) %>%
  rename(lffi = conus_1984_2022_LFFI, frq = conus_1984_2022_FRQ) %>%
  group_by(ID) %>%
  # Calculate LFFI in the buffer, ONLY using the pixels that experienced fire (i.e., not NA values)
  # Why: we want to know the fire free interval where fire did occur, not the average over the whole buffer which would be biased towards unburned areas
  # If you do it the other way (i.e., turn NAs into 39 first) you get biased high values
  # Either way, if you remove all the points whre lffi is NA, there is a significant relationship between lffi and Ho, so I am comfortable with this approach
  summarise(lffi = median(lffi, na.rm = TRUE)) %>%
  # Any remaining NA values are the result of no fire occurring in the buffer during the 39 year period so replace NA with 39 (interpretation: longest fire free interval from 1984-2022 is 39 years if no fires occurred)
  mutate(lffi = ifelse(is.na(lffi), 39, lffi))

# Confirm no NAs
stopifnot(all(!is.na(lffi_df$lffi)))

# Combine into a single dataframe
fire_mod_df <- 
  # Combine data frames
  groupveg_mode %>% 
  rename(groupveg = GROUPVEG) %>%          
  left_join(vdep_vals, by = "ID") %>%
  left_join(contemporary_fire_vals, by = "ID") %>%
  select(-ID) %>%
  # Add SampleIDs
  mutate(SampleID = coords_proj$SampleID) %>%
  # Add contemporary vegetation group (buffered)
  left_join(evt_mod_df, by = "SampleID") %>%
  # Add frg
  left_join(frg_df, by = "SampleID")  %>%
  # Create factors 
  mutate(
    frg = factor(frg, levels = rev(c("High Frequency + High Severity", "High Frequency + Low/Moderate Severity", "Intermediate Frequency + High Severity", "Intermediate Frequency + Low/Moderate Severity", "Infrequent Fire Regime", "No Fire Regime"))),
    groupveg = as.character(groupveg),
  ) 
```

# Combine rasters into a raster stack
This can take a while because of the fire rasters
```r
env_list <- 
  list(
    NDVI = ndvi,
    bio1 = bio1,
    gHM = ghm,
    tmean_dif = tmean_dif,
    csi_past = csi_past, 
    Q = q_r,
    pfs_replac = fire_list$pfs_replac,
    pfs_mixed =fire_list$pfs_mixed, 
    pfs_surfac = fire_list$pfs_surfac, 
    fri = fire_list$fri, 
    vdep = fire_list$vdep
  )

# Project all rasters to EPSG:3310 and mask with CA
env_list <- imap(env_list, ~{
  r <- terra::project(.x, "epsg:3310", method = "bilinear")
  r <- mask(r, ca_proj)
  names(r) <- .y
  return(r)
}, .progress = TRUE)

# Output path for model rasters
path <- here("data", "env", "model_rasters")

# Make model rasters dir
dir.create(path, showWarnings = FALSE)

# Write out rasters
iwalk(env_list, ~writeRaster(.x, here(path, paste0(.y, ".tif")), overwrite = TRUE))

# Snap all rasters to same grid and resolution
# Note: these rasters are only used for visualization, so these changes don't affect the models
max_res <- map_dbl(env_list, ~res(.x)[1]) %>% which.max()
target_raster <- env_list[[max_res]] # coarsest resolution
resampled_list <- imap(env_list, ~{
  # Resample to match target
  terra::resample(.x, target_raster, method = "bilinear")
})

# Create raster stack
env_stack <- 
  resampled_list %>% 
  rast() 
```

# Create model data
```{r}
# Model variables
mod_vars <- c("NDVI", "gHM", "bio1", "Q", "csi_past", "glacier", "tmean_dif", "frg", "lffi", "vdep", "groupveg", "evt", "x", "y")

# Combine dataframes
mod_df <- 
  get_coords() %>% 
  left_join(het) %>%
  left_join(env_mod_df) %>%
  left_join(paleoclim_mod_df) %>%
  left_join(q_mod_df) %>%
  left_join(glacier_mod_df) %>%
  left_join(tmean_mod_df) %>%
  left_join(fire_mod_df) %>%
  left_join(evt_mod_df) %>%
  dplyr::select(Ho, SampleID, all_of(mod_vars))

# Check sample number
# Note: het has 259 samples, but two of them are potentially flipped ("Scelocci_CAS213197", "Scelocci_CAS214858"), so we remove the from the model (they are not included in coords)
stopifnot(nrow(mod_df) == 257)

# Confirm no NAs:
stopifnot(all(colSums(is.na(mod_df)) == 0))

# Convert to sf
mod_sf <- 
  get_coords(sf = TRUE) %>% 
  right_join(mod_df) %>% 
  st_transform(3310) %>%
  select(-x, -y) %>%
  bind_cols(st_coordinates(.)) %>%
  rename(x = X, y = Y)
```

```{r}
# Export model df
export_df <- 
  st_drop_geometry(mod_df) %>% 
  dplyr::select(SampleID, everything())
write_csv(export_df, here(outpath, "model_df.csv"))
```

## Check for correlation between variables
```{r}
# Check correlations
# Compute the correlation matrix
numeric_mod_vars <-
  mod_df %>% 
  select(all_of(mod_vars), -x, -y) %>%
  select(where(is.numeric)) %>% 
  colnames()

r <- cor(mod_df[, numeric_mod_vars], use = "pairwise.complete.obs")
# Set the diagonal and lower triangle to NA
r[lower.tri(r, diag = TRUE)] <- NA
r

# Find the pairs with correlations greater than 0.6
high_corr <- which(abs(r) > 0.6, arr.ind = TRUE)
# Print the pairs
high_corr_pairs <- data.frame(
  Var1 = rownames(r)[high_corr[, 1]],
  Var2 = rownames(r)[high_corr[, 2]],
  Correlation = r[high_corr]
) %>%
arrange(desc(abs(Correlation)))

# No pairs with correlation > 0.6
high_corr_pairs
```

```{r}
#Make nice correlation plot
r <- stats::cor(mod_df[, setdiff(numeric_mod_vars, c("x", "y"))], use = "pairwise.complete.obs")
rownames(r) <- colnames(r) <- make_pretty_names(rownames(r))
r[lower.tri(r, diag = TRUE)] <- NA

library(corrplot)
png(here(plotpath, "correlation_plot.png"), width = 10*300, height = 10*300, units = "px", res = 300)
corrplot(r, method = "ellipse", type = "upper", tl.col = "black", tl.srt = 45, diag = FALSE, addCoef.col = "black", number.cex = 0.8)
dev.off()
```


```{r}
# Checking R squared for BPS vegetation group
groupveg_rsq <-  
  map(numeric_mod_vars, ~{
    mod <- lm(as.formula(paste0(.x, " ~ groupveg")), data = mod_df)
    rsq <- summary(mod)$adj.r.squared
    data.frame(variable = .x, rsq = rsq)
  }) %>%
  bind_rows() %>%
  arrange(desc(rsq))

# Filter for high R2 (>50% variance explained by vegetation group)
groupveg_rsq %>% filter(rsq > 0.5) # Only fire related variables

# Checking R squared for current vegetation group
evt_rsq  <- 
  map(numeric_mod_vars, ~{
    mod <- lm(as.formula(paste0(.x, " ~ evt")), data = mod_df)
    rsq <- summary(mod)$adj.r.squared
    data.frame(variable = .x, rsq = rsq)
  }) %>%
  bind_rows() %>%
  arrange(desc(rsq))

# Filter for high R2 (>50% variance explained by vegetation group)
evt_rsq %>% filter(rsq > 0.5) # No variables (gHM + vdep are close to 40%)

# Check correlation between groupveg and evt
mean(mod_df$groupveg == mod_df$evt)  # Only 42% match

# Checking R squared for fire categories
fire_severity_cat_rsq  <- 
  map(numeric_mod_vars, ~{
    mod <- lm(as.formula(paste0(.x, " ~ fire_severity_cat")), data = mod_df)
    rsq <- summary(mod)$adj.r.squared
    data.frame(variable = .x, rsq = rsq)
  }) %>%
  bind_rows() %>%
  arrange(desc(rsq))

# Filter for high R2
fire_severity_cat_rsq %>% filter(rsq > 0.5) # No

fire_frequency_cat_rsq  <- 
  map(numeric_mod_vars, ~{
    mod <- lm(as.formula(paste0(.x, " ~ fire_frequency_cat")), data = mod_df)
    rsq <- summary(mod)$adj.r.squared
    data.frame(variable = .x, rsq = rsq)
  }) %>%
  bind_rows() %>%
  arrange(desc(rsq))

# Filter for high R2
fire_frequency_cat_rsq %>% filter(rsq > 0.5) 
```

```{r}
# Check correlation for BPS vs EVT vegetation group 
# Just check intersection of levels
veg_levels <- expand.grid(
  unique(mod_df$groupveg),
  unique(mod_df$evt)
)

veg_corr <- pmap(veg_levels, function(Var1, Var2) {
  tibble(
    veg_class1 = Var1,
    veg_class2 = Var2,
    cor = cor(
      as.integer(mod_df$groupveg == Var1),
      as.integer(mod_df$evt == Var2),
      method = "pearson"
    )
  )
}) %>%
  bind_rows()

veg_corr %>% arrange(desc(cor))

# Check correlation between fire frequency and severity groups
fire_levels <- expand.grid(
  unique(mod_df$fire_severity_cat),
  unique(mod_df$fire_frequency_cat)
)
fire_corr <- pmap(fire_levels, function(Var1, Var2) {
  tibble(
    fire_severity_cat = Var1,
    fire_frequency_cat = Var2,
    cor = cor(
      as.integer(mod_df$fire_severity_cat == Var1),
      as.integer(mod_df$fire_frequency_cat == Var2),
      method = "pearson"
    )
  )
}) %>%
  bind_rows()

fire_corr %>% arrange(desc(abs(cor))) # High correlations
```

```{r}
pdf(here(plotpath, "vegetation_plots.pdf"))
ggplot(mod_df, aes(x = groupveg, y = fire_frequency)) +
  geom_boxplot() +
  theme_minimal() 
ggplot(mod_df, aes(x = groupveg, y = Ho)) + 
  geom_boxplot() +
  theme_minimal() 
ggplot(mod_df, aes(x = fire_frequency, y = Ho)) +
  geom_point(aes(col = groupveg)) +
  geom_smooth(method = "lm") +
  theme_minimal() 
ggplot(mod_df, aes(x = fire_severity, y = Ho)) +
  geom_point(aes(col = groupveg)) +
  geom_smooth(method = "lm") +
  theme_minimal() 
dev.off()
```

# Run Ho models

```{r}
# Model formula variables
mod_vars <- c("glacier", "NDVI", "gHM", "bio1", "csi_past", "tmean_dif",  "fire_severity_cat", "fire_frequency_cat", "fire_severity_recent", "fire_frequency_recent", "vdep", "groupveg", "evt", "lgm", "lig", "eh", "mh", "lh", "Q")

# GLS can't handle 0 distances, so take the average of the coordinates (mode for veg type)
mode_val <- function(x) {
  ux <- na.omit(x)
  ux[which.max(tabulate(match(ux, unique(ux))))]
}
cat_mod_vars <- mod_df %>% 
  select(all_of(mod_vars)) %>%
  select(where(~is.character(.x) | is.factor(.x))) %>%
  colnames()
numeric_mod_vars <- setdiff(mod_vars, cat_mod_vars)

mod_sf_unique <- 
  mod_sf %>%
  group_by(x, y) %>%
  summarise(
    across(all_of(c(numeric_mod_vars, "Ho")), ~ mean(.x, na.rm = TRUE)),
    across(all_of(cat_mod_vars), ~ mode_val(.x)),
    .groups = "drop"
  )

# Scale data
mod_sf_unique_scaled <-
  mod_sf_unique %>%
  mutate(
    across(
      all_of(setdiff(numeric_mod_vars, "glacier")),
      ~ as.numeric(scale(.x))   # <- drop matrix
    )
  )

# Confirm that all glacier values are 0 and 1
stopifnot(all(mod_sf_unique_scaled$glacier %in% c(0, 1)))

# Convert glacier to factor
mod_sf_unique_scaled$glacier_factor <- as.factor(mod_sf_unique$glacier)
mod_sf_unique$glacier_factor <- as.factor(mod_sf_unique$glacier)

# Set reference levels for groupveg and evt and fire
mod_sf_unique_scaled$groupveg <- relevel(as.factor(mod_sf_unique_scaled$groupveg), ref = "Shrubland")
mod_sf_unique_scaled$evt <- relevel(as.factor(mod_sf_unique_scaled$evt), ref = "Shrubland")

# If fire severity category and frequency category are included, set reference levels
if ("fire_severity_cat" %in% colnames(mod_sf_unique_scaled)) mod_sf_unique_scaled$fire_severity_cat <- relevel(as.factor(mod_sf_unique_scaled$fire_severity_cat), ref = "No Fire Regime")
if ("fire_frequency_cat" %in% colnames(mod_sf_unique_scaled)) mod_sf_unique_scaled$fire_frequency_cat <- relevel(as.factor(mod_sf_unique_scaled$fire_frequency_cat), ref = "No Fire Regime")

# Confirm no NA
stopifnot(all(colSums(is.na(mod_sf_unique_scaled)) == 0))
```

```{r}
f <- formula("Ho ~ glacier + gHM + bio1 + csi_past + tmean_dif + fire_frequency_cat + fire_severity_recent + vdep + groupveg + evt")

# Run OLS and get residuals
ols_model <- lm(f, data = mod_sf_unique)
summary(ols_model)
AIC(ols_model)
mod_sf_unique$ols_resids <- residuals(ols_model)
broom::tidy(ols_model) %>% filter(p.value < 0.05) %>% arrange(desc(abs(estimate)))

# Fit variograms to the residuals
ols_variogram <- variogram(ols_resids ~ 1, data = mod_sf_unique)
ols_vgm <- map(c("Exp", "Sph", "Gau"), ~fit.variogram(ols_variogram, model = vgm(.x)))
names(ols_vgm) <- c("Exp", "Sph", "Gau")

pdf(here(plotpath, "ols_variogram.pdf"))
plot(ols_variogram, ols_vgm[["Exp"]], main = "OLS Residuals Variogram (Exponential)")
plot(ols_variogram, ols_vgm[["Sph"]], main = "OLS Residuals Variogram (Spherical)")
plot(ols_variogram, ols_vgm[["Gau"]], main = "OLS Residuals Variogram (Gaussian)")
dev.off()
# Based on visual inspection, the exponential model seems to fit best

# Fit a GLS model
# Note: nugget=FALSE assumes residuals at the same location are perfectly correlated, which should align with heterozygosity, since you don't expect to have much microscale variation in heterozygosity.
# nugget=TRUE allows unstructured microscale variance (random noise at zero distance), which can absorb spatial signal and cause predictor effects to appear insignificant. 
gls_model <- gls(f, correlation = corExp(form = ~ x + y, nugget = FALSE), data = mod_sf_unique, method = "ML") # Fit with ML to compare AIC later
summary(gls_model)
# Use type = "normalized" to get residuals decorrelated and standardized by the fitted 
# spatial covariance matrix; this removes modeled spatial structure and makes them 
# suitable for Moran's I. If you don't do this Moran's I will appear higher than it actually is.
mod_sf_unique$gls_resids <- residuals(gls_model, type = "normalized")

# Plot residuals
plot_resids <- function(resids){
  # Remove max residual because it is hard to see spatial patterns with it
  ggplot(mod_sf_unique[mod_sf_unique[[resids]] != max(mod_sf_unique[[resids]]), ]) +
    geom_sf(data = ca_proj, col = NA) +
    geom_sf(aes(col = .data[[resids]])) +
    theme_void() +
    scale_color_viridis_c(option = "turbo", end = 0.1, begin = 0.9, direction = -1) +
    labs(title = paste("Residuals:", resids))
}

pdf(here(plotpath, "all_model_residuals.pdf"), width = 9, height = 4)
plt_list <- list(
  plot_resids("ols_resids"),
  plot_resids("gls_resids")
)
plot_grid(plotlist = plt_list, nrow = 1)
dev.off()
# The GLS model residuals look better than the OLS residuals

# Compare AIC
AIC(ols_model, gls_model) 

# Compare significant variables (p < 0.05) for each model using map
models <- list(OLS = ols_model, GLS = gls_model)
sig_vars_by_model <- map(models, ~ broom::tidy(.x) %>%
  filter(!term %in% c("lambda", "rho", "(Intercept)"), p.value < 0.05) %>%
  pull(term))
print(sig_vars_by_model) 
```

```{r}
# Test for spatial autocorrelation in residuals using Moran's I
library(spdep)

# Coordinates matrix
coords <- cbind(mod_sf_unique$x, mod_sf_unique$y)

# 8 nearest neighbors (robust for irregular sampling)
knn_8 <- knearneigh(coords, k = 8)
nb_8  <- knn2nb(knn_8)
lw_8  <- nb2listw(nb_8, style = "W")

# Moran's I for OLS residuals
moran_ols <- moran.test(mod_sf_unique$ols_resids, lw_8)

# Moran's I for GLS normalized residuals
moran_gls <- moran.test(mod_sf_unique$gls_resids, lw_8)

print(moran_ols)
print(moran_gls)
```

# Test model variable combinations
```{r}
library(MuMIn)
# Set NA action to "na.fail" to ensure models with missing data are not considered
options(na.action = "na.fail")

# Test all model combinations
dredge_result <- dredge(ols_model)
# Get top models based on delta AICc < 2
top_models <- get.models(dredge_result, subset = delta < 2)
# Get variables for each of top models
top_vars <- 
  top_models %>%
  map(~broom::tidy(.x) %>%
        filter(!term %in% c("(Intercept)")) %>%
        pull(term))

print(top_vars[[1]])
```

```{r}
# Global formula
f_full <- Ho ~ glacier + NDVI + gHM + bio1 + csi_past + tmean_dif + fire_severity + fire_frequency + vdep + groupveg + evt

# Global GLS model (use ML for AIC-based selection)
gls_full <- gls(
  f_full,
  data = mod_sf_unique_scaled,
  method = "ML",
  correlation = corExp(form = ~ x + y, nugget = FALSE)
)

# All subsets model selection on the GLS
dredge_res <- dredge(gls_full, trace = TRUE)

# Look at the top of the table
head(dredge_res)

# Get top models based on delta AICc < 2
top_gls_models <- get.models(dredge_res, subset = delta < 2)

# Get variables used in each of the top models (excluding intercept)
top_vars_gls <- 
  top_gls_models %>%
  map(~ broom::tidy(.x) %>%
        filter(term != "(Intercept)") %>%
        pull(term))

# Print variables in top models
top_vars_gls
top_vars_gls[[1]]

# Get best model
best_gls <- get.models(dredge_res, subset = 1)[[1]]
summary(best_gls)

# Refit best model with REML
final_gls <- update(best_gls, method = "REML")
final_gls %>% broom::tidy() %>% mutate(p.value = round(p.value, 3))
```

# Fit final model 
```{r}
mod_sf_unique_scaled$fire_recent <- exp(mod_sf_unique_scaled$fire_severity_recent)
f <- formula("Ho ~ bio1 + csi_past + tmean_dif + glacier + gHM + fire_severity + fire_severity_recent + vdep + evt")
lm(f, data = mod_df) %>% summary()
lm(f, data = mod_df) %>% car::vif()

f <- formula("Ho ~ bio1 + csi_past + tmean_dif + NDVI + glacier + gHM + fire_recent + vdep + evt") # Remove groupveg to get p-value
veg0_model <- gls(f, correlation = corExp(form = ~ x + y, nugget = FALSE), data = mod_sf_unique_scaled, method = "ML")
f <- formula("Ho ~ bio1 + csi_past + tmean_dif + NDVI + glacier + gHM + fire_recent + vdep + groupveg + evt")
veg_model <- gls(f, correlation = corExp(form = ~ x + y, nugget = FALSE), data = mod_sf_unique_scaled, method = "ML")
veg_model %>% broom::tidy() %>% mutate(p.value = round(p.value, 3)) %>% filter(p.value < 0.05)
# Emmean test for groupveg
anova(veg0_model, veg_model) # Not significant difference; groupveg does not improve model
emmeans(veg_model, specs = "groupveg", mode = "appx-satterthwaite") %>%
  contrast(method = "tukey") %>%
  data.frame() %>% 
  filter(p.value < 0.05)# No individual contrasts are significant
emmeans(veg_model, specs = "evt", mode = "appx-satterthwaite") %>%
  contrast(method = "tukey") %>%
  data.frame() %>% 
  filter(p.value < 0.05)# No individual contrasts are significant

f <- formula("Ho ~ bio1 + csi_past + tmean_dif + NDVI + glacier + gHM + fire_recent + vdep + evt")
noveg0_model <- gls(f, correlation = corExp(form = ~ x + y, nugget = FALSE), data = mod_sf_unique_scaled, method = "ML")
f <- formula("Ho ~ bio1 + csi_past + tmean_dif + NDVI + glacier + gHM + fire_recent + fire_severity + vdep + evt")
noveg_model <- gls(f, correlation = corExp(form = ~ x + y, nugget = FALSE), data = mod_sf_unique_scaled, method = "ML")
noveg_model %>% broom::tidy() %>% mutate(p.value = round(p.value, 3)) %>% filter(p.value < 0.05)
anova(noveg0_model, noveg_model) # p = 0.04; fire severity significantly improves model

AIC(veg_model, noveg_model)
anova(noveg_model, veg_model) # No significant difference


f <- formula("Ho ~ bio1 + csi_past + tmean_dif + glacier + gHM + fire_severity_cat + fire_severity_recent + vdep + evt")

# Fit model
final_model <- gls(f, correlation = corExp(form = ~ x + y, nugget = FALSE), data = mod_sf_unique_scaled, method = "REML")
summary(final_model)

# Double check OLS
ols_final <- lm(f, data = mod_sf_unique_scaled)
summary(ols_final)

# Compare OLS vs GLS
AIC(ols_final, final_model)
final_model <- ols_final

# Calculate VIF
performance::check_collinearity(final_model) # From performance package
#  groupveg 7.82 [6.33, 9.73]     2.80      0.13     [0.10, 0.16] # Moderate collinearity

# Results
mod_result <- broom::tidy(final_model)
mod_result

# Significant variables
mod_result %>% filter(p.value < 0.05) %>% arrange(desc(abs(estimate))) 
sig_vars <- mod_result %>% filter(p.value < 0.05, !term %in% c("(Intercept)", "lambda")) %>% pull(term)
print(sig_vars)

# You can't get a traditional R2 for spatial models, but you can get an alternative R2
cor(predict(final_model), mod_sf_unique$Ho)^2 # 0.45
cor(predict(final_model), mod_sf_unique$Ho) # 0.67
```
```{r}
# ANOVA and emmeans test for groupveg
anova(final_model) # Note: fire severity is not significant here because of vegetation group confounding
# Mode = "appx-satterthwaite" for GLS models
emmeans_gls <- emmeans(final_model, specs = "groupveg", mode = "appx-satterthwaite")
tukey_gls <- contrast(emmeans_gls, method = "tukey")
tukey_gls # No individual contrasts are significant

```

```{r}
# TEMPORARY TEST FIRE VARIALBE
test_vars <- function(f){
  f <- formula(f)
  # ML to compare AIC
  mod <- gls(f, correlation = corExp(form = ~ x + y, nugget = FALSE), data = mod_sf_unique_scaled, method = "ML")
  mod_result <- broom::tidy(mod)
  print(AIC(mod))
  print(cor(predict(mod), mod_sf_unique$Ho)^2)
  mod_result %>% select(term, estimate, p.value) %>% mutate(p.value = round(p.value, 3))
}

test_vars("Ho ~ fire_pc + groupveg")
test_vars("Ho ~ fire_severity + groupveg")
test_vars("Ho ~ fire_frequency + groupveg")


test_vars("Ho ~ groupveg + csi_past + bio1 + tmean_dif + glacier + gHM + NDVI + Q") # NO FIRE
test_vars("Ho ~ fire_pc + groupveg + csi_past + bio1 + tmean_dif + glacier + gHM + NDVI + Q")
test_vars("Ho ~ fire_severity + groupveg + csi_past + bio1 + tmean_dif + glacier + gHM + NDVI + Q")
test_vars("Ho ~ fire_frequency + groupveg + csi_past + bio1 + tmean_dif + glacier + gHM + NDVI + Q")


test_vars("Ho ~ groupveg + csi_past + bio1 + tmean_dif + glacier + gHM + NDVI") # NO FIRE
test_vars("Ho ~ fire_pc + groupveg + csi_past + bio1 + tmean_dif + glacier + gHM + NDVI")
test_vars("Ho ~ fire_severity + groupveg + csi_past + bio1 + tmean_dif + glacier + gHM + NDVI")
test_vars("Ho ~ fire_frequency + groupveg + csi_past + bio1 + tmean_dif + glacier + gHM + NDVI")

# Test with OLS
test_vars <- function(f){
  f <- formula(f)
  mod <- lm(f, data = mod_sf_unique_scaled)
  mod_result <- broom::tidy(mod)
  print(AIC(mod))
  print(cor(predict(mod), mod_sf_unique$Ho)^2)
  mod_result %>% select(term, estimate, p.value) %>% mutate(p.value = round(p.value, 3))
}

test_vars("Ho ~ fire_pc + groupveg")
test_vars("Ho ~ fire_severity + groupveg")
test_vars("Ho ~ fire_frequency + groupveg")

test_vars("Ho ~ fire_pc + groupveg + csi_past + bio1 + tmean_dif + glacier + gHM + NDVI + Q")
test_vars("Ho ~ fire_severity + groupveg + csi_past + bio1 + tmean_dif + glacier + gHM + NDVI + Q")
test_vars("Ho ~ fire_frequency + groupveg + csi_past + bio1 + tmean_dif + glacier + gHM + NDVI + Q")

test_vars("Ho ~ fire_pc + groupveg + csi_past + bio1 + tmean_dif + glacier + gHM + NDVI")
test_vars("Ho ~ fire_severity + groupveg + csi_past + bio1 + tmean_dif + glacier + gHM + NDVI")
test_vars("Ho ~ fire_frequency + groupveg + csi_past + bio1 + tmean_dif + glacier + gHM + NDVI")
```

```{r}
q_f <- formula("Q ~ fire_frequency_cat + groupveg + csi_past + bio1 + tmean_dif + glacier + gHM + NDVI + evt + fire_severity_recent")

# OLS
q_mod_ols <- lm(q_f, data = mod_df)
summary(q_mod_ols)
broom::tidy(q_mod_ols) %>% select(term, estimate, p.value)
ols_resids <- residuals(q_mod_ols)

# Fit and plot semivariogram
q_variogram <- variogram(ols_resids ~ 1, data = mod_sf_unique_scaled)
q_vgm <- map(c("Exp", "Sph", "Gau"), ~fit.variogram(q_variogram, model = vgm(.x)))
names(q_vgm) <- c("Exp", "Sph", "Gau")
pdf(here(plotpath, "q_ols_variogram.pdf"))
plot(q_variogram, q_vgm[["Exp"]], main = "Q OLS Residuals Variogram (Exponential)")
plot(q_variogram, q_vgm[["Sph"]], main = "Q OLS Residuals Variogram (Spherical)")
plot(q_variogram, q_vgm[["Gau"]], main = "Q OLS Residuals Variogram (Gaussian)")
dev.off()

# Coordinates matrix
coords <- cbind(mod_sf_unique$x, mod_sf_unique$y)

# 8 nearest neighbors (robust for irregular sampling)
knn_8 <- knearneigh(coords, k = 8)
nb_8  <- knn2nb(knn_8)
lw_8  <- nb2listw(nb_8, style = "W")

# Moran's I for OLS residuals
moran_ols <- moran.test(residuals(q_mod_ols), lw_8)
moran_ols

# Fit GLS
q_mod_gls <- gls(q_f, correlation = corSpher(form = ~ x + y, nugget = FALSE), data = mod_sf_unique_scaled, method = "REML")
broom::tidy(q_mod_gls) %>% select(term, estimate, p.value)

# Moran's I for GLS normalized residuals
moran_gls <- moran.test(residuals(q_mod_gls), lw_8)

print(moran_ols)
print(moran_gls)
```

# Figures
```{r}
# Define a color scale function using MetBrewer with an argument to extend the scale for better visualization
scale_met_cont <- function(values, aesthetic = c("fill", "colour"), extend_scale = 0.1, r_min  = NULL, r_max = NULL) {
  aesthetic <- match.arg(aesthetic)
  if (is.null(r_min)) r_min <- min(values, na.rm = TRUE)
  if (is.null(r_max)) r_max <- max(values, na.rm = TRUE)
  rng   <- r_max - r_min
  lims  <- c(r_min - rng*extend_scale, r_max + rng*extend_scale)

  pal <- rev(MetBrewer::met.brewer("Hiroshige", type = "continuous"))
  if (aesthetic == "fill") {
    ggplot2::scale_fill_gradientn(colors = pal, limits = lims, oob = scales::squish)
  } else {
    ggplot2::scale_colour_gradientn(colors = pal, limits = lims, oob = scales::squish)
  }
}

# Make residual plots for model
plot_mod <- function(x, y, df, model, 
                     mod_vars = c("NDVI","gHM","bio1","csi_past","glacier","tmean_dif", "fire_severity", "vdep", "evt"),
                     color_var = "x", residual = TRUE) {

  # Calculate residuals for y variable given x variable
  if (residual) df[[y]] <- residuals(model) + coef(model)[[x]] * as.numeric(scale(df[[x]]))

  # Choose whether to color by x or y variable
  col_var <- if (color_var == "x") x else y

  # Extract beta and p-value for x variable
  model_summary <- broom::tidy(model)
  # Convert to scientific notation for plotting
  beta <- formatC(model_summary$estimate[model_summary$term == x], format = "e", digits = 2)
  pval <- model_summary$p.value[model_summary$term == x]
  pval <- ifelse(pval < 0.001, "< 0.001", paste0("= ", round(pval, 3)))

  plt <-
    ggplot(df, aes(x = .data[[x]], y = .data[[y]])) +
    geom_point(aes(col = .data[[col_var]]), size = 2) +
    geom_smooth(method = "lm", colour = "black") +
    scale_met_cont(df[[col_var]], aesthetic = "colour", extend_scale = 0.1) +
    annotate("text", x = Inf, y = Inf, hjust = 1.1, vjust = 1.5,
             label = paste0("β = ", beta, "\n", "p ", pval), size = 6) +
    xlab(stringr::str_to_sentence(make_pretty_names(x))) +
    theme_classic() +
    theme(legend.position = "none",
          axis.title = element_text(size = 16),
          axis.text  = element_text(size = 14))

  if (residual) {
    plt + labs(y = "Partial heterozygosity")
  } else {
    plt + labs(y = "Observed heterozygosity")
  }

}

# Load polygons for plotting
ca_proj <- get_ca() %>% sf::st_transform(3310) # California boundaries
range_map <- get_range() # Sceloporus range map

# Loaded wingen map
wingen_r <- 
  terra::rast(here(outpath, "Ho_kriged_masked.tif")) %>% 
  mask(range_map) %>% 
  mask(ca_proj) %>%
  as.data.frame(xy = TRUE, na.rm = TRUE) 
names(wingen_r)[3] <- "layer"  # ensure 'layer' is the raster value column

# Isolate significant vs all variables
sig_vars <- c("csi_past", "tmean_dif", "bio1", "fire_severity")
mod_vars <- c("csi_past", "tmean_dif", "bio1", "glacier",  "gHM", "fire_severity", "fire_severity_recent", "vdep", "evt")

# A: wingen map of heterozygosity 
ho_plot <- ggplot(wingen_r) +
  geom_sf(data = ca_proj, fill = "gray90", colour = NA, size = 2) +
  geom_tile(aes(x = x, y = y, fill = layer), colour = NA) +
  scale_met_cont(wingen_r$layer, aesthetic = "fill", extend_scale = 0.1) +
  labs(fill = "Heterozygosity") +
  theme_void() +
  geom_sf(data = mod_sf_unique, shape = 21, aes(fill = Ho), colour = "#00000087", size = 2) +
  theme(
    plot.margin = margin(1, 1, 1, 1, unit = "mm"),
    legend.position = "inside",              # turn on "inside" mode
    legend.justification = c(1, 1),          # anchor legend's top-right corner
    legend.position.inside = c(0.9, 0.95),   # place near panel's top-right
    legend.key.height = unit(0.9, "cm"),
    legend.key.width  = unit(0.9, "cm"),
    legend.title = element_text(size = 16),
    legend.text  = element_text(size = 14)
  )

# B: partial residuals 
plts_residual_y <- purrr::map(
  sig_vars, 
  # Use unscaled data for plotting
  ~plot_mod(.x, "Ho", mod_sf_unique, model = final_model, color_var = "y", residual = TRUE)
)
# B: Raw values
plts_raw_y <- purrr::map(
  sig_vars, 
  # Use unscaled data for plotting
  ~plot_mod(.x, "Ho", mod_sf_unique, model = final_model, color_var = "y", residual = FALSE)
)

# C: predicted vs observed 
pred_vs_obs <-
  # Use unscaled data for plotting
  ggplot(mod_sf_unique, aes(x = predict(final_model), y = Ho)) +
  geom_point(aes(col = Ho), shape = 16, size = 2) +
  geom_smooth(method = "lm", colour = "black") +
  scale_met_cont(c(mod_sf_unique$Ho, wingen_r$layer), aesthetic = "colour", extend_scale = 0.1) +
  theme_classic() +
  labs(x = "Predicted heterozygosity", y = "Observed heterozygosity") +
  ggpubr::stat_cor(size = 6) +
  theme(legend.position = "none",
        axis.title = element_text(size = 16),
        axis.text  = element_text(size = 14))

pred_vs_obs_adj <- cowplot::plot_grid(pred_vs_obs, NULL, nrow = 1, rel_widths = c(0.9, 0.1))

plt2_residual <- cowplot::plot_grid(plotlist = plts_residual_y, ncol = 2)
plt2_raw <- cowplot::plot_grid(plotlist = plts_raw_y, ncol = 2)
plt1 <- cowplot::plot_grid(plotlist = list(ho_plot, pred_vs_obs_adj), ncol = 1,
                           labels = c("A", "C"), rel_heights = c(0.6, 0.45), label_size = 18)

png(here(plotpath, "figure_Ho_residual.png"), height = 10.5, width = 16, res = 300, units = "in")
cowplot::plot_grid(plt1, plt2_residual, ncol = 2, labels = c("", "B"),  rel_widths = c(0.33, 0.66), align = "h", label_size = 18)
dev.off()

png(here(plotpath, "figure_Ho_raw.png"), height = 10, width = 16, res = 300, units = "in")
cowplot::plot_grid(plt1, plt2_raw, ncol = 2,labels = c("", "B"),  rel_widths = c(0.33, 0.66), align = "h", label_size = 18)
dev.off()
```


```{r}
pop_df <- 
  get_pops() %>%
  left_join(get_coords(sf = TRUE)) %>% 
  st_as_sf() %>%
  st_transform(3310) %>%
  bind_cols(st_coordinates(.)) %>%
  st_drop_geometry() %>%
  group_by(X, Y) %>%
  # Take the mean but it doesn't matter since all samples in a pop have the same cluster
  summarise(cluster = mean(as.numeric(cluster))) %>%
  mutate(cluster = factor(cluster)) %>%
  rename(x = X, y = Y) %>%
  ungroup() %>%
  right_join(mod_sf_unique) %>%
  st_as_sf()

# Color scales
group1 <- c("8", "6", "2")
group2 <- c("4", "3", "7")
group3 <- c("1", "5")
hiroshige_cols <- rev(c("#042346", "#295580", "#6bb2cc", "#9cd3dd", "#fecd6d", "#f7a957", "#ec7b4c", "#f04f32"))
cs1 <- hiroshige_cols[1:3]
names(cs1) <- group1
cs2 <- hiroshige_cols[4:6]
names(cs2) <- group2
cs3 <- hiroshige_cols[7:8]
names(cs3) <- group3
pop_cols <- c(cs1, cs2, cs3)

# Make residual plots for model
plot_mod_pop <- function(x, y, df, model, 
                     mod_vars = c("NDVI","gHM","bio1","csi_past","glacier","tmean_dif", "pfs_replac", "pfs_mixed", "pfs_surfac", "fri", "vdep","Q"), residual = TRUE) {

  # Calculate residuals for y variable given x variable
  if (residual) df[[y]] <- residuals(model) + coef(model)[[x]] * as.numeric(scale(df[[x]]))

  # Extract beta and p-value for x variable
  model_summary <- broom::tidy(model)
  # Convert to scientific notation for plotting
  beta <- formatC(model_summary$estimate[model_summary$term == x], format = "e", digits = 2)
  pval <- model_summary$p.value[model_summary$term == x]
  pval <- ifelse(pval < 0.001, "< 0.001", paste0("= ", round(pval, 3)))

  plt <-
    ggplot(df, aes(x = .data[[x]], y = .data[[y]])) +
    geom_point(aes(col = cluster), size = 2, alpha = 0.8) +
    geom_smooth(method = "lm", colour = "black") +
    scale_color_manual(values = pop_cols) +
    annotate("text", x = Inf, y = Inf, hjust = 1.1, vjust = 1.5,
             label = paste0("β = ", beta, "\n", "p ", pval), size = 6) +
    xlab(stringr::str_to_sentence(make_pretty_names(x))) +
    theme_classic() +
    theme(legend.position = "none",
          axis.title = element_text(size = 16),
          axis.text  = element_text(size = 14))

  if (residual) {
    plt + labs(y = "Partial heterozygosity")
  } else {
    plt + labs(y = "Observed heterozygosity")
  }

}

# Isolate significant vs all variables
sig_vars <- c("csi_past", "Q", "tmean_dif", "bio1", "pfs_surfac", "log_fri")
mod_vars <- c("NDVI", "gHM", "bio1", "csi_past", "glacier", "tmean_dif", "pfs_replac", "pfs_mixed", "pfs_surfac", "log_fri", "vdep", "Q")

# A: population plot
pop_label <- 
  st_drop_geometry(pop_df) %>% 
  group_by(cluster) %>%
  summarize(x = median(x), y = median(y)) %>%
  st_as_sf(coords = c("x", "y"), crs = 3310) 

pop_plot <-
  ggplot() +
  geom_sf(data = ca, col = NA) +
  geom_sf(data = pop_df, aes(fill = cluster), cex = 2, pch = 21, col = "NA")+
  #geom_point(data = pop_label, aes(x = x, y = y, fill = cluster), cex = 7, pch = 21, alpha = 0.7) +
  #geom_text(data = pop_label, aes(x = x, y = y, label = cluster), cex = 5) +
  labs(col = "Population") +
  scale_fill_manual(values = pop_cols)  +
  theme_void() +
  theme(
    legend.position = "none"
  )
# B: partial residuals 
plts_residual_y <- purrr::map(
  sig_vars, 
  # Use unscaled data for plotting
  ~plot_mod_pop(.x, "Ho", pop_df, model = final_model, residual = TRUE)
)

# B: Raw values
plts_raw_y <- purrr::map(
  sig_vars, 
  # Use unscaled data for plotting
  ~plot_mod_pop(.x, "Ho", pop_df, model = final_model, residual = FALSE)
)

# C: predicted vs observed 
pred_vs_obs <-
  # Use unscaled data for plotting
  ggplot(pop_df, aes(x = predict(final_model), y = Ho)) +
  geom_point(aes(col = cluster), shape = 16, size = 2) +
  geom_smooth(method = "lm", colour = "black") +
  scale_color_manual(values = pop_cols) +
  theme_classic() +
  labs(x = "Predicted heterozygosity", y = "Observed heterozygosity") +
  ggpubr::stat_cor(size = 6) +
  theme(legend.position = "none",
        axis.title = element_text(size = 16),
        axis.text  = element_text(size = 14))

pred_vs_obs_adj <- cowplot::plot_grid(pred_vs_obs, NULL, nrow = 1, rel_widths = c(0.9, 0.1))

plt2_residual <- cowplot::plot_grid(plotlist = plts_residual_y, ncol = 2)
plt2_raw <- cowplot::plot_grid(plotlist = plts_raw_y, ncol = 2)
plt1 <- cowplot::plot_grid(plotlist = list(pop_plot, pred_vs_obs_adj), ncol = 1,
                           labels = c("A", "C"), rel_heights = c(0.6, 0.4), label_size = 18)

png(here(plotpath, "figure_Ho_residual_pop.png"), height = 3.45*4*300*0.9, width = 4.1*4*300*0.9, res = 300)
cowplot::plot_grid(plt1, plt2_residual, ncol = 2,labels = c("", "B"),  rel_widths = c(0.4, 0.6), align = "h", label_size = 18)
dev.off()

png(here(plotpath, "figure_Ho_raw_pop.png"), height = 3.5*4*300*0.9, width = 4.1*4*300*0.9, res = 300)
cowplot::plot_grid(plt1, plt2_raw, ncol = 2,labels = c("", "B"),  rel_widths = c(0.4, 0.6), align = "h", label_size = 18)
dev.off()
```

```{r}
# Plot points on map colored by variable
sig_vars <- c("csi_past", "Q", "tmean_dif", "bio1", "pfs_surfac", "log_fri")
plot_points_map <- function(var, points_sf) {
  plt1 <- 
    ggplot() +
    geom_sf(data = ca_proj, fill = "gray90", colour = NA, size = 0.5) +
    geom_sf(data = points_sf, aes_string(col = var), size = 2) +
    scale_met_cont(points_sf[[var]], aesthetic = "colour", extend_scale = 0.1) +
    theme_void() +
    labs(title = stringr::str_to_sentence(make_pretty_names(var)), col = "") +
    theme(legend.position = "top", plot.title = element_text(hjust = 0.5))

  # Plot variable vs Ho
  plt2 <- plot_mod(var, "Ho", points_sf, model = final_model, color_var = "x", residual = FALSE) + theme(plot.title = element_blank())

  cowplot::plot_grid(plt2, plt1, nrow = 1, align = "hv")

}
png(here(plotpath, "significant_variable_maps.png"), width = 5*2, height = 5.5*length(sig_vars), res = 300, unit = "in")
plt_list <- map(sig_vars, ~(plot_points_map(.x, mod_sf_unique)))
cowplot::plot_grid(plotlist = plt_list, ncol = 1, align = "hv")
dev.off()
```

```{r}
# For each variable, build a model with all other variables and compare the residuals to the full model residuals
calc_var_resid <- function(x, mod_vars, df){
  other_vars <- setdiff(mod_vars, x)
  sub_f <- reformulate(other_vars, response = "Ho")
  f <- reformulate(mod_vars, response = "Ho")
  sub_model <- gls(sub_f, correlation = corExp(form = ~ x + y, nugget = FALSE), data = df)
  full_model <- gls(f, correlation = corExp(form = ~ x + y, nugget = FALSE), data = df)
  #sub_model <- lm(sub_f, data = df)
  #full_model <- lm(f, data = df)

  # Difference in fit
  sub_pred <- predict(sub_model)
  full_pred <- predict(full_model)
  pred_change <- full_pred - sub_pred

  return(tibble(
    pred_change = pred_change, 
    variable = x, 
    x = df$x, 
    y = df$y)
  )
}

sig_vars <- c("csi_past", "Q", "tmean_dif", "bio1", "pfs_surfac", "log_fri")

var_resid <- map(sig_vars, ~calc_var_resid(.x, mod_vars, mod_sf_unique), .progress = TRUE)

var_resid_df <- 
  var_resid %>% 
  bind_rows() %>% 
  st_as_sf(coords = c("x", "y"), crs = 3310) %>%
  group_by(variable) %>%
  mutate(pred_change_scaled = scale(pred_change)[,1]) 

pair_plot <- function(x, label = ""){
  y = "pred_change"
  change <- var_resid_df %>% filter(variable == x) %>% pull(y)
  df <- mod_sf_unique %>% mutate(change = change)
  plt_scatter <-
    ggplot(df, aes_string(x = x, y = "Ho")) +
    geom_point(aes(col = change), alpha = 0.7) +
    scico::scale_color_scico(palette = "roma", midpoint = 0, direction = -1) +
    geom_smooth(method = "lm", colour = "black") +
    theme_classic() +
    labs(x = make_pretty_names(x), y = "Observed Heterozygosity", col = y) +
    theme(legend.position = "none", plot.title = element_text(hjust = 0.5))

  plt_map <-
    ggplot(df) +
    geom_sf(data = ca_proj, fill = "white", colour = "gray", size = 0.5) +
    geom_sf(aes(col = change), pch = 16, size = 1.5, alpha = 0.9) +
    scico::scale_color_scico(palette = "roma", midpoint = 0, direction = -1) +
    theme_void() +
    labs(col = "Δ Predicted\nHeterozygosity") +
    theme(legend.position = "right", plot.title = element_text(hjust = 0.5))

  cowplot::plot_grid(plt_scatter, plt_map, nrow = 1, labels = c(label, ""), align = "h", rel_widths = c(0.45, 0.55))
}

png(here(plotpath, "variable_importance_map.png"), width = 3.5*2*2, height = 3*length(sig_vars)/2, units = "in", res = 300)
plts <- map(1:length(sig_vars), ~pair_plot(sig_vars[.x], label = LETTERS[.x]))
cowplot::plot_grid(plotlist = plts, ncol = 2)
dev.off()
```

Maps of significant variables
```{r}
plot_var <- function(x){
  ggplot() +
    geom_sf(data = ca_proj, fill = "gray90", colour = NA, size = 0.5) +
    geom_sf(data = mod_sf_unique, aes_string(fill = x), size = 2, shape = 21, color = "#00000087") +
    scale_met_cont(mod_sf_unique[[x]], aesthetic = "fill", extend_scale = 0.1) +
    theme_void() +
    labs(title = stringr::str_to_sentence(make_pretty_names(x)), fill = "") +
    theme(plot.title = element_text(hjust = 0.5))
}

sig_vars <- c("csi_past", "Q", "tmean_dif", "pfs_replac", "pfs_surfac", "log_fri")
pdf(here(plotpath, "significant_variable_maps_sf.pdf"), width = 4*3, height = 4*2)
plts <- map(sig_vars, ~plot_var(.x))
cowplot::plot_grid(plotlist = plts, nrow = 2)
dev.off()
```

# DUMP
```{r}
df = mod_df

med_fit <- lm(Ho ~ veg_PC1 + veg_PC2 + veg_PC3 + bio1 + csi_past + glacier + gHM + tmean_dif + fire_severity + fire_recent +, data = df)
summary(med_fit)
```

```{r}
pdf(here(plotpath, "fire_groupveg_plots.pdf"), width = 5, height = 5)
ggplot(data = mod_sf_unique, aes(x = fire_severity, y = Ho)) +
  geom_point(aes(col = groupveg)) +
  geom_smooth(method = "lm") +
  theme_classic() 

ggplot(data = mod_sf_unique, aes(x = fire_severity, y = Ho)) +
  geom_point(aes(col = evt)) +
  geom_smooth(method = "lm") +
  theme_classic() 

# Boxplot with significance comparisons using ggpubr
ggboxplot(mod_sf_unique, x = "groupveg", y = "Ho",
          color = "groupveg", palette = "jco") +
  stat_compare_means(method = "anova") +
  stat_compare_means(label = "p.format", method = "t.test",
                     ref.group = "shrubland", label.y.npc = 0.9) +
  theme_classic()
dev.off()
```
