

# Libraries

```{r}
library("here")
library("tidyverse")
library("ggpubr")
library("terra")
library("sf")
library("gridExtra")
library("rpaleoclim")
library("MuMIn")
library("cowplot")
library("performance")
library("gstat")
library("sp")
library("furrr")
library("MetBrewer")
library("nlme")
library("broom.mixed")
```

# Paths for outputs
```{r}
outpath <- here("analysis", "genetic_diversity", "outputs")
plotpath <- here("analysis", "genetic_diversity", "plots")
```

# Custom functions
```{r}
source(here("general_functions.R"))
source(here("analysis", "genetic_diversity", "functions_genetic_diversity.R"))
source(here("analysis", "genetic_diversity", "functions_paleovars.R"))
```

# California polygon for plotting
```{r}
ca <- get_ca()
ca_proj <- ca %>% st_transform(3310)
```

# Coordinates
```{r}
coords <- get_coords(sf = TRUE)
coords_proj <- coords %>% st_transform(crs = st_crs(ca_proj))
```

# Heterozygosity data
```{r}
# Get genome-wide heterozygosity
het <- get_het("58-Sceloporus_chr.het") 

# Make sure you have the expected number of samples
stopifnot(nrow(het) == 259)

# Count number of sites used to calculate heterozygosity
mean(het$`N(NM)`, na.rm = TRUE) # 59955192
```

# Hypothesis 1: Genetic diversity explained by paleoclimate stability

```{r}
# Get current climate data (NOTE: not using paleoclim data because they use CHELSA V1.2 instead of V2.1)
bioclim <- rast(here("data", "env", "california_chelsa_bioclim_1981-2010_V.2.1.tif"))
cur <- bioclim[["CHELSA_bio1_1981-2010_V.2.1"]]
names(cur) <- "cur"

# Get LGM climate data (NOTE: not using paleoclim data because they use CHELSA V1.2 instead of V2.1)
lgm <- rast(here("data", "env", "california_chelsa_pmip_ccsm4.tif"))
names(lgm) <- "lgm"

# Get paleoclimate data (BIO1) from rpaleoclim and convert from C/10 to C
# Using time periods back until about ~ 1 mya (when Sceloporus diverged in S. CA)
#
# code	period	bp	source
# cur	Current (1979 – 2013)		CHELSA
# lh	Late Holocene: Meghalayan	4.2-0.3 ka	Fordham  et al. 2017
# mh	Mid Holocene: Northgrippian	8.326-4.2 ka	Fordham  et al. 2017
# eh	Early Holocene: Greenlandian	11.7-8.326 ka	Fordham  et al. 2017
# yds	Pleistocene: Younger Dryas Stadial	12.9-11.7 ka	Fordham  et al. 2017
# ba	Pleistocene: Bølling-Allerød	14.7-12.9 ka	Fordham  et al. 2017
# hs1	Pleistocene: Heinrich Stadial 1	17.0-14.7 ka	Fordham  et al. 2017
# lgm	Pleistocene: Last Glacial Maximum	ca. 21 ka	CHELSA
# lig	Pleistocene: Last Interglacial	ca. 130 ka	Otto-Bliesner et al. 2006
# mis19	Pleistocene: MIS19	ca. 787 ka	Brown et al. 2018
paleovars <- get_paleovars(c("lh", "mh", "eh", "yds", "ba", "hs1", "lig", "mis19"))
paleobio1 <- map(paleovars, ~.x$bio_1) %>% rast()/10

# Stack CHELSA rasters
chelsabio1 <- list(cur, lgm)

# Resample the higher resolution data (CHELSA) to match the lower resolution (rpaleoclim)
# This is the most conservative approach to avoid over-interpreting fine-scale variation 
csi_stack <- 
  c(as.list(chelsabio1), as.list(paleobio1)) %>% 
  map(~resample(.x, paleobio1[[1]]), method = "bilinear") %>% 
  rast()

# Calculate climate stability based on the standard deviation over the stack of rasters
# Don't remove NAs, use imputation later (NA values are at some points along the coast)
# If you remove NAs you will get some outliers because some time periods have data and others do not
csi_past <- app(csi_stack, sd, na.rm = FALSE)

# Scale CSI raster from 0 to 1 and subtract from 1 to get higher values = higher stability
csi_past[] <- 1 - scales::rescale(csi_past[], to = c(0, 1), na.rm = TRUE)

# Make a data frame with the climate stability values and holocene values for each sample
paleoclim_mod_df <- 
  data.frame(
    SampleID = coords$SampleID,
    csi_past = terra::extract(csi_past, coords, ID = FALSE)[,1],
    lh = terra::extract(paleobio1[["lh"]], coords, ID = FALSE)[,1],
    mh = terra::extract(paleobio1[["mh"]], coords, ID = FALSE)[,1],
    eh = terra::extract(paleobio1[["eh"]], coords, ID = FALSE)[,1],
    lgm = terra::extract(lgm, coords, ID = FALSE)[,1],
    lig = terra::extract(paleobio1[["lig"]], coords, ID = FALSE)[,1]
  )

# Check NA values (should be <10)
print(colSums(is.na(paleoclim_mod_df)))

# Some coastal points have NA values, so use a bilinear approach to get take the average of the four nearest raster cells
paleoclim_mod_df$csi_past <- bilinear_impute(paleoclim_mod_df$csi_past, csi_past, coords)
paleoclim_mod_df$mh <- bilinear_impute(paleoclim_mod_df$mh, csi_past, coords)
paleoclim_mod_df$eh <- bilinear_impute(paleoclim_mod_df$eh, csi_past, coords)
paleoclim_mod_df$lgm <- bilinear_impute(paleoclim_mod_df$lgm, csi_past, coords)

# Re-check NA values (should be 0)
print(colSums(is.na(paleoclim_mod_df)))
```

# Hypothesis 2: Genetic diversity explained by admixture
```{r}
# Read in admixture results produced by analysis/admixture/notebook_admixture.Rmd
q <- read_csv(here("analysis", "admixture", "outputs", "Q8.csv"))
q_r <- rast(here("analysis", "admixture", "outputs", "max_admix.tif"))

q_mod_df <- 
  q %>%
  dplyr::select(SampleID, Q) %>%
  # Calculate 1 - Q to make an "admixture score" whereby higher values = more admixture
  mutate(Q = 1 - Q)
```

# Hypothesis 3: Genetic diversity explained by current environment 
```{r}
# Read in contemporary environmental layers downloaded using the scripts in data_processing/env
bio1 <- bioclim[["CHELSA_bio1_1981-2010_V.2.1"]] # mean annual temperature
ndvi <- rast(here("data", "env", "california_ndvi_mean_2000_2020.tif")) # mean NDVI
ghm <- rast(here("data", "env", "california_global_human_modification_2016.tif")) # global human modification index

# Combine into df
env_mod_df <- data.frame(
  SampleID = coords$SampleID,
  NDVI = terra::extract(ndvi, coords, ID = FALSE)[, "NDVI"],
  gHM = terra::extract(ghm, coords, ID = FALSE)[,"gHM"],
  bio1 = terra::extract(bio1, coords, ID = FALSE)[,"CHELSA_bio1_1981-2010_V.2.1"]
)

# Check NA values
print(colSums(is.na(env_mod_df)))

# Some coastal points have NA values for gHM and bio1, so use a bilinear approach to get take the average of the four nearest raster cells
# Note: This function only imputes missing values, not all values
env_mod_df$gHM <- bilinear_impute(env_mod_df$gHM, ghm, coords)
env_mod_df$bio1 <- bilinear_impute(env_mod_df$bio1, bio1, coords)

# Double check NA values (should be 0)
print(colSums(is.na(env_mod_df)))
```

# Hypothesis 4: Genetic diversity explained by glaciation
```{r}
# Read in LGM polygon provided by Michelle Koo (MVZ)
lgm_glaciers <- st_read(here("data", "env", "Glaciotect_NA_lgm", "shp", "LGM_maxextent.shp"))

# Transform to same CRS as coords
# Note: have to first intersect and then transform otherwise you get a node error
lgm_glaciers <- st_intersection(lgm_glaciers, st_transform(ca, crs = st_crs(lgm_glaciers)))
lgm_glaciers <- st_transform(lgm_glaciers, crs = st_crs(ca))

# Add a new column to coords indicating whether each point is inside the LGM glacial extent
glacier_coords <- coords %>% st_transform(st_crs(lgm_glaciers))
glacier_mod_df <- 
  coords %>%
  mutate(glacier = as.numeric(lengths(st_within(glacier_coords, lgm_glaciers)) > 0)) %>%
  mutate(glacier = factor(glacier, levels = c(0, 1), labels = c("OutsideGlacier", "InsideGlacier"))) %>%
  st_drop_geometry()
```

# Hypothesis 5: Genetic diversity explained by recent climate change
PRISM mean temperature data from historical data: https://prism.oregonstate.edu/historical/
```{r}
# Get historical prism data (https://prism.oregonstate.edu/historical/)
# Processed by scripts in data_processing/env
get_prism <- function(year, output = "df") {
  # Create PRISM path
  prism_path <- here("data", "env", "prism", "PRISM_tmean_stable_4kmM3", year)

  # List annual file
  file <- list.files(prism_path, pattern = paste0(year, "_*.bil.bil$"), full.names = TRUE)

  # Load data
  raster <- terra::rast(file)

  if (output == "raster"){
    return(raster)
  } else {
    vals <- terra::extract(raster, st_transform(coords, st_crs(raster)), ID = FALSE)
    names(vals) <- "tmean"
    df <- data.frame(year = year, tmean = vals, SampleID = coords$SampleID)
    return(df)
  }
}

# Loop over the years of interest
historical_years <- 1895:1905 # Historical decade
current_years <- 2010:2020 # Contemporary decade

# Make dataframe of average temperature by year for each sample
tmean_his <- 
  map(
    historical_years, 
    ~get_prism(year = .x), 
    .progress =TRUE
  ) %>% 
  bind_rows()  

tmean_cur <- 
  map(
    current_years, 
    ~get_prism(year = .x), 
    .progress =TRUE
  ) %>% 
  bind_rows()  

# Calculate mean temperature for each sample over the decade
tmean_his <- 
  tmean_his %>% 
  group_by(SampleID) %>% 
  summarise(tmean_his = mean(tmean, na.rm = TRUE))

tmean_cur <- 
  tmean_cur %>% 
  group_by(SampleID) %>% 
  summarise(tmean_cur = mean(tmean, na.rm = TRUE))

# Get rasters per year and average over the decade
tmean_his_r <-
   map(historical_years, ~get_prism(year = .x, output = "raster"), .progress =TRUE) %>% 
   rast() %>% 
   mean()

tmean_cur_r <- 
  map(current_years, ~get_prism(year = .x, output = "raster"), .progress =TRUE) %>% 
  rast() %>% 
  mean()

# Calculate difference between current and historical
tmean_dif_r <- tmean_cur_r - tmean_his_r

# Crop to california and mask
tmean_dif <- crop(tmean_dif_r, ca)
tmean_dif <- mask(tmean_dif, st_transform(ca, st_crs(tmean_dif_r)))

# Calculate change
tmean_mod_df <- 
  left_join(tmean_his, tmean_cur) %>% 
  mutate(tmean_dif = tmean_cur - tmean_his) %>%
  right_join(coords) %>% 
  dplyr::select(SampleID, tmean_dif, tmean_cur)

# Check for NAs
tmean_mod_df %>% 
  summarise(across(everything(), ~sum(is.na(.))))
```

# Hypothesis 6: Genetic diversity explained by current vegetation type
```{r}
ca_proj_buffer <-  ca_proj %>% st_transform(5070) %>% st_buffer(dist = 10000) # buffer to ensure full coverage when cropping 

evt_og <- rast("analysis/wildfire/LF2016_EVT_200_CONUS/Tif/LC16_EVT_200.tif") 
evt <- crop(evt_og, ca_proj_buffer)
# EVT PHYS is EVT physiogomy 
# It is a good balance between detail and generalization that still captures relevant vegetation types for sceloporus habitat
activeCat(evt) <- "EVT_PHYS"
names(evt) <- "evt"

# Extract EVT values in a 100 m buffer around each point
coords_buffer <- coords_proj %>% st_buffer(dist = 100)

# Get counts of each EVT class per ID
evt_vals <- terra::extract(evt, coords_buffer, na.rm = FALSE, ID = TRUE)

evt_counts <-
  evt_vals  %>% 
  # Remove open water and roads
  filter(!evt %in% c("Open Water", "Developed-Roads")) %>%
  count(ID, evt, name = "n")     # counts per ID x class

# This keeps all top classes per ID (i.e., ties are retained)
evt_mod_df_with_ties <-
  evt_counts %>%
  group_by(ID) %>%
  slice_max(order_by = n, n = 1, with_ties = TRUE) %>%
  ungroup()

# IDs that have ties for the top class
ties <-
  evt_mod_df_with_ties %>%
  add_count(ID, name = "n_top") %>%  # how many "winners" for that ID
  filter(n_top > 1) %>%
  arrange(ID, desc(n), evt)

# For ties, take the class of the original coordinates
tie_ids <- unique(ties$ID)
tie_coords <- coords_proj[tie_ids, ]
tie_evt_vals <- 
  terra::extract(evt, tie_coords, na.rm = FALSE) %>%
  mutate(ID = tie_ids)

# Replace tied classes with the value from the original coordinates
evt_mod_df <-
  evt_mod_df_with_ties %>%
  filter(!(ID %in% tie_ids)) %>%  # remove tied IDs
  bind_rows(tie_evt_vals %>% select(ID, evt)) %>%  # add back in tied IDs with original coord values
  arrange(ID) %>%
  mutate(SampleID = coords_proj$SampleID) %>%
  select(-ID) 

# Get counts in each category
evt_mod_df %>% 
  count(evt) %>%
  arrange(desc(n))
#    evt                            n
#    <fct>                      <int>
#  1 Shrubland                     83
#  2 Developed                     41
#  3 Conifer                       38
#  4 Exotic Herbaceous             26
#  5 Hardwood                      16
#  6 Riparian                      14
#  7 Sparsely Vegetated             9
#  8 Developed-Medium Intensity     7
#  9 Agricultural                   7
# 10 Exotic Tree-Shrub              6
# 11 Developed-Low Intensity        5
# 12 Grassland                      3
# 13 Developed-Roads                2
# Based on these counts, I think it is safe to collapse all developed classes into one "Developed" category

# Replace any evt classes that are developed with just "Developed"
# Doing this collapsing after the counting to avoid losing information about the other vegetation types present in the buffer
evt_mod_df <-
  evt_mod_df %>%
  mutate(evt = as.character(evt)) %>%
  mutate(evt = ifelse(grepl("Developed", evt), "Developed", evt)) %>%
  select(-n)

# Print unique cats
evt_mod_df %>% count(evt)

# Check NAs
evt_mod_df %>%
  summarise(across(everything(), ~sum(is.na(.))))
```

# Hypothesis 7: Genetic diversity explained by historical fire regime
```{r}
# Historical fire regime data
# Fire Return Interval (FRI)
ca_proj_buffer <-  ca_proj %>% st_transform(5070) %>% st_buffer(dist = 10000) # buffer to ensure full coverage when cropping 

fri_og <- rast("analysis/wildfire/LF2016_FRI_200_CONUS/Tif/LC16_FRI_200.tif") 
fri <- crop(fri_og, ca_proj_buffer)

# Percent Fire Severity (PFS) 
pfs_og <- rast("analysis/wildfire/LF2016_PFS_200_CONUS/Tif/LC16_PFS_200.tif")
pfs <- crop(pfs_og, ca_proj_buffer)

# Vegetation Departure (VDep)
vdep_og <- rast("analysis/wildfire/LF2016_VDep_200_CONUS/Tif/LC16_VDep_200.tif")
vdep <- crop(vdep_og, ca_proj_buffer)

# Fire regime group (FRG) + vegetation group (BPS)
bps_og <- rast("analysis/wildfire/LF2016_BPS_200_CONUS/Tif/LC16_BPS_200.tif")
bps <- crop(bps_og, ca_proj_buffer)

# Longest fire free interval from 1984-2022
lffi_og <- rast("analysis/wildfire/conus_fire_history_metrics_1984_2020/conus_1984_2020_LFFI.tif")
lffi <- crop(lffi_og, ca_proj_buffer)
names(lffi) <- "lffi"

# Recent fire frequency (FRQ) from 1984-2022
frq_og <- rast("analysis/wildfire/conus_fire_history_metrics_1984_2020/conus_1984_2020_FRQ.tif")
frq <- crop(frq_og, ca_proj_buffer)
names(frq) <- "frq"

# Activate relevant categories for each raster
activeCat(fri)<- "FRI_ALLFIR"
activeCat(vdep) <- "LABEL"
activeCat(bps) <- "GROUPVEG"

# Only using percent replacement for fire severity (high severity fire) - other PFS categories are low/moderate and together those percentages (low + moderate + high) sum to 100, so high severity captures the relevant variation (high = 100 - low + moderate)
pfs_replac <- pfs
activeCat(pfs_replac) <- "PRC_REPLAC"
```

```{r}
# Create list of historical fire variables
fire_list <- list(pfs_replac, fri)
names(fire_list) <- c("pfs_replac", "fri")

# Make buffered points
coords_buffer <- coords_proj %>% st_buffer(dist = 100)

# Get vegetation data
groupveg_vals <- terra::extract(bps, coords_buffer, na.rm = FALSE)

# Count each vegetation type within buffer
groupveg_count <- 
  groupveg_vals %>%
  filter(GROUPVEG != "Open Water") %>%  # remove open water points
  count(ID, GROUPVEG) %>%               # count points per vegetation type
  group_by(ID) 

# Get mode vegetation type per ID
groupveg_mode_with_ties <- 
  groupveg_count %>%
  group_by(ID) %>%
  slice_max(order_by = n, n = 1, with_ties = TRUE) %>%
  ungroup() %>%
  select(ID, GROUPVEG)

# Pull out tied values and replace with the coordinate value
tied_ids <- 
  groupveg_mode_with_ties %>%
  group_by(ID) %>%
  filter(n() > 1) %>%
  pull(ID) %>%
  unique()

if (length(tied_ids) > 0) {
  tied_coords <- coords_proj[tied_ids, ]
  tied_groupveg_vals <- terra::extract(bps, tied_coords, na.rm = FALSE)
  tied_groupveg_vals$ID <- tied_ids # ensure ID column is correct
  # Replace in groupveg_mode
  groupveg_mode <- 
    groupveg_mode_with_ties %>%
    filter(!(ID %in% tied_ids)) %>%
    bind_rows(tied_groupveg_vals) %>%
    arrange(ID)
}

# Confirm that number of rows matches number of coords
stopifnot(nrow(groupveg_mode) == nrow(coords_proj))

# Get unique vegetation values for each coordinate to get a vector of allowed GROUPVEG types
allowed_groupveg <- unique(terra::extract(bps, coords_proj, na.rm = FALSE)$GROUPVEG)
# Remove open water 
allowed_groupveg <- allowed_groupveg[allowed_groupveg != "Open Water"]
# Anything that is -9999 and NOT in allowed_groupveg will be coded as NA (since it is not sceloporus habitat), any other value of -9999 will get a 0 (no fire regime)
print(allowed_groupveg)

# Create fire regime groups based on FRI and PFS based on an adaptation of bins from LANDFIRE
# NEW FRG bins (Too fine scale): https://www.landfire.gov/sites/default/files/DataDictionary/LF2016/LF16_FRGADD.pdf
# Adapted based on OLD FRG bins: https://www.landfire.gov/sites/default/files/DataDictionary/LF2014/LF14_FRGADD.pdf
frg_mode_with_ties <-
  # Extract values
  extract(rast(fire_list), coords_buffer, ID = TRUE)  %>%
  # Get mode vegetation type in buffer
  mutate(groupveg = groupveg_vals$GROUPVEG) %>%
  # Convert to numeric
  mutate(pfs_replac = as.numeric(as.character(pfs_replac)), fri = as.numeric(as.character(fri))) %>%
  # Replace -9999 with NAs
  mutate(replac_clean = ifelse(pfs_replac == -9999, NA, pfs_replac), fri_clean = ifelse(fri == -9999, NA, fri)) %>%
  mutate(
    frg = 
      case_when(
        # If vegetation group is not in allowed list, code as NA
        !(groupveg %in% allowed_groupveg) ~ NA_character_,
        # If FRI is NA and vegetation type is in allowed list, code as No Fire Regime
        is.na(fri_clean) & (groupveg %in% allowed_groupveg) ~ "No Fire Regime",
        # New FRG bins
        fri_clean <= 35 & replac_clean >= 66.7 ~ "High Frequency + High Severity",
        fri_clean <= 35 & replac_clean < 66.7 ~ "High Frequency + Low/Moderate Severity",
        fri_clean > 35 & fri_clean < 200 & replac_clean >= 66.7 ~ "Intermediate Frequency + High Severity",
        fri_clean > 35 & fri_clean < 200 & replac_clean < 66.7 ~ "Intermediate Frequency + Low/Moderate Severity",
        fri_clean >= 200 ~ "Infrequent Fire Regime",
        TRUE ~ NA_character_
      )
  ) %>%
  # Remove NA values before counting
  drop_na(frg) %>%
  # Count each fire regime group per ID
  count(ID, frg, name = "n") %>%
  group_by(ID) %>%
  # Get mode fire regime group per ID (keeping ties)
  slice_max(order_by = n, n = 1, with_ties = TRUE) %>%
  ungroup() 

# Identify IDs that have ties
frg_ties <-
  frg_mode_with_ties %>%
  add_count(ID, name = "n_top") %>%   # number of "winners" for that ID
  filter(n_top > 1) %>%
  arrange(ID, desc(n), frg)

tie_ids <- unique(frg_ties$ID)

# For tied IDs, use the original (unbuffered) coordinate value as the tiebreaker
if (length(tie_ids) > 0) {
  tie_coords <- coords_proj[tie_ids, ]

  tie_frg_vals <-
    terra::extract(rast(fire_list), tie_coords, na.rm = FALSE) %>%
    mutate(pfs_replac = as.numeric(as.character(pfs_replac)), fri = as.numeric(as.character(fri))) %>%
    mutate(replac_clean = ifelse(pfs_replac == -9999, NA, pfs_replac), fri_clean = ifelse(fri == -9999, NA, fri)) %>%
    # No vegetation check here because these are all in allowed vegetation types since they were selected from the main coords
    mutate(
      frg = 
        case_when(
          is.na(fri_clean) ~ "No Fire Regime",
          fri_clean <= 35 & replac_clean >= 66.7 ~ "High Frequency + High Severity",
          fri_clean <= 35 & replac_clean < 66.7 ~ "High Frequency + Low/Moderate Severity",
          fri_clean > 35 & fri_clean < 200 & replac_clean >= 66.7 ~ "Intermediate Frequency + High Severity",
          fri_clean > 35 & fri_clean < 200 & replac_clean < 66.7 ~ "Intermediate Frequency + Low/Moderate Severity",
          fri_clean >= 200 ~ "Infrequent Fire Regime",
          TRUE ~ NA_character_
        ),
      ID = tie_ids
    ) %>%
    select(ID, frg)

  # Replace tied IDs in the buffer-mode result with point-based category
  frg_mode_resolved <-
    frg_mode_with_ties %>%
    filter(!(ID %in% tie_ids)) %>%          # drop tied IDs
    bind_rows(tie_frg_vals) %>%             # add back resolved values
    arrange(ID)
} else {
  frg_mode_resolved <- frg_mode_with_ties %>% arrange(ID)
}

stopifnot(nrow(frg_mode_resolved) == nrow(coords_proj))  # confirm number of rows matches number of coords

# Create dataframe with SampleID and fire regime group
frg_df <- 
  frg_mode_resolved %>%
  mutate(SampleID = coords_proj$SampleID) %>%
  select(SampleID, frg) %>%
  # Set factor levels
  mutate(frg = factor(frg, levels = c(
    "No Fire Regime",
    "Infrequent Fire Regime",
    "Intermediate Frequency + Low/Moderate Severity",
    "Intermediate Frequency + High Severity",
    "High Frequency + Low/Moderate Severity",
    "High Frequency + High Severity"
  )))

# Get vegetation departure data
vdep_vals <-
  terra::extract(vdep, coords_buffer) %>%
  rename(vdep = LABEL) %>%
  mutate(vdep_num = as.numeric(as.character(vdep))) %>%
  mutate(vdep = 
    case_when(
      is.na(vdep_num) & vdep == "Water" ~ NA,
      is.na(vdep_num) & vdep == "Barren or Sparse" ~ 0,
      is.na(vdep_num) & vdep == "Urban" ~ 100,
      is.na(vdep_num) & vdep == "Agriculture" ~ 100,
      TRUE ~ vdep_num
    )
  ) %>%
  group_by(ID) %>%
  summarise(vdep = median(vdep, na.rm = TRUE))  %>%
  mutate(SampleID = coords_proj$SampleID)


# Get recent fire data
# (Not same scale as evt or vegetation group, so just take median over buffer)
recent_fire_vals <-
  terra::extract(c(lffi, frq), coords_buffer, na.rm = FALSE, ID = TRUE) %>%
  group_by(ID) %>%
  mutate(lffi_nona = ifelse(is.na(lffi), 36, lffi)) %>%
  # Calculate LFFI in the buffer, ONLY using the pixels that experienced fire (i.e., not NA values)
  # Why: we want to know the fire free interval where fire did occur, not the average over the whole buffer which would be biased towards unburned areas
  # If you do it the other way (i.e., turn NAs into 39 first) you get biased high values
  # Either way, if you remove all the points whre lffi is NA, there is a significant relationship between lffi and Ho, so I am comfortable with this approach
  mutate(frq = ifelse(is.na(frq), 0, frq)) %>% # Treat NA as 0
  summarise(
    lffi = median(lffi, na.rm = TRUE),
    lffi_nona = median(lffi_nona, na.rm = TRUE),
    frq_median = median(frq, na.rm = TRUE),
    burned = any(frq > 0, na.rm = TRUE),
    frq_max = max(frq, na.rm = TRUE)
  ) %>%
  # Use median to calculate whether there was any recent fire (frq_median > 0)
  # Don't use frq_max for bining because that would code any point that had a single pixel burned as "recent fire" and that is too sensitive (I have tested both ways and bin(frq_max) is not significant)
  mutate(frq_bin = factor(ifelse(frq_median > 0, 1, 0), levels = c(0,1), labels = c("Unburned", "Burned"))) %>%
  mutate(lffi2 = ifelse(is.na(lffi), 36, lffi)) 
  
# Combine into a single dataframe
fire_mod_df <- 
  # Combine data frames
  groupveg_mode %>% 
  rename(groupveg = GROUPVEG) %>%          
  left_join(vdep_vals, by = "ID") %>%
  left_join(recent_fire_vals, by = "ID") %>%
  select(-ID) %>%
  # Add SampleIDs
  mutate(SampleID = coords_proj$SampleID) %>%
  # Add frg
  left_join(frg_df, by = "SampleID")  %>%
  # Create simple frequency column 
  mutate(fire_frq = case_when(
    grepl("High Frequency", frg) ~ "High Frequency",
    grepl("Intermediate Frequency", frg) ~ "Intermediate Frequency",
    frg == "Infrequent Fire Regime" ~ "Infrequent/No Fire Regime",
    frg == "No Fire Regime" ~ "Infrequent/No Fire Regime",
    TRUE ~ NA_character_
  )) %>%
  # Create factors 
  mutate(
    fire_frg = factor(frg, levels = rev(c("High Frequency + High Severity", "High Frequency + Low/Moderate Severity", "Intermediate Frequency + High Severity", "Intermediate Frequency + Low/Moderate Severity", "Infrequent Fire Regime", "No Fire Regime"))),
    fire_frq = factor(fire_frq, levels = rev(c("High Frequency", "Intermediate Frequency", "Infrequent/No Fire Regime"))),
    groupveg = as.character(groupveg),
    fire_recent = frq_bin
  ) 
```

# Combine rasters into a raster stack
This can take a while because of the fire rasters
```r
env_list <- 
  list(
    NDVI = ndvi,
    bio1 = bio1,
    gHM = ghm,
    tmean_dif = tmean_dif,
    csi_past = csi_past, 
    Q = q_r,
    pfs_replac = fire_list$pfs_replac,
    pfs_mixed =fire_list$pfs_mixed, 
    pfs_surfac = fire_list$pfs_surfac, 
    fri = fire_list$fri, 
    vdep = fire_list$vdep
  )

# Project all rasters to EPSG:3310 and mask with CA
env_list <- imap(env_list, ~{
  r <- terra::project(.x, "epsg:3310", method = "bilinear")
  r <- mask(r, ca_proj)
  names(r) <- .y
  return(r)
}, .progress = TRUE)

# Output path for model rasters
path <- here("data", "env", "model_rasters")

# Make model rasters dir
dir.create(path, showWarnings = FALSE)

# Write out rasters
iwalk(env_list, ~writeRaster(.x, here(path, paste0(.y, ".tif")), overwrite = TRUE))

# Snap all rasters to same grid and resolution
# Note: these rasters are only used for visualization, so these changes don't affect the models
max_res <- map_dbl(env_list, ~res(.x)[1]) %>% which.max()
target_raster <- env_list[[max_res]] # coarsest resolution
resampled_list <- imap(env_list, ~{
  # Resample to match target
  terra::resample(.x, target_raster, method = "bilinear")
})

# Create raster stack
env_stack <- 
  resampled_list %>% 
  rast() 
```

# Create model data
```{r}
# Model variables
mod_vars <- c("NDVI", "gHM", "bio1", "csi_past", "glacier", "tmean_dif", "fire_frg", "fire_frq", "fire_recent", "vdep", "groupveg", "evt", "x", "y", "Q")

# Combine dataframes
mod_df <- 
  get_coords() %>% 
  left_join(het) %>%
  left_join(env_mod_df) %>%
  left_join(paleoclim_mod_df) %>%
  left_join(q_mod_df) %>%
  left_join(glacier_mod_df) %>%
  left_join(tmean_mod_df) %>%
  left_join(fire_mod_df) %>%
  left_join(evt_mod_df) %>%
  dplyr::select(Ho, SampleID, all_of(mod_vars))

# Check sample number
# Note: het has 259 samples, but two of them are potentially flipped ("Scelocci_CAS213197", "Scelocci_CAS214858"), so we remove the from the model (they are not included in coords)
stopifnot(nrow(mod_df) == 257)

# Confirm no NAs:
stopifnot(all(colSums(is.na(mod_df)) == 0))

# Convert to sf
mod_sf <- 
  get_coords(sf = TRUE) %>% 
  right_join(mod_df) %>% 
  st_transform(3310) %>%
  select(-x, -y) %>%
  bind_cols(st_coordinates(.)) %>%
  rename(x = X, y = Y)
```

```{r}
# Export model df
export_df <- 
  st_drop_geometry(mod_df) %>% 
  dplyr::select(SampleID, everything())
write_csv(export_df, here(outpath, "model_df.csv"))
```

## Check for correlation between variables
```{r}
# Check correlations
# Compute the correlation matrix
numeric_mod_vars <-
  mod_df %>% 
  select(all_of(mod_vars), -x, -y) %>%
  select(where(is.numeric)) %>% 
  colnames()

r <- cor(mod_df[, numeric_mod_vars], use = "pairwise.complete.obs")
# Set the diagonal and lower triangle to NA
r[lower.tri(r, diag = TRUE)] <- NA
r

# Find the pairs with correlations greater than 0.7
high_corr <- which(abs(r) > 0.6, arr.ind = TRUE)
# Print the pairs
high_corr_pairs <- data.frame(
  Var1 = rownames(r)[high_corr[, 1]],
  Var2 = rownames(r)[high_corr[, 2]],
  Correlation = r[high_corr]
) %>%
arrange(desc(abs(Correlation)))

# gHM and vdep have a correlation > 0.6
high_corr_pairs
# Decided to keep both variables since they represent different processes
```

```{r}
#Make nice correlation plot
r <- stats::cor(mod_df[, setdiff(numeric_mod_vars, c("x", "y"))], use = "pairwise.complete.obs")
rownames(r) <- colnames(r) <- make_pretty_names(rownames(r))
r[lower.tri(r, diag = TRUE)] <- NA

library(corrplot)
png(here(plotpath, "correlation_plot.png"), width = 10*300, height = 10*300, units = "px", res = 300)
corrplot(r, method = "ellipse", type = "upper", tl.col = "black", tl.srt = 45, diag = FALSE, addCoef.col = "black", number.cex = 0.8)
dev.off()
```

```{r}
# Checking R squared for BPS vegetation group
groupveg_rsq <-  
  map(numeric_mod_vars, ~{
    mod <- lm(as.formula(paste0(.x, " ~ groupveg")), data = mod_df)
    rsq <- summary(mod)$adj.r.squared
    data.frame(variable = .x, rsq = rsq)
  }) %>%
  bind_rows() %>%
  arrange(desc(rsq))

# Filter for high R2 (>50% variance explained by vegetation group)
groupveg_rsq %>% filter(rsq > 0.5) # Only fire related variables

# Checking R squared for current vegetation group
evt_rsq  <- 
  map(numeric_mod_vars, ~{
    mod <- lm(as.formula(paste0(.x, " ~ evt")), data = mod_df)
    rsq <- summary(mod)$adj.r.squared
    data.frame(variable = .x, rsq = rsq)
  }) %>%
  bind_rows() %>%
  arrange(desc(rsq))

# Filter for high R2 (>50% variance explained by vegetation group)
evt_rsq %>% filter(rsq > 0.5) # No variables >50%
evt_rsq %>% filter(rsq > 0.4) # But vdep is 45% and gHM is 41%, likely due to correlation with the Developed vegetation group

# Check correlation between groupveg and evt
mean(mod_df$groupveg == mod_df$evt)  # Only 42% match

# Checking R squared for fire categories
fire_frg_cat_rsq  <- 
  map(numeric_mod_vars, ~{
    mod <- lm(as.formula(paste0(.x, " ~ fire_frg")), data = mod_df)
    rsq <- summary(mod)$adj.r.squared
    data.frame(variable = .x, rsq = rsq)
  }) %>%
  bind_rows() %>%
  arrange(desc(rsq))

# Filter for high R2
fire_frg_cat_rsq %>% filter(rsq > 0.5) # None

# Checking R squared for glacier
glacier_rsq <-  
  map(numeric_mod_vars, ~{
    mod <- lm(as.formula(paste0(.x, " ~ glacier")), data = mod_df)
    rsq <- summary(mod)$adj.r.squared
    data.frame(variable = .x, rsq = rsq)
  }) %>%
  bind_rows() %>%
  arrange(desc(rsq))

# Filter for high R2 (>50% variance explained by glacier)
glacier_rsq %>% filter(rsq > 0.5) # None
```

```{r}
# Check correlation for BPS vs EVT vegetation group 
# Just check intersection of levels
veg_levels <- expand.grid(
  unique(mod_df$groupveg),
  unique(mod_df$evt)
)

veg_corr <- pmap(veg_levels, function(Var1, Var2) {
  tibble(
    veg_class1 = Var1,
    veg_class2 = Var2,
    cor = cor(
      as.integer(mod_df$groupveg == Var1),
      as.integer(mod_df$evt == Var2),
      method = "pearson"
    )
  )
}) %>%
  bind_rows()

veg_corr %>% arrange(desc(cor))
```

## Transform and scale data for modeling
```{r}
# Model formula variables
mod_vars <- c("glacier", "NDVI", "gHM", "bio1", "csi_past", "tmean_dif",  "fire_frg", "fire_frq", "fire_recent", "vdep", "groupveg", "evt", "Q")

# GLS can't handle 0 distances, so take the average of the coordinates (mode for veg type)
mode_val <- function(x) {
  ux <- na.omit(x)
  ux[which.max(tabulate(match(ux, unique(ux))))]
}
cat_mod_vars <- mod_df %>% 
  select(all_of(mod_vars)) %>%
  select(where(~is.character(.x) | is.factor(.x))) %>%
  colnames()
numeric_mod_vars <- setdiff(mod_vars, cat_mod_vars)

mod_sf_unique <- 
  mod_sf %>%
  group_by(x, y) %>%
  summarise(
    across(all_of(c(numeric_mod_vars, "Ho")), ~ mean(.x, na.rm = TRUE)),
    across(all_of(cat_mod_vars), ~ mode_val(.x)),
    .groups = "drop"
  )

# Scale data
mod_sf_unique_scaled <-
  mod_sf_unique %>%
  mutate(
    across(
      all_of(numeric_mod_vars),
      ~ as.numeric(scale(.x))
    )
  )

# Set reference levels for groupveg, evt, and fire variables
mod_sf_unique_scaled$groupveg <- relevel(as.factor(mod_sf_unique_scaled$groupveg), ref = "Shrubland")
mod_sf_unique_scaled$evt <- relevel(as.factor(mod_sf_unique_scaled$evt), ref = "Shrubland")
mod_sf_unique_scaled$fire_frg <- relevel(as.factor(mod_sf_unique_scaled$fire_frg), ref = "No Fire Regime")
mod_sf_unique_scaled$fire_frq <- relevel(as.factor(mod_sf_unique_scaled$fire_frq), ref = "Infrequent/No Fire Regime") # Note: setting high frequency as reference because this variable is used to test for an interaction with recent fire (hypothesis: recent fire suppression will negatively effect Ho in high frequency fire regimes, but not in other cases). We do an emmeans post-hoc test to compare all levels, so the choice of reference here is not critical.

# Confirm no NA
stopifnot(all(colSums(is.na(mod_sf_unique_scaled)) == 0))

```

```{r}
# After visual inspection of the data I observed a non-linear relationship between csi_past and Ho
# So I am testing a quadratic term for csi_past
model_linear <- lm(Ho ~ NDVI + glacier + gHM + bio1 + csi_past + tmean_dif + evt + fire_frg + fire_recent, data = mod_sf_unique)
model_poly <- lm(Ho ~ NDVI + glacier + gHM + bio1 + poly(csi_past, 2, raw = TRUE) + tmean_dif + evt + fire_frg + fire_recent, data = mod_sf_unique)
anova(model_linear, model_poly) # The polynomial model is significantly better
```

## Fit initial OLS and GLS models
```{r}
f <- formula("Ho ~ NDVI + glacier + gHM + bio1 + poly(csi_past, 2, raw = TRUE) + tmean_dif + evt + fire_frg + fire_recent")

# Run OLS and get residuals
ols_model <- lm(f, data = mod_sf_unique)
summary(ols_model)
AIC(ols_model)
mod_sf_unique$ols_resids <- residuals(ols_model)
broom::tidy(ols_model) %>% filter(p.value < 0.05) %>% arrange(desc(abs(estimate)))

# Fit variograms to the residuals
ols_variogram <- variogram(ols_resids ~ 1, data = mod_sf_unique)
ols_vgm <- map(c("Exp", "Sph", "Gau"), ~fit.variogram(ols_variogram, model = vgm(.x)))
names(ols_vgm) <- c("Exp", "Sph", "Gau")

pdf(here(plotpath, "ols_variogram.pdf"))
plot(ols_variogram, ols_vgm[["Exp"]], main = "OLS Residuals Variogram (Exponential)")
plot(ols_variogram, ols_vgm[["Sph"]], main = "OLS Residuals Variogram (Spherical)")
plot(ols_variogram, ols_vgm[["Gau"]], main = "OLS Residuals Variogram (Gaussian)")
dev.off()
# Based on visual inspection, the exponential model seems to fit best

# Fit a GLS model
# Note: nugget=FALSE assumes residuals at the same location are perfectly correlated, which should align with heterozygosity, since you don't expect to have much microscale variation in heterozygosity.
# nugget=TRUE allows unstructured microscale variance (random noise at zero distance), which can absorb spatial signal and cause predictor effects to appear insignificant. 
gls_model <- gls(f, correlation = corExp(form = ~ x + y, nugget = FALSE), data = mod_sf_unique, method = "ML") # Fit with ML to compare AIC later
summary(gls_model)
# Use type = "normalized" to get residuals decorrelated and standardized by the fitted 
# spatial covariance matrix; this removes modeled spatial structure and makes them 
# suitable for Moran's I. If you don't do this Moran's I will appear higher than it actually is.
mod_sf_unique$gls_resids <- residuals(gls_model, type = "normalized")

# Plot residuals
plot_resids <- function(resids){
  # Remove max residual because it is hard to see spatial patterns with it
  ggplot(mod_sf_unique[mod_sf_unique[[resids]] != max(mod_sf_unique[[resids]]), ]) +
    geom_sf(data = ca_proj, col = NA) +
    geom_sf(aes(col = .data[[resids]])) +
    theme_void() +
    scale_color_viridis_c(option = "turbo", end = 0.1, begin = 0.9, direction = -1) +
    labs(title = paste("Residuals:", resids))
}

pdf(here(plotpath, "all_model_residuals.pdf"), width = 9, height = 4)
plt_list <- list(
  plot_resids("ols_resids"),
  plot_resids("gls_resids")
)
plot_grid(plotlist = plt_list, nrow = 1)
dev.off()
# The GLS model residuals look the same as the OLS residuals

# Compare AIC
anova(gls_model, ols_model) # GLS significantly better, but absolute AIC difference is small

# Compare significant variables (p < 0.05) for each model using map
models <- list(OLS = ols_model, GLS = gls_model)
sig_vars_by_model <- map(models, ~ broom::tidy(.x) %>%
  filter(!term %in% c("lambda", "rho", "(Intercept)"), p.value < 0.05) %>%
  pull(term))
print(sig_vars_by_model) # Same significant variables for both models
```

```{r}
# Test for spatial autocorrelation in residuals using Moran's I
library(spdep)

# Coordinates matrix
coords <- cbind(mod_sf_unique$x, mod_sf_unique$y)

# 8 nearest neighbors (robust for irregular sampling)
knn_8 <- knearneigh(coords, k = 8)
nb_8  <- knn2nb(knn_8)
lw_8  <- nb2listw(nb_8, style = "W")

# Moran's I for OLS residuals
moran_ols <- moran.test(mod_sf_unique$ols_resids, lw_8)

# Moran's I for GLS normalized residuals
moran_gls <- moran.test(mod_sf_unique$gls_resids, lw_8)

print(moran_ols) # 0.140
print(moran_gls) # 0.144
```

# Fit full model 1
```{r}
f_0 <- formula("Ho ~ NDVI + glacier + gHM + bio1 + poly(csi_past, 2, raw = TRUE) + tmean_dif + evt + fire_frg + fire_recent + vdep")

# Fit model
gls_model0 <- gls(f_0, correlation = corExp(form = ~ x + y, nugget = FALSE), data = mod_sf_unique_scaled, method = "ML")

# Calculate VIF
performance::check_collinearity(gls_model0) 
# High Correlation
#  Term   VIF     VIF 95% CI adj. VIF Tolerance Tolerance 95% CI
#   evt 16.07 [12.99, 19.94]     4.01      0.06     [0.05, 0.08]
# Because of high VIF for evt, and knowing that 45% of the variation VDep and 40% of the variation of NDVI is explained by evt (and gHM is also correlated with VDep), we are going to run the model with EVT, but without VDep or NDVI. I am making this choice because EVT is an important measure of habitat structure and gHM is a more direct measure of human impact compared to VDep which describes many different kinds of change on the same scale and NDVI which is not directly interpretable in terms of habitat structure.
# I have tried just removing VDep or NDVI (individually), but the VIFs are still high .
f_1 <- formula("Ho ~ glacier + gHM + bio1 + poly(csi_past, 2, raw = TRUE) + tmean_dif + evt + fire_frg + fire_recent")

# Re-fit GLS
gls_model1 <- gls(f_1, correlation = corExp(form = ~ x + y, nugget = FALSE), data = mod_sf_unique_scaled, method = "ML")

# Re-calculate VIF
performance::check_collinearity(gls_model1) 
# No high correlation issues

# Double check OLS
ols_model1 <- lm(f_1, data = mod_sf_unique_scaled)

# Compare OLS vs GLS
anova(gls_model1, ols_model1) # GLS significantly better, but absolute AIC difference is small

# Refit final model with REML
gls_model1_final <- gls(f_1, correlation = corExp(form = ~ x + y, nugget = FALSE), data = mod_sf_unique_scaled, method = "REML")

# Results
mod_result <- broom::tidy(gls_model1_final)
mod_result

# Significant variables
mod_result %>% filter(p.value < 0.05) %>% arrange(desc(abs(estimate))) 
sig_vars <- mod_result %>% filter(p.value < 0.05, !term %in% c("(Intercept)", "lambda")) %>% pull(term)
print(sig_vars)

# You can't get a traditional R2 for spatial models, but you can get an alternative R2
cor(predict(gls_model1_final), mod_sf_unique$Ho)^2 # 0.52
cor(predict(gls_model1_final), mod_sf_unique$Ho) # 0.72
```

```{r}
# Mode = "df.error" for GLS models

# Test for differences between vegetation groups
emmeans_gls <- emmeans(gls_model1_final, specs = "evt", mode = "df.error")
tukey_gls <- contrast(emmeans_gls, method = "tukey", adjust = "fdr")
tukey_gls %>% data.frame() %>% filter(p.value < 0.05)
# 0 rows

# Test for differences between fire regime groups
emmeans_gls <- emmeans(gls_model1_final, specs = "fire_frg", mode = "df.error")
tukey_gls <- contrast(emmeans_gls, method = "tukey", adjust = "fdr")
tukey_gls %>% data.frame() %>% filter(p.value < 0.05)
#                                                                              contrast
# 1                   No Fire Regime - (Intermediate Frequency + Low/Moderate Severity)
# 2                           No Fire Regime - (Intermediate Frequency + High Severity)
# 3                           No Fire Regime - (High Frequency + Low/Moderate Severity)
# 4 (Intermediate Frequency + High Severity) - (High Frequency + Low/Moderate Severity)
# 5         (Intermediate Frequency + High Severity) - (High Frequency + High Severity)
#        estimate           SE  df   t.ratio     p.value
# 1 -0.0002871168 1.083897e-04 231 -2.648931 0.032369211
# 2 -0.0003732519 8.288612e-05 231 -4.503190 0.000159363
# 3 -0.0002544819 8.419149e-05 231 -3.022656 0.013941343
# 4  0.0001187700 3.924938e-05 231  3.026035 0.013941343
# 5  0.0002005696 7.881223e-05 231  2.544905 0.034748506
```

# Fit full model 2: interaction between fire frequency and recent fire

```{r}
f_2 <- formula("Ho ~ glacier + gHM + bio1 + poly(csi_past, 2, raw = TRUE) + tmean_dif + evt + fire_frq * fire_recent") 

# Fit model
gls_model2 <- gls(f_2, correlation = corExp(form = ~ x + y, nugget = FALSE), data = mod_sf_unique_scaled, method = "ML")

# Double check OLS
ols_model2 <- lm(f_2, data = mod_sf_unique_scaled)

# Compare OLS vs GLS
anova(gls_model2, ols_model2) # GLS significantly better, but absolute AIC difference is small

# Compare GLS models
anova(gls_model1, gls_model2) # No significant difference between models

# Calculate VIF
performance::check_collinearity(gls_model2) # From performance package
# No high correlation issues (other than interaction term, which just depends on the reference level set)

# Refit model with REML
gls_model2_final <- gls(f_2, correlation = corExp(form = ~ x + y, nugget = FALSE), data = mod_sf_unique_scaled, method = "REML")

# Results
mod_result <- broom::tidy(gls_model2)
mod_result

# Significant variables
mod_result %>% filter(p.value < 0.05) %>% arrange((p.value))
sig_vars <- mod_result %>% filter(p.value < 0.05, !term %in% c("(Intercept)", "lambda")) %>% pull(term)
print(sig_vars)

# You can't get a traditional R2 for spatial models, but you can get an alternative R2
cor(predict(gls_model2), mod_sf_unique$Ho)^2 # 0.0.52
cor(predict(gls_model2), mod_sf_unique$Ho) # 0.72
```

```{r}
# !Most important test: test for differences between recent fire vs no recent fire within each fire frequency group
emm <- emmeans(gls_model2, ~ fire_recent | fire_frq, mode = "df.error")
contrast(emm, method = "pairwise") # No adjustment because there are only two levels for fire
# fire_frq = Infrequent/No Fire Regime:
#  contrast           estimate       SE  df t.ratio p.value
#  Unburned - Burned  2.15e-04 0.000168 234   1.285  0.2001

# fire_frq = Intermediate Frequency:
#  contrast           estimate       SE  df t.ratio p.value
#  Unburned - Burned  5.34e-05 0.000049 234   1.089  0.2771

# fire_frq = High Frequency:
#  contrast           estimate       SE  df t.ratio p.value
#  Unburned - Burned -1.15e-04 0.000051 234  -2.251  0.0253

# Test for differences between vegetation groups
emmeans_gls <- emmeans(gls_model2, specs = "evt", mode = "df.error")
tukey_gls <- contrast(emmeans_gls, method = "tukey", adjust = "fdr")
tukey_gls %>% data.frame() %>% filter(p.value < 0.05) 
# 0 rows 

# Test for differences between fire regime groups
emmeans_gls <- emmeans(gls_model2, specs = "fire_frq", mode = "df.error")
tukey_gls <- contrast(emmeans_gls, method = "tukey", adjust = "fdr")
tukey_gls %>% data.frame() %>% filter(p.value < 0.05) 
# NOTE: Results may be misleading due to interactions (because you average over levels of fire_recent)
#                                               contrast      estimate
# 1 (Infrequent/No Fire Regime) - Intermediate Frequency -3.320759e-04
# 2         (Infrequent/No Fire Regime) - High Frequency -2.492403e-04
# 3              Intermediate Frequency - High Frequency  8.283566e-05
#             SE  df   t.ratio     p.value
# 1 9.287118e-05 234 -3.575662 0.001273024
# 2 9.470142e-05 234 -2.631853 0.013585296
# 3 3.871773e-05 234  2.139476 0.033432306
```

# Figures
```{r}
# Define a color scale function using MetBrewer with an argument to extend the scale for better visualization
scale_met_cont <- function(values, aesthetic = c("fill", "colour"), extend_scale = 0.1, r_min  = NULL, r_max = NULL) {
  aesthetic <- match.arg(aesthetic)
  if (is.null(r_min)) r_min <- min(values, na.rm = TRUE)
  if (is.null(r_max)) r_max <- max(values, na.rm = TRUE)
  rng   <- r_max - r_min
  lims  <- c(r_min - rng*extend_scale, r_max + rng*extend_scale)

  pal <- rev(MetBrewer::met.brewer("Hiroshige", type = "continuous"))
  if (aesthetic == "fill") {
    ggplot2::scale_fill_gradientn(colors = pal, limits = lims, oob = scales::squish)
  } else {
    ggplot2::scale_colour_gradientn(colors = pal, limits = lims, oob = scales::squish)
  }
}

df = mod_sf_unique
model = gls_model2
x = "fire_recent"
y = "Ho"
residual = TRUE
color_var = "y"
# Make residual plots for model
plot_mod <- function(x, y, df, model, mod_vars,
                     col_var = "Ho_original", residual = TRUE) {

  # Calculate residuals for y variable given x variable
  # Keep a copy of original Ho
  df$Ho_original <- df$Ho
  if (residual) {
    # Scale all model variables in df that are numeric
    numeric_mod_vars <- names(df)[sapply(df, is.numeric)]
    df_scaled <- 
      df %>%
      mutate(across(all_of(numeric_mod_vars),~ as.numeric(scale(.x))))

    # Residuals
    r <- residuals(model)

    # Model matrix + coefficients, aligned
    X <- model.matrix(terms(model), data = df_scaled)
    b <- coef(model)

    # Align in case ordering differs
    b <- b[colnames(X)]

    # Calculate residuals
    cols <- grepl(x, colnames(X))
    if (all(!cols)) {
      stop(paste("Variable", x, "not found in model"))
    }
    sub_X <- X[, cols, drop = FALSE] 
    sub_b <- b[cols]
    stopifnot(all(colnames(sub_X) == names(sub_b)))
    df[[y]] <- r + drop(sub_X %*% sub_b)
  }

  # Extract beta and p-value for x variable
  model_summary <- broom::tidy(model)
  b_annotation <- 
    model_summary %>%
    dplyr::filter(grepl(x, term)) %>%
    dplyr::mutate(
      beta = formatC(estimate, format = "e", digits = 2),
      p_lab = ifelse(p.value < 0.001, paste0("p[", row_number(), "] < 0.001"), paste0("p[", row_number(), "] == ", round(p.value, 3))),
      line = paste0("beta[", row_number(), "] == ", beta, "*','~", p_lab)
    )

  if (x == "csi_past") {
    
    label_txt <- paste0("atop(", paste(b_annotation$line, collapse = ","), ")")
    
    plt <-
      ggplot(df, aes(x = .data[[x]], y = .data[[y]])) +
      geom_point(aes(col = .data[[col_var]]), size = 2, show.legend = FALSE) +
      geom_smooth(method = "lm", colour = "black", formula = y ~ poly(x, 2)) +
      annotate("text", x = max(df[[x]], na.rm = TRUE), y = min(df[[y]], na.rm = TRUE), label = label_txt, size = 5, parse = TRUE, vjust = -0.1, hjust = 1) 

  } else if (x == "fire_frq") {
    
    emm  <- emmeans(model, specs = x, mode = "df.error")
    pairs_df <- 
      pairs(emm, adjust = "fdr") %>% summary(infer = TRUE) %>% as.data.frame() %>%
      separate(contrast, c("group1","group2"), " - ") %>%
      # Remove just the parentheses symbols from group1 and 2
      mutate_at(vars(group1, group2), ~str_replace_all(.x, "[()]", "")) %>%
      mutate(annotations = cut(p.value, c(-Inf,.001,.01,.05,Inf), c("***","**","*","ns"))) %>%
      mutate(xmin = group1, xmax = group2) 
      
    df$new_x <- factor(df[[x]], levels = levels(df[[x]]), labels = c("No Fire/Low", "Intermediate", "High"))
    pairs_df$xmin <- factor(pairs_df$xmin, levels = levels(df[[x]]), labels = c("No Fire/Low", "Intermediate", "High"))
    pairs_df$xmax <- factor(pairs_df$xmax, levels = levels(df[[x]]), labels = c("No Fire/Low", "Intermediate", "High"))
    
    pairs_df$y_position <- max(df[[y]], na.rm=TRUE) * (0.8 + 0.08*seq_len(nrow(pairs_df)))

    plt <- 
      ggplot(df, aes(new_x, .data[[y]])) +
      geom_boxplot() +
      geom_jitter(aes(col = .data[[col_var]]), size = 2, width = 0.2, show.legend = FALSE) +
      ggpubr::stat_pvalue_manual(pairs_df,
        label = "annotations",
        xmin = "xmin", xmax = "xmax",
        y.position = "y_position",
        size = 8,
        tip.length = 0,
        vjust = 0.6
      )
  } else if (x == "fire_recent") {

    dodge_w <- 0.8
    frq_lvls <- levels(df$fire_frq)
    rec_lvls <- levels(df$fire_recent)
    nrec <- length(rec_lvls)
    off <- setNames(((seq_len(nrec) - (nrec + 1)/2) * (dodge_w / nrec)), rec_lvls)  # for 2 lvls: -0.2, +0.2

    emm <- emmeans(model, ~ fire_recent | fire_frq, mode = "df.error")
    ymax_df <- df %>% dplyr::group_by(fire_frq) %>% dplyr::summarise(ymax = max(.data[[y]], na.rm=TRUE), .groups="drop")

    pairs_df <- contrast(emm, "pairwise") %>% as.data.frame() %>%
      separate(contrast, c("group1","group2"), " - ") %>%
      mutate(
        across(c(group1, group2), ~ stringr::str_replace_all(.x, "[()]", "")),
        annotations = cut(p.value, c(-Inf,.001,.01,.05,Inf), c("***","**","*","ns")),
        xbase = as.numeric(factor(fire_frq, levels = frq_lvls)),
        xmin  = xbase + off[group1],
        xmax  = xbase + off[group2]
      ) %>%
      filter(annotations != "ns") %>%
      left_join(ymax_df, by="fire_frq") %>%
      group_by(fire_frq) %>% 
      mutate(y_position = ymax * (1.01 + 0.08*(row_number()-1))) %>% 
      ungroup()

    df$new_fire_frq <- factor(df$fire_frq, levels = levels(df$fire_frq), labels = c("No Fire/Low", "Intermediate", "High"))
    pairs_df$fire_frq <- factor(pairs_df$fire_frq, levels = levels(df$fire_frq), labels = c("No Fire/Low", "Intermediate", "High"))

    plt <-
      ggplot(df, aes(x = new_fire_frq, y = .data[[y]])) +
      geom_boxplot(aes(fill = fire_recent), position = position_dodge(dodge_w), outlier.shape = NA) +
      geom_point(
        aes(color = .data[[col_var]], group = fire_recent),
        position = position_jitterdodge(jitter.width = 0.15, dodge.width = dodge_w),
        size = 2, show.legend = FALSE
      ) +
      ggpubr::stat_pvalue_manual(pairs_df,
        label = "annotations", xmin = "xmin", xmax = "xmax",  y.position = "y_position",
        size = 10, tip.length = 0, vjust = 0.6
      ) +
      guides(col = "none") +
      scale_fill_manual(values = c("Burned" = "gray", "Unburned" = "white")) +
      labs(fill = "Recent fire") 
    
  } else {

    label_txt <- paste(b_annotation$line, collapse = ",")

    plt <-
      ggplot(df, aes(x = .data[[x]], y = .data[[y]])) +
      geom_point(aes(col = .data[[col_var]]), size = 2, show.legend = FALSE) +
      geom_smooth(method = "lm", colour = "black", formula = y ~ x) +
      annotate("text", x = max(df[[x]], na.rm = TRUE), y = min(df[[y]], na.rm = TRUE), label = label_txt, size = 5, parse = TRUE, vjust = -0.1, hjust = 1) 
  }
  
  x_label <- ifelse(x == "fire_recent", "fire_frq", x) # For recent fire, X is fire regime
  plt <- 
    plt + 
    xlab(stringr::str_to_sentence(make_pretty_names(x_label))) +
    theme_classic() +
    theme(axis.title = element_text(size = 16), axis.text  = element_text(size = 14)) +
    scale_met_cont(df[[col_var]], aesthetic = "colour", extend_scale = 0.1) 

  # Add back legend to fire recent
  if (x == "fire_recent") {
    plt <- 
      plt +
      theme(
        legend.position = "inside",
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 14),
        legend.justification = c(1, 1),
        legend.position.inside = c(0.4, 0.95),
      )
  }

  if (residual) {
    plt + labs(y = "Partial residual heterozygosity")
  } else {
    plt + labs(y = "Observed heterozygosity")
  }

}

# Load polygons for plotting
ca_proj <- get_ca() %>% sf::st_transform(3310) # California boundaries
range_map <- get_range() # Sceloporus range map

# Loaded wingen map
wingen_r <- 
  terra::rast(here(outpath, "Ho_kriged_masked.tif")) %>% 
  mask(range_map) %>% 
  mask(ca_proj) %>%
  as.data.frame(xy = TRUE, na.rm = TRUE) 
names(wingen_r)[3] <- "layer"  # ensure 'layer' is the raster value column

# Isolate significant vs all variables
vars_to_plot <- c("csi_past", "tmean_dif", "fire_frq", "fire_recent")
mod_vars <- c("csi_past", "tmean_dif", "bio1", "glacier", "gHM", "fire_frq", "fire_recent", "evt")

# A: wingen map of heterozygosity 
ho_plot <- ggplot(wingen_r) +
  geom_sf(data = ca_proj, fill = "gray90", colour = NA, size = 2) +
  geom_tile(aes(x = x, y = y, fill = layer), colour = NA) +
  scale_met_cont(wingen_r$layer, aesthetic = "fill", extend_scale = 0.1) +
  labs(fill = "Heterozygosity") +
  theme_void() +
  geom_sf(data = mod_sf_unique, shape = 21, aes(fill = Ho), colour = "#00000087", size = 2) +
  theme(
    plot.margin = margin(1, 1, 1, 1, unit = "mm"),
    legend.position = "inside",              # turn on "inside" mode
    legend.justification = c(1, 1),          # anchor legend's top-right corner
    legend.position.inside = c(0.9, 0.95),   # place near panel's top-right
    legend.key.height = unit(0.8, "cm"),
    legend.key.width  = unit(0.8, "cm"),
    legend.title = element_text(size = 16),
    legend.text  = element_text(size = 14)
  )

# B: partial residuals 
plts_residual_y <- purrr::map(
  vars_to_plot,
  # Use unscaled data for plotting
  ~plot_mod(.x, "Ho", mod_sf_unique, model = gls_model2_final, col_var = "Ho_original", residual = TRUE)
)

# B: Raw values
plts_raw_y <- purrr::map(
  vars_to_plot,
  # Use unscaled data for plotting
  ~plot_mod(.x, "Ho", mod_sf_unique, model = gls_model2_final, col_var = "Ho_original", residual = FALSE)
)

# C: predicted vs observed 
pred_vs_obs <-
  # Use unscaled data for plotting
  ggplot(mod_sf_unique, aes(x = predict(gls_model2), y = Ho)) +
  geom_point(aes(col = Ho), shape = 16, size = 2) +
  geom_smooth(method = "lm", colour = "black") +
  scale_met_cont(c(mod_sf_unique$Ho, wingen_r$layer), aesthetic = "colour", extend_scale = 0.1) +
  theme_classic() +
  labs(x = "Predicted heterozygosity", y = "Observed heterozygosity") +
  ggpubr::stat_cor(size = 6) +
  theme(legend.position = "none",
        axis.title = element_text(size = 16),
        axis.text  = element_text(size = 14))

pred_vs_obs_adj <- cowplot::plot_grid(pred_vs_obs, NULL, nrow = 1, rel_widths = c(0.9, 0.1))

plt2_residual <- cowplot::plot_grid(plotlist = plts_residual_y, ncol = 2)
plt2_raw <- cowplot::plot_grid(plotlist = plts_raw_y, ncol = 2)
plt1 <- cowplot::plot_grid(plotlist = list(ho_plot, pred_vs_obs_adj), ncol = 1,
                           labels = c("A", "C"), rel_heights = c(0.6, 0.45), label_size = 18)

png(here(plotpath, "figure_Ho_residual.png"), height = 10*0.9, width = 16*0.9, res = 300, units = "in")
cowplot::plot_grid(plt1, plt2_residual, ncol = 2, labels = c("", "B"),  rel_widths = c(0.33, 0.66), align = "h", label_size = 18)
dev.off()

png(here(plotpath, "figure_Ho_raw.png"), height = 10, width = 16, res = 300, units = "in")
cowplot::plot_grid(plt1, plt2_raw, ncol = 2,labels = c("", "B"),  rel_widths = c(0.33, 0.66), align = "h", label_size = 18)
dev.off()
```

```{r}
# Bin csi_past and count admixed individuals in each bin
1- range(bin_df$Q) # Reminder that Q is ancestry score (so actually 1 - max(original Q))
bin_df <- 
  mod_df %>%
  # Admixed = greater than 10% other population ancestry
  mutate(admix = as.numeric(Q > 0.10)) %>%
  mutate(csi_past_bin = cut(csi_past, breaks = 5, include.lowest = TRUE))
  
prop_df <- 
  bin_df %>%
  count(csi_past_bin, admix, name = "n") %>%
  group_by(csi_past_bin) %>%
  mutate(prop = n / sum(n),
         admix = factor(admix, levels = c(0,1), labels = c("Non-admixed", "Admixed"))) %>%
  ungroup()

bin_df$admix <- factor(bin_df$admix, levels = c(0,1), labels = c("Non-admixed", "Admixed"))


# get midpoints for labels
bin_levels <- levels(bin_df$csi_past_bin)
bin_mids <- sapply(bin_levels, function(s) {
  nums <- as.numeric(gsub("\\(|\\]|\\[|\\)", "", unlist(strsplit(s, ","))))
  mean(nums)
})
bin_lab <- sprintf("%.2f", bin_mids)  # change decimals here

# Get high and low of each label
bin_max <- sapply(bin_levels, function(s) {
  nums <- as.numeric(gsub("\\(|\\]|\\[|\\)", "", unlist(strsplit(s, ","))))
  max(nums)
})

total_count <- 
  prop_df %>%
  group_by(csi_past_bin) %>%
  summarise(total = sum(n)) 

bin_csi_plt <-
  ggplot(bin_df, aes(x = csi_past, y = Ho)) +
  geom_point(aes(col = admix)) +
  geom_smooth(method = "lm", colour = "black", formula = y ~ poly(x, 2)) +
  geom_vline(xintercept = bin_max, linetype = "dashed", colour = "gray50") +
  theme_classic()  +
  labs(x = "Paleoclimate stability", y = "Heterozygosity", col = "") +
  scale_x_continuous(labels = bin_lab, breaks = as.numeric(bin_lab), expand = c(0,0)) +
  scale_color_manual(values = c("Non-admixed" = "red2", "Admixed" = "skyblue2")) +
  theme(
    axis.title = element_text(size = 18),
    axis.text  = element_text(size = 16),
    legend.text = element_text(size = 16),
    legend.position = "top",
  )

csi_plt_count <- 
  ggplot(prop_df, aes(x = csi_past_bin)) +
  geom_col(aes(y = prop, fill = admix)) +
  # Add text with counts to bars
  geom_text(data = prop_df, aes(label = n, y = ifelse(admix == "Admixed", prop + 0.1, 0.9), col = admix), size = 7, fontface = 2) +
  scale_fill_manual(values = c("Non-admixed" = "red2", "Admixed" = "skyblue2")) +
  scale_color_manual(values = c("Non-admixed" = "pink1", "Admixed" = "lightskyblue1")) +
  scale_x_discrete(labels = bin_lab, expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic() +
  labs(x = "Binned paleoclimate stability", y = "Proportion", fill = "") +
  theme(
    legend.position = "none",
    axis.title = element_text(size = 18),
    axis.text  = element_text(size = 16),
    legend.text = element_text(size = 16)
  )

# REPEAT FOR FIRE
bin_df$fire_frq <- factor(bin_df$fire_frq, levels = levels(bin_df$fire_frq), labels = c("No Fire/Low", "Intermediate",  "High"))

fire_prop_df <- 
  bin_df %>%
  count(fire_frq, admix, name = "n") %>%
  group_by(fire_frq) %>%
  mutate(prop = n / sum(n),
         admix = factor(admix, levels = c(0,1), labels = c("Non-admixed", "Admixed"))) %>%
  ungroup()

bin_df$admix <- factor(bin_df$admix, levels = c(0,1), labels = c("Non-admixed", "Admixed"))

total_count <- 
  fire_prop_df %>%
  group_by(fire_frq) %>%
  summarise(total = sum(n)) 

fire_plt <-
  ggplot(bin_df, aes(x = fire_frq, y = Ho)) +
  geom_boxplot(aes(col = admix)) +
  geom_jitter(aes(col = admix), size = 2, position = position_jitterdodge(), alpha = 0.5) +
  theme_classic()  +
  labs(x = "Historical fire frequency", y = "Heterozygosity", col = "") +
  scale_color_manual(values = c("Non-admixed" = "red2", "Admixed" = "skyblue2")) +
  theme(
    axis.title = element_text(size = 18),
    axis.text  = element_text(size = 16),
    legend.text = element_text(size = 16),
    legend.position = "top",
  )


fire_plt_count <- 
  ggplot(fire_prop_df, aes(x = fire_frq)) +
  geom_col(aes(y = prop, fill = admix)) +
  # Add text with counts to bars
  geom_text(data = fire_prop_df, aes(label = n, y = ifelse(admix == "Admixed", prop + 0.1, 0.9), col = admix), size = 7, fontface = 2) +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_manual(values = c("Non-admixed" = "red2", "Admixed" = "skyblue2")) +
  scale_color_manual(values = c("Non-admixed" = "pink1", "Admixed" = "lightskyblue1")) +
  labs(x = "Fire frequency", y = "Proportion", fill = "", col = "") +
  theme(
    legend.position = "none",
    axis.title = element_text(size = 18),
    axis.text  = element_text(size = 16),
    legend.text = element_text(size = 16)
  )

# Combine
admix_fire_plt <- plot_grid(fire_plt, fire_plt_count, ncol = 1, align = "v", rel_heights = c(0.6, 0.4))
admix_csi_plt <- plot_grid(bin_csi_plt, csi_plt_count, ncol = 1, align = "v", rel_heights = c(0.6, 0.4))

png(here(plotpath, "fire_vs_admix.png"), height = 7, width = 12, res = 300, units = "in")
plot_grid(admix_csi_plt, admix_fire_plt, nrow = 1, labels = c("A", "B"), label_size = 18)
dev.off()

# Binomial model for admix
mod <- glm(admix ~ poly(csi_past, 2, raw = TRUE) + fire_frq*fire_recent, data = bin_df, family = "binomial")
car::Anova(mod, 2)

# Emmeans comparisons between bins
# Test for differences between fire regime groups
emm <- emmeans(mod, specs = "fire_frq", mode = "df.error")
tk <- contrast(emm, method = "tukey", adjust = "fdr")
summary(tk, infer = TRUE) %>% filter(p.value < 0.05) # Intermediate - High

# Test for differences in csi_past bins
emm <- emmeans(mod, specs = "csi_past", mode = "df.error")
summary(emm, infer = TRUE)

# Test for difference in fire recent within fire frq
emm <- emmeans(mod, ~ fire_recent | fire_frq, mode = "df.error")
contrast(emm, method = "pairwise") # Nothing

tab <- xtabs(n ~ csi_past_bin + admix, data = prop_df)
chisq.test(tab)

set.seed(1)

# observed difference: intermediate vs max of others
obs <- with(bin_df, {
  p_int <- mean(admix == "Admixed" & csi_past_bin == "(0.544,0.686]") /
           mean(csi_past_bin == "(0.544,0.686]")
  p_oth <- mean(admix == "Admixed" & csi_past_bin != "(0.544,0.686]") /
           mean(csi_past_bin != "(0.544,0.686]")
  p_int - p_oth
})

perm <- replicate(10000, {
  perm_bin <- sample(bin_df$csi_past_bin)
  p_int <- mean(bin_df$admix == "Admixed" & perm_bin == "(0.544,0.686]") /
           mean(perm_bin == "(0.544,0.686]")
  p_oth <- mean(bin_df$admix == "Admixed" & perm_bin != "(0.544,0.686]") /
           mean(perm_bin != "(0.544,0.686]")
  p_int - p_oth
})

p_value <- mean(perm >= obs)
p_value

```


```{r}
# Check high outlier
admix_df <- read_csv(here("analysis", "admixture", "outputs", "Q8.csv"))
outlier <- 
  mod_df %>% 
  filter(fire_recent == "Burned") %>% 
  filter(Ho == max(Ho)) %>% 
  select(Q, fire_recent, SampleID) 

q8 <- read_table("/mnt/sda1/Anusha/GitHub/ccgpscelop/analysis/admixture/outputs/58-Sceloporus_pruned_0.6_thinned_10kb_chr.8.Q", col_names = FALSE)
id <- read_table("/mnt/sda1/Anusha/GitHub/ccgpscelop/analysis/admixture/outputs/58-Sceloporus_pruned_0.6_thinned_10kb_chr.fam", col_names = FALSE)
admix_df %>% filter(SampleID %in% outlier$SampleID)
q8[which(id$X2 %in% outlier$SampleID),]
```

```{r}
pop_df <- 
  get_pops() %>%
  left_join(get_coords(sf = TRUE)) %>% 
  st_as_sf() %>%
  st_transform(3310) %>%
  bind_cols(st_coordinates(.)) %>%
  st_drop_geometry() %>%
  group_by(X, Y) %>%
  # Take the mean but it doesn't matter since all samples in a pop have the same cluster
  summarise(cluster = mean(as.numeric(cluster))) %>%
  mutate(cluster = factor(cluster)) %>%
  rename(x = X, y = Y) %>%
  ungroup() %>%
  right_join(mod_sf_unique) %>%
  st_as_sf()

# Color scales
group1 <- c("8", "6", "2")
group2 <- c("4", "3", "7")
group3 <- c("1", "5")
hiroshige_cols <- rev(c("#042346", "#295580", "#6bb2cc", "#9cd3dd", "#fecd6d", "#f7a957", "#ec7b4c", "#f04f32"))
cs1 <- hiroshige_cols[1:3]
names(cs1) <- group1
cs2 <- hiroshige_cols[4:6]
names(cs2) <- group2
cs3 <- hiroshige_cols[7:8]
names(cs3) <- group3
pop_cols <- c(cs1, cs2, cs3)

# Make residual plots for model
plot_mod_pop <- function(x, y, df, model, 
                     mod_vars = c("NDVI","gHM","bio1","csi_past","glacier","tmean_dif", "pfs_replac", "pfs_mixed", "pfs_surfac", "fri", "vdep","Q"), residual = TRUE) {

  # Calculate residuals for y variable given x variable
  if (residual) df[[y]] <- residuals(model) + coef(model)[[x]] * as.numeric(scale(df[[x]]))

  # Extract beta and p-value for x variable
  model_summary <- broom::tidy(model)
  # Convert to scientific notation for plotting
  beta <- formatC(model_summary$estimate[model_summary$term == x], format = "e", digits = 2)
  pval <- model_summary$p.value[model_summary$term == x]
  pval <- ifelse(pval < 0.001, "< 0.001", paste0("= ", round(pval, 3)))

  plt <-
    ggplot(df, aes(x = .data[[x]], y = .data[[y]])) +
    geom_point(aes(col = cluster), size = 2, alpha = 0.8) +
    geom_smooth(method = "lm", colour = "black") +
    scale_color_manual(values = pop_cols) +
    annotate("text", x = Inf, y = Inf, hjust = 1.1, vjust = 1.5,
             label = paste0("β = ", beta, "\n", "p ", pval), size = 5) +
    xlab(stringr::str_to_sentence(make_pretty_names(x))) +
    theme_classic() +
    theme(legend.position = "none",
          axis.title = element_text(size = 16),
          axis.text  = element_text(size = 14))

  if (residual) {
    plt + labs(y = "Partial heterozygosity")
  } else {
    plt + labs(y = "Observed heterozygosity")
  }

}

# Isolate significant vs all variables
sig_vars <- c("csi_past", "Q", "tmean_dif", "bio1", "pfs_surfac", "log_fri")
mod_vars <- c("NDVI", "gHM", "bio1", "csi_past", "glacier", "tmean_dif", "pfs_replac", "pfs_mixed", "pfs_surfac", "log_fri", "vdep", "Q")

# A: population plot
pop_label <- 
  st_drop_geometry(pop_df) %>% 
  group_by(cluster) %>%
  summarize(x = median(x), y = median(y)) %>%
  st_as_sf(coords = c("x", "y"), crs = 3310) 

pop_plot <-
  ggplot() +
  geom_sf(data = ca, col = NA) +
  geom_sf(data = pop_df, aes(fill = cluster), cex = 2, pch = 21, col = "NA")+
  #geom_point(data = pop_label, aes(x = x, y = y, fill = cluster), cex = 7, pch = 21, alpha = 0.7) +
  #geom_text(data = pop_label, aes(x = x, y = y, label = cluster), cex = 5) +
  labs(col = "Population") +
  scale_fill_manual(values = pop_cols)  +
  theme_void() +
  theme(
    legend.position = "none"
  )
# B: partial residuals 
plts_residual_y <- purrr::map(
  sig_vars, 
  # Use unscaled data for plotting
  ~plot_mod_pop(.x, "Ho", pop_df, model = final_model, residual = TRUE)
)

# B: Raw values
plts_raw_y <- purrr::map(
  sig_vars, 
  # Use unscaled data for plotting
  ~plot_mod_pop(.x, "Ho", pop_df, model = final_model, residual = FALSE)
)

# C: predicted vs observed 
pred_vs_obs <-
  # Use unscaled data for plotting
  ggplot(pop_df, aes(x = predict(final_model), y = Ho)) +
  geom_point(aes(col = cluster), shape = 16, size = 2) +
  geom_smooth(method = "lm", colour = "black") +
  scale_color_manual(values = pop_cols) +
  theme_classic() +
  labs(x = "Predicted heterozygosity", y = "Observed heterozygosity") +
  ggpubr::stat_cor(size = 6) +
  theme(legend.position = "none",
        axis.title = element_text(size = 16),
        axis.text  = element_text(size = 14))

pred_vs_obs_adj <- cowplot::plot_grid(pred_vs_obs, NULL, nrow = 1, rel_widths = c(0.9, 0.1))

plt2_residual <- cowplot::plot_grid(plotlist = plts_residual_y, ncol = 2)
plt2_raw <- cowplot::plot_grid(plotlist = plts_raw_y, ncol = 2)
plt1 <- cowplot::plot_grid(plotlist = list(pop_plot, pred_vs_obs_adj), ncol = 1,
                           labels = c("A", "C"), rel_heights = c(0.6, 0.4), label_size = 18)

png(here(plotpath, "figure_Ho_residual_pop.png"), height = 3.45*4*300*0.9, width = 4.1*4*300*0.9, res = 300)
cowplot::plot_grid(plt1, plt2_residual, ncol = 2,labels = c("", "B"),  rel_widths = c(0.4, 0.6), align = "h", label_size = 18)
dev.off()

png(here(plotpath, "figure_Ho_raw_pop.png"), height = 3.5*4*300*0.9, width = 4.1*4*300*0.9, res = 300)
cowplot::plot_grid(plt1, plt2_raw, ncol = 2,labels = c("", "B"),  rel_widths = c(0.4, 0.6), align = "h", label_size = 18)
dev.off()
```

```{r}
# Plot points on map colored by variable
sig_vars <- c("csi_past", "Q", "tmean_dif", "bio1", "pfs_surfac", "log_fri")
plot_points_map <- function(var, points_sf) {
  plt1 <- 
    ggplot() +
    geom_sf(data = ca_proj, fill = "gray90", colour = NA, size = 0.5) +
    geom_sf(data = points_sf, aes_string(col = var), size = 2) +
    scale_met_cont(points_sf[[var]], aesthetic = "colour", extend_scale = 0.1) +
    theme_void() +
    labs(title = stringr::str_to_sentence(make_pretty_names(var)), col = "") +
    theme(legend.position = "top", plot.title = element_text(hjust = 0.5))

  # Plot variable vs Ho
  plt2 <- plot_mod(var, "Ho", points_sf, model = final_model, color_var = "x", residual = FALSE) + theme(plot.title = element_blank())

  cowplot::plot_grid(plt2, plt1, nrow = 1, align = "hv")

}
png(here(plotpath, "significant_variable_maps.png"), width = 5*2, height = 5.5*length(sig_vars), res = 300, unit = "in")
plt_list <- map(sig_vars, ~(plot_points_map(.x, mod_sf_unique)))
cowplot::plot_grid(plotlist = plt_list, ncol = 1, align = "hv")
dev.off()
```

```{r}
# For each variable, build a model with all other variables and compare the residuals to the full model residuals
calc_var_resid <- function(x, mod_vars, df){
  other_vars <- setdiff(mod_vars, x)
  sub_f <- reformulate(other_vars, response = "Ho")
  f <- reformulate(mod_vars, response = "Ho")
  sub_model <- gls(sub_f, correlation = corExp(form = ~ x + y, nugget = FALSE), data = df)
  full_model <- gls(f, correlation = corExp(form = ~ x + y, nugget = FALSE), data = df)
  #sub_model <- lm(sub_f, data = df)
  #full_model <- lm(f, data = df)

  # Difference in fit
  sub_pred <- predict(sub_model)
  full_pred <- predict(full_model)
  pred_change <- full_pred - sub_pred

  return(tibble(
    pred_change = pred_change, 
    variable = x, 
    x = df$x, 
    y = df$y)
  )
}

sig_vars <- c("csi_past", "Q", "tmean_dif", "bio1", "pfs_surfac", "log_fri")

var_resid <- map(sig_vars, ~calc_var_resid(.x, mod_vars, mod_sf_unique), .progress = TRUE)

var_resid_df <- 
  var_resid %>% 
  bind_rows() %>% 
  st_as_sf(coords = c("x", "y"), crs = 3310) %>%
  group_by(variable) %>%
  mutate(pred_change_scaled = scale(pred_change)[,1]) 

pair_plot <- function(x, label = ""){
  y = "pred_change"
  change <- var_resid_df %>% filter(variable == x) %>% pull(y)
  df <- mod_sf_unique %>% mutate(change = change)
  plt_scatter <-
    ggplot(df, aes_string(x = x, y = "Ho")) +
    geom_point(aes(col = change), alpha = 0.7) +
    scico::scale_color_scico(palette = "roma", midpoint = 0, direction = -1) +
    geom_smooth(method = "lm", colour = "black") +
    theme_classic() +
    labs(x = make_pretty_names(x), y = "Observed Heterozygosity", col = y) +
    theme(legend.position = "none", plot.title = element_text(hjust = 0.5))

  plt_map <-
    ggplot(df) +
    geom_sf(data = ca_proj, fill = "white", colour = "gray", size = 0.5) +
    geom_sf(aes(col = change), pch = 16, size = 1.5, alpha = 0.9) +
    scico::scale_color_scico(palette = "roma", midpoint = 0, direction = -1) +
    theme_void() +
    labs(col = "Δ Predicted\nHeterozygosity") +
    theme(legend.position = "right", plot.title = element_text(hjust = 0.5))

  cowplot::plot_grid(plt_scatter, plt_map, nrow = 1, labels = c(label, ""), align = "h", rel_widths = c(0.45, 0.55))
}

png(here(plotpath, "variable_importance_map.png"), width = 3.5*2*2, height = 3*length(sig_vars)/2, units = "in", res = 300)
plts <- map(1:length(sig_vars), ~pair_plot(sig_vars[.x], label = LETTERS[.x]))
cowplot::plot_grid(plotlist = plts, ncol = 2)
dev.off()
```

Maps of significant variables
```{r}
plot_var <- function(x){
  ggplot() +
    geom_sf(data = ca_proj, fill = "gray90", colour = NA, size = 0.5) +
    geom_sf(data = mod_sf_unique, aes_string(fill = x), size = 2, shape = 21, color = "#00000087") +
    scale_met_cont(mod_sf_unique[[x]], aesthetic = "fill", extend_scale = 0.1) +
    theme_void() +
    labs(title = stringr::str_to_sentence(make_pretty_names(x)), fill = "") +
    theme(plot.title = element_text(hjust = 0.5))
}

sig_vars <- c("csi_past", "Q", "tmean_dif", "pfs_replac", "pfs_surfac", "log_fri")
pdf(here(plotpath, "significant_variable_maps_sf.pdf"), width = 4*3, height = 4*2)
plts <- map(sig_vars, ~plot_var(.x))
cowplot::plot_grid(plotlist = plts, nrow = 2)
dev.off()
```



# Figures for MVZ

```{r}
library(tidyverse)
library(here)
library(ggpubr)
library(rstatix)
outpath = here("analysis", "genetic_diversity", "outputs")
plotpath = here("analysis", "genetic_diversity", "plots")
mod_df <- read_csv(here(outpath, "model_df.csv"))
gg_df <- 
  mod_df %>%
  mutate(fire_recent = round(fire_recent, 0)) %>%
  mutate(fire_recent_cat = case_when(
    fire_recent == 0 ~ "No fire since 1984",
    fire_recent > 0 ~ "≥1 fire since 1984"
  )) %>%
  mutate(fire_frq = case_when(
    grepl("High Frequency", fire_frg) ~ "High",
    grepl("Intermediate Frequency", fire_frg) ~ "Intermediate",
    fire_frg == "Infrequent Fire Regime" ~ "Low/No Fire",
    fire_frg == "No Fire Regime" ~ "Low/No Fire",
    TRUE ~ NA_character_
  )) %>%
  mutate(fire_frg = gsub("\\+ ", "\\+\n", fire_frg)) %>%
  mutate(
    fire_frq = factor(fire_frq, levels = rev(c("High", "Intermediate", "Low/No Fire"))),
    fire_recent_cat = factor(fire_recent_cat, levels = c("No fire since 1984", "≥1 fire since 1984")),
    fire_frg = factor(fire_frg, levels = rev(c("High Frequency +\nHigh Severity", "High Frequency +\nLow/Moderate Severity", "Intermediate Frequency +\nHigh Severity", "Intermediate Frequency +\nLow/Moderate Severity", "Infrequent Fire Regime", "No Fire Regime")))
  )

common_evt <- gg_df %>% count(evt) %>% arrange(desc(n)) %>% filter(n > 10) %>% pull(evt)
gg_df <- gg_df %>% mutate(evt_simple = ifelse(evt %in% common_evt, evt, "Other"))

mod_sf_unique_scaled$partial_resid_csi <-
  resid(gls_model1) +
  coef(gls_model1)["csi_past"] * mod_sf_unique_scaled$csi_past

b <- coef(gls_model1poly)
mod_sf_unique_scaled$partial_resid_csi_poly <-
  resid(gls_model1poly) +
  b["poly(csi_past, 2, raw = TRUE)1"] * mod_sf_unique_scaled$csi_past +
  b["poly(csi_past, 2, raw = TRUE)2"] * mod_sf_unique_scaled$csi_past^2

# Calc pseudo R2 for poly model
cor(predict(gls_model1poly), mod_sf_unique$Ho) # 0.72

plt_poly <- 
  ggplot(mod_df, aes(x = csi_past, y = Ho)) +
  geom_point(aes(col = csi_past)) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2, raw = TRUE)) +
  theme_classic() +scale_color_viridis_c(option = "turbo") +
  labs(x = "Paleoclimate stability", y = "Heterozygosity") +
  theme(
    axis.title = element_text(size = 16),
    axis.text  = element_text(size = 14)
  )

# Predicted vs observed plot
pred_obs_plt <-
  ggplot(mod_sf_unique_scaled, aes(x = predict(gls_model1poly), y = Ho)) +
  geom_point( size = 2) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black") +
  geom_smooth(method = "lm", color = "blue") +
  labs(x = "Predicted heterozygosity", y = "Observed heterozygosity") +
  theme_classic() +
  stat_cor(method = "pearson", label.x.npc = "left", label.y.npc = "top", size = 6) +
  theme(
    axis.title = element_text(size = 16),
    axis.text  = element_text(size = 14)
  )

gls_model_simple <- gls(Ho ~ poly(csi_past, 2) + tmean_dif + bio1 + fire_frg + fire_recent, correlation = corExp(form = ~ x + y, nugget = FALSE), data = mod_sf_unique_scaled, method = "ML")
cor(predict(gls_model_simple), mod_sf_unique$Ho) # 0.72

lm(Ho ~ x + y, data = mod_sf_unique_scaled) %>% summary() # R2 = 0.35
lm(Ho ~ poly(csi_past, 2) + tmean_dif + bio1 + evt + fire_frg, data =mod_sf_unique_scaled) %>% summary() # R2 = 50

lm(Ho ~ poly(csi_past, 2) + tmean_dif + bio1 + evt + fire_frg + x + y, data =mod_sf_unique_scaled) %>% summary() # even including x and y, variables are still significant

map_csi_past <-
  ggplot(mod_sf_unique) +
  geom_sf(data = ca_proj, fill = "gray90", colour = NA, size = 2) +
  geom_sf(aes(col = csi_past), size = 2) +
  scale_met_cont(mod_sf$csi_past, aesthetic = "col") +
  labs(col = "Paleoclimate\nstability") +
  theme_void() +
  theme(
    legend.position = "inside",
    legend.justification = c(1, 1),
    legend.position.inside = c(0.9, 0.95),
    legend.title = element_text(size = 14),
    legend.text  = element_text(size = 12)
  )

mod_sf_unique$partial_resid_csi_poly <- mod_sf_unique_scaled$partial_resid_csi_poly
model_summary <- broom::tidy(gls_model1poly)

b1 <- model_summary %>%
  dplyr::filter(term == "poly(csi_past, 2, raw = TRUE)1")
b2 <- model_summary %>%
  dplyr::filter(term == "poly(csi_past, 2, raw = TRUE)2")

beta1 <- formatC(b1$estimate, format = "e", digits = 2)
beta2 <- formatC(b2$estimate, format = "e", digits = 2)

p1_lab <- if (b1$p.value < 0.001) {
  "p[1] < 0.001"
} else {
  paste0("p[1] == ", round(b1$p.value, 3))
}

p2_lab <- if (b2$p.value < 0.001) {
  "p[2] < 0.001"
} else {
  paste0("p[2] == ", round(b2$p.value, 3))
}

line1 <- paste0("beta[1] == ", beta1, "*','~", p1_lab)
line2 <- paste0("beta[2] == ", beta2, "*','~", p2_lab)

label_txt <- paste0("atop(", line1, ",", line2, ")")

plt_poly_partial <- 
  ggplot(mod_sf_unique, aes(x = csi_past, y = partial_resid_csi_poly)) +
  geom_point(aes(col = csi_past)) +
  geom_smooth(
    method = "lm",
    formula = y ~ poly(x, 2, raw = TRUE),
    col = "black"
  ) +
  theme_classic() +
  scale_met_cont(mod_sf$csi_past, aesthetic = "col") +
  annotate(
    "text", x = Inf, y = Inf, hjust = 1, vjust = 1,
    label = label_txt, parse = TRUE, size = 5
  ) +
  labs(
    x = "Paleoclimate stability",
    y = "Partial residual heterozygosity"
  ) +
  theme(
    axis.title = element_text(size = 16),
    axis.text  = element_text(size = 14),
    legend.position = "none"
  )


csi_past_plt <- cowplot::plot_grid(plt_poly_partial, map_csi_past, nrow = 1, rel_widths = c(1, 0.8))

q_plt <- ggplot(filter(mod_sf_unique, Q > 0.01), aes(x = csi_past, y = Q)) +
  geom_point(aes(col = csi_past)) +
  geom_smooth(method = "lm", col = "black",  formula = y ~ poly(x, 2, raw = TRUE)) +
  theme_classic() +
  scale_met_cont(mod_sf$csi_past, aesthetic = "col") +
  labs(x = "Paleoclimate stability", y = "Inbreeding coefficient (Q)") +
  theme(
    axis.title = element_text(size = 16),
    axis.text  = element_text(size = 14)
  )

range(mod_sf_unique$Q)

mod_sf_unique <- mod_sf_unique %>%
  mutate(
    admix = if_else(Q > 0.001, "Admixed", "Non-admixed"),
    csi_past_bin = cut(csi_past, breaks = 8, include.lowest = TRUE)
  )

# get midpoints for labels
bin_levels <- levels(mod_sf_unique$csi_past_bin)
bin_mids <- sapply(bin_levels, function(s) {
  nums <- as.numeric(gsub("\\(|\\]|\\[|\\)", "", unlist(strsplit(s, ","))))
  mean(nums)
})
bin_lab <- sprintf("%.2f", bin_mids)  # change decimals here

count_df <- mod_sf_unique %>%
  count(csi_past_bin, admix, name = "n") %>%
  group_by(csi_past_bin) %>%
  mutate(prop = n / sum(n),
         admix = factor(admix, levels = c("Non-admixed", "Admixed"))) %>%
  ungroup()

count_df_fire <- mod_sf_unique %>%
  count(fire_frg, admix, name = "n") %>%
  group_by(fire_frg) %>%
  mutate(prop = n / sum(n),
         admix = factor(admix, levels = c("Non-admixed", "Admixed"))) %>%
  ungroup()

q_plt_prop <- 
  ggplot(count_df, aes(x = csi_past_bin, y = prop, fill = admix)) +
  geom_col() +
  scale_x_discrete(labels = bin_lab) +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic() +
  scale_fill_manual(values = c("Non-admixed" = "tomato2", "Admixed" = "skyblue2")) +
  labs(x = "Binned paleoclimate stability", y = "Proportion of individuals", fill = "") +
  theme(
    legend.position = "top",
    axis.title = element_text(size = 18),
    axis.text  = element_text(size = 16),
    legend.text = element_text(size = 16)
  )

total_count <- count_df %>%
  group_by(csi_past_bin) %>%
  summarise(total = sum(n)) 

q_plt_count <- 
  ggplot(count_df, aes(x = csi_past_bin)) +
  geom_col(aes(y = prop, fill = admix)) +
  # Add text with counts to bars
  #geom_text(data = total_count, aes(label = total), y = 0.9, size = 6, color = "white") +
  scale_x_discrete(labels = bin_lab) +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic() +
  scale_fill_manual(values = c("Non-admixed" = "red2", "Admixed" = "skyblue2")) +
  labs(x = "Binned paleoclimate stability", y = "Number of individuals", fill = "") +
  theme(
    legend.position = "top",
    axis.title = element_text(size = 18),
    axis.text  = element_text(size = 16),
    legend.text = element_text(size = 16)
  )

pdf(here(plotpath, "mvz.pdf"), width = 5.5, height = 5)
q_plt_count
dev.off()

pdf(here(plotpath, "mvz.pdf"), width = 9, height = 4.5)
q_plt_count
csi_past_plt
q_plt
dev.off()

mod_sf_unique$partial_resid_csi_notpoly <-
  resid(gls_model1) +
  coef(gls_model1)["csi_past"] * mod_sf_unique_scaled$csi_past

plt_notpoly_raw <- 
  ggplot(mod_sf_unique, aes(x = csi_past, y = Ho)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x) +
  labs(x = "Paleoclimate stability", y = "Partial residual heterozygosity") +
  theme_classic() + 
  theme(
    axis.title = element_text(size = 16),
    axis.text  = element_text(size = 14)
  )
plt_poly_raw <- 
  ggplot(mod_sf_unique, aes(x = csi_past, y = Ho)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2, raw = TRUE)) +
  labs(x = "Paleoclimate stability", y = "Partial residual heterozygosity") +
  theme_classic() + 
  theme(
    axis.title = element_text(size = 16),
    axis.text  = element_text(size = 14)
  )


plt_notpoly <- 
  ggplot(mod_sf_unique, aes(x = csi_past, y = partial_resid_csi_notpoly)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x) +
  labs(x = "Paleoclimate stability", y = "Partial residual heterozygosity") +
  theme_classic() + 
  theme(
    axis.title = element_text(size = 16),
    axis.text  = element_text(size = 14)
  )
plt_poly <- 
  ggplot(mod_sf_unique, aes(x = csi_past, y = partial_resid_csi_poly)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2, raw = TRUE)) +
  labs(x = "Paleoclimate stability", y = "Partial residual heterozygosity") +
  theme_classic() + 
  theme(
    axis.title = element_text(size = 16),
    axis.text  = element_text(size = 14)
  )


mod_sf_unique$partial_resid_tmean_dif <-
  resid(gls_model1poly) +
  coef(gls_model1poly)["tmean_dif"] * mod_sf_unique_scaled$tmean_dif

# Get coefficient and p-value for annotation
model_summary <- broom::tidy(gls_model1poly)
beta <- formatC(model_summary$estimate[model_summary$term == "tmean_dif"], format = "e", digits = 2)
pval <- model_summary$p.value[model_summary$term == "tmean_dif"] %>% round(2)

tmean_dif_plt <-
  ggplot(mod_sf_unique, aes(x = tmean_dif, y = partial_resid_tmean_dif)) +
  geom_point(col = "orange2") +
  geom_smooth(method = "lm", col = "black") +
  theme_classic() +
  labs(x = "Contemporary temperature change", y = "Partial residual heterozygosity") +
  annotate("text", x = Inf, y = Inf, hjust = 1.1, vjust = 1.5,
           label = paste0("beta == ", beta, "*','~p == ", pval), parse = TRUE, size = 6) +
  theme(
    axis.title = element_text(size = 16),
    axis.text  = element_text(size = 14),
    legend.position = "none"
  )
pdf(here(plotpath, "mvz.pdf"), width = 5, height = 5)
tmean_dif_plt
dev.off()
```

```{r}
# -------------------------------------------------------------------
# Recreate tuk_df from gls_model1poly (this is what you deleted)
# -------------------------------------------------------------------

# helper to clean factor labels consistently
clean_lvl <- function(x) {
  x %>%
    as.character() %>%
    str_replace_all("[\r\n]+", " ") %>%  # remove line breaks
    str_squish() %>%                    # collapse whitespace
    str_remove("^\\(") %>%              # drop leading "("
    str_remove("\\)$")                  # drop trailing ")"
}


# emmeans + Tukey (FDR) on the GLS model
emm_fire <- emmeans(gls_model1poly, ~ fire_frg, mode = "df.error")
tuk <- pairs(emm_fire, adjust = "fdr")

# Convert to a "tuk_df" equivalent with group1/group2 columns + stars
tuk_df <- as.data.frame(tuk) %>%
  tidyr::separate(contrast, into = c("group1", "group2"), sep = " - ", remove = FALSE) %>%
  mutate(
    stars = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*",
      TRUE ~ "ns"
    )
  )

# If you only want significant comparisons:
tuk_df <- tuk_df %>% filter(stars != "ns")

# -------------------------------------------------------------------
# Map comparisons to x positions for plotting brackets on ggplot
# -------------------------------------------------------------------
gg_df2 <- gg_df %>%
  mutate(fire_frg_clean = clean_lvl(fire_frg),
         fire_frg_clean = factor(
           fire_frg_clean,
           levels = clean_lvl(levels(fire_frg))  # keep same ordering as fire_frg
         ))

lvls_key <- levels(gg_df2$fire_frg_clean)

tuk_df2 <- tuk_df %>%
  mutate(
    group1_key = clean_lvl(group1),
    group2_key = clean_lvl(group2),
    x1 = match(group1_key, lvls_key),
    x2 = match(group2_key, lvls_key),
    xmin = pmin(x1, x2),
    xmax = pmax(x1, x2),
    xmid = (xmin + xmax) / 2
  ) %>%
  filter(!is.na(xmin), !is.na(xmax))


# -------------------------------------------------------------------
# Plot (kept): fire_frg vs Ho with significant Tukey brackets
# -------------------------------------------------------------------
mod_sf_unique$Ho_fire_frg_partial <- resid(gls_model1poly) +
  coef(gls_model1poly)["fire_frgInfrequent Fire Regime"] * (mod_sf_unique_scaled$fire_frg == "Infrequent Fire Regime") +
  coef(gls_model1poly)["fire_frgIntermediate Frequency + Low/Moderate Severity"] * (mod_sf_unique_scaled$fire_frg == "Intermediate Frequency + Low/Moderate Severity") +
  coef(gls_model1poly)["fire_frgIntermediate Frequency + High Severity"] * (mod_sf_unique_scaled$fire_frg == "Intermediate Frequency + High Severity") +
  coef(gls_model1poly)["fire_frgHigh Frequency + Low/Moderate Severity"] * (mod_sf_unique_scaled$fire_frg == "High Frequency + Low/Moderate Severity") +
  coef(gls_model1poly)["fire_frgHigh Frequency + High Severity"] * (mod_sf_unique_scaled$fire_frg == "High Frequency + High Severity")
mod_sf_unique <-  
  mod_sf_unique %>%
  mutate(fire_frg_lb = gsub("\\+ ", "\\+\n", as.character(fire_frg))) %>%
  mutate(
    fire_frg_lb = factor(fire_frg_lb, levels = rev(c("High Frequency +\nHigh Severity", "High Frequency +\nLow/Moderate Severity", "Intermediate Frequency +\nHigh Severity", "Intermediate Frequency +\nLow/Moderate Severity", "Infrequent Fire Regime", "No Fire Regime")))
  )
unique(mod_sf_unique$fire_frg_lb)


# y positions for brackets (stack them)
y_base <- max(mod_sf_unique$Ho_fire_frg_partial, na.rm = TRUE) - 0.0001
y_step <- 0.08 * diff(range(mod_sf_unique$Ho_fire_frg_partial, na.rm = TRUE))  # tweak if needed

tuk_df2 <- tuk_df2 %>%
  arrange(xmin, xmax) %>%
  mutate(
    y = y_base + row_number() * y_step,
    y_tick = y - 0.25 * y_step
  )


bxp2 <-
  ggplot(mod_sf_unique, aes(x = fire_frg_lb, y = Ho_fire_frg_partial)) +
  geom_boxplot(col = NA) +
  geom_boxplot(data = filter(mod_sf_unique, fire_frg != "Intermediate Frequency + High Severity")) +
  geom_boxplot(data = filter(mod_sf_unique, fire_frg == "Intermediate Frequency + High Severity"), col = "tomato2") +
  geom_jitter(data = filter(mod_sf_unique, fire_frg != "Intermediate Frequency + High Severity"), width = 0.2, alpha = 0.5) +
  geom_jitter(data = filter(mod_sf_unique, fire_frg == "Intermediate Frequency + High Severity"), col = "tomato2", width = 0.2, alpha = 0.5) +
  theme_classic() +
  labs(x = "Historical fire regime", y = "Partial residual heterozygosity") +
  geom_segment(data = tuk_df2,
               aes(x = xmin, xend = xmax, y = y, yend = y),
               inherit.aes = FALSE, linewidth = 0.6) +
  geom_segment(data = tuk_df2,
               aes(x = xmin, xend = xmin, y = y_tick, yend = y),
               inherit.aes = FALSE, linewidth = 0.6) +
  geom_segment(data = tuk_df2,
               aes(x = xmax, xend = xmax, y = y_tick, yend = y),
               inherit.aes = FALSE, linewidth = 0.6) +
  geom_text(data = tuk_df2,
            aes(x = xmid, y = y - 0.00005, label = stars),
            inherit.aes = FALSE, vjust = 0, size = 8) +
  coord_cartesian(clip = "off") +
  theme(
    axis.title = element_text(size = 16),
    axis.text  = element_text(size = 14),
    axis.text.x = element_text(angle = 20, hjust = 1, size = 14),
    plot.margin = margin(5.5, 5.5, 5.5, 25)
  )

# If you want to save:
ggsave(here(plotpath, "fire_frg_tukey_gls.pdf"), bxp2, width = 10, height = 6)

```

```{r}
df = mod_sf_unique
df <- df %>% mutate(admixed = ifelse(Q > 0.01, 1, 0))

# admixed is binomial so use glm with family = binomial
admix_fit <- glm(admixed ~ csi_past + glacier + gHM + bio1 + tmean_dif + evt + fire_recent * fire_frq, data = df, family = "binomial")
car::Anova(admix_fit, type = 2)
admix_fit %>% broom::tidy() %>% filter(p.value < 0.05)

# Emmeans check on fire_recent
emmeans(admix_fit, ~ fire_recent | fire_frq, mode = "df.error") |>
  pairs() %>%  
  summary(infer = TRUE)
```


# DUMP
```{r}
df = mod_df

med_fit <- lm(Ho ~ veg_PC1 + veg_PC2 + veg_PC3 + bio1 + csi_past + glacier + gHM + tmean_dif + fire_severity + fire_recent +, data = df)
summary(med_fit)
```

```{r}
pdf(here(plotpath, "fire_groupveg_plots.pdf"), width = 5, height = 5)
ggplot(data = mod_sf_unique, aes(x = fire_severity, y = Ho)) +
  geom_point(aes(col = groupveg)) +
  geom_smooth(method = "lm") +
  theme_classic() 

ggplot(data = mod_sf_unique, aes(x = fire_severity, y = Ho)) +
  geom_point(aes(col = evt)) +
  geom_smooth(method = "lm") +
  theme_classic() 

# Boxplot with significance comparisons using ggpubr
ggboxplot(mod_sf_unique, x = "groupveg", y = "Ho",
          color = "groupveg", palette = "jco") +
  stat_compare_means(method = "anova") +
  stat_compare_means(label = "p.format", method = "t.test",
                     ref.group = "shrubland", label.y.npc = 0.9) +
  theme_classic()
dev.off()
```
