# Libraries
```{r}
library("here")
library("tidyverse")
library("ggpubr")
library("terra")
library("sf")
library("gridExtra")
library("rpaleoclim")
library("MuMIn")
library("cowplot")
library("performance")
library("gstat")
library("sp")
library("furrr")
library("MetBrewer")
library("nlme")
library("broom.mixed")
```

# Paths for outputs
```{r}
outpath <- here("analysis", "genetic_diversity", "outputs")
plotpath <- here("analysis", "genetic_diversity", "plots")
```

# Custom functions
```{r}
source(here("general_functions.R"))
source(here("analysis", "genetic_diversity", "functions_genetic_diversity.R"))
source(here("analysis", "genetic_diversity", "functions_paleovars.R"))
```

# California polygon for plotting
```{r}
ca <- get_ca()
ca_proj <- ca %>% st_transform(3310)
```

# Coordinates
```{r}
coords <- get_coords(sf = TRUE)
coords_proj <- coords %>% st_transform(crs = st_crs(ca_proj))
```

# Heterozygosity data
```{r}
# Get genome-wide heterozygosity
het <- get_het("58-Sceloporus_chr.het") 

# Make sure you have the expected number of samples
stopifnot(nrow(het) == 259)

# Count number of sites used to calculate heterozygosity
mean(het$`N(NM)`, na.rm = TRUE) # 59955192
```

# Hypothesis 1: Genetic diversity explained by paleoclimate stability

```{r}
# Get current climate data (NOTE: not using paleoclim data because they use CHELSA V1.2 instead of V2.1)
bioclim <- rast(here("data", "env", "california_chelsa_bioclim_1981-2010_V.2.1.tif"))
cur <- bioclim[["CHELSA_bio1_1981-2010_V.2.1"]]
names(cur) <- "cur"

# Get LGM climate data (NOTE: not using paleoclim data because they use CHELSA V1.2 instead of V2.1)
lgm <- rast(here("data", "env", "california_chelsa_pmip_ccsm4.tif"))
names(lgm) <- "lgm"

# Get paleoclimate data (BIO1) from rpaleoclim and convert from C/10 to C
# Using time periods back until about ~ 1 mya (when Sceloporus diverged in S. CA)
#
# code	period	bp	source
# cur	Current (1979 – 2013)		CHELSA
# lh	Late Holocene: Meghalayan	4.2-0.3 ka	Fordham  et al. 2017
# mh	Mid Holocene: Northgrippian	8.326-4.2 ka	Fordham  et al. 2017
# eh	Early Holocene: Greenlandian	11.7-8.326 ka	Fordham  et al. 2017
# yds	Pleistocene: Younger Dryas Stadial	12.9-11.7 ka	Fordham  et al. 2017
# ba	Pleistocene: Bølling-Allerød	14.7-12.9 ka	Fordham  et al. 2017
# hs1	Pleistocene: Heinrich Stadial 1	17.0-14.7 ka	Fordham  et al. 2017
# lgm	Pleistocene: Last Glacial Maximum	ca. 21 ka	CHELSA
# lig	Pleistocene: Last Interglacial	ca. 130 ka	Otto-Bliesner et al. 2006
# mis19	Pleistocene: MIS19	ca. 787 ka	Brown et al. 2018
paleovars <- get_paleovars(c("lh", "mh", "eh", "yds", "ba", "hs1", "lig", "mis19"))
paleobio1 <- map(paleovars, ~.x$bio_1) %>% rast()/10

# Stack CHELSA rasters
chelsabio1 <- list(cur, lgm)

# Resample the higher resolution data (CHELSA) to match the lower resolution (rpaleoclim)
# This is the most conservative approach to avoid over-interpreting fine-scale variation 
csi_stack <- 
  c(as.list(chelsabio1), as.list(paleobio1)) %>% 
  map(~resample(.x, paleobio1[[1]]), method = "bilinear") %>% 
  rast()

# Calculate climate stability based on the standard deviation over the stack of rasters
# Don't remove NAs, use imputation later (NA values are at some points along the coast)
# If you remove NAs you will get some outliers because some time periods have data and others do not
csi_past <- app(csi_stack, sd, na.rm = FALSE)

# Scale CSI raster from 0 to 1 and subtract from 1 to get higher values = higher stability
csi_past[] <- 1 - scales::rescale(csi_past[], to = c(0, 1), na.rm = TRUE)

# Make a data frame with the climate stability values
paleoclim_mod_df <- 
  data.frame(
    SampleID = coords$SampleID,
    csi_past = terra::extract(csi_past, coords, ID = FALSE)[,1]
  )

# Check NA values (should be <10)
print(colSums(is.na(paleoclim_mod_df)))

# Some coastal points have NA values, so use a bilinear approach to get take the average of the four nearest raster cells
paleoclim_mod_df$csi_past <- bilinear_impute(paleoclim_mod_df$csi_past, csi_past, coords)

# Re-check NA values (should be 0)
print(colSums(is.na(paleoclim_mod_df)))
```

# Hypothesis 2: Genetic diversity explained by admixture
```{r}
# Read in admixture results produced by analysis/admixture/notebook_admixture.Rmd
q <- read_csv(here("analysis", "admixture", "outputs", "Q8.csv"))
q_r <- rast(here("analysis", "admixture", "outputs", "max_admix.tif"))

q_mod_df <- 
  q %>%
  dplyr::select(SampleID, Q) %>%
  # Calculate 1 - Q to make an "admixture score" whereby higher values = more admixture
  mutate(Q = 1 - Q)
```

# Hypothesis 3: Genetic diversity explained by current environment 
```{r}
# Read in contemporary environmental layers downloaded using the scripts in data_processing/env
bio1 <- bioclim[["CHELSA_bio1_1981-2010_V.2.1"]] # mean annual temperature
ndvi <- rast(here("data", "env", "california_ndvi_mean_2000_2020.tif")) # mean NDVI
ghm <- rast(here("data", "env", "california_global_human_modification_2016.tif")) # global human modification index

# Combine into df
env_mod_df <- data.frame(
  SampleID = coords$SampleID,
  NDVI = terra::extract(ndvi, coords, ID = FALSE)[, "NDVI"],
  gHM = terra::extract(ghm, coords, ID = FALSE)[,"gHM"],
  bio1 = terra::extract(bio1, coords, ID = FALSE)[,"CHELSA_bio1_1981-2010_V.2.1"]
)

# Check NA values
print(colSums(is.na(env_mod_df)))

# Some coastal points have NA values for gHM and bio1, so use a bilinear approach to get take the average of the four nearest raster cells
# Note: This function only imputes missing values, not all values
env_mod_df$gHM <- bilinear_impute(env_mod_df$gHM, ghm, coords)
env_mod_df$bio1 <- bilinear_impute(env_mod_df$bio1, bio1, coords)

# Double check NA values (should be 0)
print(colSums(is.na(env_mod_df)))
```

# Hypothesis 4: Genetic diversity explained by glaciation
```{r}
# Read in LGM polygon provided by Michelle Koo (MVZ)
lgm_glaciers <- st_read(here("data", "env", "Glaciotect_NA_lgm", "shp", "LGM_maxextent.shp"))

# Transform to same CRS as coords
# Note: have to first intersect and then transform otherwise you get a node error
lgm_glaciers <- st_intersection(lgm_glaciers, st_transform(ca, crs = st_crs(lgm_glaciers)))
lgm_glaciers <- st_transform(lgm_glaciers, crs = st_crs(ca))

# Add a new column to coords indicating whether each point is inside the LGM glacial extent
glacier_coords <- coords %>% st_transform(st_crs(lgm_glaciers))
glacier_mod_df <- 
  coords %>%
  mutate(glacier = as.numeric(lengths(st_within(glacier_coords, lgm_glaciers)) > 0)) %>%
  st_drop_geometry()
```

# Hypothesis 5: Genetic diversity explained by recent climate change
PRISM mean temperature data from historical data: https://prism.oregonstate.edu/historical/
```{r}
# Get historical prism data (https://prism.oregonstate.edu/historical/)
# Processed by scripts in data_processing/env
get_prism <- function(year, output = "df") {
  # Create PRISM path
  prism_path <- here("data", "env", "prism", "PRISM_tmean_stable_4kmM3", year)

  # List annual file
  file <- list.files(prism_path, pattern = paste0(year, "_*.bil.bil$"), full.names = TRUE)

  # Load data
  raster <- terra::rast(file)

  if (output == "raster"){
    return(raster)
  } else {
    vals <- terra::extract(raster, st_transform(coords, st_crs(raster)), ID = FALSE)
    names(vals) <- "tmean"
    df <- data.frame(year = year, tmean = vals, SampleID = coords$SampleID)
    return(df)
  }
}

# Loop over the years of interest
historical_years <- 1895:1905 # Historical decade
current_years <- 2010:2020 # Contemporary decade

# Make dataframe of average temperature by year for each sample
tmean_his <- 
  map(
    historical_years, 
    ~get_prism(year = .x), 
    .progress =TRUE
  ) %>% 
  bind_rows()  

tmean_cur <- 
  map(
    current_years, 
    ~get_prism(year = .x), 
    .progress =TRUE
  ) %>% 
  bind_rows()  

# Calculate mean temperature for each sample over the decade
tmean_his <- 
  tmean_his %>% 
  group_by(SampleID) %>% 
  summarise(tmean_his = mean(tmean, na.rm = TRUE))

tmean_cur <- 
  tmean_cur %>% 
  group_by(SampleID) %>% 
  summarise(tmean_cur = mean(tmean, na.rm = TRUE))

# Get rasters per year and average over the decade
tmean_his_r <-
   map(historical_years, ~get_prism(year = .x, output = "raster"), .progress =TRUE) %>% 
   rast() %>% 
   mean()

tmean_cur_r <- 
  map(current_years, ~get_prism(year = .x, output = "raster"), .progress =TRUE) %>% 
  rast() %>% 
  mean()

# Calculate difference between current and historical
tmean_dif_r <- tmean_cur_r - tmean_his_r

# Crop to california and mask
tmean_dif <- crop(tmean_dif_r, ca)
tmean_dif <- mask(tmean_dif, st_transform(ca, st_crs(tmean_dif_r)))

# Calculate change
tmean_mod_df <- 
  left_join(tmean_his, tmean_cur) %>% 
  mutate(tmean_dif = tmean_cur - tmean_his) %>%
  right_join(coords) %>% 
  dplyr::select(SampleID, tmean_dif, tmean_cur)

# Check for NAs
tmean_mod_df %>% 
  summarise(across(everything(), ~sum(is.na(.))))
```

# Hypothesis 6: Genetic diversity explained by wildland fire frequency
USGS Combined wildland fire datasets for the United States and certain territories, 1800s-Present (summary rasters): https://www.sciencebase.gov/catalog/item/61aa5483d34eb622f699df85
```{r}
# Load wildfire frequency raster created using data_processing/get_wildfire.R
fire <- rast(here("data", "env", "california_fire_frequency.tif"))
names(fire) <- "fire"

# Extract fire values at sample locations
fire_mod_df <- data.frame(
  SampleID = coords$SampleID,
  fire = terra::extract(fire, coords, ID = FALSE)[, "fire"]
)

# Check NA values
print(colSums(is.na(fire_mod_df)))
```

# Combine rasters into a raster stack
This can take a while because of the fire raster
```{r}
env_list <- 
  list(
    NDVI = ndvi,
    bio1 = bio1,
    gHM = ghm,
    tmean_dif = tmean_dif,
    csi_past = csi_past, 
    Q = q_r,
    fire = fire
  )

# Project all rasters to EPSG:3310 and mask with CA
env_list <- imap(env_list, ~{
  r <- terra::project(.x, "epsg:3310", method = "bilinear")
  r <- mask(r, ca_proj)
  names(r) <- .y
  return(r)
}, .progress = TRUE)

# Output path for model rasters
path <- here("data", "env", "model_rasters")

# Make model rasters dir
dir.create(path, showWarnings = FALSE)

# Write out rasters
iwalk(env_list, ~writeRaster(.x, here(path, paste0(.y, ".tif")), overwrite = TRUE))

# Snap all rasters to same grid and resolution
# Note: these rasters are only used for visualization, so these changes don't affect the models
max_res <- map_dbl(env_list, ~res(.x)[1]) %>% which.max()
target_raster <- env_list[[max_res]] # coarsest resolution
resampled_list <- imap(env_list, ~{
  # Resample to match target
  terra::resample(.x, target_raster, method = "bilinear")
})

# Create raster stack
env_stack <- 
  resampled_list %>% 
  rast() 
```

# Create model data
```{r}
# Model variables
mod_vars <- c("NDVI", "gHM", "bio1", "Q", "csi_past", "glacier", "tmean_dif", "fire", "x", "y")

# Combine dataframes
mod_df <- 
  get_coords() %>% 
  left_join(het) %>%
  left_join(env_mod_df) %>%
  left_join(paleoclim_mod_df) %>%
  left_join(q_mod_df) %>%
  left_join(glacier_mod_df) %>%
  left_join(tmean_mod_df) %>%
  left_join(fire_mod_df) %>%
  dplyr::select(Ho, SampleID, all_of(mod_vars))

# Check sample number
# Note: het has 259 samples, but two of them are potentially flipped ("Scelocci_CAS213197", "Scelocci_CAS214858"), so we remove the from the model (they are not included in coords)
stopifnot(nrow(mod_df) == 257)

# Confirm no NAs:
stopifnot(all(colSums(is.na(mod_df)) == 0))

# Convert to sf
mod_sf <- 
  get_coords(sf = TRUE) %>% 
  right_join(mod_df) %>% 
  st_transform(3310)
```

```{r}
# Export model df
export_df <- 
  st_drop_geometry(mod_df) %>% 
  dplyr::select(SampleID, everything())
write_csv(export_df, here(outpath, "model_df.csv"))
```

## Check for correlation between variables
```{r}
# Check correlations
# Compute the correlation matrix
r <- cor(mod_df[, mod_vars], use = "pairwise.complete.obs")
# Set the diagonal and lower triangle to NA
r[lower.tri(r, diag = TRUE)] <- NA

# Find the pairs with correlations greater than 0.6
high_corr <- which(abs(r) > 0.6, arr.ind = TRUE)
# Print the pairs
high_corr_pairs <- data.frame(
  Var1 = rownames(r)[high_corr[, 1]],
  Var2 = rownames(r)[high_corr[, 2]],
  Correlation = r[high_corr]
)

# No highly correlated pairs (other than x/y, which doesn't matter)
high_corr_pairs
```

```{r}
#Make nice correlation plot
r <- stats::cor(mod_df[, setdiff(mod_vars, c("x", "y"))], use = "pairwise.complete.obs")
rownames(r) <- colnames(r) <- make_pretty_names(rownames(r))
r[lower.tri(r, diag = TRUE)] <- NA

library(corrplot)
png(here(plotpath, "correlation_plot.png"), width = 10*300, height = 10*300, units = "px", res = 300)
corrplot(r, method = "ellipse", type = "upper", tl.col = "black", tl.srt = 45, diag = FALSE, addCoef.col = "black", number.cex = 0.8)
dev.off()
```

# Run Ho models

```{r}
# Model formula variables
mod_vars <- c("glacier", "NDVI", "gHM", "bio1", "csi_past", "tmean_dif", "fire", "Q")

# GLS can't handle 0 distances, so take the average of the coordinates
mod_sf_unique <- 
  mod_sf %>%
  group_by(x, y) %>%
  summarise_at(c(mod_vars, "Ho"), mean, na.rm = TRUE) %>%
  ungroup()

# Scale data
mod_sf_unique_scaled <-
  mod_sf_unique %>%
  mutate(
    across(
      all_of(setdiff(mod_vars, "glacier")),
      ~ as.numeric(scale(.x))   # <- drop matrix
    )
  )

# Confirm that all glacier values are 0 and 1
stopifnot(all(mod_sf_unique_scaled$glacier %in% c(0, 1)))

# Convert glacier to factor
mod_sf_unique_scaled$glacier_factor <- as.factor(mod_sf_unique$glacier)
mod_sf_unique$glacier_factor <- as.factor(mod_sf_unique$glacier)

# Run OLS and get residuals
f <- formula(paste0("Ho ~", paste(mod_vars, collapse = " + ")))
ols_model <- lm(f, data = mod_sf_unique)
summary(ols_model)
AIC(ols_model)
mod_sf_unique$ols_resids <- residuals(ols_model)
broom::tidy(ols_model) %>% filter(p.value < 0.05) %>% arrange(desc(abs(estimate)))

# Fit variograms to the residuals
ols_variogram <- variogram(ols_resids ~ 1, data = mod_sf_unique)
ols_vgm <- map(c("Exp", "Sph", "Gau"), ~fit.variogram(ols_variogram, model = vgm(.x)))
names(ols_vgm) <- c("Exp", "Sph", "Gau")

pdf(here(plotpath, "ols_variogram.pdf"))
plot(ols_variogram, ols_vgm[["Exp"]], main = "OLS Residuals Variogram (Exponential)")
plot(ols_variogram, ols_vgm[["Sph"]], main = "OLS Residuals Variogram (Spherical)")
plot(ols_variogram, ols_vgm[["Gau"]], main = "OLS Residuals Variogram (Gaussian)")
dev.off()
# Based on visual inspection, the exponential model seems to fit best

# Fit a GLS model
# Note: nugget=FALSE assumes residuals at the same location are perfectly correlated, which should align with heterozygosity, since you don't expect to have much microscale variation in heterozygosity.
# nugget=TRUE allows unstructured microscale variance (random noise at zero distance), which can absorb spatial signal and cause predictor effects to appear insignificant. 
gls_model <- gls(f, correlation = corExp(form = ~ x + y, nugget = FALSE), data = mod_sf_unique)
summary(gls_model)
# Use type = "normalized" to get residuals decorrelated and standardized by the fitted 
# spatial covariance matrix; this removes modeled spatial structure and makes them 
# suitable for Moran's I. If you don't do this Moran's I will appear higher than it actually is.
mod_sf_unique$gls_resids <- residuals(gls_model, type = "normalized")

# Plot residuals
plot_resids <- function(resids){
  # Remove max residual because it is hard to see spatial patterns with it
  ggplot(mod_sf_unique[mod_sf_unique[[resids]] != max(mod_sf_unique[[resids]]), ]) +
    geom_sf(data = ca_proj, col = NA) +
    geom_sf(aes(col = .data[[resids]])) +
    theme_void() +
    scale_color_viridis_c(option = "turbo", end = 0.1, begin = 0.9, direction = -1) +
    labs(title = paste("Residuals:", resids))
}

pdf(here(plotpath, "all_model_residuals.pdf"), width = 9, height = 4)
plt_list <- list(
  plot_resids("ols_resids"),
  plot_resids("gls_resids")
)
plot_grid(plotlist = plt_list, nrow = 1)
dev.off()
# The GLS model residuals look better than the OLS residuals

# Compare AIC
# You will get warning about the models not having the same number of observations, but the number of observations is the same for all models, so this is not an issue 
all_equal_nobs <- length(unique(c(
  nobs(ols_model), nobs(gls_model)
))) == 1  
stopifnot(all_equal_nobs)

AIC(ols_model, gls_model) # OLS model has lowest AIC, but it has spatial autocorrelation in the residuals

# Compare significant variables (p < 0.05) for each model using map
models <- list(OLS = ols_model, GLS = gls_model)
sig_vars_by_model <- map(models, ~ broom::tidy(.x) %>%
  filter(!term %in% c("lambda", "rho", "(Intercept)"), p.value < 0.05) %>%
  pull(term))
print(sig_vars_by_model)
# $OLS
# [1] "bio1"      "csi_past"  "tmean_dif" "fire"      "Q"        

# $GLS
# [1] "bio1"      "csi_past"  "tmean_dif" "fire"      "Q"       
```

# Fit final model 
```{r}
# Create final formula
f <- formula("Ho ~ glacier + NDVI + gHM + bio1 + csi_past + tmean_dif + fire + Q")

# Fit model
final_model <- gls(f, correlation = corExp(form = ~ x + y, nugget = FALSE), data = mod_sf_unique_scaled)

# Results
mod_result <- broom::tidy(final_model)
mod_result

# Significant variables
mod_result %>% filter(p.value < 0.05) %>% arrange(desc(abs(estimate))) 
sig_vars <- mod_result %>% filter(p.value < 0.05, !term %in% c("(Intercept)", "lambda")) %>% pull(term)
print(sig_vars)

# You can't get a traditional R2 for spatial models, but you can get an alternative R2
cor(predict(final_model), mod_sf_unique$Ho)^2 # 0.46
cor(predict(final_model), mod_sf_unique$Ho) # 0.68
```

# Figures
```{r}
# Define a color scale function using MetBrewer with an argument to extend the scale for better visualization
scale_met_cont <- function(values, aesthetic = c("fill", "colour"), extend_scale = 0.1) {
  aesthetic <- match.arg(aesthetic)
  r_min <- min(values, na.rm = TRUE)
  r_max <- max(values, na.rm = TRUE)
  rng   <- r_max - r_min
  lims  <- c(r_min - rng*extend_scale, r_max + rng*extend_scale)

  pal <- rev(MetBrewer::met.brewer("Hiroshige", type = "continuous"))
  if (aesthetic == "fill") {
    ggplot2::scale_fill_gradientn(colors = pal, limits = lims, oob = scales::squish)
  } else {
    ggplot2::scale_colour_gradientn(colors = pal, limits = lims, oob = scales::squish)
  }
}

# Make residual plots for model
plot_mod <- function(x, y, df, model, 
                     mod_vars = c("NDVI","gHM","bio1","csi_past","glacier","tmean_dif","fire","Q"),
                     color_var = "x") {

  # Calculate residuals for y variable given x variable
  df[[y]] <- residuals(model) + coef(model)[[x]] * as.numeric(scale(df[[x]]))

  # Choose whether to color by x or y variable
  col_var <- if (color_var == "x") x else y

  ggplot(df, aes(x = .data[[x]], y = .data[[y]])) +
    geom_point(aes(col = .data[[col_var]])) +
    geom_smooth(method = "lm", colour = "black") +
    scale_met_cont(df[[col_var]], aesthetic = "colour", extend_scale = 0.1) +
    #ggpubr::stat_cor() +
    labs(x = stringr::str_to_sentence(make_pretty_names(x)), 
      y = stringr::str_to_sentence("Residual heterozygosity")) +
    theme_classic() +
    theme(legend.position = "none",
          axis.title = element_text(size = 14),
          axis.text  = element_text(size = 12))

}

# Load polygons for plotting
ca_proj <- get_ca() %>% sf::st_transform(3310) # California boundaries
range_map <- get_range() # Sceloporus range map

# Loaded wingen map
wingen_r <- 
  terra::rast(here(outpath, "Ho_kriged_masked.tif")) %>% 
  mask(range_map) %>% 
  mask(ca_proj) %>%
  as.data.frame(xy = TRUE, na.rm = TRUE) 
names(wingen_r)[3] <- "layer"  # ensure 'layer' is the raster value column

# Isolate significant vs all variables
sig_vars <- c("csi_past", "tmean_dif", "bio1", "fire", "Q")
mod_vars <- c("NDVI", "gHM", "bio1", "csi_past", "glacier", "tmean_dif", "fire", "Q")

# A: wingen map of heterozygosity 
ho_plot <- ggplot(wingen_r) +
  geom_sf(data = ca_proj, fill = "gray90", colour = NA, size = 2) +
  geom_tile(aes(x = x, y = y, fill = layer), colour = NA) +
  scale_met_cont(wingen_r$layer, aesthetic = "fill", extend_scale = 0.1) +
  labs(fill = "Heterozygosity") +
  theme_void() +
  geom_sf(data = mod_sf_unique, shape = 21, aes(fill = Ho), colour = "#00000087", size = 2) +
  theme(
    plot.margin = margin(1, 1, 1, 1, unit = "mm"),
    legend.position = "inside",              # turn on "inside" mode
    legend.justification = c(1, 1),          # anchor legend's top-right corner
    legend.position.inside = c(0.9, 0.9),   # place near panel's top-right
    legend.key.height = unit(0.5, "cm"),
    legend.key.width  = unit(0.5, "cm"),
    legend.title = element_text(size = 12),
    legend.text  = element_text(size = 10)
  )

# B: partial residuals 
plts_residual_y <- purrr::map(
  sig_vars, 
  # Use unscaled data for plotting
  ~plot_mod(.x, "Ho", mod_sf_unique, model = final_model, color_var = "y")
)

# C: predicted vs observed 
pred_vs_obs <-
  # Use unscaled data for plotting
  ggplot(mod_sf_unique, aes(x = predict(final_model), y = Ho)) +
  geom_point(aes(col = Ho), shape = 1) +
  geom_smooth(method = "lm", colour = "black") +
  scale_met_cont(c(mod_sf_unique$Ho, wingen_r$layer), aesthetic = "colour", extend_scale = 0.1) +
  theme_classic() +
  labs(x = "Predicted heterozygosity", y = "Observed heterozygosity") +
  ggpubr::stat_cor(size = 5) +
  theme(legend.position = "none",
        axis.title = element_text(size = 14),
        axis.text  = element_text(size = 12))

pred_vs_obs_adj <- cowplot::plot_grid(pred_vs_obs, NULL, nrow = 1, rel_widths = c(0.9, 0.1))

plt1 <- cowplot::plot_grid(plotlist = plts_residual_y, ncol = 2)
plt2 <- cowplot::plot_grid(plotlist = list(ho_plot, pred_vs_obs_adj), ncol = 1,
                           labels = c("A", "C"), rel_heights = c(0.6, 0.4), label_size = 18)

png(here(plotpath, "figure_Ho.png"), height = 2.95*4*300*0.9, width = 3.55*4*300*0.9, res = 300)
cowplot::plot_grid(plt2, plt1, ncol = 2,labels = c("", "B"),  rel_widths = c(0.4, 0.6), align = "h", label_size = 18)
dev.off()
```