
```{r}
library("here")
library("tidyverse")
library("ggpubr")
library("terra")
library("sf")
library("gridExtra")
library("rpaleoclim")
library("spdep")
library("spatialreg")
library("MuMIn")
library("cowplot")
library("performance")
library("gstat")
library("sp")
library("furrr")
library("MetBrewer")

source(here("general_functions.R"))
source(here("analysis", "genetic_diversity", "functions_genetic_diversity.R"))
source(here("analysis", "genetic_diversity", "functions_paleovars.R"))
ca <- get_ca()
outpath <- here("analysis", "genetic_diversity", "outputs")
plotpath <- here("analysis", "genetic_diversity", "plots")

coords <- get_coords(sf = TRUE)
dem <- get_dem(r= TRUE)
```

```{r}
# Get genome-wide heterozygosity
het <- get_het("58-Sceloporus_chr.het") 

# Make sure you have the expected number of samples
stopifnot(nrow(het) == 259)

# Count number of sites used to calculate heterozygosity
mean(het$`N(NM)`, na.rm = TRUE) # 59955192

# Get adaptive heterozygosity
genes <- 
  format_het(here("analysis", "gea", "outputs", "genes.het"), callable_sites = 1) %>%
  mutate(Ho_gea = 1 - `O(HOM)`/`N(NM)`) %>%
  rename(SampleID = IID) %>%
  dplyr::select(SampleID, Ho_gea) 
```

# Hypothesis 1: Genetic diversity explained by paleoclimate stability

```{r}
# Get current climate data (NOTE: not using paleoclim data because they use V1.2 instead of V2.1)
bioclim <- rast(here("data", "env", "california_chelsa_bioclim_1981-2010_V.2.1.tif"))
cur <- bioclim[["CHELSA_bio1_1981-2010_V.2.1"]]
names(cur) <- "cur"

# Get LGM climate data (NOTE: not using paleoclim data because they use V1.2 instead of V2.1)
lgm <- rast(here("data", "env", "california_chelsa_pmip_ccsm4.tif"))
names(lgm) <- "lgm"

# Get paleoclimate data and convert from C/10 to C
lig <- get_paleovars("lig", cache = TRUE)$lig$bio_1/10 
paleovars <- get_paleovars(c("lh", "mh", "eh", "yds", "ba", "hs1", "lig", "mis19"))
paleobio1 <- map(paleovars, ~.x$bio_1) %>% rast()/10

# Calculate climate change over different time periods
cur_lgm <- cur - lgm
names(cur_lgm) <- "cur_lgm"
# Note: LIG is lower resolution than current and LGM data, so we need to resample it
lgm_lig <- lgm - resample(lig, lgm)
names(lgm_lig) <- "lgm_lig"

# Calculate climate stability
chelsabio1 <- list(cur, lgm)
# Resample the higher resolution data (CHELSA) to match the lower resolution (rpaleoclim)
# This is the most conservative approach to avoid over-interpreting fine-scale variation 
csi_stack <- 
  c(as.list(chelsabio1), as.list(paleobio1)) %>% 
  map(~resample(.x, paleobio1[[1]]), method = "bilinear") %>% 
  rast()

# Calculate the standard deviation over the stack of rasters
# Don't remove NAs, use imputation later (NA values are some points along the coast)
# If you remove NAs you will get some outliers because some time periods have data and others do not
csi_past <- app(csi_stack, sd, na.rm = FALSE)

# Scale CSI raster from 0 to 1 and subtract from 1 to get higher values = more stable
csi_past[] <- 1 - scales::rescale(csi_past[], to = c(0, 1), na.rm = TRUE)

# Make a data frame with the climate change values
paleoclim_mod_df <- 
  data.frame(
    SampleID = coords$SampleID,
    cur_lgm = terra::extract(cur_lgm, coords, ID = FALSE)[,1],
    lgm_lig = terra::extract(lgm_lig, coords, ID = FALSE)[,1],
    csi_past = terra::extract(csi_past, coords, ID = FALSE)[,1]
  )

# Check NA values
print(colSums(is.na(paleoclim_mod_df)))

# Some coastal points have NA values, so use a bilinear approach to get take the average of the four nearest raster cells
paleoclim_mod_df$csi_past <- bilinear_impute(paleoclim_mod_df$csi_past, csi_past, coords)
paleoclim_mod_df$cur_lgm <- bilinear_impute(paleoclim_mod_df$cur_lgm, cur_lgm, coords)
paleoclim_mod_df$lgm_lig <- bilinear_impute(paleoclim_mod_df$lgm_lig, lgm_lig, coords)

# Double check NA values
print(colSums(is.na(paleoclim_mod_df)))

# All paleovariables are highly correlated
cor(paleoclim_mod_df[, - 1], use = "pairwise.complete.obs")
```
```{r}
pdf(here(plotpath, "paleoclim_test.pdf"), width = 10, height = 10)
test_df <- left_join(paleoclim_mod_df, het) %>% left_join(coords) %>% st_as_sf()

ggplot(test_df) +
  geom_sf(data = ca) +
  geom_sf(aes(col = csi_past, geometry = geometry)) +
  theme_void() +
  scale_color_viridis_c(option = "turbo", end = 0.1, begin = 0.9, direction = -1) 

ggplot(test_df) +
  geom_point(aes(x = csi_past, y = Ho), inherit.aes = FALSE) +
  geom_smooth(aes(x = csi_past, y = Ho), method = "lm", col = "green") +
  stat_smooth(aes(x = csi_past, y = Ho), method = "gam", formula = y ~ s(x, k = 3), col = "blue") +
  stat_smooth(aes(x = csi_past, y = Ho), method = "lm", formula = y ~ poly(x, 2), col = "red", linetype = "dashed") +
  theme_classic()
dev.off()
```

# Hypothesis 2: Genetic diversity explained by admixture
```{r}
q <- read_csv(here("analysis", "admixture", "outputs", "Q8.csv"))

q_mod_df <- 
  q %>%
  dplyr::select(SampleID, Q) %>%
  # Do 1- to make the values more intuitive
  mutate(Q = 1 - Q)
```

# Hypothesis 3: Genetic diversity explained by current environment 
```{r}
bioclim <- rast(here("data", "env", "california_chelsa_bioclim_1981-2010_V.2.1.tif"))
bio1 <- bioclim[["CHELSA_bio1_1981-2010_V.2.1"]]
ndvi <- rast(here("data", "env", "california_ndvi_mean_2000_2020.tif"))
ghm <- rast(here("data", "env", "california_global_human_modification_2016.tif"))

env_mod_df <- data.frame(
  SampleID = coords$SampleID,
  NDVI = terra::extract(ndvi, coords, ID = FALSE)[, "NDVI"],
  gHM = terra::extract(ghm, coords, ID = FALSE)[,"gHM"],
  bio1 = terra::extract(bio1, coords, ID = FALSE)[,"CHELSA_bio1_1981-2010_V.2.1"]
)

# Check NA values
print(colSums(is.na(env_mod_df)))

# Some coastal points have NA values for gHM and bio1, so use a bilinear approach to get take the average of the four nearest raster cells
env_mod_df$gHM <- bilinear_impute(env_mod_df$gHM, ghm, coords)
env_mod_df$bio1 <- bilinear_impute(env_mod_df$bio1, bio1, coords)

# Double check NA values
print(colSums(is.na(env_mod_df)))
```

# Hypothesis 4: Genetic diversity explained by glaciation
```{r}
lgm_glaciers <- st_read(here("data", "env", "Glaciotect_NA_lgm", "shp", "LGM_maxextent.shp"))
# Note: have to first intersect and then transform otherwise you get a node error
lgm_glaciers <- st_intersection(lgm_glaciers, st_transform(ca, crs = st_crs(lgm_glaciers)))
lgm_glaciers <- st_transform(lgm_glaciers, crs = st_crs(ca))
# Add a new column to coords indicating whether each point is inside the lgm shapefile
coords <- get_coords(sf = TRUE) %>% st_transform(crs = st_crs(lgm_glaciers))
glacier_mod_df <- 
  coords %>%
  mutate(glacier = as.numeric((lengths(st_within(coords, lgm_glaciers)) > 0))) %>%
  st_drop_geometry()
```

# Hypothesis 5: Genetic diversity explained by recent climate change
PRISM mean temperature data from historical data: https://prism.oregonstate.edu/historical/

```{r}
coords <- get_coords(sf = TRUE)
# Get prism data
get_prism <- function(year, output = "df") {
  # Create PRISM path
  prism_path <- here("data", "env", "prism", "PRISM_tmean_stable_4kmM3", year)

  # List annual file
  file <- list.files(prism_path, pattern = paste0(year, "_*.bil.bil$"), full.names = TRUE)

  # Load data
  raster <- terra::rast(file)

  if (output == "raster"){
    return(raster)
  } else {
    vals <- terra::extract(raster, st_transform(coords, st_crs(raster)), ID = FALSE)
    names(vals) <- "tmean"
    df <- data.frame(year = year, tmean = vals, SampleID = coords$SampleID)
    return(df)
  }
}

# Loop over the years of interest
historical_years <- 1895:1905
current_years <- 2010:2020
## Get dataframes
tmean_his <- map(historical_years, ~get_prism(year = .x), .progress =TRUE) %>% bind_rows()  
tmean_his <- tmean_his %>% group_by(SampleID) %>% summarise(tmean_his = mean(tmean, na.rm = TRUE))
tmean_cur <- map(current_years, ~get_prism(year = .x), .progress =TRUE) %>% bind_rows()  
tmean_cur <- tmean_cur %>% group_by(SampleID) %>% summarise(tmean_cur = mean(tmean, na.rm = TRUE))
# Get rasters
tmean_his_r <- map(historical_years, ~get_prism(year = .x, output = "raster"), .progress =TRUE) %>% rast() %>% mean()
tmean_cur_r <- map(current_years, ~get_prism(year = .x, output = "raster"), .progress =TRUE) %>% rast() %>% mean()
tmean_dif_r <- tmean_cur_r - tmean_his_r
tmean_dif_r <- crop(tmean_dif_r, ca)
tmean_dif_r <- mask(tmean_dif_r, st_transform(ca, st_crs(tmean_dif_r)))

# Calculate change
tmean_mod_df <- 
  left_join(tmean_his, tmean_cur) %>% 
  mutate(tmean_dif = tmean_cur - tmean_his) %>%
  right_join(coords) %>% 
  dplyr::select(SampleID, tmean_dif, tmean_cur)

# Create and save raster
tmean_dif <- tmean_dif_r %>% project("epsg:3310")
names(tmean_dif) <- "tmean_dif"
writeRaster(tmean_dif, here("data", "env", "tmean_dif.tif"), overwrite = TRUE)
```

# Combe rasters
Combine into a raster stack:
```{r}
env_list <- 
  list(
    NDVI = ndvi,
    bio1 = bio1,
    gHM = ghm,
    tmean_dif = tmean_dif,
    cur_lgm = cur_lgm,
    lgm_lig = lgm_lig,
    csi_past = csi_past
  )

env_list <- imap(env_list, ~{
  .x <- terra::project(.x, "epsg:3310")
  names(.x) <- .y
  return(.x)
}, .progress = TRUE)

path <- here("data", "env", "model_rasters")
# Make model rasters dir
dir.create(path, showWarnings = FALSE)
iwalk(env_list, ~writeRaster(.x, here(path, paste0(.y, ".tif")), overwrite = TRUE))
```

# Create model data
```{r}
mod_vars <- c("NDVI", "gHM", "bio1", "Q", "csi_past", "glacier", "tmean_dif", "cur_lgm", "lgm_lig", "x", "y")

# Combine dataframes
mod_df <- 
  get_coords() %>% 
  left_join(het) %>%
  #left_join(genes) %>%
  left_join(env_mod_df) %>%
  left_join(paleoclim_mod_df) %>%
  left_join(q_mod_df) %>%
  left_join(glacier_mod_df) %>%
  left_join(tmean_mod_df) %>%
  dplyr::select(Ho, SampleID, all_of(mod_vars)) %>%
  drop_na(Ho)

# Check sample numebr
stopifnot(nrow(mod_df) == 259)

# Confirm no NAs:
stopifnot(all(colSums(is.na(mod_df)) == 0))

# Scale the variables (except glacier)
mod_df_scaled <- 
  mod_df %>% 
  mutate(across(c(all_of(mod_vars), -glacier), scale))

# Convert to sf
mod_sf <- get_coords(sf = TRUE) %>% right_join(mod_df_scaled) %>% st_transform(3310)
```

```{r}
export_df <- 
  st_drop_geometry(mod_df) %>% 
  dplyr::select(SampleID, everything())
write_csv(export_df, here(outpath, "model_df.csv"))
```

## Check for correlation between variables
```{r}
# Check correlations
# Compute the correlation matrix
r <- cor(mod_df_scaled[, mod_vars], use = "pairwise.complete.obs")
# Set the diagonal and lower triangle to NA
r[lower.tri(r, diag = TRUE)] <- NA
# Find the pairs with correlations greater than 0.6
high_corr <- which(abs(r) > 0.6, arr.ind = TRUE)
# Print the pairs
high_corr_pairs <- data.frame(
  Var1 = rownames(r)[high_corr[, 1]],
  Var2 = rownames(r)[high_corr[, 2]],
  Correlation = r[high_corr]
)
high_corr_pairs

# cur_lgm, csi_past, and lgm_lig are all correlated

```

# Run Ho models

## Check for spatial autocorrelation
```{r}
# Uncorrelated set of model variables
mod_vars <- c("NDVI", "gHM", "bio1", "csi_past", "glacier_factor", "tmean_dif", "Q")

# Make glacier into a factor
mod_df_scaled$glacier_factor <- as.factor(mod_df_scaled$glacier)
mod_sf$glacier_factor <- as.factor(mod_sf$glacier)

# Plot relationships between each predictor and Ho
pdf(here(plotpath, "predictor_vs_Ho.pdf"), width = 3, height = 3)
walk(mod_vars, ~print(gglm(.x, "Ho", mod_df_scaled)))
dev.off()

# Past climate stability appears to have a nonlinear relationship with Ho, so we will use a polynomial term for it
# Compare AIC for model with and without polynomial term
f <- formula("Ho ~ NDVI + gHM + bio1 + csi_past + glacier_factor + tmean_dif + Q")
f_poly <- formula("Ho ~ NDVI + gHM + bio1 + poly(csi_past, 2) + glacier_factor + tmean_dif + Q")
mod <- lm(f, data = mod_df_scaled)
mod_poly <- lm(f_poly, data = mod_df_scaled)
AIC(mod, mod_poly) # The polynomial model is better

# Run OLS and get residuals
f <- formula("Ho ~ NDVI + gHM + bio1 + poly(csi_past, 2) + glacier_factor + tmean_dif + Q")
ols_model <- lm(f, data = mod_df_scaled)
AIC(ols_model)
mod_sf$ols_resids <- residuals(ols_model)
broom::tidy(ols_model) %>% filter(p.value < 0.05) %>% arrange(desc(abs(estimate)))

# Fit variograms to the residuals
ols_variogram <- variogram(ols_resids ~ 1, data = mod_sf)
ols_vgm <- map(c("Exp", "Sph", "Gau"), ~fit.variogram(ols_variogram, model = vgm(.x)))
names(ols_vgm) <- c("Exp", "Sph", "Gau")

pdf(here(plotpath, "ols_variogram.pdf"))
plot(ols_variogram, ols_vgm[["Exp"]], main = "OLS Residuals Variogram (Exponential)")
plot(ols_variogram, ols_vgm[["Sph"]], main = "OLS Residuals Variogram (Spherical)")
plot(ols_variogram, ols_vgm[["Gau"]], main = "OLS Residuals Variogram (Gaussian)")
dev.off()

# Based on visual inspection, the exponential model seems to fit best
# Get the range of the variogram 
vgm_range <- ols_vgm[["Exp"]] %>% filter(model == "Exp") %>% pull(range)
print(vgm_range)

# Create a spatial weights matrix using the range from the variogram
nb <- dnearneigh(mod_sf, d1 = 0, d2 = vgm_range)
listw <- nb2listw(nb, style = "W", zero.policy = TRUE)
pdf(here(plotpath, "spatial_weights.pdf"))
plot(nb, st_coordinates(mod_sf), main = paste0("Neighborhoods (Range = ", round(vgm_range, 2), ")"))
dev.off()

# Fit an SEM model
sem_model <- errorsarlm(f, data = mod_sf, listw = listw, zero.policy = TRUE)
summary(sem_model)
mod_sf$sem_resids <- residuals(sem_model)
mod_sf$sem_fit <- fitted(sem_model)

# Fit an SLM model
slm_model <- lagsarlm(f, data = mod_sf, listw = listw, zero.policy = TRUE)
summary(slm_model)
mod_sf$slm_resids <- residuals(slm_model)
mod_sf$slm_fit <- fitted(slm_model)
# https://crd230.github.io/lab8.html

# Calculate Moran's I before and after accounting for spatial autocorrelation
moran.test(mod_sf$ols_resids, listw, zero.policy = TRUE)$estimate # 0.38
moran.test(mod_sf$sem_resids, listw, zero.policy = TRUE)$estimate # -0.04
moran.test(mod_sf$slm_resids, listw, zero.policy = TRUE)$estimate # 0.30

# Compare AIC
AIC(ols_model, sem_model, slm_model) # SEM model is the best choice
```

```{r}
# Final model variables
final_vars <- c("NDVI", "gHM", "bio1", "csi_past", "glacier_factor", "tmean_dif", "Q")

# Create formula
f <- formula("Ho ~ NDVI + gHM + bio1 + poly(csi_past, 2) + glacier_factor + tmean_dif + Q")

# Fit model
best_model <- errorsarlm(f, data = mod_sf, listw = listw, zero.policy = TRUE)

# Results
mod_result <- broom::tidy(best_model)
mod_result

# Significant variables
mod_result %>% filter(p.value < 0.05) %>% arrange(desc(abs(estimate))) 

# You can't get a traditional R2 for spatial models, but you can get an alternative R2
r2_alternative <- cor(best_model$fitted.values, best_model$y)^2
print(r2_alternative) # 61%
```

## Variance partioning analysis
```{r}
all_vars <- c("NDVI", "gHM", "bio1", "cur_lgm", "lgm_lig", "csi_past", "glacier", "tmean_dif", "Q")
predictor_sets <- mod_df_scaled %>% select(all_of(all_vars)) %>% as.list()
response <- mod_df_scaled$Ho
varpart_result <- lm_variance_partition(predictor_sets, response, scale_vars =TRUE)
```

## Plots 
```{r}
# Model variables for plotting (glacier is not treated as a factor for plotting)
mod_vars <- c("NDVI", "gHM", "bio1", "csi_past", "glacier", "tmean_dif", "Q")
mod <- best_model

# Confirm treating glacier as a factor vs numeric makes no difference because it is binary
lm(Ho ~ glacier, data = mod_df_scaled) %>% summary()
lm(Ho ~ glacier_factor, data = mod_df_scaled) %>% summary()


# Plot predicted vs observed
plt1 <- 
  map(mod_vars, ~{
    ggplot(mod_df) +
      geom_point(aes(x = predict(mod), y = Ho, col = .data[[.x]])) +
      #geom_abline(intercept = 0, slope = 1) +
      geom_smooth(aes(x = predict(mod), y = Ho), method = "lm") +
      theme_classic() +
      labs(x = "Predicted", y = "Observed") +
      scale_color_viridis_c(option = "mako", end = 0.1, begin = 0.9)
  })

# Plot values across california
plt2 <-
  map(mod_vars, ~{
    ggplot(mod_df) +
      geom_sf(data = ca) +
      geom_point(aes(x = x, y = y, col = .data[[.x]])) +
      theme_void() +
      scale_color_viridis_c(option = "mako", end = 0.1, begin = 0.9)
  })

# Plot Ho vs variable
plt3 <- map(mod_vars, ~{
  ggplot(mod_df, aes(x = .data[[.x]], y = Ho)) +
    geom_point(aes(col = .data[[.x]])) +
    geom_smooth(method = "lm") +
    theme_classic() +
    labs(x = gsub("\n", " ", .x), y = "Ho") +
    scale_color_viridis_c(option = "mako", end = 0.1, begin = 0.9) +
    theme(legend.position = "none") +
    stat_cor(label.y = min(gg_df[["Ho"]], na.rm = TRUE))
})

# Plot Ho
plt4 <-
  map(1:length(mod_vars), ~{
    ggplot(mod_df) +
      geom_sf(data = ca) +
      geom_point(aes(x = x, y = y, col = Ho)) +
      theme_void() +
      scale_color_viridis_c(option ="mako", end = 0.1, begin = 0.9, direction = -1)
})

pdf(here(plotpath, "ho.pdf"), width = 4.5, height = 4.5)
ggplot(mod_df) +
  geom_sf(data = ca) +
  geom_point(aes(x = x, y = y, col = Ho)) +
  theme_void() +
  scale_color_viridis_c(option ="mako") +
  theme(
    plot.margin = margin(t = 1, r = 1, b = 1, l = 1, unit = "mm"),
    legend.position = c(0.85, 0.6),  # (x, y) in npc coords (0-1)
    legend.justification = c(1, 0)
  )
dev.off()

# use cowplot to combine plots
pdf(here(plotpath, "linear_model_plot_Ho.pdf"), width = 10, height = 25)
ordered_plts <- pmap(list(plt3, plt1, plt2), ggpubr::ggarrange, common.legend = TRUE, legend = "right", nrow = 1)
do.call(cowplot::plot_grid, c(ordered_plts, list(nrow = length(plt1))))
ordered_plts <- pmap(list(plt3, plt1, plt4), ggpubr::ggarrange, nrow = 1)
do.call(cowplot::plot_grid, c(ordered_plts, list(nrow = length(plt1))))
dev.off()
```

```{r}

col_function <- function() scale_color_gradientn(colors = rev(MetBrewer::met.brewer("Hiroshige", type = "continuous")))
fill_function <- function() scale_fill_gradientn(colors = rev(MetBrewer::met.brewer("Hiroshige", type = "continuous")))
# col_function <- function() scale_color_viridis_c(option = "mako")
# fill_function <- function() scale_fill_viridis_c(option = "mako")
plot_poly <- function(x, y, df, partial = FALSE, mod_vars = c("NDVI", "gHM", "bio1", "csi_past", "glacier", "tmean_dif", "Q"), col = "x") {
  if (partial) {
    f_partial_y <- paste(y, "~", paste(setdiff(mod_vars, x), collapse = " + "))
    if (x != "csi_past") f_partial_y <- gsub("csi_past", "poly(csi_past, 2)", f_partial_y)
    f_partial_x <- paste(x, "~", paste(setdiff(mod_vars, x), collapse = " + "))
    if (x != "csi_past") f_partial_x <- gsub("csi_past", "poly(csi_past, 2)", f_partial_x)
    df[[x]] <- residuals(lm(f_partial_x, data = df))
    df[[y]] <- residuals(lm(f_partial_y, data = df))
  }

  col_var <- ifelse(col == "x", x, y)
  plt <-
    ggplot(df, aes(x = .data[[x]], y = .data[[y]], col = .data[[col_var]])) +
    geom_point() +
    theme_classic() +
    col_function() +
    labs(x = make_pretty_names(x), y = make_pretty_names(y)) +
    theme(legend.position = "none") 

  if (x == "csi_past") {
    plt <- plt + geom_smooth(method = "lm", formula = y ~ poly(x, 2), col = "black")
  } else {
    plt <- plt + geom_smooth(method = "lm", col = "black")
  }
  
}

sig_vars <- c("csi_past", "tmean_dif", "bio1", "Q")
plts_partial <- map(sig_vars, ~plot_poly(.x, "Ho", mod_df, partial = TRUE))
plts_raw <- map(sig_vars, ~plot_poly(.x, "Ho", mod_df, partial = FALSE))

pred_vs_obs <- 
  ggplot(mod_df, aes(x = predict(best_model), y = Ho)) +
  geom_point(aes(col = Ho)) +
  geom_smooth(method = "lm", col = "black") +
  col_function() +
  theme_classic() +
  labs(x = "Predicted heterozygosity", y = "Observed heterozygosity") +
  stat_cor(label.y = min(mod_df[["Ho"]], na.rm = TRUE)) +
  theme(legend.position = "none")

ho_plot <- 
  ggplot(mod_df) +
  geom_sf(data = ca) +
  geom_point(aes(x = x, y = y, fill = Ho), col = "black", cex = 2, pch = 21) +
  theme_void() +
  fill_function() +
  labs(fill = "Heterozygosity") +
  theme(
    plot.margin = margin(t = 1, r = 1, b = 1, l = 1, unit = "mm"),
    legend.position = c(0.9, 0.55),  # (x, y) in npc coords (0-1)
    legend.justification = c(1, 0), 
    legend.key.height = unit(0.5, "cm"),
    legend.key.width = unit(0.5, "cm"),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8)
  )

env_plot <-
  map(sig_vars, ~{
    ggplot(mod_df) +
      geom_sf(data = ca) +
      geom_point(aes(x = x, y = y, fill = .data[[.x]]), cex = 2, col = "black", pch = 21) +
      theme_void() +
      labs(fill = gsub(" ", "\n", make_pretty_names(.x))) +
      fill_function() +
      theme(
        plot.margin = margin(t = 1, r = 1, b = 1, l = 1, unit = "mm"),
        legend.position = c(0.9, 0.55),  # (x, y) in npc coords (0-1)
        legend.justification = c(1, 0), 
        legend.key.height = unit(0.5, "cm"),
        legend.key.width = unit(0.5, "cm"),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8)
      )
  })
  
pltlist1b <- c(list(pred_vs_obs), plts_partial)
pltlist1a <- c(list(pred_vs_obs), plts_raw)
pltlist2 <- c(list(ho_plot), env_plot)
pdf(here(plotpath, "partial_plots_Ho.pdf"), width = 16, height = 9)
a <- plot_grid(plotlist = pltlist1a, nrow = 1)
b <- plot_grid(plotlist = pltlist1b, nrow = 1)
d <- plot_grid(plotlist = pltlist2, nrow = 1, align = "hv")
plot_grid(a, b, d, ncol = 1)
dev.off()

pdf(here(plotpath, "figure_Ho.pdf"), width = 8, height = 20)
a <- plot_grid(plotlist = pltlist1a, ncol = 1)
b <- plot_grid(plotlist = pltlist2, ncol = 1)
plot_grid(a, b, nrow = 1)
dev.off()

# Repeated but params adjusted slightly
ho_plot <- 
  ggplot(mod_df) +
  geom_sf(data = ca) +
  geom_point(aes(x = x, y = y, fill = Ho), col = "black", cex = 1.5, pch = 21) +
  theme_void() +
  fill_function() +
  labs(fill = "Heterozygosity") +
  theme(
    plot.margin = margin(t = 1, r = 1, b = 1, l = 1, unit = "mm"),
    legend.position = c(0.9, 0.6),  # (x, y) in npc coords (0-1)
    legend.justification = c(1, 0), 
    legend.key.height = unit(0.5, "cm"),
    legend.key.width = unit(0.5, "cm"),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8)
  )


pltlist1 <- c(list(custom_plot(weighted_krig)), list(pred_vs_obs))
pdf(here(plotpath, "figure_Ho2.pdf"), height = 7.5, width = 12)
plt1 <- plot_grid(plotlist = plts_raw, nrow = 2)
plt2 <- plot_grid(plotlist = pltlist1, ncol = 1, labels = c("A", "C"))
plot_grid(plt2, plt1, ncol = 2, rel_widths = c(0.33, 0.66), labels = c("", "B"))
dev.off()
```


```{r}
marginal_effect_plot <- function(var, model, data, listw, response = "Ho") {
  # Sequence for var: from min to max observed
  x_seq <- seq(min(data[[var]]), max(data[[var]]), length.out = 100)
  
  # Build newdata: hold all other predictors constant
  newdata <- data[rep(1, 100), ]
  newdata[[var]] <- x_seq
  
  # Set all other numeric predictors to their mean
  numeric_vars <- sapply(data, is.numeric)
  numeric_vars <- setdiff(names(data)[numeric_vars], c(var, response, "x", "y"))
  for (v in numeric_vars) {
    newdata[[v]] <- mean(data[[v]])
  }
  
  # Set factors to their reference level (first level)
  factor_vars <- sapply(data, is.factor)
  for (v in names(data)[factor_vars]) {
    newdata[[v]] <- factor(levels(data[[v]])[1], levels = levels(data[[v]]))
  }
  
  # Drop row names to avoid region.id conflict
  rownames(newdata) <- NULL
  
  # Predict using spatial error model
  newdata$fit <- predict(model, newdata = newdata, listw = listw)
  
  # Plot predicted marginal effect with raw data
  ggplot() +
    geom_point(data = data, aes_string(x = var, y = response), alpha = 0.5) +
    geom_line(data = newdata, aes_string(x = var, y = "fit"), color = "blue", size = 1.2) +
    labs(
      x = make_pretty_names(var),
      y = "Predicted Heterozygosity"
    ) +
    theme_classic()
}

# Apply to all significant  variables
marginal_plots <- map(sig_vars, ~marginal_effect_plot(.x, best_model, mod_df_scaled, listw))

# Combine into a panel
library(cowplot)
pdf(here(plotpath, "partial_effects_Ho.pdf"), width = 12, height = 3)
plot_grid(plotlist = marginal_plots, nrow = 1, align = "hv")
dev.off()

pltlist1b <- c(list(pred_vs_obs), partial_plots)
pltlist1a <- c(list(pred_vs_obs), plts_raw)
pltlist2 <- c(list(ho_plot), env_plot)
pdf(here(plotpath, "partial_plots_Ho.pdf"), width = 16, height = 9)
a <- plot_grid(plotlist = pltlist1a, nrow = 1)
b <- plot_grid(plotlist = pltlist1b, nrow = 1)
d <- plot_grid(plotlist = pltlist2, nrow = 1, align = "hv")
plot_grid(a, b, d, ncol = 1)
dev.off()

```
```{r}
model_Ho <- best_model
```


# Run Adaptive Ho models

## Check for spatial autocorrelation
```{r}
# Run OLS and get residuals
f <- as.formula(paste("Ho_gea ~", paste(mod_vars, collapse = " + ")))
ols_model <- lm(f, data = mod_df_scaled)
mod_sf$ols_resids <- residuals(ols_model)

# Fit variograms to the residuals
ols_variogram <- variogram(ols_resids ~ 1, data = mod_sf)
ols_vgm <- map(c("Exp", "Sph", "Gau"), ~fit.variogram(ols_variogram, model = vgm(.x)))
names(ols_vgm) <- c("Exp", "Sph", "Gau")

pdf(here(plotpath, "ols_variogram_gea.pdf"))
plot(ols_variogram, ols_vgm[["Exp"]], main = "OLS Residuals Variogram (Exponential)")
plot(ols_variogram, ols_vgm[["Sph"]], main = "OLS Residuals Variogram (Spherical)")
plot(ols_variogram, ols_vgm[["Gau"]], main = "OLS Residuals Variogram (Gaussian)")
dev.off()

# Based on visual inspection, the exponential model seems to fit best
# Get the range of the variogram (3 times the range parameter)
vgm_range <- ols_vgm[["Exp"]] %>% filter(model == "Exp") %>% pull(range)
print(vgm_range)

# Create a spatial weights matrix using the range from the variogram
nb <- dnearneigh(mod_sf, d1 = 0, d2 = vgm_range)
listw <- nb2listw(nb, style = "W", zero.policy = TRUE)

# Fit an spatial model
sp_model <- lagsarlm(f, data = mod_sf, listw = listw, zero.policy = TRUE)
summary(sp_model)
mod_sf$sp_resids <- residuals(sp_model)

# Calculate Moran's I before and after accounting for spatial autocorrelation
moran.test(mod_sf$ols_resids, listw)$estimate
moran.test(mod_sf$sp_resids, listw)$estimate
```

```{r}
full_model <- lagsarlm(f, data = mod_sf, listw = listw, zero.policy = TRUE)

# Variables are NDVI + glacier + Q
broom::tidy(full_model) %>% filter(p.value < 0.05)

best_model <- full_model
summary(best_model)

# You can't get a traditional R2 for spatial models, but you can get an alternative R2
r2_alternative <- cor(sem_model$fitted.values, sem_model$y)^2
r2_alternative # 43%
```

```{r}
# DELETE Later
pdf(here(plotpath, "partial_plots.pdf"), width = 4, height = 4)
ggpartial("csi_past", "Ho", mod_vars, mod_df, col = "csi_past")
ggpartial("tmean_dif", "Ho", mod_vars, mod_df, col = "tmean_dif")
ggpartial("bio1", "Ho", mod_vars, mod_df, col = "bio1")
dev.off()
```

## Variance partioning analysis
```{r}
all_vars <- c("NDVI", "gHM", "bio1", "cur_lgm", "lgm_lig", "csi_past", "glacier", "tmean_dif", "Q")
predictor_sets <- mod_df_scaled %>% select(all_of(all_vars)) %>% as.list()
response <- mod_df_scaled$Ho_gea
varpart_result_gea <- lm_variance_partition(predictor_sets, response, scale_vars =TRUE)
```

## Plots and tables
```{r}
# Turn into a nice gt table
library(gt)

best_model_table <- 
  broom::tidy(best_model) %>%
  data.frame() %>%
  mutate(term = map_chr(term, ~make_pretty_names(.x))) %>%
  filter(term != "(Intercept)") %>%
  gt() %>%
  tab_header(title = "Full model") %>%
  gtExtras::gt_hulk_col_numeric(estimate, trim = TRUE) %>%
  fmt_scientific(
    columns = c("estimate", "std.error", "statistic", "p.value"),
    decimals = 2
  ) %>%
  # bold significant rows
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = c("p.value"),
      rows = p.value < 0.05
    )
  )

# export to html
gt::gtsave(best_model_table, filename = here(plotpath, "best_model_table_Ho_gea.html"))
```

```{r}
# Make pretty plots 
mod <- best_model

# Plot predicted vs observed
plt1 <- 
  map(mod_vars, ~{
    ggplot(mod_df) +
      geom_point(aes(x = predict(mod), y = Ho, col = .data[[.x]])) +
      #geom_abline(intercept = 0, slope = 1) +
      geom_smooth(aes(x = predict(mod), y = Ho), method = "lm") +
      theme_classic() +
      labs(x = "Predicted", y = "Observed") +
      scale_color_viridis_c(option = "mako", end = 0.1, begin = 0.9)
  })

# Plot values across california
plt2 <-
  map(mod_vars, ~{
    ggplot(mod_df) +
      geom_sf(data = ca) +
      geom_point(aes(x = x, y = y, col = .data[[.x]])) +
      theme_void() +
      scale_color_viridis_c(option = "mako", end = 0.1, begin = 0.9)
  })

# Plot Ho vs variable
plt3 <- map(mod_vars, ~{
  ggplot(mod_df, aes(x = .data[[.x]], y = Ho)) +
    geom_point(aes(col = .data[[.x]])) +
    geom_smooth(method = "lm") +
    theme_classic() +
    labs(x = gsub("\n", " ", .x), y = "Ho") +
    scale_color_viridis_c(option = "mako", end = 0.1, begin = 0.9) +
    theme(legend.position = "none") +
    stat_cor(label.y = min(gg_df[["Ho"]], na.rm = TRUE))
})

# Plot Ho
plt4 <-
  map(1:length(mod_vars), ~{
    ggplot(mod_df) +
      geom_sf(data = ca) +
      geom_point(aes(x = x, y = y, col = Ho)) +
      theme_void() +
      scale_color_viridis_c(option ="mako", end = 0.1, begin = 0.9, direction = -1)
})

pdf(here(plotpath, "ho.pdf"), width = 4.5, height = 4.5)
ggplot(mod_df) +
  geom_sf(data = ca) +
  geom_point(aes(x = x, y = y, col = Ho)) +
  theme_void() +
  scale_color_viridis_c(option ="mako") +
  theme(plot.margin = margin(t = 1, r = 1, b = 1, l = 1, unit = "mm"))
dev.off()

# use cowplot to combine plots
pdf(here(plotpath, "linear_model_plot_Ho_gea.pdf"), width = 10, height = 25)
ordered_plts <- pmap(list(plt3, plt1, plt2), ggpubr::ggarrange, common.legend = TRUE, legend = "right", nrow = 1)
do.call(cowplot::plot_grid, c(ordered_plts, list(nrow = length(plt1))))
ordered_plts <- pmap(list(plt3, plt1, plt4), ggpubr::ggarrange, nrow = 1)
do.call(cowplot::plot_grid, c(ordered_plts, list(nrow = length(plt1))))
dev.off()
```

```{r}
model_Ho_gea <- best_model
```

# Future climate change
```{r}
future_ho <- 
  # NOTE: CSI was inverted (1 - CSI) before
  df_future %>%
  left_join(mod_df) %>%
  # REMOVE CHANNEL ISLAND
  filter(SampleID != "Scelocci_CCGPMC_MW01-3-14") %>%
  pivot_longer(c("Ho", "Ho_gea"), names_to = "metric", values_to = "Ho") %>%
  mutate(metric = case_when(
    metric == "Ho" ~ "Genome-wide diversity",
    metric == "Ho_gea" ~ "Adaptive diversity"
  )) %>%
  mutate(metric = factor(metric, levels = c("Genome-wide diversity", "Adaptive diversity"))) 

plt <- 
  ggplot(future_ho, aes(x = csi, y = Ho, col = csi)) +
  geom_point() +
  ggh4x::facet_grid2(metric ~ model_name, scales = "free", switch = "y", axes = "all") +
  theme_classic() +
  labs(x = "Climate stability", y = "") +
  geom_smooth(method = "lm", se = FALSE, col = "black") +
  stat_cor(label.y = min(future_ho$Ho, na.rm = TRUE)) +
  scale_color_viridis_c(option = "mako") +
  theme(
    strip.background = element_blank(),
    strip.text.x = element_text(hjust = 0, size = 14),
    strip.placement = "outside",
    strip.text.y = element_text(size = 14),
    axis.title = element_text(size = 14)
  ) 

pdf(here(plotpath, "future_plot.pdf"), width = 13, height = 6)
plt
dev.off()
```

# Figures
```{r}
# Variance partitioning plot
gg_df <- 
  varpart_result$summary_table %>%
  select(-r2) %>%
  pivot_longer(cols = c("adj_r2", "unique_contribution"), names_to = "metric", values_to = "value") %>%
  mutate(metric = recode(metric, adj_r2 = "Total", unique_contribution = "Unique")) %>%
  mutate(value = as.numeric(gsub("%", "", value))/100) %>%
  mutate(predictor_set = make_pretty_names(predictor_set)) %>%
  mutate(Ho = "Ho")

gg_df_gea <- 
  varpart_result_gea$summary_table %>%
  select(-r2) %>%
  pivot_longer(cols = c("adj_r2", "unique_contribution"), names_to = "metric", values_to = "value") %>%
  mutate(metric = recode(metric, adj_r2 = "Total", unique_contribution = "Unique")) %>%
  mutate(value = as.numeric(gsub("%", "", value))/100) %>%
  mutate(predictor_set = make_pretty_names(predictor_set)) %>%
  mutate(Ho = "Ho_gea")

gg_df <- 
  bind_rows(gg_df, gg_df_gea) %>% 
  mutate(Ho = 
    case_when(
      Ho == "Ho" ~ "Genome-wide",
      Ho == "Ho_gea" ~ "Adaptive"
    )
  ) %>%
  mutate(Ho = factor(Ho, levels = c("Genome-wide", "Adaptive")))

ylgnbl <- RColorBrewer::brewer.pal(9, "YlGnBu")
ylgnbl_lite <- (ylgnbl[-c(1, 7:9)])

varpart_plt <- 
  ggplot(gg_df, aes(x = metric, y = predictor_set, fill = value)) +
  geom_tile(color = "white", lwd = 1) +
  geom_text(aes(label = scales::percent(value, accuracy = 1)), color = "black") +
  scale_fill_gradientn(colors = ylgnbl_lite) +
  theme_minimal() +
  labs(x = "", y = "", fill = "Variance\nExplained") +
  scale_x_discrete(position = "top") +
  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        strip.text.x = element_text(size = 14, face = "bold"),
        strip.placement = "outside") +
  facet_wrap(~Ho, nrow = 1)

# Correlation plot
mod_vars <- c("NDVI", "gHM", "bio1", "cur_lgm", "lgm_lig", "csi_past", "glacier", "tmean_dif", "Q")
cormat <- cor(mod_df[, mod_vars])

cor_df <- 
  cormat %>%
  as.data.frame() %>%
  rownames_to_column(var = "var1") %>%
  pivot_longer(cols = -var1, names_to = "var2", values_to = "cor") %>%
  filter(!is.na(cor)) %>%
  mutate(var1_pretty = make_pretty_names(var1), var2_pretty = make_pretty_names(var2)) %>%
  mutate(cor = case_when(var1 == var2 ~ NA, TRUE ~ cor))

cor_plt <- 
  ggplot(cor_df, aes(x = var1_pretty, y = var2_pretty, fill = cor)) +
  geom_tile(color = "white", lwd = 1) +
  geom_text(aes(label = round(cor, 2)), color = "black") +
  scale_fill_gradient2(
    low = "cyan3",
    mid = "white",
    high = "orange",
    midpoint = 0,
    limits = c(-1, 1)
  ) +
  theme_minimal() +
  labs(x = "", y = "", fill = "Correlation") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 12),
        axis.text.y = element_text(size = 12))

# Plot variables correlated with Ho that are not in the final model
names(mod_vars) <- mod_vars
confounded_vars <- map(mod_vars, ~cor.test(mod_df$Ho, mod_df[[.x]])$p.value) < 0.05
confounded_vars <- names(confounded_vars[confounded_vars])
confounded_vars <- confounded_vars[!(confounded_vars %in% c("csi_past", "Q"))]

missig_plt <-
  mod_df %>%
  pivot_longer(all_of(confounded_vars)) %>%
  mutate(name = make_pretty_names(name)) %>%
  ggplot(aes(x = value, y = Ho)) +
  geom_point(aes(col = cur_lgm)) +
  geom_smooth(method = "lm") +
  theme_classic() +
  labs(x = "", y = "Ho", col = "Paleoclimate change\n(LGM - LIG)") +
  scale_color_viridis_c(option = "mako") +
  facet_wrap(~name, scales = "free", strip.position = "bottom", nrow = 2) +
  theme(
    strip.placement = "outside", 
    strip.background = element_blank(), 
    strip.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  ) +
  stat_cor(label.y = min(mod_df$Ho, na.rm = TRUE)) 

names(mod_vars) <- mod_vars
confounded_vars <- map(mod_vars, ~cor.test(mod_df$Ho_gea, mod_df[[.x]])$p.value) < 0.05
confounded_vars <- names(confounded_vars[confounded_vars])
confounded_vars <- confounded_vars[!(confounded_vars %in% names(model_Ho_gea$coefficients))]
length(confounded_vars) # No confounded vars for Ho_gea
# missig_plt_gea <-
#   mod_df %>%
#   pivot_longer(all_of(confounded_vars)) %>%
#   mutate(name = make_pretty_names(name)) %>%
#   ggplot(aes(x = value, y = Ho)) +
#   geom_point(aes(col = lgm_lig)) +
#   geom_smooth(method = "lm") +
#   theme_classic() +
#   labs(x = "", y = "Ho", col = "Paleoclimate change\n(LGM - LIG)") +
#   scale_color_viridis_c(option = "mako") +
#   facet_wrap(~name, scales = "free", strip.position = "bottom", nrow = 1) +
#   theme(strip.placement = "outside", strip.background = element_blank()) +
#   stat_cor(label.y = min(mod_df$Ho, na.rm = TRUE)) 

##combo <- plot_grid(cor_plt, varpart_plt, nrow = 1, rel_widths = c(5, 3), labels = c("A", "B"))
##plot_grid(combo, missig_plt, nrow = 2, rel_heights = c(0.6, 0.35), labels = c("", "D"))


combo <- plot_grid(cor_plt, varpart_plt, nrow = 1, rel_widths = c(5, 4), labels = c("A", "B"))
pdf(here(plotpath, "confounding.pdf"), width = 17, height = 13)
plot_grid(combo, missig_plt, nrow = 2, rel_heights = c(0.4, 0.5), labels = c("", "D"))
dev.off()
```

```{r}
ca <- get_ca()

colpal = "plasma"
sig_vars <- c("csi_past", "Q")
plot_map <- function(var){
  ggplot(st_as_sf(mod_df)) +
    geom_sf(data = ca) +
    geom_sf(aes(fill = .data[[var]]), pch = 21, col = "black", cex = 2) +
    theme_void() +
    scale_fill_viridis_c(option = colpal) +
    labs(fill = gsub(" ", "\n", make_pretty_names(var)))
}

predicted_Ho <- data.frame(SampleID = mod_sf$SampleID, predicted = model_Ho$fitted, metric = "Ho")
predicted_Ho_gea <- data.frame(SampleID = mod_sf$SampleID, predicted = model_Ho_gea$fitted, metric = "Ho_gea")
mod_df$predicted <- model_Ho$fitted
mod_df$predicted_gea <- model_Ho_gea$fitted

gg_df <- 
  mod_df %>%
  pivot_longer(c("Ho", "Ho_gea"), names_to = "metric", values_to = "Ho") %>%
  left_join(bind_rows(predicted_Ho, predicted_Ho_gea))

mod_plt <-
  ggplot(mod_df, aes(x = predicted, y = Ho)) +
  geom_point(aes(fill = Ho), pch = 21, col = "black", alpha = 0.8, cex = 2) +
  geom_smooth(method = "lm", color = "black") +
  theme_classic() +
  labs(x = "Predicted genome-wide heterozygosity", y = "Observed genome-wide heterozygosity") +
  scale_fill_viridis_c(option = colpal) +
  stat_cor(label.y = min(mod_df$Ho, na.rm = TRUE)) +
  theme(legend.position = "none") 

mod_plt_gea <-
  ggplot(mod_df, aes(x = predicted_gea, y = Ho_gea)) +
  geom_point(aes(fill = Ho), pch = 21, col = "black", alpha = 0.8, cex = 2) +
  geom_smooth(method = "lm", color = "black") +
  theme_classic() +
  labs(x = "Predicted adaptive heterozygosity", y = "Observed adaptive heterozygosity") +
  scale_fill_viridis_c(option = colpal) +
  stat_cor(label.y = min(mod_df$Ho_gea, na.rm = TRUE)) +
  theme(legend.position = "none") 

map_plts <- map(c("Ho", sig_vars), plot_map)
map_plts_gea <- map(c("Ho_gea", sig_vars), plot_map)
#lm_plts <- map(sig_vars, ~gglm(.x, "Ho", mod_df))
partial_plts <- map(sig_vars, ~ggpartial(.x, "Ho", sig_vars, mod_df, col = "Ho", cex = 2) + scale_fill_viridis_c(option = colpal))
partial_plts_gea <- map(sig_vars, ~ggpartial(.x, "Ho_gea", sig_vars, mod_df, col = "Ho_gea", cex = 2) + scale_fill_viridis_c(option = colpal))

#plts <- c(map_plts, list(mod_plt), lm_plts, list(blank = ggplot() + theme_void()), partial_plts)
plts <- do.call(plot_grid, c(map_plts[1], partial_plts, list(mod_plt), nrow = 1))
plts_gea <- do.call(plot_grid, c(map_plts_gea[1], partial_plts_gea, list(mod_plt_gea), nrow = 1))

#labels <- LETTERS[1:(length(plts) - 1)]
#labels[which(names(plts) == "blank")] <- ""

pdf(here(plotpath, "main_figure.pdf"), width = 17, height = 9) 
plot_grid(plts, plts_gea, nrow = 2)
dev.off()
```

# Committee meeting

```{r}
pdf(here(plotpath, "committee_meeting.pdf"), width = 15, height = 3.5)
sig_vars <- c("csi_past", "tmean_dif", "bio1", "Q")
partial_plts <- map(sig_vars, ~ggpartial(.x, "Ho", sig_vars, mod_df, col = "Ho", cex = 2) + scale_fill_viridis_c(option = colpal))
do.call(plot_grid, c(partial_plts, list(nrow = 1)))
dev.off()

pdf(here(plotpath, "committee_meeting.pdf"), width = 5, height = 4)
plot_map("Ho") + scale_fill_viridis_c(option = "mako")
dev.off()
```
# Test of what variables are no longer significant after accounting for spatial autocorrelation

In the lm() bio1/tmean_dif is significant but in the spatial model it is not.
```{r}
# Compare significant variables from ols and sem models
summary(ols_model) %>% broom::tidy() %>% filter(term != "(Intercept)", term != "lambda") %>% filter(p.value < 0.05)
summary(sem_model) %>% broom::tidy() %>% filter(term != "(Intercept)", term != "lambda") %>% filter(p.value < 0.05)

# All variables except NDVI have M > 0.5, lgm_lig has the highest M (0.7)
names(mod_vars) <- mod_vars
map(mod_vars, ~moran.test(pull(mod_sf, .x), listw)$estimate) %>% 
  bind_rows(.id = "var") %>% 
  arrange(desc(`Moran I statistic`))
```

# Biodiversity comparison
```{r}
richness <- rast(here("data", "biodiversity_mapping", "Reptiles_total_richness.tif"))
df <- 
  left_join(get_coords(sf = TRUE), get_het()) %>%
  mutate(`Reptile species richness` = extract(richness, get_coords(sf = TRUE), ID = FALSE)[,1]) %>%
  mutate(`Reptile species richness` = as.numeric(as.character(`Reptile species richness`))) %>%
  drop_na(Ho) %>%
  # Drop Island individual 
  # REMOVE CHANNEL ISLAND SAMPLE
  filter(SampleID != "Scelocci_CCGPMC_MW01-3-14")

pdf(here(plotpath, "richness_vs_ho.pdf"), width = 13, height =4)
plt1 <- 
  gglm("Reptile species richness", "Ho", df) + 
  scale_color_viridis_c(option = "plasma")
plt2 <-
  ggplot(df) +
  geom_sf(data = ca) +
  geom_sf(aes(col = Ho)) +
  scale_color_viridis_c(option = "plasma") +
  theme_void()
plt3 <-
  ggplot(df) +
  geom_sf(data = ca) +
  geom_sf(aes(col = `Reptile species richness`)) +
  labs(col = "Reptile\nspecies\nrichness") +
  scale_color_viridis_c(option = "plasma") +
  theme_void()
plot_grid(plt1, plt2, plt3, nrow = 1)
dev.off()
```