# Libraries
```{r}
library("here")
library("tidyverse")
library("ggpubr")
library("terra")
library("sf")
library("gridExtra")
library("rpaleoclim")
library("MuMIn")
library("cowplot")
library("performance")
library("gstat")
library("sp")
library("furrr")
library("MetBrewer")
library("nlme")
library("broom.mixed")
```

# Paths for outputs
```{r}
outpath <- here("analysis", "genetic_diversity", "outputs")
plotpath <- here("analysis", "genetic_diversity", "plots")
```

# Custom functions
```{r}
source(here("general_functions.R"))
source(here("analysis", "genetic_diversity", "functions_genetic_diversity.R"))
source(here("analysis", "genetic_diversity", "functions_paleovars.R"))
```

# California polygon for plotting
```{r}
ca <- get_ca()
ca_proj <- ca %>% st_transform(3310)
```

# Coordinates
```{r}
coords <- get_coords(sf = TRUE)
coords_proj <- coords %>% st_transform(crs = st_crs(ca_proj))
```

# Heterozygosity data
```{r}
# Get genome-wide heterozygosity
het <- get_het("58-Sceloporus_chr.het") 

# Make sure you have the expected number of samples
stopifnot(nrow(het) == 259)

# Count number of sites used to calculate heterozygosity
mean(het$`N(NM)`, na.rm = TRUE) # 59955192
```

# Hypothesis 1: Genetic diversity explained by paleoclimate stability

```{r}
# Get current climate data (NOTE: not using paleoclim data because they use CHELSA V1.2 instead of V2.1)
bioclim <- rast(here("data", "env", "california_chelsa_bioclim_1981-2010_V.2.1.tif"))
cur <- bioclim[["CHELSA_bio1_1981-2010_V.2.1"]]
names(cur) <- "cur"

# Get LGM climate data (NOTE: not using paleoclim data because they use CHELSA V1.2 instead of V2.1)
lgm <- rast(here("data", "env", "california_chelsa_pmip_ccsm4.tif"))
names(lgm) <- "lgm"

# Get paleoclimate data (BIO1) from rpaleoclim and convert from C/10 to C
# Using time periods back until about ~ 1 mya (when Sceloporus diverged in S. CA)
#
# code	period	bp	source
# cur	Current (1979 – 2013)		CHELSA
# lh	Late Holocene: Meghalayan	4.2-0.3 ka	Fordham  et al. 2017
# mh	Mid Holocene: Northgrippian	8.326-4.2 ka	Fordham  et al. 2017
# eh	Early Holocene: Greenlandian	11.7-8.326 ka	Fordham  et al. 2017
# yds	Pleistocene: Younger Dryas Stadial	12.9-11.7 ka	Fordham  et al. 2017
# ba	Pleistocene: Bølling-Allerød	14.7-12.9 ka	Fordham  et al. 2017
# hs1	Pleistocene: Heinrich Stadial 1	17.0-14.7 ka	Fordham  et al. 2017
# lgm	Pleistocene: Last Glacial Maximum	ca. 21 ka	CHELSA
# lig	Pleistocene: Last Interglacial	ca. 130 ka	Otto-Bliesner et al. 2006
# mis19	Pleistocene: MIS19	ca. 787 ka	Brown et al. 2018
paleovars <- get_paleovars(c("lh", "mh", "eh", "yds", "ba", "hs1", "lig", "mis19"))
paleobio1 <- map(paleovars, ~.x$bio_1) %>% rast()/10

# Stack CHELSA rasters
chelsabio1 <- list(cur, lgm)

# Resample the higher resolution data (CHELSA) to match the lower resolution (rpaleoclim)
# This is the most conservative approach to avoid over-interpreting fine-scale variation 
csi_stack <- 
  c(as.list(chelsabio1), as.list(paleobio1)) %>% 
  map(~resample(.x, paleobio1[[1]]), method = "bilinear") %>% 
  rast()

# Calculate climate stability based on the standard deviation over the stack of rasters
# Don't remove NAs, use imputation later (NA values are at some points along the coast)
# If you remove NAs you will get some outliers because some time periods have data and others do not
csi_past <- app(csi_stack, sd, na.rm = FALSE)

# Scale CSI raster from 0 to 1 and subtract from 1 to get higher values = higher stability
csi_past[] <- 1 - scales::rescale(csi_past[], to = c(0, 1), na.rm = TRUE)

# Make a data frame with the climate stability values
paleoclim_mod_df <- 
  data.frame(
    SampleID = coords$SampleID,
    csi_past = terra::extract(csi_past, coords, ID = FALSE)[,1]
  )

# Check NA values (should be <10)
print(colSums(is.na(paleoclim_mod_df)))

# Some coastal points have NA values, so use a bilinear approach to get take the average of the four nearest raster cells
paleoclim_mod_df$csi_past <- bilinear_impute(paleoclim_mod_df$csi_past, csi_past, coords)

# Re-check NA values (should be 0)
print(colSums(is.na(paleoclim_mod_df)))
```

# Hypothesis 2: Genetic diversity explained by admixture
```{r}
# Read in admixture results produced by analysis/admixture/notebook_admixture.Rmd
q <- read_csv(here("analysis", "admixture", "outputs", "Q8.csv"))
q_r <- rast(here("analysis", "admixture", "outputs", "max_admix.tif"))

q_mod_df <- 
  q %>%
  dplyr::select(SampleID, Q) %>%
  # Calculate 1 - Q to make an "admixture score" whereby higher values = more admixture
  mutate(Q = 1 - Q)
```

# Hypothesis 3: Genetic diversity explained by current environment 
```{r}
# Read in contemporary environmental layers downloaded using the scripts in data_processing/env
bio1 <- bioclim[["CHELSA_bio1_1981-2010_V.2.1"]] # mean annual temperature
ndvi <- rast(here("data", "env", "california_ndvi_mean_2000_2020.tif")) # mean NDVI
ghm <- rast(here("data", "env", "california_global_human_modification_2016.tif")) # global human modification index

# Combine into df
env_mod_df <- data.frame(
  SampleID = coords$SampleID,
  NDVI = terra::extract(ndvi, coords, ID = FALSE)[, "NDVI"],
  gHM = terra::extract(ghm, coords, ID = FALSE)[,"gHM"],
  bio1 = terra::extract(bio1, coords, ID = FALSE)[,"CHELSA_bio1_1981-2010_V.2.1"]
)

# Check NA values
print(colSums(is.na(env_mod_df)))

# Some coastal points have NA values for gHM and bio1, so use a bilinear approach to get take the average of the four nearest raster cells
# Note: This function only imputes missing values, not all values
env_mod_df$gHM <- bilinear_impute(env_mod_df$gHM, ghm, coords)
env_mod_df$bio1 <- bilinear_impute(env_mod_df$bio1, bio1, coords)

# Double check NA values (should be 0)
print(colSums(is.na(env_mod_df)))
```

# Hypothesis 4: Genetic diversity explained by glaciation
```{r}
# Read in LGM polygon provided by Michelle Koo (MVZ)
lgm_glaciers <- st_read(here("data", "env", "Glaciotect_NA_lgm", "shp", "LGM_maxextent.shp"))

# Transform to same CRS as coords
# Note: have to first intersect and then transform otherwise you get a node error
lgm_glaciers <- st_intersection(lgm_glaciers, st_transform(ca, crs = st_crs(lgm_glaciers)))
lgm_glaciers <- st_transform(lgm_glaciers, crs = st_crs(ca))

# Add a new column to coords indicating whether each point is inside the LGM glacial extent
glacier_coords <- coords %>% st_transform(st_crs(lgm_glaciers))
glacier_mod_df <- 
  coords %>%
  mutate(glacier = as.numeric(lengths(st_within(glacier_coords, lgm_glaciers)) > 0)) %>%
  st_drop_geometry()
```

# Hypothesis 5: Genetic diversity explained by recent climate change
PRISM mean temperature data from historical data: https://prism.oregonstate.edu/historical/
```{r}
# Get historical prism data (https://prism.oregonstate.edu/historical/)
# Processed by scripts in data_processing/env
get_prism <- function(year, output = "df") {
  # Create PRISM path
  prism_path <- here("data", "env", "prism", "PRISM_tmean_stable_4kmM3", year)

  # List annual file
  file <- list.files(prism_path, pattern = paste0(year, "_*.bil.bil$"), full.names = TRUE)

  # Load data
  raster <- terra::rast(file)

  if (output == "raster"){
    return(raster)
  } else {
    vals <- terra::extract(raster, st_transform(coords, st_crs(raster)), ID = FALSE)
    names(vals) <- "tmean"
    df <- data.frame(year = year, tmean = vals, SampleID = coords$SampleID)
    return(df)
  }
}

# Loop over the years of interest
historical_years <- 1895:1905 # Historical decade
current_years <- 2010:2020 # Contemporary decade

# Make dataframe of average temperature by year for each sample
tmean_his <- 
  map(
    historical_years, 
    ~get_prism(year = .x), 
    .progress =TRUE
  ) %>% 
  bind_rows()  

tmean_cur <- 
  map(
    current_years, 
    ~get_prism(year = .x), 
    .progress =TRUE
  ) %>% 
  bind_rows()  

# Calculate mean temperature for each sample over the decade
tmean_his <- 
  tmean_his %>% 
  group_by(SampleID) %>% 
  summarise(tmean_his = mean(tmean, na.rm = TRUE))

tmean_cur <- 
  tmean_cur %>% 
  group_by(SampleID) %>% 
  summarise(tmean_cur = mean(tmean, na.rm = TRUE))

# Get rasters per year and average over the decade
tmean_his_r <-
   map(historical_years, ~get_prism(year = .x, output = "raster"), .progress =TRUE) %>% 
   rast() %>% 
   mean()

tmean_cur_r <- 
  map(current_years, ~get_prism(year = .x, output = "raster"), .progress =TRUE) %>% 
  rast() %>% 
  mean()

# Calculate difference between current and historical
tmean_dif_r <- tmean_cur_r - tmean_his_r

# Crop to california and mask
tmean_dif <- crop(tmean_dif_r, ca)
tmean_dif <- mask(tmean_dif, st_transform(ca, st_crs(tmean_dif_r)))

# Calculate change
tmean_mod_df <- 
  left_join(tmean_his, tmean_cur) %>% 
  mutate(tmean_dif = tmean_cur - tmean_his) %>%
  right_join(coords) %>% 
  dplyr::select(SampleID, tmean_dif, tmean_cur)

# Check for NAs
tmean_mod_df %>% 
  summarise(across(everything(), ~sum(is.na(.))))
```

# Hypothesis 6: Genetic diversity explained by fire
```{r}
# Historical fire regime data
# Fire Return Interval (FRI)
ca_proj_buffer <-  ca_proj %>% st_transform(5070) %>% st_buffer(dist = 10000) # buffer to ensure full coverage when cropping 
rasters
fri_og <- rast("analysis/wildfire/LF2016_FRI_200_CONUS/Tif/LC16_FRI_200.tif") 
fri <- crop(fri_og, ca_proj_buffer)

# Percent Replacement (PFS)
pfs_og <- rast("analysis/wildfire/LF2016_PFS_200_CONUS/Tif/LC16_PFS_200.tif")
pfs <- crop(pfs_og, ca_proj_buffer)

# Vegetation Departure (VDep)
vdep_og <- rast("analysis/wildfire/LF2016_VDep_200_CONUS/Tif/LC16_VDep_200.tif")
vdep <- crop(vdep_og, ca_proj_buffer)

# Fire regime group (FRG)
bps_og <- rast("analysis/wildfire/LF2016_BPS_200_CONUS/Tif/LC16_BPS_200.tif")
bps <- crop(bps_og, ca_proj_buffer)

# Activate relevant categories for each raster
activeCat(fri)<- "FRI_ALLFIR"

activeCat(vdep) <- "LABEL"

pfs_replac <- pfs
pfs_mixed <- pfs
pfs_surfac <- pfs
activeCat(pfs_replac) <- "PRC_REPLAC"
activeCat(pfs_mixed) <- "PRC_MIXED"
activeCat(pfs_surfac) <- "PRC_SURFAC"

activeCat(bps) <- "GROUPVEG"
```

```{r}
fire_list <- list(pfs_replac, pfs_mixed, pfs_surfac, fri, vdep)
names(fire_list) <- c("pfs_replac", "pfs_mixed", "pfs_surfac", "fri", "vdep")

coords_fire_proj <- st_transform(get_coords(sf = TRUE), crs = st_crs(fri))
fire_mod_df <- 
  # Extract fire rasters
  map(fire_list, ~terra::extract(.x, coords_fire_proj)[,2], .progress = TRUE) %>% 
  bind_cols() %>%
  # Add vegetation
  mutate(groupveg = terra::extract(bps, coords_fire_proj)[,2]) %>%
  # Add SampleIDs
  mutate(SampleID = coords_fire_proj$SampleID) %>%
  # Make categorical variables for fire regime
  mutate(
    fri_cat = case_when(
      fri == -9999 ~ "No Fire Regime",                 # NA means no modeled fire
      fri <= 35 ~ "Frequent Fire (≤35 yrs)",
      fri <= 200 ~ "Infrequent Fire (35–200 yrs)",
      fri > 200 ~ "Very Infrequent Fire (>200 yrs)"
    )
  ) %>%
  mutate(
    pfs_replac_cat = case_when(
      pfs_replac == -9999 ~ "No Fire Regime",
      #pfs_replac == -9999 ~ "Low Replacement (<33%)", # Treat as 0% replacement, otherwise total overlap with FRI no fire regime
      pfs_replac < 33 ~ "Low Replacement (<33%)",
      pfs_replac < 67 ~ "Moderate Replacement (33–66%)",
      pfs_replac >= 67 ~ "High Replacement (>66%)"
    )
  ) %>%
  # Convert -9999 to 0 for pfs values (-9999 are areas with no fire, so 0% of area burned at any severity)
  mutate(across(starts_with("pfs"), ~ifelse(.x == -9999, 0, .x))) %>%
  # Combined mixed and low PFS into a single "low" category and rename pfs_replac to "high"
  mutate(
    pfs_sum = pfs_replac + pfs_mixed + pfs_surfac,
    pfs_low = (pfs_mixed + pfs_surfac)/pfs_sum,
    pfs_high = pfs_replac/pfs_sum
  ) %>%
  # Convert -9999 to NA for fri (these will be dealt with later)
  mutate(fri = ifelse(fri == -9999, NA, fri)) %>%
  # Convert vdep to fully numeric by coding categories (urban/agriculture/etc.) as 101 (highest departure)
  # A warning is normal here because of coercion to numeric
  mutate(
    vdep = as.numeric(as.character(vdep)),
    vdep = ifelse(is.na(vdep), 101, vdep)
  ) 

# Check which points have "Open Water" vegetation
# These seem to be next to lakes
coords[fire_mod_df$groupveg == "Open Water",]

# Confirm that pfs_low and pfs_high are perfectly colinear
cor(fire_mod_df$pfs_low, fire_mod_df$pfs_high, use = "complete.obs") # -1

# Impute Open Water values for extracted values based on neighbors
na_i <- which(fire_mod_df$groupveg == "Open Water")
na_coords <- coords_fire_proj[na_i, ]

# Get the cell that contains each coordinate
xy <- na_coords %>% st_coordinates()

groupveg_na_vals <-
  map_vec(1:nrow(xy), ~{
    # Get focal cell
    focal_cell <- cellFromXY(bps, xy[.x, , drop = FALSE])
    stopifnot(!is.na(focal_cell))

    # Get neighboring (bordering) cells
    # directions = 4 (rook neighbors) or 8 (queen neighbors incl. diagonals)
    nbr_cells <- adjacent(bps, cells = focal_cell, directions = 8, pairs = FALSE)
    nbr_cells <- as.vector(nbr_cells) # make into vector
    
    # Extract values from those neighboring cells
    nbr_vals <- bps[nbr_cells]

    # Combine in df
    nbr_df <- 
      data.frame(
        groupveg = nbr_vals[, "GROUPVEG"]
      ) %>%
      # Remove Open water
      filter(groupveg != "Open Water")

    if (nrow(nbr_df) == 0){
      # If no neighboring values, return NA
      return(NA)
    }

    # Get most common value of neighboring values
    mode_val <- 
      nbr_df %>%
      count(groupveg) %>%
      arrange(desc(n)) %>%
      slice(1) %>%
      pull(groupveg)
    return(mode_val)
  })

fire_mod_df$groupveg[fire_mod_df$groupveg == "Open Water"] <- groupveg_na_vals
unique(fire_mod_df$groupveg)

# Count NA values again
fire_mod_df %>% 
  summarise(across(everything(), ~sum(is.na(.))))

# Replace final NA values with 0 because these are locations that likely never burn (e.g., urban areas/very sparse vegetation)
# # From email: "If the BpS is not Sparsely Vegetated or Barren/Water/Rock/Snow, the BpS model does not include fire because model contributors did not think that fire was a critical disturbance shaping the BpS. This is most common in very wet or very dry ecosystems. For these systems an FRI of 0 may be appropriate."
fire_mod_df <- 
  fire_mod_df %>%
  mutate(fri = ifelse(is.na(fri), 0, fri))

# Check for NAs
fire_mod_df %>% 
  summarise(across(everything(), ~sum(is.na(.))))

# Check correlations between metrics (with a log correction because of skew)
cor(log(fire_mod_df[, c("pfs_replac", "pfs_surfac", "pfs_low", "pfs_high", "fri")] + 1), use = "pairwise.complete.obs")
# In the end I am choosing: 
# fire severity = pfs_replac (proportion of high severity fires) which is almost perfectly correlated with pfs_low (proportion of low + mixed severity fires), but 0 values can be either high proportion of low severity fire or no fire at all
# fire frequency = fri (fire return interval)
fire_mod_df <-
  fire_mod_df %>%
  mutate(
    fire_severity = log1p(pfs_replac),
    fire_frequency = log1p(fri),
    fire_severity_cat = factor(pfs_replac_cat, levels = c("No Fire Regime", "Low Replacement (<33%)", "Moderate Replacement (33–66%)", "High Replacement (>66%)")),
    fire_frequency_cat = factor(fri_cat, levels = c("No Fire Regime", "Frequent Fire (≤35 yrs)", "Infrequent Fire (35–200 yrs)",  "Very Infrequent Fire (>200 yrs)"))
  )

# These two variables are correlated at 0.8, which is high
cor(fire_mod_df$fire_severity, fire_mod_df$fire_frequency, use = "pairwise.complete.obs") 

# Run PCA on fire_severity and fire_frequency
fire_pc <- prcomp((fire_mod_df[, c("fire_frequency", "fire_severity")]), center = TRUE, scale. = TRUE)$x[,1]
# Flip so increasing fire pc = increasing fire severity/decreasing frequency
fire_mod_df$fire_pc <- fire_pc * -1

# Count NA values again
fire_mod_df %>% 
  summarise(across(everything(), ~sum(is.na(.)))) %>%
  select(where(~. > 0))
```

```{r}
test_df <- 
  get_het() %>% 
  right_join(fire_mod_df) %>% 
  left_join(get_coords(sf = TRUE)) %>%
  st_as_sf() %>%
  left_join(q_mod_df)
 
pdf(here(plotpath, "fire_variables.pdf"), width = 6.5, height = 4)
ggplot(test_df, aes(x = fire_severity, y = Ho)) +
  geom_point(aes(col = fire_severity_cat)) +
  geom_smooth(method = "lm") +
  theme_classic() + 
  labs(x = "log1p(% High Severity Fire)", col = "Bins") +
  stat_cor()

ggplot(test_df, aes(x = fire_frequency, y = Ho)) +
  geom_point(aes(col = fire_frequency_cat)) +
  geom_smooth(method = "lm") +
  theme_classic() + 
  labs(x = "log1p(Fire Return Interval)", col = "Bins") +
  stat_cor()

ggplot(test_df, aes(x = fire_pc, y = Ho)) +
  geom_point(aes(col = fire_frequency_cat, pch = fire_severity_cat)) +
  geom_smooth(method = "lm") +
  theme_classic() + 
  labs(x = "Fire PCA", col = "Frequency category", pch = "Severity category") +
  stat_cor()

ggplot(test_df) +
  geom_sf(data = ca) +
  geom_sf(aes(col = fire_frequency_cat, pch = fire_severity_cat)) +
  labs(col = "Frequency category", pch = "Severity category") +
  theme_void()

ggplot(test_df, aes(x = fire_pc, y = Ho)) +
  geom_point(aes(col = groupveg)) +
  geom_smooth(method = "lm") +
  theme_classic() + 
  labs(x = "Fire PCA", col = "Vegetation group") +
  stat_cor()
dev.off()

lm(
  Ho ~ fire_pc + groupveg + Q,
  data = test_df
) %>%
summary()

mod <- lm(
  Ho ~ pfs_replac_cat + fri_cat + groupveg,
  data = test_df
)
summary(mod)

library(emmeans)
emmeans_pfs <- emmeans(mod, specs = "pfs_replac_cat")
tukey_pfs <- contrast(emmeans_pfs, method = "tukey")
tukey_pfs

# Tukey pairwise comparisons for Fire Return Interval category
emmeans_fri <- emmeans(mod, specs = "fri_cat")
tukey_fri <- contrast(emmeans_fri, method = "tukey")
tukey_fri

count(fire_mod_df, fri_cat, pfs_replac_cat)
```

# Combine rasters into a raster stack
This can take a while because of the fire rasters
```r
env_list <- 
  list(
    NDVI = ndvi,
    bio1 = bio1,
    gHM = ghm,
    tmean_dif = tmean_dif,
    csi_past = csi_past, 
    Q = q_r,
    pfs_replac = fire_list$pfs_replac,
    pfs_mixed =fire_list$pfs_mixed, 
    pfs_surfac = fire_list$pfs_surfac, 
    fri = fire_list$fri, 
    vdep = fire_list$vdep
  )

# Project all rasters to EPSG:3310 and mask with CA
env_list <- imap(env_list, ~{
  r <- terra::project(.x, "epsg:3310", method = "bilinear")
  r <- mask(r, ca_proj)
  names(r) <- .y
  return(r)
}, .progress = TRUE)

# Output path for model rasters
path <- here("data", "env", "model_rasters")

# Make model rasters dir
dir.create(path, showWarnings = FALSE)

# Write out rasters
iwalk(env_list, ~writeRaster(.x, here(path, paste0(.y, ".tif")), overwrite = TRUE))

# Snap all rasters to same grid and resolution
# Note: these rasters are only used for visualization, so these changes don't affect the models
max_res <- map_dbl(env_list, ~res(.x)[1]) %>% which.max()
target_raster <- env_list[[max_res]] # coarsest resolution
resampled_list <- imap(env_list, ~{
  # Resample to match target
  terra::resample(.x, target_raster, method = "bilinear")
})

# Create raster stack
env_stack <- 
  resampled_list %>% 
  rast() 
```

# Create model data
```{r}
# Model variables
mod_vars <- c("NDVI", "gHM", "bio1", "Q", "csi_past", "glacier", "tmean_dif", "fire_severity", "fire_frequency", "fire_severity_cat", "fire_frequency_cat", "fire_pc", "vdep", "groupveg", "x", "y")

# Combine dataframes
mod_df <- 
  get_coords() %>% 
  left_join(het) %>%
  left_join(env_mod_df) %>%
  left_join(paleoclim_mod_df) %>%
  left_join(q_mod_df) %>%
  left_join(glacier_mod_df) %>%
  left_join(tmean_mod_df) %>%
  left_join(fire_mod_df) %>%
  dplyr::select(Ho, SampleID, all_of(mod_vars))

# Check sample number
# Note: het has 259 samples, but two of them are potentially flipped ("Scelocci_CAS213197", "Scelocci_CAS214858"), so we remove the from the model (they are not included in coords)
stopifnot(nrow(mod_df) == 257)

# Confirm no NAs:
stopifnot(all(colSums(is.na(mod_df)) == 0))

# Convert to sf
mod_sf <- 
  get_coords(sf = TRUE) %>% 
  right_join(mod_df) %>% 
  st_transform(3310) %>%
  select(-x, -y) %>%
  bind_cols(st_coordinates(.)) %>%
  rename(x = X, y = Y)
```

```{r}
# Export model df
export_df <- 
  st_drop_geometry(mod_df) %>% 
  dplyr::select(SampleID, everything())
write_csv(export_df, here(outpath, "model_df.csv"))
```

## Check for correlation between variables
```{r}
# Check correlations
# Compute the correlation matrix
numeric_mod_vars <-
  mod_df %>% 
  select(all_of(mod_vars)) %>%
  select(where(is.numeric)) %>% 
  colnames()

r <- cor(mod_df[, numeric_mod_vars], use = "pairwise.complete.obs")
# Set the diagonal and lower triangle to NA
r[lower.tri(r, diag = TRUE)] <- NA
r

# Find the pairs with correlations greater than 0.6
high_corr <- which(abs(r) > 0.6, arr.ind = TRUE)
# Print the pairs
high_corr_pairs <- data.frame(
  Var1 = rownames(r)[high_corr[, 1]],
  Var2 = rownames(r)[high_corr[, 2]],
  Correlation = r[high_corr]
)

# fire_severity and log_fri have a correlation of 0.67
high_corr_pairs
```

```{r}
#Make nice correlation plot
r <- stats::cor(mod_df[, setdiff(numeric_mod_vars, c("x", "y"))], use = "pairwise.complete.obs")
rownames(r) <- colnames(r) <- make_pretty_names(rownames(r))
r[lower.tri(r, diag = TRUE)] <- NA

library(corrplot)
png(here(plotpath, "correlation_plot.png"), width = 10*300, height = 10*300, units = "px", res = 300)
corrplot(r, method = "ellipse", type = "upper", tl.col = "black", tl.srt = 45, diag = FALSE, addCoef.col = "black", number.cex = 0.8)
dev.off()
```

```{r}
# Check for confounding with vegetation group and fire
anova(lm(fire_frequency ~ groupveg, data = mod_df)) # Significant ANOVA
lsr::etaSquared(lm(fire_frequency ~ groupveg, data = mod_df)) # Eta squared = 0.55 -> high confounding (>50% variance explained by veg group)

lm(Ho ~ fire_frequency, data = mod_df) %>% summary() # Significant effect of fire frequency alone
lm(Ho ~ fire_frequency + groupveg, data = mod_df) %>% summary() # No significant effect of fire frequency when controlling for vegetation group
lm(Ho ~ fire_severity + groupveg, data = mod_df) %>% summary() # Significant effect of fire severity when controlling for vegetation group
lm(Ho ~ fire_severity + groupveg + csi_past + bio1 + tmean_dif + glacier + gHM + NDVI, data = mod_df) %>% summary() # Fire severity significant
lm(Ho ~ groupveg + csi_past + bio1 + tmean_dif + Q + glacier + gHM + NDVI + Q, data = mod_df) %>% summary() # Adding Q makes fire severity non-significant

lm(Ho ~ fire_severity + fire_frequency + csi_past + bio1 + tmean_dif + glacier + gHM, data = mod_df) %>% summary()


lm(Q ~ fire_severity + csi_past + bio1 + tmean_dif + vdep + groupveg + gHM + NDVI + glacier + fire_frequency, data = mod_df) %>% summary() 
# Q is significantly positively affected by fire_severity and vegetation and negatively affected by NDVI

pdf(here(plotpath, "vegetation_plots.pdf"))
ggplot(mod_df, aes(x = groupveg, y = fire_frequency)) +
  geom_boxplot() +
  theme_minimal() 
ggplot(mod_df, aes(x = groupveg, y = Ho)) + 
  geom_boxplot() +
  theme_minimal() 
ggplot(mod_df, aes(x = fire_frequency, y = Ho)) +
  geom_point(aes(col = groupveg)) +
  geom_smooth(method = "lm") +
  theme_minimal() 
ggplot(mod_df, aes(x = fire_severity, y = Ho)) +
  geom_point(aes(col = groupveg)) +
  geom_smooth(method = "lm") +
  theme_minimal() 
dev.off()
```

# Run Ho models

```{r}
# Model formula variables
mod_vars <- c("glacier", "NDVI", "gHM", "bio1", "csi_past", "tmean_dif",  "fire_severity", "fire_frequency", "vdep", "groupveg", "Q")

# GLS can't handle 0 distances, so take the average of the coordinates (mode for veg type)
mode_val <- function(x) {
  ux <- na.omit(x)
  ux[which.max(tabulate(match(ux, unique(ux))))]
}
cat_mod_vars <- mod_df %>% 
  select(all_of(mod_vars)) %>%
  select(where(~is.character(.x) | is.factor(.x))) %>%
  colnames()
numeric_mod_vars <- setdiff(mod_vars, cat_mod_vars)

mod_sf_unique <- 
  mod_sf %>%
  group_by(x, y) %>%
  summarise(
    across(all_of(c(numeric_mod_vars, "Ho")), ~ mean(.x, na.rm = TRUE)),
    across(all_of(cat_mod_vars), ~ mode_val(.x)),
    .groups = "drop"
  )

# Scale data
mod_sf_unique_scaled <-
  mod_sf_unique %>%
  mutate(
    across(
      all_of(setdiff(numeric_mod_vars, "glacier")),
      ~ as.numeric(scale(.x))   # <- drop matrix
    )
  )

# Confirm that all glacier values are 0 and 1
stopifnot(all(mod_sf_unique_scaled$glacier %in% c(0, 1)))

# Convert glacier to factor
mod_sf_unique_scaled$glacier_factor <- as.factor(mod_sf_unique$glacier)
mod_sf_unique$glacier_factor <- as.factor(mod_sf_unique$glacier)

# Confirm no NA
stopifnot(all(colSums(is.na(mod_sf_unique_scaled)) == 0))
```

```{r}
# Run OLS and get residuals
f <- formula(paste0("Ho ~", paste(mod_vars, collapse = " + ")))
ols_model <- lm(f, data = mod_sf_unique)
summary(ols_model)
AIC(ols_model)
mod_sf_unique$ols_resids <- residuals(ols_model)
broom::tidy(ols_model) %>% filter(p.value < 0.05) %>% arrange(desc(abs(estimate)))

# Fit variograms to the residuals
ols_variogram <- variogram(ols_resids ~ 1, data = mod_sf_unique)
ols_vgm <- map(c("Exp", "Sph", "Gau"), ~fit.variogram(ols_variogram, model = vgm(.x)))
names(ols_vgm) <- c("Exp", "Sph", "Gau")

pdf(here(plotpath, "ols_variogram.pdf"))
plot(ols_variogram, ols_vgm[["Exp"]], main = "OLS Residuals Variogram (Exponential)")
plot(ols_variogram, ols_vgm[["Sph"]], main = "OLS Residuals Variogram (Spherical)")
plot(ols_variogram, ols_vgm[["Gau"]], main = "OLS Residuals Variogram (Gaussian)")
dev.off()
# Based on visual inspection, the exponential model seems to fit best

# Fit a GLS model
# Note: nugget=FALSE assumes residuals at the same location are perfectly correlated, which should align with heterozygosity, since you don't expect to have much microscale variation in heterozygosity.
# nugget=TRUE allows unstructured microscale variance (random noise at zero distance), which can absorb spatial signal and cause predictor effects to appear insignificant. 
gls_model <- gls(f, correlation = corExp(form = ~ x + y, nugget = FALSE), data = mod_sf_unique)
summary(gls_model)
# Use type = "normalized" to get residuals decorrelated and standardized by the fitted 
# spatial covariance matrix; this removes modeled spatial structure and makes them 
# suitable for Moran's I. If you don't do this Moran's I will appear higher than it actually is.
mod_sf_unique$gls_resids <- residuals(gls_model, type = "normalized")

# Plot residuals
plot_resids <- function(resids){
  # Remove max residual because it is hard to see spatial patterns with it
  ggplot(mod_sf_unique[mod_sf_unique[[resids]] != max(mod_sf_unique[[resids]]), ]) +
    geom_sf(data = ca_proj, col = NA) +
    geom_sf(aes(col = .data[[resids]])) +
    theme_void() +
    scale_color_viridis_c(option = "turbo", end = 0.1, begin = 0.9, direction = -1) +
    labs(title = paste("Residuals:", resids))
}

pdf(here(plotpath, "all_model_residuals.pdf"), width = 9, height = 4)
plt_list <- list(
  plot_resids("ols_resids"),
  plot_resids("gls_resids")
)
plot_grid(plotlist = plt_list, nrow = 1)
dev.off()
# The GLS model residuals look better than the OLS residuals

# Compare AIC
# You will get warning about the models not having the same number of observations, but the number of observations is the same for all models, so this is not an issue 
all_equal_nobs <- length(unique(c(
  nobs(ols_model), nobs(gls_model)
))) == 1  
stopifnot(all_equal_nobs)

AIC(ols_model, gls_model) # OLS model has lowest AIC, but it has spatial autocorrelation in the residuals

# Compare significant variables (p < 0.05) for each model using map
models <- list(OLS = ols_model, GLS = gls_model)
sig_vars_by_model <- map(models, ~ broom::tidy(.x) %>%
  filter(!term %in% c("lambda", "rho", "(Intercept)"), p.value < 0.05) %>%
  pull(term))
print(sig_vars_by_model) 
```

```{r}
library(spdep)

# Coordinates matrix
coords <- cbind(mod_sf_unique$x, mod_sf_unique$y)

# 8 nearest neighbors (robust for irregular sampling)
knn_8 <- knearneigh(coords, k = 8)
nb_8  <- knn2nb(knn_8)
lw_8  <- nb2listw(nb_8, style = "W")

# Moran's I for OLS residuals
moran_ols <- moran.test(mod_sf_unique$ols_resids, lw_8)
print(moran_ols)

# Moran's I for GLS normalized residuals
moran_gls <- moran.test(mod_sf_unique$gls_resids, lw_8)
print(moran_gls)
```

# Test model variable combinations
```{r}
library(MuMIn)
# Set NA action to "na.fail" to ensure models with missing data are not considered
options(na.action = "na.fail")

# Test all model combinations
dredge_result <- dredge(ols_model)
# Get top models based on delta AICc < 2
top_models <- get.models(dredge_result, subset = delta < 2)
# Get variables for each of top models
top_vars <- 
  top_models %>%
  map(~broom::tidy(.x) %>%
        filter(!term %in% c("(Intercept)")) %>%
        pull(term))

print(top_vars[[1]])
```

```{r}
# Global formula
f_full <- Ho ~ glacier + NDVI + gHM + bio1 + csi_past + tmean_dif + fire_severity + fire_frequency + vdep + groupveg + Q 

# Global GLS model (use ML for AIC-based selection)
gls_full <- gls(
  f_full,
  data = mod_sf_unique_scaled,
  method = "ML",
  correlation = corExp(form = ~ x + y, nugget = FALSE)
)

# All subsets model selection on the GLS
dredge_res <- dredge(gls_full, trace = TRUE)

# Look at the top of the table
head(dredge_res)

# Get top models based on delta AICc < 2
top_gls_models <- get.models(dredge_res, subset = delta < 2)

# Get variables used in each of the top models (excluding intercept)
top_vars_gls <- 
  top_gls_models %>%
  map(~ broom::tidy(.x) %>%
        filter(term != "(Intercept)") %>%
        pull(term))

# Print variables in top models
top_vars_gls

# Get best model
best_gls <- get.models(dredge_res, subset = 1)[[1]]
summary(best_gls)

# Refit best model with REML
final_gls <- update(best_gls, method = "REML")
summary(final_gls)
```

# Fit final model 
```{r}
# Create final formula
f <- formula("Ho ~ bio1 + csi_past + tmean_dif + fire_frequency + gHM + NDVI + glacier + groupveg + vdep")

# Fit model
final_model <- gls(f, correlation = corExp(form = ~ x + y, nugget = FALSE), data = mod_sf_unique_scaled, method = "REML") 

# Calculate VIF
vif_values <- car::vif(final_model)
stopifnot(!any(vif_values > 5)) # Check for multicollinearity issues

# Results
mod_result <- broom::tidy(final_model)
mod_result

# Significant variables
mod_result %>% filter(p.value < 0.05) %>% arrange(desc(abs(estimate))) 
sig_vars <- mod_result %>% filter(p.value < 0.05, !term %in% c("(Intercept)", "lambda")) %>% pull(term)
print(sig_vars)

# You can't get a traditional R2 for spatial models, but you can get an alternative R2
cor(predict(final_model), mod_sf_unique$Ho)^2 # 0.45
cor(predict(final_model), mod_sf_unique$Ho) # 0.67
```

# Figures
```{r}
# Define a color scale function using MetBrewer with an argument to extend the scale for better visualization
scale_met_cont <- function(values, aesthetic = c("fill", "colour"), extend_scale = 0.1, r_min  = NULL, r_max = NULL) {
  aesthetic <- match.arg(aesthetic)
  if (is.null(r_min)) r_min <- min(values, na.rm = TRUE)
  if (is.null(r_max)) r_max <- max(values, na.rm = TRUE)
  rng   <- r_max - r_min
  lims  <- c(r_min - rng*extend_scale, r_max + rng*extend_scale)

  pal <- rev(MetBrewer::met.brewer("Hiroshige", type = "continuous"))
  if (aesthetic == "fill") {
    ggplot2::scale_fill_gradientn(colors = pal, limits = lims, oob = scales::squish)
  } else {
    ggplot2::scale_colour_gradientn(colors = pal, limits = lims, oob = scales::squish)
  }
}

# Make residual plots for model
plot_mod <- function(x, y, df, model, 
                     mod_vars = c("NDVI","gHM","bio1","csi_past","glacier","tmean_dif", "pfs_replac", "pfs_mixed", "pfs_surfac", "fri", "vdep","Q"),
                     color_var = "x", residual = TRUE) {

  # Calculate residuals for y variable given x variable
  if (residual) df[[y]] <- residuals(model) + coef(model)[[x]] * as.numeric(scale(df[[x]]))

  # Choose whether to color by x or y variable
  col_var <- if (color_var == "x") x else y

  # Extract beta and p-value for x variable
  model_summary <- broom::tidy(model)
  # Convert to scientific notation for plotting
  beta <- formatC(model_summary$estimate[model_summary$term == x], format = "e", digits = 2)
  pval <- model_summary$p.value[model_summary$term == x]
  pval <- ifelse(pval < 0.001, "< 0.001", paste0("= ", round(pval, 3)))

  plt <-
    ggplot(df, aes(x = .data[[x]], y = .data[[y]])) +
    geom_point(aes(col = .data[[col_var]])) +
    geom_smooth(method = "lm", colour = "black") +
    scale_met_cont(df[[col_var]], aesthetic = "colour", extend_scale = 0.1) +
    annotate("text", x = Inf, y = Inf, hjust = 1.1, vjust = 1.5,
             label = paste0("β = ", beta, "\n", "p ", pval), size = 6) +
    xlab(stringr::str_to_sentence(make_pretty_names(x))) +
    theme_classic() +
    theme(legend.position = "none",
          axis.title = element_text(size = 16),
          axis.text  = element_text(size = 14))

  if (residual) {
    plt + labs(y = "Partial heterozygosity")
  } else {
    plt + labs(y = "Observed heterozygosity")
  }

}

# Load polygons for plotting
ca_proj <- get_ca() %>% sf::st_transform(3310) # California boundaries
range_map <- get_range() # Sceloporus range map

# Loaded wingen map
wingen_r <- 
  terra::rast(here(outpath, "Ho_kriged_masked.tif")) %>% 
  mask(range_map) %>% 
  mask(ca_proj) %>%
  as.data.frame(xy = TRUE, na.rm = TRUE) 
names(wingen_r)[3] <- "layer"  # ensure 'layer' is the raster value column

# Isolate significant vs all variables
sig_vars <- c("csi_past", "Q", "tmean_dif", "bio1", "fire_frequency")
mod_vars <- c("NDVI", "gHM", "bio1", "csi_past", "glacier", "tmean_dif", "fire_frequency", "vdep", "Q")

# A: wingen map of heterozygosity 
ho_plot <- ggplot(wingen_r) +
  geom_sf(data = ca_proj, fill = "gray90", colour = NA, size = 2) +
  geom_tile(aes(x = x, y = y, fill = layer), colour = NA) +
  scale_met_cont(wingen_r$layer, aesthetic = "fill", extend_scale = 0.1) +
  labs(fill = "Heterozygosity") +
  theme_void() +
  geom_sf(data = mod_sf_unique, shape = 21, aes(fill = Ho), colour = "#00000087", size = 2) +
  theme(
    plot.margin = margin(1, 1, 1, 1, unit = "mm"),
    legend.position = "inside",              # turn on "inside" mode
    legend.justification = c(1, 1),          # anchor legend's top-right corner
    legend.position.inside = c(0.9, 0.9),   # place near panel's top-right
    legend.key.height = unit(0.9, "cm"),
    legend.key.width  = unit(0.9, "cm"),
    legend.title = element_text(size = 16),
    legend.text  = element_text(size = 14)
  )

# B: partial residuals 
plts_residual_y <- purrr::map(
  sig_vars, 
  # Use unscaled data for plotting
  ~plot_mod(.x, "Ho", mod_sf_unique, model = final_model, color_var = "y", residual = TRUE)
)
# B: Raw values
plts_raw_y <- purrr::map(
  sig_vars, 
  # Use unscaled data for plotting
  ~plot_mod(.x, "Ho", mod_sf_unique, model = final_model, color_var = "y", residual = FALSE)
)


# C: predicted vs observed 
pred_vs_obs <-
  # Use unscaled data for plotting
  ggplot(mod_sf_unique, aes(x = predict(final_model), y = Ho)) +
  geom_point(aes(col = Ho), shape = 16, size = 2) +
  geom_smooth(method = "lm", colour = "black") +
  scale_met_cont(c(mod_sf_unique$Ho, wingen_r$layer), aesthetic = "colour", extend_scale = 0.1) +
  theme_classic() +
  labs(x = "Predicted heterozygosity", y = "Observed heterozygosity") +
  ggpubr::stat_cor(size = 6) +
  theme(legend.position = "none",
        axis.title = element_text(size = 16),
        axis.text  = element_text(size = 14))

pred_vs_obs_adj <- cowplot::plot_grid(pred_vs_obs, NULL, nrow = 1, rel_widths = c(0.9, 0.1))

plt2_residual <- cowplot::plot_grid(plotlist = plts_residual_y, ncol = 2)
plt2_raw <- cowplot::plot_grid(plotlist = plts_raw_y, ncol = 2)
plt1 <- cowplot::plot_grid(plotlist = list(ho_plot, pred_vs_obs_adj), ncol = 1,
                           labels = c("A", "C"), rel_heights = c(0.6, 0.4), label_size = 18)

png(here(plotpath, "figure_Ho_residual.png"), height = 3.45*4*300*0.9, width = 4.1*4*300*0.9, res = 300)
cowplot::plot_grid(plt1, plt2_residual, ncol = 2,labels = c("", "B"),  rel_widths = c(0.4, 0.6), align = "h", label_size = 18)
dev.off()

png(here(plotpath, "figure_Ho_raw.png"), height = 3.5*4*300*0.9, width = 4.1*4*300*0.9, res = 300)
cowplot::plot_grid(plt1, plt2_raw, ncol = 2,labels = c("", "B"),  rel_widths = c(0.4, 0.6), align = "h", label_size = 18)
dev.off()
```


```{r}
pop_df <- 
  get_pops() %>%
  left_join(get_coords(sf = TRUE)) %>% 
  st_as_sf() %>%
  st_transform(3310) %>%
  bind_cols(st_coordinates(.)) %>%
  st_drop_geometry() %>%
  group_by(X, Y) %>%
  # Take the mean but it doesn't matter since all samples in a pop have the same cluster
  summarise(cluster = mean(as.numeric(cluster))) %>%
  mutate(cluster = factor(cluster)) %>%
  rename(x = X, y = Y) %>%
  ungroup() %>%
  right_join(mod_sf_unique) %>%
  st_as_sf()

# Color scales
group1 <- c("8", "6", "2")
group2 <- c("4", "3", "7")
group3 <- c("1", "5")
hiroshige_cols <- rev(c("#042346", "#295580", "#6bb2cc", "#9cd3dd", "#fecd6d", "#f7a957", "#ec7b4c", "#f04f32"))
cs1 <- hiroshige_cols[1:3]
names(cs1) <- group1
cs2 <- hiroshige_cols[4:6]
names(cs2) <- group2
cs3 <- hiroshige_cols[7:8]
names(cs3) <- group3
pop_cols <- c(cs1, cs2, cs3)

# Make residual plots for model
plot_mod_pop <- function(x, y, df, model, 
                     mod_vars = c("NDVI","gHM","bio1","csi_past","glacier","tmean_dif", "pfs_replac", "pfs_mixed", "pfs_surfac", "fri", "vdep","Q"), residual = TRUE) {

  # Calculate residuals for y variable given x variable
  if (residual) df[[y]] <- residuals(model) + coef(model)[[x]] * as.numeric(scale(df[[x]]))

  # Extract beta and p-value for x variable
  model_summary <- broom::tidy(model)
  # Convert to scientific notation for plotting
  beta <- formatC(model_summary$estimate[model_summary$term == x], format = "e", digits = 2)
  pval <- model_summary$p.value[model_summary$term == x]
  pval <- ifelse(pval < 0.001, "< 0.001", paste0("= ", round(pval, 3)))

  plt <-
    ggplot(df, aes(x = .data[[x]], y = .data[[y]])) +
    geom_point(aes(col = cluster), size = 2, alpha = 0.8) +
    geom_smooth(method = "lm", colour = "black") +
    scale_color_manual(values = pop_cols) +
    annotate("text", x = Inf, y = Inf, hjust = 1.1, vjust = 1.5,
             label = paste0("β = ", beta, "\n", "p ", pval), size = 6) +
    xlab(stringr::str_to_sentence(make_pretty_names(x))) +
    theme_classic() +
    theme(legend.position = "none",
          axis.title = element_text(size = 16),
          axis.text  = element_text(size = 14))

  if (residual) {
    plt + labs(y = "Partial heterozygosity")
  } else {
    plt + labs(y = "Observed heterozygosity")
  }

}

# Isolate significant vs all variables
sig_vars <- c("csi_past", "Q", "tmean_dif", "bio1", "pfs_surfac", "log_fri")
mod_vars <- c("NDVI", "gHM", "bio1", "csi_past", "glacier", "tmean_dif", "pfs_replac", "pfs_mixed", "pfs_surfac", "log_fri", "vdep", "Q")

# A: population plot
pop_label <- 
  st_drop_geometry(pop_df) %>% 
  group_by(cluster) %>%
  summarize(x = median(x), y = median(y)) %>%
  st_as_sf(coords = c("x", "y"), crs = 3310) 

pop_plot <-
  ggplot() +
  geom_sf(data = ca, col = NA) +
  geom_sf(data = pop_df, aes(fill = cluster), cex = 2, pch = 21, col = "NA")+
  #geom_point(data = pop_label, aes(x = x, y = y, fill = cluster), cex = 7, pch = 21, alpha = 0.7) +
  #geom_text(data = pop_label, aes(x = x, y = y, label = cluster), cex = 5) +
  labs(col = "Population") +
  scale_fill_manual(values = pop_cols)  +
  theme_void() +
  theme(
    legend.position = "none"
  )
# B: partial residuals 
plts_residual_y <- purrr::map(
  sig_vars, 
  # Use unscaled data for plotting
  ~plot_mod_pop(.x, "Ho", pop_df, model = final_model, residual = TRUE)
)

# B: Raw values
plts_raw_y <- purrr::map(
  sig_vars, 
  # Use unscaled data for plotting
  ~plot_mod_pop(.x, "Ho", pop_df, model = final_model, residual = FALSE)
)

# C: predicted vs observed 
pred_vs_obs <-
  # Use unscaled data for plotting
  ggplot(pop_df, aes(x = predict(final_model), y = Ho)) +
  geom_point(aes(col = cluster), shape = 16, size = 2) +
  geom_smooth(method = "lm", colour = "black") +
  scale_color_manual(values = pop_cols) +
  theme_classic() +
  labs(x = "Predicted heterozygosity", y = "Observed heterozygosity") +
  ggpubr::stat_cor(size = 6) +
  theme(legend.position = "none",
        axis.title = element_text(size = 16),
        axis.text  = element_text(size = 14))

pred_vs_obs_adj <- cowplot::plot_grid(pred_vs_obs, NULL, nrow = 1, rel_widths = c(0.9, 0.1))

plt2_residual <- cowplot::plot_grid(plotlist = plts_residual_y, ncol = 2)
plt2_raw <- cowplot::plot_grid(plotlist = plts_raw_y, ncol = 2)
plt1 <- cowplot::plot_grid(plotlist = list(pop_plot, pred_vs_obs_adj), ncol = 1,
                           labels = c("A", "C"), rel_heights = c(0.6, 0.4), label_size = 18)

png(here(plotpath, "figure_Ho_residual_pop.png"), height = 3.45*4*300*0.9, width = 4.1*4*300*0.9, res = 300)
cowplot::plot_grid(plt1, plt2_residual, ncol = 2,labels = c("", "B"),  rel_widths = c(0.4, 0.6), align = "h", label_size = 18)
dev.off()

png(here(plotpath, "figure_Ho_raw_pop.png"), height = 3.5*4*300*0.9, width = 4.1*4*300*0.9, res = 300)
cowplot::plot_grid(plt1, plt2_raw, ncol = 2,labels = c("", "B"),  rel_widths = c(0.4, 0.6), align = "h", label_size = 18)
dev.off()
```

```{r}
# Plot points on map colored by variable
sig_vars <- c("csi_past", "Q", "tmean_dif", "bio1", "pfs_surfac", "log_fri")
plot_points_map <- function(var, points_sf) {
  plt1 <- 
    ggplot() +
    geom_sf(data = ca_proj, fill = "gray90", colour = NA, size = 0.5) +
    geom_sf(data = points_sf, aes_string(col = var), size = 2) +
    scale_met_cont(points_sf[[var]], aesthetic = "colour", extend_scale = 0.1) +
    theme_void() +
    labs(title = stringr::str_to_sentence(make_pretty_names(var)), col = "") +
    theme(legend.position = "top", plot.title = element_text(hjust = 0.5))

  # Plot variable vs Ho
  plt2 <- plot_mod(var, "Ho", points_sf, model = final_model, color_var = "x", residual = FALSE) + theme(plot.title = element_blank())

  cowplot::plot_grid(plt2, plt1, nrow = 1, align = "hv")

}
png(here(plotpath, "significant_variable_maps.png"), width = 5*2, height = 5.5*length(sig_vars), res = 300, unit = "in")
plt_list <- map(sig_vars, ~(plot_points_map(.x, mod_sf_unique)))
cowplot::plot_grid(plotlist = plt_list, ncol = 1, align = "hv")
dev.off()
```

```{r}
# For each variable, build a model with all other variables and compare the residuals to the full model residuals
calc_var_resid <- function(x, mod_vars, df){
  other_vars <- setdiff(mod_vars, x)
  sub_f <- reformulate(other_vars, response = "Ho")
  f <- reformulate(mod_vars, response = "Ho")
  sub_model <- gls(sub_f, correlation = corExp(form = ~ x + y, nugget = FALSE), data = df)
  full_model <- gls(f, correlation = corExp(form = ~ x + y, nugget = FALSE), data = df)
  #sub_model <- lm(sub_f, data = df)
  #full_model <- lm(f, data = df)

  # Difference in fit
  sub_pred <- predict(sub_model)
  full_pred <- predict(full_model)
  pred_change <- full_pred - sub_pred

  return(tibble(
    pred_change = pred_change, 
    variable = x, 
    x = df$x, 
    y = df$y)
  )
}

sig_vars <- c("csi_past", "Q", "tmean_dif", "bio1", "pfs_surfac", "log_fri")

var_resid <- map(sig_vars, ~calc_var_resid(.x, mod_vars, mod_sf_unique), .progress = TRUE)

var_resid_df <- 
  var_resid %>% 
  bind_rows() %>% 
  st_as_sf(coords = c("x", "y"), crs = 3310) %>%
  group_by(variable) %>%
  mutate(pred_change_scaled = scale(pred_change)[,1]) 

pair_plot <- function(x, label = ""){
  y = "pred_change"
  change <- var_resid_df %>% filter(variable == x) %>% pull(y)
  df <- mod_sf_unique %>% mutate(change = change)
  plt_scatter <-
    ggplot(df, aes_string(x = x, y = "Ho")) +
    geom_point(aes(col = change), alpha = 0.7) +
    scico::scale_color_scico(palette = "roma", midpoint = 0, direction = -1) +
    geom_smooth(method = "lm", colour = "black") +
    theme_classic() +
    labs(x = make_pretty_names(x), y = "Observed Heterozygosity", col = y) +
    theme(legend.position = "none", plot.title = element_text(hjust = 0.5))

  plt_map <-
    ggplot(df) +
    geom_sf(data = ca_proj, fill = "white", colour = "gray", size = 0.5) +
    geom_sf(aes(col = change), pch = 16, size = 1.5, alpha = 0.9) +
    scico::scale_color_scico(palette = "roma", midpoint = 0, direction = -1) +
    theme_void() +
    labs(col = "Δ Predicted\nHeterozygosity") +
    theme(legend.position = "right", plot.title = element_text(hjust = 0.5))

  cowplot::plot_grid(plt_scatter, plt_map, nrow = 1, labels = c(label, ""), align = "h", rel_widths = c(0.45, 0.55))
}

png(here(plotpath, "variable_importance_map.png"), width = 3.5*2*2, height = 3*length(sig_vars)/2, units = "in", res = 300)
plts <- map(1:length(sig_vars), ~pair_plot(sig_vars[.x], label = LETTERS[.x]))
cowplot::plot_grid(plotlist = plts, ncol = 2)
dev.off()
```

Maps of significant variables
```{r}
plot_var <- function(x){
  ggplot() +
    geom_sf(data = ca_proj, fill = "gray90", colour = NA, size = 0.5) +
    geom_sf(data = mod_sf_unique, aes_string(fill = x), size = 2, shape = 21, color = "#00000087") +
    scale_met_cont(mod_sf_unique[[x]], aesthetic = "fill", extend_scale = 0.1) +
    theme_void() +
    labs(title = stringr::str_to_sentence(make_pretty_names(x)), fill = "") +
    theme(plot.title = element_text(hjust = 0.5))
}

sig_vars <- c("csi_past", "Q", "tmean_dif", "pfs_replac", "pfs_surfac", "log_fri")
pdf(here(plotpath, "significant_variable_maps_sf.pdf"), width = 4*3, height = 4*2)
plts <- map(sig_vars, ~plot_var(.x))
cowplot::plot_grid(plotlist = plts, nrow = 2)
dev.off()
```