
```{r}
library("here")
library("tidyverse")
library("ggpubr")
library("terra")
library("sf")
library("gridExtra")
library("rpaleoclim")
library("spdep")
library("spatialreg")
library("MuMIn")
library("cowplot")
library("performance")
library("gstat")
library("sp")
library("furrr")
library("MetBrewer")

source(here("general_functions.R"))
source(here("analysis", "genetic_diversity", "functions_genetic_diversity.R"))
source(here("analysis", "genetic_diversity", "functions_paleovars.R"))
ca <- get_ca()
ca_proj <- ca %>% st_transform(3310)
outpath <- here("analysis", "genetic_diversity", "outputs")
plotpath <- here("analysis", "genetic_diversity", "plots")

coords <- get_coords(sf = TRUE)
coords_proj <- coords %>% st_transform(crs = st_crs(ca_proj))
dem <- get_dem(r= TRUE)
```

```{r}
# Get genome-wide heterozygosity
het <- get_het("58-Sceloporus_chr.het") 

# Make sure you have the expected number of samples
stopifnot(nrow(het) == 259)

# Count number of sites used to calculate heterozygosity
mean(het$`N(NM)`, na.rm = TRUE) # 59955192

# Get adaptive heterozygosity
genes <- 
  format_het(here("analysis", "gea", "outputs", "genes.het"), callable_sites = 1) %>%
  mutate(Ho_gea = 1 - `O(HOM)`/`N(NM)`) %>%
  rename(SampleID = IID) %>%
  dplyr::select(SampleID, Ho_gea) 
```

# Hypothesis 1: Genetic diversity explained by paleoclimate stability

```{r}
# Get current climate data (NOTE: not using paleoclim data because they use V1.2 instead of V2.1)
bioclim <- rast(here("data", "env", "california_chelsa_bioclim_1981-2010_V.2.1.tif"))
cur <- bioclim[["CHELSA_bio1_1981-2010_V.2.1"]]
names(cur) <- "cur"

# Get LGM climate data (NOTE: not using paleoclim data because they use V1.2 instead of V2.1)
lgm <- rast(here("data", "env", "california_chelsa_pmip_ccsm4.tif"))
names(lgm) <- "lgm"

# Get paleoclimate data and convert from C/10 to C
lig <- get_paleovars("lig", cache = TRUE)$lig$bio_1/10 
paleovars <- get_paleovars(c("lh", "mh", "eh", "yds", "ba", "hs1", "lig", "mis19"))
paleobio1 <- map(paleovars, ~.x$bio_1) %>% rast()/10

# Calculate climate change over different time periods
cur_lgm <- cur - lgm
names(cur_lgm) <- "cur_lgm"
# Note: LIG is lower resolution than current and LGM data, so we need to resample to make them comparable
# 
lgm_lig <- resample(lgm, lig) - lig
names(lgm_lig) <- "lgm_lig"

# Stack CHELSA rasters
chelsabio1 <- list(cur, lgm)

# Resample the higher resolution data (CHELSA) to match the lower resolution (rpaleoclim)
# This is the most conservative approach to avoid over-interpreting fine-scale variation 
csi_stack <- 
  c(as.list(chelsabio1), as.list(paleobio1)) %>% 
  map(~resample(.x, paleobio1[[1]]), method = "bilinear") %>% 
  rast()

# Calculate the standard deviation over the stack of rasters
# Don't remove NAs, use imputation later (NA values are some points along the coast)
# If you remove NAs you will get some outliers because some time periods have data and others do not
csi_past <- app(csi_stack, sd, na.rm = FALSE)

# Scale CSI raster from 0 to 1 and subtract from 1 to get higher values = more stable
csi_past[] <- 1 - scales::rescale(csi_past[], to = c(0, 1), na.rm = TRUE)

# Make a data frame with the climate change values
paleoclim_mod_df <- 
  data.frame(
    SampleID = coords$SampleID,
    cur_lgm = terra::extract(cur_lgm, coords, ID = FALSE)[,1],
    lgm_lig = terra::extract(lgm_lig, coords, ID = FALSE)[,1],
    csi_past = terra::extract(csi_past, coords, ID = FALSE)[,1]
  )

# Check NA values
print(colSums(is.na(paleoclim_mod_df)))

# Some coastal points have NA values, so use a bilinear approach to get take the average of the four nearest raster cells
paleoclim_mod_df$csi_past <- bilinear_impute(paleoclim_mod_df$csi_past, csi_past, coords)
paleoclim_mod_df$cur_lgm <- bilinear_impute(paleoclim_mod_df$cur_lgm, cur_lgm, coords)
paleoclim_mod_df$lgm_lig <- bilinear_impute(paleoclim_mod_df$lgm_lig, lgm_lig, coords)

# Double check NA values
print(colSums(is.na(paleoclim_mod_df)))

# All paleovariables are highly correlated
cor(paleoclim_mod_df[, - 1], use = "pairwise.complete.obs")
```

```{r}
pdf(here(plotpath, "paleoclim_test.pdf"), width = 10, height = 10)
test_df <- left_join(paleoclim_mod_df, het) %>% left_join(coords) %>% st_as_sf()

ggplot(test_df) +
  geom_sf(data = ca) +
  geom_sf(aes(col = csi_past, geometry = geometry)) +
  theme_void() +
  scale_color_viridis_c(option = "turbo", end = 0.1, begin = 0.9, direction = -1) 

ggplot(test_df) +
  geom_point(aes(x = csi_past, y = Ho), inherit.aes = FALSE) +
  geom_smooth(aes(x = csi_past, y = Ho), method = "lm", col = "green") +
  stat_smooth(aes(x = csi_past, y = Ho), method = "gam", formula = y ~ s(x, k = 3), col = "blue") +
  stat_smooth(aes(x = csi_past, y = Ho), method = "lm", formula = y ~ poly(x, 2), col = "red", linetype = "dashed") +
  theme_classic()
dev.off()
```

```{r}
hindcast <- rast(here("data", "hindcast_sdm_sceloporus_occidentalis.tif"))
pdf(here(plotpath, "hindcast_sdm_sceloporus_occidentalis.pdf"), width = 10, height = 10)
plot(hindcast, col = viridis::turbo(100))
dev.off()

# NEED TO CHECK WITH NEW PALEOCLIM DATA FROM CHELSA for cur/lgm
hindcast_mean <- mean(hindcast, na.rm = FALSE) 
paleoclim_mod_df$sdm_stability <- terra::extract(hindcast, coords, ID = FALSE)[,1]
all(!is.na(paleoclim_mod_df$sdm_stability)) 

test <- left_join(paleoclim_mod_df, het)
cor(test$sdm_stability, test$Ho, use = "pairwise.complete.obs") # 0.42
cor(test$csi_past, test$Ho, use = "pairwise.complete.obs") # 0.36
```

# Hypothesis 2: Genetic diversity explained by admixture
```{r}
q <- read_csv(here("analysis", "admixture", "outputs", "Q8.csv"))
q_r <- rast(here("analysis", "admixture", "outputs", "max_admix.tif"))

q_mod_df <- 
  q %>%
  dplyr::select(SampleID, Q) %>%
  # Do 1- to make the values more intuitive
  mutate(Q = 1 - Q)
```

# Hypothesis 3: Genetic diversity explained by current environment 
```{r}
bioclim <- rast(here("data", "env", "california_chelsa_bioclim_1981-2010_V.2.1.tif"))
bio1 <- bioclim[["CHELSA_bio1_1981-2010_V.2.1"]]
ndvi <- rast(here("data", "env", "california_ndvi_mean_2000_2020.tif"))
ghm <- rast(here("data", "env", "california_global_human_modification_2016.tif"))

env_mod_df <- data.frame(
  SampleID = coords$SampleID,
  NDVI = terra::extract(ndvi, coords, ID = FALSE)[, "NDVI"],
  gHM = terra::extract(ghm, coords, ID = FALSE)[,"gHM"],
  bio1 = terra::extract(bio1, coords, ID = FALSE)[,"CHELSA_bio1_1981-2010_V.2.1"]
)

# Check NA values
print(colSums(is.na(env_mod_df)))

# Some coastal points have NA values for gHM and bio1, so use a bilinear approach to get take the average of the four nearest raster cells
# Note: This function only imputes missing values, not all values
env_mod_df$gHM <- bilinear_impute(env_mod_df$gHM, ghm, coords)
env_mod_df$bio1 <- bilinear_impute(env_mod_df$bio1, bio1, coords)

# Double check NA values
print(colSums(is.na(env_mod_df)))
```

# Hypothesis 4: Genetic diversity explained by glaciation
```{r}
lgm_glaciers <- st_read(here("data", "env", "Glaciotect_NA_lgm", "shp", "LGM_maxextent.shp"))

# Note: have to first intersect and then transform otherwise you get a node error
lgm_glaciers <- st_intersection(lgm_glaciers, st_transform(ca, crs = st_crs(lgm_glaciers)))
lgm_glaciers <- st_transform(lgm_glaciers, crs = st_crs(ca))

# Add a new column to coords indicating whether each point is inside the lgm shapefile
coords <- get_coords(sf = TRUE) %>% st_transform(crs = st_crs(lgm_glaciers))
glacier_mod_df <- 
  coords %>%
  mutate(glacier = as.numeric((lengths(st_within(coords, lgm_glaciers)) > 0))) %>%
  st_drop_geometry()
```

# Hypothesis 5: Genetic diversity explained by recent climate change
PRISM mean temperature data from historical data: https://prism.oregonstate.edu/historical/

```{r}
coords <- get_coords(sf = TRUE)
# Get prism data
get_prism <- function(year, output = "df") {
  # Create PRISM path
  prism_path <- here("data", "env", "prism", "PRISM_tmean_stable_4kmM3", year)

  # List annual file
  file <- list.files(prism_path, pattern = paste0(year, "_*.bil.bil$"), full.names = TRUE)

  # Load data
  raster <- terra::rast(file)

  if (output == "raster"){
    return(raster)
  } else {
    vals <- terra::extract(raster, st_transform(coords, st_crs(raster)), ID = FALSE)
    names(vals) <- "tmean"
    df <- data.frame(year = year, tmean = vals, SampleID = coords$SampleID)
    return(df)
  }
}

# Loop over the years of interest
historical_years <- 1895:1905
current_years <- 2010:2020
## Get dataframes
tmean_his <- map(historical_years, ~get_prism(year = .x), .progress =TRUE) %>% bind_rows()  
tmean_his <- tmean_his %>% group_by(SampleID) %>% summarise(tmean_his = mean(tmean, na.rm = TRUE))
tmean_cur <- map(current_years, ~get_prism(year = .x), .progress =TRUE) %>% bind_rows()  
tmean_cur <- tmean_cur %>% group_by(SampleID) %>% summarise(tmean_cur = mean(tmean, na.rm = TRUE))
# Get rasters
tmean_his_r <- map(historical_years, ~get_prism(year = .x, output = "raster"), .progress =TRUE) %>% rast() %>% mean()
tmean_cur_r <- map(current_years, ~get_prism(year = .x, output = "raster"), .progress =TRUE) %>% rast() %>% mean()
tmean_dif_r <- tmean_cur_r - tmean_his_r
tmean_dif_r <- crop(tmean_dif_r, ca)
tmean_dif_r <- mask(tmean_dif_r, st_transform(ca, st_crs(tmean_dif_r)))

# Calculate change
tmean_mod_df <- 
  left_join(tmean_his, tmean_cur) %>% 
  mutate(tmean_dif = tmean_cur - tmean_his) %>%
  right_join(coords) %>% 
  dplyr::select(SampleID, tmean_dif, tmean_cur)

# Check for NAs
tmean_mod_df %>% 
  summarise(across(everything(), ~sum(is.na(.))))

# Create and save raster
tmean_dif <- tmean_dif_r %>% project("epsg:3310")
names(tmean_dif) <- "tmean_dif"
writeRaster(tmean_dif, here("data", "env", "tmean_dif.tif"), overwrite = TRUE)
writeRaster(tmean_his_r, here("data", "env", "tmean_his.tif"), overwrite = TRUE)
writeRaster(tmean_cur_r, here("data", "env", "tmean_cur.tif"), overwrite = TRUE)
```

# Combine rasters
Combine into a raster stack:
```{r}
env_list <- 
  list(
    NDVI = ndvi,
    bio1 = bio1,
    gHM = ghm,
    tmean_dif = tmean_dif,
    cur_lgm = cur_lgm,
    lgm_lig = lgm_lig,
    csi_past = csi_past, 
    Q = q_r
  )

# Project all rasters to EPSG:3310 and mask with CA
env_list <- imap(env_list, ~{
  r <- terra::project(.x, "epsg:3310", method = "bilinear")
  r <- mask(r, ca_proj)
  names(r) <- .y
  return(r)
}, .progress = TRUE)


path <- here("data", "env", "model_rasters")
# Make model rasters dir
dir.create(path, showWarnings = FALSE)
iwalk(env_list, ~writeRaster(.x, here(path, paste0(.y, ".tif")), overwrite = TRUE))

# Snap all rasters to same grid and resolution
max_res <- map_dbl(env_list, ~res(.x)[1]) %>% which.max()
target_raster <- env_list[[max_res]] # coarsest resolution
resampled_list <- imap(env_list, ~{
  # Resample to match target
  terra::resample(.x, target_raster, method = "bilinear")
})

env_stack <- 
  resampled_list %>% 
  rast() 
```

# Create model data
```{r}
mod_vars <- c("NDVI", "gHM", "bio1", "Q", "csi_past", "glacier", "tmean_dif", "x", "y")

# Combine dataframes
mod_df <- 
  get_coords() %>% 
  left_join(het) %>%
  #left_join(genes) %>%
  left_join(env_mod_df) %>%
  left_join(paleoclim_mod_df) %>%
  left_join(q_mod_df) %>%
  left_join(glacier_mod_df) %>%
  left_join(tmean_mod_df) %>%
  dplyr::select(Ho, SampleID, all_of(mod_vars))

# Check values that have NA for Ho
mod_df %>% filter(is.na(Ho)) %>% select(SampleID)

# Drop Ho NA values
mod_df <-
  mod_df %>%
  drop_na(Ho) 

# Check sample number
stopifnot(nrow(mod_df) == 259)

# Confirm no NAs:
stopifnot(all(colSums(is.na(mod_df)) == 0))

# Scale the variables (except glacier)
mod_df_scaled <- 
  mod_df %>% 
  mutate(across(c(all_of(mod_vars), -glacier), scale))

# Convert to sf
mod_sf <- get_coords(sf = TRUE) %>% right_join(mod_df_scaled) %>% st_transform(3310)
```

```{r}
export_df <- 
  st_drop_geometry(mod_df) %>% 
  dplyr::select(SampleID, everything())
write_csv(export_df, here(outpath, "model_df.csv"))
```

## Check for correlation between variables
```{r}
# Check correlations
# Compute the correlation matrix
r <- cor(mod_df_scaled[, mod_vars], use = "pairwise.complete.obs")
# Set the diagonal and lower triangle to NA
r[lower.tri(r, diag = TRUE)] <- NA

# Find the pairs with correlations greater than 0.6
high_corr <- which(abs(r) > 0.6, arr.ind = TRUE)
# Print the pairs
high_corr_pairs <- data.frame(
  Var1 = rownames(r)[high_corr[, 1]],
  Var2 = rownames(r)[high_corr[, 2]],
  Correlation = r[high_corr]
)
high_corr_pairs
```

```{r}
#Make nice correlation plot
r <- stats::cor(mod_df_scaled[, setdiff(mod_vars, c("x", "y"))], use = "pairwise.complete.obs")
rownames(r) <- colnames(r) <- make_pretty_names(rownames(r))
r[lower.tri(r, diag = TRUE)] <- NA

library(corrplot)
png(here(plotpath, "correlation_plot.png"), width = 10*300, height = 10*300, units = "px", res = 300)
corrplot(r, method = "ellipse", type = "upper", tl.col = "black", tl.srt = 45, diag = FALSE, addCoef.col = "black", number.cex = 0.8)
dev.off()
```

# Run Ho models

## Compare spatial models
```{r}
# Uncorrelated set of model variables
mod_vars <- c("glacier_factor", "NDVI", "gHM", "bio1", "csi_past", "tmean_dif", "Q")

# Make glacier into a factor
mod_df_scaled$glacier_factor <- as.factor(mod_df_scaled$glacier)
mod_sf$glacier_factor <- as.factor(mod_sf$glacier)

# Plot relationships between each predictor and Ho
pdf(here(plotpath, "predictor_vs_Ho.pdf"), width = 3, height = 3)
walk(mod_vars[-1], ~print(gglm(.x, "Ho", mod_df_scaled)))
dev.off()

# GLS can't handle 0 distances, so take the average of the coordinates for comparing between different models
mod_sf_unique <- 
  mod_sf %>%
  group_by(x, y) %>%
  mutate(glacier_factor = as.numeric(as.character(glacier_factor))) %>%
  summarise_at(c(mod_vars, "Ho"), mean, na.rm = TRUE) 

# Run OLS and get residuals
f <- formula("Ho ~ NDVI + gHM +  csi_past + glacier_factor + Q + tmean_dif + bio1")
ols_model <- lm(f, data = mod_sf_unique)
summary(ols_model)
AIC(ols_model)
mod_sf_unique$ols_resids <- residuals(ols_model)
broom::tidy(ols_model) %>% filter(p.value < 0.05) %>% arrange(desc(abs(estimate)))

# Fit variograms to the residuals
ols_variogram <- variogram(ols_resids ~ 1, data = mod_sf_unique)
ols_vgm <- map(c("Exp", "Sph", "Gau"), ~fit.variogram(ols_variogram, model = vgm(.x)))
names(ols_vgm) <- c("Exp", "Sph", "Gau")

pdf(here(plotpath, "ols_variogram.pdf"))
plot(ols_variogram, ols_vgm[["Exp"]], main = "OLS Residuals Variogram (Exponential)")
plot(ols_variogram, ols_vgm[["Sph"]], main = "OLS Residuals Variogram (Spherical)")
plot(ols_variogram, ols_vgm[["Gau"]], main = "OLS Residuals Variogram (Gaussian)")
dev.off()

# Based on visual inspection, the exponential model seems to fit best
# Get the range of the variogram 
vgm_range <- ols_vgm[["Exp"]] %>% filter(model == "Exp") %>% pull(range)
print(vgm_range)

# Create a spatial weights matrix using the range from the variogram
nb <- dnearneigh(mod_sf_unique, d1 = 0, d2 = vgm_range)
listw <- nb2listw(nb, style = "W", zero.policy = TRUE)
pdf(here(plotpath, "spatial_weights.pdf"))
plot(nb, st_coordinates(mod_sf_unique), main = paste0("Neighborhoods (Range = ", round(vgm_range, 2), ")"))
dev.off()

# Fit an SEM model
sem_model <- errorsarlm(f, data = mod_sf_unique, listw = listw, zero.policy = TRUE)
summary(sem_model)
mod_sf_unique$sem_resids <- residuals(sem_model)

# Fit an SLM model
slm_model <- lagsarlm(f, data = mod_sf_unique, listw = listw, zero.policy = TRUE)
summary(slm_model)
mod_sf_unique$slm_resids <- residuals(slm_model)
# https://crd230.github.io/lab8.html

# Fit a GLS model
library(nlme)
# Note: nugget=FALSE assumes residuals at the same location are perfectly correlated, which aligns with expectations for heterozygosity, since you don't expect to have much microscale variation in heterozygosity.
# nugget=TRUE allows unstructured microscale variance (random noise at zero distance), which can absorb spatial signal and cause predictor effects to appear insignificant. 
gls_model <- gls(f, correlation = corExp(form = ~ x + y, nugget = FALSE), data = mod_sf_unique)
summary(gls_model)
# Use type = "normalized" to get residuals decorrelated and standardized by the fitted 
# spatial covariance matrix; this removes modeled spatial structure and makes them 
# suitable for Moran's I. If you don't do this Moran's I will appear higher than it actually is.
mod_sf_unique$gls_resids <- residuals(gls_model, type = "normalized")

# Test with nugget = TRUE
nugget_model <- gls(f, correlation = corExp(form = ~ x + y, nugget = TRUE), data = mod_sf_unique)
mod_sf_unique$nugget_resids <- residuals(nugget_model, type = "normalized")

# Plot residuals
plot_resids <- function(resids){
  # Remove max residual because it is hard to see spatial patterns with it
  ggplot(mod_sf_unique[mod_sf_unique[[resids]] != max(mod_sf_unique[[resids]]), ]) +
    geom_sf(data = ca_proj, col = NA) +
    geom_sf(aes(col = .data[[resids]])) +
    theme_void() +
    scale_color_viridis_c(option = "turbo", end = 0.1, begin = 0.9, direction = -1) +
    labs(title = paste("Residuals:", resids))
}

pdf(here(plotpath, "all_model_residuals.pdf"), width = 12, height = 8)
plt_list <- list(
  plot_resids("ols_resids"),
  plot_resids("sem_resids"),
  plot_resids("slm_resids"),
  plot_resids("gls_resids"),
  plot_resids("nugget_resids")
)
plot_grid(plotlist = plt_list, nrow = 2)
dev.off()

# Compare AIC
# You will get warning about the models not having the same number of observations, but the number of observations is the same for all models, so this is not an issue 
all_equal_nobs <- length(unique(c(
  nobs(sem_model), nobs(slm_model), nobs(ols_model), nobs(gls_model), nobs(nugget_model)
))) == 1  
stopifnot(all_equal_nobs)

AIC(ols_model, sem_model, slm_model, gls_model, nugget_model) # SEM model is the best fit


# Compare significant variables
# Compare significant variables (p < 0.05) for each model using map
models <- list(OLS = ols_model, SEM = sem_model, SLM = slm_model, GLS = gls_model, GLS_nugget = nugget_model)
library(broom.mixed)
sig_vars_by_model <- map(models, ~ broom::tidy(.x) %>%
  filter(!term %in% c("lambda", "rho", "(Intercept)"), p.value < 0.05) %>%
  pull(term))
print(sig_vars_by_model)

# GLS adjusts standard errors for residual spatial autocorrelation, making it appropriate when 
# the goal is inference on predictors (e.g., testing relationships between environmental variables 
# and genome-wide heterozygosity). It models a smooth spatial covariance structure to correct 
# inflated Type I error rates without changing the interpretation of fixed effects. In contrast, 
# SEM explicitly models latent spatial processes (e.g., unmeasured environmental or historical 
# factors) and absorbs their influence into a spatial error term. While this can further reduce 
# residual spatial autocorrelation, it also partitions variance away from the predictors, potentially 
# leading to smaller coefficients and reduced significance. Since we are not trying to model 
# unobserved spatial processes themselves, SEM is overkill and may obscure the effects of interest.
```

Model comparison table:
```{r}
model_list <- list(
  OLS = ols_model,
  SEM = sem_model,
  SLM = slm_model,
  GLS = gls_model,
  GLS_nugget = nugget_model
)

aic_tbl <- AIC(ols_model, sem_model, slm_model, gls_model, nugget_model) |>
  as.data.frame() |>
  rownames_to_column("raw_name") |>
  mutate(Model = recode(raw_name,
                        "ols_model" = "OLS",
                        "sem_model" = "SEM",
                        "slm_model" = "SLM",
                        "gls_model" = "GLS",
                        "nugget_model" = "GLS_nugget")) |>
  select(Model, df, AIC)

morans_I <- c(
  OLS        = unname(moran.test(mod_sf_unique$ols_resids,    listw, zero.policy = TRUE)$estimate[["Moran I statistic"]]),
  SEM        = unname(moran.test(mod_sf_unique$sem_resids,    listw, zero.policy = TRUE)$estimate[["Moran I statistic"]]),
  SLM        = unname(moran.test(mod_sf_unique$slm_resids,    listw, zero.policy = TRUE)$estimate[["Moran I statistic"]]),
  GLS        = unname(moran.test(mod_sf_unique$gls_resids,    listw, zero.policy = TRUE)$estimate[["Moran I statistic"]]),
  GLS_nugget = unname(moran.test(mod_sf_unique$nugget_resids, listw, zero.policy = TRUE)$estimate[["Moran I statistic"]])
)

sig_tbl <- 
  enframe(sig_vars_by_model, name = "Model", value = "Significant Variables") %>%
  mutate(`Significant Variables` = map_chr(`Significant Variables`,
                                           ~ if (length(.x)) paste(.x, collapse = ", ") else NA_character_))

model_levels <- c("OLS", "SEM", "SLM", "GLS", "GLS_nugget")

model_comparison_tbl <- 
  aic_tbl %>%
  mutate(`Moran's I` = unname(morans_I[Model])) %>%
  left_join(sig_tbl, by = "Model") %>%
  mutate(Model = factor(Model, levels = model_levels)) %>%
  arrange(Model)

model_comparison_tbl
```
```{r}
# Final model variables
final_vars <- c("NDVI", "gHM", "bio1", "csi_past", "glacier_factor", "tmean_dif", "Q")

# Create formula
f <- formula(paste0("Ho ~", paste(final_vars, collapse = " + ")))

# Fit model
best_model <-  gls(f, correlation = corExp(form = ~ x + y, nugget = FALSE), data = mod_sf_unique)

# Results
mod_result <- broom::tidy(best_model)
mod_result

# Significant variables
mod_result %>% filter(p.value < 0.05) %>% arrange(desc(abs(estimate))) 
sig_vars <- mod_result %>% filter(p.value < 0.05, !term %in% c("(Intercept)", "lambda")) %>% pull(term)
print(sig_vars)

# You can't get a traditional R2 for spatial models, but you can get an alternative R2
cor(predict(best_model), mod_sf_unique$Ho)^2 # 0.42
cor(predict(best_model), mod_sf_unique$Ho) # 0.65
```

## Variance partioning analysis
```{r}
all_vars <- c("NDVI", "gHM", "bio1", "csi_past", "glacier", "tmean_dif", "Q")
predictor_sets <- mod_df_scaled %>% dplyr::select(all_of(all_vars)) %>% as.list()
response <- mod_df_scaled$Ho
varpart_result <- lm_variance_partition(predictor_sets, response, scale_vars =TRUE)
```

## Plots 
```{r}
# Model variables for plotting (glacier is not treated as a factor for plotting)
mod_vars <- c("NDVI", "gHM", "bio1", "csi_past", "glacier", "tmean_dif", "Q")
mod <- best_model

# Confirm treating glacier as a factor vs numeric makes no difference because it is binary
lm(Ho ~ glacier, data = mod_df_scaled) %>% summary()
lm(Ho ~ glacier_factor, data = mod_df_scaled) %>% summary()

# Created unscaled plot
mod_sf_unique_unscaled <- 
  mod_df %>%
  group_by(x, y) %>%
  summarise_at(c(mod_vars, "Ho"), mean, na.rm = TRUE)

# Plot predicted vs observed
plt1 <- 
  map(mod_vars, ~{
    ggplot(mod_sf_unique_unscaled) +
      geom_point(aes(x = predict(mod), y = Ho, col = .data[[.x]])) +
      #geom_abline(intercept = 0, slope = 1) +
      geom_smooth(aes(x = predict(mod), y = Ho), method = "lm") +
      theme_classic() +
      labs(x = "Predicted", y = "Observed") +
      scale_color_viridis_c(option = "mako", end = 0.1, begin = 0.9)
  })

# Plot values across california
plt2 <-
  map(mod_vars, ~{
    ggplot(mod_sf_unique_unscaled) +
      geom_sf(data = ca) +
      geom_point(aes(x = x, y = y, col = .data[[.x]])) +
      theme_void() +
      scale_color_viridis_c(option = "mako", end = 0.1, begin = 0.9)
  })

# Plot Ho vs variable
plt3 <- map(mod_vars, ~{
  ggplot(mod_sf_unique_unscaled, aes(x = .data[[.x]], y = Ho)) +
    geom_point(aes(col = .data[[.x]])) +
    geom_smooth(method = "lm") +
    theme_classic() +
    labs(x = gsub("\n", " ", .x), y = "Ho") +
    scale_color_viridis_c(option = "mako", end = 0.1, begin = 0.9) +
    theme(legend.position = "none") +
    stat_cor(label.y = min(mod_df[["Ho"]], na.rm = TRUE))
})

# Plot Ho
plt4 <-
  map(1:length(mod_vars), ~{
    ggplot(mod_sf_unique_unscaled) +
      geom_sf(data = ca) +
      geom_point(aes(x = x, y = y, col = Ho)) +
      theme_void() +
      scale_color_viridis_c(option ="mako", end = 0.1, begin = 0.9, direction = -1)
})

pdf(here(plotpath, "ho.pdf"), width = 4.5, height = 4.5)
ggplot(mod_sf_unique_unscaled) +
  geom_sf(data = ca) +
  geom_point(aes(x = x, y = y, col = Ho)) +
  theme_void() +
  scale_color_viridis_c(option ="mako") +
  theme(
    plot.margin = margin(t = 1, r = 1, b = 1, l = 1, unit = "mm"),
    legend.position = c(0.85, 0.6),  # (x, y) in npc coords (0-1)
    legend.justification = c(1, 0)
  )
dev.off()

# use cowplot to combine plots
pdf(here(plotpath, "linear_model_plot_Ho.pdf"), width = 10, height = 25)
ordered_plts <- pmap(list(plt3, plt1, plt2), ggpubr::ggarrange, common.legend = TRUE, legend = "right", nrow = 1)
do.call(cowplot::plot_grid, c(ordered_plts, list(nrow = length(plt1))))
ordered_plts <- pmap(list(plt3, plt1, plt4), ggpubr::ggarrange, nrow = 1)
do.call(cowplot::plot_grid, c(ordered_plts, list(nrow = length(plt1))))
dev.off()
```

```{r}
scale_met_cont <- function(values, aesthetic = c("fill", "colour"), extend_scale = 0.1) {
  aesthetic <- match.arg(aesthetic)
  r_min <- min(values, na.rm = TRUE)
  r_max <- max(values, na.rm = TRUE)
  rng   <- r_max - r_min
  lims  <- c(r_min - rng*extend_scale, r_max + rng*extend_scale)

  pal <- rev(MetBrewer::met.brewer("Hiroshige", type = "continuous"))
  if (aesthetic == "fill") {
    ggplot2::scale_fill_gradientn(colors = pal, limits = lims, oob = scales::squish)
  } else {
    ggplot2::scale_colour_gradientn(colors = pal, limits = lims, oob = scales::squish)
  }
}


plot_mod <- function(x, y, df, type = "residual", model = NULL, 
                     mod_vars = c("NDVI","gHM","bio1","csi_past","glacier","tmean_dif","Q"),
                     color_var = c("x","y")) {

  if (type == "partial") {
    f_rhs <- paste(setdiff(mod_vars, x), collapse = " + ")
    f_py  <- stats::as.formula(paste(y, "~", f_rhs))
    f_px  <- stats::as.formula(paste(x, "~", f_rhs))
    df[[x]] <- residuals(lm(f_px, data = df))
    df[[y]] <- residuals(lm(f_py, data = df))
  } else if (type == "residual") {
    # partial residuals for y
    stopifnot(!is.null(model))
    # note; for this residuals should not be normalized
    df[[y]] <- residuals(model) + coef(model)[[x]] * as.numeric(scale(df[[x]]))
  }

  col_var <- if (color_var == "x") x else y

  ggplot(df, aes(x = .data[[x]], y = .data[[y]], colour = .data[[col_var]])) +
    geom_point() +
    geom_smooth(method = "lm", colour = "black") +
    scale_met_cont(df[[col_var]], aesthetic = "colour", extend_scale = 0.1) +
    ggpubr::stat_cor() +
    labs(x = stringr::str_to_sentence(make_pretty_names(x)), 
      y = stringr::str_to_sentence(paste0(type, " heterozygosity"))) +
    theme_classic() +
    theme(legend.position = "none",
          axis.title = element_text(size = 14),
          axis.text  = element_text(size = 10))

}

ca_proj <- get_ca() %>% sf::st_transform(3310)
range_map <- get_range()
wingen_r <- 
  terra::rast(here(outpath, "Ho_kriged_masked.tif")) %>% 
  mask(range_map) %>% 
  as.data.frame(xy = TRUE, na.rm = TRUE) 
names(wingen_r)[3] <- "layer"  # ensure 'layer' is the raster value column

sig_vars <- c("csi_past", "tmean_dif", "bio1", "Q")
mod_vars <- c("NDVI", "gHM", "bio1", "csi_past", "glacier", "tmean_dif", "Q")

# A: map
ho_plot <- ggplot(wingen_r) +
  geom_sf(data = ca_proj, fill = "gray90", colour = NA) +
  geom_tile(aes(x = x, y = y, fill = layer), colour = NA) +
  scale_met_cont(wingen_r$layer, aesthetic = "fill", extend_scale = 0.1) +
  labs(fill = "Heterozygosity") +
  theme_void() +
  geom_sf(data = mod_sf_unique, shape = 21, aes(fill = Ho), colour = "#00000087", size = 1.3) +
  theme(
    plot.margin = margin(1, 1, 1, 1, unit = "mm"),
    legend.position = c(0.9, 0.6),
    legend.justification = c(1, 0),
    legend.key.height = unit(0.5, "cm"),
    legend.key.width  = unit(0.5, "cm"),
    legend.title = element_text(size = 12),
    legend.text  = element_text(size = 10)
  )


# B: partial residuals 
plts_residual_y <- purrr::map(
  sig_vars, 
  ~plot_mod(.x, "Ho", mod_sf_unique_unscaled, type = "residual", model = best_model, color_var = "y")
)


# C: predicted vs observed 
pred_vs_obs <-
  ggplot(mod_sf_unique_unscaled, aes(x = predict(best_model), y = Ho)) +
  geom_point(aes(colour = Ho), shape = 1) +
  geom_smooth(method = "lm", colour = "black") +
  scale_met_cont(c(mod_sf_unique_unscaled$Ho, wingen_r$layer), aesthetic = "colour", extend_scale = 0.1) +
  theme_classic() +
  labs(x = "Predicted heterozygosity", y = "Observed heterozygosity") +
  ggpubr::stat_cor() +
  theme(legend.position = "none",
        axis.title = element_text(size = 13),
        axis.text  = element_text(size = 10))

pred_vs_obs_adj <- cowplot::plot_grid(pred_vs_obs, NULL, nrow = 1, rel_widths = c(0.9, 0.1))

plt1 <- cowplot::plot_grid(plotlist = plts_residual_y, nrow = 2)
plt2 <- cowplot::plot_grid(plotlist = list(ho_plot, pred_vs_obs_adj), ncol = 1,
                           labels = c("A", "C"), rel_heights = c(0.6, 0.4))
pdf(here(plotpath, "figure_Ho.pdf"), height = 8, width = 12)
cowplot::plot_grid(plt2, plt1, ncol = 2, rel_widths = c(0.33, 0.67), labels = c("", "B"))
dev.off()

png(here(plotpath, "figure_Ho.png"), height = 8*300, width = 12*300, res = 300)
cowplot::plot_grid(plt2, plt1, ncol = 2, rel_widths = c(0.33, 0.67), labels = c("", "B"))
dev.off()
```

# Make map of predictions
```{r}
pred_vars <- sig_vars

# Get coefficients (excluding the spatial error term) 
betas <- coef(best_model)

# Just use significant variables
betas <- betas[c("(Intercept)", pred_vars)]

# Reorder env_stack layers to match betas and scale
pred_stack <- scale(env_stack[[pred_vars]])

# Apply the fixed effects projection
pred <- 
  map(pred_vars, ~{
    # Get the raster for the variable
    r <- env_stack[[.x]]
    # Apply the fixed effects projection
    r * betas[.x]
  }) %>% 
  rast() %>%
  sum() %>%
  # Add the intercept
  `+`(betas["(Intercept)"])

names(pred) <- "Ho_pred"
range_map <- get_range()
pred <- mask(pred, range_map)

hiroshige_cols <- MetBrewer::met.brewer("Hiroshige", type = "continuous", n = 100, direction = -1)

ggrast <- function(x) {
  df <- as.data.frame(x, xy = TRUE)
  val_col <- names(x)[1]  # assume single-layer raster for prediction

  ggplot(df) +
    geom_sf(data = ca_proj, col = NA) +
    geom_raster( aes(x = x, y = y, fill = .data[[val_col]])) + 
    theme_void()
}


png(here(plotpath, "Ho_pred.png"), width = 1500, height = 1500, res = 300)
ggrast(pred) + 
  labs(fill = "Predicted\nHeterozygosity") +
  scale_fill_gradientn(colors = hiroshige_cols, na.value = "transparent") 
dev.off()
```

```{r}
model_Ho <- best_model
```

# Run Adaptive Ho models

## Check for spatial autocorrelation
```{r}
# Run OLS and get residuals
f <- as.formula(paste("Ho_gea ~", paste(mod_vars, collapse = " + ")))
ols_model <- lm(f, data = mod_df_scaled)
mod_sf$ols_resids <- residuals(ols_model)

# Fit variograms to the residuals
ols_variogram <- variogram(ols_resids ~ 1, data = mod_sf)
ols_vgm <- map(c("Exp", "Sph", "Gau"), ~fit.variogram(ols_variogram, model = vgm(.x)))
names(ols_vgm) <- c("Exp", "Sph", "Gau")

pdf(here(plotpath, "ols_variogram_gea.pdf"))
plot(ols_variogram, ols_vgm[["Exp"]], main = "OLS Residuals Variogram (Exponential)")
plot(ols_variogram, ols_vgm[["Sph"]], main = "OLS Residuals Variogram (Spherical)")
plot(ols_variogram, ols_vgm[["Gau"]], main = "OLS Residuals Variogram (Gaussian)")
dev.off()

# Based on visual inspection, the exponential model seems to fit best
# Get the range of the variogram (3 times the range parameter)
vgm_range <- ols_vgm[["Exp"]] %>% filter(model == "Exp") %>% pull(range)
print(vgm_range)

# Create a spatial weights matrix using the range from the variogram
nb <- dnearneigh(mod_sf, d1 = 0, d2 = vgm_range)
listw <- nb2listw(nb, style = "W", zero.policy = TRUE)

# Fit an spatial model
sp_model <- lagsarlm(f, data = mod_sf, listw = listw, zero.policy = TRUE)
summary(sp_model)
mod_sf$sp_resids <- residuals(sp_model)

# Calculate Moran's I before and after accounting for spatial autocorrelation
moran.test(mod_sf$ols_resids, listw)$estimate
moran.test(mod_sf$sp_resids, listw)$estimate
```

```{r}
full_model <- lagsarlm(f, data = mod_sf, listw = listw, zero.policy = TRUE)

# Variables are NDVI + glacier + Q
broom::tidy(full_model) %>% filter(p.value < 0.05)

best_model <- full_model
summary(best_model)

# You can't get a traditional R2 for spatial models, but you can get an alternative R2
r2_alternative <- cor(sem_model$fitted.values, sem_model$y)^2
r2_alternative # 43%
```

```{r}
# DELETE Later
pdf(here(plotpath, "partial_plots.pdf"), width = 4, height = 4)
ggpartial("csi_past", "Ho", mod_vars, mod_df, col = "csi_past")
ggpartial("tmean_dif", "Ho", mod_vars, mod_df, col = "tmean_dif")
ggpartial("bio1", "Ho", mod_vars, mod_df, col = "bio1")
dev.off()
```

## Variance partioning analysis
```{r}
all_vars <- c("NDVI", "gHM", "bio1", "cur_lgm", "lgm_lig", "csi_past", "glacier", "tmean_dif", "Q")
predictor_sets <- mod_df_scaled %>% select(all_of(all_vars)) %>% as.list()
response <- mod_df_scaled$Ho_gea
varpart_result_gea <- lm_variance_partition(predictor_sets, response, scale_vars =TRUE)
```

## Plots and tables
```{r}
# Turn into a nice gt table
library(gt)

best_model_table <- 
  broom::tidy(best_model) %>%
  data.frame() %>%
  mutate(term = map_chr(term, ~make_pretty_names(.x))) %>%
  filter(term != "(Intercept)") %>%
  gt() %>%
  tab_header(title = "Full model") %>%
  gtExtras::gt_hulk_col_numeric(estimate, trim = TRUE) %>%
  fmt_scientific(
    columns = c("estimate", "std.error", "statistic", "p.value"),
    decimals = 2
  ) %>%
  # bold significant rows
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = c("p.value"),
      rows = p.value < 0.05
    )
  )

# export to html
gt::gtsave(best_model_table, filename = here(plotpath, "best_model_table_Ho_gea.html"))
```

```{r}
# Make pretty plots 
mod <- best_model

# Plot predicted vs observed
plt1 <- 
  map(mod_vars, ~{
    ggplot(mod_df) +
      geom_point(aes(x = predict(mod), y = Ho, col = .data[[.x]])) +
      #geom_abline(intercept = 0, slope = 1) +
      geom_smooth(aes(x = predict(mod), y = Ho), method = "lm") +
      theme_classic() +
      labs(x = "Predicted", y = "Observed") +
      scale_color_viridis_c(option = "mako", end = 0.1, begin = 0.9)
  })

# Plot values across california
plt2 <-
  map(mod_vars, ~{
    ggplot(mod_df) +
      geom_sf(data = ca) +
      geom_point(aes(x = x, y = y, col = .data[[.x]])) +
      theme_void() +
      scale_color_viridis_c(option = "mako", end = 0.1, begin = 0.9)
  })

# Plot Ho vs variable
plt3 <- map(mod_vars, ~{
  ggplot(mod_df, aes(x = .data[[.x]], y = Ho)) +
    geom_point(aes(col = .data[[.x]])) +
    geom_smooth(method = "lm") +
    theme_classic() +
    labs(x = gsub("\n", " ", .x), y = "Ho") +
    scale_color_viridis_c(option = "mako", end = 0.1, begin = 0.9) +
    theme(legend.position = "none") +
    stat_cor(label.y = min(gg_df[["Ho"]], na.rm = TRUE))
})

# Plot Ho
plt4 <-
  map(1:length(mod_vars), ~{
    ggplot(mod_df) +
      geom_sf(data = ca) +
      geom_point(aes(x = x, y = y, col = Ho)) +
      theme_void() +
      scale_color_viridis_c(option ="mako", end = 0.1, begin = 0.9, direction = -1)
})

pdf(here(plotpath, "ho.pdf"), width = 4.5, height = 4.5)
ggplot(mod_df) +
  geom_sf(data = ca) +
  geom_point(aes(x = x, y = y, col = Ho)) +
  theme_void() +
  scale_color_viridis_c(option ="mako") +
  theme(plot.margin = margin(t = 1, r = 1, b = 1, l = 1, unit = "mm"))
dev.off()

# use cowplot to combine plots
pdf(here(plotpath, "linear_model_plot_Ho_gea.pdf"), width = 10, height = 25)
ordered_plts <- pmap(list(plt3, plt1, plt2), ggpubr::ggarrange, common.legend = TRUE, legend = "right", nrow = 1)
do.call(cowplot::plot_grid, c(ordered_plts, list(nrow = length(plt1))))
ordered_plts <- pmap(list(plt3, plt1, plt4), ggpubr::ggarrange, nrow = 1)
do.call(cowplot::plot_grid, c(ordered_plts, list(nrow = length(plt1))))
dev.off()
```

```{r}
model_Ho_gea <- best_model
```

# Future climate change
```{r}
future_ho <- 
  # NOTE: CSI was inverted (1 - CSI) before
  df_future %>%
  left_join(mod_df) %>%
  # REMOVE CHANNEL ISLAND
  filter(SampleID != "Scelocci_CCGPMC_MW01-3-14") %>%
  pivot_longer(c("Ho", "Ho_gea"), names_to = "metric", values_to = "Ho") %>%
  mutate(metric = case_when(
    metric == "Ho" ~ "Genome-wide diversity",
    metric == "Ho_gea" ~ "Adaptive diversity"
  )) %>%
  mutate(metric = factor(metric, levels = c("Genome-wide diversity", "Adaptive diversity"))) 

plt <- 
  ggplot(future_ho, aes(x = csi, y = Ho, col = csi)) +
  geom_point() +
  ggh4x::facet_grid2(metric ~ model_name, scales = "free", switch = "y", axes = "all") +
  theme_classic() +
  labs(x = "Climate stability", y = "") +
  geom_smooth(method = "lm", se = FALSE, col = "black") +
  stat_cor(label.y = min(future_ho$Ho, na.rm = TRUE)) +
  scale_color_viridis_c(option = "mako") +
  theme(
    strip.background = element_blank(),
    strip.text.x = element_text(hjust = 0, size = 14),
    strip.placement = "outside",
    strip.text.y = element_text(size = 14),
    axis.title = element_text(size = 14)
  ) 

pdf(here(plotpath, "future_plot.pdf"), width = 13, height = 6)
plt
dev.off()
```

# Figures
```{r}
# Variance partitioning plot
gg_df <- 
  varpart_result$summary_table %>%
  select(-r2) %>%
  pivot_longer(cols = c("adj_r2", "unique_contribution"), names_to = "metric", values_to = "value") %>%
  mutate(metric = recode(metric, adj_r2 = "Total", unique_contribution = "Unique")) %>%
  mutate(value = as.numeric(gsub("%", "", value))/100) %>%
  mutate(predictor_set = make_pretty_names(predictor_set)) %>%
  mutate(Ho = "Ho")

gg_df_gea <- 
  varpart_result_gea$summary_table %>%
  select(-r2) %>%
  pivot_longer(cols = c("adj_r2", "unique_contribution"), names_to = "metric", values_to = "value") %>%
  mutate(metric = recode(metric, adj_r2 = "Total", unique_contribution = "Unique")) %>%
  mutate(value = as.numeric(gsub("%", "", value))/100) %>%
  mutate(predictor_set = make_pretty_names(predictor_set)) %>%
  mutate(Ho = "Ho_gea")

gg_df <- 
  bind_rows(gg_df, gg_df_gea) %>% 
  mutate(Ho = 
    case_when(
      Ho == "Ho" ~ "Genome-wide",
      Ho == "Ho_gea" ~ "Adaptive"
    )
  ) %>%
  mutate(Ho = factor(Ho, levels = c("Genome-wide", "Adaptive")))

ylgnbl <- RColorBrewer::brewer.pal(9, "YlGnBu")
ylgnbl_lite <- (ylgnbl[-c(1, 7:9)])

varpart_plt <- 
  ggplot(gg_df, aes(x = metric, y = predictor_set, fill = value)) +
  geom_tile(color = "white", lwd = 1) +
  geom_text(aes(label = scales::percent(value, accuracy = 1)), color = "black") +
  scale_fill_gradientn(colors = ylgnbl_lite) +
  theme_minimal() +
  labs(x = "", y = "", fill = "Variance\nExplained") +
  scale_x_discrete(position = "top") +
  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        strip.text.x = element_text(size = 14, face = "bold"),
        strip.placement = "outside") +
  facet_wrap(~Ho, nrow = 1)

# Correlation plot
mod_vars <- c("NDVI", "gHM", "bio1", "cur_lgm", "lgm_lig", "csi_past", "glacier", "tmean_dif", "Q")
cormat <- cor(mod_df[, mod_vars])

cor_df <- 
  cormat %>%
  as.data.frame() %>%
  rownames_to_column(var = "var1") %>%
  pivot_longer(cols = -var1, names_to = "var2", values_to = "cor") %>%
  filter(!is.na(cor)) %>%
  mutate(var1_pretty = make_pretty_names(var1), var2_pretty = make_pretty_names(var2)) %>%
  mutate(cor = case_when(var1 == var2 ~ NA, TRUE ~ cor))

cor_plt <- 
  ggplot(cor_df, aes(x = var1_pretty, y = var2_pretty, fill = cor)) +
  geom_tile(color = "white", lwd = 1) +
  geom_text(aes(label = round(cor, 2)), color = "black") +
  scale_fill_gradient2(
    low = "cyan3",
    mid = "white",
    high = "orange",
    midpoint = 0,
    limits = c(-1, 1)
  ) +
  theme_minimal() +
  labs(x = "", y = "", fill = "Correlation") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 12),
        axis.text.y = element_text(size = 12))

# Plot variables correlated with Ho that are not in the final model
names(mod_vars) <- mod_vars
confounded_vars <- map(mod_vars, ~cor.test(mod_df$Ho, mod_df[[.x]])$p.value) < 0.05
confounded_vars <- names(confounded_vars[confounded_vars])
confounded_vars <- confounded_vars[!(confounded_vars %in% c("csi_past", "Q"))]

missig_plt <-
  mod_df %>%
  pivot_longer(all_of(confounded_vars)) %>%
  mutate(name = make_pretty_names(name)) %>%
  ggplot(aes(x = value, y = Ho)) +
  geom_point(aes(col = cur_lgm)) +
  geom_smooth(method = "lm") +
  theme_classic() +
  labs(x = "", y = "Ho", col = "Paleoclimate change\n(LGM - LIG)") +
  scale_color_viridis_c(option = "mako") +
  facet_wrap(~name, scales = "free", strip.position = "bottom", nrow = 2) +
  theme(
    strip.placement = "outside", 
    strip.background = element_blank(), 
    strip.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  ) +
  stat_cor(label.y = min(mod_df$Ho, na.rm = TRUE)) 

names(mod_vars) <- mod_vars
confounded_vars <- map(mod_vars, ~cor.test(mod_df$Ho_gea, mod_df[[.x]])$p.value) < 0.05
confounded_vars <- names(confounded_vars[confounded_vars])
confounded_vars <- confounded_vars[!(confounded_vars %in% names(model_Ho_gea$coefficients))]
length(confounded_vars) # No confounded vars for Ho_gea
# missig_plt_gea <-
#   mod_df %>%
#   pivot_longer(all_of(confounded_vars)) %>%
#   mutate(name = make_pretty_names(name)) %>%
#   ggplot(aes(x = value, y = Ho)) +
#   geom_point(aes(col = lgm_lig)) +
#   geom_smooth(method = "lm") +
#   theme_classic() +
#   labs(x = "", y = "Ho", col = "Paleoclimate change\n(LGM - LIG)") +
#   scale_color_viridis_c(option = "mako") +
#   facet_wrap(~name, scales = "free", strip.position = "bottom", nrow = 1) +
#   theme(strip.placement = "outside", strip.background = element_blank()) +
#   stat_cor(label.y = min(mod_df$Ho, na.rm = TRUE)) 

##combo <- plot_grid(cor_plt, varpart_plt, nrow = 1, rel_widths = c(5, 3), labels = c("A", "B"))
##plot_grid(combo, missig_plt, nrow = 2, rel_heights = c(0.6, 0.35), labels = c("", "D"))


combo <- plot_grid(cor_plt, varpart_plt, nrow = 1, rel_widths = c(5, 4), labels = c("A", "B"))
pdf(here(plotpath, "confounding.pdf"), width = 17, height = 13)
plot_grid(combo, missig_plt, nrow = 2, rel_heights = c(0.4, 0.5), labels = c("", "D"))
dev.off()
```

```{r}
ca <- get_ca()

colpal = "plasma"
sig_vars <- c("csi_past", "Q")
plot_map <- function(var){
  ggplot(st_as_sf(mod_df)) +
    geom_sf(data = ca) +
    geom_sf(aes(fill = .data[[var]]), pch = 21, col = "black", cex = 2) +
    theme_void() +
    scale_fill_viridis_c(option = colpal) +
    labs(fill = gsub(" ", "\n", make_pretty_names(var)))
}

predicted_Ho <- data.frame(SampleID = mod_sf$SampleID, predicted = model_Ho$fitted, metric = "Ho")
predicted_Ho_gea <- data.frame(SampleID = mod_sf$SampleID, predicted = model_Ho_gea$fitted, metric = "Ho_gea")
mod_df$predicted <- model_Ho$fitted
mod_df$predicted_gea <- model_Ho_gea$fitted

gg_df <- 
  mod_df %>%
  pivot_longer(c("Ho", "Ho_gea"), names_to = "metric", values_to = "Ho") %>%
  left_join(bind_rows(predicted_Ho, predicted_Ho_gea))

mod_plt <-
  ggplot(mod_df, aes(x = predicted, y = Ho)) +
  geom_point(aes(fill = Ho), pch = 21, col = "black", alpha = 0.8, cex = 2) +
  geom_smooth(method = "lm", color = "black") +
  theme_classic() +
  labs(x = "Predicted genome-wide heterozygosity", y = "Observed genome-wide heterozygosity") +
  scale_fill_viridis_c(option = colpal) +
  stat_cor(label.y = min(mod_df$Ho, na.rm = TRUE)) +
  theme(legend.position = "none") 

mod_plt_gea <-
  ggplot(mod_df, aes(x = predicted_gea, y = Ho_gea)) +
  geom_point(aes(fill = Ho), pch = 21, col = "black", alpha = 0.8, cex = 2) +
  geom_smooth(method = "lm", color = "black") +
  theme_classic() +
  labs(x = "Predicted adaptive heterozygosity", y = "Observed adaptive heterozygosity") +
  scale_fill_viridis_c(option = colpal) +
  stat_cor(label.y = min(mod_df$Ho_gea, na.rm = TRUE)) +
  theme(legend.position = "none") 

map_plts <- map(c("Ho", sig_vars), plot_map)
map_plts_gea <- map(c("Ho_gea", sig_vars), plot_map)
#lm_plts <- map(sig_vars, ~gglm(.x, "Ho", mod_df))
partial_plts <- map(sig_vars, ~ggpartial(.x, "Ho", sig_vars, mod_df, col = "Ho", cex = 2) + scale_fill_viridis_c(option = colpal))
partial_plts_gea <- map(sig_vars, ~ggpartial(.x, "Ho_gea", sig_vars, mod_df, col = "Ho_gea", cex = 2) + scale_fill_viridis_c(option = colpal))

#plts <- c(map_plts, list(mod_plt), lm_plts, list(blank = ggplot() + theme_void()), partial_plts)
plts <- do.call(plot_grid, c(map_plts[1], partial_plts, list(mod_plt), nrow = 1))
plts_gea <- do.call(plot_grid, c(map_plts_gea[1], partial_plts_gea, list(mod_plt_gea), nrow = 1))

#labels <- LETTERS[1:(length(plts) - 1)]
#labels[which(names(plts) == "blank")] <- ""

pdf(here(plotpath, "main_figure.pdf"), width = 17, height = 9) 
plot_grid(plts, plts_gea, nrow = 2)
dev.off()
```

# Committee meeting

```{r}
pdf(here(plotpath, "committee_meeting.pdf"), width = 15, height = 3.5)
sig_vars <- c("csi_past", "tmean_dif", "bio1", "Q")
partial_plts <- map(sig_vars, ~ggpartial(.x, "Ho", sig_vars, mod_df, col = "Ho", cex = 2) + scale_fill_viridis_c(option = colpal))
do.call(plot_grid, c(partial_plts, list(nrow = 1)))
dev.off()

pdf(here(plotpath, "committee_meeting.pdf"), width = 5, height = 4)
plot_map("Ho") + scale_fill_viridis_c(option = "mako")
dev.off()
```
# Test of what variables are no longer significant after accounting for spatial autocorrelation

In the lm() bio1/tmean_dif is significant but in the spatial model it is not.
```{r}
# Compare significant variables from ols and sem models
summary(ols_model) %>% broom::tidy() %>% filter(term != "(Intercept)", term != "lambda") %>% filter(p.value < 0.05)
summary(sem_model) %>% broom::tidy() %>% filter(term != "(Intercept)", term != "lambda") %>% filter(p.value < 0.05)

# All variables except NDVI have M > 0.5, lgm_lig has the highest M (0.7)
names(mod_vars) <- mod_vars
map(mod_vars, ~moran.test(pull(mod_sf, .x), listw)$estimate) %>% 
  bind_rows(.id = "var") %>% 
  arrange(desc(`Moran I statistic`))
```

# Biodiversity comparison
```{r}
richness <- rast(here("data", "biodiversity_mapping", "Reptiles_total_richness.tif"))
df <- 
  left_join(get_coords(sf = TRUE), get_het()) %>%
  mutate(`Reptile species richness` = extract(richness, get_coords(sf = TRUE), ID = FALSE)[,1]) %>%
  mutate(`Reptile species richness` = as.numeric(as.character(`Reptile species richness`))) %>%
  drop_na(Ho) %>%
  # Drop Island individual 
  # REMOVE CHANNEL ISLAND SAMPLE
  filter(SampleID != "Scelocci_CCGPMC_MW01-3-14")

pdf(here(plotpath, "richness_vs_ho.pdf"), width = 13, height =4)
plt1 <- 
  gglm("Reptile species richness", "Ho", df) + 
  scale_color_viridis_c(option = "plasma")
plt2 <-
  ggplot(df) +
  geom_sf(data = ca) +
  geom_sf(aes(col = Ho)) +
  scale_color_viridis_c(option = "plasma") +
  theme_void()
plt3 <-
  ggplot(df) +
  geom_sf(data = ca) +
  geom_sf(aes(col = `Reptile species richness`)) +
  labs(col = "Reptile\nspecies\nrichness") +
  scale_color_viridis_c(option = "plasma") +
  theme_void()
plot_grid(plt1, plt2, plt3, nrow = 1)
dev.off()
```