

# Libraries

```{r}
library("here")
library("tidyverse")
library("ggpubr")
library("terra")
library("sf")
library("gridExtra")
library("rpaleoclim")
library("MuMIn")
library("cowplot")
library("performance")
library("gstat")
library("sp")
library("furrr")
library("MetBrewer")
library("nlme")
library("broom.mixed")
library("emmeans")
library("interactions")
```

# Paths for outputs
```{r}
outpath <- here("analysis", "genetic_diversity", "outputs")
plotpath <- here("analysis", "genetic_diversity", "plots")
```

# Custom functions
```{r}
source(here("general_functions.R"))
source(here("analysis", "genetic_diversity", "functions_genetic_diversity.R"))
source(here("analysis", "genetic_diversity", "functions_paleovars.R"))
```

# California polygon for plotting
```{r}
ca <- get_ca()
ca_proj <- ca %>% st_transform(3310)
```

# Coordinates
```{r}
coords <- get_coords(sf = TRUE)
coords_proj <- coords %>% st_transform(crs = st_crs(ca_proj))

# Calculate average sampling distance
dist_mat <- st_distance(coords_proj)
avg_dist <- median(dist_mat[lower.tri(dist_mat)])
# Convert to km
avg_dist_km <- round(as.numeric(avg_dist) / 1000, 0)
avg_dist_km # 333
```

# Heterozygosity data
```{r}
# Get genome-wide heterozygosity
het <- get_het("58-Sceloporus_chr.het") 

# Make sure you have the expected number of samples
stopifnot(nrow(het) == 259)

# Count number of sites used to calculate heterozygosity
mean(het$`N(NM)`, na.rm = TRUE) # 59955192
```

# Hypothesis 1: Genetic diversity explained by paleoclimate stability

```{r}
# Get current climate data (NOTE: not using paleoclim data because they use CHELSA V1.2 instead of V2.1)
bioclim <- rast(here("data", "env", "california_chelsa_bioclim_1981-2010_V.2.1.tif"))
cur <- bioclim[["CHELSA_bio1_1981-2010_V.2.1"]]
names(cur) <- "cur"

# Get LGM climate data (NOTE: not using paleoclim data because they use CHELSA V1.2 instead of V2.1)
lgm <- rast(here("data", "env", "california_chelsa_pmip_ccsm4.tif"))
names(lgm) <- "lgm"

# Get paleoclimate data (BIO1) from rpaleoclim and convert from C/10 to C
# Using time periods back until about ~ 1 mya (when Sceloporus diverged in S. CA)
#
# code	period	bp	source
# cur	Current (1979 – 2013)		CHELSA
# lh	Late Holocene: Meghalayan	4.2-0.3 ka	Fordham  et al. 2017
# mh	Mid Holocene: Northgrippian	8.326-4.2 ka	Fordham  et al. 2017
# eh	Early Holocene: Greenlandian	11.7-8.326 ka	Fordham  et al. 2017
# yds	Pleistocene: Younger Dryas Stadial	12.9-11.7 ka	Fordham  et al. 2017
# ba	Pleistocene: Bølling-Allerød	14.7-12.9 ka	Fordham  et al. 2017
# hs1	Pleistocene: Heinrich Stadial 1	17.0-14.7 ka	Fordham  et al. 2017
# lgm	Pleistocene: Last Glacial Maximum	ca. 21 ka	CHELSA
# lig	Pleistocene: Last Interglacial	ca. 130 ka	Otto-Bliesner et al. 2006
# mis19	Pleistocene: MIS19	ca. 787 ka	Brown et al. 2018
paleovars <- get_paleovars(c("lh", "mh", "eh", "yds", "ba", "hs1", "lig", "mis19"))
paleobio1 <- map(paleovars, ~.x$bio_1) %>% rast()/10

# Stack CHELSA rasters
chelsabio1 <- list(cur, lgm)

# Resample the higher resolution data (CHELSA) to match the lower resolution (rpaleoclim)
# This is the most conservative approach to avoid over-interpreting fine-scale variation 
csi_stack <- 
  c(as.list(chelsabio1), as.list(paleobio1)) %>% 
  map(~resample(.x, paleobio1[[1]]), method = "bilinear") %>% 
  rast()

# Calculate climate stability based on the standard deviation over the stack of rasters
# Don't remove NAs, use imputation later (NA values are at some points along the coast)
# If you remove NAs you will get some outliers because some time periods have data and others do not
csi_past <- app(csi_stack, sd, na.rm = FALSE)

# Scale CSI raster from 0 to 1 and subtract from 1 to get higher values = higher stability
csi_past[] <- 1 - scales::rescale(csi_past[], to = c(0, 1), na.rm = TRUE)

# Make a data frame with the climate stability values and holocene values for each sample
paleoclim_mod_df <- 
  data.frame(
    SampleID = coords$SampleID,
    csi_past = terra::extract(csi_past, coords, ID = FALSE)[,1],
    lh = terra::extract(paleobio1[["lh"]], coords, ID = FALSE)[,1],
    mh = terra::extract(paleobio1[["mh"]], coords, ID = FALSE)[,1],
    eh = terra::extract(paleobio1[["eh"]], coords, ID = FALSE)[,1],
    lgm = terra::extract(lgm, coords, ID = FALSE)[,1],
    lig = terra::extract(paleobio1[["lig"]], coords, ID = FALSE)[,1]
  )

# Check NA values (should be <10)
print(colSums(is.na(paleoclim_mod_df)))

# Some coastal points have NA values, so use a bilinear approach to get take the average of the four nearest raster cells
paleoclim_mod_df$csi_past <- bilinear_impute(paleoclim_mod_df$csi_past, csi_past, coords)
paleoclim_mod_df$mh <- bilinear_impute(paleoclim_mod_df$mh, csi_past, coords)
paleoclim_mod_df$eh <- bilinear_impute(paleoclim_mod_df$eh, csi_past, coords)
paleoclim_mod_df$lgm <- bilinear_impute(paleoclim_mod_df$lgm, csi_past, coords)

# Re-check NA values (should be 0)
print(colSums(is.na(paleoclim_mod_df)))
```

# Hypothesis 3: Genetic diversity explained by current environment 
```{r}
# Read in contemporary environmental layers downloaded using the scripts in data_processing/env
bio1 <- bioclim[["CHELSA_bio1_1981-2010_V.2.1"]] # mean annual temperature
ghm <- rast(here("data", "env", "california_global_human_modification_2016.tif")) # global human modification index
# Global Aridity Index values need to be multiplied by 0.0001 to retrieve the values in the correct units (see documentation: https://doi.org/10.6084/m9.figshare.7504448)
# Note: higher AI valules = more humid, lower AI = more arid
ai <- rast(here("data", "env", "Global-AI_ET0__annual_v3_1", "ai_v31_yr.tif")) %>% crop(ca) * 0.0001

# Combine into df
env_mod_df <- data.frame(
  SampleID = coords$SampleID,
  gHM = terra::extract(ghm, coords, ID = FALSE)[,"gHM"],
  bio1 = terra::extract(bio1, coords, ID = FALSE)[,"CHELSA_bio1_1981-2010_V.2.1"],
  ai = terra::extract(ai, coords, ID = FALSE)[,"awi_pm_sr_yr"]
)

# Check NA values
print(colSums(is.na(env_mod_df)))

# Some coastal points have NA values for gHM and bio1, so use a bilinear approach to get take the average of the four nearest raster cells
# Note: This function only imputes missing values, not all values
env_mod_df$gHM <- bilinear_impute(env_mod_df$gHM, ghm, coords)
env_mod_df$bio1 <- bilinear_impute(env_mod_df$bio1, bio1, coords)

# Double check NA values (should be 0)
print(colSums(is.na(env_mod_df)))
```

# Hypothesis 4: Genetic diversity explained by glaciation
```{r}
# Read in LGM polygon provided by Michelle Koo (MVZ)
lgm_glaciers <- st_read(here("data", "env", "Glaciotect_NA_lgm", "shp", "LGM_maxextent.shp"))

# Transform to same CRS as coords
# Note: have to first intersect and then transform otherwise you get a node error
lgm_glaciers <- st_intersection(lgm_glaciers, st_transform(ca, crs = st_crs(lgm_glaciers)))
lgm_glaciers <- st_transform(lgm_glaciers, crs = st_crs(ca))

# Add a new column to coords indicating whether each point is inside the LGM glacial extent
glacier_coords <- coords %>% st_transform(st_crs(lgm_glaciers))
glacier_mod_df <- 
  coords %>%
  mutate(glacier = as.numeric(lengths(st_within(glacier_coords, lgm_glaciers)) > 0)) %>%
  mutate(glacier = factor(glacier, levels = c(0, 1), labels = c("OutsideGlacier", "InsideGlacier"))) %>%
  st_drop_geometry()
```

# Hypothesis 5: Genetic diversity explained by recent climate change
PRISM mean temperature data from historical data: https://prism.oregonstate.edu/historical/
```{r}
# Get historical prism data (https://prism.oregonstate.edu/historical/)
# Processed by scripts in data_processing/env
get_prism <- function(year, output = "df") {
  # Create PRISM path
  prism_path <- here("data", "env", "prism", "PRISM_tmean_stable_4kmM3", year)

  # List annual file
  file <- list.files(prism_path, pattern = paste0(year, "_*.bil.bil$"), full.names = TRUE)

  # Load data
  raster <- terra::rast(file)

  if (output == "raster"){
    return(raster)
  } else {
    vals <- terra::extract(raster, st_transform(coords, st_crs(raster)), ID = FALSE)
    names(vals) <- "tmean"
    df <- data.frame(year = year, tmean = vals, SampleID = coords$SampleID)
    return(df)
  }
}

# Loop over the years of interest
historical_years <- 1895:1905 # Historical decade
current_years <- 2010:2020 # Contemporary decade

# Make dataframe of average temperature by year for each sample
tmean_his <- 
  map(
    historical_years, 
    ~get_prism(year = .x), 
    .progress =TRUE
  ) %>% 
  bind_rows()  

tmean_cur <- 
  map(
    current_years, 
    ~get_prism(year = .x), 
    .progress =TRUE
  ) %>% 
  bind_rows()  

# Calculate mean temperature for each sample over the decade
tmean_his <- 
  tmean_his %>% 
  group_by(SampleID) %>% 
  summarise(tmean_his = mean(tmean, na.rm = TRUE))

tmean_cur <- 
  tmean_cur %>% 
  group_by(SampleID) %>% 
  summarise(tmean_cur = mean(tmean, na.rm = TRUE))

# Get rasters per year and average over the decade
tmean_his_r <-
   map(historical_years, ~get_prism(year = .x, output = "raster"), .progress =TRUE) %>% 
   rast() %>% 
   mean()

tmean_cur_r <- 
  map(current_years, ~get_prism(year = .x, output = "raster"), .progress =TRUE) %>% 
  rast() %>% 
  mean()

# Calculate difference between current and historical
tmean_dif_r <- tmean_cur_r - tmean_his_r

# Crop to california and mask
tmean_dif <- crop(tmean_dif_r, ca)
tmean_dif <- mask(tmean_dif, st_transform(ca, st_crs(tmean_dif_r)))

# Calculate change
tmean_mod_df <- 
  left_join(tmean_his, tmean_cur) %>% 
  mutate(tmean_dif = tmean_cur - tmean_his) %>%
  right_join(coords) %>% 
  dplyr::select(SampleID, tmean_dif, tmean_cur)

# Check for NAs
tmean_mod_df %>% 
  summarise(across(everything(), ~sum(is.na(.))))
```

# Hypothesis 6: Genetic diversity explained by current vegetation type 
```{r}
# Read in pre-processed LANDFIRE data produced by analysis/genetic_diversity/p2_processing_fire_data.R
landfire <- read_csv(here(outpath, "landfire_data.csv"))

evt_mod_df <- 
  landfire %>%
  select(SampleID, evt) 
```


# Hypothesis 7: Genetic diversity explained by historical fire regime
```{r}
fire_mod_df <- 
  landfire %>%
  select(SampleID, fire_frq, fire_recent)
```

# Create model data
```{r}
# Model variables
mod_vars <- c("gHM", "bio1", "ai", "csi_past", "glacier", "tmean_dif", "fire_frq", "fire_recent", "evt", "x", "y")

# Combine dataframes
mod_df <- 
  get_coords() %>% 
  left_join(het) %>%
  left_join(env_mod_df) %>%
  left_join(paleoclim_mod_df) %>%
  left_join(glacier_mod_df) %>%
  left_join(tmean_mod_df) %>%
  left_join(fire_mod_df) %>%
  left_join(evt_mod_df) %>%
  dplyr::select(Ho, SampleID, all_of(mod_vars))

# Check sample number
# Note: het has 259 samples, but two of them are potentially flipped ("Scelocci_CAS213197", "Scelocci_CAS214858"), so we remove the from the model (they are not included in coords)
stopifnot(nrow(mod_df) == 257)

# Confirm no NAs:
stopifnot(all(colSums(is.na(mod_df)) == 0))

# Convert to sf
mod_sf <- 
  get_coords(sf = TRUE) %>% 
  right_join(mod_df) %>% 
  st_transform(3310) %>%
  dplyr::select(-x, -y) %>%
  bind_cols(st_coordinates(.)) %>%
  dplyr::rename(x = X, y = Y)
```

```{r}
# Export model df
export_df <- 
  st_drop_geometry(mod_df) %>% 
  dplyr::select(SampleID, everything())
write_csv(export_df, here(outpath, "model_df.csv"))
#mod_df <- read_csv(here(outpath, "model_df.csv"))
#mod_sf <- coords_proj %>% left_join(mod_df)
```

## Check for correlation between variables
```{r}
# Check correlations
# Compute the correlation matrix
numeric_mod_vars <-
  mod_df %>% 
  dplyr::select(all_of(mod_vars), -x, -y) %>%
  dplyr::select(where(is.numeric)) %>% 
  colnames()

r <- cor(mod_df[, numeric_mod_vars], use = "pairwise.complete.obs")
# Set the diagonal and lower triangle to NA
r[lower.tri(r, diag = TRUE)] <- NA
r

# Find the pairs with correlations greater than 0.7
high_corr <- which(abs(r) > 0.6, arr.ind = TRUE)
# Print the pairs
high_corr_pairs <- data.frame(
  Var1 = rownames(r)[high_corr[, 1]],
  Var2 = rownames(r)[high_corr[, 2]],
  Correlation = r[high_corr]
) %>%
arrange(desc(abs(Correlation)))

# No highly correlated pairs
high_corr_pairs
```

```{r}
#-----------------------------
# Helpers: detect variable type
#-----------------------------
is_cat <- function(x) is.factor(x) || is.character(x) || is.logical(x)
is_cont <- function(x) is.numeric(x) || is.integer(x)

#-----------------------------
# Build the three diagnostics
# A) cont vs cont correlation (Pearson)
# B) cont ~ cat adj R^2 (one-way ANOVA via lm)
# C) cat vs cat association (Cramér's V)
#-----------------------------
make_diag_tables <- function(mod_df, cor_vars, min_n = 10){
  cor_vars <- intersect(cor_vars, names(mod_df))

  cont_vars <- cor_vars[map_lgl(mod_df[cor_vars], is_cont)]
  cat_vars  <- cor_vars[map_lgl(mod_df[cor_vars], is_cat)]

  # ---- A: continuous vs continuous (correlation)
  A_pairs <- combn(cont_vars, 2, simplify = FALSE)
  A_df <- map_dfr(A_pairs, \(p){
    v1 <- p[1]; v2 <- p[2]
    x <- mod_df[[v1]]; y <- mod_df[[v2]]
    ok <- is.finite(x) & is.finite(y)
    n <- sum(ok)
    if (n < min_n) {
      return(tibble(var1=v1, var2=v2, n=n, metric="r", value=NA_real_, p_value=NA_real_))
    }
    ct <- suppressWarnings(cor.test(x[ok], y[ok], method = "pearson"))
    tibble(var1=v1, var2=v2, n=n, metric="r", value=unname(ct$estimate), p_value=ct$p.value)
  })

  # ---- B: continuous ~ categorical (adj R^2)
  B_df <- expand_grid(cont = cont_vars, cat = cat_vars) %>%
    mutate(res = pmap(list(cont, cat), \(cont, cat){
      df <- tibble(y = mod_df[[cont]], g = as.factor(mod_df[[cat]]))
      ok <- is.finite(df$y) & !is.na(df$g)
      df <- df[ok, , drop = FALSE]

      if (nrow(df) < min_n || nlevels(df$g) < 2) {
        return(tibble(n=nrow(df), metric="adj_rsq", value=NA_real_, p_value=NA_real_))
      }

      m <- lm(y ~ g, data = df)
      sm <- summary(m)
      # overall p-value for group effect
      p <- tryCatch(anova(m)[["Pr(>F)"]][1], error = function(e) NA_real_)
      tibble(n=stats::nobs(m), metric="rsq", value=sm$r.squared, p_value=p)
    })) %>%
    unnest(res) %>%
    transmute(var1 = cont, var2 = cat, n, metric, value, p_value)

  # ---- C: categorical vs categorical (Cramér's V)
  C_pairs <- combn(cat_vars, 2, simplify = FALSE)
  C_df <- map_dfr(C_pairs, \(p){
    v1 <- p[1]; v2 <- p[2]
    x <- as.factor(mod_df[[v1]])
    y <- as.factor(mod_df[[v2]])
    ok <- !is.na(x) & !is.na(y)
    x <- droplevels(x[ok]); y <- droplevels(y[ok])
    n <- length(x)

    if (n < min_n || nlevels(x) < 2 || nlevels(y) < 2) {
      return(tibble(var1=v1, var2=v2, n=n, metric="cramers_v", value=NA_real_, p_value=NA_real_))
    }

    tbl <- table(x, y)
    chi <- suppressWarnings(chisq.test(tbl, correct = FALSE))
    # Cramér's V
    r <- nrow(tbl); k <- ncol(tbl)
    V <- sqrt(as.numeric(chi$statistic) / (sum(tbl) * (min(r - 1, k - 1))))
    tibble(var1=v1, var2=v2, n=sum(tbl), metric="cramers_v", value=V, p_value=chi$p.value)
  })

    # --- Make A symmetric + add diagonal NA so empties are on diagonal
  A_levels <- cont_vars
  A_df <- A_df %>%
    bind_rows(transmute(A_df, var1 = var2, var2 = var1)) %>%   # mirror
    bind_rows(tibble(var1 = A_levels, var2 = A_levels,          # diagonal
                     n = NA_integer_, metric = "r",
                     value = NA_real_, p_value = NA_real_)) %>%
    mutate(
      var1 = factor(var1, levels = A_levels),
      var2 = factor(var2, levels = A_levels)
    )

  # --- Make C symmetric + add diagonal NA so empties are on diagonal
  C_levels <- cat_vars
  C_df <- C_df %>%
    bind_rows(transmute(C_df, var1 = var2, var2 = var1)) %>%   # mirror
    bind_rows(tibble(var1 = C_levels, var2 = C_levels,          # diagonal
                     n = NA_integer_, metric = "cramers_v",
                     value = NA_real_, p_value = NA_real_)) %>%
    mutate(
      var1 = factor(var1, levels = C_levels),
      var2 = factor(var2, levels = C_levels)
    )

  list(A_df = A_df, B_df = B_df, C_df = C_df,
       cont_vars = cont_vars, cat_vars = cat_vars)
}

#-----------------------------
# Plotting functions
#-----------------------------
plot_A_cont_cont <- function(df){
  ggplot(df, aes(var1, var2, fill = value)) +
    geom_tile(color = "white", linewidth = 0.3) +
    geom_text(aes(label = ifelse(is.na(value), "", sprintf("%.2f", value))), size = 4) +
    scale_fill_gradientn(
      name = "Pearson's r",
      colours = c(viridis::mako(3, direction = 1, begin = 0.4, end = 0.9), "white", viridis::rocket(3, direction = -1, begin = 0.4, end = 0.9)),
      limits = c(-1, 1),
      oob = scales::squish,
      na.value = "grey95"
    ) +
    scale_x_discrete(labels = make_pretty_names) +
    scale_y_discrete(labels = make_pretty_names) +
    coord_fixed() +
    labs(x = NULL, y = NULL) +
    theme_minimal(base_size = 11) +
    theme(axis.text.x = element_text(angle = 30, hjust = 1),
          panel.grid = element_blank())
}

plot_C_cat_cat <- function(df){
  ggplot(df, aes(var1, var2, fill = value)) +
    geom_tile(color = "white", linewidth = 0.3) +
    geom_text(aes(label = ifelse(is.na(value), "", sprintf("%.2f", value))), size = 4) +
    scale_fill_viridis_c(name = "Cramér's V", option = "rocket", limits = c(0, 1), begin = 0.2, direction = -1, na.value = "grey90") +
    scale_x_discrete(labels = make_pretty_names) +
    scale_y_discrete(labels = make_pretty_names) +
    coord_fixed() +
    labs(x = NULL, y = NULL) +
    theme_minimal(base_size = 11) +
    theme(axis.text.x = element_text(angle = 30, hjust = 1),
          panel.grid = element_blank())
}


plot_B_cont_cat <- function(df){
  ggplot(df, aes(make_pretty_names(var2), make_pretty_names(var1), fill = value)) +
    geom_tile(color = "white", linewidth = 0.3) +
    geom_text(aes(label = ifelse(is.na(value), "", sprintf("%.2f", value))), size = 4) +
    scale_fill_viridis_c(name = "R²", option = "rocket", limits = c(-0.001, 1), begin = 0.2, direction = -1, na.value = "grey90") +
    labs(x = NULL, y = NULL) +
    theme_minimal(base_size = 11) +
    theme(axis.text.x = element_text(angle = 30, hjust = 1),
          panel.grid = element_blank())
} 

#-----------------------------
# RUN (your variables)
#-----------------------------
cor_vars <- c("csi_past", "bio1", "ai", "tmean_dif", "fire_frq", "fire_recent", "gHM", "glacier", "evt")

out <- make_diag_tables(mod_df, cor_vars, min_n = 10)

p_cont_cont <- plot_A_cont_cont(out$A_df) + theme(axis.text = element_text(size = 12), text = element_text(size = 14))
p_cont_cat <- plot_B_cont_cat(out$B_df) + theme(axis.text = element_text(size = 12), text = element_text(size = 14))
p_cat_cat <- plot_C_cat_cat(out$C_df) + theme(axis.text = element_text(size = 12), text = element_text(size = 14)) 

group1 <- plot_grid(p_cont_cont, p_cat_cat, ncol = 1, labels = c("A", "B"), align = "hv", label_size = 20)
png(here(plotpath, "variable_correlation_diagnostics.png"), width = 7, height = 12, units = "in", res = 300)
cowplot::plot_grid(p_cont_cont, p_cat_cat, p_cont_cat, ncol = 1, labels = c("A", "B", "C"), label_size = 18, align ="hv")
dev.off()

# No apparent issues with multicollinearity
```

## Transform and scale data for modeling
```{r}
# Model formula variables
mod_vars <- c("glacier", "gHM", "bio1", "ai", "csi_past", "tmean_dif", "fire_frq", "fire_recent", "evt")

# GLS can't handle 0 distances, so take the average of the coordinates (mode for veg type)
mode_val <- function(x) {
  ux <- na.omit(x)
  ux[which.max(tabulate(match(ux, unique(ux))))]
}
cat_mod_vars <- mod_df %>% 
  select(all_of(mod_vars)) %>%
  select(where(~is.character(.x) | is.factor(.x))) %>%
  colnames()
numeric_mod_vars <- setdiff(mod_vars, cat_mod_vars)

# Count number of duplicate coordinates
dup_coords <-
  mod_sf %>%
  st_coordinates() %>%
  as.data.frame() %>%
  group_by(X, Y) %>%
  tally() %>%
  filter(n > 1)
print(paste("Number of duplicate coordinates:", nrow(dup_coords)))

mod_sf_unique <- 
  mod_sf %>%
  group_by(x, y) %>%
  summarise(
    across(all_of(c(numeric_mod_vars, "Ho")), ~ mean(.x, na.rm = TRUE)),
    across(all_of(cat_mod_vars), ~ mode_val(.x)),
    .groups = "drop"
  )

# Set factor levels for fire_frq/fire_frg
mod_sf_unique$fire_frq <- factor(mod_sf_unique$fire_frq, levels = c("No Fire/Low Frequency", "Intermediate Frequency", "High Frequency"))
mod_sf_unique$fire_recent <- factor(mod_sf_unique$fire_recent, levels = c("Unburned", "Burned"))

# Set reference levels for groupveg, evt, and fire variables
mod_sf_unique$evt <- relevel(as.factor(mod_sf_unique$evt), ref = "Shrubland")
mod_sf_unique$fire_frq <- relevel(as.factor(mod_sf_unique$fire_frq), ref = "No Fire/Low Frequency") # Note: setting high frequency as reference because this variable is used to test for an interaction with recent fire (hypothesis: recent fire suppression will negatively effect Ho in high frequency fire regimes, but not in other cases). We do an emmeans post-hoc test to compare all levels, so the choice of reference here is not critical.

# Scale data
mod_sf_unique_scaled <-
  mod_sf_unique %>%
  mutate(
    across(
      all_of(numeric_mod_vars),
      ~ as.numeric(scale(.x))
    )
  )

# Confirm no NA
stopifnot(all(colSums(is.na(mod_sf_unique_scaled)) == 0))

# Save mod_sf_unique and mod_sf_unique_scaled objects
saveRDS(mod_sf_unique, here(outpath, "mod_sf_unique.rds"))
saveRDS(mod_sf_unique_scaled, here(outpath, "mod_sf_unique_scaled.rds"))
```

```{r}
# After visual inspection of the data I observed a non-linear relationship between csi_past and Ho
# So I am testing a quadratic term for csi_past
model_linear <- lm(Ho ~ glacier + gHM + bio1 + ai + csi_past + tmean_dif + evt + fire_frq + fire_recent, data = mod_sf_unique)
model_poly <- lm(Ho ~ glacier + gHM + bio1 + ai+ poly(csi_past, 2, raw = TRUE) + tmean_dif + evt + fire_frq + fire_recent, data = mod_sf_unique)

AIC(model_linear, model_poly) # The polynomial model has a lower AIC
anova(model_linear, model_poly) # The polynomial model is significantly better
```

```{r}
# I am interested in whether the effect of recent fire is dependent on the historical fire regime, so I am testing an interaction between fire_recent and fire_frq.
model_nointeract <- lm(Ho ~ glacier + gHM + ai + poly(csi_past, 2, raw = TRUE) + bio1 + tmean_dif + evt + fire_frq * fire_recent, data = mod_sf_unique)
model_interact <- lm(Ho ~ glacier + gHM + ai+ poly(csi_past, 2, raw = TRUE) + bio1 * tmean_dif + evt + fire_frq * fire_recent, data = mod_sf_unique)

AIC(model_nointeract, model_interact) # The interaction model has a lower AIC
anova(model_nointeract, model_interact) # The interaction model is significantly better
```

```{r}
# I am also interested in whether the effect of recent climate change is dependent on the temperature experienced by the population, so I am testing an interaction between bio1 and tmean_dif.
model_nointeract <- lm(Ho ~ glacier + gHM + ai + poly(csi_past, 2, raw = TRUE) + bio1 + tmean_dif + evt + fire_frq * fire_recent, data = mod_sf_unique)
model_interact <- lm(Ho ~ glacier + gHM + ai+ poly(csi_past, 2, raw = TRUE) + bio1 * tmean_dif + evt + fire_frq * fire_recent, data = mod_sf_unique)

AIC(model_nointeract, model_interact) # The interaction model has a lower AIC
anova(model_nointeract, model_interact) # The interaction model is significantly better
```

## Fit initial OLS and GLS models
```{r}
f <- formula("Ho ~ glacier + gHM + ai + poly(csi_past, 2, raw = TRUE) + bio1 * tmean_dif + evt + fire_frq * fire_recent")

# Run OLS and get residuals
ols_model <- lm(f, data = mod_sf_unique)
summary(ols_model)
AIC(ols_model)
mod_sf_unique$ols_resids <- residuals(ols_model)
broom::tidy(ols_model) %>% filter(p.value < 0.05) %>% arrange(desc(abs(estimate)))

# Fit variograms to the residuals
ols_variogram <- variogram(ols_resids ~ 1, data = mod_sf_unique)
ols_vgm <- map(c("Exp", "Sph", "Gau"), ~fit.variogram(ols_variogram, model = vgm(.x)))
names(ols_vgm) <- c("Exp", "Sph", "Gau")

pdf(here(plotpath, "ols_variogram.pdf"))
plot(ols_variogram, ols_vgm[["Exp"]], main = "OLS Residuals Variogram (Exponential)")
plot(ols_variogram, ols_vgm[["Sph"]], main = "OLS Residuals Variogram (Spherical)")
plot(ols_variogram, ols_vgm[["Gau"]], main = "OLS Residuals Variogram (Gaussian)")
dev.off()
# Based on visual inspection, the exponential model seems to fit best

# Fit a GLS model
# Note: nugget=FALSE assumes residuals at the same location are perfectly correlated, which should align with heterozygosity, since you don't expect to have much microscale variation in heterozygosity.
# nugget=TRUE allows unstructured microscale variance (random noise at zero distance), which can absorb spatial signal and cause predictor effects to appear insignificant. 
gls_model <- gls(f, correlation = corExp(form = ~ x + y, nugget = FALSE), data = mod_sf_unique, method = "ML") # Fit with ML to compare AIC later
summary(gls_model)
# Use type = "normalized" to get residuals decorrelated and standardized by the fitted 
# spatial covariance matrix; this removes modeled spatial structure and makes them 
# suitable for Moran's I. If you don't do this Moran's I will appear higher than it actually is.
mod_sf_unique$gls_resids <- residuals(gls_model, type = "normalized")

# Plot residuals
plot_resids <- function(resids){
  # Remove max residual because it is hard to see spatial patterns with it
  ggplot(mod_sf_unique[mod_sf_unique[[resids]] != max(mod_sf_unique[[resids]]), ]) +
    geom_sf(data = ca_proj, col = NA) +
    geom_sf(aes(col = .data[[resids]])) +
    theme_void() +
    scale_color_viridis_c(option = "turbo", end = 0.1, begin = 0.9, direction = -1) +
    labs(title = paste("Residuals:", resids))
}

pdf(here(plotpath, "all_model_residuals.pdf"), width = 9, height = 4)
plt_list <- list(
  plot_resids("ols_resids"),
  plot_resids("gls_resids")
)
plot_grid(plotlist = plt_list, nrow = 1)
dev.off()
# The GLS model residuals look the same as the OLS residuals

# Compare AIC
anova(gls_model, ols_model) # GLS significantly better, but absolute AIC difference is small

# Compare significant variables (p < 0.05) for each model using map
models <- list(OLS = ols_model, GLS = gls_model)
sig_vars_by_model <- map(models, ~ broom::tidy(.x) %>%
  filter(!term %in% c("lambda", "rho", "(Intercept)"), p.value < 0.05) %>%
  pull(term))
print(sig_vars_by_model) # Same significant variables for both models
```


# Fit full model
```{r}
f_final <- formula("Ho ~ glacier + gHM + ai + poly(csi_past, 2, raw = TRUE) + bio1 * tmean_dif + evt + fire_frq * fire_recent")

# Fit model
gls_model <- gls(f_final, correlation = corExp(form = ~ x + y, nugget = FALSE), data = mod_sf_unique_scaled, method = "ML")
ols_model <- lm(f_final, data = mod_sf_unique_scaled)

performance::check_collinearity(gls_model) 
performance::check_collinearity(ols_model) 
# High VIF for fire interaction term, but that is expected; otherwise all our <10
# (note: VIF may be high for fire_recent/interaction because of the imbalance in the number of burned vs unburned samples within the no fire/low frequency group)

# Compare OLS vs GLS
anova(gls_model, ols_model) # GLS significantly better, but absolute AIC difference is small
AIC(gls_model) - AIC(ols_model) # -7.83

# Coordinates matrix
coords_mat <- cbind(mod_sf_unique$x, mod_sf_unique$y)

# 8 nearest neighbors (robust for irregular sampling)
knn_8 <- knearneigh(coords_mat, k = 8)
nb_8  <- knn2nb(knn_8)
lw_8  <- nb2listw(nb_8, style = "W")

# Moran's I for OLS residuals
moran_ols <- moran.test(residuals(ols_model), lw_8)

# Moran's I for GLS normalized residuals
moran_gls <- moran.test(residuals(gls_model), lw_8)

print(moran_ols) # 0.15
print(moran_gls) # 0.15

# Refit final model with OLS since no gains in spatial autocorrelation and OLS is more interpretable (no spatial parameters like lambda)
final_model <- lm(f_final, data = mod_sf_unique_scaled)

# Results
mod_result <- broom::tidy(final_model)
mod_result %>% arrange(p.value)

# Significant variables
mod_result %>% filter(p.value < 0.05) %>% arrange(desc(abs(estimate))) 
sig_vars <- mod_result %>% filter(p.value < 0.05, !term %in% c("(Intercept)", "lambda")) %>% pull(term)
print(sig_vars)

r2 <- round(summary(final_model)$adj.r.squared, 4)
print(paste("R²:", r2))
```

```{r}
# !Most important test: test for differences between recent fire vs no recent fire within each fire frequency group
emm <- emmeans(final_model, ~ fire_recent | fire_frq)
contrast(emm, method = "pairwise") # No adjustment because there are only two levels for fire
# fire_frq = No Fire/Low Frequency:
#  contrast           estimate       SE  df t.ratio p.value
#  Unburned - Burned  2.10e-04 1.64e-04 232   1.277  0.2029

# fire_frq = Intermediate Frequency:
#  contrast           estimate       SE  df t.ratio p.value
#  Unburned - Burned  3.11e-05 4.80e-05 232   0.647  0.5180

# fire_frq = High Frequency:
#  contrast           estimate       SE  df t.ratio p.value
#  Unburned - Burned -1.14e-04 5.11e-05 232  -2.235  0.0264

# Results are averaged over the levels of: glacier, evt 

# Test for differences between vegetation groups
emm <- emmeans(final_model, specs = "evt")
tk <- contrast(emm, method = "tukey", adjust = "tukey") %>% as.data.frame()
common_evt <- 
  mod_sf_unique_scaled %>% 
  group_by(evt) %>% 
  summarise(n = n()) %>%
  arrange(desc(n)) %>% 
  filter(n > 10) %>% 
  pull(evt)
tk %>%
  filter(p.value < 0.05)

tk %>% 
  filter(grepl(paste(paste0(common_evt, " - "), collapse = "|"), contrast)) %>%
  filter(grepl(paste(paste0(" - ", common_evt), collapse = "|"), contrast)) %>%
  arrange(p.value) %>%
  head()

# Test for differences between fire regime groups
emm <- emmeans(final_model, specs = "fire_frq")
contrast(emm, method = "tukey", adjust = "tukey")
#  contrast                                          estimate       SE  df
#  (No Fire/Low Frequency) - Intermediate Frequency -2.99e-04 9.07e-05 232
#  (No Fire/Low Frequency) - High Frequency         -2.27e-04 9.26e-05 232
#  Intermediate Frequency - High Frequency           7.21e-05 3.92e-05 232
#  t.ratio p.value
#   -3.297  0.0032
#   -2.451  0.0396
#    1.843  0.1581

# Comparison with interaction term
emm <- emmeans(final_model, ~ fire_frq | fire_recent)
contrast(emm, method = "pairwise", adjust = "tukey")
# fire_recent = Unburned:
#  contrast                                          estimate       SE  df
#  (No Fire/Low Frequency) - Intermediate Frequency -2.10e-04 6.08e-05 232
#  (No Fire/Low Frequency) - High Frequency         -6.51e-05 6.06e-05 232
#  Intermediate Frequency - High Frequency           1.45e-04 4.14e-05 232
#  t.ratio p.value
#   -3.450  0.0019
#   -1.075  0.5308
#    3.496  0.0016

# fire_recent = Burned:
#  contrast                                          estimate       SE  df
#  (No Fire/Low Frequency) - Intermediate Frequency -3.88e-04 1.65e-04 232
#  (No Fire/Low Frequency) - High Frequency         -3.89e-04 1.67e-04 232
#  Intermediate Frequency - High Frequency          -4.53e-07 6.00e-05 232
#  t.ratio p.value
#   -2.355  0.0505
#   -2.327  0.0541
#   -0.008  1.0000

# Results are averaged over the levels of: glacier, evt 
# P value adjustment: tukey method for comparing a family of 3 estimates 
```

## Test to see which variable makes GHM no longer significant
```{r}
f_test <- formula("Ho ~ glacier + gHM + ai + poly(csi_past, 2, raw = TRUE) + tmean_dif * bio1 + evt + fire_frq * fire_recent")

vars_to_test <- c("glacier", "ai", "poly(csi_past, 2, raw = TRUE)", "tmean_dif", "bio1", "evt", "fire_frq", "fire_recent", "fire_frq:fire_recent")

ghm_significance <- map_dfr(vars_to_test, ~{
  var <- .x
  f_mod <- as.formula(paste("Ho ~ gHM +", paste(setdiff(vars_to_test, var), collapse = " + ")))
  mod <- lm(f_mod, data = mod_sf_unique_scaled)
  tidy_mod <- broom::tidy(mod)
  ghm_row <- tidy_mod %>% filter(term == "gHM")
  data.frame(removed_var = var, estimate = ghm_row$estimate, p_value = ghm_row$p.value)
})

ghm_significance %>% 
  filter(p_value < 0.05)
#                     removed_var     estimate    p_value
# 1 poly(csi_past, 2, raw = TRUE) 5.368732e-05 0.02264467
```

# Figures
```{r}
# Define a color scale function using MetBrewer with an argument to extend the scale for better visualization
scale_met_cont <- function(values, aesthetic = c("fill", "colour"), extend_scale = 0.1, r_min  = NULL, r_max = NULL) {
  aesthetic <- match.arg(aesthetic)
  if (is.null(r_min)) r_min <- min(values, na.rm = TRUE)
  if (is.null(r_max)) r_max <- max(values, na.rm = TRUE)
  rng   <- r_max - r_min
  lims  <- c(r_min - rng*extend_scale, r_max + rng*extend_scale)

  pal <- rev(MetBrewer::met.brewer("Hiroshige", type = "continuous"))
  if (aesthetic == "fill") {
    ggplot2::scale_fill_gradientn(colors = pal, limits = lims, oob = scales::squish)
  } else {
    ggplot2::scale_colour_gradientn(colors = pal, limits = lims, oob = scales::squish)
  }
}

df = mod_sf_unique
model = final_model
x = "fire_frq"
y = "Ho"
col_var = "Ho_original"
residual = TRUE
plot_mod <- function(x, y, df, model, mod_vars,
                     col_var = "Ho_original", residual = TRUE, interact = TRUE, bump_legend = 0 ) {

  df$Ho_original <- df$Ho

  if (residual) {
    numeric_mod_vars <- names(df)[sapply(df, is.numeric)]
    df_scaled <- 
      df %>%
      mutate(across(all_of(numeric_mod_vars), ~ as.numeric(scale(.x))))

    r <- residuals(model)
    X <- model.matrix(terms(model), data = df_scaled)
    b <- coef(model)
    b <- b[colnames(X)]

    if (x %in% c("bio1", "tmean_dif")) {
      cols <- grepl(paste0("(^", x, "$)|(^", x, ":)|(:", x, "$)"), colnames(X))
    } else {
      cols <- grepl(x, colnames(X))
    }

    if (all(!cols)) stop(paste("Variable", x, "not found in model"))

    sub_X <- X[, cols, drop = FALSE]
    sub_b <- b[cols]

    if (!all(colnames(sub_X) == names(sub_b))) {
      print(colnames(sub_X))
      print(names(sub_b))
      stop("Column names don't match")
    }

    df[[y]] <- r + drop(sub_X %*% sub_b)
  }

  model_summary <- broom::tidy(model)

  fmt_beta <- function(x, digits = 2) {
    s <- sprintf(paste0("%.", digits, "e"), x)
    mant <- sub("e.*", "", s)
    expo <- sub(".*e([+-]?[0-9]+)", "\\1", s)
    paste0(mant, "%*%10^", expo)
  }

  b_annotation <- 
    model_summary %>%
    dplyr::filter(grepl(x, term)) %>%
    dplyr::mutate(
      beta = fmt_beta(estimate),
      p_lab = ifelse(
        p.value < 0.001,
        paste0("p[", row_number(), "] < 0.001"),
        paste0("p[", row_number(), "] == ", signif(p.value, 3))
      ),
      line = paste0("beta[", row_number(), "] == ", beta, "~\",\"~", p_lab)
    )

  if (x == "csi_past") {

    if (nrow(b_annotation) == 1) {
      label_txt <- b_annotation$line[1]
    } else {
      label_txt <- Reduce(function(a, b) paste0("atop(", a, ",", b, ")"), b_annotation$line)
    }

    plt <-
      ggplot(df, aes(x = .data[[x]], y = .data[[y]])) +
      geom_point(aes(col = .data[[col_var]]), size = 2, show.legend = FALSE) +
      geom_smooth(method = "lm", colour = "black", formula = y ~ poly(x, 2)) +
      annotate("text", x = max(df[[x]], na.rm = TRUE), y = min(df[[y]], na.rm = TRUE), label = label_txt, size = 5, parse = TRUE, vjust = -0.1, hjust = 1)

  } else if (x == "fire_frq" && interact) {
    dodge_w <- 0.8
    frq_lvls <- levels(df$fire_frq)
    rec_lvls <- levels(df$fire_recent)

    df$new_fire_frq <- factor(df$fire_frq, levels = levels(df$fire_frq),
                              labels = c("No Fire/Low", "Intermediate", "High"))

    off <- setNames(
      ((seq_along(frq_lvls) - (length(frq_lvls) + 1) / 2) * (dodge_w / length(frq_lvls))),
      frq_lvls
    )

    emm <- emmeans(model, ~ fire_frq | fire_recent)

    pairs_df <- contrast(emm, method = "pairwise", adjust = "tukey") %>%
      as.data.frame() %>%
      tidyr::separate(contrast, c("group1","group2"), " - ") %>%
      dplyr::mutate(
        dplyr::across(c(group1, group2), ~ stringr::str_replace_all(.x, "[()]", "")),
        annotations = cut(p.value, c(-Inf, .001, .01, .05, Inf), c("***","**","*","ns")),
        xbase = as.numeric(factor(fire_recent, levels = rec_lvls)),
        xmin  = xbase + off[group1],
        xmax  = xbase + off[group2]
      ) %>%
      dplyr::filter(annotations != "ns") %>%
      dplyr::group_by(fire_recent) %>%
      dplyr::ungroup()

    y_height_adj <- ifelse(residual, 0.5, 0.9)
    pairs_df$y_position <- max(df[[y]], na.rm=TRUE) * (y_height_adj + 0.08*seq_len(nrow(pairs_df)))

    plt <-
      ggplot(df, aes(x = fire_recent, y = .data[[y]], fill = new_fire_frq)) +
      geom_boxplot(position = position_dodge(dodge_w), outlier.shape = NA) +
      geom_point(
        aes(color = .data[[col_var]], group = new_fire_frq),
        position = position_jitterdodge(jitter.width = 0.3, dodge.width = dodge_w),
        size = 2, show.legend = FALSE
      ) +
      ggpubr::stat_pvalue_manual(
        pairs_df,
        label = "annotations",
        xmin = "xmin", xmax = "xmax",
        y.position = "y_position",
        size = 8, tip.length = 0, vjust = 0.6,
        inherit.aes = FALSE
      ) +
      labs(x = "Contemporary fire occurrence", fill = "Historical fire frequency") +
      scale_fill_manual(values = c("white", "gray60", "gray30"))

    scale_y <- ifelse(residual, 1.1, 1.4)
    plt <- plt + ylim(NA, max(df[[y]]) * scale_y)

  } else if (x == "fire_frq") {
    emm  <- emmeans(model, specs = x)
    pairs_df <- 
      pairs(emm, adjust = "tukey") %>% summary(infer = TRUE) %>% as.data.frame() %>%
      separate(contrast, c("group1","group2"), " - ") %>%
      mutate_at(vars(group1, group2), ~str_replace_all(.x, "[()]", "")) %>%
      mutate(annotations = cut(p.value, c(-Inf,.001,.01,.05,Inf), c("***","**","*","ns"))) %>%
      dplyr::filter(annotations != "ns") %>%
      mutate(xmin = group1, xmax = group2)

    df$new_x <- factor(df[[x]], levels = levels(df[[x]]), labels = c("No Fire/Low", "Intermediate", "High"))
    pairs_df$xmin <- factor(pairs_df$xmin, levels = levels(df[[x]]), labels = c("No Fire/Low", "Intermediate", "High"))
    pairs_df$xmax <- factor(pairs_df$xmax, levels = levels(df[[x]]), labels = c("No Fire/Low", "Intermediate", "High"))

    y_height_adj <- ifelse(residual, 0.8, 1)
    pairs_df$y_position <- max(df[[y]], na.rm = TRUE) * (y_height_adj + 0.08*seq_len(nrow(pairs_df)))

    plt <- 
      ggplot(df, aes(new_x, .data[[y]])) +
      geom_boxplot() +
      geom_jitter(aes(col = .data[[col_var]]), size = 2, width = 0.2, show.legend = FALSE) +
      ggpubr::stat_pvalue_manual(
        pairs_df,
        label = "annotations",
        xmin = "xmin", xmax = "xmax",
        y.position = "y_position",
        size = 8,
        tip.length = 0,
        vjust = 0.6
      )

  } else if (x == "fire_recent") {

    dodge_w <- 0.8
    frq_lvls <- levels(df$fire_frq)
    rec_lvls <- levels(df$fire_recent)
    nrec <- length(rec_lvls)
    off <- setNames(((seq_len(nrec) - (nrec + 1)/2) * (dodge_w / nrec)), rec_lvls)

    emm <- emmeans(model, ~ fire_recent | fire_frq)
    ymax_df <- df %>% dplyr::group_by(fire_frq) %>% dplyr::summarise(ymax = max(.data[[y]], na.rm=TRUE), .groups="drop")

    pairs_df <- contrast(emm, "pairwise") %>% as.data.frame() %>%
      separate(contrast, c("group1","group2"), " - ") %>%
      mutate(
        across(c(group1, group2), ~ stringr::str_replace_all(.x, "[()]", "")),
        annotations = cut(p.value, c(-Inf,.001,.01,.05,Inf), c("***","**","*","ns")),
        xbase = as.numeric(factor(fire_frq, levels = frq_lvls)),
        xmin  = xbase + off[group1],
        xmax  = xbase + off[group2]
      ) %>%
      filter(annotations != "ns") %>%
      left_join(ymax_df, by="fire_frq") %>%
      group_by(fire_frq) %>%
      mutate(y_position = ymax * (1 + 0.08*(row_number()-1))) %>%
      ungroup()

    df$new_fire_frq <- factor(df$fire_frq, levels = levels(df$fire_frq), labels = c("No Fire/Low", "Intermediate", "High"))
    pairs_df$fire_frq <- factor(pairs_df$fire_frq, levels = levels(df$fire_frq), labels = c("No Fire/Low", "Intermediate", "High"))

    plt <-
      ggplot(df, aes(x = new_fire_frq, y = .data[[y]])) +
      geom_boxplot(aes(fill = fire_recent), position = position_dodge(dodge_w), outlier.shape = NA) +
      geom_point(
        aes(color = .data[[col_var]], group = fire_recent),
        position = position_jitterdodge(jitter.width = 0.3, dodge.width = dodge_w),
        size = 2, show.legend = FALSE
      ) +
      guides(col = "none") +
      scale_fill_manual(values = c("Burned" = "gray", "Unburned" = "white")) +
      labs(fill = "Contemporary fire\noccurrence")

    if (nrow(pairs_df) != 0) {
      plt <- plt +
        ggpubr::stat_pvalue_manual(
          pairs_df,
          label = "annotations", xmin = "xmin", xmax = "xmax", y.position = "y_position",
          size = 10, tip.length = 0, vjust = 0.6
        )
    }

    scale_y <- ifelse(residual, NA, 1.4)
    plt <- plt + ylim(NA, max(df[[y]])*scale_y)

  } else if (x %in% c("tmean_dif", "bio1") && interact) {

    mod_raw <- lm(formula(model), data = df)
    fmt <- function(x) paste0(round(x, 1), " °C")

    if (x == "tmean_dif") {
      jn <- johnson_neyman(mod_raw, pred = tmean_dif, modx = bio1)
      jn_bounds <- jn$bounds
      breaks <- c(min(df$bio1, na.rm = TRUE), jn_bounds[1], jn_bounds[2], max(df$bio1, na.rm = TRUE))
      labels <- c(
        paste0(fmt(breaks[1]), " – ", fmt(breaks[2])),
        paste0(fmt(breaks[2]), " – ", fmt(breaks[3])),
        paste0(fmt(breaks[3]), " – ", fmt(breaks[4]))
      )
      df$mod_bin <- cut(df$bio1, breaks = breaks, include.lowest = TRUE)
      df$mod_bin <- factor(df$mod_bin, labels = labels)
      df$mod_bin <- factor(df$mod_bin, levels = rev(levels(df$mod_bin)))
      cols <- rev(c("navy", "gray60", "tomato2"))
      shapes <- rev(c(25, 21, 24))

      plt <-
        ggplot(df, aes(x = tmean_dif, y = .data[[y]])) +
        geom_point(size = 2, aes(fill = .data[[col_var]], pch = mod_bin, color = mod_bin)) +
        geom_smooth(data = filter(df, bio1 < breaks[2]), method = "lm", fill = "navy", col = "navy", alpha = 0.2) +
        geom_smooth(data = filter(df, bio1 > breaks[3]), method = "lm", fill = "tomato2", col = "tomato2", alpha = 0.2) +
        labs(color = "Contemporary\ntemperature", pch = "Contemporary\ntemperature", x = make_pretty_names("tmean_dif")) +
        scale_met_cont(df[[col_var]], aesthetic = "fill", extend_scale = 0.1) +
        guides(fill = "none") +
        scale_shape_manual(values = shapes) +
        scale_color_manual(values = cols)

      scale_y <- ifelse(residual, 1.1, 1.3)
      plt <- plt + ylim(NA, max(df[[y]], na.rm = TRUE) * scale_y)

    } else if (x == "bio1") {
      jn <- johnson_neyman(mod_raw, pred = bio1, modx = tmean_dif)
      jn_bounds <- jn$bounds
      breaks <- c(min(df$tmean_dif, na.rm = TRUE), jn_bounds[1], jn_bounds[2], max(df$tmean_dif, na.rm = TRUE))
      labels <- c(
        paste0(fmt(breaks[1]), " – ", fmt(breaks[2])),
        paste0(fmt(breaks[2]), " – ", fmt(breaks[3])),
        paste0(fmt(breaks[3]), " – ", fmt(breaks[4]))
      )
      df$mod_bin <- cut(df$tmean_dif, breaks = breaks, include.lowest = TRUE)
      df$mod_bin <- factor(df$mod_bin, labels = labels)
      df$mod_bin <- factor(df$mod_bin, levels = rev(levels(df$mod_bin)))
      cols <- rev(c("navy", "gray60", "tomato2"))
      shapes <- rev(c(25, 21, 24))

      plt <-
        ggplot(df, aes(x = bio1, y = .data[[y]])) +
        geom_point(size = 2, aes(fill = .data[[col_var]], pch = mod_bin, color = mod_bin)) +
        geom_smooth(data = filter(df, tmean_dif < breaks[2]), method = "lm", fill = "navy", col = "navy", alpha = 0.2) +
        geom_smooth(data = filter(df, tmean_dif > breaks[3]), method = "lm", fill = "tomato2", col = "tomato2", alpha = 0.2) +
        labs(color = "Temperature\nchange", pch = "Temperature\nchange", x = make_pretty_names("bio1")) +
        scale_met_cont(df[[col_var]], aesthetic = "fill", extend_scale = 0.1) +
        guides(fill = "none") +
        scale_shape_manual(values = shapes) +
        scale_color_manual(values = cols)

      scale_y <- ifelse(residual, 1.1, 1.3)
      plt <- plt + ylim(NA, max(df[[y]], na.rm = TRUE) * scale_y)
    }

  } else if (x == "evt" | x == "glacier") {
    emm  <- emmeans(model, specs = x)
    pairs_df <- 
      pairs(emm, adjust = "tukey") %>% summary(infer = TRUE) %>% as.data.frame() %>%
      separate(contrast, c("group1","group2"), " - ") %>%
      mutate_at(vars(group1, group2), ~str_replace_all(.x, "[()]", "")) %>%
      mutate(annotations = cut(p.value, c(-Inf,.001,.01,.05,Inf), c("***","**","*","ns"))) %>%
      mutate(xmin = group1, xmax = group2) %>%
      filter(annotations != "ns")

    if (nrow(pairs_df) == 0) {
      message(paste("No significant pairwise differences for", x))
      plt <- 
        ggplot(df, aes(.data[[x]], .data[[y]])) +
        geom_boxplot() +
        geom_jitter(aes(col = .data[[col_var]]), size = 2, width = 0.2, show.legend = FALSE)
    } else {
      y_height_adj <- ifelse(residual, 0.8, 1)
      pairs_df$y_position <- max(df[[y]], na.rm=TRUE) * (y_height_adj + 0.08*seq_len(nrow(pairs_df)))

      plt <- 
        ggplot(df, aes(.data[[x]], .data[[y]])) +
        geom_boxplot() +
        geom_jitter(aes(col = .data[[col_var]]), size = 2, width = 0.15, show.legend = FALSE) +
        ggpubr::stat_pvalue_manual(
          pairs_df,
          label = "annotations",
          xmin = "xmin", xmax = "xmax",
          y.position = "y_position",
          size = 8,
          tip.length = 0,
          vjust = 0.6
        )
    }

  } else {

    if (nrow(b_annotation) == 1) {
      label_txt <- b_annotation$line[1]
    } else {
      label_txt <- Reduce(function(a, b) paste0("atop(", a, ",", b, ")"), b_annotation$line)
    }

    plt <-
      ggplot(df, aes(x = .data[[x]], y = .data[[y]])) +
      geom_point(aes(col = .data[[col_var]]), size = 2, show.legend = FALSE) +
      geom_smooth(method = "lm", colour = "black", formula = y ~ x) +
      annotate("text", x = max(df[[x]], na.rm = TRUE), y = min(df[[y]], na.rm = TRUE), label = label_txt, size = 5, parse = TRUE, vjust = -0.1, hjust = 1)
  }

  x_label <-
    case_when(
      x == "fire_recent" ~ "fire_frq",
      x == "fire_frq" & interact ~ "fire_recent",
      TRUE ~ x
    )

  plt <- 
    plt +
    xlab(stringr::str_to_sentence(make_pretty_names(x_label)) %>%
    stringr::str_replace_all("°c", "°C")) +
    theme_classic() +
    theme(axis.title = element_text(size = 16), axis.text = element_text(size = 14)) 
      
  if (!(x %in% c("tmean_dif", "bio1") && interact)) {
    plt <- plt + scale_met_cont(df[[col_var]], aesthetic = "colour", extend_scale = 0.1)
  }

  legend.position.inside.x <-
    case_when(
      x == "fire_recent" ~ 0.48,
      x == "fire_frq" & interact ~ 0.61,
      x == "tmean_dif" & interact ~ 0.5,
      x == "bio1" & interact ~ 0.4
    )
  if (residual) legend.position.inside.x <- legend.position.inside.x + 0.1

  if (x == "fire_recent" | (x == "fire_frq" & interact) | (x %in% c("tmean_dif", "bio1") & interact)) {
    plt <- 
      plt +
      theme(
        legend.position = "inside",
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 14),
        legend.justification = c(1, 1),
        legend.background = element_rect(fill = rgb(1,1,1,0.5), color = "black"),
        legend.position.inside = c(legend.position.inside.x + bump_legend, 1)
      )
  }

  if (x == "evt") {
    plt <- plt + theme(axis.text.x = element_text(angle = 30, hjust = 1))
  }

  plt <- plt + theme(plot.margin = margin(2, 2, 2, 2, unit = "mm"))

  if (residual) {
    plt + labs(y = "Partial residual heterozygosity")
  } else {
    plt + labs(y = "Observed heterozygosity")
  }
}

# Load polygons for plotting
ca_proj <- get_ca() %>% sf::st_transform(3310) # California boundaries
range_map <- get_range() # Sceloporus range map

# Loaded wingen map
wingen_r <- 
  terra::rast(here(outpath, "Ho_kriged_masked.tif")) %>% 
  mask(range_map) %>% 
  mask(ca_proj) %>%
  as.data.frame(xy = TRUE, na.rm = TRUE) 
names(wingen_r)[3] <- "layer"  # ensure 'layer' is the raster value column

# Isolate variables to plot vs all variables
vars_to_plot <- c("csi_past", "tmean_dif", "fire_frq", "fire_recent")
mod_vars <- c("csi_past", "tmean_dif", "ai", "bio1", "glacier", "gHM", "fire_frq", "fire_recent", "evt")

# Read in png image of sceloporus
img <- png::readPNG(here("sceloporus_image.png"))

# A: wingen map of heterozygosity 
ho_plot <- ggplot(wingen_r) +
  geom_sf(data = ca_proj, fill = "gray90", colour = NA, size = 2) +
  geom_tile(aes(x = x, y = y, fill = layer), colour = NA) +
  scale_met_cont(c(mod_sf_unique$Ho, wingen_r$layer), aesthetic = "fill", extend_scale = 0.1) +
  labs(fill = "Heterozygosity") +
  theme_void() +
  geom_sf(data = mod_sf_unique, shape = 21, aes(fill = Ho), colour = "#00000087", size = 2) +
  theme(
    plot.margin = margin(1, 1, 1, 1, unit = "mm"),
    legend.position = "inside",              # turn on "inside" mode
    legend.justification = c(1, 1),          # anchor legend's top-right corner
    legend.position.inside = c(0.9, 0.95),   # place near panel's top-right
    legend.key.height = unit(0.8, "cm"),
    legend.key.width  = unit(0.8, "cm"),
    legend.title = element_text(size = 16),
    legend.text  = element_text(size = 14)
  )

png(here("genetic_diversity_map.png"), height = 6, width = 6, res = 300, units = "in")
ho_plot
dev.off()

# Add image of sceloporus to the plot
library(grid)
img_grob <- rasterGrob(as.raster(img), interpolate = TRUE)

xr <- range(wingen_r$x, na.rm = TRUE)
yr <- range(wingen_r$y, na.rm = TRUE)

ho_plot_annotated <- ho_plot +
  annotation_custom(
    img_grob,
    xmin = xr[1] - 0.1 * diff(xr),                               
    xmax = xr[1] + 0.75 * diff(xr),               
    ymin = yr[1] - 0.2 * diff(yr),                 
    ymax = yr[1] + 0.3 * diff(yr)               
  )
  
# B: partial residuals 
plts_residual_y <- purrr::map(
  vars_to_plot,
  # Use unscaled data for plotting
  # Keep fire_frq simple
  ~{
    interact <- ifelse(.x == "fire_frq", FALSE, TRUE)
    plot_mod(.x, "Ho", mod_sf_unique, model = final_model, col_var = "Ho_original", residual = TRUE, interact = interact) + 
    theme(plot.margin = margin(7, 7, 7, 7))
  }
)

# B: Raw values
plts_raw_y <- purrr::map(
  vars_to_plot,
  # Use unscaled data for plotting
  # Keep fire_frq simple
  ~{
    interact <- ifelse(.x == "fire_frq", FALSE, TRUE)
    plot_mod(.x, "Ho", mod_sf_unique, model = final_model, col_var = "Ho_original", residual = FALSE, interact = interact) + 
    theme(plot.margin = margin(7, 7, 7, 7))
  }
)

# C: predicted vs observed 
pred_vs_obs <-
  # Use unscaled data for plotting
  ggplot(mod_sf_unique, aes(x = predict(final_model), y = Ho)) +
  geom_point(aes(col = Ho), shape = 16, size = 2) +
  geom_smooth(method = "lm", colour = "black") +
  scale_met_cont(c(mod_sf_unique$Ho, wingen_r$layer), aesthetic = "colour", extend_scale = 0.1) +
  theme_classic() +
  labs(x = "Predicted heterozygosity", y = "Observed heterozygosity") +
  ggpubr::stat_cor(size = 6) +
  theme(legend.position = "none",
        axis.title = element_text(size = 16),
        axis.text  = element_text(size = 14))

pred_vs_obs_adj <- cowplot::plot_grid(pred_vs_obs, NULL, nrow = 1, rel_widths = c(0.9, 0.1))

plt2_residual <- cowplot::plot_grid(plotlist = plts_residual_y, ncol = 2, align = "hv")
plt2_raw <- cowplot::plot_grid(plotlist = plts_raw_y, ncol = 2, align = "hv")
plt1 <- cowplot::plot_grid(plotlist = list(ho_plot_annotated, pred_vs_obs_adj), ncol = 1,
                           labels = c("A", "C"), rel_heights = c(0.6, 0.45), label_size = 18)

png(here(plotpath, "figure_Ho_residual.png"), height = 10*0.9, width = 16*0.9, res = 300, units = "in")
cowplot::plot_grid(plt1, plt2_residual, ncol = 2, labels = c("", "B"),  rel_widths = c(0.33, 0.66), align = "h", label_size = 18)
dev.off()

plt2_raw <- cowplot::plot_grid(plotlist = plts_raw_y, ncol = 2, align = "hv")
plt <- cowplot::plot_grid(plt1, plt2_raw, ncol = 2,labels = c("", "B"),  rel_widths = c(0.33, 0.66), align = "h", label_size = 18)
png(here(plotpath, "figure_model_results.png"), height = 10, width = 16, res = 300, units = "in")
plt
dev.off()

pdf(here(plotpath, "figure_model_results.pdf"), height = 10*0.9, width = 16*0.9)
plt
dev.off()
```

```{r}
# Plot all variables
final_model_vars <- c("csi_past", "tmean_dif", "tmean_dif", "bio1", "bio1", "ai", "fire_frq", "fire_recent", "fire_frq", "glacier", "gHM", "evt")
interact <- rep(TRUE, length(final_model_vars))
interact[3] <- FALSE
interact[5] <- FALSE
interact[7] <- FALSE

# Partial residuals 
plts_residual_y <- 
  purrr::map2(
    final_model_vars,
    interact,
    # Use unscaled data for plotting
    ~plot_mod(.x, "Ho", mod_sf_unique, model = final_model, col_var = "Ho_original", residual = TRUE, interact = .y)
  ) %>% 
  map(~ .x + theme(plot.margin = margin(7, 7, 7, 7)))

# B: Raw values
plts_raw_y <- 
  purrr::map2(
    final_model_vars,
    interact, 
    # Use unscaled data for plotting
    ~plot_mod(.x, "Ho", mod_sf_unique, model = final_model, col_var = "Ho_original", residual = FALSE, bump_legend = 0.05, interact = .y)
  ) %>% 
  map(~ .x + theme(plot.margin = margin(7, 7, 7, 7)))

plt_residual <- cowplot::plot_grid(plotlist = plts_residual_y, ncol = 3, align = "v", labels = LETTERS[1:length(plts_residual_y)], label_size = 18, rel_widths = 1)

plt_raw <- cowplot::plot_grid(plotlist = plts_raw_y, ncol = 3, align = "v", labels = LETTERS[1:length(plts_raw_y)], label_size = 18)

png(here(plotpath, "figure_Ho_allvars_residual.png"), height = 16, width = 15, res = 300, units = "in")
plt_residual
dev.off()

png(here(plotpath, "figure_Ho_allvars_raw.png"), height = 16, width = 15, res = 300, units = "in")
plt_raw
dev.off()
```

## Spatial scale analysis
```{r}
# term-specific contributions (centered)
terms_mat <- predict(final_model, type = "terms")
term_names <- colnames(terms_mat)
# Remove glacier and gHM (not significant)
term_names <- term_names[!term_names %in% c("glacier", "gHM", "ai")]

# helper: compute variogram + fitted model + fit line for one term
one_term <- function(term, model = "Gau"){
  df_term <- mod_sf_unique_scaled
  df_term$signal <- as.numeric(terms_mat[, term])

  vg <- variogram(signal ~ 1, data = df_term)
  fit <- fit.variogram(vg, vgm(model))

  vg$term <- term

  fit_line <- variogramLine(fit, maxdist = max(vg$dist)) %>%
    mutate(term = term)

  tibble(
    term = term,
    range_m = fit$range[2],
    range_km = fit$range[2] / 1000,
    sill = fit$psill[2],
    nugget = fit$psill[1],
    vg = list(vg),
    fit_line = list(fit_line),
    fit = list(fit)
  )
}

res <- map_dfr(term_names, one_term, model = "Gau")

ranges_tbl <- res %>%
  select(term, range_km, nugget, sill) %>%
  arrange(range_km)

print(ranges_tbl)

vg_all <- bind_rows(res$vg)
fit_lines <- bind_rows(res$fit_line)

pdf(here(plotpath, "figure_variograms_all_terms.pdf"), height = 5, width = 8)
ggplot(vg_all, aes(dist / 1000, gamma, col = term)) +
  geom_point(size = 1.6, alpha = 0.8) +
  geom_line(data = fit_lines, aes(dist / 1000, gamma, col = term), linewidth = 1) +
  labs(x = "Distance (km)", y = "Semivariance", col = "Term") +
  theme_bw()
dev.off()

```

```{r}
fit_var <- function(spdf, var){
  f <- as.formula(paste(var, "~ 1"))
  vg <- variogram(f, spdf)
  fits <- list(
    Gau = tryCatch(fit.variogram(vg, vgm("Gau")), error = function(e) NULL),
    Sph = tryCatch(fit.variogram(vg, vgm("Sph")), error = function(e) NULL),
    Exp = tryCatch(fit.variogram(vg, vgm("Exp")), error = function(e) NULL)
  ) %>% purrr::compact()
  if (length(fits) == 0) return(tibble(model = NA_character_, range_km = NA_real_, vg = list(vg), fit_line = list(NULL)))
  aic <- purrr::map_dbl(fits, function(fit){
    n <- nrow(vg); k <- length(fit$psill) + 1; rss <- attr(fit, "SSErr")
    n * log(rss / n) + 2 * k
  })
  best_name <- names(fits)[which.min(aic)]
  best_fit <- fits[[best_name]]
  fit_line <- variogramLine(best_fit, maxdist = max(vg$dist))
  tibble(model = best_name, range_km = best_fit$range[2] / 1000, vg = list(vg), fit_line = list(fit_line))
}

cat_variogram_range <- function(spdf, class_var){
  lvls <- spdf[[class_var]] %>% as.factor() %>% levels()
  purrr::map_dfr(lvls, function(lvl){
    spdf$ind <- as.numeric(as.factor(spdf[[class_var]]) == lvl)
    fit_var(spdf, "ind") %>% mutate(level = lvl)
  }) %>% mutate(var = class_var) %>% select(var, level, model, range_km, vg, fit_line)
}

scale_test <- function(spdf, num_vars, cat_vars){
  range_num <- purrr::map_dfr(num_vars, function(v){
    fit_var(spdf, v) %>% mutate(var = v, level = NA_character_)
  }) %>% select(var, level, model, range_km, vg, fit_line)
  range_cat <- purrr::map_dfr(cat_vars, function(v) cat_variogram_range(spdf, v))
  bind_rows(range_num, range_cat) %>% arrange(var, level)
}

num_vars <- c("csi_past", "tmean_dif", "bio1")
cat_vars <- c("fire_frq", "fire_recent")
res_ind <- 
  scale_test(mod_sf_unique_scaled, num_vars, cat_vars) %>% 
  mutate(term = ifelse(is.na(level), var, paste0(var, ": ", level)))
terms_to_keep <- c("csi_past", "tmean_dif", "bio1", "fire_frq: Intermediate Frequency", "fire_recent: Burned")

vg_ind <- 
  purrr::imap_dfr(res_ind$vg, ~ mutate(.x, term = res_ind$term[.y])) %>% 
  filter(term %in% terms_to_keep) %>% 
  mutate(term = make_pretty_names(term, units = FALSE))

fit_ind <- 
  purrr::imap_dfr(res_ind$fit_line, ~ mutate(.x, term = res_ind$term[.y])) %>% 
  filter(term %in% terms_to_keep) %>% mutate(term = make_pretty_names(term, units = FALSE))

vlines_ind <- 
  res_ind %>% 
  filter(term %in% terms_to_keep) %>% 
  mutate(term = make_pretty_names(term, units = FALSE)) %>% 
  distinct(term, range_km)

p_ind <- ggplot(vg_ind, aes(dist/1000, gamma, col = term)) +
  geom_point(size = 1.6, alpha = 0.8) +
  geom_line(data = fit_ind, aes(dist/1000, gamma, col = term), linewidth = 1) +
  labs(x = "Distance (km)", y = "Semivariance of predictor", col = "Predictor") +
  theme_classic() +
  theme(legend.position = "right")

terms_mat <- predict(final_model, type = "terms")
term_names <- setdiff(colnames(terms_mat), c("glacier", "ai", "gHM", "bio1:tmean_dif"))

one_term <- function(term){
  df_term <- mod_sf_unique_scaled
  df_term$signal <- as.numeric(terms_mat[, term])
  vg <- variogram(signal ~ 1, data = df_term)

  fits <- list(
    Gau = tryCatch(fit.variogram(vg, vgm("Gau")), error = function(e) NULL),
    Sph = tryCatch(fit.variogram(vg, vgm("Sph")), error = function(e) NULL),
    Exp = tryCatch(fit.variogram(vg, vgm("Exp")), error = function(e) NULL)
  ) %>% purrr::compact()

  if (length(fits) == 0) return(tibble(term = term, model = NA_character_, range_km = NA_real_, vg = list(vg), fit_line = list(NULL)))

  aic <- purrr::map_dbl(fits, function(fit){
    n <- nrow(vg); k <- length(fit$psill) + 1; rss <- attr(fit, "SSErr")
    n * log(rss / n) + 2 * k
  })

  best_name <- names(fits)[which.min(aic)]
  best_fit <- fits[[best_name]]
  fit_line <- variogramLine(best_fit, maxdist = max(vg$dist))

  tibble(term = term, model = best_name, range_km = best_fit$range[2]/1000, vg = list(vg), fit_line = list(fit_line))
}

res_term <- 
  purrr::map_dfr(term_names, one_term) %>% 
  filter(!term %in% c("evt","fire_frq:fire_recent")) %>% 
  mutate(term = make_pretty_names(term, units = FALSE))

vg_term <- purrr::imap_dfr(res_term$vg, ~ mutate(.x, term = res_term$term[.y]))
fit_term <- purrr::imap_dfr(res_term$fit_line, ~ mutate(.x, term = res_term$term[.y]))
vlines_term <- res_term %>% distinct(term, range_km)

p_term <- ggplot(vg_term, aes(dist/1000, gamma, col = term)) +
  geom_point(size = 1.6, alpha = 0.8) +
  geom_line(data = fit_term, aes(dist/1000, gamma, col = term), linewidth = 1) +
  labs(x = "Distance (km)", y = "Semivariance of predictor contribution", col = "Predictor") +
  theme_classic() +
  theme(legend.position = "none")

# Add range points and labels
range_pts_ind <- vlines_ind %>% left_join(fit_ind, by = "term") %>% mutate(d = abs((dist/1000) - range_km)) %>% group_by(term) %>% slice_min(d, n = 1, with_ties = FALSE) %>% ungroup()
range_pts_term <- vlines_term %>% left_join(fit_term, by = "term") %>% mutate(d = abs((dist/1000) - range_km)) %>% group_by(term) %>% slice_min(d, n = 1, with_ties = FALSE) %>% ungroup()

p_ind <- p_ind +
  ggrepel::geom_label_repel(data = range_pts_ind, aes(x = range_km, y = gamma, label = sprintf("%.0f km", range_km), col = term), size = 4, fill = "white", label.size = 0.35, label.r = unit(0.15, "lines"), nudge_y = max(vg_ind$gamma, na.rm = TRUE) * 0.08, force = 3, force_pull = 0.2, box.padding = 0.6, point.padding = 0.5, segment.size = 0.9, segment.alpha = 0.8, show.legend = FALSE) +
  geom_point(data = range_pts_ind, aes(x = range_km, y = gamma, col = term), shape = 21, fill = "white", stroke = 1.2, size = 3.5, inherit.aes = FALSE) 

p_term <- p_term +
  ggrepel::geom_label_repel(data = range_pts_term, aes(x = range_km, y = gamma, label = sprintf("%.0f km", range_km), col = term), size = 4, fill = "white", label.size = 0.35, label.r = unit(0.15, "lines"), nudge_y = max(vg_term$gamma, na.rm = TRUE) * 0.08, force = 3, force_pull = 0.2, box.padding = 0.6, point.padding = 0.5, segment.size = 0.9, segment.alpha = 0.8, show.legend = FALSE) +
  geom_point(data = range_pts_term, aes(x = range_km, y = gamma, col = term), shape = 21, fill = "white", stroke = 1.2, size = 3.5, inherit.aes = FALSE) 

# Set colors
all_terms <- union(unique(vg_ind$term), unique(vg_term$term))
hiroshige_cols <- c("#6bb2cc", "#042346", "#f04f32", "#f7a957", "#295580")
names(hiroshige_cols) <- all_terms
p_ind <- p_ind + scale_color_manual(values = hiroshige_cols)
p_term <- p_term + scale_color_manual(values = hiroshige_cols)

# Combine plots side by side with shared legend
p_ind_noleg <- p_ind + theme(legend.position = "none")
p_ind <- p_ind + theme(legend.title = element_text(size = 14), legend.text = element_text(size = 12))
leg <- cowplot::get_legend(p_ind + theme(legend.position = "right"))
# Increase x axis label size
p_ind_noleg <- p_ind_noleg + theme(axis.title = element_text(size = 14), axis.text = element_text(size = 12), plot.margin = margin(5.5, 5.5, 5.5, 25))
p_term <- p_term + theme(axis.title = element_text(size = 14), axis.text = element_text(size = 12), plot.margin = margin(5.5, 5.5, 5.5, 25))

png(here(plotpath, "figure_variograms_side_by_side.png"), height = 8, width = 10, units = "in", res = 300)
cowplot::plot_grid(cowplot::plot_grid(p_ind_noleg, p_term, ncol = 1, align = "hv", labels = c("A","B"), label_size = 18), leg, nrow = 1, rel_widths = c(1,0.5), align = "v")
dev.off()
```

## Admixture plot
```{r}
`%||%` <- function(a, b) if (!is.null(a)) a else b

col_pal      <- c("Q ≤ 0.10" = "skyblue2", "Q > 0.10" = "red2")
col_pal_text <- c("Q ≤ 0.10" = "skyblue1", "Q > 0.10" = "red2")

bin_labels <- function(x, breaks = 5, digits = 2) {
  b <- cut(x, breaks = breaks, include.lowest = TRUE)
  lev <- levels(b)

  mids <- sapply(lev, \(s) {
    nums <- as.numeric(gsub("\\(|\\]|\\[|\\)", "", unlist(strsplit(s, ","))))
    mean(nums)
  })
  maxs <- sapply(lev, \(s) {
    nums <- as.numeric(gsub("\\(|\\]|\\[|\\)", "", unlist(strsplit(s, ","))))
    max(nums)
  })

  list(bin = b, lab = sprintf(paste0("%.", digits, "f"), mids), max = maxs)
}

admix_prop_long <- function(df, x_group, q = "Q", thresh = 0.10) {
  df %>%
    group_by({{ x_group }}) %>%
    summarise(
      n = dplyr::n(),
      n_hi = sum(.data[[q]] > thresh, na.rm = TRUE),
      prop_hi = n_hi / n,
      .groups = "drop"
    ) %>%
    transmute(
      {{ x_group }},
      `Q ≤ 0.10` = 1 - prop_hi,
      `Q > 0.10` = prop_hi
    ) %>%
    tidyr::pivot_longer(c(`Q ≤ 0.10`, `Q > 0.10`), names_to = "admix_bin", values_to = "prop") %>%
    mutate(admix_bin = factor(admix_bin, levels = c("Q ≤ 0.10", "Q > 0.10")))
}

admix_label_df <- function(df, x_group, q = "Q", thresh = 0.10) {
  df %>%
    group_by({{ x_group }}) %>%
    summarise(
      n = dplyr::n(),
      n_hi = sum(.data[[q]] > thresh, na.rm = TRUE),
      prop_hi = n_hi / n,
      .groups = "drop"
    )
}

plot_continuous_pair <- function(bin_df, x, y, breaks = 5, poly2 = FALSE, xlab = NULL, ylab = NULL,
                                 q = "Q", thresh = 0.10) {
  bl  <- bin_labels(bin_df[[x]], breaks = breaks)
  df2 <- bin_df %>% mutate(x_bin = bl$bin)

  prop_df  <- admix_prop_long(df2, x_bin, q = q, thresh = thresh)
  label_df <- admix_label_df(df2, x_bin, q = q, thresh = thresh)

  form <- if (poly2) y ~ poly(x, 2) else y ~ x

  p_scatter <-
    ggplot(df2, aes(x = .data[[x]], y = .data[[y]])) +
    geom_point(aes(colour = .data[[q]]), size = 2) +
    geom_smooth(method = "lm", colour = "black", formula = form) +
    geom_vline(xintercept = bl$max, linetype = "dashed", colour = "gray50") +
    theme_classic() +
    labs(x = xlab %||% x, y = ylab %||% y, colour = "Admixture score") +
    scale_x_continuous(labels = bl$lab, breaks = as.numeric(bl$lab), expand = c(0, 0)) +
    scale_colour_gradient(low = "skyblue2", high = "red2") +
    theme(
      axis.title = element_text(size = 18),
      axis.text  = element_text(size = 16),
      legend.title = element_text(size = 16),
      legend.text  = element_text(size = 16),
      legend.position = "top",
      legend.key.width = unit(1, "cm")
    )

  p_count <-
    ggplot(prop_df, aes(x = x_bin)) +
    geom_col(aes(y = prop, fill = admix_bin)) +
    geom_text(
      data = label_df,
      aes(x = x_bin, y = prop_hi + 0.1, label = paste0(n_hi, "/", n)),
      colour = "red2", size = 7, inherit.aes = FALSE
    ) +
    scale_fill_manual(values = col_pal) +
    scale_x_discrete(labels = bl$lab, expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0), limits = c(0, 1)) +
    theme_classic() +
    labs(x = paste0("Binned ", xlab %||% x), y = "P(admixture score > 0.10)", fill = "") +
    theme(
      legend.position = "none",
      axis.title = element_text(size = 18),
      axis.text  = element_text(size = 16)
    )

  list(df = df2, prop_df = prop_df, plt_scatter = p_scatter, plt_count = p_count, plt = plot_grid(p_scatter, p_count, ncol = 1, align = "v", rel_heights = c(0.6, 0.45)))
}

plot_categorical_pair <- function(bin_df, x, y, xlab = NULL, ylab = NULL, q = "Q", thresh = 0.10) {
  prop_df  <- admix_prop_long(bin_df, .data[[x]], q = q, thresh = thresh)
  label_df <- admix_label_df(bin_df, .data[[x]], q = q, thresh = thresh)

  p_box <-
    ggplot(bin_df, aes(x = .data[[x]], y = .data[[y]])) +
    geom_boxplot() +
    geom_jitter(aes(colour = .data[[q]]), size = 2, width = 0.2) +
    theme_classic() +
    labs(x = xlab %||% x, y = ylab %||% y, colour = "Admixture score") +
    scale_colour_gradient(low = "skyblue2", high = "red2") +
    theme(
      axis.title = element_text(size = 18),
      axis.text  = element_text(size = 16),
      legend.title = element_text(size = 16),
      legend.text  = element_text(size = 16),
      legend.position = "top",
      legend.key.width = unit(1, "cm")
    )

  p_count <-
    ggplot(prop_df, aes(x = .data[[x]])) +
    geom_col(aes(y = prop, fill = admix_bin)) +
    geom_text(
      data = label_df,
      aes(x = .data[[x]], y = prop_hi + 0.1, label = paste0(n_hi, "/", n)),
      colour = "red2", size = 7, inherit.aes = FALSE
    ) +
    theme_classic() +
    scale_y_continuous(expand = c(0, 0), limits = c(0, 1)) +
    scale_fill_manual(values = col_pal) +
    labs(x = xlab %||% x, y = "P(admixture score > 0.10)", fill = "") +
    theme(
      legend.position = "none",
      axis.title = element_text(size = 18),
      axis.text  = element_text(size = 16)
    )

  list(df = bin_df, prop_df = prop_df, plt = plot_grid(p_box, p_count, ncol = 1, align = "v", rel_heights = c(0.6, 0.45)))
}

q_df <- read_csv(here("analysis", "admixture", "outputs", "Q8.csv"))
bin_df <-
  mod_df %>%
  left_join(q_df %>% select(SampleID, Q), by = "SampleID") %>%
  mutate(Q = 1 - Q) %>%
  mutate(
    admix = factor(as.numeric(Q > 0.10), levels = c(0, 1), labels = c("Q ≤ 0.10", "Q > 0.10")),
    fire_frq = factor(
      fire_frq,
      levels = c("No Fire/Low Frequency", "Intermediate Frequency", "High Frequency"),
      labels = c("No Fire/Low", "Intermediate", "High")
    )
  )

breaks = 4
admix_csi   <- plot_continuous_pair(bin_df, x = "csi_past",  y = "Ho", breaks = breaks, poly2 = TRUE,
                                    xlab = "Paleoclimate stability", ylab = "Heterozygosity", q = "Q", thresh = 0.10)

admix_bio1  <- plot_continuous_pair(bin_df, x = "bio1",      y = "Ho", breaks = breaks, poly2 = FALSE,
                                    xlab = "Contemporary temperature", ylab = "Heterozygosity", q = "Q", thresh = 0.10)

admix_tdiff <- plot_continuous_pair(bin_df, x = "tmean_dif", y = "Ho", breaks = breaks, poly2 = FALSE,
                                    xlab = "Contemporary climate change", ylab = "Heterozygosity", q = "Q", thresh = 0.10)

admix_fire  <- plot_categorical_pair(bin_df, x = "fire_frq", y = "Ho",
                                     xlab = "Historical fire frequency", ylab = "Heterozygosity", q = "Q", thresh = 0.10)

admix_recent_high <- plot_categorical_pair(filter(bin_df, fire_frq == "High"), x = "fire_recent", y = "Ho",
                                           xlab = "Contemporary fire occurrence (Historical fire frequency = High)", ylab = "Heterozygosity",
                                           q = "Q", thresh = 0.10)

admix_fire_plt        <- admix_fire$plt
admix_csi_plt         <- admix_csi$plt
admix_bio1_plt        <- admix_bio1$plt
admix_tmean_dif_plt   <- admix_tdiff$plt
admix_fire_recent_plt <- admix_recent_high$plt

png(here(plotpath, "admix_all_plots.png"), height = 16, width = 20, res = 300, units = "in")
plot_grid(admix_csi_plt, admix_fire_plt, admix_tmean_dif_plt, admix_fire_recent_plt, admix_bio1_plt,
          nrow = 2, labels = LETTERS[1:5], label_size = 18)
dev.off()
```

```{r}
x = "fire_frq"
y = "Ho"
q = "Q"
thresh = 0.10

bin_df$fire_recent <- factor(bin_df$fire_recent, levels = c("Unburned", "Burned"))
prop_df  <- 
  bin_df %>%
    group_by(fire_frq, fire_recent) %>%
    summarise(
      n = dplyr::n(),
      n_hi = sum(.data[[q]] > thresh, na.rm = TRUE),
      prop_hi = n_hi / n,
      .groups = "drop"
    ) %>%
    transmute(
      fire_frq, fire_recent,
      `Q ≤ 0.10` = 1 - prop_hi,
      `Q > 0.10` = prop_hi
    ) %>%
    tidyr::pivot_longer(c(`Q ≤ 0.10`, `Q > 0.10`), names_to = "admix_bin", values_to = "prop") %>%
    mutate(admix_bin = factor(admix_bin, levels = c("Q ≤ 0.10", "Q > 0.10")))

label_df <- 
  bin_df %>%
  group_by(fire_frq, fire_recent) %>%
  summarise(
    n = dplyr::n(),
    n_hi = sum(.data[[q]] > thresh, na.rm = TRUE),
    prop_hi = n_hi / n,
    .groups = "drop"
  )

p_box <-
  ggplot(bin_df, aes(x = .data[[x]], y = .data[[y]], fill = factor(fire_recent))) +
  geom_boxplot(position = position_dodge(0.8), alpha = 0.7) +
  geom_jitter(
    aes(colour = .data[[q]]),
    position = position_jitterdodge(jitter.width = 0.4, dodge.width = 0.8),
    size = 2) +
  theme_classic() +
  labs(x = "Historical fire frequency", y = "Observed heterozygosity", fill = "Recent fire", colour = "Admixture score") +
  scale_colour_gradient(low = "skyblue2", high = "red2") +
  scale_fill_manual(values = c("white", "gray")) +
  # remove fill guide
  guides(col = "none") +
  theme(
    axis.title = element_text(size = 18),
    axis.text = element_text(size = 16),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 16),
    legend.position = "top",
    # Make legend wider
    legend.key.width = unit(1, "cm")
  )


p_count <-
  ggplot(prop_df, aes(x = fire_recent, y = prop, fill = admix_bin)) +
  geom_col() +
  geom_text(
    data = label_df,
    aes(x = fire_recent, y = prop_hi + 0.05, label = paste0(n_hi, "/", n)),
    colour = "red2", size = 6, inherit.aes = FALSE, vjust = 0.01
  ) +
  theme_classic() +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1)) +
  scale_fill_manual(values = col_pal) +
  facet_wrap(~ fire_frq, strip.position = "bottom") +
  labs(x = "", y = "P(admixture score > 0.10)", fill = "") +
  theme(
    legend.position = "none",
    axis.title = element_text(size = 18),
    axis.text.x = element_text(size = 16, angle = 30, hjust = 1),
    axis.text.y  = element_text(size = 16),
    strip.placement = "outside",
    strip.text = element_text(size = 18, color = "black"),
    strip.background = element_blank(),
    plot.margin = margin(5.5, 5.5, -15, 25)
  )

admix_fire_interaction_plt <- cowplot::plot_grid(p_box, p_count, ncol = 1, align = "v", rel_heights = c(0.6, 0.45))

admix_csi <- plot_continuous_pair(bin_df, x = "csi_past",  y = "Ho", breaks = 5, poly2 = TRUE,
                                    xlab = "Paleoclimate stability", ylab = "Heterozygosity", q = "Q", thresh = 0.10)
admix_csi_scatter <- admix_csi$plt_scatter + 
  theme(
    legend.position = "top", 
    legend.text = element_text(size = 18), 
    legend.title = element_text(size = 18), 
    legend.key.width = unit(1.5, "cm"))
admix_csi_scatter_noleg <- admix_csi_scatter + theme(legend.position = "none", plot.margin = margin(40, 0, 5.5, 0))
admix_csi_count <- admix_csi$plt_count
admix_csi_plt <- plot_grid(admix_csi_scatter_noleg, admix_csi_count, ncol = 1, align = "v", rel_heights = c(0.6, 0.45))
admix_legend <- cowplot::get_plot_component(admix_csi_scatter, "guide-box", return_all = TRUE)

group1 <- plot_grid(admix_csi_plt, admix_fire_interaction_plt, nrow = 1, labels = c("A", "B"), label_size = 18, align = "v")
plt <- cowplot::plot_grid(admix_legend[[4]], group1, nrow = 2, rel_heights = c(0.1, 1))

png(here(plotpath, "figure_admixture.png"), height = 8,5, width = 13, res = 300, units = "in")
plt
dev.off()

pdf(here(plotpath, "figure_admixture.pdf"), height = 8.5, width = 13)
plt
dev.off()
```

## Climate change interaction plot

```{r}
library(ggeffects)

# Refit model with raw data to get more interpretable interaction plots
mod_raw <- lm(formula(final_model), data = mod_sf_unique)
b <- coef(mod_raw)

# Johnson-Neyman test to find regions of significance for the interaction
jn <- johnson_neyman(mod_raw, pred = tmean_dif, modx = bio1)
jn_bounds <- jn$bounds

# Write out individuals with
cold_ind <- mod_sf %>% filter(bio1 < jn_bounds["Lower"]) %>% pull(SampleID)
hot_ind  <- mod_sf %>% filter(bio1 > jn_bounds["Higher"]) %>% pull(SampleID)
write_tsv(data.frame(SampleID = cold_ind), here(outpath, "cold_individuals.txt"), col_names = FALSE)
write_tsv(data.frame(SampleID = hot_ind), here(outpath, "hot_individuals.txt"), col_names = FALSE)

# Calculate switch point
# The switch point occurs where the marginal effect of tmean_dif on Ho = 0
# Taking the derivative with respect to tmean_dif:
#   dY / d(tmean_dif) = beta_tmean_dif + beta_(bio1:tmean_dif) * bio1
# Setting this equal to zero and solving for bio1:
#  0 = beta_tmean_dif + beta_(bio1:tmean_dif) * bio1
#   bio1* = -beta_tmean_dif / beta_(bio1:tmean_dif)
# This value of bio1 defines the point at which the effect of tmean_dif changes sign.
switch_point <- -b["tmean_dif"] / b["bio1:tmean_dif"]

# Get slopes
b <- coef(mod_raw)
V <- vcov(mod_raw)
bio_seq <- seq(
  min(mod_sf_unique$bio1, na.rm = TRUE),
  max(mod_sf_unique$bio1, na.rm = TRUE),
  length.out = 200
)

# Create data frame of bio1 values
slope_df <- 
  data.frame(bio1 = bio_seq) %>%
  # Compute slope: dY/d(tmean_dif) = β1 + β_interaction * bio1
  dplyr::mutate(
    slope = b["tmean_dif"] + b["bio1:tmean_dif"] * bio1
  )

# Compute standard error for linear combination
slope_df$se <- purrr::map_dbl(slope_df$bio1, function(z) {
  a <- c("tmean_dif" = 1, "bio1:tmean_dif" = z)
  sqrt(t(a) %*% V[names(a), names(a)] %*% a)
})

# 95% confidence intervals
crit <- qt(0.975, df = df.residual(mod_raw))
slope_df <- slope_df %>%
  dplyr::mutate(
    conf.low  = slope - crit * se,
    conf.high = slope + crit * se
  )

# Split bio1 into min, halfway between min and switch point, switch point, halfway between switch point and max, and max
# compute breakpoints ONCE
bio_min  <- min(mod_sf_unique$bio1, na.rm = TRUE)
bio_max  <- max(mod_sf_unique$bio1, na.rm = TRUE)

mid_low  <- (bio_min + switch_point) / 2
mid_high <- (switch_point + bio_max) / 2

# breaks <- c(bio_min, mid_low, switch_point, mid_high, bio_max)

# fmt <- function(x) paste0(round(x, 1), " °C")
# labels <- c(
#   paste0(fmt(bio_min), " – ", fmt(mid_low)),
#   paste0(fmt(mid_low), " – ", fmt(switch_point)),
#   paste0(fmt(switch_point), " – ", fmt(mid_high)),
#   paste0(fmt(mid_high), " – ", fmt(bio_max))
# )

breaks <- seq(bio_min, bio_max, length.out = 5)

fmt <- function(x) paste0(round(x, 1), " °C")
labels <- c(
  paste0(fmt(breaks[1]), " – ", fmt(breaks[2])),
  paste0(fmt(breaks[2]), " – ", fmt(breaks[3])),
  paste0(fmt(breaks[3]), " – ", fmt(breaks[4])),
  paste0(fmt(breaks[4]), " – ", fmt(breaks[5]))
)

interact_df <- mod_sf_unique %>%
  mutate(
    bio1_bin = cut(
      bio1,
      breaks = breaks,
      include.lowest = TRUE,
      labels = labels
    )
  )

slope_df <-
  slope_df %>%
  mutate(
    bio1_bin = cut(
      bio1,
      breaks = breaks,
      include.lowest = TRUE,
      labels = labels
    )
  )

# Get midpoint of each break (or you will have five intervals)
midbreaks <- (breaks[-1] + breaks[-length(breaks)]) / 2
preds <- ggeffects::ggpredict(
  mod_raw,
  terms = c(
    "tmean_dif [all]",
    paste0("bio1 [", paste(midbreaks, collapse = ", "), "]")
  ),
  data = mod_sf_unique
)
 
 #(#000249), Model Ultramarine (#0F4392), Hot Kiss (#FF4949) and Rich Red (#DD1717). The RGB and CMYK values of the c

#095786	(9,87,134)
#0090ab	(0,144,171)
#e6cea0	(230,206,160)
#db4646	(219,70,70)
#ae0000

# Bin raster by breaks and assign colors
bio1_binned <- 
  bio1 %>%
  project("epsg:3310") %>%
  mask(range_map) %>%
  as.data.frame(xy = TRUE) %>%
  mutate(
    bio1_bin = cut(
      `CHELSA_bio1_1981-2010_V.2.1`,
      breaks = breaks,
      include.lowest = TRUE,
      labels = labels
    )
  ) %>%
  drop_na()

pos_min <- jn_bounds["Higher"]
neg_max <- jn_bounds["Lower"]

cols <- c("navyblue", "skyblue1", "tomato1", "tomato4")
plt_slope <-
  ggplot(slope_df, aes(x = bio1, y = slope)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = bio1_bin), alpha = 0.2) +
  geom_line(linewidth = 1.2, aes(col = bio1_bin)) +
  geom_hline(yintercept = 0, col = "gray") +
  geom_vline(xintercept = breaks, linetype = "dashed", color = "gray") +
  # geom_vline(xintercept = switch_point, linetype = "solid", color = "darkgray", alpha = 0.8) +
  # annotate("text", x = switch_point, y = max(slope_df$slope), color = "darkgray", vjust = 0, hjust = 1.05, label = paste0(round(switch_point, 1), " °C")) +
  geom_vline(xintercept = neg_max, linetype = "solid", color = "navy", alpha = 0.8) +
  annotate("text", x = neg_max, y = max(slope_df$slope), color = "navy", vjust = 0, hjust = 1.05, label = paste0(round(neg_max, 2), " °C")) +
  geom_vline(xintercept = pos_min, linetype = "solid", color = "tomato3", alpha = 0.8) +
  annotate("text", x = pos_min, y = max(slope_df$slope), color = "tomato3", vjust = 0, hjust = 1.05, label = paste0(round(pos_min, 2), " °C")) +
  scale_color_manual(values = cols) +
  scale_fill_manual(values = cols) +
  scale_x_continuous(expand = c(0, 0)) +
  theme_classic() +
  labs(
    x = make_pretty_names("bio1"),
    y = "Effect of contemporary climate\nchange on heterozygosity"
  ) +
  # bltr
  theme(
    legend.position = "none", 
    plot.margin = margin(25, 10, 1, 10),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )

plt_interact <-
  ggplot(preds, aes(x = x, y = predicted, color = group, fill = group)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2, color = NA) +
  theme_classic() +
  labs(
    x = make_pretty_names("tmean_dif"),
    y = "Predicted heterozygosity",
  ) +
  theme(
    legend.position = "none",
    plot.margin = margin(10, 10, 10, 10),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  ) +
  scale_color_manual(values = cols) +
  scale_fill_manual(values = cols) 

plt_map <-
  ggplot(interact_df) +
  geom_sf(data = ca, fill = "white") +
  geom_raster(data = bio1_binned, aes(x = x, y = y, fill = bio1_bin), inherit.aes = FALSE, alpha = 0.7, show.legend = FALSE) +
  geom_sf(data = ca_proj, fill = NA, color = "black", linewidth = 0.3) +
  geom_sf(aes(fill = bio1_bin), size = 1.5, pch = 21) +
  scale_fill_manual(values = cols) +
  theme_void() +
  labs(fill = "Contemporary\ntemperature") +
  theme(
    legend.position = "inside", 
    legend.position.inside = c(0.75, 0.8), 
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11),
    plot.margin = margin(5, 5, 5, 5)
  )

plt_raw <-
  ggplot(interact_df, aes(x = tmean_dif, y = Ho, col = bio1_bin)) +
  geom_point(alpha = 0.8, size = 1) +
  geom_smooth(method = "lm", se = TRUE, col = "black") +
  theme_minimal() +
  labs(x = make_pretty_names("tmean_dif"), y = "Observed\nheterozygosity") +
  theme(legend.title = element_blank()) + 
  facet_wrap(~ bio1_bin, nrow = 1) +
  scale_color_manual(values = cols) +
  #scale_x_continuous(expand = c(0, 0)) +
  theme_classic() +
  theme(
    legend.position = "none", 
    panel.background = element_rect(fill = "white", color = "black", linewidth = 1),
    axis.line = element_blank(),
    strip.background = element_blank(),
    strip.text = element_text(size = 10),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10),
    #bltr
    plot.margin = margin(1, 10, 10, 20)

  ) + 
  stat_cor(aes(label = after_stat(r.label)), label.x = min(interact_df$tmean_dif), label.y = max(interact_df$Ho), size = 3.5, vjust = 1.5, method = "pearson")

b <- coef(lm(formula(final_model), data = mod_sf_unique))

slope_rast <- project(bio1, "epsg:3310")
# Mask out temperatures outside the range of our data
slope_rast[slope_rast < min(mod_sf_unique$bio1, na.rm = TRUE)] <- NA
slope_rast[slope_rast > max(mod_sf_unique$bio1, na.rm = TRUE)] <- NA
slope_rast <-  b["tmean_dif"] + b["bio1:tmean_dif"] * slope_rast
slope_rast <- mask(slope_rast, range_map)
slope_df <- as.data.frame(slope_rast, xy = TRUE)
names(slope_df)[3] <- "slope"

d <- max(abs(c(min(slope_df$slope), max(slope_df$slope))))
plt_map_slope <- 
  ggplot() +
  geom_raster(data = slope_df, aes(x, y, fill = slope)) +
  geom_sf(data = ca_proj, fill = NA, color = "black", linewidth = 0.3) +
  scale_fill_gradient2(
    low = "skyblue2",
    mid = "#f8f8c9",
    high = "tomato2",
    midpoint = 0,
    name = "Effect of contemporary\nclimate change on\nheterozygosity",
    # Set range
    limits = c(-d, d)
  ) +
  theme_void() +
  theme(
    legend.position = "inside", 
    legend.position.inside = c(0.82, 0.78), 
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11),
    legend.key.height = unit(0.4, "cm"),
    # Order: top, right, bottom, left
    plot.margin = margin(5, 20, 5, 5)
  )


group1 <- plot_grid(plt_map,  plt_map_slope, nrow = 1, labels = c("C", "D"))
group2 <- plot_grid(plt_slope, plt_raw, ncol = 1, labels = c("A", "B"), align = "v", rel_heights = c(0.5, 0.4))
plt <- plot_grid(group2, group1, ncol = 1, rel_heights = c(0.6, 0.4))

png(here(plotpath, "figure_climate_change_interaction.png"), width = 7, height = 9, units = "in", res = 300)
plt
dev.off()

pdf(here(plotpath, "figure_climate_change_interaction.pdf"), width = 9, height = 5.1)
plt
dev.off()

```
 
# Create supplementary tables
```{r}
# ============================================================
# Make Tables S3–S7 from final_model + emmeans contrasts
# Writes one TSV per table
# ============================================================

# ---- helpers ----
sig_stars <- function(p) {
  case_when(
    p < 0.001 ~ "***",
    p < 0.01  ~ "**",
    p < 0.05  ~ "*",
    p < 0.1   ~ "·",
    TRUE      ~ ""
  )
}

fmt_e <- function(x, digits = 2) formatC(x, format = "e", digits = digits)

# For Table S2 term prettifying (tailored to your exact model terms)
pretty_term_S2 <- function(term) {
  case_when(
    term == "(Intercept)" ~ "Intercept",
    term == "glacierInsideGlacier" ~ "Glacier (inside/outside)",
    term == "gHM" ~ "Human modification",
    term == "bio1" ~ "Contemporary temperature",
    term == "tmean_dif" ~ "Contemporary climate change",
    term == "bio1:tmean_dif" ~ "Contemporary temperature × Contemporary climate change",

    term == "poly(csi_past, 2, raw = TRUE)1" ~ "Paleoclimate stability (linear)",
    term == "poly(csi_past, 2, raw = TRUE)2" ~ "Paleoclimate stability (quadratic)",

    str_detect(term, "^evt") ~ paste0("Contemporary vegetation: ", str_replace(term, "^evt", "")),

    term == "fire_frqIntermediate Frequency" ~ "Historical fire frequency (Intermediate)",
    term == "fire_frqHigh Frequency" ~ "Historical fire frequency (High)",
    term == "fire_recentBurned" ~ "Contemporary fire occurrence (Burned)",
    term == "fire_frqIntermediate Frequency:fire_recentBurned" ~ "Fire frequency × Contemporary fire occurrence (Intermediate)",
    term == "fire_frqHigh Frequency:fire_recentBurned" ~ "Fire frequency × Contemporary fire occurrence (High)",

    term == "ai" ~ "Aridity index",
    TRUE ~ term
  )
}

# output folder (change if you want)
out_dir <- here::here("analysis", "genetic_diversity", "outputs", "tables")
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

# ============================================================
# Table S2 | Final model statistics
# ============================================================

term_order <- c(
  "(Intercept)",
  "poly(csi_past, 2, raw = TRUE)1",
  "poly(csi_past, 2, raw = TRUE)2",
  "bio1",
  "tmean_dif",
  "bio1:tmean_dif",
  "fire_frqIntermediate Frequency",
  "fire_frqHigh Frequency",
  "fire_recentBurned",
  "fire_frqIntermediate Frequency:fire_recentBurned",
  "fire_frqHigh Frequency:fire_recentBurned",
  paste0("evt", c("Agricultural", "Conifer", "Developed", "Exotic Herbaceous", "Exotic Tree-Shrub", "Grassland", "Hardwood", "Riparian", "Sparsely Vegetated")),
  "glacierInsideGlacier",
  "gHM",
  "ai"
)

tab_S3 <-
  broom::tidy(final_model) %>%
  slice(match(term_order, term)) %>%
  transmute(
    Predictor = pretty_term_S2(term),
    Estimate = fmt_e(estimate, 2),
    `Std. Error` = fmt_e(std.error, 2),
    `t-value` = round(statistic, 3),
    `p-value` = paste0(format.pval(p.value, digits = 3, eps = 0.001), " ", sig_stars(p.value)) %>% str_trim()
  ) 
readr::write_tsv(tab_S3, file.path(out_dir, "Table_S3_final_model_statistics.tsv"))

# ============================================================
# Table S4 | EMM contrasts for historical fire frequency
# ============================================================
emm_frq <- emmeans(final_model, ~ fire_frq, mode = "df.error")

tab_S4 <-
  contrast(emm_frq, method = "tukey", adjust = "tukey") %>%
  as.data.frame() %>%
  transmute(
    Contrast = contrast,
    Estimate = fmt_e(estimate, 2),
    `Std. Error` = fmt_e(SE, 2),
    df = df,
    `t-value` = round(t.ratio, 3),
    `p-value` = paste0(format.pval(p.value, digits = 3, eps = 0.001), " ", sig_stars(p.value)) %>% str_trim()
  )

readr::write_tsv(tab_S4, file.path(out_dir, "Table_S4_fire_frq_contrasts.tsv"))

# ============================================================
# Table S5 | EMM contrasts for contemporary fire within each fire_frq
# (No adjustment; 2 levels only)
# ============================================================
emm_recent_by_frq <- emmeans(final_model, ~ fire_recent | fire_frq, mode = "df.error")

tab_S5 <-
  contrast(emm_recent_by_frq, method = "pairwise") %>%
  as.data.frame() %>%
  transmute(
    `Historical fire frequency` = fire_frq,
    Contrast = contrast,
    Estimate = fmt_e(estimate, 2),
    `Std. Error` = fmt_e(SE, 2),
    df = df,
    `t-value` = round(t.ratio, 3),
    `p-value` = paste0(format.pval(p.value, digits = 3, eps = 0.001), " ", sig_stars(p.value)) %>% str_trim()
  )

readr::write_tsv(tab_S5, file.path(out_dir, "Table_S5_fire_recent_within_fire_frq.tsv"))

# ============================================================
# Table S6 | EMM contrasts for fire_frq within each fire_recent
# (You said you’re using tukey)
# ============================================================
emm_frq_by_recent <- emmeans(final_model, ~ fire_frq | fire_recent, mode = "df.error")

tab_S6 <-
  contrast(emm_frq_by_recent, method = "pairwise", adjust = "tukey") %>%
  as.data.frame() %>%
  transmute(
    `Contemporary fire occurrence` = fire_recent,
    Contrast = contrast,
    Estimate = fmt_e(estimate, 2),
    `Std. Error` = fmt_e(SE, 2),
    df = df,
    `t-value` = round(t.ratio, 3),
    `p-value` = paste0(format.pval(p.value, digits = 3, eps = 0.001), " ", sig_stars(p.value)) %>% str_trim()
  )

readr::write_tsv(tab_S6, file.path(out_dir, "Table_S6_fire_frq_within_fire_recent.tsv"))

# ============================================================
# Table S7 | EMM contrasts for contemporary vegetation (all pairs, Tukey)
# ============================================================
emm_evt <- emmeans(final_model, ~ evt, mode = "df.error")

tab_S7 <-
  contrast(emm_evt, method = "tukey", adjust = "tukey") %>%
  as.data.frame() %>%
  transmute(
    Contrast = contrast,
    Estimate = fmt_e(estimate, 2),
    `Std. Error` = fmt_e(SE, 2),
    df = df,
    `t-value` = round(t.ratio, 3),
    `p-value` = paste0(format.pval(p.value, digits = 4, eps = 0.0001), " ", sig_stars(p.value)) %>% str_trim()
  )

readr::write_tsv(tab_S7, file.path(out_dir, "Table_S7_evt_contrasts.tsv"))

# ============================================================
# Done: prints paths
# ============================================================
list.files(out_dir, pattern = "^Table_S[3-7].*\\.tsv$", full.names = TRUE)

```

# Variance partioning analysis

```{r}
# Calculate shared variance for a set of main-effect blocks
block_varpart_exact_shared <- function(model, blocks) {
  stopifnot(inherits(model, "lm"))

  get_R2 <- function(m) summary(m)$r.squared
  all_terms <- attr(terms(model), "term.labels")

  block_names <- names(blocks)
  block_terms <- blocks

  # R2 for a model using the union of terms in a subset of blocks
  r2_subset <- function(use_blocks) {
    rhs <- unique(unlist(block_terms[use_blocks]))
    rhs <- intersect(rhs, all_terms)

    m <- update(model, as.formula(paste(". ~", paste(rhs, collapse = " + "))))
    get_R2(m)
  }

  # Unique contribution of each block
  uniq <- purrr::map_dbl(block_names, \(b) {
    r2_subset(block_names) - r2_subset(setdiff(block_names, b))
  })

  # R2 of each block by itself
  full <- purrr::map_dbl(block_names, \(b) {
    r2_subset(b)
  })

  tibble(
    block = block_names,
    R2_unique = uniq,
    R2_full = full
  )
}

# -------------------------------------------------------------------
# 1. Fit reduced models to isolate interaction increments
# -------------------------------------------------------------------

# Main-effects-only model
no_interact_model <- update(final_model, . ~ . - fire_frq:fire_recent - bio1:tmean_dif)

# Main effects + fire interaction only
fire_interact_model <- update(no_interact_model, . ~ . + fire_frq:fire_recent)

# Main effects + climate interaction only
climate_interact_model <- update(no_interact_model, . ~ . + bio1:tmean_dif)

# R2 values
R2_main <- summary(no_interact_model)$r.squared
R2_full <- summary(final_model)$r.squared

R2_interaction_increment_fire <-
  summary(fire_interact_model)$r.squared - R2_main

R2_interaction_increment_climate <-
  summary(climate_interact_model)$r.squared - R2_main

# -------------------------------------------------------------------
# 2. Variance partitioning for main effects only
# -------------------------------------------------------------------

blocks_nointeract <- list(
  "fire_frq" = "fire_frq",
  "fire_recent" = "fire_recent",
  "evt" = "evt",
  "ai" = "ai",
  "bio1" = "bio1",
  "tmean_dif" = "tmean_dif",
  "poly(csi_past, 2, raw = TRUE)" = "poly(csi_past, 2, raw = TRUE)",
  "glacier" = "glacier",
  "gHM" = "gHM"
)

names(blocks_nointeract) <- make_pretty_names(names(blocks_nointeract), units = FALSE)

vp <- block_varpart_exact_shared(no_interact_model, blocks_nointeract)

# -------------------------------------------------------------------
# 3. Shared + unexplained variance
# -------------------------------------------------------------------

# Shared variance is whatever remains after:
# - unique main effects
# - unique fire interaction increment
# - unique climate interaction increment
shared_R2 <- R2_full -
  sum(vp$R2_unique) -
  R2_interaction_increment_fire -
  R2_interaction_increment_climate

unexplained_R2 <- 1 - R2_full

# Check sums
stopifnot(isTRUE(all.equal(
  sum(vp$R2_unique) +
    R2_interaction_increment_fire +
    R2_interaction_increment_climate +
    shared_R2 +
    unexplained_R2,
  1
)))

stopifnot(isTRUE(all.equal(
  sum(vp$R2_unique) +
    R2_interaction_increment_fire +
    R2_interaction_increment_climate +
    shared_R2,
  R2_full
)))

# -------------------------------------------------------------------
# 4. Plot
# -------------------------------------------------------------------

plot_df <- vp %>%
  transmute(
    component = block,
    pct_total = 100 * R2_unique,
    kind = "Unique"
  ) %>%
  bind_rows(
    tibble(
      component = "Historical fire × contemporary fire occurrence",
      pct_total = 100 * R2_interaction_increment_fire,
      kind = "Interaction"
    )
  ) %>%
  bind_rows(
    tibble(
      component = "Contemporary temperature × temperature change",
      pct_total = 100 * R2_interaction_increment_climate,
      kind = "Interaction"
    )
  ) %>%
  bind_rows(
    tibble(
      component = "Shared",
      pct_total = 100 * shared_R2,
      kind = "Shared"
    )
  ) %>%
  bind_rows(
    tibble(
      component = "Unexplained",
      pct_total = 100 * unexplained_R2,
      kind = "Unexplained"
    )
  ) %>%
  arrange(factor(kind, levels = c("Unique", "Interaction", "Shared", "Unexplained")), pct_total) %>%
  mutate(component = factor(component, levels = unique(component)))

plot_df %>%
  group_by(kind, grepl("Contemporary", component)) %>%
  summarize(sum_pct = sum(pct_total), .groups = "drop")

fill_vals <- plot_df %>%
  filter(!component %in% c("Shared", "Unexplained")) %>%
  pull(pct_total)

png(here(plotpath, "variance_partioning.png"),
    width = 8, height = 3.5, res = 300, units = "in")

ggplot(
  plot_df,
  aes(
    x = "Varpart",
    y = component,
    fill = ifelse(component %in% c("Shared", "Unexplained"), NA, pct_total)
  )
) +
  geom_tile(color = "white", linewidth = 0.8) +
  geom_text(
    aes(label = paste0(round(pct_total, 2), "%")),
    size = 4,
    color = "black"
  ) +
  scale_met_cont(values = fill_vals, extend_scale = 0.2) +
  labs(x = NULL, y = NULL, fill = "% Variance") +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid = element_blank(),
    legend.position = "right",
    axis.text.y = element_text(size = 10, color = "black")
  )

dev.off()

# -------------------------------------------------------------------
# 5. Output tables
# -------------------------------------------------------------------

vp %>%
  mutate(across(c(R2_unique, R2_full), ~ round(.x * 100, 2)))

tibble(
  component = c(
    "Historical fire × contemporary fire occurrence",
    "Contemporary temperature × temperature change",
    "Shared",
    "Unexplained"
  ),
  pct_total = round(100 * c(
    R2_interaction_increment_fire,
    R2_interaction_increment_climate,
    shared_R2,
    unexplained_R2
  ), 3)
)

# Variance explained by non-significant variables
vp %>% 
  filter(block %in% c("Aridity index", "Glacier (inside/outside)", "Human modification")) %>%
  pull(R2_unique) %>% 
  sum() * 100

# Total R2
round(summary(final_model)$r.squared * 100, 2)
# Adjusted R2
round(summary(final_model)$adj.r.squared * 100, 2)
```

# Plot maps of variables
```{r}
plot_points_map <- function(var, df) {
  df$fire_frq <- factor(
    df$fire_frq,
    levels = c("No Fire/Low Frequency","Intermediate Frequency","High Frequency"),
    labels = c("Low/No Fire","Intermediate","High")
  )
  plt_var <- 
    ggplot() +
    geom_sf(data = ca_proj) +
    theme_void() +
    theme(
      legend.position = "right",
      plot.title = element_text(hjust = 0.5),
      plot.margin = margin(2, 2, 2, 2),
      legend.title = element_text(size = 14),
      legend.text = element_text(size = 12)
    ) 
  if (is.numeric(df[[var]])){
    plt_var <- 
      plt_var <- plt_var + 
      geom_sf(data = df, aes_string(fill = var), size = 2, shape = 21) +
      scale_met_cont(df[[var]], aesthetic = "fill", extend_scale = 0.1)
  } else {
    if(length(unique(df[[var]])) <= 3){
      plt_var <- 
        plt_var + 
        geom_sf(data = df, aes_string(fill = var, shape = var), size = 2) +
        scale_shape_manual(values = c(21, 24, 22)[1:length(unique(df[[var]]))]) + 
        labs(shape = gsub(" ", "\n", make_pretty_names(var))) + 
        scale_fill_viridis_d(option = "turbo", begin = 0.25, end = 0.75, direction = -1)
    } else {
      plt_var <- 
        plt_var + 
        geom_sf(data = df, aes_string(fill = var), size = 2, shape = 21) + 
        scale_fill_viridis_d(option = "turbo", begin = 0.1, end = 0.9, direction = -1)
    }    
  }
  plt_var <- plt_var + labs(fill = gsub(" ", "\n", make_pretty_names(var)))
}

final_model_vars <- unique(final_model_vars)
png(here(plotpath, "final_model_variable_maps.png"), height = 15, width = 11, res = 300, unit = "in")
plt_list <- map(final_model_vars, ~(plot_points_map(.x, mod_sf_unique)))
cowplot::plot_grid(plotlist = plt_list, ncol = 2, align = "hv", labels = LETTERS[1:length(final_model_vars)], label_size = 14)
dev.off()
```