---
title: "SDM"
output:
  github_document:
    toc: TRUE
---
```{r}
library("rpaleoclim")
library("here")
library("tidyverse")
library("ggpubr")
library("terra")
library("sf")
library("gridExtra")
library("rpaleoclim")
source("general_functions.R")
source(here("analysis", "genetic_diversity", "genetic_diversity.R"))
source(here("analysis", "sdm", "paleovars.R"))
ca <- get_ca()
pdfpath <- here("analysis", "sdm", "pdfs")
```

```{r}
# get paleovars
env_list <- get_paleovars()
paleokey <- get_paleokey()

# Drop mis19
env_list <- env_list[names(env_list) != "mis19"]

# Assign the loaded and processed data to individual variables
list2env(env_list, envir = .GlobalEnv)

# Create example map (divide by 10 to convert to C)
stk <- c(env_list$cur$bio_1, env_list$lgm$bio_1)/10
names(stk) <- c("Current", "LGM")
env_df <-
  as.data.frame(stk, xy = TRUE) %>%
  mutate(Change = Current - LGM)
```

```{r}
plt2 <-
  ggplot(env_df) +
  geom_sf(data = ca) +
  geom_raster(aes(x = x, y = y, fill = Current)) +
  scale_fill_viridis_c(option = "viridis", direction = -1, limits = c(-15, 30)) +
  theme_void()

plt1 <-
  ggplot(env_df) +
  geom_sf(data = ca) +
  geom_raster(aes(x = x, y = y, fill = LGM)) +
  scale_fill_viridis_c(option = "viridis", direction = -1, limits = c(-15, 30)) +
  theme_void()

plt3 <-
  ggplot(env_df) +
  geom_sf(data = ca) +
  geom_raster(aes(x = x, y = y, fill = Change)) +
  labs(fill = "Change in temp\nsince the LGM") +
  scale_fill_viridis_c(option = "rocket") +
  theme_void()

grid.arrange(plt1, plt2, plt3, nrow = 1)
```


```{r}
# get coordinates
df <- get_coords()

# get genetic diversity measures
het <- get_het()
roh <- get_roh()

# combine together
df <- left_join(df, het) %>% left_join(roh)

# use data list to get vars
env_df_vals <- map(env_list, ~ data.frame(df, extract(.x, df[,c("x", "y")], ID = FALSE)))
diff_vars <- names(env_list) %>% setdiff("cur")
diff_vals <- map(diff_vars, ~ {
  diff <- cur - get(.x)
  return(data.frame(df, extract(diff, df[,c("x", "y")], ID = FALSE)))
})
env_df_vals <- c(env_df_vals, diff_vals)
names(env_df_vals) <- c(names(env_list), paste0("dif_", diff_vars))

# Join env_df_vals and het
env_df_vals <- map(env_df_vals, ~ left_join(.x, het))
```

# Correlation tests
```{r}
# check for strong correlations
# only strong correlation is with the late holocene/bio2
map(env_df_vals, ~ {
  cordf <- .x %>% dplyr::select(Ho, contains("bio")) %>% drop_na(Ho)
  strong <- cor(cordf, use = "complete.obs")["Ho", -1]
  which(abs(strong) > 0.5)
})
```
```{r}
cor_vals <- 
  map(env_df_vals, ~ {
    cordf <- .x %>% dplyr::select(Ho, contains("bio")) %>% drop_na(Ho)
    res <- cor(cordf, use = "complete.obs")["Ho", -1]
  }) 

map(cor_vals, ~ which(abs(.x) > 0.6))

cor_df <- 
  cor_vals %>% 
  bind_rows(.id = "timeperiod") %>% 
  pivot_longer(contains("bio")) %>%
  mutate(timeperiod = factor(timeperiod, levels = names(env_df_vals))) %>%
  mutate(bioclim = as.factor(as.numeric(gsub("bio_", "", name))))

pdf(here(pdfpath, "correlation_heatmap.pdf"), width = 5, height = 3.75)
ggplot(cor_df, aes(x = bioclim, y = timeperiod, fill = value)) +
  geom_tile() +
  labs(fill = expression(r[Ho])) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  theme_minimal()

ggplot(cor_df, aes(x = bioclim, y = timeperiod, fill = abs(value))) +
  geom_tile() +
  scale_fill_viridis_c(option = "rocket", direction = -1) +
  labs(fill = expression(r[Ho])) +
  theme_minimal()
dev.off()
```

```{r}
# confirmed no correlation with ROH:
map(env_df_vals, ~ {
  r <- cor(.x %>% dplyr::select(froh, contains("bio")) %>% drop_na(froh))["froh", -1]
  which(abs(r) > 0.6)
})
```

# Linear models

## Linear models with all variables (excluding correlated pairs)
```{r}

# Run linear models (dropping correlated variables for models, but adding them back in for visualization)
run_lm <- function(df, model_name){
  safe_collinear <- possibly(spatialEco::collinear)
  drop_vars <- safe_collinear(df %>% dplyr::select(contains("bio")), p = 0.6)
  uncorrelated <- names(df %>% dplyr::select(contains("bio"))) %>% setdiff(drop_vars)
  if (is.null(uncorrelated)) {
    uncorrelated <- names(df %>% dplyr::select(contains("bio")))
    correlated <- data.frame(corvars = NA, var = uncorrelated, model = model_name)
  } else {
    cormat <- cor(df %>% dplyr::select(contains("bio")))
    correlated <- 
      map(uncorrelated, ~ data.frame(corvars = names(which(abs(cormat[.x,]) > 0.5)), var = .x)) %>% 
      bind_rows() %>%
      mutate(model = model_name)
  }

  fHo <- formula(paste0("Ho ~", paste0(uncorrelated, collapse = " + ")))
  mod <- summary(lm(fHo, data=df))

  result <- 
    data.frame(mod$coefficients) %>% 
    mutate(model = model_name) %>%
    mutate(adj_rsqr = mod$adj.r.squared) %>%
    mutate(var = rownames(.)) %>% 
    filter(var != "(Intercept)") %>%
    left_join(correlated) 
  return(result)
}

lm_result <- imap(env_df_vals, \(x, idx) run_lm(x, idx)) %>% bind_rows()

sigvars <- 
  lm_result %>% 
  dplyr::select(-corvars) %>% 
  distinct() %>% 
  filter(`Pr...t..` < 0.05) %>%
  arrange(desc(Estimate))
```

```{r}
plts <-
  map(unique(lm_result$model), ~ {
    sig_df <-
    lm_result %>% 
    filter(model == .x) %>%
    filter(`Pr...t..` < 0.05)

    sig_env_df <- 
      env_df_vals[[.x]] %>% 
      pivot_longer(contains("bio"), names_to = "var") %>% 
      #include correlated vars
      filter(var %in% unique(sig_df$corvars))

    title <- paste0(.x, " (adj R^2 = ", round(unique(sig_df$adj_rsqr), 2), ")" )

    if (nrow(sig_env_df) == 0) return(NULL)
    plt <-
    ggplot(sig_env_df, aes(value, Ho)) +
      geom_point(alpha = 0.7) +
      facet_wrap(~var, scales = "free", ncol = 5, nrow = 3) +
      stat_regline_equation(
        aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
        formula = y ~ x
      ) +
      geom_smooth(method = "lm") +
      stat_cor(label.y = 0) +
      theme_classic() +
      ggtitle(title)

    plt
  })

pdf(here(pdfpath, "env_sigvars_with_corvars.pdf"), width = 20, height = 15)
walk(plts, print)
dev.off()

plts <-
  map(unique(lm_result$model), ~ {
    sig_df <- lm_result %>% filter(model == .x) %>% filter(`Pr...t..` < 0.05)
    sig_env_df <- env_df_vals[[.x]] %>% pivot_longer(contains("bio"), names_to = "var") %>% filter(var %in% unique(sig_df$var))
    title <- paste0(.x, " (adj R^2 = ", round(unique(sig_df$adj_rsqr), 2), ")" )

    if (nrow(sig_env_df) == 0) return(NULL)
    plt <-
    ggplot(sig_env_df, aes(value, Ho)) +
      geom_point(alpha = 0.7) +
      facet_wrap(~var, scales = "free", nrow = 1) +
      stat_regline_equation(
        aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
        formula = y ~ x
      ) +
      geom_smooth(method = "lm") +
      stat_cor(label.y = 0) +
      theme_classic() +
      ggtitle(title)

    plt
  })

```

## PC-based linear models

```{r}
# PCA of all variables 
env_wide <- 
  bind_rows(env_df_vals, .id = "timeperiod") %>% 
  dplyr::select("Ho", contains("bio"), "timeperiod") %>%
  drop_na("Ho") %>%
  # Filter out rows where all bio values are NA (only one sample for hs1)
  filter(!rowSums(is.na(dplyr::select(., contains("bio"))) == ncol(dplyr::select(., contains("bio"))))) %>%
  pivot_longer(contains("bio")) %>% 
  pivot_wider(names_from = "timeperiod", values_from = value) %>% 
  pivot_wider(names_from = "name", values_from = names(env_df_vals)) %>%
  drop_na(contains("bio"))

# Just change variables
pca_change <- prcomp(env_wide %>% dplyr::select(contains("dif")), scale = TRUE)
plot(pca_change)

# calculate when 90% of the variance is explained
varexp <- cumsum(pca_change$sdev^2) / sum(pca_change$sdev^2)
npcs <- which(varexp > 0.8)[1]
  
# Run lm with PCs
pcs_change <- data.frame(pca_change$x[,1:npcs])
mod <-summary(lm(env_wide$Ho ~ ., pcs_change))
mod_df <- - data.frame(mod$coefficients) 
adj_rsq <- mod$adj.r.squared

# Get loadings for significant PCs
sigpcs <- 
  mod %>% 
  filter(`Pr...t..` < 0.05) %>%
  mutate(var = row.names(.)) %>% 
  filter(var != "(Intercept)") %>%
  pull(var)

ggdf <-
  data.frame(pcs_change, Ho = env_wide$Ho) %>%
  pivot_longer(contains("PC")) %>%
  filter(name %in% sigpcs)

ggplot(ggdf, aes(value, Ho)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation(
    aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~~~")),
    formula = y ~ x
  ) +
  facet_wrap(~name, scales = "free") +
  stat_cor(label.y = 0) +
  theme_classic()


loadings <- data.frame(pca_change$rotation[,1:npcs], var = row.names(pca_change$rotation))

important_vars <-  map(sigpcs, ~{
  # Get loadings for the PC
  subloadings <- loadings %>% dplyr::select(all_of(.x), var)

  # Identify the variable with the highest loading
  bigvar <- loadings %>% arrange(desc(.x)) %>% head(1) %>% pull(var)

  # Get variables correlated with the bigvar
  bigcor <- 
    env_wide %>% 
    dplyr::select(contains("dif")) %>%
    cor(use = "complete.obs") %>% 
    .[bigvar,] 
    
  bigcorvars <- names(which(abs(bigcor) > 0.7))
}) %>% unlist() %>% unique()


important_df <- 
  env_wide %>% 
  dplyr::select(contains("dif")) %>%
  pivot_longer(contains("dif")) %>% 
  filter(name %in% important_vars) %>%
  mutate(splitname = name) %>%
  separate(splitname, into = c("prefix", "timeperiod", "bio", "bion"), sep = "_") %>%
  mutate(bio = paste0("bio_", bion)) 

ggplot(important_df, aes(value, Ho)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation(
    aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~~~")),
    formula = y ~ x
  ) +
  facet_wrap(~name, scales = "free") +
  stat_cor(label.y = 0) +
  theme_classic()

biokey <- get_biokey()
count_df <- left_join(important_df, biokey) %>% dplyr::select(-value) %>% distinct()
count(count_df, vartype)
count(count_df, bio) %>% arrange(desc(n))

# Just single variables (i.e., not change/dif variables)
pca_change <- prcomp(env_wide %>% dplyr::select(-contains("dif")), scale = TRUE)
plot(pca_change)
pcs_change <- data.frame(pca_change$x[,1:6])
summary(lm(env_wide$Ho ~ ., pcs_change))

loadings <- data.frame(pca_change$rotation[,1:6])
loadings %>% arrange(desc(PC1)) %>% head()

ggplot() +
  geom_point(aes(x = pcs_change$PC2, y =env_wide$Ho)) +
  geom_smooth(aes(x = pcs_change$PC2, y =env_wide$Ho), method = "lm") +
  stat_cor(label.y = 0) +
  theme_classic()


```

## Showing off BIO2
```{r}
mod_df <- env_df_vals[["dif_lh"]]
plt2 <-
  ggplot(mod_df) +
    geom_sf(data = ca) +
    geom_point(aes(x = x, y = y, color = bio_2), cex = 2) +
    labs(color = "Change\nin mean\ndiurnal\nrange") +
    scale_color_viridis_c(option = "magma", direction = -1) +
    theme_void() 

plt3 <-
  ggplot(mod_df) +
    geom_sf(data = ca) +
    geom_point(aes(x = x, y = y, color = Ho), cex = 2) +
    theme_void() +
    scale_color_viridis_c(option = "magma")

plt1 <- 
  ggplot(mod_df, aes(bio_2, Ho)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation(
    aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~~~")),
    formula = y ~ x
  ) +
  xlab("Change in mean diurnal range") +
  stat_cor(label.y = 0) +
  xlim(-7.5, NA) +
  theme_classic()

pdf(here(pdfpath, "bio2.pdf"), width = 15, height = 5)
grid.arrange(plt1, plt2, plt3, nrow = 1)
dev.off()
```

Notes on residuals:
- Lower genetic diversity than expected in very isolated sample
- Higher genetic diversity than expected in transverse ranges (contact/admixture zone?)
```{r}
# plot residuals
no_na <- mod_df %>% drop_na(Ho)
mod <- lm(Ho ~ bio_2, no_na)
no_na$residuals <- residuals(mod)

resid <- left_join(get_coords(sf = TRUE), no_na, by = "SampleID")

ggplot(resid) +
  geom_sf(data = ca) +
  geom_point(aes(x = x, y = y, fill = residuals), cex = 3, color = "#00000087", pch = 21) +
  theme_void() +
  scale_fill_gradient2(low = "#0086a7", mid = "beige", high = "#f9900d", midpoint = 0)
```


### Evaluating BIO2 in other time periods
```{r}
mega_df <- 
  bind_rows(env_df_vals, .id = "timeperiod") %>% 
  mutate(timeperiod = factor(timeperiod, levels = names(env_df_vals)))

mega_df2 <- 
  mega_df %>% 
  filter(grepl("dif", timeperiod)) %>%
  mutate(id = gsub("dif_", "", timeperiod)) %>%
  left_join(paleokey) %>%
  mutate(dif = paste0("Current to ", period, "\n(", time, ")")) %>%
  mutate(dif = factor(dif, levels = unique(dif)))

ggplot(mega_df2, aes(bio_2, Ho)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation(
    aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~~~")),
    formula = y ~ x
  ) +
  facet_wrap(~dif, scales = "free", nrow = 2) +
  xlab("Change in mean diurnal range") +
  stat_cor(label.y = 0) +
  theme_classic()
```

### Evaluating variables correlated with bio2
```{r}
wide_df <- 
  mega_df %>% 
  # drop MI19 because of missing values
  filter(!grepl("mis19", timeperiod)) %>%
  # note: pivot wider requires rows be uniquely identified (so sampleID is included here)
  dplyr::select(timeperiod, Ho, SampleID, contains("bio")) %>%
  pivot_wider(names_from = "timeperiod", values_from = contains("bio"))

bio2_cors <- cor(wide_df %>% dplyr::select(contains("bio")), use = "complete.obs")[,"bio_2_dif_lh"]
bio2_cors_names <- names(bio2_cors[abs(bio2_cors) > 0.6])
#bio1 - mean annual temperature
#bio6 - min temperature of coldest month (bio6)
#bio5 - max temperature of warmest month (bio5)
#bio7 - temperature annual range (bio5-bio6)
#bio2 - mean diurnal range (mean of monthly (max temp - min temp)
#bio12 - annual precipitation
#bio13 - precipitation of wettest month

gg_df <- 
  wide_df %>% 
  pivot_longer(all_of(bio2_cors_names)) %>%
  left_join(get_coords())

plt1 <- 
  ggplot(gg_df, aes(value, Ho)) +
    geom_point() +
    geom_smooth(method = "lm") +
    stat_regline_equation(
      aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~~~")),
      formula = y ~ x
    ) +
    facet_wrap(~name, scales = "free") +
    stat_cor(label.y = 0) +
    theme_classic() +
    xlab("value")

plt_list <- 
  gg_df %>%
  split(.$name) %>%
  map(~ {
    limit <- max(abs(.x[["value"]])) * c(-1, 1)
    ggplot(.x) +
        geom_sf(data = ca, fill = "darkgray") +
        labs(color = "value") +
        geom_point(aes(x = x, y = y, color = value), cex = 2) +
        scale_color_distiller(type = "div", limit = limit) +
        facet_wrap(~name) +
        theme_void()}
        )

plt3 <-
  ggplot(gg_df) +
    geom_sf(data = ca) +
    geom_point(aes(x = x, y = y, color = Ho)) +
    ggtitle("E. Genetic diversity") +
    theme_void() +
    scale_color_viridis_c(option = "magma")  

pdf(here(pdfpath, "correlated_vars.pdf"), width = 10, height = 10)
print(plt1)
do.call(grid.arrange, c(c(plt_list), nrow = 3))
print(plt3)
dev.off()
```

### Evaluationg all diff_vals
```{r}
# Calculate the difference between all unique combinations of env_list
coords <- get_coords()
combos <- data.frame(t(combn(names(env_list), 2)))
dif_vals <- map2(combos[,1], combos[,2], ~{
  diff <- env_list[[.x]] - env_list[[.y]]
  names(diff) <- unique(c(names(env_list[[.x]]), names(env_list[[.y]])))
  vals <- extract(diff, coords[,c("x", "y")], ID = FALSE)
  df <- data.frame(vals, comparison = paste0(.x, "_", .y), SampleID = coords$SampleID)
  return(df)
}) %>% bind_rows()

dif_ls <- left_join(dif_vals, get_het()) %>% group_by(comparison) %>% group_split()
dif_vals <- left_join(dif_vals, get_het()) %>% pivot_longer(contains("bio"), names_to = "bioclim")
cor_vals <- 
  map(dif_ls, ~ {
    cordf <- .x %>% dplyr::select(Ho, contains("bio")) %>% drop_na(Ho)
    res <- cor(cordf, use = "complete.obs")["Ho", -1]
  }) 
names(cor_vals) <- unique(dif_vals$comparison)

cor_df <- 
  cor_vals %>% 
  bind_rows(.id = "timeperiod") %>% 
  pivot_longer(contains("bio")) %>%
  mutate(bioclim = as.factor(as.numeric(gsub("bio_", "", name))))

# WARNING I THINK LABELS OF TIMEPERIODS/BIOCLIMS ARE OFF FIX THIS
pdf(here(pdfpath, "all_comparisons.pdf"), width = 13, height = 10)
ggplot(cor_df, aes(x = bioclim, y = timeperiod, fill = value)) +
  geom_tile() +
  labs(fill = expression(r[Ho])) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  theme_minimal()

ggplot(cor_df, aes(x = bioclim, y = timeperiod, fill = abs(value))) +
  geom_tile() +
  scale_fill_viridis_c(option = "rocket", direction = -1) +
  labs(fill = expression(r[Ho])) +
  theme_minimal()
dev.off()
```

### Evaluating variance in variables 
```{r}
var_df <- 
  mega_df %>%
  group_by(SampleID) %>%
  mutate(across(contains("bio"), ~ var(., na.rm =TRUE))) 

var_df_long <- 
  var_df %>%
  pivot_longer(contains("bio"))

pdf(here(pdfpath, "var.pdf"), width = 13, height = 10)
ggplot(var_df_long, aes(value, Ho)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation(
    aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~~~")),
    formula = y ~ x
  ) +
  facet_wrap(~name, scales = "free") +
  stat_cor(label.y = 0) +
  theme_classic() +
  xlab("value")
dev.off()
```

### big lazy model
```{r}
# NOTE: this way does not work well, results in low R^2 (PCS)
no_na_df <- wide_df %>% drop_na(contains("bio"), Ho)
pca <- prcomp(drop_na(dplyr::select(no_na_df, contains("bio"))))
pcs <- pca$x[,1:10]
#make a screeplot
screeplot(pca, type="lines", main="Screeplot of PCA")

mod_df_pca <- data.frame(Ho = no_na_df$Ho, pcs)
summary(lm(Ho ~ . , data = mod_df_pca))

all_cors <- cor(no_na_df %>% dplyr::select(contains("bio")))

# get variables with biggest PCA contributions to PC1, 2, and 3
# NOTE: this way does not work well, results in low R^2 (Variables selection based on PCs)
vars <- 
  as.data.frame(pca$rotation[,1:10]) %>%
  mutate(var = rownames(.)) %>%
  pivot_longer(-var, names_to = "PC", values_to = "loading") %>%
  group_by(PC) %>%
  filter(loading == max(loading)) %>%
  pull(var)
f <- formula(paste("Ho ~", paste(vars, collapse = " + ")))
summary(lm(f, data = no_na_df))
```

```{r}
# THIS DOESN"T WOKR
# get all cors and get absolute value (has to be absolute for clustering!)
all_cors <- abs(cor(dplyr::select(no_na_df, contains("bio"))))

# Convert the correlation matrix to a distance matrix (1 - |correlation|)
dist_matrix <- as.dist(1 - all_cors)

# Perform hierarchical clustering using the distance matrix
hc <- hclust(dist_matrix)

# Cut the dendrogram to create clusters, using a height that corresponds to the correlation threshold
# The height for cutting is 1 - threshold because the distance is 1 - correlation
clusters <- cutree(hc, h = 1 - 0.6)

# Group the variable names based on the clustering result
grouped_vars <- split(names(clusters), clusters)

# Get the var from each group with the strongest correlation with Ho
selected_vars <- 
  map_chr(grouped_vars, ~ {
    if (length(.x) == 1) return(.x)
    vars <- no_na_df %>% dplyr::select(all_of(.x))
    # find var with strongest correlation with Ho
    Ho_cor <- cor(vars, no_na_df$Ho, use = "complete.obs")
    var <- row.names(Ho_cor)[which.max(abs(Ho_cor))]
    return(var)
  }) %>% as.vector()

cor(no_na_df$bio_3_dif_ba, no_na_df$bio_7_dif_lig)
grouped_vars[map_lgl(grouped_vars, ~"bio_3_dif_ba" %in% .x)]
grouped_vars[map_lgl(grouped_vars, ~"bio_7_dif_lig" %in% .x)]

# check that all cors are below 0.6
check_df <- dplyr::select(no_na_df, all_of(selected_vars))
check_cor <- cor(check_df, use = "complete.obs")
diag(check_cor) <- 0
all(abs(check_cor) < 0.6)
max(check_cor)

# run model
mod_df_no_cor <- no_na_df %>% dplyr::select(all_of(selected_vars), Ho)
mod <- lm(Ho ~ ., data = mod_df_no_cor)
summary(mod)
df <- data.frame(predicted = mod$fitted.values, observed = mod_df_no_cor$Ho)

ggplot(df, aes(x = observed, y = predicted)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation(
    aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
    formula = y ~ x
  ) +
  stat_cor(label.y = 0) +
  theme_classic()
```

```{r}
# get all cors and get absolute value 
all_cors <- abs(cor(dplyr::select(no_na_df, contains("bio"))))

# for each variable get the variables it is correlated with
cor_vars <- 
  map(1:nrow(all_cors), ~ {
    cor_vars <- names(which(all_cors[.x,] > 0.6))
    return(cor_vars)
  })

#define a function to merge overlapping vectors
merge_overlaps <- function(lst) {
  for (i in seq_along(lst)) {
    for (j in seq_along(lst)) {
      if (i != j && any(lst[[i]] %in% lst[[j]])) {
        # Merge i and j and remove j from the list
        lst[[i]] <- unique(c(lst[[i]], lst[[j]]))
        lst <- lst[-j]
        # Recursively call the function until there are no overlaps left
        return(merge_overlaps(lst))
      }
    }
  }
  # Return the list if there are no overlaps left
  lst
}

# Use the function to merge overlaps
lst <- merge_overlaps(cor_vars)

summary(lm(Ho ~ bio_2_dif_lh + bio_8_dif_lig + bio_2_dif_lig, no_na_df))
```

```{r}
# Playing around with vars
mod_df <- data.frame(wide_df, var = var_df$bio_1) %>% drop_na(Ho, var, bio_2_dif_lh, bio_9_dif_lh, bio_19_dif_lig, bio_2_hs1, bio_3_dif_lh)
mod <- (lm(Ho ~ var + bio_2_dif_lh + bio_9_dif_lh + bio_19_dif_lig + bio_9_dif_lh + bio_2_hs1 + bio_3_dif_lh, data = mod_df))
summary(mod)
df <- data.frame(predicted = mod$fitted.values, observed = mod_df$Ho)

ggplot(df, aes(x = observed, y = predicted)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation(
    aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
    formula = y ~ x
  ) +
  stat_cor(label.y = 0) +
  theme_classic()
```

```{r}
mod <- lm(wide_df$Ho ~ ., data = select(wide_df, contains("bio")))
library(MuMIn)
library(glmnet)

no_na_df <- wide_df %>% drop_na()
# Fit Elastic Net model (alpha = 0.5 combines Ridge and Lasso)
x = as.matrix(select(no_na_df, contains("bio")))
y = no_na_df$Ho
elastic_net_model <- cv.glmnet(x, y, alpha = 0.5, standardize = TRUE)

# Plot the cross-validated mean squared error
plot(elastic_net_model)

# Get the best lambda (penalty parameter)
best_lambda <- elastic_net_model$lambda.min
print(best_lambda)

# Fit the model with the best lambda
final_model <- glmnet(x, y, alpha = 0.5, lambda = best_lambda)

# Get the coefficients of the final model
coef(final_model)

# Extract the indices of the selected variables
coefs <- coef(final_model) 
coefs <- as.vector(coef(final_model))
names(coefs) <- rownames(coef(final_model))
selected_vars <- names(which(coefs != 0))
selected_vars <- selected_vars[selected_vars != "(Intercept)"]
print(selected_vars)

summary(lm(wide_df$Ho ~ ., data = select(wide_df, all_of(selected_vars))))
```

### Testing for effects of population structure
```{r}
source(here("analysis", "admixture", "admixture.R"))
admix <- get_Q(K = 5)
head(admix)
coords <- get_coords()
mod_df <- left_join(mod_df, admix)

# bio2 is still significant after controling for clusters
summary(lm(Ho ~ cluster + bio_2, data = mod_df))
# nothing is significant once an interaction term is added but I think that is due to low power
summary(lm(Ho ~ cluster + bio_2, data = mod_df))

# look at heterozygosity by cluster and remove largest cluster
mod_df %>% group_by(cluster) %>% summarise(mean = mean(Ho, na.rm = TRUE)) %>% arrange(desc(mean))
# when you remove the southern pop with really high genetic diversity (cluster 3), the pattern still exists (albeit weaker)
plt1 <- ggplot(mod_df[mod_df$cluster != 3, ], aes(bio_2, Ho)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation(
    aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~~~")),
    formula = y ~ x
  ) +
  stat_cor(label.y = 0) +
  theme_classic()

plt2 <- ggplot(mod_df[mod_df$cluster != 3, ]) +
  geom_sf(data = ca) +
  geom_point(aes(x = x, y = y, color = bio_2), cex = 2) +
  labs(color = "bio_2") +
  theme_void() +
  scale_color_viridis_c(option = "magma", direction = -1)

plt3 <- ggplot(mod_df[mod_df$cluster != 3, ]) +
  geom_sf(data = ca) +
  geom_point(aes(x = x, y = y, color = Ho), cex = 2) +
  theme_void() +
  scale_color_viridis_c(option = "magma")

grid.arrange(plt1, plt2, plt3, nrow = 1)
```

### Checking with bouzid data
```{r}
vcf <- vcfR::read.vcfR(here("data", "bouzid", "populations_r20.haplotypes.filtered_m70_randomSNP.vcf"))
coords <- read_table(here("data", "bouzid", "Scelop.coord"), col_names =FALSE) %>% rename(x = X1, y = X2)
dos <- algatr::vcf_to_dosage(vcf)
dos_to_het <- function(dos){
  dos[dos == 1] <- 1
  dos[dos != 1] <- 0
  het <- rowMeans(dos, na.rm = TRUE)
  return(het)
}
df <- data.frame(coords, Ho = dos_to_het(dos))

ggplot(df) +
  geom_sf(data = ca) +
  geom_point(aes(x = x, y = y, color = Ho)) +
  scale_color_viridis_c(option = "magma") +
  theme_void()

lh <- load_paleoclim(here("data", "env", "LH_v1_2_5m.zip"))
cur <- load_paleoclim(here("data", "env", "CHELSA_cur_V1_2B_r2_5m.zip"))
dif <- cur - lh
lh_vals <- extract(dif, df[,c("x", "y")], ID = FALSE)

mod_df <- data.frame(df, lh_vals)

ggplot(mod_df, aes(bio_2, Ho)) +
    geom_point(alpha = 0.7) +
    stat_regline_equation(
      aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
      formula = y ~ x
    ) +
    geom_smooth(method = "lm") +
    stat_cor(label.y = 0) +
    theme_classic() 

#bio5 (max temperature of warmest month)
#bio6 (min temperature of coldest month)
```

### Testing for current environment effects
```{r}
df <-
 mega_df %>% 
 filter(timeperiod %in% c("dif_lh", "cur"))%>% 
 dplyr::select("Ho", "bio_2", "timeperiod", "SampleID") %>% 
 pivot_wider(names_from = timeperiod, values_from = bio_2) %>% 
 drop_na()

summary(lm(Ho ~ cur + dif_lh, data = df))
```


### Looking at change over time
```{r}
# create start and end times (approx) for each period
paleokey_split <- 
  paleokey %>%
  filter(id != "mis19") %>%
  mutate(time = 
    case_when(
      period == "Current" ~ "1.979-2.013 ka", 
      .default = time
      )
    ) %>%
  separate(time, into = c("start_time", "end_time"), sep = "-", convert = TRUE, fill = "right") %>%
  mutate(across(c(start_time, end_time), ~ case_when(
    str_detect(.x, "ka") ~ as.numeric(str_remove(str_remove(.x, "ka"), "ca.")) * 1000,
    str_detect(.x, "Ma") ~ as.numeric(str_remove(str_remove(.x, "Ma"), "ca.")) * 1000000,
    TRUE ~ as.numeric(.x)
  )))

# get values
change_df <- 
  mega_df %>%
  left_join(paleokey_split, by = c("timeperiod" ="id")) %>%
  pivot_longer(c(start_time, end_time), names_to = "start_end", values_to = "time") %>%
  drop_na(time)

mean_change <-
  change_df %>%
  group_by(time) %>%
  summarize(bio_2 = mean(bio_2, na.rm =TRUE)) %>%
  drop_na(time)

ggplot() +
  geom_line(data = change_df, aes(x = time, y = bio_2, group = SampleID, col = "Samples"), alpha = 0.3) +
  geom_line(data = mean_change, aes(x = time, y = bio_2, col = "Mean"), lwd = 2) +
  scale_color_manual(values = c("#007c8f", "darkgray")) + 
  labs(col = "") +
  scale_x_log10()  +
  theme_classic()
```


# SDM

BIOMOD: https://biomodhub.github.io/biomod2/articles/examples_1_mainFunctions.html

```{r}
library(countrycode)
library(CoordinateCleaner)
library(rgbif)
library(sp)
library(here)
library(raster)
library(dismo)
library(tidyverse)
library(viridis)
library(biomod2)
library(sf)
library(terra)
library(ggpubr)
# devtools::install_github('adamlilith/enmSdm', dependencies=TRUE)
library(enmSdm)
library(RStoolbox)
source(here("analysis", "sdm", "bias.R"))
source(here("analysis", "sdm", "sdm.R"))
source(here("analysis", "sdm", "get_coords.R"))
source(here("analysis", "sdm", "paleovars.R"))
source(here("general_functions.R"))

# set maxent path
# put maxent.jar in your working directory (what you get when you run getwd()) or set manually: https://rdrr.io/cran/biomod2/man/BIOMOD_ModelingOptions.html
Sys.setenv(JAVA_HOME='/usr/lib/jvm/java-11-openjdk-amd64')

# get env data
env_list <- get_paleovars()
env_list_agg <- map(env_list, ~aggregate(.x, 2))
cur <- env_list[["cur"]]
lh <- env_list[["lh"]]

# pca
selected_vars <- pc_var_selection(cur)
```

```{r}
# get coordinates
occ_coords <- get_occ_coords("Sceloporus occidentalis", data_source = "both", occ_limit = 10000, cache = TRUE)

# get or create bias background layer
bias_class <- get_bias("Sceloporus occidentalis", rank = "class", limit = 10000, n = 10000, cache = TRUE)

envlayers <- cur[[selected_vars]]
envlayers <- rast(here("data", "env", "CLEANED_ENVDATA_NOCOR.tif"))

# run SDM
sdm_mod <- 
  sdm(
    coordinates = occ_coords,
    envlayers = envlayers, 
    biasdat = bias_class, 
    models = "RF", 
    BIOMOD_name = "BIOMOD", 
    tune = FALSE
    )

sdm_raster <- sdm_mod$raster
#writeRaster(sdm_raster, here("analysis", "sdm", "outputs", "sdm.tif"))
sdm_raster <- rast(here("analysis", "sdm", "outputs", "sdm.tif"))

cutoff <- 
  sdm_mod$eval %>% 
  filter(metric.eval == "TSS") %>% 
  pull(cutoff) 
  
bin_raster <- sdm_raster
bin_raster[bin_raster > cutoff/1000] <- 1
bin_raster[bin_raster <= cutoff/1000] <- 0
plot(bin_raster)

# hindcast SDMs
hindcast_sdms <- 
  map(
    env_list[c("lig", "lgm", "hs1", "ba", "yds", "eh", "mh", "lh", "cur")],
   ~hindcast(sdm_mod$myBiomodModelOut, new_env = .x[[selected_vars]])
   )
hindcast_sdms <- rast(hindcast_sdms)
writeRaster(hindcast_sdms, here("outputs", "hindcast_sdms.tif"))

# calculate change and stability
# TODO: PLAY AROUND WITH TIMEPERIOD USED TO CALCULATE DIF_SDM AND STABILITY
cur_sdm <- sdm_mod$raster 
dif_sdm <- cur_sdm - hindcast_sdms[["lh"]]

hindcast_subset <- subset(hindcast_sdms, c("hs1", "ba", "yds", "eh", "mh", "lh", "cur"))
stability <- app(hindcast_subset, mean, na.rm =TRUE)
variability <- app(hindcast_subset, sd, na.rm =TRUE)
plot(stability, col = mako(100))
plot(variability, col = mako(100))
plot(dif_sdm, col = viridis(100))

trend <- global(hindcast_subset, mean, na.rm = TRUE)
trend <- data.frame(mean = trend, tp = rownames(trend)) %>% mutate(tp = factor(tp, levels = names(hindcast_subset)))
ggplot(trend) + geom_line(aes(x = tp, y = mean, group = 1))

# compare to genetic diversity
het <- get_het()
roh <- get_roh()
coords <- get_coords(sf = TRUE) %>% left_join(het) %>% left_join(roh)
change_vals <- extract(dif_sdm, coords, ID = FALSE)
stability_vals <- extract(stability, coords, ID = FALSE)
cur_vals <- extract(cur_sdm, coords, ID = FALSE)
# df <- data.frame(cur = cur_vals, coords)
df <- data.frame(change = change_vals[,1], stability = stability_vals[,1],  cur = cur_vals[,1], coords)
ggplot(df, aes(x = change, y = Ho)) +
  geom_point() +
  stat_regline_equation(
      aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
      formula = y ~ x
    ) +
  geom_smooth(method = "lm") +
  stat_cor(label.y = 0) +
  theme_classic()

ggplot(df, aes(x = stability, y = Ho)) +
  geom_point() +
  stat_regline_equation(
      aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
      formula = y ~ x
    ) +
  geom_smooth(method = "lm") +
  stat_cor(label.y = 0) +
  theme_classic()

ggplot(df, aes(x = cur, y = Ho)) +
  geom_point() +
  stat_regline_equation(
      aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
      formula = y ~ x
    ) +
  geom_smooth(method = "lm") +
  stat_cor(label.y = 0) +
  theme_classic()

ggplot(df, aes(x = change, y = froh0)) +
  geom_point() +
  stat_regline_equation(
      aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
      formula = y ~ x
    ) +
  geom_smooth(method = "lm") +
  stat_cor(label.y = 0) +
  theme_classic()

ggplot(df, aes(x = cur, y = froh0)) +
  geom_point() +
  stat_regline_equation(
      aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
      formula = y ~ x
    ) +
  geom_smooth(method = "lm") +
  stat_cor(label.y = 0) +
  theme_classic()

# grouped
df_group <- 
  df %>% 
  mutate(stability = stability > 3) %>%
  drop_na(stability)

# play around with this:
ggplot(df_group) +
  geom_boxplot(aes(x = stability, y = froh0))
```

```{r}
# presentation plots
ggplot(as.data.frame(sdm_raster, xy = TRUE)) +
  geom_sf(data = ca, fill = "darkgray") +
  geom_raster(aes(x = x, y = y, fill = .data[[names(sdm_raster)[1]]])) +
  scale_fill_viridis_c(option = "mako") +
  labs(fill = "Suitability") +
  theme_void()

ggplot(as.data.frame(stability, xy = TRUE)) +
  geom_sf(data = ca, fill = "darkgray") +
  geom_raster(aes(x = x, y = y, fill = .data[[names(stability)[1]]])) +
  scale_fill_viridis_c(option = "mako", limits =c(0,1)) +
  labs(fill = "Stability") +
  theme_void()
```

```{r}
# is stability correlated with change in bio2
dif_bio2 <- env_list[["cur"]][["bio_2"]] -  env_list[["lh"]][["bio_2"]]
dif_vals <- extract(dif_bio2, coords, ID = FALSE)
df$change_bio2 <- dif_vals[,1]

# check correlations with Ho
cor(df$change_bio2, df$Ho, use = "pairwise.complete.obs")
cor(df$stability, df$Ho, use = "pairwise.complete.obs")

# correlated, but not strongly
cor(df$change_bio2, df$stability, use = "pairwise.complete.obs")

# check for statistical suppression/whether both variables are significant when both are included
mod1 <- lm(Ho ~ change_bio2, data = df)
summary(mod1)
mod2 <- lm(Ho ~ stability, data = df)
summary(mod2)
mod3 <- lm(Ho ~ stability + change_bio2, data = df)
summary(mod3)
# stability is no longer significant when bio2 is included
```

```{r}
sdm_vals <- extract(hindcast_sdms, coords, ID = FALSE)
df <- 
  data.frame(sdm_vals, coords) %>% 
  pivot_longer(names(sdm_vals)) %>%
  mutate(name = factor(name, levels = names(sdm_vals)))

ggplot(df, aes(x = value, y = Ho)) +
  geom_point() +
  facet_wrap(~name) +
  stat_regline_equation(
      aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
      formula = y ~ x
    ) +
  geom_smooth(method = "lm") +
  stat_cor(label.y = 0) +
  theme_classic()

ggplot(filter(df, name %in% c("ba", "eh", "mh", "lh", "cur")), aes(x = value, y = Ho)) +
  geom_point() +
  facet_wrap(~name) +
  stat_regline_equation(
      aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
      formula = y ~ x
    ) +
  geom_smooth(method = "lm") +
  stat_cor(label.y = 0) +
  theme_classic()

ggplot(filter(df, name %in% c("lh", "cur")), aes(x = value, y = froh0)) +
  geom_point() +
  facet_wrap(~name) +
  stat_regline_equation(
      aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
      formula = y ~ x
    ) +
  geom_smooth(method = "lm") +
  stat_cor(label.y = 0) +
  theme_classic()
```

```{r}
# is sdm current value correlated with change in bio2
df <- coords
dif_bio2 <-env_list[["cur"]][["bio_2"]] -  env_list[["lh"]][["bio_2"]]
df$change_bio2 <-  extract(dif_bio2, coords, ID = FALSE)[,1]
df$sdm_vals <- extract(hindcast_sdms[["cur"]], coords, ID = FALSE)[,1]

# confirm correlations with Ho
cor(df$change_bio2, df$Ho, use = "pairwise.complete.obs")
cor(df$sdm_vals, df$Ho, use = "pairwise.complete.obs")

# correlated, but not strongly (-0.42)
cor(df$change_bio2, df$sdm_vals, use = "pairwise.complete.obs")

# check for statistical suppression/whether both variables are significant when both are included
mod1 <- lm(Ho ~ change_bio2, data = df)
summary(mod1)
mod2 <- lm(Ho ~ sdm_vals, data = df)
summary(mod2)
mod3 <- lm(Ho ~ sdm_vals + change_bio2, data = df)
summary(mod3)
# sdm value is not significant when bio2 is included or by itself
```


```{r}
# is sdm ba value correlated with change in bio2 (noticed a correlation)
df <- coords
dif_bio2 <-env_list[["cur"]][["bio_2"]] -  env_list[["lh"]][["bio_2"]]
df$change_bio2 <-  extract(dif_bio2, coords, ID = FALSE)[,1]
df$sdm_vals <- extract(hindcast_sdms[["cur"]], coords, ID = FALSE)[,1]

# confirm correlations with Ho
cor(df$change_bio2, df$Ho, use = "pairwise.complete.obs")
cor(df$sdm_vals, df$Ho, use = "pairwise.complete.obs")

# correlated, but not strongly (-0.42)
cor(df$change_bio2, df$sdm_vals, use = "pairwise.complete.obs")

# check for statistical suppression/whether both variables are significant when both are included
mod1 <- lm(Ho ~ change_bio2, data = df)
summary(mod1)
mod2 <- lm(Ho ~ sdm_vals, data = df)
summary(mod2)
mod3 <- lm(Ho ~ sdm_vals + change_bio2, data = df)
summary(mod3)
# sdm value is close to significant when bio2 is included
```

# Recent climate
Download PRISM max temperature data from historical data: https://prism.oregonstate.edu/historical/
```{r}
caladapt <- map(list.files(here("data", "env", "caladapt"), pattern = ".tif", full.names = TRUE), rast) %>% rast()

names(caladapt)

caladapt1950 <- caladapt[["livneh_year_Tmax_clip_15Oct2014.1950"]]
caladapt2013 <- caladapt[["livneh_year_Tmax_clip_15Oct2014.2013"]]
dif <- caladapt2013 - caladapt1950
plot(dif)
```

```{r}
# Function to calculate yearly average from monthly files
coords <- get_coords(sf = TRUE)
year = 1895
calculate_yearly_avg <- function(year  = 1895, type = "historical", output = "df") {
  if (type == "historical") directory_path <- here::here("data", "env", "prism", "PRISM_tmax_stable_4kmM3_189501_198012_bil")
  if (type == "current") directory_path <- here::here("data", "env", "prism", "PRISM_tmax_stable_4kmM3_198101_202308_bil")
  monthly_files <- list.files(directory_path, pattern = "*.bil.bil$", full.names = TRUE)
  monthly_files <- monthly_files[grepl(paste0("PRISM_tmax_stable_4kmM3_", year), monthly_files)]
  if (length(monthly_files) > 0) {
    rasters <- purrr::map(monthly_files, terra::rast)
    rasters <- terra::rast(rasters)
    if (output == "raster"){
      rasters <- mean(rasters, na.rm = TRUE)
      rm(rasters)
      gc()
    } else {
      vals <- terra::extract(rasters, st_transform(coords, st_crs(rasters)))
      annual_avg <- rowMeans(vals, na.rm = TRUE)
      return(data.frame(year = year, tmax = annual_avg, SampleID = coords$SampleID))
      rm(rasters)
      gc()
    }
    
  } else {
    return(NULL)
  }
}

# Loop over the years of interest
years <- 1895:1925 # Update this to the range of years you have data for
tmax_his <- map(1895:1925, ~calculate_yearly_avg(.x, type = "historical"), .progress =TRUE) %>% bind_rows()  
tmax_his <- tmax_his %>% group_by(SampleID) %>% summarise(tmax_his = mean(tmax, na.rm = TRUE))
tmax_cur <- map(1986:2016, ~calculate_yearly_avg(.x, type = "current"), .progress =TRUE) %>% bind_rows()  
tmax_cur <- tmax_cur %>% group_by(SampleID) %>% summarise(tmax_cur = mean(tmax, na.rm = TRUE))
tmax <- 
  left_join(tmax_his, tmax_cur) %>% 
  mutate(dif = tmax_cur - tmax_his) %>%
  right_join(coords) %>% 
  st_as_sf()

ggplot(tmax) +
  geom_sf(data = ca) +
  geom_sf(aes(color = dif)) +
  scale_color_viridis_c(option = "magma", direction = -1) +
  theme_void()

tmax_df <- left_join(tmax, get_het())

# Difference not signif by itself
summary(lm(Ho ~ dif, data =tmax_df))

# But once you add cur, it is significant and ahs a negative relationsihp
summary(lm(Ho ~ dif + tmax_cur, data = tmax_df))
# Close also for tmax_his, but note tmax_cur+tmax_his are super correlated
summary(lm(Ho ~ dif + tmax_his, data = tmax_df))

# Both tmax_cur and tmax_his are significant alone
summary(lm(Ho ~ tmax_his, data =tmax_df))
summary(lm(Ho ~ tmax_cur, data =tmax_df))

# Check with BIO2
source(here("analysis", "sdm", "paleovars.R"))
paleovars <- rast(here("data", "env", "paleoclim", "paleoclim.tif"))
bio2_dif <- paleovars$cur_bio_2 - paleovars$lh_bio_2
bio2_dif_vals <- terra::extract(bio2_dif, coords, ID = FALSE)[,1]
bio2_dif_df <- data.frame(SampleID = coords$SampleID, bio2_dif = bio2_dif_vals)

# dif is no longer significant once you account for bio2_dif, but tmax_his and tmax_cur still are significant
# ...likely just a elevation affect idk
tmax_df <- left_join(tmax_df, bio2_dif_df)
summary(lm(Ho ~ bio2_dif + dif + tmax_cur, data = tmax_df))
summary(lm(Ho ~ bio2_dif + dif + tmax_his, data = tmax_df))
# They aren't strongly correlated so it is likely not a suppression effect
cor(tmax_df$bio2_dif, tmax_df$dif)

plot(tmax_df$dif, tmax_df$Ho)
```

```{r}
# Compute the long-term average from 1895 to 1925
long_term_avg <- app(stack(annual_averages), mean)
```

```{r}
tmax_his <- map(1895:1925, ~calculate_yearly_avg(.x, type = "historical", output = "raster"), .progress =TRUE) %>% rast()  
tmax_cur <- map(1986:2016, ~calculate_yearly_avg(.x, type = "current"), .progress =TRUE) %>% bind_rows()  
```

# pastclimate

This data is not high enough resolution but keeping it here in case I want it in the future

```{r}
library("pastclim")
library("here")
set_data_path(path_to_nc = here("data", "env", "pastclim"))

# downloads all variables (annual) by default
download_dataset(dataset = "Beyer2020", annual = TRUE)
```

It is possible to also load a time series of rasters with the function region_series. In this case, the function returns a SpatRasterDataset, with each variable as a sub-dataset:
```{r}
ca <- get_ca()
climate_region <- region_series(
  time_bp = list(min = -120000, max = 0),
  bio_variables = c("bio05", "biome"),
  dataset = "Beyer2020",
  crop = ca
)

climate_region
```

Note that terra stores dates in years as AD, not BP. You can inspect the times in years BP with:
```{r}
time_bp(climate_region)
```

We can then plot the time series of a given variable (we relabel the plots to use years bp):
```{r}
terra::plot(climate_region$bio05[[1:5]], main = time_bp(climate_region))
```

To plot all climate variables for a given time step, we can slice the time series:
```{r}
slice_10k <- slice_region_series(climate_region, time_bp = -10000)
terra::plot(slice_10k)
```

Get ice sheets:
```{r}
terra::plot(climate_region$biome[[1]], main = time_bp(climate_region))
ice_id <- get_biome_classes("Beyer2020") %>% filter(category == "Land ice") %>% pull(id)

# check if there is any ice
prop_ice <- map(1:nlyr(climate_region$biome), ~global(climate_region$biome[[.x]] == ice_id, mean, na.rm = TRUE)) %>% bind_rows()
prop_ice
which_ice <- which(prop_ice$mean > 0)

ice <- climate_region$biome*0
ice[climate_region$biome == ice_id] <- 1

ice_df <- 
  as.data.frame(ice[[which_ice]], xy = TRUE, ID = FALSE) %>%
  dplyr::select(-ID)
names(ice_df) <- c("x", "y", time_bp(ice[[which_ice]]))
ice_long <- 
  pivot_longer(ice_df, -c("x", "y")) %>% 
  mutate(name = factor(name, levels = time_bp(ice))) %>%
  mutate(value = case_when(value == 1 ~ "ice", value == 0 ~ "no ice", .default = NA))

ggplot(ice_long) +
  geom_raster(aes(x = x, y = y, fill = value)) +
  facet_wrap(~name) +
  scale_fill_manual(values = c("#d5d5d5", "#009969"), na.value = NA) +
  theme_void()

# unfortunately the resolution is not very high for ca

# combine ice layers
ice_mean <- mean(ice[[which_ice]], na.rm = TRUE)
names(ice_mean) <- "ice"
plot(ice_mean)
```


# Look at future climate change

To download all bioclim data:
Generate envidatS3paths.txt file here: https://envicloud.wsl.ch/#/?prefix=chelsa%2Fchelsa_V2%2FGLOBAL%2Fclimatologies%2F
wget --no-host-directories --force-directories --input-file=envidatS3paths.txt

```{r}
# Get current climate
paleovars <- get_paleovars()
cur <- paleovars[["cur"]]
cur <- crop(cur, ca)
cur <- mask(cur, ca)

# Get BIO1 and convert
cur <- cur[["bio_1"]]/1000

# Get future climate
future2 <- rast(here("data", "env", "future", "CHELSA_bio1_2041-2070_ipsl-cm6a-lr_ssp585_V.2.1.tif"))
# Get future climate
future <- rast(here("data", "env", "future", "CHELSA_bio1_2041-2070_gfdl-esm4_ssp585_V.2.1.tif"))
future <- crop(future, ca)
future <- mask(future, ca)

future <- crop(future, ca)
future <- mask(future, ca)

# Convert future to same extent as cur
future <- resample(future, cur)

# Calculate climate change
cc <- future - cur

ggplot(as.data.frame(cc, xy = TRUE)) +
  geom_sf(data = ca) +
  geom_raster(aes(x = x, y = y, fill = .data[[names(cc)[1]]])) +
  labs(fill = "Future change\nin temperature")
  scale_fill_viridis_c(option = "magma", na.value = NA) +
  theme_void()
```


# FINAL LM MODELS

## Look at change in climate in regions that were glaciated
```{r}
# get current/lh climate
paleovars <- get_paleovars()
cur <- paleovars[["cur"]]
lh <- paleovars[["lh"]]
dif <- cur - lh
ice_resamp <- terra::resample(ice_mean, dif, method = "max")
dif_resamp <- terra::resample(dif, ice_mean)
ice_vect <- st_as_sf(as.polygons(ice_resamp))
ice_vect <- ice_vect[ice_vect$ice == 1, ]

df <- 
  cbind(as.data.frame(dif, ID = FALSE, xy =TRUE), as.data.frame(ice_resamp, ID = FALSE)) %>%
  pivot_longer(names(dif)) %>%
  mutate(ice = case_when(ice == 1 ~ "ice", ice == 0 ~ "no ice", .default = NA)) %>%
  drop_na(ice)

ggplot(filter(df, name %in% c("bio_1", "bio_2", "bio_5"))) +
  geom_boxplot(aes(x = ice, y = value, fill = ice)) +
  facet_wrap(~name, scales = "free")

df_resamp <- 
  cbind(as.data.frame(dif_resamp, ID = FALSE, xy =TRUE), as.data.frame(ice_mean, ID = FALSE)) %>%
  pivot_longer(names(dif)) %>%
  mutate(ice = case_when(ice == 1 ~ "ice", ice == 0 ~ "no ice", .default = NA)) %>%
  drop_na(ice)

ggplot(filter(df_resamp, name %in% c("bio_1", "bio_2", "bio_5"))) +
  geom_boxplot(aes(x = ice, y = value, fill = ice)) +
  facet_wrap(~name, scales = "free")

ggplot(filter(df_resamp, name == "bio_2")) +
  geom_raster(aes(x = x, y = y, fill = value)) +
  geom_sf(data = ice_vect, fill = NA, col = "black", lty = "dashed", lwd = 1) +
  scale_fill_viridis_c(option = "magma", na.value = NA) +
  theme_void()

```

## Adding Q-values
```{r}
# cluster = final cluster assignment for SampleID given max Q
# PopGroup = Q values for each cluster
mod_df <- env_df_vals[["dif_lh"]]
q <- read_csv(here("analysis", "admixture", "outputs", "Q6.csv"))
q <- 
  q %>%
  # only keep the pop group corresponding to the cluster
  mutate(PopGroup = as.numeric(gsub("V", "", PopGroup))) %>%
  filter(cluster == PopGroup) 

# plot residuals
no_na <- 
  mod_df %>% 
  drop_na(Ho) %>%
  left_join(q, by = "SampleID")

ggplot(no_na) +
  geom_sf(data = ca) +
  geom_point(aes(x = x, y = y, fill = Q), cex = 3, color = "#00000087", pch = 21) +
  theme_void() +
  scale_fill_viridis_c()

# Use the Q of the identified cluster (the max Q) to represent admixture (idea: lower max Q = more admixture)
mod_q <- lm(Ho ~ bio_2 + Q, no_na)

# Q is significant
summary(mod_q)

# Plot Ho vs Q
ggplot(no_na, aes(x = 1/Q, y = Ho)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation(
    aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~~~")),
    formula = y ~ x
  ) +
  stat_cor(label.y = 0) +
  theme_classic()

no_na$residuals <- residuals(mod_q)
ggplot(no_na) +
  geom_sf(data = ca) +
  geom_point(aes(x = x, y = y, fill = residuals), cex = 3, color = "#00000087", pch = 21) +
  theme_void() +
  scale_fill_gradient2(low = "#0086a7", mid = "beige", high = "#f9900d", midpoint = 0)

# create null model
no_na$null <- 1
mod_null <- lm(Ho ~ null, no_na)
no_na$residuals_null <- residuals(mod_null)
ggplot(no_na) +
  geom_sf(data = ca) +
  geom_point(aes(x = x, y = y, fill = residuals_null), cex = 3, color = "#00000087", pch = 21) +
  theme_void() +
  scale_fill_gradient2(low = "#0086a7", mid = "beige", high = "#f9900d", midpoint = 0)

mod_bio2 <- lm(Ho ~ bio_2, no_na)
no_na$residuals_bio2 <- residuals(mod_bio2)

mod_q <- lm(Ho ~ Q, no_na)
no_na$residuals_q <- residuals(mod_q)

mod_q_bio2 <- lm(Ho ~ Q + bio_2, no_na)
no_na$residuals_q_bio2 <- residuals(mod_q_bio2)

no_na$delta_residuals_q <- abs(no_na$residuals_bio2) - abs(no_na$residuals_q_bio2)
no_na$delta_residuals_bio2 <- abs(no_na$residuals_q) -  abs(no_na$residuals_q_bio2)

ggdf <-
  no_na %>%
  pivot_longer(c("delta_residuals_bio2", "delta_residuals_q")) 
ggplot(ggdf) +
  geom_sf(data = ca) +
  geom_point(aes(x = x, y = y, fill = value), cex = 3, color = "#00000087", pch = 21) +
  facet_wrap(~name) +
  theme_void() +
  scale_fill_gradient2(low = "#0086a7", mid = "beige", high = "#f9900d", midpoint = 0)

ggdf <-
  no_na %>%
  pivot_longer(c("residuals_bio2", "residuals_q", "residuals_q_bio2")) 
ggplot(ggdf) +
  geom_sf(data = ca) +
  geom_point(aes(x = x, y = y, fill = value), cex = 3, color = "#00000087", pch = 21) +
  facet_wrap(~name) +
  theme_void() +
  scale_fill_gradient2(low = "#0086a7", mid = "beige", high = "#f9900d", midpoint = 0)
```

## EVERYTHING
```{r}
elev <- get_dem()
coords <- get_coords(sf = TRUE)

# Climate data
paleovars <- get_paleovars()
bio2_dif <- paleovars$cur$bio_2 - paleovars$lh$bio_2
bio2_dif_vals <- terra::extract(bio2_dif, coords, ID = FALSE)[,1]
clim <- data.frame(SampleID = coords$SampleID, bio_2 = bio2_dif_vals)

# Q data
q <- read_csv(here("analysis", "admixture", "outputs", "Q6.csv"))
q <- 
  q %>%
  # only keep the pop group corresponding to the cluster
  mutate(PopGroup = as.numeric(gsub("V", "", PopGroup))) %>%
  filter(cluster == PopGroup) 

# Glacier data
lgm_glaciers <- st_read(here("data", "env", "Glaciotect_NA_lgm", "shp", "LGM_maxextent.shp"))
# Note: have to first intersect and then transform otherwise you get a node error
lgm_glaciers <- st_intersection(lgm_glaciers, st_transform(ca, crs = st_crs(lgm_glaciers)))
lgm_glaciers <- st_transform(lgm_glaciers, crs = st_crs(ca))
# Add a new column to coords indicating whether each point is inside the lgm shapefile
coords <- get_coords(sf = TRUE)
glacier <- 
  coords %>%
  mutate(glacier = lengths(st_within(coords, lgm_glaciers)) > 0) %>%
  mutate(glacier = ifelse(glacier, "Inside", "Outside")) %>%
  mutate(glacier = factor(glacier, levels = c("Inside", "Outside")))

# make model df
mod_df <- 
  get_coords() %>%
  left_join(get_het()) %>%
  left_join(glacier) %>%
  left_join(clim) %>% 
  left_join(q) %>%
  left_join(elev) %>%
  drop_na(Ho, glacier, Q, bio_2) 

mod_df_elev <- 
  mod_df %>%
  drop_na(elev)

mod <- lm(Ho ~ bio_2 + Q + glacier + elev, mod_df_elev)

# Confirm: elev is not significant after adding bio2 and based on AIC bio2 alone is about the same
summary(lm(Ho ~ elev, mod_df_elev))
summary(lm(Ho ~ bio_2 + elev, mod_df_elev))
AIC(lm(Ho ~ elev, mod_df_elev))
AIC(lm(Ho ~ bio_2 + elev, mod_df_elev))
AIC(lm(Ho ~ bio_2, mod_df_elev))

mod <- lm(Ho ~ bio_2 + Q + glacier, mod_df)
print(summary(mod)$adj.r.squared)
result <- summary(mod)$coefficients

# Plot predicted vs observed values
qplt1 <- 
  ggplot(mod_df) +
  geom_point(aes(x = predict(mod), y = Ho, col = Q)) +
  geom_abline(intercept = 0, slope = 1) +
  geom_smooth(aes(x = predict(mod), y = Ho), method = "lm") +
  theme_classic() +
  labs(x = "Predicted", y = "Observed") +
  scale_color_viridis_c(option = "viridis", end = 0.1, begin = 0.9)

gplt1 <- 
  ggplot(mod_df) +
  geom_point(aes(x = predict(mod), y = Ho, col = glacier)) +
  geom_smooth(aes(x = predict(mod), y = Ho), method = "lm") +
  theme_classic() +
  labs(x = "Predicted", y = "Observed") +
  scale_color_manual(values = c("skyblue2", "lightgray"))

bio2plt1 <-
  ggplot(mod_df) +
  geom_point(aes(x = predict(mod), y = Ho, col = bio_2)) +
  geom_smooth(aes(x = predict(mod), y = Ho), method = "lm") +
  geom_smooth(aes(x = predict(mod), y = Ho), method = "lm") +
  theme_classic() +
  labs(x = "Predicted", y = "Observed") +
  scale_color_viridis_c(option = "inferno", end = 0.1, begin = 0.9)

# Plot values across california
qplt2 <-
  ggplot(mod_df) +
  geom_sf(data = ca) +
  geom_point(aes(x = x, y = y, col = Q)) +
  theme_void() +
  scale_color_viridis_c(option = "viridis", end = 0.1, begin = 0.9)

gplt2 <-
  ggplot(mod_df) +
  geom_sf(data = ca) +
  geom_point(aes(x = x, y = y, col = glacier)) +
  theme_void() +
  scale_color_manual(values = c("skyblue2", "gray"))

bio2plt2 <-
  ggplot(mod_df) +
  geom_sf(data = ca) +
  geom_point(aes(x = x, y = y, col = bio_2)) +
  theme_void() +
  scale_color_viridis_c(option = "inferno", end = 0.1, begin = 0.9)

# Plot Ho vs variable
qplt3 <-
  ggplot(mod_df, aes(x = Q, y = Ho, col = Q)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_classic() +
  scale_color_viridis_c(option = "viridis", end = 0.1, begin = 0.9)


gplt3 <- 
  ggplot(mod_df) +
  geom_boxplot(aes(x = glacier, y = Ho, fill = glacier)) +
  geom_jitter(aes(x = glacier, y = Ho), alpha = 0.5) +
  theme_classic() +
  scale_fill_manual(values = c("skyblue2", "lightgray"))

bio2plt3 <-
  ggplot(mod_df, aes(x = bio_2, y = Ho, col = bio_2)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_classic() +
  scale_color_viridis_c(option = "inferno", end = 0.1, begin = 0.9)

# use cowplot to combine plots
pdf(here("analysis", "sdm", "pdfs", "linear_model_plot.pdf"), width = 13, height = 10)
cowplot::plot_grid(qplt2, gplt2, bio2plt2,  qplt3, gplt3, bio2plt3, qplt1, gplt1, bio2plt1, ncol = 3)
dev.off() 

# Plot Ho

mod_df <-
   mod_df %>% 
   mutate(explain =
    case_when(
      glacier == "Inside" ~ "Inside glaciated region",
      Q < 0.8 ~ "High admixture"
    )
   )

ggplot(mod_df) + 
  geom_sf(data = ca) +
  geom_point(data = filter(mod_df, is.na(explain)), aes(x = x, y = y, fill = Ho), col = "black", pch = 21, cex = 3) +
  geom_point(data = drop_na(mod_df, explain), aes(x = x, y = y, fill = Ho, col = explain), pch = 21, cex = 3, stroke = 1) +
  scale_fill_viridis_c(option = "inferno") +
  scale_color_manual(values = c("green", "skyblue2")) +
  theme_void()
```
