---
title: "SDM"
output:
  github_document:
    toc: TRUE
---
```{r}
library("rpaleoclim")
library("here")
library("tidyverse")
library("ggpubr")
library("terra")
library("sf")
library("gridExtra")
library("rpaleoclim")
source("general_functions.R")
source(here("analysis", "genetic_diversity", "genetic_diversity.R"))
ca <- get_ca()
```
```{r}
# get paleovars
env_list <- get_paleovars()

# Assign the loaded and processed data to individual variables
list2env(env_list, envir = .GlobalEnv)
```

# running linear models
```{r}
# get coordinates
df <- get_coords()

# get genetic diversity measures
het <- get_het()
roh <- get_roh()

# combine together
df <- left_join(df, het) %>% left_join(roh)

# use data list to get vars
env_df_vals <- map(env_list, ~ data.frame(df, extract(.x, df[,c("x", "y")], ID = FALSE)))
diff_vars <- names(env_list) %>% setdiff("cur")
diff_vals <- map(diff_vars, ~ {
  diff <- cur - get(.x)
  return(data.frame(df, extract(diff, df[,c("x", "y")], ID = FALSE)))
})

env_df_vals <- c(env_df_vals, diff_vals)
names(env_df_vals) <- c(paleovars, paste0("dif_", diff_vars))

# check for strong correlations
# only strong correlation is with the late holocene/bio2
map(env_df_vals, ~ {
  cordf <- .x %>% select(Ho, contains("bio")) %>% drop_na(Ho)
  strong <- cor(cordf, use = "complete.obs")["Ho", -1]
  which(abs(strong) > 0.5)
})

cor_vals <- 
  map(env_df_vals, ~ {
    cordf <- .x %>% select(Ho, contains("bio")) %>% drop_na(Ho)
    res <- cor(cordf, use = "complete.obs")["Ho", -1]
  }) 

map(cor_vals, ~ which(abs(.x) > 0.6))

cor_df <- 
  cor_vals %>% 
  bind_rows(.id = "timeperiod") %>% 
  pivot_longer(contains("bio")) %>%
  mutate(timeperiod = factor(timeperiod, levels = names(cor_vals))) %>%
  mutate(bioclim = as.factor(as.numeric(gsub("bio_", "", name))))

pdf("correlation_heatmap.pdf", width = 5, height = 3.75)
ggplot(cor_df, aes(x = bioclim, y = timeperiod, fill = value)) +
  geom_tile() +
  labs(fill = expression(r[Ho])) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  theme_minimal()

ggplot(cor_df, aes(x = bioclim, y = timeperiod, fill = abs(value))) +
  geom_tile() +
  scale_fill_viridis_c(option = "rocket", direction = -1) +
  labs(fill = expression(r[Ho])) +
  theme_minimal()
dev.off()
```

```{r}
# confirmed no correlation with ROH:
map(env_df_vals, ~ {
  r <- cor(.x %>% select(froh, contains("bio")) %>% drop_na(froh))["froh", -1]
  which(abs(r) > 0.6)
})
```

```{r}
run_lm <- function(df, model_name){
  safe_collinear <- possibly(spatialEco::collinear)
  drop_vars <- safe_collinear(df %>% select(contains("bio")), p = 0.6)
  uncorrelated <- names(df %>% select(contains("bio"))) %>% setdiff(drop_vars)
  if (is.null(uncorrelated)) {
    uncorrelated <- names(df %>% select(contains("bio")))
    correlated <- data.frame(corvars = NA, var = uncorrelated, model = model_name)
  } else {
    cormat <- cor(df %>% select(contains("bio")))
    correlated <- 
      map(uncorrelated, ~ data.frame(corvars = names(which(abs(cormat[.x,]) > 0.5)), var = .x)) %>% 
      bind_rows() %>%
      mutate(model = model_name)
  }

  fHo <- formula(paste0("Ho ~", paste0(uncorrelated, collapse = " + ")))
  mod <- summary(lm(fHo, data=df))

  result <- 
    data.frame(mod$coefficients) %>% 
    mutate(model = model_name) %>%
    mutate(adj_rsqr = mod$adj.r.squared) %>%
    mutate(var = rownames(.)) %>% 
    filter(var != "(Intercept)") %>%
    left_join(correlated) 
  return(result)
}

lm_result <- imap(env_df_vals, \(x, idx) run_lm(x, idx)) %>% bind_rows()

sigvars <- lm_result %>% select(-corvars) %>% distinct() %>% filter(`Pr...t..` < 0.05) %>% arrange(desc(Estimate))
```

```{r}
plts <-
map(unique(lm_result$model), ~ {
  sig_df <-
   lm_result %>% 
   filter(model == .x) %>%
   filter(`Pr...t..` < 0.05)

  sig_env_df <- 
    env_df_vals[[.x]] %>% 
    pivot_longer(contains("bio"), names_to = "var") %>% 
    #include correlated vars
    filter(var %in% unique(sig_df$corvars))

  title <- paste0(.x, " (adj R^2 = ", round(unique(sig_df$adj_rsqr), 2), ")" )

  if (nrow(sig_env_df) == 0) return(NULL)
  plt <-
  ggplot(sig_env_df, aes(value, Ho)) +
    geom_point(alpha = 0.7) +
    facet_wrap(~var, scales = "free", ncol = 5, nrow = 3) +
    stat_regline_equation(
      aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
      formula = y ~ x
    ) +
    geom_smooth(method = "lm") +
    stat_cor(label.y = 0) +
    theme_classic() +
    ggtitle(title)

  plt
})

pdf("env_sigvars_with_corvars.pdf", width = 20, height = 15)
walk(plts, print)
dev.off()

plts <-
map(unique(lm_result$model), ~ {
  sig_df <- lm_result %>% filter(model == .x) %>% filter(`Pr...t..` < 0.05)
  sig_env_df <- env_df_vals[[.x]] %>% pivot_longer(contains("bio"), names_to = "var") %>% filter(var %in% unique(sig_df$var))
  title <- paste0(.x, " (adj R^2 = ", round(unique(sig_df$adj_rsqr), 2), ")" )

  if (nrow(sig_env_df) == 0) return(NULL)
  plt <-
  ggplot(sig_env_df, aes(value, Ho)) +
    geom_point(alpha = 0.7) +
    facet_wrap(~var, scales = "free", nrow = 1) +
    stat_regline_equation(
      aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
      formula = y ~ x
    ) +
    geom_smooth(method = "lm") +
    stat_cor(label.y = 0) +
    theme_classic() +
    ggtitle(title)

  plt
})

```

```{r}
# showing off bio2
mod_df <- env_df_vals[["dif_lh"]]

plt2 <-
ggplot(mod_df) +
  geom_sf(data = ca) +
  geom_point(aes(x = x, y = y, color = bio_2), cex = 2) +
  labs(color = "Change\nin mean\ndiurnal\nrange") +
  scale_color_viridis_c(option = "magma", direction = -1) +
  theme_void() 

plt3 <-
ggplot(mod_df) +
  geom_sf(data = ca) +
  geom_point(aes(x = x, y = y, color = Ho), cex = 2) +
  theme_void() +
  scale_color_viridis_c(option = "magma")
  

plt1 <- ggplot(mod_df, aes(bio_2, Ho)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation(
    aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~~~")),
    formula = y ~ x
  ) +
  xlab("Change in mean diurnal range") +
  stat_cor(label.y = 0) +
  xlim(-7.5, NA) +
  theme_classic()

pdf(here("bio2.pdf"), width = 15, height = 5)
grid.arrange(plt1, plt2, plt3, nrow = 1)
dev.off()
```

### Evaluating other time periods
```{r}
mega_df <- bind_rows(env_df_vals, .id = "timeperiod") %>% mutate(timeperiod = factor(timeperiod, levels = names(env_df_vals)))
mega_df2 <- 
  mega_df %>% 
  filter(grepl("dif", timeperiod)) %>%
  mutate(id = gsub("dif_", "", timeperiod)) %>%
  left_join(paleokey) %>%
  mutate(dif = paste0("Current to ", period, "\n(", time, ")")) %>%
  mutate(dif = factor(dif, levels = unique(dif)))

ggplot(mega_df2, aes(bio_2, Ho)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation(
    aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~~~")),
    formula = y ~ x
  ) +
  facet_wrap(~dif, scales = "free", nrow = 2) +
  xlab("Change in mean diurnal range") +
  stat_cor(label.y = 0) +
  theme_classic()
```


### Evaluating variables correlated with bio2
```{r}
wide_df <- 
  mega_df %>% 
  # drop MI19 because of missing values
  filter(!grepl("mis19", timeperiod)) %>%
  # note: pivot wider requires rows be uniquely identified (so sampleID is included here)
  select(timeperiod, Ho, SampleID, contains("bio")) %>%
  pivot_wider(names_from = "timeperiod", values_from = contains("bio"))

bio2_cors <- cor(wide_df %>% select(contains("bio")), use = "complete.obs")[,"bio_2_dif_lh"]
bio2_cors_names <- names(bio2_cors[abs(bio2_cors) > 0.6])
#bio1 - mean annual temperature
#bio6 - min temperature of coldest month (bio6)
#bio5 - max temperature of warmest month (bio5)
#bio7 - temperature annual range (bio5-bio6)
#bio2 - mean diurnal range (mean of monthly (max temp - min temp)
#bio12 - annual precipitation
#bio13 - precipitation of wettest month

gg_df <- 
  wide_df %>% 
  pivot_longer(all_of(bio2_cors_names)) %>%
  left_join(get_coords())

plt1 <- ggplot(gg_df, aes(value, Ho)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation(
    aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~~~")),
    formula = y ~ x
  ) +
  facet_wrap(~name, scales = "free") +
  stat_cor(label.y = 0) +
  theme_classic() +
  xlab("value")

plt_list <- gg_df %>%
  split(.$name) %>%
  map(~ {
    limit <- max(abs(.x[["value"]])) * c(-1, 1)
    ggplot(.x) +
        geom_sf(data = ca, fill = "darkgray") +
        labs(color = "value") +
        geom_point(aes(x = x, y = y, color = value), cex = 2) +
        scale_color_distiller(type = "div", limit = limit) +
        facet_wrap(~name) +
        theme_void()}
        )

plt3 <-
ggplot(gg_df) +
  geom_sf(data = ca) +
  geom_point(aes(x = x, y = y, color = Ho)) +
  ggtitle("E. Genetic diversity") +
  theme_void() +
  scale_color_viridis_c(option = "magma")  

pdf("correlated_vars.pdf", width = 10, height = 10)
print(plt1)
do.call(grid.arrange, c(c(plt_list), nrow = 3))
print(plt3)
dev.off()
```

### Evaluationg all diff_vals
```{r}
# Calculate the difference between all unique combinations of env_list
coords <- get_coords()
combos <- data.frame(t(combn(names(env_list), 2)))
dif_vals <- map2(combos[,1], combos[,2], ~{
  diff <- env_list[[.x]] - env_list[[.y]]
  names(diff) <- unique(c(names(env_list[[.x]]), names(env_list[[.y]])))
  vals <- extract(diff, coords[,c("x", "y")], ID = FALSE)
  df <- data.frame(vals, comparison = paste0(.x, "_", .y), SampleID = coords$SampleID)
  return(df)
}) %>% bind_rows()

dif_ls <- left_join(dif_vals, get_het()) %>% group_by(comparison) %>% group_split()
dif_vals <- left_join(dif_vals, get_het()) %>% pivot_longer(contains("bio"), names_to = "bioclim")
cor_vals <- 
  map(dif_ls, ~ {
    cordf <- .x %>% select(Ho, contains("bio")) %>% drop_na(Ho)
    res <- cor(cordf, use = "complete.obs")["Ho", -1]
  }) 
names(cor_vals) <- unique(dif_vals$comparison)

cor_df <- 
  cor_vals %>% 
  bind_rows(.id = "timeperiod") %>% 
  pivot_longer(contains("bio")) %>%
  mutate(bioclim = as.factor(as.numeric(gsub("bio_", "", name))))

# WARNING I THINK LABELS OF TIMEPERIODS/BIOCLIMS ARE OFF FIX THIS
pdf("all_comparisons.pdf", width = 13, height = 10)
ggplot(cor_df, aes(x = bioclim, y = timeperiod, fill = value)) +
  geom_tile() +
  labs(fill = expression(r[Ho])) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  theme_minimal()

ggplot(cor_df, aes(x = bioclim, y = timeperiod, fill = abs(value))) +
  geom_tile() +
  scale_fill_viridis_c(option = "rocket", direction = -1) +
  labs(fill = expression(r[Ho])) +
  theme_minimal()
dev.off()
```

### Evaluating variance
```{r}
var_df <- 
  mega_df %>%
  group_by(SampleID) %>%
  mutate(across(contains("bio"), ~ var(., na.rm =TRUE))) %>%
  pivot_longer(contains("bio"))

pdf("var.pdf", width = 13, height = 10)
ggplot(var_df, aes(value, Ho)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation(
    aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~~~")),
    formula = y ~ x
  ) +
  facet_wrap(~name, scales = "free") +
  stat_cor(label.y = 0) +
  theme_classic() +
  xlab("value")
dev.off()
```
### big lazy model
```{r}
no_na_df <- wide_df %>% drop_na(contains("bio"), Ho)
pca <- prcomp(drop_na(select(no_na_df, contains("bio"))))
pcs <- pca$x[,1:3]
#make a screeplot
screeplot(pca, type="lines", main="Screeplot of PCA")

mod_df_pca <- data.frame(Ho = no_na_df$Ho, pcs)

summary(lm(Ho ~ . , data = mod_df_pca))

all_cors <- cor(no_na_df %>% select(contains("bio")))

```

```{r}
# get all cors and get absolute value (has to be absolute for clustering!)
all_cors <- abs(cor(select(no_na_df, contains("bio"))))

# Convert the correlation matrix to a distance matrix (1 - |correlation|)
dist_matrix <- as.dist(1 - all_cors)

# Perform hierarchical clustering using the distance matrix
hc <- hclust(dist_matrix)

# Cut the dendrogram to create clusters, using a height that corresponds to the correlation threshold
# The height for cutting is 1 - threshold because the distance is 1 - correlation
clusters <- cutree(hc, h = 1 - 0.6)

# Group the variable names based on the clustering result
grouped_vars <- split(names(clusters), clusters)

# Get the var from each group with the strongest correlation with Ho
selected_vars <- 
map_chr(combined_groups, ~ {
  if (length(.x) == 1) return(.x)
  vars <- no_na_df %>% select(all_of(.x))
  # find var with strongest correlation with Ho
  Ho_cor <- cor(vars, no_na_df$Ho, use = "complete.obs")
  var <- row.names(Ho_cor)[which.max(abs(Ho_cor))]
  return(var)
}) %>% as.vector()


# check that all cors are below 0.6
check_df <- select(no_na_df, all_of(selected_vars))
check_cor <- cor(check_df, use = "complete.obs")
diag(check_cor) <- 0
all(abs(check_cor) < 0.6)

# run model
mod_df_no_cor <- no_na_df %>% select(all_of(selected_vars), Ho)
mod <- lm(Ho ~ ., data = mod_df_no_cor)
summary(mod)

df <- data.frame(predicted = mod$fitted.values, observed = mod_df_no_cor$Ho)

ggplot(df, aes(x = observed, y = predicted)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation(
    aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
    formula = y ~ x
  ) +
  stat_cor(label.y = 0) +
  theme_classic()
```

### Testing for effects of population structure
```{r}
source(here("analysis", "admixture", "admixture.R"))
admix <- get_Q(K = 5)
head(admix)
coords <- get_coords()
mod_df <- left_join(mod_df, admix)

# bio2 is still significant after controling for clusters
summary(lm(Ho ~ cluster + bio_2, data = mod_df))
# nothing is significant once an interaction term is added but I think that is due to low power
summary(lm(Ho ~ cluster + bio_2, data = mod_df))

# look at heterozygosity by cluster and remove largest cluster
mod_df %>% group_by(cluster) %>% summarise(mean = mean(Ho, na.rm = TRUE)) %>% arrange(desc(mean))
# when you remove the southern pop with really high genetic diversity (cluster 3), the pattern still exists (albeit weaker)
plt1 <- ggplot(mod_df[mod_df$cluster != 3, ], aes(bio_2, Ho)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation(
    aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~~~")),
    formula = y ~ x
  ) +
  stat_cor(label.y = 0) +
  theme_classic()

plt2 <- ggplot(mod_df[mod_df$cluster != 3, ]) +
  geom_sf(data = ca) +
  geom_point(aes(x = x, y = y, color = bio_2), cex = 2) +
  labs(color = "bio_2") +
  theme_void() +
  scale_color_viridis_c(option = "magma", direction = -1)

plt3 <- ggplot(mod_df[mod_df$cluster != 3, ]) +
  geom_sf(data = ca) +
  geom_point(aes(x = x, y = y, color = Ho), cex = 2) +
  theme_void() +
  scale_color_viridis_c(option = "magma")

grid.arrange(plt1, plt2, plt3, nrow = 1)
```

### Checking with bouzid data
```{r}
vcf <- vcfR::read.vcfR("../wingen/paperex/empex/data/populations_r20.haplotypes.filtered_m70_randomSNP.vcf")
coords <- read_table("../wingen/paperex/empex/data/Scelop.coord", col_names =FALSE) %>% rename(x = X1, y = X2)
dos <- algatr::vcf_to_dosage(vcf)
dos_to_het <- function(dos){
  dos[dos == 1] <- 1
  dos[dos != 1] <- 0
  het <- rowMeans(dos, na.rm = TRUE)
  return(het)
}
df <- data.frame(coords, Ho = dos_to_het(dos))

ggplot(df) +
  geom_sf(data = ca) +
  geom_point(aes(x = x, y = y, color = Ho)) +
  scale_color_viridis_c(option = "magma") +
  theme_void()

lh <- load_paleoclim(here("data", "env", "LH_v1_2_5m.zip"))
cur <- load_paleoclim(here("data", "env", "CHELSA_cur_V1_2B_r2_5m.zip"))
dif <- cur - lh
lh_vals <- extract(dif, df[,c("x", "y")], ID = FALSE)

mod_df <- data.frame(df, lh_vals)

ggplot(mod_df, aes(bio_2, Ho)) +
    geom_point(alpha = 0.7) +
    stat_regline_equation(
      aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
      formula = y ~ x
    ) +
    geom_smooth(method = "lm") +
    stat_cor(label.y = 0) +
    theme_classic() 

#bio5 (max temperature of warmest month)
#bio6 (min temperature of coldest month)
```

### Testing for current environment effects
```{r}

df <-
 mega_df %>% 
 filter(timeperiod %in% c("dif_lh", "cur"))%>% 
 select("Ho", "bio_2", "timeperiod", "SampleID") %>% 
 pivot_wider(names_from = timeperiod, values_from = bio_2) %>% 
 drop_na()

summary(lm(Ho ~ cur + dif_lh, data = df))
```


### Looking at change over time
```{r}
names(mega_df)
paleokey_split <- 
  paleokey %>%
  mutate(time = 
    case_when(
      period == "Current" ~ "1.979-2.013 ka", 
      .default = time
      )
    ) %>%
  separate(time, into = c("start_time", "end_time"), sep = "-", convert = TRUE, fill = "right") %>%
  mutate(across(c(start_time, end_time), ~ case_when(
    str_detect(.x, "ka") ~ as.numeric(str_remove(str_remove(.x, "ka"), "ca.")) * 1000,
    str_detect(.x, "Ma") ~ as.numeric(str_remove(str_remove(.x, "Ma"), "ca.")) * 1000000,
    TRUE ~ as.numeric(.x)
  )))

change_df <- 
  mega_df %>%
  left_join(paleokey_split, by = c("timeperiod" ="id")) %>%
  pivot_longer(c(start_time, end_time), names_to = "start_end", values_to = "time") 

mean_change <-
  change_df %>%
  group_by(time) %>% summarize(bio_2 = mean(bio_2, na.rm =TRUE))

ggplot() +
  geom_line(data = change_df, aes(x = time, y = bio_2, group = SampleID, col = "Samples"), alpha = 0.3) +
  geom_line(data = mean_change, aes(x = time, y = bio_2, col = "Mean"), lwd = 2) +
  scale_color_manual(values = c("#007c8f", "darkgray")) + 
  labs(col = "") +
  scale_x_log10()  +
  theme_classic()
```


# SDM
```{r}
library(countrycode)
library(CoordinateCleaner)
library(rgbif)
library(sp)
library(here)
library(raster)
library(dismo)
library(tidyverse)
library(viridis)
library(biomod2)
library(sf)
library(terra)
library(ggpubr)
# devtools::install_github('adamlilith/enmSdm', dependencies=TRUE)
library(enmSdm)
library(RStoolbox)
source(here("analysis", "sdm", "bias.R"))
source(here("analysis", "sdm", "sdm.R"))
source(here("analysis", "sdm", "get_coords.R"))
source(here("analysis", "sdm", "paleovars.R"))
source(here("general_functions.R"))

# set maxent path
# put maxent.jar in your working directory (what you get when you run getwd()) or set manually: https://rdrr.io/cran/biomod2/man/BIOMOD_ModelingOptions.html
Sys.setenv(JAVA_HOME='/usr/lib/jvm/java-11-openjdk-amd64')

# get env data
env_list <- get_paleovars()
env_list_agg <- map(env_list, ~aggregate(.x, 2))
cur <- env_list[["cur"]]
lh <- env_list[["lh"]]

# pca
selected_vars <- pc_var_selection(cur)

# get coordinates
occ_coords <- get_occ_coords("Sceloporus occidentalis", data_source = "both", occ_limit = 10000, cache = TRUE)

# create bias background layer
bias_class <- make_biasbkg(key = "Sceloporus occidentalis", rank = "class", lyr = get_lyr(), limit = 1000, n = 10000)
bias_class <- get_bias("Sceloporus occidentalis", rank = "class")

# run SDM
sdm_mod <- sdm(occ_coords, cur[[selected_vars]], biasdat = bias_class, models = "RF", nbkg = 10000, output = FALSE, file.name = "", tune = TRUE)

# hindcast
lh_sdm <- hindcast(sdm_mod$myBiomodModelOut, new_env = lh[[selected_vars]])
hindcast_sdms <- map(env_list[c("lig", "lgm", "hs1", "ba", "yds", "eh", "mh", "lh", "cur")], ~hindcast(sdm_mod$myBiomodModelOut, new_env = .x[[selected_vars]]))
hindcast_sdms <- rast(hindcast_sdms)

# calculate change and stability
# TODO: PLAY AROUND WITH TIMEPERIOD USED TO CALCULATE DIF_SDM AND STABILITY
cur_sdm <- sdm_mod$raster 
dif_sdm <- cur_sdm - hindcast_sdms[["lh"]]
stability <- sum(subset(hindcast_sdms, c("eh", "mh", "lh", "cur")))
plot(stability)
plot(dif_sdm)
 
# compare to genetic diversity
het <- get_het()
roh <- get_roh()
coords <- get_coords(sf = TRUE) %>% left_join(het) %>% left_join(roh)
change_vals <- extract(dif_sdm, coords, ID = FALSE)
stability_vals <- extract(stability, coords, ID = FALSE)
df <- data.frame(change = change_vals[,1], stability = stability_vals[,1],  coords)
ggplot(df, aes(x = change, y = Ho)) +
  geom_point() +
  stat_regline_equation(
      aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
      formula = y ~ x
    ) +
  geom_smooth(method = "lm") +
  stat_cor(label.y = 0) +
  theme_classic()

ggplot(df, aes(x = stability, y = Ho)) +
  geom_point() +
  stat_regline_equation(
      aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
      formula = y ~ x
    ) +
  geom_smooth(method = "lm") +
  stat_cor(label.y = 0) +
  theme_classic()

ggplot(df, aes(x = change, y = froh0)) +
  geom_point() +
  stat_regline_equation(
      aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
      formula = y ~ x
    ) +
  geom_smooth(method = "lm") +
  stat_cor(label.y = 0) +
  theme_classic()

# grouped
df_group <- 
  df %>% 
  mutate(stability = stability > 3) %>%
  drop_na(stability)

# play around with this:
ggplot(df_group) +
  geom_boxplot(aes(x = stability, y = froh0))
```

```{r}
# is stability correlated with change in bio2
dif_bio2 <-env_list[["cur"]][["bio_2"]] -  env_list[["lh"]][["bio_2"]]
dif_vals <- extract(dif_bio2, coords, ID = FALSE)
df$change_bio2 <- dif_vals[,1]

# confirm correlations with Ho
cor(df$change_bio2, df$Ho, use = "pairwise.complete.obs")
cor(df$stability, df$Ho, use = "pairwise.complete.obs")

# correlated, but not strongly
cor(df$change_bio2, df$stability, use = "pairwise.complete.obs")

# check for statistical suppression/whether both variables are significant when both are included
mod1 <- lm(Ho ~ change_bio2, data = df)
summary(mod1)
mod2 <- lm(Ho ~ stability, data = df)
summary(mod2)
mod3 <- lm(Ho ~ stability + change_bio2, data = df)
summary(mod3)
# stability is no longer significant when bio2 is included
```


# Recent climate
Download PRISM max temperature data from historical data: https://prism.oregonstate.edu/historical/
```{r}
caladapt <- map(list.files(here("data", "env", "caladapt"), pattern = ".tif", full.names = TRUE), rast) %>% rast()

names(caladapt)

caladapt1950 <- caladapt[["livneh_year_Tmax_clip_15Oct2014.1950"]]
caladapt2013 <- caladapt[["livneh_year_Tmax_clip_15Oct2014.2013"]]
dif <- caladapt2013 - caladapt1950
plot(dif)
```

```{r}
# Function to calculate yearly average from monthly files
coords <- get_coords(sf = TRUE)
year = 1895
calculate_yearly_avg <- function(year  = 1895, type = "historical", output = "df") {
  if (type == "historical") directory_path <- here::here("data", "env", "prism", "PRISM_tmax_stable_4kmM3_189501_198012_bil")
  if (type == "current") directory_path <- here::here("data", "env", "prism", "PRISM_tmax_stable_4kmM3_198101_202308_bil")
  monthly_files <- list.files(directory_path, pattern = "*.bil.bil$", full.names = TRUE)
  monthly_files <- monthly_files[grepl(paste0("PRISM_tmax_stable_4kmM3_", year), monthly_files)]
  if (length(monthly_files) > 0) {
    rasters <- purrr::map(monthly_files, terra::rast)
    rasters <- terra::rast(rasters)
    if (output == "raster"){
      rasters <- mean(rasters, na.rm = TRUE)
      rm(rasters)
      gc()
    } else {
      vals <- terra::extract(rasters, st_transform(coords, st_crs(rasters)))
      annual_avg <- rowMeans(vals, na.rm = TRUE)
      return(data.frame(year = year, tmax = annual_avg, SampleID = coords$SampleID))
      rm(rasters)
      gc()
    }
    
  } else {
    return(NULL)
  }
}

# Loop over the years of interest
years <- 1895:1925 # Update this to the range of years you have data for
tmax_his <- map(1895:1925, ~calculate_yearly_avg(.x, type = "historical"), .progress =TRUE) %>% bind_rows()  
tmax_his <- tmax_his %>% group_by(SampleID) %>% summarise(tmax_his = mean(tmax, na.rm = TRUE))
tmax_cur <- map(1986:2016, ~calculate_yearly_avg(.x, type = "current"), .progress =TRUE) %>% bind_rows()  
tmax_cur <- tmax_cur %>% group_by(SampleID) %>% summarise(tmax_cur = mean(tmax, na.rm = TRUE))
tmax <- 
left_join(tmax_his, tmax_cur) %>% 
mutate(dif = tmax_cur - tmax_his) %>%
right_join(coords) %>% st_as_sf()

ggplot(tmax) +
  geom_sf(data = ca) +
  geom_sf(aes(color = dif)) +
  scale_color_viridis_c(option = "magma", direction = -1) +
  theme_void()

tmax <- left_join(tmax, froh)

summary(lm(froh0 ~ dif, data =tmax))
summary(lm(froh0 ~ tmax_cur, data =tmax))
```

```{r}
# Compute the long-term average from 1895 to 1925
long_term_avg <- app(stack(annual_averages), mean)
```

```{r}
tmax_his <- map(1895:1925, ~calculate_yearly_avg(.x, type = "historical", output = "raster"), .progress =TRUE) %>% rast()  
tmax_cur <- map(1986:2016, ~calculate_yearly_avg(.x, type = "current"), .progress =TRUE) %>% bind_rows()  
```

# pastclimate

This data is not high enough resolution but keeping it here in case I want it in the future

```{r}
library("pastclim")
library("here")
set_data_path(path_to_nc = here("data", "env", "pastclim"))

# downloads all variables (annual) by default
download_dataset(dataset = "Beyer2020", annual = TRUE)
```

It is possible to also load a time series of rasters with the function region_series. In this case, the function returns a SpatRasterDataset, with each variable as a sub-dataset:
```{r}
ca <- vect(get_ca())
climate_region <- region_series(
  time_bp = list(min = -15000, max = 0),
  bio_variables = c("bio05", "biome"),
  dataset = "Beyer2020",
  crop = ca
)
climate_region
```

Note that terra stores dates in years as AD, not BP. You can inspect the times in years BP with:
```{r}
time_bp(climate_region)
```

We can then plot the time series of a given variable (we relabel the plots to use years bp):
```{r}
terra::plot(climate_region$bio05, main = time_bp(climate_region))
```

To plot all climate variables for a given time step, we can slice the time series:
```{r}
slice_10k <- slice_region_series(climate_region, time_bp = -10000)
terra::plot(slice_10k)
```

Get ice sheets:
```{r}
terra::plot(climate_region$biome, main = time_bp(climate_region))
ice_id <- get_biome_classes("Beyer2020") %>% filter(category == "Land ice") %>% pull(id)
ice <- climate_region$biome*0

# check if there is any ice
global(any(climate_region$biome == ice_id), sum, na.rm = TRUE)
ice[climate_region$biome == ice_id] <- 1
plot(ice)

# unfortunately the resolution is not high enough
```