```{r}
library(tidyverse)
library(here)
library(sf)
source(here("general_functions.R"))
outpath <- here("analysis", "gea", "outputs")
plotpath <- here("analysis", "gea", "plots")
```

```{r}
# Get pi in 50kb windows centered on each GEA SNP
gea_pi <- 
  read_table(here(outpath, "bio1ndvi_gea_genes_nonsyn_flanking_pi_by_pop.tsv")) %>% 
  rename(CHROM = chrom) %>%
  mutate(pop = gsub("k8_pop", "", population)) 

# Get pi for all windows in each population
pops <- 1:8
names(pops) <- pops
pop_window_files <- map(pops, ~ here(outpath, paste0("k8_pop", .x, "_50kb_windowpi.windowed.pi")))
wpi <- 
  map(pop_window_files, read_table) %>% 
  bind_rows(.id = "pop") %>%
  filter(grepl("chr", CHROM)) %>%
  select(CHROM, BIN_START, BIN_END, PI, pop)

stopifnot(all(gea_pi$pop %in% wpi$pop))

# Compute empirical p-values by comparing each GEA SNP's pi to the distribution of pi in windows from the same chromosome and population
# left-tailed: lower pi than expected
combos <- expand.grid(chr_i = unique(gea_pi$CHROM), pop_i = unique(gea_pi$pop))
p_values <- pmap(combos, \(chr_i, pop_i) {
  # Get the GEA SNPs for this chromosome and population
  snp_pi <- gea_pi %>% filter(CHROM == chr_i, pop == pop_i)
  
  # Get the null distribution of pi for this chromosome and population
  null_pi <- wpi %>% filter(CHROM == chr_i, pop == pop_i) %>% pull(PI)
  
  # Calculate empirical p-values for each GEA SNP in this group
  snp_pi %>%
    mutate(
      p_lower = map_dbl(center_pi, ~ (sum(null_pi <= .x, na.rm = TRUE) + 1) / (length(na.omit(null_pi)) + 1)),
      p_upper = map_dbl(center_pi, ~ (sum(null_pi >= .x, na.rm = TRUE) + 1) / (length(na.omit(null_pi)) + 1))
    )
  }, .progress = TRUE) %>%
  bind_rows()

gea_with_p <- 
  p_values %>%
  mutate(
    p_lower_adj = p.adjust(p_lower, method = "fdr"),
    p_upper_adj = p.adjust(p_upper, method = "fdr")
  )

gea_with_p %>% 
  filter(p_lower < 0.05) %>%
  distinct(snp_id)

gea_with_p %>% 
  filter(p_lower_adj < 0.05) %>%
  distinct(snp_id)

gea_with_p %>% 
  filter(p_upper_adj < 0.05) %>%
  distinct(snp_id)

# Plot results
null_df <- wpi %>%
  filter(!is.na(PI), grepl("chr", CHROM)) %>%
  transmute(CHROM, pop, pi = PI, source = "Background windows")

gea_df <- gea_with_p %>%
  filter(!is.na(center_pi), grepl("chr", CHROM)) %>%
  transmute(CHROM, pop, pi = center_pi, source = "GEA SNP windows")

plot_df <- bind_rows(null_df, gea_df) %>%
  mutate(
    source = factor(source, levels = c("Background windows", "GEA SNP windows")),
    pop = factor(paste0("Population ", pop)),
    CHROM = factor(CHROM, levels = paste0("chr", 1:10))
  )

png(here(plotpath, "window_pi_distributions.png"), width = 13, height = 9, units = "in", res = 300)
ggplot() +
  geom_density(
    data = filter(plot_df, source == "Background windows"),
    aes(x = pi, col = "Background windows"),
    fill = "grey80",
    alpha = 0.8
  ) +
  geom_density(
    data = filter(plot_df, source == "GEA SNP windows"),
    aes(x = pi, color = "Non-synonymous GEA SNP windows"),
    linewidth = 0.8
  ) +
  facet_grid(pop ~ CHROM, scales = "free_y") +
  labs(
    x = expression(pi),
    y = "Density",
    col = ""
  ) +
  theme_classic() +
  theme(
    legend.position = "top", 
    strip.background = element_blank(), 
    panel.border = element_rect(),
    legend.text = element_text(size = 16)
  ) +
  scale_color_manual(values = c("gray50", "tomato2"))
dev.off()


```

# OLD
```{r}

gea_pi <- read_table(here(outpath, "bio1ndvi_gea_genes_nonsyn_50kb_pi.tsv")) %>% rename(CHROM = chrom) 
gw_pi <- read_table(here(outpath, "58-Sceloporus_50kb_windowpi.windowed.pi")) %>% filter(grepl("chr", CHROM))


pdf(here(plotpath, "window_pi_distributions.pdf"), width = 8, height = 4)
ggplot() +
  geom_density(data = gw_pi, aes(x = PI), fill = "lightblue", alpha = 0.5) +
  geom_density(data = gea_pi, aes(x = pi), fill = "salmon", alpha = 0.5) +
  theme_classic() +
  labs(x = expression("Mean " * pi * " (50kb window)"), y = "Density") +
  theme(legend.title = element_blank()) +
  scale_fill_manual(values = c("lightblue", "salmon"), labels = c("Genome-wide", "GEA genes"))
dev.off()

```

```{r}
hotcold_pi <- read_table(here(outpath, "bio1ndvi_gea_genes_nonsyn_flanking_pi_hot_cold.tsv")) %>% rename(CHROM = chrom)
gw_cold_pi <- read_table(here(outpath, "cold_50kb_windowpi.windowed.pi")) %>% filter(grepl("chr", CHROM))
gw_hot_pi <- read_table(here(outpath, "hot_50kb_windowpi.windowed.pi")) %>% filter(grepl("chr", CHROM))

aggregate(delta_pi ~ group, hotcold_pi, median, na.rm = TRUE)

head(hotcold_pi)

# center_pi_sum is the total π across variant sites, whereas center_pi normalizes by window length to estimate nucleotide diversity (π) per base

pdf(here(plotpath, "window_pi_distributions.pdf"), width = 8, height = 4)
ggplot() +
  geom_density(data = gw_cold_pi, aes(x = PI, fill = "Cold"), alpha = 0.5) +
  geom_density(data = gw_hot_pi, aes(x = PI, fill = "Hot"), alpha = 0.5) +
  geom_density(data = hotcold_pi, aes(x = center_pi, col = group), alpha = 0.5) +
  #geom_boxplot() +
  theme_classic() +
  scale_color_manual(values = c("lightblue", "salmon"), labels = c("Cold", "Hot")) +
  scale_fill_manual(values = c("lightblue", "salmon"), labels = c("Cold", "Hot")) 
dev.off()


pdf(here(plotpath, "window_pi_distributions.pdf"), width = 8, height = 4)
ggplot(hotcold_pi, aes(x = delta_pi, col = group)) +
  geom_density() +
  #geom_boxplot() +
  theme_classic()
dev.off()
```

```{r}
hotcold_pi <- read_table(here(outpath, "bio1ndvi_gea_genes_nonsyn_flanking_pi_hot_cold.tsv")) %>% rename(CHROM = chrom)
gw_cold_pi <- read_table(here(outpath, "cold_50kb_windowpi.windowed.pi")) %>% filter(grepl("chr", CHROM))
gw_hot_pi <- read_table(here(outpath, "hot_50kb_windowpi.windowed.pi")) %>% filter(grepl("chr", CHROM))

null_cold <- gw_cold_pi$PI
null_hot <- gw_hot_pi$PI

snp_pi <- 
  hotcold_pi %>%
  mutate(p_empirical = case_when(
    group == "cold" ~ map_dbl(center_pi, ~ (sum(null_cold <= .x) + 1) / (length(null_cold) + 1)),
    group == "hot"  ~ map_dbl(center_pi, ~ (sum(null_hot  <= .x) + 1) / (length(null_hot) + 1))
  )) 

sig <-
  snp_pi %>%
  filter(group == "cold" & p_empirical < 0.05 & center_pi < mean(null_cold) + 3 * sd(null_cold)) %>%
  arrange(center_pi) %>%
  head(10)

i  <- 
  wpi %>% 
  count(CHROM) %>% 
  arrange(desc(n)) %>% 
  # Pull out long scaffolds (chr and Scaffold_13)
  filter(n > 1000) %>% 
  pull(CHROM)

gg_df <-
  wpi %>%
  filter(CHROM %in% i) %>%
  mutate(CUMSUM = cumsum(BIN_START)) %>%
  mutate(CHROM = factor(CHROM, levels = i))


flank <- 1000000

goi_plot_df <- 
  sig %>%
  select(CHROM, pos, window_start = center_start, window_end = center_end) 

plot_df <- 
  goi_plot_df %>%
  inner_join(gg_df, by = "CHROM") %>%
  filter(BIN_START >= window_start, BIN_START <= window_end) %>%
  mutate(rel_pos = BIN_START - pos)

# summary across SNPs at each relative position
ribbon_df <- 
  plot_df %>%
  summarise(
    mean_pi = mean(PI, na.rm = TRUE),
    sd_pi = sd(PI, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    ymin = mean_pi - sd_pi,
    ymax = mean_pi + sd_pi
  )

mean_pi_global <- mean(plot_df$PI, na.rm = TRUE)

genes_nonsyn <- get_gea_genes(nonsyn = TRUE) %>% filter(locus %in% sig$snp_id)

genes_goi <- genes %>% filter(gene_name %in% unique(genes_nonsyn$gene_name)) %>% distinct()
# FIGURE OUT WHY TRPV ISNT INCLUDED

png(here(plotpath, "windowed_pi_100kb_around_goi_snps.png"), width = 8, height = 8, units = "in", res = 300)
ggplot(plot_df, aes(x = BIN_START, y = PI)) +
  geom_rect(data = ribbon_df, aes(ymin = ymin, ymax = ymax, xmin = -Inf, xmax = Inf), fill = "skyblue2", alpha = 0.2, inherit.aes = FALSE) +
  geom_rect(data = genes_goi, aes(xmin = gene_start, xmax = gene_end, ymin = -Inf, ymax = Inf), fill = "#94e6c0", alpha = 0.1, inherit.aes = FALSE) +
  geom_hline(yintercept = ribbon_df$mean_pi, linetype = "dashed", col = "skyblue2") +
  geom_line(linewidth = 0.5) +
  geom_vline(aes(xintercept = pos), col = "tomato2", linetype = "solid") +
  scale_x_continuous(labels = scales::scientific) +
  theme_classic() +
  facet_wrap(~ gene_name, scales = "free_x", ncol = 3) +
  labs(
    x = "Position relative to focal SNP (bp)",
    y = expression("Mean " * pi * " (50 kb window)")
  ) +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(size = 9)
  )

dev.off()
```
```


```{r}

source(here("analysis", "gea", "functions_selection_stats.R"))
genes <- get_gene_structure()
```

```{r}
library(tidyverse)

# read data
hotcold_pi <- read_table(here(outpath, "bio1ndvi_gea_genes_nonsyn_flanking_pi_hot_cold.tsv")) %>%
  rename(CHROM = chrom)

wpi <- read_table(here(outpath, "cold_50kb_windowpi.windowed.pi")) %>%
  filter(grepl("chr", CHROM))

null_cold <- wpi$PI %>% na.omit()

# empirical p-value for unusually low diversity in the cold group
snp_pi <- hotcold_pi %>%
  mutate(
    p_empirical = case_when(
      group == "cold" ~ map_dbl(center_pi, ~ (sum(null_cold <= .x) + 1) / (length(null_cold) + 1)),
      group == "hot"  ~ map_dbl(center_pi, ~ (sum(null_hot  <= .x) + 1) / (length(null_hot) + 1)),
      TRUE ~ NA_real_
    )
  )

# significant cold SNPs in low-diversity regions
sig <- 
  snp_pi %>%
  pivot_wider(names_from = group, values_from = p_empirical) %>%
  filter(cold < 0.05 & !(hot < 0.05)) #0



```


```{r}
wpi_cold <- read_table(here(outpath, "cold_50kb_windowpi.windowed.pi")) %>% filter(grepl("chr", CHROM)) %>% mutate(pop = "Cold") %>% mutate(PI = scale(PI)[,1])
wpi_hot <- read_table(here(outpath, "hot_50kb_windowpi.windowed.pi")) %>% filter(grepl("chr", CHROM)) %>% mutate(pop = "Hot") %>% mutate(PI = scale(PI)[,1])
wpi <- bind_rows(wpi_cold, wpi_hot) 

gg_df <-
  wpi %>%
  filter(grepl("chr", CHROM)) %>%
  mutate(CUMSUM = cumsum(BIN_START)) %>%
  mutate(CHROM = factor(CHROM, levels = i)) 

goi <- read_csv(here(outpath, "genes_of_interest.csv"))

flank <- 1000000

goi_plot_df <- goi %>%
  mutate(
    CHROM = scaffold,
    pos = position,
    window_start = pmax(0, pos - flank),
    window_end = pos + flank,
    snp_id = paste0(gene_name, ":", CHROM, ":", scales::comma(pos))
  ) %>%
  select(gene_name, CHROM, pos, window_start, window_end, snp_id)

plot_df <- 
  goi_plot_df %>%
  inner_join(gg_df, by = "CHROM") %>%
  filter(BIN_START >= window_start, BIN_START <= window_end) %>%
  mutate(rel_pos = BIN_START - pos)

# summary across SNPs at each relative position
ribbon_df <- 
  plot_df %>%
  summarise(
    mean_pi = mean(PI, na.rm = TRUE),
    sd_pi = sd(PI, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    ymin = mean_pi - sd_pi,
    ymax = mean_pi + sd_pi
  )

mean_pi_global <- mean(plot_df$PI, na.rm = TRUE)

genes_goi <- genes %>% filter(gene_name %in% goi$gene_name) %>% filter(gene_name %in% plot_df$gene_name)
# FIGURE OUT WHY TRPV ISNT INCLUDED

png(here(plotpath, "windowed_pi_100kb_around_goi_snps.png"), width = 8, height = 8, units = "in", res = 300)
ggplot(plot_df, aes(x = BIN_START, y = PI)) +
  geom_rect(data = ribbon_df, aes(ymin = ymin, ymax = ymax, xmin = -Inf, xmax = Inf), fill = "skyblue2", alpha = 0.2, inherit.aes = FALSE) +
  geom_rect(data = genes_goi, aes(xmin = gene_start, xmax = gene_end, ymin = -Inf, ymax = Inf), fill = "gray", alpha = 0.1, inherit.aes = FALSE) +
  geom_hline(data = ribbon_df, aes(yintercept = mean_pi), linetype = "dashed", color = "skyblue2") +
  geom_line(linewidth = 0.5, aes(col = pop)) +
  geom_vline(aes(xintercept = pos), col = "black", linetype = "solid") +
  scale_x_continuous(labels = scales::scientific) +
  theme_classic() +
  facet_wrap(~ gene_name, scales = "free_x", ncol = 3) +
  labs(
    x = "Position relative to focal SNP (bp)",
    y = expression("Mean " * pi * " (50 kb window)")
  ) +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(size = 9)
  )

dev.off()


pdf(here(plotpath, "windowed_pi.pdf"), width = 12, height = 10)
ggplot(gg_df) +
  #geom_vline(aes(xintercept = BIN_START, col = PI), alpha = 0.5) +
  geom_line(aes(x = BIN_START, y = PI, col = pop), alpha = 0.5, lwd = 0.5) +
  scale_x_continuous(labels = scales::comma, expand = c(0, 0)) +
  scale_y_continuous(labels = scales::comma) +
  theme_classic() +
  facet_wrap(~CHROM, scales = "fixed", ncol = 1) +
  theme(axis.text.x = element_blank(), 
        axis.title.x = element_blank(), 
        axis.line.x = element_blank(),
        strip.background = element_blank(),
        strip.text.x = element_text(hjust = 0, size = 10), legend.position = "bottom") +
  labs(y = expression("Mean " * pi * " (10kb window)"), col =expression("Mean " * pi * " (10kb window)"))
dev.off()

#subtract cold from hot to see if there are any clear differences in π around the SNPs of interest
gg_df2 <- 
  wpi_hot %>% 
  rename(Hot = PI) %>%
  left_join(wpi_cold, by = c("CHROM", "BIN_START", "BIN_END")) %>%
  rename(Cold = PI) %>%
  mutate(delta_pi = Hot - Cold) %>%
  distinct(CHROM, BIN_START, delta_pi)

pdf(here(plotpath, "windowed_pi.pdf"), width = 20, height = 10)
ggplot(gg_df2) +
  #geom_vline(aes(xintercept = BIN_START, col = PI), alpha = 0.5) +
  geom_line(aes(x = BIN_START, y = delta_pi), alpha = 0.5, lwd = 0.5) +
  scale_x_continuous(labels = scales::comma, expand = c(0, 0)) +
  scale_y_continuous(labels = scales::comma) +
  theme_classic() +
  facet_wrap(~CHROM, scales = "fixed", ncol = 1) +
  theme(axis.text.x = element_blank(), 
        axis.title.x = element_blank(), 
        axis.line.x = element_blank(),
        strip.background = element_blank(),
        strip.text.x = element_text(hjust = 0, size = 10), legend.position = "bottom") +
  labs(y = expression("Mean " * pi * " (10kb window)"), col =expression("Mean " * pi * " (10kb window)"))
dev.off()
```

```{r}

pops <- 1:8
names(pops) <- pops
pop_window_files <- map(pops, ~ here(outpath, paste0("k8_pop", .x, "_50kb_windowpi.windowed.pi")))
wpi <- map(pop_window_files, read_table) %>% bind_rows(.id = "pop") 

gg_df <-
  wpi %>%
  group_by(pop) %>%
  # Scale pi
  mutate(PI = scale(PI)[,1]) %>%
  ungroup() %>%
  filter(grepl("chr", CHROM)) %>%
  mutate(CUMSUM = cumsum(BIN_START)) %>%
  mutate(CHROM = factor(CHROM, levels = i)) 

goi <- read_csv(here(outpath, "genes_of_interest.csv"))

flank <- 500000

goi_plot_df <- goi %>%
  mutate(
    CHROM = scaffold,
    pos = position,
    window_start = pmax(0, pos - flank),
    window_end = pos + flank,
    snp_id = paste0(gene_name, ":", CHROM, ":", scales::comma(pos))
  ) %>%
  select(gene_name, CHROM, pos, window_start, window_end, snp_id)

plot_df <- 
  goi_plot_df %>%
  inner_join(gg_df, by = "CHROM") %>%
  filter(BIN_START >= window_start, BIN_START <= window_end) %>%
  mutate(rel_pos = BIN_START - pos)

# summary across SNPs at each relative position
ribbon_df <- 
  plot_df %>%
  summarise(
    mean_pi = mean(PI, na.rm = TRUE),
    sd_pi = sd(PI, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    ymin = mean_pi - sd_pi,
    ymax = mean_pi + sd_pi
  )

mean_pi_global <- mean(plot_df$PI, na.rm = TRUE)

genes_goi <- genes %>% filter(gene_name %in% goi$gene_name) %>% filter(gene_name %in% plot_df$gene_name)
# FIGURE OUT WHY TRPV ISNT INCLUDED

png(here(plotpath, "windowed_pi_100kb_around_goi_snps.png"), width = 8, height = 8, units = "in", res = 300)
ggplot(plot_df, aes(x = BIN_START, y = PI)) +
  geom_rect(data = ribbon_df, aes(ymin = ymin, ymax = ymax, xmin = -Inf, xmax = Inf), alpha = 0.2, inherit.aes = FALSE) +
  geom_rect(data = genes_goi, aes(xmin = gene_start, xmax = gene_end, ymin = -Inf, ymax = Inf), fill = "gray", alpha = 0.1, inherit.aes = FALSE) +
  geom_hline(data = ribbon_df, aes(yintercept = mean_pi), linetype = "dashed") +
  geom_line(linewidth = 0.5, aes(col = pop)) +
  geom_vline(aes(xintercept = pos), col = "black", linetype = "solid") +
  scale_x_continuous(labels = scales::scientific) +
  theme_classic() +
  facet_wrap(~ gene_name, scales = "free_x", ncol = 3) +
  labs(
    x = "Position",
    y = expression("Mean " * pi * " (50 kb window)")
  ) +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(size = 9)
  )

dev.off()
```

```{r}
library(tidyverse)
library(here)
library(readr)

pop_df <- read_csv(here("analysis", "admixture", "outputs", "Q8.csv"))
cold_individuals <- read_table(here("analysis", "genetic_diversity", "outputs", "cold_individuals.txt"), col_names = "SampleID")
pops_with_cold <- pop_df %>%
  filter(SampleID %in% cold_individuals$SampleID) %>%
  pull(cluster)

# read pops that are not cold 
pops <- setdiff(1:8, pops_with_cold)
names(pops) <- pops

pop_window_files <- map(pops, ~ here(outpath, paste0("k8_pop", .x, "_50kb_windowpi.windowed.pi")))

wpi_pop <- 
  map2(pop_window_files, names(pops), ~ read_table(.x) %>%
    filter(grepl("chr", CHROM)) %>%
    mutate(pop = .y)
  ) %>%
  bind_rows()

# read cold population
wpi_cold <- 
  read_table(here(outpath, "cold_50kb_windowpi.windowed.pi")) %>%
  filter(grepl("chr", CHROM)) %>%
  mutate(pop = "Cold")

# summarize the 8 comparison populations for each window
pop_summary <- 
  wpi_pop %>%
  group_by(CHROM, BIN_START, BIN_END) %>%
  summarise(
    mean_other = mean(PI, na.rm = TRUE),
    sd_other   = sd(PI, na.rm = TRUE),
    min_other  = min(PI, na.rm = TRUE),
    n_other    = n(),
    .groups = "drop"
  )

# join cold to summary of other populations
cold_vs_others <- 
  wpi_cold %>%
  select(CHROM, BIN_START, BIN_END, cold_PI = PI) %>%
  left_join(pop_summary, by = c("CHROM", "BIN_START", "BIN_END")) %>%
  arrange(CHROM, BIN_START) %>%
  mutate(
    # strict rule: lower than every other population
    lower_than_all = cold_PI < min_other,
    
    # optional z-score / one-sided p-value
    z = (cold_PI - mean_other) / sd_other,
    p_lower = pnorm(z),   # probability of being this low or lower
    p_adj = p.adjust(p_lower, method = "BH")
  )

# find runs of >= 2 consecutive windows where cold is lower than all others
cold_runs <- 
  cold_vs_others %>%
  group_by(CHROM) %>%
  arrange(BIN_START, .by_group = TRUE) %>%
  mutate(
    consec = BIN_START == lag(BIN_START, default = first(BIN_START)) + 50000,
    new_run = cumsum(!coalesce(consec, FALSE)),
    run_id = paste(CHROM, new_run, sep = "_")
  ) %>%
  ungroup() %>%
  group_by(run_id) %>%
  mutate(
    run_len = n()
  ) %>%
  ungroup()

# keep only runs where every window is lower than all others and run length >= 2
candidate_regions <- 
  cold_runs %>%
  group_by(run_id) %>%
  filter(all(p_adj < 0.05), n() >= 2) %>%
  ungroup()

candidate_regions
```

```{r}
library(tidyverse)
library(ggplot2)

#----------------------------
# 1. Collapse candidate windows into regions
#----------------------------
regions <- 
  candidate_regions %>%
  group_by(run_id, CHROM) %>%
  summarise(
    region_start = min(BIN_START),
    region_end   = max(BIN_END),
    .groups = "drop"
  ) %>%
  mutate(
    plot_start = pmax(1, region_start - 500000),
    plot_end   = region_end + 500000,
    region = paste0(CHROM, ":", scales::comma(region_start), "-", scales::comma(region_end))
  )

#----------------------------
# 2. Combine cold + pop 1:8 into one plotting dataframe
#----------------------------
# IMPORTANT:
# use raw PI for both if you want direct comparison
wpi_cold_raw <- 
  read_table(here(outpath, "cold_50kb_windowpi.windowed.pi"), col_types = cols(
    CHROM = col_character(),
    BIN_START = col_double(),
    BIN_END = col_double(),
    N_VARIANTS = col_double(),
    PI = col_double()
  )) %>%
  filter(grepl("chr", CHROM)) %>%
  mutate(pop = "Cold")

wpi_all <- 
  bind_rows(wpi_pop, wpi_cold_raw) %>%
  mutate(midpoint = (BIN_START + BIN_END) / 2)

#----------------------------
# 3. Keep only windows that fall inside each region +/- 500 kb
#----------------------------
plot_df <- 
  regions %>%
  select(run_id, CHROM, region, plot_start, plot_end, region_start, region_end) %>%
  inner_join(wpi_all, by = "CHROM") %>%
  filter(BIN_END >= plot_start, BIN_START <= plot_end) %>%
  mutate(
    region = factor(region, levels = unique(regions$region))
  )

#----------------------------
# 4. Prepare genes overlapping each plotted region
#----------------------------
genes_plot <- 
  regions %>%
  select(run_id, CHROM, region, plot_start, plot_end) %>%
  inner_join(
    genes %>%
      rename(CHROM = scaffold),
    by = "CHROM"
  ) %>%
  #filter(!grepl("egapxtmp", gene_name)) %>% # filter out weird gene models that are probably artifacts
  filter(gene_end >= plot_start, gene_start <= plot_end) %>%
  mutate(
    gene_plot_start = pmax(gene_start, plot_start),
    gene_plot_end   = pmin(gene_end, plot_end)
  ) %>%
  mutate(gene_label = if_else(
    is.na(gene_name) & str_detect(full_name, "description="),
    str_extract(full_name, "(?<=description=)[^;]+"),
    gene_name
  )) %>%
  group_by(region) %>%
  arrange(gene_plot_start, .by_group = TRUE) %>%
  mutate(gene_y = rep(seq(-0.08, -0.20, length.out = 4), length.out = n())) %>%
  ungroup()

#----------------------------
# 5. Get a little room below for genes
#----------------------------
ymin_plot <- min(plot_df$PI, na.rm = TRUE)
ymax_plot <- max(plot_df$PI, na.rm = TRUE)
yrange <- ymax_plot - ymin_plot

gene_base <- ymin_plot - 0.22 * yrange
gene_text <- ymin_plot - 0.14 * yrange

genes_plot <- genes_plot %>%
  mutate(
    gene_y = gene_base + (gene_y - min(gene_y, na.rm = TRUE))
  )

#----------------------------
# 6. Plot
#----------------------------
png(here(plotpath, "windowed_pi_cold_vs_others.png"), width = 9, height = 30, units = "in", res = 300)
ggplot(plot_df, aes(x = midpoint, y = PI, color = pop, group = pop)) +
  geom_line(alpha = 0.7) +
  geom_rect(
    data = regions,
    aes(
      xmin = region_start,
      xmax = region_end,
      ymin = -Inf,
      ymax = Inf
    ),
    inherit.aes = FALSE,
    fill = "grey70",
    alpha = 0.15
  ) +
  # geom_segment(
  #   data = genes_plot,
  #   aes(
  #     x = gene_plot_start,
  #     xend = gene_plot_end,
  #     y = gene_y,
  #     yend = gene_y
  #   ),
  #   inherit.aes = FALSE,
  #   linewidth = 1,
  #   color = "black"
  # ) +
  # geom_text(
  #   data = genes_plot,
  #   aes(
  #     x = (gene_plot_start + gene_plot_end) / 2,
  #     y = min(plot_df$PI),
  #     label = gene_name
  #   ),
  #   inherit.aes = FALSE,
  #   angle = 45,
  #   hjust = 0,
  #   vjust = -0.3,
  #   size = 3,
  #   color = "black"
  # ) +
  facet_wrap(~region, scales = "free", ncol = 1) +
  scale_x_continuous(labels = scales::comma) +
  #coord_cartesian(ylim = c(gene_base - 0.05 * yrange, ymax_plot)) +
  labs(
    x = "Genomic position",
    y = expression(pi),
    color = "Population"
  ) +
  theme_bw() +
  theme(
    strip.text = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
dev.off()
```

