```{r}
library(here)
library(tidyverse)
library(sf)
library(terra)
library(ggpubr)
library(algatr)
library(viridis)
library(tigris)
library(raster)
library(igraph)
library(sp)
library(gridExtra)
library(wingen)
library(furrr)
source(here("analysis", "chromenv", "functions_chromenv.R"))
source(here("general_functions.R"))
outpath <- here("analysis", "chromenv", "outputs")
plotpath <- here("analysis", "chromenv", "plots")

library(gstat)
library(spatialreg)
library(spdep)
# SPATIAL
fit_vgm <- function(data, mod_vars){
  f <- as.formula(paste("Ho ~", paste(mod_vars, collapse = " + ")))
  ols_model <- lm(f, data = data)
  data$ols_resids <- residuals(ols_model)
  ols_variogram <- variogram(ols_resids ~ 1, data = data)
  ols_vgm <- fit.variogram(ols_variogram, model = vgm("Exp"))
  range <- ols_vgm %>% filter(model == "Exp") %>% pull(range)
}

fit_sem <- function(data, mod_vars, moran_test = TRUE){
  f <- as.formula(paste("Ho ~", paste(mod_vars, collapse = " + ")))

  # Fit variogram
  vgm_range <- fit_vgm(data, mod_vars)

  # Create neighborhood list
  nb <- dnearneigh(data, d1 = 0, d2 = vgm_range)
  listw <- nb2listw(nb, style = "W", zero.policy = TRUE)

  # Fit lm
  lm_model <- lm(f, data = data)
  data$lm_resids <- residuals(lm_model)

  # Test for autocorrelation based on Moran's I of residuals
  if (moran_test) {
    m <- moran.test(data$lm_resids, listw, zero.policy = TRUE)
    m <- data.frame(estimate = m$estimate[1], p.value = m$p.value)
    
    if(m$p.value < 0.05){
      sem_model <- errorsarlm(f, data = data, listw = listw, zero.policy = TRUE)
      return(list(mod = sem_model, range = vgm_range, moran = m))
    } else {
      return(list(mod = lm_model, range = NA, moran = m))
    }
  }

  sem_model <- errorsarlm(f, data = data, listw = listw, zero.policy = TRUE)
  return(list(mod = sem_model, range = vgm_range))
}
```
```{r}
# Get all annotated genes
all_genes <- get_genes()

# Get genes of interest (genes with SNPs identified by GEA that have functional importance)
genes <- get_goi()

# Get moving window Tajima's D
tjd <- get_tjd()

# Get model data.frame
mod_df <- get_moddf()
```

```{r}
pca_files <- list.files(outpath, pattern = "pca_chr.*\\.eigenvec$", full.names = TRUE)
chr_names <- str_extract(pca_files, "chr\\d+") 
mod_df <- read_csv(here(outpath, "model_df.csv"))
pca_chr <- 
  map2(pca_files, chr_names, ~read_table(.x, col_names = FALSE) %>% mutate(scaffold = .y)) %>%
  bind_rows() %>%
  rename(SampleID = X2, PC1 = X3, PC2 = X4, PC3 = X5) %>%
  right_join(get_coords(sf = TRUE)) %>%
  st_as_sf() %>%
  drop_na() %>%
  left_join(mod_df, by = "SampleID") 

# Read in the data
library(furrr)
plan(multisession, workers = 10)
files <- list.files(outpath, pattern = ".eigenvec", full.names = TRUE)
names(files) <- basename(files)
ls <- future_map(files, ~read_table(.x, col_names = FALSE), .progress = TRUE)
ls_format <- future_map(ls, ~mutate_all(.x, as.character))
plan(sequential)

files[grepl[]]

df <- 
  ls_format %>%
  bind_rows(.id = "file") %>%
  rename(FID = X1, SampleID = X2, PC1 = X3, PC2 = X4) %>%
  mutate_at(c("PC1", "PC2"), as.numeric) %>%
  mutate(file = str_remove(str_remove(file, ".eigenvec"), "pca_")) %>%
  separate(col = file, into = c("scaffold", "BIN_START", "BIN_END"), sep = "_") %>%
  left_join(mod_df) %>%
  mutate(BIN_START = as.numeric(BIN_START)) %>%
  mutate(BIN_END = BIN_START + 500000) %>%
  mutate(scaffold = factor(scaffold, levels = paste0("chr", 1:11))) 
```
```{r}
mod_vars = c("tmean_dif", "gHM", "bio1", "NDVI", "glacier", "csi_past")
scaled_sf <- 
  df %>%
  group_by(scaffold, BIN_START, BIN_END) %>%
  mutate_at(c(mod_vars, "PC1", "PC2", "Q"), scale) %>%
  ungroup() %>%
  drop_na(all_of(mod_vars), Ho) %>%
  st_as_sf(coords = c("x", "y"), crs = 4326) %>%
  st_transform(3310)

plan(multisession, workers = 10)
sem_df <-
  scaled_sf %>%
  group_by(scaffold, BIN_START) %>%
  group_split() %>%
  future_map(~{
    if (all(is.na(.x$csi_past))) return(NULL)
    result <- fit_sem(.x, c("tmean_dif", "gHM", "bio1", "NDVI", "glacier", "csi_past"))
    mod <- result$mod
    psuedo_rsq <- cor.test(mod$fitted, .x$PC1, use = "pairwise.complete.obs")$estimate^2
    result <- 
      broom::tidy(mod) %>% 
      mutate(
        BIN_START = .x$BIN_START[1], 
        scaffold = .x$scaffold[1], 
        rsq = psuedo_rsq, 
        range = result$range,
        moran = result$moran$estimate,
        moran.p = result$moran$p.value
      )
    return(result)
  }, .progress = TRUE) %>%
  compact() %>%
  bind_rows() %>%
  mutate(p_adj = p.adjust(p.value, method = "fdr")) %>%
  mutate(sig = ifelse(p_adj < 0.05, estimate, NA)) %>%
  mutate(BIN_START = as.numeric(BIN_START)) %>%
  filter(term != "(Intercept)", term != "lambda") %>%
  rename(var = term) %>% 
  mutate(BIN_END = BIN_START + 500000)
plan(sequential)

summary_tb <-
  sem_df %>%
  group_by(var) %>%
  summarize(`Prop(p < 0.05)` = round(mean(p_adj < 0.05), 2), `Mean effect` = round(mean(estimate), 2)) %>%
  arrange(desc(abs(`Mean effect`)), desc(`Prop(p < 0.05)`)) %>%
  mutate(var = make_pretty_names(var)) %>%
  rename(Variable = var) %>%
  pivot_longer(-Variable) %>%
  mutate(Variable = factor(Variable, levels = rev(unique(Variable))))

colors <- get_colors()
names(colors) <- make_pretty_names(names(colors))

pdf(here(plotpath, "sem_pca_summary.pdf"), width = 5, height = 3)
ggplot(summary_tb, aes(x = name, y = Variable)) +
  geom_tile(aes(fill = Variable), color = "white", lwd = 1) +
  geom_text(aes(label = value)) +
  scale_fill_manual(values = colors) +
  theme_minimal() +
  scale_x_discrete(position = "top") +
  labs(x = "", y = "") +
  theme(
    legend.position = "none",
    panel.grid = element_blank()
  )
dev.off()

scaff_rect <- get_scaff_rect(sem_df)
pdf(here(plotpath, "sem_pca.pdf"), width = 12, height = 9)
ggplot(sem_df) +
  geom_rect(aes(xmin = BIN_START, xmax = BIN_END, ymin = 0, ymax = 1, fill = log(range))) +
  geom_rect(data = scaff_rect, aes(xmin = BIN_START, xmax = BIN_END), ymin = 0, ymax = 1, col = "black", fill = NA) +
  scaffold_theme() +
  ggtitle("log(variogram range)") +
  scale_fill_viridis_c(option = "magma")

ggplot(sem_df) +
  geom_rect(aes(xmin = BIN_START, xmax = BIN_END, ymin = 0, ymax = 1, fill = moran)) +
  geom_rect(data = scaff_rect, aes(xmin = BIN_START, xmax = BIN_END), ymin = 0, ymax = 1, col = "black", fill = NA) +
  scaffold_theme() +
  ggtitle("Moran's I") +
  scale_fill_viridis_c(option = "magma")

plotchrom_all(sem_df)
dev.off()
```


nonsyn <- 
  get_nonsyn() %>% 
  filter(scaffold %in% unique(sem_df$scaffold))

# Get genes
genes <- 
  get_genes() %>%
  mutate(BIN_START = floor(start / 500000) * 500000) %>%
  mutate(BIN_END = BIN_START + 500000) %>%
  filter(scaffold %in% unique(sem_df$scaffold)) %>%
  mutate(scaffold = factor(scaffold, levels = unique(sem_df$scaffold)))

genes_bin <-
  genes %>%
  group_by(scaffold, BIN_START, BIN_END) %>%
  summarise(ngenes = n()) %>%
  mutate(scaffold = factor(scaffold, levels = unique(sem_df$scaffold)))

# Create bins for BIN_START
nonsyn <- 
  get_nonsyn() %>%
  mutate(BIN_START = floor(position / 500000) * 500000)  %>%
  mutate(BIN_END = BIN_START + 500000) %>%
  filter(scaffold %in% unique(sem_df$scaffold)) %>%
  mutate(scaffold = factor(scaffold, levels = unique(sem_df$scaffold)))

# Calculate number of nonsynonymous mutations per bin of 100k
nonsyn_bin <- 
  nonsyn %>% 
  group_by(scaffold, BIN_START, BIN_END) %>% 
  summarise(nonsyn = n()) %>%
  filter(scaffold %in% unique(sem_df$scaffold))

# Create bins for BIN_START
syn <- 
  get_syn() %>%
  mutate(BIN_START = floor(position / 500000) * 500000)  %>%
  mutate(BIN_END = BIN_START + 500000) %>%
  filter(scaffold %in% unique(sem_df$scaffold)) %>%
  mutate(scaffold = factor(scaffold, levels = unique(sem_df$scaffold)))

# Calculate number of nonsynonymous mutations per bin of 100k
syn_bin <- 
  syn %>% 
  group_by(scaffold, BIN_START, BIN_END) %>% 
  summarise(syn = n()) %>%
  filter(scaffold %in% unique(sem_df$scaffold))

syn_nonsyn <-
  full_join(syn_bin, nonsyn_bin, by = c("scaffold", "BIN_START", "BIN_END")) %>%
  # DON"T MAKE SYN 0 OR YOU WILL GET INFINITY
 # mutate(syn = ifelse(is.na(syn), 0, syn)) %>%
  mutate(nonsyn = ifelse(is.na(nonsyn), 0, nonsyn)) %>%
  mutate(ratio = nonsyn / syn)

pdf(here(plotpath, "nonsyn.pdf"), width = 12, height = 8)
ggplot(sem_df) +
  geom_rect(aes(xmin = BIN_START, xmax = BIN_END, ymin = 0, ymax = max(syn_nonsyn$ratio, na.rm = TRUE), fill = rsq), color = NA) +
  geom_line(data = syn_nonsyn, aes(x = BIN_START, y = ratio), col = "white", lwd = 0.5) +
  scaffold_theme() +
  ggtitle("RSQ vs NONSYN:SYN") +
  scale_fill_viridis_c(option = "inferno")

ggplot(sem_df) +
  geom_rect(aes(xmin = BIN_START, xmax = BIN_END, ymin = 0, ymax = max(log(genes_bin$ngenes)), fill = rsq), color = NA) +
  geom_line(data = genes_bin, aes(x = BIN_START, y = log(ngenes)), col = "white", lwd = 0.5) +
  scaffold_theme() +
  ggtitle("RSQ vs gene counts") +
  scale_fill_viridis_c(option = "inferno")
  
ggplot(sem_df) +
  geom_rect(aes(xmin = BIN_START, xmax = BIN_END, ymin = 0, ymax = max(log(nonsyn_bin$nonsyn)), fill = rsq), color = NA) +
  geom_line(data = nonsyn_bin, aes(x = BIN_START, y = log(nonsyn)), col = "white", lwd = 0.5) +
  scaffold_theme() +
  ggtitle("RSQ vs non-synonymous counts") +
  scale_fill_viridis_c(option = "inferno")

ggplot(nonsyn_bin) +
  geom_rect(aes(xmin = BIN_START, xmax = BIN_END, ymin = 0, ymax = 1, fill = (nonsyn)), color = NA) +
  scaffold_theme() +
  ggtitle("Non-synonymous counts") +
  scale_fill_viridis_c(option = "inferno")

ggplot(nonsyn_bin) +
  geom_rect(aes(xmin = BIN_START, xmax = BIN_END, ymin = 0, ymax = 1, fill = log(nonsyn)), color = NA) +
  scaffold_theme() +
  ggtitle("Non-synonymous counts") +
  scale_fill_viridis_c(option = "inferno")

ggplot(genes_bin) +
  geom_rect(aes(xmin = BIN_START, xmax = BIN_END, ymin = 0, ymax = 1, fill = log(ngenes)), color = NA) +
  scaffold_theme() +
  ggtitle("Gene counts") +
  scale_fill_viridis_c(option = "inferno")

ggplot(genes_bin) +
  geom_rect(aes(xmin = BIN_START, xmax = BIN_END, ymin = 0, ymax = 1, fill = (ngenes)), color = NA) +
  scaffold_theme() +
  ggtitle("Gene counts") +
  scale_fill_viridis_c(option = "inferno")
  
dev.off()

test <-
  syn_nonsyn %>% 
  filter(scaffold == "chr5") %>% 
  ungroup() %>%
  filter(ratio == max(ratio, na.rm = TRUE))

nonsyn %>% 
  filter(scaffold == "chr5", position > test$BIN_START, position < test$BIN_END)

genes %>%
  filter(grepl("\\[", full_name)) %>%
  filter(scaffold == "chr5", start > test$BIN_START, start < test$BIN_END)

csi_past_df <-
  df %>%
  group_by(scaffold, BIN_START) %>%
  group_split() %>%
  map(~{
    if (all(is.na(.x$csi_past))) return(NULL)
    result <- cor.test(.x$csi_past, .x$Ho, use = "pairwise.complete.obs")
    tibble(BIN_START = .x$BIN_START[1], scaffold = .x$scaffold[1], r = result$estimate, p = result$p.value, var = "csi_past")
  }, .progress = TRUE) %>%
  compact() %>%
  bind_rows() %>%
  mutate(p_adj = p.adjust(p, method = "fdr")) %>%
  mutate(r_sig = ifelse(p_adj < 0.05, r, NA)) %>%
  mutate(BIN_START = as.numeric(BIN_START))

tmean_dif_df <-
  df %>%
  group_by(scaffold, BIN_START) %>%
  group_split() %>%
  map(~{
    if (all(is.na(.x$csi_past))) return(NULL)
    result <- cor.test(.x$tmean_dif, .x$Ho, use = "pairwise.complete.obs")
    tibble(BIN_START = .x$BIN_START[1], scaffold = .x$scaffold[1], r = result$estimate, p = result$p.value, var = "tmean_dif")
  }, .progress = TRUE) %>%
  compact() %>%
  bind_rows() %>%
  mutate(r_sig = ifelse(p < 0.05, r, NA)) %>%
  mutate(BIN_START = as.numeric(BIN_START))

  
pdf(here(plotpath, "windowed_env.pdf"), width = 12, height = 8)
ggplot(csi_past_df) +
  geom_vline(aes(xintercept = BIN_START, col = r_sig), alpha = 1) +
  scaffold_theme() +
  scale_color_gradient2(midpoint = 0, low = "cyan3", mid = "white", high = "orange2", limits = c(-0.3, 0.3))

ggplot(tmean_dif_df) +
  geom_vline(aes(xintercept = BIN_START, col = r_sig), alpha = 1) +
  scaffold_theme() +
  scale_color_gradient2(midpoint = 0, low = "cyan3", mid = "white", high = "orange2", limits = c(-0.3, 0.3))
dev.off()


# Define the variables to test

cor.test2 <- function(x, y = "Ho", data){
  result <- cor.test(data[[x]], data[[y]], use = "pairwise.complete.obs")
  tibble(r = result$estimate, p = result$p.value, var = x)
}

names(mod_df)
vars <- c("tmean_dif", "gHM", "lgm_lig", "cur_lgm", "bio1", "csi_past")

var_df <-
  df %>%
  # MAKE CUR_LGM AND LGM_LIG ABSOLUTE VALUES SO THEY CAN BE COMPARED
  mutate(cur_lgm = abs(cur_lgm), lgm_lig = abs(lgm_lig)) %>%
  group_by(scaffold, BIN_START) %>%
  group_split() %>%
  map(~{
    if (all(is.na(.x$csi_past))) return(NULL)
    result <-
      map_dfr(vars, \(var) cor.test2(x = var, y = "Ho", data = .x)) %>%
      mutate(BIN_START = .x$BIN_START[1], scaffold = .x$scaffold[1])
  }, .progress = TRUE) %>%
  compact() %>%
  bind_rows() %>%
  mutate(p_adj = p.adjust(p, method = "fdr")) %>%
  mutate(r_sig = ifelse(p_adj < 0.05, r, NA)) %>%
  mutate(BIN_START = as.numeric(BIN_START))



csi_past_df <- 
  var_df %>%
  filter(!(var %in% c("lgm_lig", "cur_lgm"))) 

# For each scaffold/bin, determine which variable has the highest correlation with Ho
top <-
  csi_past_df %>%
  ungroup() %>%
  group_by(scaffold, BIN_START) %>%
  mutate(r_abs = abs(r_sig)) %>%
  select(scaffold, BIN_START, var, r_abs) %>%
  mutate(top_r = case_when(all(is.na(r_abs)) ~ NA, r_abs == max(r_abs, na.rm = TRUE) ~ var, TRUE ~ NA)) %>%
  drop_na(top_r)

neg <-
  csi_past_df %>%
  ungroup() %>%
  filter(r_sig < 0) %>%
  group_by(scaffold, BIN_START) %>%
  mutate(r_abs = abs(r_sig)) %>%
  select(scaffold, BIN_START, var, r_abs) %>%
  mutate(top_r = case_when(all(is.na(r_abs)) ~ NA, r_abs == max(r_abs, na.rm = TRUE) ~ var, TRUE ~ NA)) %>%
  drop_na(top_r)

unique <-
  csi_past_df %>%
  ungroup() %>%
  group_by(scaffold, BIN_START) %>%
  mutate(r_abs = abs(r_sig)) %>%
  select(scaffold, BIN_START, var, r_abs) %>%
  # Filter for bins with only one significant correlation
  filter(sum(!is.na(r_abs)) == 1) %>%
  # Drop bins with no significant correlations so you only get the one significant var
  drop_na()

unique_neg <-
  csi_past_df %>%
  ungroup() %>%
  group_by(scaffold, BIN_START) %>%
  select(scaffold, BIN_START, var, r_sig) %>%
  # Filter for bins with only one significant correlation
  filter(sum(!is.na(r_sig)) == 1) %>%
  # Only keep negative correlations
  filter(r_sig < 0) %>%
  # Drop bins with no significant correlations so you only get the one significant var
  drop_na()

pdf(here(plotpath, "windowed_env_top.pdf"), width = 12, height = 8)
ggplot(top) +
  geom_vline(aes(xintercept = BIN_START, col = top_r), alpha = 1) +
  geom_vline(data = genes, aes(xintercept = start), col = "black", lwd = 0.5, lty = "dashed") +
  scaffold_theme() +
  ggtitle("Top correlation with Ho")

ggplot(neg) +
  geom_vline(aes(xintercept = BIN_START, col = top_r), alpha = 1) +
  geom_vline(data = genes, aes(xintercept = start), col = "black", lwd = 0.5, lty = "dashed") +
  scaffold_theme() +
  ggtitle("Top negative correlation with Ho")

ggplot(unique) +
  geom_vline(aes(xintercept = BIN_START, col = var), alpha = 1) +
  geom_vline(data = genes, aes(xintercept = start), col = "black", lwd = 0.5, lty = "dashed") +
  scaffold_theme() +
  ggtitle("Unique significant correlation with Ho")

ggplot(unique_neg) +
  geom_vline(aes(xintercept = BIN_START, col = var), alpha = 1) +
  geom_vline(data = genes, aes(xintercept = start), col = "black", lwd = 0.5, lty = "dashed") +
  scaffold_theme() +
  ggtitle("Unique negative significant correlation with Ho")
dev.off()



# Test if you can disentangel cur_lgm and lgm_lig
change_df <- 
  var_df %>%
  filter(!(var %in% "csi_past")) 

# For each scaffold/bin, determine which variable has the highest correlation with Ho
top <-
  change_df %>%
  ungroup() %>%
  group_by(scaffold, BIN_START) %>%
  mutate(r_abs = abs(r_sig)) %>%
  select(scaffold, BIN_START, var, r_abs) %>%
  mutate(top_r = case_when(all(is.na(r_abs)) ~ NA, r_abs == max(r_abs, na.rm = TRUE) ~ var, TRUE ~ NA)) %>%
  drop_na(top_r)

neg <-
  change_df %>%
  ungroup() %>%
  filter(r_sig < 0) %>%
  group_by(scaffold, BIN_START) %>%
  mutate(r_abs = abs(r_sig)) %>%
  select(scaffold, BIN_START, var, r_abs) %>%
  mutate(top_r = case_when(all(is.na(r_abs)) ~ NA, r_abs == max(r_abs, na.rm = TRUE) ~ var, TRUE ~ NA)) %>%
  drop_na(top_r)

unique <-
  change_df %>%
  ungroup() %>%
  group_by(scaffold, BIN_START) %>%
  mutate(r_abs = abs(r_sig)) %>%
  select(scaffold, BIN_START, var, r_abs) %>%
  # Filter for bins with only one significant correlation
  filter(sum(!is.na(r_abs)) == 1) %>%
  # Drop bins with no significant correlations so you only get the one significant var
  drop_na()

unique_neg <-
  change_df %>%
  ungroup() %>%
  group_by(scaffold, BIN_START) %>%
  select(scaffold, BIN_START, var, r_sig) %>%
  # Filter for bins with only one significant correlation
  filter(sum(!is.na(r_sig)) == 1) %>%
  # Only keep negative correlations
  filter(r_sig < 0) %>%
  # Drop bins with no significant correlations so you only get the one significant var
  drop_na()

unique_neg %>% group_by(var) %>% count()

tjd_mean <- tjd %>% group_by(scaffold, BIN_START) %>% summarise(TajimaD = mean(TajimaD, na.rm = TRUE))

pdf(here(plotpath, "windowed_env_disentangle_paleochange.pdf"), width = 12, height = 8)
ggplot(top) +
  geom_vline(aes(xintercept = BIN_START, col = top_r), alpha = 1) +
  geom_vline(data = genes, aes(xintercept = start), col = "black", lwd = 0.5, lty = "dashed") +
  #geom_line(data = tjd_mean, aes(x = BIN_START, y = TajimaD), col = "black", lwd = 0.5) +
  scaffold_theme() +
  ggtitle("Top correlation with Ho")

ggplot(neg) +
  geom_vline(aes(xintercept = BIN_START, col = top_r), alpha = 1) +
  geom_vline(data = genes, aes(xintercept = start), col = "black", lwd = 0.5, lty = "dashed") +
  scaffold_theme() +
  ggtitle("Top negative correlation with Ho")

ggplot(unique) +
  geom_vline(aes(xintercept = BIN_START, col = var), alpha = 1) +
  geom_vline(data = genes, aes(xintercept = start), col = "black", lwd = 0.5, lty = "dashed") +
  scaffold_theme() +
  ggtitle("Unique significant correlation with Ho")

ggplot(unique_neg) +
  geom_vline(aes(xintercept = BIN_START, col = var), alpha = 1) +
  geom_vline(data = genes, aes(xintercept = start), col = "black", lwd = 0.5, lty = "dashed") +
  scaffold_theme() +
  ggtitle("Unique negative significant correlation with Ho")
dev.off()


# CHROMOSOME 7
chr7 <- 
  lm_df %>% 
  left_join(rename(avg_het, avg_Ho = Ho)) %>%
  filter(scaffold == "chr7") %>%
  filter(rsq < 0.1, avg_Ho > 0.1)

start_end <- range(chr7$BIN_START)
genes <- get_genes()
chr7_genes <- 
  genes %>% 
  filter(scaffold == "chr7", start > start_end[1], start < start_end[2]) %>%
  # Drop genes with no ID
  filter(grepl("\\[", full_name)) %>%
  mutate(gene_size = end - start) %>%
  as_tibble()
chr7_genes %>% arrange(desc(gene_size)) %>% pull(full_name)

chr7_nonsyn <- get_nonsyn() %>% filter(scaffold == "chr7", position > start_end[1], position < start_end[2])
chr7_nonsyn %>% count(csq)

# Count the number of nonsynonymous mutations per gene

library(fuzzyjoin)
chr7_nonsyn_gene <- fuzzy_inner_join(
  chr7_genes,
  chr7_nonsyn,
  by = c("start" = "position", "end" = "position"),
  match_fun = list(`<=`, `>=`)
) %>%
  group_by(full_name, start, end) %>%
  summarise(nonsyn_count = n(), .groups = "drop") %>%
  arrange(desc(nonsyn_count))

chr7_nonsyn_gene %>% dplyr::select(full_name, nonsyn_count) %>% arrange(desc(nonsyn_count))

pdf(here(plotpath, "chr7.pdf"), width = 12, height = 8)
ggplot(lm_df) +
  #geom_rect(aes(xmin = BIN_START, xmax = BIN_END, ymin = 0, ymax = max(avg_het$Ho, na.rm = TRUE), fill = rsq)) +
  #geom_rect(aes(xmin = start_end[1], xmax = start_end[2], ymin = 0, ymax = max(avg_het$Ho, na.rm = TRUE)), fill = "red", alpha = 0.5) +
  geom_vline(aes(xintercept = BIN_START, col = rsq)) +
  geom_line(data = avg_het, aes(x = BIN_START, y = Ho), col = "white", lwd = 0.5) +
  scaffold_theme() +
  ggtitle("RSQ") +
  scale_fill_viridis_c(option = "magma")
dev.off()


# LD window
ld_chr5 <- read_table(here("analysis", "chromenv", "ld_chr5.ld"))
ld_bin <- 
  ld_chr5 %>% 
  mutate(BIN_START = floor(BP_A / 500000) * 500000) %>% 
  mutate(BIN_END = BIN_START + 500000) %>%
  group_by(BIN_START, BIN_END) %>% 
  summarise(mean = mean(R2, na.rm = TRUE)) %>%
  mutate(scaffold = "chr5")

pdf(here(plotpath, "ld.pdf"), width = 12, height = 2)
ggplot(ld_bin) +
  geom_rect(aes(xmin = BIN_START, xmax = BIN_END, ymin = 0, ymax = 1, fill = mean), color = NA) +
  scaffold_theme() +
  scale_fill_viridis_c(option = "magma") +
  ggtitle("LD")
dev.off()