```{r}
library("rpaleoclim")
library("here")
library("tidyverse")
library("ggpubr")
library("terra")
library("sf")
library("gridExtra")
library("rpaleoclim")
library("spdep")
library("spatialreg")
library("MuMIn")

source("general_functions.R")
source(here("analysis", "genetic_diversity", "genetic_diversity.R"))
source(here("analysis", "sdm", "paleovars.R"))

source(here("general_functions.R"))
source(here("analysis", "genetic_diversity", "genetic_diversity.R"))
source(here("ccgp_test", "get_ccgp_data.R"))

ca <- get_ca()
pdfpath <- here("analysis", "ccgp_test", "pdfs")
```

# Get heterozygosity Data
```{r}
het <- get_ccgp_het()
coords <- get_ccgp_coords(sf = TRUE)

df <- 
  inner_join(het, coords, by = c("project", "SampleID")) %>%
  drop_na(project, SampleID)

taxa_df <- 
  data.frame(project = df$project) %>%
  mutate(genus = gsub("[^A-Za-z]", "", project)) %>%
  distinct()

library(taxize)
taxize_options(ncbi_sleep = 0.4)
a <- classification(taxa_df$genus, db = "ncbi")
b <- map(1:length(a), ~a[[.x]])
names(b) <- taxa_df$project
taxa_df <- 
  bind_rows(b, .id = "project") %>%
  filter(rank %in% c("kingdom", "phylum", "class", "order", "family"))
```

# Hypothesis 1: Genetic diversity explained by paleoclimate

```{r}
# Get paleoclimate data
# Created using paleoclimate.Rmd
paleovars <- rast(here("data", "env", "paleoclim", "paleoclim.tif"))

# Get important var for change in climate (based on paleoclimate.Rmd)
bio1_dif_lig <- paleovars$cur_bio_1 - paleovars$lig_bio_1
paleoclim_mod_df <- 
  data.frame(
    coords,
    paleo_change = terra::extract(bio1_dif_lig, coords, ID = FALSE)[,1]
    )
```

# Hypothesis 2: Genetic diversity explained by past climate stability

### Climate stability (CSI)

Original paper: 
Herrando-Moraira, S., Nualart, N., Galbany-Casals, M. et al. Climate Stability Index maps, a global high resolution cartography of climate stability from Pliocene to 2100. Sci Data 9, 48 (2022). https://doi.org/10.1038/s41597-022-01144-5

Note: Bioclimatic variables do not contain Bio 2 or Bio 5, but they do contain other variables correlated with those variables

```{r, fig.width = 10, fig.height = 2.5}
csi_past <- rast(here("data", "env", "stability", "Layers", "past", "csi_past.tif"))
csi_past <- crop(csi_past, ca)
csi_past <- mask(csi_past, ca)

files <- list.files(here("data", "env", "stability", "Layers", "future"), full.names = TRUE, pattern = "csi*")
csi_future <- rast(map(files, rast))
csi_future <- crop(csi_future, ca)
csi_future <- mask(csi_future, ca)

csi_stack <- c(csi_past, csi_future)
names(csi_stack) <- c("Past",  "SSP 1-2.6", "SSP 2-4.5", "SSP 3-7.0", "SSP 5-8.5")
writeRaster(csi_stack, here("data", "env", "stability", "csi.tif"), overwrite = TRUE)
csi_df <- 
  terra::as.data.frame(csi_stack, xy = TRUE) %>% 
  pivot_longer(-c(x, y), 
  names_to = "model", 
  values_to = "CSI")
```

```{r, fig.height = 4, fig.width = 4.5}
csi_vals <- extract(csi_past, coords)
df_past <- data.frame(csi_past = csi_vals[,1], coords)
```

```{r, fig.height = 4, fig.width = 16}
csi_vals <- extract(csi_future, coords, ID = FALSE)
df_future <- 
  data.frame(1 - csi_vals, coords) %>% 
  pivot_longer(starts_with("csi"), names_to = "future_model", values_to = "csi") %>%
  mutate(model_name = case_when(
    future_model == "csi_future_ssp126" ~ "SSP 1-2.6",
    future_model == "csi_future_ssp245" ~ "SSP 2-4.5",
    future_model == "csi_future_ssp370" ~ "SSP 3-7.0",
    future_model == "csi_future_ssp585" ~ "SSP 5-8.5"
  ))
```

```{r}
# Do 1 - to make the values more intuitive
csi_mod_df <- data.frame(csi = 1 - extract(csi_past, coords, ID = FALSE), coords)
```

# Hypothesis 3: Genetic diversity explained by admixture
```{r}
```

# Hypothesis 4: Genetic diversity explained by current environment 
```{r}
env_data <- rast(here("data", "env", "california_env_stack.tif"))
env_data$BurnDate <- as.numeric(env_data$BurnDate > 0)
env_mod_df <- data.frame(terra::extract(env_data, coords, ID = FALSE))
env_mod_df$SampleID <- coords$SampleID
env_mod_df$project <- coords$project
```

```{r}
coords_buffer <- st_buffer(coords, 10000)
gHM_buffer <- extract(env_data[["gHM"]], coords_buffer, fun = mean, na.rm = TRUE)

env_mod_df$gHM <- gHM_buffer$gHM
env_mod_df$buffer <- 10000
```

# Hypothesis 5: Genetic diversity explained by glaciation
```{r}
lgm_glaciers <- st_read(here("data", "env", "Glaciotect_NA_lgm", "shp", "LGM_maxextent.shp"))
# Note: have to first intersect and then transform otherwise you get a node error
lgm_glaciers <- st_intersection(lgm_glaciers, st_transform(ca, crs = st_crs(lgm_glaciers)))
lgm_glaciers <- st_transform(lgm_glaciers, crs = st_crs(ca))
# Add a new column to coords indicating whether each point is inside the lgm shapefile
glacier_mod_df <- 
  coords %>%
  mutate(glacier = as.numeric((lengths(st_within(st_transform(coords, st_crs(lgm_glaciers)), lgm_glaciers)) > 0)))
```

# Hypothesis 6: Genetic diversity explained by recent climate change
Download PRISM mean temperature data from historical data: https://prism.oregonstate.edu/historical/

Data can be downloaded for all years using:
wget -r ftp://prism.oregonstate.edu/monthly/tmean/

mv prism.oregonstate.edu/monthly/tmean/* .

for year in {1895..2024}
do
  unzip '/media/wanglab/798f0e01-89d1-4d0f-8ed8-ef323be70ab91/Anusha/GitHub/ccgpscelop/data/env/prism/TEMP/'$year'/*.zip' -d '/media/wanglab/798f0e01-89d1-4d0f-8ed8-ef323be70ab91/Anusha/GitHub/ccgpscelop/data/env/prism/TEMP/'$year
done

```{r}
year = 1895

# Get prism data
get_prism <- function(year  = 1895, output = "df") {
  # Create PRISM path
  prism_path <- here("data", "env", "prism", "PRISM_tmean_stable_4kmM3", year)

  # List annual file
  file <- list.files(prism_path, pattern = paste0(year, "_*.bil.bil$"), full.names = TRUE)

  # Load data
  raster <- terra::rast(file)

  if (output == "raster"){
    return(raster)
  } else {
    vals <- terra::extract(raster, st_transform(coords, st_crs(raster)), ID = FALSE)
    names(vals) <- "tmean"
    df <- data.frame(year = year, tmean = vals, SampleID = coords$SampleID)
    return(df)
  }
}

# Loop over the years of interest
tmean_his <- map(1895:1905, get_prism, .progress =TRUE) %>% bind_rows()  
tmean_his <- tmean_his %>% group_by(SampleID) %>% summarise(tmean_his = mean(tmean, na.rm = TRUE))
tmean_cur <- map(2010:2020, ~get_prism(year = .x), .progress =TRUE) %>% bind_rows()  
tmean_cur <- tmean_cur %>% group_by(SampleID) %>% summarise(tmean_cur = mean(tmean, na.rm = TRUE))

tmean_his_r <- map(1895:1905, get_prism, .progress =TRUE, output= "raster") %>% rast() %>% mean()
tmean_cur_r <- map(2010:2020, get_prism, .progress =TRUE, output= "raster") %>% rast() %>% mean()
tmean_dif_r <- tmean_cur_r - tmean_his_r
tmean_dif_r <- crop(tmean_dif_r, ca)
tmean_dif_r <- mask(tmean_dif_r, st_transform(ca, st_crs(tmean_dif_r)))

tmean_mod_df <- 
  left_join(tmean_his, tmean_cur) %>% 
  mutate(tmean_dif = tmean_cur - tmean_his) %>%
  right_join(coords) %>% 
  dplyr::select(SampleID, project, tmean_dif)
```

```{r}
env_list <- list(
  paleo_change = bio1_dif_lig, 
  csi_past = 1 - csi_past, 
  env_data = env_data, 
  glacier = lgm_glaciers, 
  tmean_dif = tmean_dif_r
  )
```

# Models
```{r}
mod_vars <- c("NDVI", "gHM", "paleo_change", "CHELSA_bio1_1981.2010_V.2.1", "csi_past", "glacier", "tmean_dif")

mod_df <- 
  df %>% 
  left_join(env_mod_df) %>%
  left_join(paleoclim_mod_df) %>%
  #left_join(q_mod_df) %>%
  left_join(csi_mod_df) %>%
  left_join(st_drop_geometry(glacier_mod_df)) %>%
  left_join(tmean_mod_df) %>%
  dplyr::select(Ho, SampleID, geometry, project, all_of(mod_vars)) %>%
  drop_na(Ho) %>%
  # REMOVE CHANNEL ISLAND
  filter(SampleID != "Scelocci_CCGPMC_MW01-3-14")
  
# Count NAs in each column
have_nas <- colSums(is.na(mod_df))[colSums(is.na(mod_df)) > 0]
have_nas

# NDVI has some NA values (< 3), impute to the value of the closest sample
mod_sf <- mod_df %>% st_as_sf() %>% st_transform(3310)
close_impute <- function(var, mod_sf){
  map_dbl(1:nrow(mod_sf), \(i){ 
    if (!is.na(st_drop_geometry(mod_sf[i, var]))) return(mod_sf[i, var][[1]])
    # Get the closest sample that is not NA
    nearest <- st_nearest_feature(mod_sf[i,], drop_na(mod_sf[-i,]), 1)
    # Get the value of the closest sample
    return(drop_na(mod_sf)[nearest, var][[1]])
  }) 
}

vars <- names(have_nas)
names(vars) <- vars
mod_sf_imputed <- map(vars, ~close_impute(.x, mod_sf), .progress = TRUE) 
mod_df_imputed <- bind_cols(dplyr::select(mod_df, -all_of(vars)), bind_cols(mod_sf_imputed))

#  Drop the rest of NAs
mod_df_imputed <- mod_df_imputed %>% drop_na(all_of(mod_vars))

# Confirm no NAs:
colSums(is.na(mod_df_imputed))[colSums(is.na(mod_df_imputed)) > 0]

# Scale the variables (except glacier)
mod_df_imputed_scaled <- 
  mod_df_imputed %>% 
  mutate(across(c(all_of(mod_vars), -glacier), scale))

# Convert to sf
mod_sf <- st_as_sf(mod_df) %>% st_transform(3310)
```

# Model selection

```{r}
get_best <- function(project, mod_df_imputed_scaled){
  sub_df <- mod_df_imputed_scaled %>% filter(project == !!project)

  # Full model
  # Note: included both regular and poly gHM to see which was the better fit using dredge
  mod <- lm(Ho ~  NDVI + paleo_change + poly(gHM, 2) + gHM + CHELSA_bio1_1981.2010_V.2.1 + csi_past + tmean_dif, data = sub_df, na.action = na.fail)

  summary(mod)

  # Use dredge for model selection
  dredge_result <- dredge(mod)

  # Select the models with delta AIC < 2
  best_models <- get.models(dredge_result, subset = delta < 2)

  # Look at the best models
  # Get the best model
  best_model <- get.models(dredge_result, subset = delta < 2)[[1]]

  return(best_model)
}
```
```{r}
projects <- unique(mod_df_imputed_scaled$project)
names(projects) <- projects

best_models <- map(projects, ~get_best(.x, mod_df_imputed_scaled), .progress = TRUE)

summaries <- map(best_models, broom::tidy) %>% bind_rows(.id = "project")

adjRsqr <- map_dbl(best_models, ~summary(.x)$adj.r.squared)
hist(adjRsqr)
mean(adjRsqr > 0.3, na.rm = TRUE)

sigvars <- 
  summaries %>%
  filter(term != "(Intercept)") %>%
  rename(var = term) %>%
  filter(p.value < 0.05)

count(sigvars, var)

plot_sig <- function(var, cordf, mod_df_imputed_scaled){
  sub_sig <- cordf %>% filter(var == !!var)
  sub_df <- mod_df_imputed_scaled %>% filter(project %in% sub_sig$project)
  if (grepl("gHM", var)) var <- "gHM"
  if (grepl("glacier", var)) var <- "glacier"
  plt <-
    ggplot(sub_df, aes(x = .data[[var]], y = Ho)) +
    geom_point() +
    geom_smooth(method = "lm") +
    labs(x = var, y = "Ho") +
    facet_wrap(~project, scales = "free") +
    stat_cor(label.y = 0) +
    theme_classic() 
  return(plt)
}

plts <- map(unique(sigvars$term), ~plot_sig(.x, sigvars, mod_df_imputed_scaled))

pdf(here("ccgp_test", "ccgp_mods.pdf"), width = 10, height = 10)
walk(plts, ~print(.x))
dev.off()
```

```{r}
combos <- 
  expand.grid(
    project = unique(mod_df_imputed_scaled$project),
    var = mod_vars
  )

r_df <- pmap(combos, \(project, var){
  sub_df <- mod_df_imputed_scaled %>% filter(project == !!project)
  result <- cor.test(sub_df$Ho, sub_df[[var]])
  return(data.frame(r = result$estimate, p = result$p.value, project = project, var = var))
}) %>% bind_rows()

sig_r <- 
  r_df %>% 
  filter(p < 0.05) %>%
  mutate(var = as.character(var))

count(sig_r, var)

plts <- map(unique(sig_r$var), ~plot_sig(.x, sig_r, mod_df_imputed_scaled))

pdf(here("ccgp_test", "ccgp_rsig.pdf"), width = 10, height = 10)
walk(plts, ~print(.x))
dev.off()
```

```{r}
sign_df <- 
  r_df %>% 
  mutate(sign = 
    case_when(
      r > 0 & p < 0.05 ~ "Positive",
      r < 0 & p < 0.05 ~ "Negative",
      TRUE ~ "No relationship")
  ) %>%
  select(project, sign, var) %>%
  mutate(var = as.character(var))

gg_df <-
  mod_df_imputed_scaled %>% 
  group_by(project) %>%
  mutate(Ho = scale(Ho)) %>%
  ungroup()%>%
  pivot_longer(all_of(mod_vars), names_to = "var") %>%
  left_join(sign_df) %>%
  left_join(filter(taxa_df, rank == "class"))

pdf(here("ccgp_test", "ccgp_sign.pdf"), width = 10, height = 20)
ggplot(gg_df, aes(x = value, y = Ho, group = project)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm", se = FALSE, alpha = 0.5) +
  facet_grid(var~sign)
dev.off()

gg_df2 <- gg_df %>% select(-SampleID, -geometry, -Ho) %>% distinct()

ggplot(gg_df2, aes(x = sign, fill = name)) +
  geom_bar() +
  facet_wrap(~var, scales = "free")

mod2_df <- filter(gg_df2, var == "paleo_change") %>% drop_na(sign, name)
mod2_df$sign <- as.factor(mod2_df$sign)  # Ensure 'sign' is a factor
mod2_df$name <- as.factor(mod2_df$name)  # Ensure 'name' is a factor

# Create a contingency table
contingency_table <- table(mod2_df$sign, mod2_df$name)

# Perform Chi-square test
chi_square_result <- chisq.test(contingency_table)
chi_square_result
contingency_table <- table(mod2_df$sign, mod2_df$name)
contingency_table
```
