# Adaptive index

```{r setup, include=FALSE}
# Set global options for caching
knitr::opts_chunk$set(
  cache = TRUE,
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  result = FALSE)

library(algatr)
library(here)
library(tidyverse)
library(RColorBrewer)
library(sf)
library(terra)
library(raster)
library(vegan)
library(cowplot)
library(tidyterra)

source(here("analysis", "adaptive", "adaptive_index.R"))
source(here("general_functions.R"))
```

## Specify model to examine

```{r}
gea_method = "bio1ndvi" # "pca" or "bio1ndvi"
prefix = "58-Sceloporus_bio1ndvi_gea"
path = here("analysis", "adaptive", "outputs")
```

## 1. Get input files

```{r}
# If `mod` is not in memory, run:
dat_mod <- import_rda(output_path = path, suffix = paste0(prefix, "_mod"), rds_obj = TRUE)
# If `mod` is still in memory, can run:
dat_mod <- list(mod = mod, biplot = data.frame(mod$CCA$biplot), 
                eig = data.frame(mod$CCA$eig), 
                scaledload = vegan::scores(mod, choices = 1:ncol(mod$CCA$v), tidy = TRUE), 
                unscaledload = vegan::scores(mod, choices = 1:ncol(mod$CCA$v), tidy = TRUE, scaling = 0))

# Retrieve coords for samples on which RDA was run
coords <- read_tsv(here(path, "/RDA_genes_coords.txt"), col_names = TRUE)
ca <- get_ca()
```

## 2. Select env layers

Calculate adaptive index using either BIO1 + NDVI or enviro PCs.

```{r}
if (gea_method == "pca") envlayers <- import_env_files(type = "rasterPCs", future = FALSE)
if (gea_method == "bio1ndvi") envlayers <- import_env_files(type = "ind_layers", future = FALSE)

# Check env var naming
all(names(envlayers$env_pres) == row.names(dat_mod$biplot))
# Rename env names to keep consistent with RDA results
# names(envlayers$env_pres) <- paste("env_", names(envlayers$env_pres), sep = "")
```

## 3. Calculate adaptive index

```{r}
# Adaptive index for simple RDA model
ai <- adaptive_index(biplot = dat_mod$biplot, 
                     K = ncol(dat_mod$biplot), 
                     env_pres = envlayers$env_pres, 
                     coords = coords %>% dplyr::select(x, y))
```

## 3. Export rasters of adaptive index

```{r}
if (gea_method == "pca") ai_rs <- c(terra::rast(ai[[1]]), terra::rast(ai[[2]]), terra::rast(ai[[3]]))
if (gea_method == "bio1ndvi") ai_rs <- c(terra::rast(ai[[1]]), terra::rast(ai[[2]]))

writeRaster(ai_rs, paste0(path, "/", prefix, "_adapind_load.tif"), overwrite = TRUE)
```

## 4. Build maps of adaptive index

There are a number of options within the `plot_adaptive()` function, specified using the `plot_type` argument. The first is "basic" which shows the adaptive index (always scaled 0-1) for each RDA axis.

```{r}
p_ai <- plot_adaptive(Proj_data = ai_rs, 
                      bkg = ca, 
                      to_mask = FALSE, 
                      index_name = "Adaptive index", 
                      plot_type = "basic")

# Export plot
pdf(paste0(here("analysis", "adaptive", "plots"), "/", prefix, "_ai.pdf"), width = 12, height = 8)
p_ai
dev.off()
```

You can also have the function extract adaptive index values for your sampling coordinates by setting `plot_type = "extracted_vals"`.

```{r}
p_ext <- plot_adaptive(Proj_data = ai_rs, 
                       bkg = ca, 
                       to_mask = FALSE, 
                       index_name = "Adaptive index", 
                       plot_type = "extracted_vals",
                       coords = coords %>% dplyr::select(x, y))
# Export plot
pdf(paste0(here("analysis", "adaptive", "plots"), "/", prefix, "_ai_extracted.pdf"), width = 12, height = 8)
p_ext
dev.off()
```

If you want the plotting done the way Capblancq & Forester did, you can do `plot_type = "capblancq"`.

```{r}
plot_adaptive(Proj_data = ai_rs, 
              bkg = ca, 
              to_mask = FALSE, 
              index_name = "Adaptive index", 
              plot_type = "capblancq",
              coords = coords %>% dplyr::select(x, y))
```

## 5. Build rainbow adaptive index maps

To combine across RDA axes, we can build 'rainbow' maps which assign each axis to a given R, G, or B value. This requires sampling coordinates and scaled loadings from the RDA model.

```{r}
# If `dat_mod` object already loaded, do:
# loadings = dat_mod$scaledload
# If `dat_mod` not loaded, import by doing:
loadings = read_csv(paste0(path, "/RDA_scaledloadings_", prefix, "_mod.csv"))

p_rainbow <- plot_adaptive(Proj_data = ai_rs, 
                           bkg = ca, 
                           to_mask = FALSE, 
                           index_name = "Adaptive index", 
                           plot_type = "rainbow",
                           coords = coords %>% dplyr::select(x, y),
                           loadings = loadings,
                           bkg_col = "black")

pdf(paste0(here("analysis", "adaptive", "plots"), "/", prefix, "_ai_rainbowmap.pdf"), width = 12, height = 8)
p_rainbow
dev.off()
```

Finally, you can extract values from your rainbow map for your sampling coordinates by setting `plot_type = "extracted_rainbow"`.

```{r}
p_rainbow_ext <- plot_adaptive(Proj_data = ai_rs, 
                               bkg = ca, 
                               to_mask = FALSE,
                               index_name = "Adaptive index", 
                               plot_type = "extracted_rainbow",
                               coords = coords %>% dplyr::select(x, y),
                               loadings = loadings,
                               bkg_col = "white")

pdf(paste0(here("analysis", "adaptive", "plots"), "/", prefix, "_ai_rainbowmap_extracted.pdf"), width = 12, height = 8)
p_rainbow_ext
dev.off()
```

Mask based on species range:

```{r}
env_pres <- raster::mask(env_pres, range)
env_fut <- raster::mask(env_fut, range)
```