# Adaptive index

```{r setup, include=FALSE}
# Set global options for caching
knitr::opts_chunk$set(
  cache = TRUE,
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  result = FALSE)

library(algatr)
library(here)
library(tidyverse)
library(RColorBrewer)
library(sf)
library(terra)
library(raster)
library(vegan)
library(cowplot)

source(here("analysis", "adaptive", "adaptive_index.R"))
source(here("general_functions.R"))
```

## Specify model to examine

```{r}
gea_method = "bio1ndvi" # "pca" or "bio1ndvi"

if (gea_method == "bio1ndvi") output_path = here("analysis", "adaptive", "outputs", "RDA_bio1_ndvi") # for BIO1 + NDVI
if (gea_method == "pca") output_path = here("analysis", "adaptive", "outputs", "RDA_PCA") # for rasterPCA env layers
```

## 1. Get input files

```{r}
dat_mod <- import_rda(output_path = output_path, suffix = "mod", rds_obj = TRUE)
dat_mod_pc <- import_rda(output_path = output_path, suffix = "modPCs", rds_obj = TRUE)

# Retrieve coords for samples on which RDA was run
coords <- read_tsv(here(output_path, "RDA_genes_coords.txt"), col_names = TRUE)
ca <- get_ca()
```

## 2. Select env layers

We'll calculate adaptive index using both three env PCs and also just BIO1 + NDVI.

```{r}
if (gea_method == "pca") envlayers <- import_env_files(type = "rasterPCs", future = FALSE, model = NULL, years = NULL, RCP = NULL)
if (gea_method == "bio1ndvi") envlayers <- import_env_files(type = "ind_layers", future = FALSE, model = NULL, years = NULL, RCP = NULL)

# Check that above worked
names(envlayers$env_pres)
```

## 2. Calculate adaptive index

We're going to calculate the adaptive index for both RDA runs: one without population structure correction and the other with (using top 3 PCs). We'll calculate the adaptive index using the loadings method.

```{r}
# Change layer names to keep consisent with rownames in RDA model
if (names(envlayers$env_pres) != row.names(dat_mod$biplot)) names(envlayers$env_pres) <- paste("env_", names(envlayers$env_pres), sep = "")
if (names(envlayers$env_pres) != row.names(dat_mod_pc$biplot)) names(envlayers$env_pres) <- paste("env_", names(envlayers$env_pres), sep = "")

# Check that above worked
names(envlayers$env_pres) == row.names(dat_mod$biplot)

# Adaptive index for RDA models without PC correction
ai <- adaptive_index(biplot = dat_mod$biplot, K = ncol(dat_mod$biplot), env_pres = envlayers$env_pres, 
                     coords = coords %>% dplyr::select(x, y), range = NULL, 
                     method = "loadings")

# Adaptive index for RDA models with PC correction
ai_pc <- adaptive_index(dat_mod_pc$biplot, K = ncol(dat_mod_pc$biplot), env_pres, 
                        coords = coords %>% dplyr::select(x, y), range = NULL, 
                        method = "loadings")
```

## 3. Export rasters of adaptive index

```{r}
if (gea_method == "pca") {
  ai_rs <- c(terra::rast(ai[[1]]), terra::rast(ai[[2]]), terra::rast(ai[[3]]))
  ai_pc_rs <- c(terra::rast(ai_pc[[1]]), terra::rast(ai_pc[[2]]), terra::rast(ai[[3]]))
}
if (gea_method == "bio1ndvi") {
  ai_rs <- c(terra::rast(ai[[1]]), terra::rast(ai[[2]]))
  ai_pc_rs <- c(terra::rast(ai_pc[[1]]), terra::rast(ai_pc[[2]]))
}

writeRaster(ai_rs, here(output_path, "adapind_load.tif"), overwrite = TRUE)
writeRaster(ai_pc_rs, here(output_path, "adapind_load_PCs.tif"), overwrite = TRUE)
```

## 4. Build maps of adaptive index

```{r}
# ai_rs <- raster::stack(here(output_path, "adapind_load.tif"))
# ai_pc_rs <- raster::stack(here(output_path, "adapind_load_PCs.tif"))

# ai <- list(ai_rs[[1]], air)

# No PC correction
p_ai <- plot_adaptive(Proj_data = ai, K = length(ai), bkg = ca, to_mask = FALSE, index_name = "Adaptive index", title = "RDA adaptive index, loadings method")

# With PC correction
p_ai_pc <- plot_adaptive(Proj_data = ai_pc, K = length(ai_pc), bkg = ca, to_mask = FALSE, title = "RDA adaptive index, loadings method")

ai_plots <- plot_grid(p_ai, p_ai_pc, nrow = 2)

if (gea_method == "pca") pdf(here("analysis", "adaptive", "plots", "gea_genes_pca_ai.pdf"), width = 12, height = 8)
if (gea_method == "bio1ndvi") pdf(here("analysis", "adaptive", "plots", "gea_genes_bio1ndvi_ai.pdf"), width = 12, height = 8)
p_ai
dev.off()
```

## 5. Build biplots

```{r}
# # Import vals from GEA analysis
# rdadapt <- read_csv(TODO)
# # TODO automatically uses q.values column from rdadapt
# plt_biplot <- biplot_plot(biplot, loadings = scaled_loadings, 
#                           pvals = rdadapt, sig = 0.01, biplot_type = "separate")
# biplots <- plot_grid(plt_biplot$p_rda1, plt_biplot$p_rda2, plt_biplot$p_rda3, nrow = 1)

# pdf(here("analysis", "adaptive", "plots", "gea_genes_biplot.pdf"), width = 12, height = 4)
# biplots
# dev.off()
```
