# Adaptive index

```{r setup, include=FALSE}
# Set global options for caching
knitr::opts_chunk$set(
  cache = TRUE,
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  result = FALSE)

library(algatr)
library(here)
library(tidyverse)
library(RColorBrewer)
library(sf)
library(terra)
library(raster)
library(vegan)
library(cowplot)

source(here("analysis", "adaptive", "adaptive_index.R"))
source(here("general_functions.R"))
```

## Specify model to examine

```{r}
gea_method = "bio1ndvi" # "pca" or "bio1ndvi"

if (gea_method == "bio1ndvi") output_path = here("analysis", "adaptive", "outputs", "RDA_bio1_ndvi") # for BIO1 + NDVI
if (gea_method == "pca") output_path = here("analysis", "adaptive", "outputs", "RDA_PCA") # for rasterPCA env layers
```

## 1. Get input files

```{r}
dat_mod <- import_rda(output_path = output_path, suffix = "mod", rds_obj = TRUE)
dat_mod_pc <- import_rda(output_path = output_path, suffix = "modPCs", rds_obj = TRUE)

# Retrieve coords for samples on which RDA was run
coords <- read_tsv(here(output_path, "RDA_genes_coords.txt"), col_names = TRUE)
ca <- get_ca()
```

## 2. Select env layers

We'll calculate adaptive index using both three env PCs and also just BIO1 + NDVI.

```{r}
if (gea_method == "pca") envlayers <- import_env_files(type = "rasterPCs", future = FALSEL)
if (gea_method == "bio1ndvi") envlayers <- import_env_files(type = "ind_layers", future = FALSE)
# Rename env names to keep consistent with RDA results
names(envlayers$env_pres) <- paste("env_", names(envlayers$env_pres), sep = "")

# Check that above worked
all(names(envlayers$env_pres) == row.names(dat_mod$biplot))
# Check resolution
all(res(envlayers$env_pres) == res(envlayers$env_fut))
```

## 3. Calculate adaptive index

We're going to calculate the adaptive index for both RDA runs: one without population structure correction and the other with (using top 3 PCs). We'll calculate the adaptive index using the loadings method.

```{r}
# Adaptive index for RDA models without PC correction
ai <- adaptive_index(biplot = dat_mod$biplot, K = ncol(dat_mod$biplot), env_pres = envlayers$env_pres, 
                     coords = coords %>% dplyr::select(x, y), range = NULL, 
                     method = "loadings")

# Adaptive index for RDA models with PC correction
ai_pc <- adaptive_index(dat_mod_pc$biplot, K = ncol(dat_mod_pc$biplot), env_pres, 
                        coords = coords %>% dplyr::select(x, y), range = NULL, 
                        method = "loadings")
```

## 3. Export rasters of adaptive index

```{r}
if (gea_method == "pca") {
  ai_rs <- c(terra::rast(ai[[1]]), terra::rast(ai[[2]]), terra::rast(ai[[3]]))
  ai_pc_rs <- c(terra::rast(ai_pc[[1]]), terra::rast(ai_pc[[2]]), terra::rast(ai[[3]]))
}
if (gea_method == "bio1ndvi") {
  ai_rs <- c(terra::rast(ai[[1]]), terra::rast(ai[[2]]))
  ai_pc_rs <- c(terra::rast(ai_pc[[1]]), terra::rast(ai_pc[[2]]))
}

writeRaster(ai_rs, here(output_path, "adapind_load.tif"), overwrite = TRUE)
writeRaster(ai_pc_rs, here(output_path, "adapind_load_PCs.tif"), overwrite = TRUE)
```

## 4. Build maps of adaptive index

There are a number of options within the `plot_adaptive()` function, specified using the `plot_type` argument. The first is "basic" which shows the adaptive index (always scaled 0-1) for each RDA axis.

```{r}
ai_rs <- terra::rast(here(output_path, "adapind_load.tif"))
# ai_pc_rs <- terra::rast(here(output_path, "adapind_load_PCs.tif"))

p_ai <- plot_adaptive(Proj_data = ai_rs, bkg = ca, to_mask = FALSE, 
                      index_name = "Adaptive index", plot_type = "basic")

# Export plot
if (gea_method == "pca") pdf(here("analysis", "adaptive", "plots", "pca_ai.pdf"), width = 12, height = 8)
if (gea_method == "bio1ndvi") pdf(here("analysis", "adaptive", "plots", "bio1ndvi_ai.pdf"), width = 12, height = 8)
p_ai
dev.off()
```

You can also have the function extract adaptive index values for your sampling coordinates by setting `plot_type = "extracted_vals"`.

```{r}
plot_adaptive(Proj_data = ai_rs, bkg = ca, to_mask = FALSE, 
              index_name = "Adaptive index", plot_type = "extracted_vals",
              coords = coords %>% dplyr::select(x, y))
```

If you want the plotting done the way Capblancq & Forester did, you can do `plot_type = "capblancq"`.

```{r}
plot_adaptive(Proj_data = ai_rs, bkg = ca, to_mask = FALSE, 
              index_name = "Adaptive index", plot_type = "capblancq",
              coords = coords %>% dplyr::select(x, y))
```

## 5. Build rainbow adaptive index maps

To combine across RDA axes, we can build 'rainbow' maps which assign each axis to a given R, G, or B value. This requires sampling coordinates and scaled loadings from the RDA model.

```{r}
# Import loadings in either of the following ways:
# loadings = dat_mod$scaledload
loadings = read_csv(here(output_path, "58-Sceloporus_RDA_scaledloadings_mod.csv"))

p_rainbow <- plot_adaptive(Proj_data = ai_rs, bkg = ca, to_mask = FALSE, 
                           index_name = "Adaptive index", plot_type = "rainbow",
                           coords = coords %>% dplyr::select(x, y),
                           loadings = loadings)

if (gea_method == "bio1ndvi") pdf(here("analysis", "adaptive", "plots", "bio1ndvi_ai_rainbowmap.pdf"), width = 12, height = 8)
if (gea_method == "pca") pdf(here("analysis", "adaptive", "plots", "pca_ai_rainbowmap.pdf"), width = 12, height = 8)
p_rainbow
dev.off()
```

Finally, you can extract values from your rainbow map for your sampling coordinates by setting `plot_type = "extracted_rainbow"`.

```{r}
plot_adaptive(Proj_data = ai_rs, bkg = ca, to_mask = FALSE,
              index_name = "Adaptive index", plot_type = "extracted_rainbow",
              coords = coords %>% dplyr::select(x, y),
              loadings = loadings)
```
