# Visualizing environmental layers

```{r setup, include=FALSE}
# Set global options for caching
knitr::opts_chunk$set(
  cache = TRUE,
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  result = FALSE)

library(here)
library(tidyverse)
library(tidyterra)
library(cowplot)
library(sf)
library(terra)
library(raster)
library(cowplot)

source(here("analysis", "adaptive", "adaptive_index.R"))
source(here("general_functions.R"))
```

## 1. Get data files

```{r}
# Retrieve coords for samples on which RDA was run
coords <- get_coords()
ca <- get_ca()
```

## 2. Get present and future env layers for rasterPCA and BIO1+NDVI

```{r}
pca_envlayers <- import_env_files(type = "rasterPCs", future = TRUE)
bio1ndvi_envlayers <- import_env_files(type = "ind_layers", future = TRUE)
```

## 3. Compare BIO1 layers

```{r}
# Mask future to present
mask_26 <- terra::mask(terra::rast(bio1ndvi_envlayers$env_fut$CHELSA_bio1_2071.2100_gfdl.esm4_ssp126_V.2.1),
            terra::rast(bio1ndvi_envlayers$env_pres$BIO1))
mask_85 <- terra::mask(terra::rast(bio1ndvi_envlayers$env_fut$CHELSA_bio1_2071.2100_gfdl.esm4_ssp585_V.2.1),
            terra::rast(bio1ndvi_envlayers$env_pres$BIO1))

# minbio1 = minValue(bio1ndvi_envlayers$env_pres$BIO1)
# maxbio1 = maxValue(raster::stack(mask_85))

p_bio1_pres <- ggplot() +
  geom_sf(data = ca, fill = "lightgrey") +
  geom_spatraster(data = terra::rast(bio1ndvi_envlayers$env_pres$BIO1)) +
  scale_fill_viridis_c(na.value = "transparent", limits = c(-9, 31)) +
  theme_map() +
  ggtitle("BIO1, present")

p_bio1_fut26 <- ggplot() +
  geom_sf(data = ca, fill = "lightgrey") +
  geom_spatraster(data = mask_26) +
  scale_fill_viridis_c(na.value = "transparent", limits = c(-9, 31)) +
  theme_map() +
  ggtitle("BIO1, 2071-2100 (RCP2.6)")

p_bio1_fut85 <- ggplot() +
  geom_sf(data = ca, fill = "lightgrey") +
  geom_spatraster(data = mask_85) +
  scale_fill_viridis_c(na.value = "transparent", limits = c(-9, 31)) +
  theme_map() +
  ggtitle("BIO1, 2071-2100 (RCP8.5)")

plot_grid(p_bio1_pres, p_bio1_fut26, p_bio1_fut85, nrow = 1)
```

## 3. Compare rasterPCs

```{r}
names(pca_envlayers$env_fut) <- c("env_PC1_RCP2.6", "env_PC2_RCP2.6", "env_PC3_RCP2.6",
                                  "env_PC1_RCP8.5", "env_PC2_RCP8.5", "env_PC3_RCP8.5")

p_pcs_pres <-
  ggplot() +
  geom_sf(data = ca, fill = "lightgrey") +
  geom_spatraster(data = terra::rast(pca_envlayers$env_pres)) +
  scale_fill_viridis_c(na.value = "transparent") +
  theme_map() +
  ggtitle("rasterPCs, present") +
  facet_wrap(~lyr)

p_pcs_fut26 <-
  ggplot() +
  geom_sf(data = ca, fill = "lightgrey") +
  geom_spatraster(data = terra::rast(pca_envlayers$env_fut[[1:3]])) +
  scale_fill_viridis_c(na.value = "transparent", limits = c(-12450, 32000)) +
  theme_map() +
  ggtitle("rasterPCs, 2071-2100 (RCP2.6)") +
  facet_wrap(~lyr)

p_pcs_fut85 <- 
  ggplot() +
  geom_sf(data = ca, fill = "lightgrey") +
  geom_spatraster(data = terra::rast(pca_envlayers$env_fut[[4:6]])) +
  scale_fill_viridis_c(na.value = "transparent", limits = c(-14750, 36000)) +
  theme_map() +
  ggtitle("rasterPCs, 2071-2100 (RCP8.5)") +
  facet_wrap(~lyr)

plot_grid(p_pcs_pres, p_pcs_fut26, p_pcs_fut85, nrow = 3)
```

## Compare RCPs and models

## 1. Get present env layers

```{r}
env_pcs <- terra::rast(here("data", "env", "california_chelsa_bioclim_1981-2010_V.2.1_pca.tif"))
# future_pcs <- terra::rast(paste0(here("data", "env", "future"), "/CHELSA_2041-2070_", model, "_", ssp, "_V.2.1_pca.tif"))
# names(future_pcs) <- c("PC1", "PC2", "PC3")
p_pres <- ggplot() +
  geom_spatraster(data = env_pcs) +
  theme_map() +
  facet_wrap(~lyr) +
  scale_fill_viridis_c(option = "D", na.value = "transparent") +
  ggtitle("Present PCs")
```

## 2. Define function

The following function will pull in future env PCs and build comparison plots depending on the factor you're interested in comparing: a range of years, two different models, or two RCPs.

```{r}
#' Builds faceted plots by user-defined metric to compare present & future rasterPCA results
#'
#' @param model model name (e.g., "gfdl-esm4" or c("gfdl-esm4", "ipsl-cm6a-lr"))
#' @param ssp ssp value (retrieved earlier using RCP value); MUST MATCH RCP
#' @param RCP rep conc pathway value; MUST MATCH SSP
#' @param years range of years to examine (e.g., "2041-2070" or c("2041-2070", "2071-2100"))
#' @param to_compare what to facet each plot on; options are "years", "RCP", or "model"
#' 
plot_future <- function(model, ssp, RCP, years, to_compare = "years") {
  if (to_compare == "years") {
    future_pcs1 <- terra::rast(paste0(here("data", "env", "future"), "/CHELSA_", years[1], "_", model, "_", ssp, "_V.2.1_pca.tif"))
    future_pcs2 <- terra::rast(paste0(here("data", "env", "future"), "/CHELSA_", years[2], "_", model, "_", ssp, "_V.2.1_pca.tif"))
    title1 <- paste0("Future PCs, ", model, " RCP ", RCP, ", ", years[1])
    title2 <- paste0("Future PCs, ", model, " RCP ", RCP, ", ", years[2])
  }
  if (to_compare == "model") {
    future_pcs1 <- terra::rast(paste0(here("data", "env", "future"), "/CHELSA_", years, "_", model[1], "_", ssp, "_V.2.1_pca.tif"))
    future_pcs2 <- terra::rast(paste0(here("data", "env", "future"), "/CHELSA_", years, "_", model[2], "_", ssp, "_V.2.1_pca.tif"))
    title1 <- paste0("Future PCs, ", model[1], " RCP ", RCP, ", ", years)
    title2 <- paste0("Future PCs, ", model[2], " RCP ", RCP, ", ", years)
  }
  if (to_compare == "RCP") {
    future_pcs1 <- terra::rast(paste0(here("data", "env", "future"), "/CHELSA_", years, "_", model, "_", ssp[1], "_V.2.1_pca.tif"))
    future_pcs2 <- terra::rast(paste0(here("data", "env", "future"), "/CHELSA_", years, "_", model, "_", ssp[2], "_V.2.1_pca.tif"))
    title1 <- paste0("Future PCs, ", model, " RCP ", RCP[1], ", ", years)
    title2 <- paste0("Future PCs, ", model, " RCP ", RCP[2], ", ", years)
  }
  names(future_pcs1) <- c("PC1", "PC2", "PC3")
  names(future_pcs2) <- c("PC1", "PC2", "PC3")

  return(list("future_pcs1" = future_pcs1, "title1" = title1, "future_pcs2" = future_pcs2, "title2" = title2))
}
```

## 3. Build plots and export

Define arguments.

```{r}
RCP = c(2.6, 8.5) # options are 2.6, 7, or 8.5 from CHELSA V2
model = "gfdl-esm4" # options are "gfdl-esm4" (highest priority) or "ipsl-cm6a-lr"
years = "2071-2100" # options are "2041-2070" or "2071-2100"
ssp <- c("ssp126", "ssp585")

# For single RCPs:
# if (RCP == 8.5) ssp <- "ssp585"
# if (RCP == 7) ssp <- "ssp370"
# if (RCP == 2.6) ssp <- "ssp126"
```

```{r}
to_compare = "RCP"
plot_dat <- plot_future(model = model, RCP = RCP, years = years, to_compare = to_compare)

p_fut_1 <- ggplot() +
  geom_spatraster(data = plot_dat$future_pcs1) +
  theme_map() +
  facet_wrap(~lyr) +
  scale_fill_viridis_c(option = "D", na.value = "transparent") +
  ggtitle(paste0(plot_dat$title1))

p_fut_2 <- ggplot() +
  geom_spatraster(data = plot_dat$future_pcs2) +
  theme_map() +
  facet_wrap(~lyr) +
  scale_fill_viridis_c(option = "D", na.value = "transparent") +
  ggtitle(paste0(plot_dat$title2))

p_pcs <- plot_grid(p_pres, p_fut_1, p_fut_2, nrow = 3)

# pdf(paste0(here("analysis", "adaptive", "plots"), "/env_PCs_", model, "_RCP", RCP, ".pdf"), width = 12, height = 8)
pdf(paste0(here("analysis", "adaptive", "plots"), "/env_PCs_", model, "_", to_compare, "comp.pdf"), width = 12, height = 8)
p_pcs
dev.off()
```

## 4. Comparing RCPs for BIO1

```{r}
all_pres <- terra::rast(here("data", "env", "california_chelsa_bioclim_1981-2010_V.2.1.tif"))
bio1_pres <- terra::subset(all_pres, 1) # "CHELSA_bio1_1981.2010_V.2.1"
bio1_fut_26 <- terra::rast(paste0(here("data", "env", "future", "envicloud/chelsa/chelsa_V2/GLOBAL/climatologies"), "/2071-2100/GFDL-ESM4/ssp126/bio/CHELSA_bio1_2071-2100_gfdl-esm4_ssp126_V.2.1.tif"))
bio1_fut_85 <- terra::rast(paste0(here("data", "env", "future", "envicloud/chelsa/chelsa_V2/GLOBAL/climatologies"), "/2071-2100/GFDL-ESM4/ssp585/bio/CHELSA_bio1_2071-2100_gfdl-esm4_ssp585_V.2.1.tif"))

resamp_26 <- terra::resample(bio1_fut_26, bio1_pres)
resamp_85 <- terra::resample(bio1_fut_85, bio1_pres)
crop_26 <- terra::crop(resamp_26, bio1_pres, mask = TRUE)
crop_85 <- terra::crop(resamp_85, bio1_pres, mask = TRUE)

change_26 <- resamp_26 - bio1_pres
change_85 <- resamp_85 - bio1_pres

# Plot results
p_bio1 <- ggplot() +
  geom_spatraster(data = bio1_pres) +
  theme_map() +
  scale_fill_viridis_c(option = "D", na.value = "transparent", limits = c(minmax(bio1_pres)[1], minmax(crop_85)[2])) +
  ggtitle("BIO1, present")

p_26 <- ggplot() +
  geom_spatraster(data = crop_26) +
  theme_map() +
  scale_fill_viridis_c(option = "D", na.value = "transparent", limits = c(minmax(bio1_pres)[1], minmax(crop_85)[2])) +
  ggtitle("BIO1, RCP2.6")

p_85 <- ggplot() +
  geom_spatraster(data = crop_85) +
  theme_map() +
  scale_fill_viridis_c(option = "D", na.value = "transparent", limits = c(minmax(bio1_pres)[1], minmax(crop_85)[2])) +
  ggtitle("BIO1, RCP8.5")

p_26_change <- ggplot() +
  geom_spatraster(data = change_26) +
  theme_map() +
  # facet_wrap(~lyr) +
  scale_fill_viridis_c(option = "D", na.value = "transparent", limits = c(minmax(change_26)[1], minmax(change_85)[2])) +
  ggtitle("Change in BIO1 between present and RCP2.6")

p_85_change <- ggplot() +
  geom_spatraster(data = change_85) +
  theme_map() +
  # facet_wrap(~lyr) +
  scale_fill_viridis_c(option = "D", na.value = "transparent", limits = c(minmax(change_26)[1], minmax(change_85)[2])) +
  ggtitle("Change in BIO1 between present and RCP8.5")

p1 <- plot_grid(p_bio1, p_26, nrow = 1)
p2 <- plot_grid(p_bio1, p_85, nrow = 1)

# p_bio1change <- plot_grid(p_26_change, p_85_change, nrow = 1)
# pdf(paste0(here("analysis", "adaptive", "plots"), "/BIO1_change_maps.pdf"), width = 12, height = 8)
# p_bio1change
# dev.off()

p_bigchange <- plot_grid(p1, p2, p_26_change, p_85_change, nrow = 2, rel_heights = c(1, 2))

pdf(paste0(here("analysis", "adaptive", "plots"), "/BIO1_big_change_maps.pdf"), width = 12, height = 8)
p_bigchange
dev.off()

# Stack with present NDVI layer; resample and crop
cropped <- terra::crop(terra::rast(bio1_fut), terra::ext(terra::rast(ndvi)))
resamp <- terra::resample(cropped, terra::rast(ndvi))
# TODO something weird happening below
# env_fut <- raster::stack(raster::raster(resamp), ndvi)
```
