---
title: "genomic_offset"
output:
  github_document:
    toc: TRUE
---

```{r setup, include=FALSE}
# Set global options for caching
knitr::opts_chunk$set(
  cache = TRUE,
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  result = FALSE)

library(algatr)
library(here)
library(tidyverse)
library(tidyterra)
library(RColorBrewer)
library(sf)
library(terra)
library(raster)
library(cowplot)
library(RStoolbox)

source(here("analysis", "adaptive", "adaptive_index.R"))
source(here("analysis", "adaptive", "genomic_offset.R"))
source(here("general_functions.R"))
```

## Specify model to examine

```{r}
gea_method = "bio1ndvi" # "pca" or "bio1ndvi"
path = here("analysis", "adaptive", "outputs")
prefix = "58-Sceloporus_bio1ndvi_gea" # "58-Sceloporus_bio1ndvi_gea_genes_nonsyn"
```

## 1. Get input files

```{r}
# If `dat_mod` is not in memory, run:
dat_mod <- import_rda(output_path = path, suffix = paste0(prefix, "_mod"), rds_obj = TRUE)

# Retrieve coords for samples on which RDA was run
coords <- read_tsv(here(path, "RDA_genes_coords.txt"), col_names = TRUE)
ca <- get_ca()
```

## 2. Get environmental layers: present and future

```{r}
if (gea_method == "pca") {
  envlayers <- import_env_files(type = "rasterPCs", future = TRUE)
  # Rename present env names to keep consistent with RDA results
  names(envlayers$env_pres) <- paste("env_", names(envlayers$env_pres), sep = "")
}
if (gea_method == "bio1ndvi") envlayers <- import_env_files(type = "ind_layers", future = TRUE)

# Check that above worked
all(names(envlayers$env_pres) == row.names(dat_mod$biplot))
names(envlayers$env_fut)
# Check resolution
all(res(envlayers$env_pres) == res(envlayers$env_fut))
```

## 3. Calculate genomic offset

The `genomic_offset()` function outputs a list of five 
objects: (1) Proj_pres, which is the present projection (equivalent to the adaptive index); (2) Proj_fut, which projects the adaptive index values on to future
environmental layers, (3) Proj_offset, which is the genomic offset (i.e., difference between Proj_pres and Proj_fut), (4) Proj_offset_global, which is the global
offset if multiple future environmental time points were provided, and (5) weights, which are the eigenvalues from the RDA model and are how the RDA axes values 
were weighted.

```{r}
offset <- genomic_offset(loadings = dat_mod$scaledload,
                         biplot = dat_mod$biplot, 
                         eig = dat_mod$eig, 
                         K = ncol(dat_mod$biplot), 
                         env_pres = envlayers$env_pres,
                         env_fut = envlayers$env_fut,
                         scale_env,
                         center_env,
                         mod = NULL)
```

## 3. Save genomic offset maps

```{r}
if (gea_method == "bio1ndvi") {
  offset_26 <- raster::stack(offset$Proj_offset_RCP26$RDA1, 
                             offset$Proj_offset_RCP26$RDA2)
  offset_85 <- raster::stack(offset$Proj_offset_RCP85$RDA1, 
                             offset$Proj_offset_RCP85$RDA2)
}
if (gea_method == "pca") {
  offset_26 <- raster::stack(offset$Proj_offset_RCP26$RDA1, 
                             offset$Proj_offset_RCP26$RDA2, 
                             offset$Proj_offset_RCP26$RDA3)
  offset_85 <- raster::stack(offset$Proj_offset_RCP85$RDA1, 
                             offset$Proj_offset_RCP85$RDA2, 
                             offset$Proj_offset_RCP85$RDA3)
}

offset_26_g <- raster::stack(offset$Proj_offset_global_RCP26)
offset_85_g <- raster::stack(offset$Proj_offset_global_RCP85)

writeRaster(offset_26, paste0(path, "/", prefix, "_offset_load_RCP26.tif"), overwrite = TRUE)
writeRaster(offset_26_g, paste0(path, "/", prefix, "_offset_load_global_RCP26.tif"), overwrite = TRUE)
writeRaster(offset_85, paste0(path, "/", prefix, "_offset_load_RCP85.tif"), overwrite = TRUE)
writeRaster(offset_85_g, paste0(path, "/", prefix, "_offset_load_global_RCP85.tif"), overwrite = TRUE)
```

## 4. Plot genomic offset

```{r}
# If you're starting here, run the following lines:
# offset_26 <- terra::rast(paste0(path, "/", prefix, "_offset_load_RCP26.tif"))
# offset_85 <- terra::rast(paste0(path, "/", prefix, "_offset_load_RCP85.tif"))

p_26 <- plot_offset(offset_26, bkg = ca, index_name = "Genomic offset (RCP 2.6)",
                    plot_type = "basic", viridis_option = "B", free_scales = TRUE)

p_85 <- plot_offset(offset_85, bkg = ca, index_name = "Genomic offset (RCP 8.5)",
                    plot_type = "basic", viridis_option = "B", free_scales = TRUE)

p <- plot_grid(p_26, p_85, nrow = 2)
pdf(paste0(here("analysis", "adaptive", "plots"), "/", prefix, "_offset.pdf"), width = 12, height = 8)
p
dev.off()
```

Extract offset values using the `plot_type = "extracted_vals"` argument. Coordinates must be provided.

```{r}
p_ext_26 <- plot_offset(offset_26, bkg = ca, index_name = "Genomic offset (RCP 2.6)",
            plot_type = "extracted_vals", viridis_option = "B", 
            coords = coords %>% dplyr::select(x, y))
p_ext_85 <- plot_offset(offset_85, bkg = ca, index_name = "Genomic offset (RCP 8.5)",
            plot_type = "extracted_vals", viridis_option = "B", 
            coords = coords %>% dplyr::select(x, y))

p_ext <- plot_grid(p_ext_26, p_ext_85, nrow = 2)

pdf(paste0(here("analysis", "adaptive", "plots"), "/", prefix, "_offset_extracted.pdf"), width = 12, height = 8)
p_ext
dev.off()
```

## 5. Make rainbow maps

```{r}
# For rainbow plot, you need scaled loadings from RDA model
# loadings <- dat_mod$scaledload
loadings <- read_csv(paste0(path, "/RDA_scaledloadings_", prefix, "_mod.csv"))

p_rainbow <- plot_offset(offset_85, bkg = ca, index_name = "Genomic offset (RCP 8.5)",
            plot_type = "rainbow", viridis_option = "B", coords = coords %>% dplyr::select(x, y))
pdf(paste0(here("analysis", "adaptive", "plots"), "/", prefix, "_offset_rainbow_RCP85.pdf"), width = 12, height = 8)
p_rainbow
dev.off()

p_rainbow_26 <- plot_offset(offset_26, bkg = ca, index_name = "Genomic offset (RCP 2.6)",
            plot_type = "rainbow", viridis_option = "B", coords = coords %>% dplyr::select(x, y))
pdf(paste0(here("analysis", "adaptive", "plots"), "/", prefix, "_offset_rainbow_RCP26.pdf"), width = 12, height = 8)
p_rainbow_26 
dev.off()
```

Extract rainbow map values and only plot points using `plot_type = "extracted_rainbow"`.

```{r}
p_rainbow_ext <- plot_offset(offset_85, bkg = ca, index_name = "Genomic offset (RCP 8.5)",
            plot_type = "extracted_rainbow", viridis_option = "B", coords = coords %>% dplyr::select(x, y))
pdf(paste0(here("analysis", "adaptive", "plots"), "/", prefix, "_offset_rainbow_extracted.pdf"), width = 12, height = 8)
p_rainbow_ext
dev.off()
```

## 6. Extracted offset vals on top of env raster layer

```{r}
fut_bio1_85 <- terra::rast(envlayers$env_fut$CHELSA_bio1_2071.2100_gfdl.esm4_ssp585_V.2.1)
fut_bio1_85 <- terra::crop(fut_bio1_85, ca, mask = TRUE)

ext <- terra::extract(offset_85, coords %>% dplyr::select(x, y), ID = FALSE, xy = TRUE)
tidy_ext <- ext %>% 
  pivot_longer(cols = 1:2, names_to = "axis", values_to = "value")
axis_names <- unique(tidy_ext$axis)

p1 <- ggplot() +
  geom_sf(data = ca, color = "lightgrey") +
  geom_spatraster(data = fut_bio1_85) +
  scale_fill_viridis_c(option = "D", na.value = "transparent", name = "BIO1 (°C), \nRCP 8.5") +
  geom_point(data = tidy_ext %>% filter(axis == axis_names[[1]]),
             aes(x = x, y = y, color = value), size = 3) +
  scale_color_viridis_c(option = "B", name = "Genomic offset (RDA1), \nRCP 8.5") +
  theme_map()
p2 <- ggplot() +
  geom_sf(data = ca, color = "lightgrey") +
  geom_spatraster(data = fut_bio1_85) +
  scale_fill_viridis_c(option = "D", na.value = "transparent", name = "BIO1 (°C), \nRCP 8.5") +
  geom_point(data = tidy_ext %>% filter(axis == axis_names[[2]]),
             aes(x = x, y = y, color = value), size = 3) +
  scale_color_viridis_c(option = "B", name = "Genomic offset (RDA2), \nRCP 8.5") +
  theme_map()

p_ext_bio1 <- plot_grid(p1, p2, nrow = 1)

pdf(paste0(here("analysis", "adaptive", "plots"), "/", prefix, "_offset_extracted_bio1.pdf"), width = 12, height = 8)
p_ext_bio1
dev.off()
```

## 7. Correlation between offset and bio1ndvi

For values extracted based on sampling coordinates:

```{r}
# If you're starting here, run the following lines:
offset_26 <- terra::rast(paste0(path, "/", prefix, "_offset_load_RCP26.tif"))
offset_85 <- terra::rast(paste0(path, "/", prefix, "_offset_load_RCP85.tif"))

bio1_pres <- terra::rast(envlayers$env_pres$BIO1)
bio1_fut_26 <- terra::rast(envlayers$env_fut$"CHELSA_bio1_2071.2100_gfdl.esm4_ssp126_V.2.1")
bio1_fut_85 <- terra::rast(envlayers$env_fut$"CHELSA_bio1_2071.2100_gfdl.esm4_ssp585_V.2.1")
resamp_26 <- terra::resample(bio1_fut_26, bio1_pres)
resamp_85 <- terra::resample(bio1_fut_85, bio1_pres)
crop_26 <- terra::crop(resamp_26, bio1_pres, mask = TRUE)
crop_85 <- terra::crop(resamp_85, bio1_pres, mask = TRUE)
# Calculate change present -> future for BIO1
change_26 <- resamp_26 - bio1_pres
change_85 <- resamp_85 - bio1_pres

# Extract values for sampling coordinates
change_26_vals <- terra::extract(change_26, coords %>% dplyr::select(x, y)) %>% 
  rename(BIO1change_RCP26 = `CHELSA_bio1_2071.2100_gfdl.esm4_ssp126_V.2.1`) %>% 
  dplyr::select(-ID)
change_85_vals <- terra::extract(change_85, coords %>% dplyr::select(x, y)) %>% 
  rename(BIO1change_RCP85 = `CHELSA_bio1_2071.2100_gfdl.esm4_ssp585_V.2.1`) %>% 
  dplyr::select(-ID)

sum_offset_26_vals <- terra::extract(offset_26, coords %>% dplyr::select(x, y)) %>% 
  rename(offset_RCP26_RDA1 = `58-Sceloporus_bio1ndvi_gea_offset_load_RCP26_1`, offset_RCP26_RDA2 = `58-Sceloporus_bio1ndvi_gea_offset_load_RCP26_2`) %>% 
  mutate(sum_offset_RCP26 = offset_RCP26_RDA1+offset_RCP26_RDA2) %>% 
  dplyr::select(sum_offset_RCP26, -ID)
sum_offset_85_vals <- terra::extract(offset_85, coords %>% dplyr::select(x, y)) %>% 
  rename(offset_RCP85_RDA1 = `58-Sceloporus_bio1ndvi_gea_offset_load_RCP85_1`, offset_RCP85_RDA2 = `58-Sceloporus_bio1ndvi_gea_offset_load_RCP85_2`) %>% 
  mutate(sum_offset_RCP85 = offset_RCP85_RDA1+offset_RCP85_RDA2) %>% 
  dplyr::select(sum_offset_RCP85, -ID)

rcp26_sum_vals <- bind_cols(change_26_vals, sum_offset_26_vals)
cors_sum_26 <- cor(rcp26_sum_vals, use = "na.or.complete", method = "pearson")
rcp85_sum_vals <- bind_cols(change_85_vals, sum_offset_85_vals)
cors_sum_85 <- cor(rcp85_sum_vals, use = "na.or.complete", method = "pearson")

cors_sum_26 # 0.9885295
cors_sum_85 # 0.9874386
```

Calculating the correlation across the entire raster:

```{r}
sum_offset_26 <- offset_26[[1]] + offset_26[[2]]
rcp26_layers <- raster::stack(c(change_26, sum_offset_26))
cors_lyr_26 <- raster::layerStats(rcp26_layers, stat = "pearson", na.rm = TRUE)
cors_lyr_26 <- cors_lyr_26$`pearson correlation coefficient`

sum_offset_85 <- offset_85[[1]] + offset_85[[2]]
rcp85_layers <- raster::stack(c(change_85, sum_offset_85))
cors_lyr_85 <- raster::layerStats(rcp85_layers, stat = "pearson", na.rm = TRUE)
cors_lyr_85 <- cors_lyr_85$`pearson correlation coefficient`

cors_lyr_26 # 0.9955793
cors_lyr_85 # 0.9954994
```