---
title: "genomic_offset"
output:
  github_document:
    toc: TRUE
---

```{r setup, include=FALSE}
# Set global options for caching
knitr::opts_chunk$set(
  cache = TRUE,
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  result = FALSE)

library(algatr)
library(here)
library(tidyverse)
library(tidyterra)
library(RColorBrewer)
library(sf)
library(terra)
library(raster)
library(cowplot)
library(RStoolbox)

source(here("analysis", "adaptive", "adaptive_index.R"))
source(here("analysis", "adaptive", "genomic_offset.R"))
source(here("general_functions.R"))
```

## Specify model to examine

```{r}
gea_method = "bio1ndvi" # "pca" or "bio1ndvi"
path = here("analysis", "adaptive", "outputs")
prefix = "58-Sceloporus_bio1ndvi_gea" # "58-Sceloporus_bio1ndvi_gea_genes_nonsyn"
```

## 1. Get input files

```{r}
dat_mod <- import_rda(output_path = path, suffix = paste0(prefix, "_mod"), rds_obj = TRUE)
dat_mod_pc <- import_rda(output_path = path, suffix = paste0(prefix, "_modPCs"), rds_obj = TRUE)

# Retrieve coords for samples on which RDA was run
coords <- read_tsv(here(path, "RDA_genes_coords.txt"), col_names = TRUE)
ca <- get_ca()
```

## 2. Get environmental layers: present and future

```{r}
if (gea_method == "pca") {
  envlayers <- import_env_files(type = "rasterPCs", future = TRUE)
  # Rename present env names to keep consistent with RDA results
  names(envlayers$env_pres) <- paste("env_", names(envlayers$env_pres), sep = "")
}
if (gea_method == "bio1ndvi") envlayers <- import_env_files(type = "ind_layers", future = TRUE)

# Check that above worked
all(names(envlayers$env_pres) == row.names(dat_mod$biplot))
names(envlayers$env_fut)
# Check resolution
all(res(envlayers$env_pres) == res(envlayers$env_fut))

# Get scaled and centered info
env <- terra::extract(envlayers$env_pres, coords %>% dplyr::select(x, y))
env <- scale(env, center = TRUE, scale = TRUE)
# Recover scaling coefficients
scale_env <- attr(env, 'scaled:scale')
center_env <- attr(env, 'scaled:center')
```

## 3. Calculate genomic offset

The `genomic_offset()` function outputs a list of five 
objects: (1) Proj_pres, which is the present projection (equivalent to the adaptive index); (2) Proj_fut, which projects the adaptive index values on to future
environmental layers, (3) Proj_offset, which is the genonic offset (i.e., difference between Proj_pres and Proj_fut), (4) Proj_offset_global, which is the global
offset if multiple future environmental time points were provided, and (5) weights, which are the eigenvalues from the RDA model and are how the RDA axes values 
were weighted.

```{r}
offset <- genomic_offset(loadings = dat_mod$scaledload, biplot = dat_mod$biplot, 
                         eig = dat_mod$eig, K = ncol(dat_mod$biplot), 
                         env_pres = envlayers$env_pres, env_fut = envlayers$env_fut, 
                         range = NULL, method = "loadings", scale_env, center_env, 
                         mod = NULL)

# offset_pc <- genomic_offset(loadings = dat_mod_pc$scaledload, biplot = dat_mod_pc$biplot, eig = dat_mod_pc$eig, K = ncol(dat_mod_pc$biplot), env_pres, env_fut, range = NULL, 
#                          method = "loadings", scale_env, center_env, mod = NULL)
```

## 3. Save genomic offset maps

```{r}
if (gea_method == "bio1ndvi") {
  offset_26 <- raster::stack(offset$Proj_offset_RCP26$RDA1, 
                             offset$Proj_offset_RCP26$RDA2)
  offset_85 <- raster::stack(offset$Proj_offset_RCP85$RDA1, 
                             offset$Proj_offset_RCP85$RDA2)
}
if (gea_method == "pca") {
  offset_26 <- raster::stack(offset$Proj_offset_RCP26$RDA1, 
                             offset$Proj_offset_RCP26$RDA2, 
                             offset$Proj_offset_RCP26$RDA3)
  offset_85 <- raster::stack(offset$Proj_offset_RCP85$RDA1, 
                             offset$Proj_offset_RCP85$RDA2, 
                             offset$Proj_offset_RCP85$RDA3)
}

offset_26_g <- raster::stack(offset$Proj_offset_global_RCP26)
offset_85_g <- raster::stack(offset$Proj_offset_global_RCP85)

writeRaster(offset_26, paste0(path, "/", prefix, "_offset_load_RCP26.tif"), overwrite = TRUE)
writeRaster(offset_26_g, paste0(path, "/", prefix, "_offset_load_global_RCP26.tif"), overwrite = TRUE)
writeRaster(offset_85, paste0(path, "/", prefix, "_offset_load_RCP85.tif"), overwrite = TRUE)
writeRaster(offset_85_g, paste0(path, "/", prefix, "_offset_load_global_RCP85.tif"), overwrite = TRUE)
```

## 4. Plot genomic offset

```{r}
# If you're starting here, run the following lines:
# offset_26 <- terra::rast(paste0(path, "/", prefix, "_offset_load_RCP26.tif"))
# offset_85 <- terra::rast(paste0(path, "/", prefix, "_offset_load_RCP85.tif"))

p_26 <- plot_offset(offset_26, bkg = ca, index_name = "Genomic offset (RCP 2.6)",
                    plot_type = "basic", viridis_option = "B", free_scales = TRUE)

p_85 <- plot_offset(offset_85, bkg = ca, index_name = "Genomic offset (RCP 8.5)",
                    plot_type = "basic", viridis_option = "B", free_scales = TRUE)

p <- plot_grid(p_26, p_85, nrow = 2)
pdf(paste0(here("analysis", "adaptive", "plots"), "/", prefix, "_offset.pdf"), width = 12, height = 8)
p
dev.off()
```

Extract offset values using the `plot_type = "extracted_vals"` argument. Coordinates must be provided.

```{r}
p_ext <- plot_offset(offset_85, bkg = ca, index_name = "Genomic offset (RCP 8.5)",
            plot_type = "extracted_vals", viridis_option = "B", coords = coords %>% dplyr::select(x, y))
pdf(paste0(here("analysis", "adaptive", "plots"), "/", prefix, "_offset_extracted.pdf"), width = 12, height = 8)
p_ext
dev.off()
```

## 5. Make rainbow maps

```{r}
# For rainbow plot, you need scaled loadings from RDA model
loadings <- dat_mod$scaledload
# loadings <- read_csv(here(output_path, "58-Sceloporus_RDA_scaledloadings_mod.csv"))

p_rainbow <- plot_offset(offset_85, bkg = ca, index_name = "Genomic offset (RCP 2.6)",
            plot_type = "rainbow", viridis_option = "B")
pdf(paste0(here("analysis", "adaptive", "plots"), "/", prefix, "_offset_rainbow.pdf"), width = 12, height = 8)
p_rainbow
dev.off()
```

Extract rainbow map values and only plot points using `plot_type = "extracted_rainbow"`.

```{r}

```

## 6. Extracted offset vals on top of env raster layer

```{r}
fut_bio1_85 <- terra::rast(envlayers$env_fut$CHELSA_bio1_2071.2100_gfdl.esm4_ssp585_V.2.1)
fut_bio1_85 <- terra::crop(fut_bio1_85, ca, mask = TRUE)

ext <- terra::extract(offset_85, coords %>% dplyr::select(x, y), ID = FALSE, xy = TRUE)
tidy_ext <- ext %>% 
  pivot_longer(cols = 1:2, names_to = "axis", values_to = "value")
axis_names <- unique(tidy_ext$axis)

p1 <- ggplot() +
  geom_sf(data = ca, color = "lightgrey") +
  geom_spatraster(data = fut_bio1_85) +
  scale_fill_viridis_c(option = "D", na.value = "transparent", name = "BIO1 (°C), \nRCP 8.5") +
  geom_point(data = tidy_ext %>% filter(axis == axis_names[[1]]),
             aes(x = x, y = y, color = value), size = 3) +
  scale_color_viridis_c(option = "B", name = "Genomic offset (RDA1), \nRCP 8.5") +
  theme_map()
p2 <- ggplot() +
  geom_sf(data = ca, color = "lightgrey") +
  geom_spatraster(data = fut_bio1_85) +
  scale_fill_viridis_c(option = "D", na.value = "transparent", name = "BIO1 (°C), \nRCP 8.5") +
  geom_point(data = tidy_ext %>% filter(axis == axis_names[[2]]),
             aes(x = x, y = y, color = value), size = 3) +
  scale_color_viridis_c(option = "B", name = "Genomic offset (RDA2), \nRCP 8.5") +
  theme_map()

p_ext_bio1 <- plot_grid(p1, p2, nrow = 1)

pdf(paste0(here("analysis", "adaptive", "plots"), "/", prefix, "_offset_extracted_bio1.pdf"), width = 12, height = 8)
p_ext_bio1
dev.off()
```
