---
title: "genomic_offset"
output:
  github_document:
    toc: TRUE
---

```{r setup, include=FALSE}
# Set global options for caching
knitr::opts_chunk$set(
  cache = TRUE,
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  result = FALSE)

library(algatr)
library(here)
library(tidyverse)
library(tidyterra)
library(RColorBrewer)
library(sf)
library(terra)
library(raster)
library(cowplot)
library(RStoolbox)

source(here("analysis", "adaptive", "adaptive_index.R"))
source(here("analysis", "adaptive", "genomic_offset.R"))
source(here("general_functions.R"))
```

## Specify model to examine

```{r}
gea_method = "bio1ndvi" # "pca" or "bio1ndvi"

if (gea_method == "bio1ndvi") output_path = here("analysis", "adaptive", "outputs", "RDA_bio1_ndvi") # for BIO1 + NDVI
if (gea_method == "pca") output_path = here("analysis", "adaptive", "outputs", "RDA_PCA") # for rasterPCA env layers
```

## 1. Get input files

```{r}
dat_mod <- import_rda(output_path = output_path, suffix = "mod", rds_obj = TRUE)
dat_mod_pc <- import_rda(output_path = output_path, suffix = "modPCs", rds_obj = TRUE)

# Retrieve coords for samples on which RDA was run
coords <- read_tsv(here(output_path, "RDA_genes_coords.txt"), col_names = TRUE)
ca <- get_ca()
```

## 2. Select env layers

```{r}
if (gea_method == "pca") envlayers <- import_env_files(type = "rasterPCs", future = TRUE)
if (gea_method == "bio1ndvi") envlayers <- import_env_files(type = "ind_layers", future = TRUE)

# Check that above worked
names(envlayers$env_pres)
names(envlayers$env_fut)

all(res(envlayers$env_pres) == res(envlayers$env_fut))

# Rename to keep consistent with RDA results
names(envlayers$env_pres) <- c("env_BIO1", "env_NDVI")

# Change layer names to keep consistent with rownames in RDA model
# if (all(names(envlayers$env_pres) != row.names(dat_mod$biplot))) names(envlayers$env_pres) <- paste("env_", names(envlayers$env_pres), sep = "")
# if (names(envlayers$env_pres) != row.names(dat_mod_pc$biplot)) names(envlayers$env_pres) <- paste("env_", names(envlayers$env_pres), sep = "")
```

```{r}
# Get scaled and centered info
env <- terra::extract(envlayers$env_pres, coords %>% dplyr::select(x, y))
env <- scale(env, center = TRUE, scale = TRUE)
# Recover scaling coefficients
scale_env <- attr(env, 'scaled:scale')
center_env <- attr(env, 'scaled:center')
```

## 2. Genomic offset

The `genomic_offset()` function outputs a list of five 
objects: (1) Proj_pres, which is the present projection (equivalent to the adaptive index); (2) Proj_fut, which projects the adaptive index values on to future
environmental layers, (3) Proj_offset, which is the genonic offset (i.e., difference between Proj_pres and Proj_fut), (4) Proj_offset_global, which is the global
offset if multiple future environmental time points were provided, and (5) weights, which are the eigenvalues from the RDA model and are how the RDA axes values 
were weighted.

```{r}
# Check that names are consistent
all(names(envlayers$env_pres) == row.names(dat_mod$biplot))

offset <- genomic_offset(loadings = dat_mod$scaledload, biplot = dat_mod$biplot, eig = dat_mod$eig, K = ncol(dat_mod$biplot), env_pres = envlayers$env_pres, env_fut = envlayers$env_fut, range = NULL, 
                         method = "loadings", scale_env, center_env, mod = NULL)

offset_pc <- genomic_offset(loadings = dat_mod_pc$scaledload, biplot = dat_mod_pc$biplot, eig = dat_mod_pc$eig, K = ncol(dat_mod_pc$biplot), env_pres, env_fut, range = NULL, 
                         method = "loadings", scale_env, center_env, mod = NULL)
```

## 3. Save genomic offset maps

```{r}
if (gea_method == "bio1ndvi") {
  offset_26 <- raster::stack(offset$Proj_offset_RCP26$RDA1, offset$Proj_offset_RCP26$RDA2)
  offset_85 <- raster::stack(offset$Proj_offset_RCP85$RDA1, offset$Proj_offset_RCP85$RDA2)
}
if (gea_method == "pca") {
  offset_26 <- raster::stack(offset$Proj_offset_RCP26$RDA1, offset$Proj_offset_RCP26$RDA2, offset$Proj_offset_RCP26$RDA3)
  offset_85 <- raster::stack(offset$Proj_offset_RCP85$RDA1, offset$Proj_offset_RCP85$RDA2, offset$Proj_offset_RCP85$RDA3)
}

offset_26_g <- raster::stack(offset$Proj_offset_global_RCP26)
offset_85_g <- raster::stack(offset$Proj_offset_global_RCP85)

writeRaster(offset_26, here(output_path, "offset_load_RCP26.tif"), overwrite = TRUE)
writeRaster(offset_26_g, here(output_path, "offset_load_global_RCP26.tif"), overwrite = TRUE)
writeRaster(offset_85, here(output_path, "offset_load_RCP85.tif"), overwrite = TRUE)
writeRaster(offset_85_g, here(output_path, "offset_load_global_RCP85.tif"), overwrite = TRUE)
```

## 4. Plot genomic offset

The `mask` argument is whether or not to crop (and resample) values based on the background sf provided in `bkg`.

```{r}
offset_26 <- terra::rast(here(output_path, "offset_load_RCP26.tif"))
offset_85 <- terra::rast(here(output_path, "offset_load_RCP85.tif"))

# No pop structure correction
p_pres <- plot_adaptive(offset$Proj_pres, K = length(offset$Proj_pres), bkg = ca, index_name = "Adaptive index", title = "Present values (loadings method)", to_mask = TRUE)
p_fut <- plot_adaptive(offset$Proj_fut, K = length(offset$Proj_fut), bkg = ca, index_name = "Adaptive index", title = "Future values", to_mask = TRUE)
p_offset <- plot_adaptive(offset$Proj_offset, K = length(offset$Proj_offset), bkg = ca, index_name = "Genomic offset", title = "Genomic offset", to_mask = TRUE)
offset_plots <- plot_grid(p_pres, p_fut, p_offset, nrow = 3)

pdf(here("analysis", "adaptive", "plots", "offset_plots.pdf"), width = 12, height = 8)
p
dev.off()

# With pop structure correction
p_pres_pc <- plot_adaptive(offset_pc$Proj_pres, K = length(offset_pc$Proj_pres), bkg = ca, index_name = "Adaptive index, \nPC correction", title = "Present values (loadings method)", to_mask = TRUE)
p_fut_pc <- plot_adaptive(offset_pc$Proj_fut, K = length(offset_pc$Proj_fut), bkg = ca, index_name = "Adaptive index, \nPC correction", title = "Future values", to_mask = TRUE)
p_offset_pc <- plot_adaptive(offset_pc$Proj_offset, K = length(offset_pc$Proj_offset), bkg = ca, index_name = "Genomic offset, \nPC correction", title = "Genomic offset", to_mask = TRUE)
offset_plots_pc <- plot_grid(p_pres_pc, p_fut_pc, p_offset_pc, nrow = 3)

pdf(here("analysis", "adaptive", "plots", "offset_plots_PCs.pdf"), width = 12, height = 8)
offset_plots_pc
dev.off()

colors <- c(colorRampPalette(brewer.pal(11, "Spectral")[6:5])(2), colorRampPalette(brewer.pal(11, "Spectral")[4:3])(2), colorRampPalette(brewer.pal(11, "Spectral")[2:1])(3))
ggplot() + 
  geom_sf(data = bkg, fill = "lightgrey") +
  geom_spatraster(data = offset_26) + 
  scale_fill_viridis_c(option = "B", na.value = "transparent") +
  geom_sf(data = bkg, fill = NA, size = 0.1) +
  xlab("Longitude") + 
  ylab("Latitude") +
  facet_grid(~lyr) +
  theme_map() +
  # theme_bw() +
  theme(panel.grid = element_blank(), plot.background = element_blank(), 
        panel.background = element_blank(), strip.text = element_text(size = 11))
```

Plotting offset over multiple time periods:

```{r}
# # Table global genetic offset predicted for 2050 and 2080
# RDA_proj_offset <- data.frame(rbind(rasterToPoints(offset$Proj_offset_global), 
#                                     rasterToPoints(Proj_offset_global)), 
#                                     Date = c(rep("2050", nrow(rasterToPoints(Proj_offset_global))), rep("2080", nrow(rasterToPoints(Proj_offset_global)))))

# ## Projecting genomic offset on a map
# colors <- c(colorRampPalette(brewer.pal(11, "Spectral")[6:5])(2), colorRampPalette(brewer.pal(11, "Spectral")[4:3])(2), colorRampPalette(brewer.pal(11, "Spectral")[2:1])(3))

# ggplot(data = RDA_proj_offset) + 
#     geom_sf(data = ca, fill = "lightgrey") +
#   geom_raster(aes(x = x, y = y, fill = cut(Global_offset, breaks = seq(1, 8, by = 1), include.lowest = T)), alpha = 1) + 
#   scale_fill_manual(values = colors, labels = c("1-2","2-3","3-4","4-5","5-6","6-7","7-8"), guide = guide_legend(title = "Genomic offset", title.position = "top", title.hjust = 0.5, ncol = 1, label.position = "right"), na.translate = F) +
#   geom_sf(data = ca, fill = NA, size = 0.1) +
#   xlab("Longitude") + ylab("Latitude") +
#   facet_grid(~ Date) +
#   theme_bw(base_size = 11, base_family = "Times") +
#   theme(panel.grid = element_blank(), plot.background = element_blank(), panel.background = element_blank(), strip.text = element_text(size = 11)) +
#   ggtitle(paste0(species, " genomic offset"))
```

```{r}
offset_26 <- terra::rast(here(output_path, "offset_load_RCP26.tif"))
# offset_26_g <- terra::rast(here(output_path, "offset_load_global_RCP26.tif"))

offset = offset_26

# Count number of layers
n_layers <- terra::nlyr(offset)
# Max number of layers to plot is 3, so adjust n_layers accordingly
if (n_layers > 3) {
  n_layers <- 3
}
# Scale rasters to get colors (each layer will correspond with R, G, or B in the final plot)
offsetRGB <- stack_to_rgb(offset)

# If there are fewer than 3 n_layers (e.g., <3 variables), the RGB plot won't work (because there isn't an R, G, and B)
  # To get around this, create a blank raster (i.e., a white raster), and add it to the stack
  if (n_layers < 3) {
    warning("Fewer than three non-zero coefficients provided, adding white substitute layers to RGB plot")
    # Create white raster by multiplying a layer of pcaRast by 0 and adding 255
    white_raster <- offsetRGB[[1]] * 0 + 255
  }

  # If n_layers = 2, you end up making a bivariate map
  if (n_layers == 2) {
    offsetRGB <- c(offsetRGB, white_raster)
  }

  # If n_layers = 1, you end up making a univariate map
  if (n_layers == 1) {
    offsetRGB <- c(offsetRGB, white_raster, white_raster)
  }

# terra::plotRGB(offsetRGB, r = 1, g = 2, b = 3)
ggplot() +
  geom_sf(data = ca, fill = "lightgrey") +
  geom_spatraster_rgb(data = offsetRGB, r = 3, g = 2, b = 1) +
  theme_map()

ggplot() +
  geom_sf(data = ca, fill = "lightgrey") +
  geom_spatraster(data = offset_26) +
  theme_map() +
  facet_wrap(~lyr) +
  scale_fill_viridis_c(na.value = "transparent", limits = c(440, 770))

ggplot() +
  geom_sf(data = ca, fill = "lightgrey") +
  geom_spatraster(data = offset_26[[2]]) +
  theme_map() +
  # facet_wrap(~lyr) +
  scale_fill_viridis_c(na.value = "transparent")

# gdm_plot_vars(pcaSamp, pcaRast, pcaRastRGB, coords, x = "PC1", y = "PC2", scl = scl, display_axes = display_axes)

# Plot points only, colorized based on RGB offset value
pts <- terra::extract(offsetRGB, coords %>% dplyr::select(x, y))
pts <- bind_cols(pts, coords)

ggplot() +
  geom_sf(data = ca, fill = "lightgrey") +
  geom_point(data = pts %>% pivot_longer(cols = 2:4, names_to = "axis", values_to = "offset_value"), aes(x = x, y = y, color = offset_value), size = 2) +
  scale_color_viridis_c(option = "D") +
  theme_map() +
  facet_grid(~axis)

# Plot points only, colorized by RDA1 value
fut_bio1_26 <- terra::rast(here("data", "env", "future", "envicloud/chelsa/chelsa_V2/GLOBAL/climatologies", "/2071-2100/GFDL-ESM4/ssp126/bio/", "CHELSA_bio1_2071-2100_gfdl-esm4_ssp126_V.2.1.tif"))
fut_bio1_85 <- terra::rast(here("data", "env", "future", "envicloud/chelsa/chelsa_V2/GLOBAL/climatologies", "/2071-2100/GFDL-ESM4/ssp585/bio/", "CHELSA_bio1_2071-2100_gfdl-esm4_ssp585_V.2.1.tif"))

# resamp <- terra::resample(fut_bio1_26, offset_26)
crop <- terra::crop(fut_bio1_26, ca, mask = TRUE)
pts <- terra::extract(offset_26[[1]], coords %>% dplyr::select(x, y))
pts <- bind_cols(pts, coords)

ggplot() +
  geom_sf(data = ca, fill = "lightgrey") +
  geom_spatraster(data = crop) +
  scale_fill_viridis_c(na.value = "transparent") +
  geom_point(data = pts, aes(x = x, y = y, color = offset_load_RCP26_1), size = 2) +
  scale_color_viridis_c(option = "A") +
  theme_map()

env_pcs <- terra::rast(here("data", "env", "california_chelsa_bioclim_1981-2010_V.2.1_pca.tif"))
fut_pcs_26 <- terra::rast(here("data", "env", "future", "CHELSA_2071-2100_gfdl-esm4_ssp126_V.2.1_pca.tif"))
fut_pcs_85 <- terra::rast(here("data", "env", "future", "CHELSA_2071-2100_gfdl-esm4_ssp585_V.2.1_pca.tif"))

# Plot points against future BIO1 for RCP2.6 "Annual mean temperature"

```
