---
title: "Admixture analysis"
output:
  github_document:
    toc: TRUE
---
```{r, include = FALSE}
# Set global options for caching
knitr::opts_chunk$set(
  cache = TRUE,
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  result = FALSE
)
library(here)
library(tidyverse)
library(cowplot)
library(sf)
library(terra)
library(ggpubr)
library(algatr)
library(viridis)
library(vcfR)
library(hierfstat)
library(tigris)
devtools::load_all()
theme_set(theme_cowplot())

source(here("analysis", "anne", "admixture.R"))
source(here("general_functions.R"))
```

-   [1. Import metadata]
-   [2. Import and visualize ADMIXTURE results]
-   [3. Kriging ADMIXTURE results]
-   [4. Structure boundaries]

# 1. Import metadata

```{r, echo = FALSE, message = FALSE, include=FALSE}
# Load the U.S. state boundaries data
ca <- get_ca()
# make theme for plotting
theme_custom <- 
  theme(
    strip.background = element_rect(color = "gray", fill = "gray"),
    panel.background = element_rect(color = "gray", fill = NA),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    strip.text = element_text(size = 16, face = "italic")
    )
```

```{r, include = FALSE}
# get coordinates 
coords <- read_tsv(here("analysis", "anne", "outputs", "58-Sceloporus.coords.txt"), col_names = FALSE) %>% 
  rename(SampleID = X1, x = X2, y = X3)
```

# 2. Import and visualize ADMIXTURE results

Import and plot cross-validation results for K=1-10.

```{r cv, fig.width = 13, fig.height = 6}
cv <- get_cv(output_dir = here("analysis", "anne", "outputs"), 
             prefix = "58-Sceloporus_maf05_minDP5_maxDP50_rmsamp60_mm80_rmsamp20_r60")

ggplot(cv, aes(x = K, y = cv_error, group = 1)) +
  geom_point() +
  geom_line() +
  theme_bw()
```

Import Q-values for all K values.

```{r admixture}
Kvals <- c(1:10)
Q <-
  map(Kvals, ~get_Q(K = .x, 
                   output_dir = here("analysis", "anne", "outputs"), 
                   prefix = "58-Sceloporus_maf05_minDP5_maxDP50_rmsamp60_mm80_rmsamp20_r60") %>% 
  mutate(K = .x)) %>% 
  bind_rows() %>%
  arrange(across(starts_with("X"))) %>%
  mutate(SampleID = factor(SampleID, levels = unique(SampleID))) %>%
  pivot_longer(starts_with("X"), names_to = "PopGroup", values_to = "Q")

# Join with coordinates
df <-
 left_join(Q, coords[,c("x", "y", "SampleID")], by = "SampleID") %>%
 st_as_sf(coords = c("x", "y")) %>%
 st_set_crs(st_crs(4326)) # 4326 or 3310

# Plot, faceting by K-value and remove outlier sample (likely incorrect locality)
ggplot(df %>% filter(K != 1) %>% filter(SampleID != "Scelocci_CHI1382_DAW5-46-21")) +
  geom_sf(data = ca) +
  geom_sf(aes(col = cluster)) +
  facet_wrap(~K, nrow = 2) +
  theme_void() +
  scale_color_manual(values = viridis::turbo(max(Kvals)))

barplots <- map(Kvals, ~structure_plot(qmat = get_Q(K = .x, 
                                                    output_dir = here("analysis", "anne", "outputs"),
                                                    prefix = "58-Sceloporus_maf05_minDP5_maxDP50_rmsamp60_mm80_rmsamp20_r60",
                                                    qmat_only = TRUE)) + 
                  scale_fill_manual(values = viridis::turbo(max(Kvals))))
do.call(ggarrange, c(barplots, ncol = 1))
```

Build maps and barplots for specific K values. Based on the CV error, let's look at K=5-7.

```{r, fig.width = 13, fig.height = 6}
ggplot(df %>% filter(K %in% 5:7) %>% filter(SampleID != "Scelocci_CHI1382_DAW5-46-21")) +
  geom_sf(data = ca) +
  geom_sf(aes(col = cluster)) +
  facet_wrap(~K, nrow = 1) +
  theme_void() +
  scale_color_manual(values = viridis::turbo(7))

barplots <- map(5:7, ~structure_plot(qmat = get_Q(K = .x, 
                                                  output_dir = here("analysis", "anne", "outputs"),
                                                  prefix = "58-Sceloporus_maf05_minDP5_maxDP50_rmsamp60_mm80_rmsamp20_r60",
                                                  qmat_only = TRUE)) + 
                  scale_fill_manual(values = viridis::turbo(7)))
do.call(ggarrange, c(barplots, ncol = 1))
```

Save Q-values to file. `cluster` is the final cluster assignment for SampleID given max Q.

```{r save}
Kval = 5

pops <- paste0("X", 1:Kval)
Q %>% 
  filter(K == Kval) %>% 
  filter(SampleID != "Scelocci_CHI1382_DAW5-46-21") %>% 
  filter(PopGroup %in% pops) %>% 
  write_tsv(paste0(here("analysis", "anne", "outputs"), "/Q", Kval, ".txt"), col_names = TRUE)

# Write out samples in each population for ROH calculations
Q %>%
  filter(K == Kval) %>%
  filter(SampleID != "Scelocci_CHI1382_DAW5-46-21") %>% 
  # only keep the pop group corresponding to the cluster
  mutate(PopGroup = as.numeric(gsub("X", "", PopGroup))) %>%
  filter(cluster == PopGroup) %>%
  mutate(pop = paste0("pop", cluster)) %>%
  group_by(cluster) %>%
  group_split() %>%
  map(~as.character(pull(.x, SampleID))) %>%
  iwalk(\(x, idx) write(x, here("analysis", "anne", "outputs", paste0("K", Kval, "_pop", idx, ".txt"))))
```

# 3. Kriging ADMIXTURE results

```{r krige}
algatr::load_algatr_example()
# Get raster onto which we'll krige results
krig_raster <- raster::aggregate(CA_env[[1]], fact = 6)
# envdat <- reproject(coords %>% filter(species == speciesID), 
#                     env = krig_raster,
#                     crs_code = 3310)

qmat <- get_Q(K = Kval,
              output_dir = here("analysis", "anne", "outputs"),
              prefix = "58-Sceloporus_maf05_minDP5_maxDP50_rmsamp60_mm80_rmsamp20_r60") %>% 
  filter(SampleID != "Scelocci_CHI1382_DAW5-46-21" | SampleID != "")

df <-
 left_join(qmat, coords[ ,c("x", "y", "SampleID")], by = "SampleID")

coords <- df %>% 
  dplyr::select(x, y) %>% 
  st_as_sf(coords = c("x", "y")) %>%
  st_set_crs(st_crs(4326))

krig_admix <- algatr::tess_krig(qmat = qmat %>% 
                                  dplyr::select(-c(cluster, K, SampleID, geometry)), 
                                coords = qmat$geometry, 
                                grid = envdat$env_proj)

tess_ggplot(map, plot_method = "maxQ", ggplot_fill = scale_fill_manual(values = cols))
```
# 4. Structure boundaries
