---
title: "genomic_offset"
output:
  github_document:
    toc: TRUE
---

```{r setup, include=FALSE}
# Set global options for caching
knitr::opts_chunk$set(
  cache = TRUE,
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  result = FALSE)

library(algatr)
library(here)
library(tidyverse)
library(RColorBrewer)
library(sf)
library(terra)
library(raster)
library(cowplot)
library(RStoolbox)

# source(here("analysis", "anne", "genomic_offset.R"))
source(here("general_functions.R"))
```

-   [1. Process environmental layers]
-   [2. Get input files]
-   [3. Calculate adaptive index]
-   [4. Build map of adaptive index]

# 1. Process environmental layers

## 1a. Run rasterPCA on present env layers to get model

```{r}
env_pres <- raster::stack(here("data/env/california_chelsa_bioclim_1981-2010_V.2.1.tif"))
# Rename layers (names must be consistent with future too)
names(env_pres) <- c("BIO1", "BIO10", "BIO11", "BIO12", "BIO13",
                      "BIO14", "BIO15", "BIO16", "BIO17", "BIO18", 
                      "BIO19", "BIO2", "BIO3", "BIO4", "BIO5",
                      "BIO6", "BIO7", "BIO8", "BIO9")
env_pcs <- RStoolbox::rasterPCA(scale(env_pres)) # outputs map as SpatRaster

# Specify projection is lat/long to keep consistent with how envlayers were projected
# env_proj <- raster::projectRaster(env_pcs$map, crs = "+proj=longlat") # 4326
```

## 1b. Read in and process future env layers

```{r}
files <- list.files(here("data/env/CHELSA_future/bio"), full.names = TRUE)
env_future <- raster::stack(files)
# Rename vars to keep consistent with env_pres
names(env_future) <- c("BIO1", "BIO10", "BIO11", "BIO12", "BIO13",
                      "BIO14", "BIO15", "BIO16", "BIO17", "BIO18", 
                      "BIO19", "BIO2", "BIO3", "BIO4", "BIO5",
                      "BIO6", "BIO7", "BIO8", "BIO9")

env_pcs_terra <- terra::rast(env_pcs$map)
# Crop to match extent of env PCs
cropped <- terra::crop(terra::rast(env_future), terra::ext(env_pcs_terra))
# resamp <- terra::resample(cropped, env_pcs_terra)
```

## 1c. Predict future values using rasterPCA model

```{r}
# Predict values of future bioclim layers using present rasterPCA model
future_pcs <- terra::predict(cropped,
                             model = env_pcs$model)

masked <- terra::mask(future_pcs[[1:3]], env_pcs$map[[1]])
names(masked) <- c("PC1", "PC2", "PC3")
raster::writeRaster(raster::stack(masked),
                  here("data/env/CHELSA_2041-2070_ipsl-cm6a-lr_ssp585_V.2.1_pca.tif"))
```

# 2. Get input files

## 2a. Env files

```{r}
# Env files
env_pres <- raster::stack(here("data/env/california_chelsa_bioclim_1981-2010_V.2.1_pca.tif"))
env_fut <- raster::stack(here("data/env/CHELSA_2041-2070_ipsl-cm6a-lr_ssp585_V.2.1_pca.tif"))
names(env_fut) <- names(env_pres)

ca <- get_ca()
```

## 2b. Other files

```{r}
coords <- read_tsv(here("data/ccgp_data/58-Sceloporus.coords.txt"), col_names = c("INDV", "x", "y"))
# coords <- get_coords(sf = TRUE)

# Genetic data
# TODO get RDA SNPs from genes
```

## 3. Run RDA on SNPs in genes

```{r}

```

## 4. Adaptive index

```{r}
source(here("analysis", "anne", "adaptive_index.R"))
```
## 5. Genomic offset

